### 2012-11-19 doc/levelset/level_set_method_on_quadtree_JCP_rebuttal.tex
Add Frederec's level set paper

%-------------------------------------------------------------------------------
%
% pre-Header
%
%
%-------------------------------------------------------------------------------
\documentclass[english]{article}
\usepackage[T1]{fontenc}
\usepackage[latin1]{inputenc}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{times}
\usepackage{geometry}
\usepackage{color}
\usepackage{multirow}
\usepackage{bigstrut}

\geometry{verbose,letterpaper,tmargin=1in,bmargin=1in,lmargin=1in,rmargin=1in}
\DeclareMathOperator*{\avg}{average}
\usepackage{babel}

\begin{document}
\newtheorem{thm}{Theorem}
\newtheorem{cor}{Corollary}
\newtheorem{argm}{Argument}
\newtheorem{defn}{Definition}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newenvironment{proof}[1][Proof]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}
\newenvironment{definition}[1][Definition]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}
\newenvironment{example}[1][Example]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}
\newenvironment{remark}[1][Remark]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}
\newcommand{\qed}{\nobreak \ifvmode \relax \else
\ifdim\lastskip<1.5em \hskip-\lastskip
\hskip1.5em plus0em minus0.5em \fi \nobreak
\vrule height0.75em width0.5em depth0.25em\fi}

\newcommand{\etal}{\textit{et al. }}
\newcommand{\TT}{\times 10^}

%-------------------------------------------------------------------------------
%
% Title
%
%-------------------------------------------------------------------------------
\title{A Second Order Accurate Level Set Method on Non-Graded Adaptive Cartesian Grids}
\author{Chohong Min \thanks{Mathematics Department, University of California, Santa Barbara, CA 93106.} \hspace{2cm}
Fr\'{e}d\'{e}ric Gibou \thanks{Mechanical Engineering Department \& Computer Science Department,
University of California, Santa Barbara, CA 93106.}} \maketitle
\begin{abstract}
We present a level set method on non-graded adaptive Cartesian grids, i.e.
grids for which the ratio between adjacent cells is not constrained. We use
quadtree and octree data structures to represent the grid and a simple
algorithm to generate a mesh with the finest resolution at the interface.
In particular, we present (1) a locally third order accurate
reinitialization scheme that transforms an arbitrary level set function
into a signed distance function, (2) a second order accurate
semi-Lagrangian methods to evolve the linear level set advection equation
under an externally generated velocity field, (3) a second order accurate
upwind method to evolve the nonlinear level set equation under a normal
velocity as well as to extrapolate scalar quantities across an interface in
the normal direction, and (4) a semi-implicit scheme to evolve the
interface under mean curvature. Combined, we obtain a level set method on
adaptive Cartesian grids with a negligible amount of mass loss. We propose
numerical examples in two and three spatial dimensions to demonstrate the
accuracy of the method.
\end{abstract}

%-------------------------------------------------------------------------------
%
% Introduction
%
%-------------------------------------------------------------------------------
\section{Introduction}

Many problems in science and engineering can be described by a moving free
boundary model. Examples include free surface flows, Stefan problems and
multiphase flows to cite a few. The difficulty in solving these problems
stems from the fact that: First, they involve dissimilar length scales.
Second, the boundary position must be computed as part of the solution
process. Third, the interface may be expected to undergo complex
topological changes, such as the merging or the pinching of two fronts.
Numerically, the interface that separates the two phases can be either
explicitly tracked or implicitly captured. Several classes of successful
methods exist with their own virtues and drawbacks.

Volume of Fluid methods
\cite{benson:1992:hydrocodes,benson:2002:VOF_Review,DeBar:1974:KRAKEN,Hirt_Nichols:1981:VOF,Noh_Woodward:1976:SLIC,Youngs:1984:VOF}
have the advantage of being volume preserving since the mass fraction in
each cell is being tracked. However, it is often difficult to extract
geometrical properties such as curvatures due to the fact that it is
challenging or even impossible to reconstruct a smooth enough function from
the mass fractions alone. We note however that some recent improvement in
interface reconstruction can be found in \cite{Dyadechko:2006:MOF}.

The main advantage of an explicit approach, e.g. front tracking
\cite{glimm:1999:fronttracking, Juric:1996:Dendritic_Solidification,
Juric:1998:Boiling_Flows, Tryggvason:2001:Front_Tracking_Flows_Review}, is
its accuracy. The main disadvantage is that additional treatments are
needed for handling changes in the interface's topology. In turn, the
explicit treatment of connectivity makes the method challenging to extend
to three spatial dimensions. While researchers have produced remarkable
results for a wide variety of applications using front tracking techniques,
these difficulties make this approach not ideally suited for studying
interface problems with changes in topology. Implicit representations such
as the level set method or the phase-field method represent the front as an
isocontour of a continuous function. Topological changes are consequently
handled in a straightforward fashion, and thus these methods are readily
implemented in both two and three spatial dimensions.

The main idea behind the phase-field method is to distinguish between
phases with an order parameter (or phase-field) that is constant within
each phase but varies smoothly across an interfacial region of finite
thickness. The dynamics of the phase-field is then coupled to that of the
solution in such a way that it tracks the interface motion and approximates
the sharp interface limit when the order parameter vanishes. Phase-field
methods are very popular techniques for simulating dendritic growth for
example and have produced accurate quantitative results, e.g.
\cite{Langer:1986:Phase_Field, Karma_Rappel:1997:Phase_Field_2D_3D,
Karma:2001:Phase_Field_Different_Diffusion,
Nestler:2005:Phase_Field_FEM_FDM, Schmidt:1996:CTD}. However, these methods
suffer from their own limitations: Phase-field methods have only an
approximate representation of the front location and thus the
discretization of the diffusion field is less accurate near the front,
resembling an enthalpy method \cite{chorin:1967:ANM}. Another consequence
is the stringent time step restriction imposed by such methods. Karma and
Rappel \cite{Karma_Rappel:1996:Phase_Field_Kinetics} have developed a
thin-interface limit of the phase-field model with a significant
improvement of the capillary length to interface thickness ratio
constraint; however, the time step restriction is still on the order of the
microscopic capillary length. Another disadvantage is the potential
difficulty in relating the phase-field parameters to the physical
parameters \cite{Wheeler:1993:Handbook}, although some progress is being
made for some wider class of problems
\cite{Elder_Provatas:2001:Phase_Field_Sharp_Interface}.

The main difference between the phase-field method and the level set approach
\cite{osher:1988:originallevelset, sethian:1999:levelset, osher:2002:levelset} is that the level
set method is a sharp interface model. The level set can therefore be used to exactly locate the
interface in order to apply discretizations that depend on the exact interface location.
Consequently, the sharp interface equation can be solved directly with no need for asymptotic
analysis, which makes the method potentially more attractive in developing general tool box
software for a wide range of applications. Another advantage is that only the standard time step
restrictions for stability and consistency are required, making the method significantly more
efficient. Level set methods have been extremely successful on uniform grids in the study of
physical problems such as compressible flows, incompressible flows, multiphase flows (see e.g.
\cite{osher:2002:levelset, sethian:1999:levelset} and the references therein), Epitaxial growth
(see e.g.\cite{Caflisch:1999:Island_Dynamics, Gibou:2001:Rate_Equations_Scaling_Laws,
Gibou:2001:Rate_Equations, Ratsch:2001:Island_Dynamics_Review} and the references therein) or in
image processing (see e.g. \cite{osher:2003:levelset} and the references therein). One of the main
problem of the level set method, namely its mass loss, has been partially solved with the advent of
the particle level set method of Enright \etal \cite{Enright:2002:PLS}. Within this method, the
interface is captured by the level set method and massless particles are added in order to reduce
the mass loss. The massless particles are also used in the reinitialization process for obtaining
smoother results for the reinitialized level set function. However, the use of particles adds to
the CPU and the memory requirement and cannot be applied for flows producing shocks. Rather
recently, there has been a thrust in developing level set methods on adaptive Cartesian grids. For
example Losasso \etal \cite{losasso:2004:octree} presented a particle level set based method to
simulate free surface flows on non-graded Cartesian grids. Within this method, the interface
between the liquid and the air is captured by the particle level set on a non graded octree data
structure. Other interesting work on adaptive level-set methods can be found in
\cite{droske01adaptive} and \cite{Milne:1995:thesis}.

In this paper, we present a general particle-less level set method on
non-graded Cartesian grids that produces a negligible amount of mass loss.
We apply this method to the level set evolution (1) with an externally
generated velocity field, (2) in the normal direction and (3) under mean
curvature. We also present a locally third order accurate reinitialization
scheme that transforms an arbitrary function into a sign distance function
as well as standard techniques to extrapolate a scalar quantities across an
interface in its normal direction.

\section{The Level Set Method}
The level set method, introduced by Osher and Sethian
\cite{osher:1988:originallevelset} describes a curve in two spatial
dimensions or a surface in three spatial dimensions by the zero-contour of
a higher dimensional function $\phi$, called the level set function. For
example, in two spatial dimension, a curve is define by $\{(x, y):
\phi(x,y)=0\}$. Under a velocity field $V$, the interface deforms according
to the level set equation
\begin{eqnarray}
\phi_t+V \cdot \nabla \phi = 0. \label{eq_levelset}
\end{eqnarray}
To keep the values of $ \phi $ close to those of a signed distance
function, i.e. $ |\nabla \phi| = 1 $, the reinitialization equation
introduced in Sussman \etal \cite{Sussman:1994:levelsetincompressible}
\begin{eqnarray}
\phi_{\tau} + S(\phi_o) \left( |\nabla \phi| - 1 \right) =0 \label{reinit}
\end{eqnarray}
is traditionally iterated for a few steps in fictitious time, $ \tau $.
Here $S(\phi_o)$ is a smoothed out sign function. The level set function is
used to compute the normal $$ \vec{n} = \nabla \phi/|\nabla \phi|, $$ and
the mean curvature $$\kappa = \nabla \cdot \vec{n}. $$ We refer the
interested readers to the book by Osher and Fedkiw
\cite{osher:2002:levelset} as well to the book by Sethian
\cite{sethian:1999:levelset} for more details on the level set method.

%-------------------------------------------------------------------------------
%
% Spatial Discretization
%
%-------------------------------------------------------------------------------
\section{Spatial Discretization and Refinement Criterion \label{sec_spatial_discretization}}

We use a standard quadtree (resp. octree) data structure to represent the
spatial discretization of the physical domain in two (resp. three) spatial
dimensions as depicted in figure \ref{fig_quadtree_description}: Initially
the root of the tree is associated with the entire domain, then we
recursively split each cell into four children until the desired level of
detail is achieved. This is done similarly in three spatial dimensions,
except that cells are split into eight cubes (children). We refer the
reader to the books of Samet \cite{Samet:1990:AOS,Samet:1989:TDA} for more
details on quadtree/octree data structures.

By definition, the difference of level between a parent cell and its direct
descendant is one. The level is then incremented by one for each new
generation of children.  A tree in which the difference of level between
adjacent cells is at most one is called a graded tree. Meshes associated
with graded trees are often used in the case of finite element methods in
order to produce procedures that are easier to implement. Graded Cartesian
grids are also used in the case of finite difference schemes, see for
example the work of Popinet \cite{Popinet:2003:Gerris} for the study of
incompressible flows. Graded meshes impose that extra grid cells must be
added in regions where they are not necessarily needed, consuming some
computational resources that cannot be spent elsewhere, eventually limiting
the highest level of detail that can be achieved. Moore
\cite{Moore:1995:The_Cost_of_Balancing_Quadtrees} demonstrates that the
cost of transforming an arbitrary quadtree into a graded quadtree could
involve 8 times as many grid nodes. Weiser \cite{Weiser:1981:thesis}
proposed a rough estimate for the three dimensional case and concluded that
as much as 71 times as many grid nodes could be needed for balancing
octrees. These estimates clearly represent the worse case scenarios that
seldom exist in practical simulations. However, there is still a non
negligible difference between graded and non graded grids. In addition, not
imposing any constraint on the difference of level between two adjacent
cells allows for easier/faster adaptive mesh generations.

\begin{figure}
\begin{center}\includegraphics[scale=0.3]{quadtree_description.eps}
\includegraphics[scale=0.3]{quadtree_description_2.eps}\end{center}
\caption{Discretization of a two dimensional domain (left) and its quadtree
representation (right). The entire domain corresponds to the root of the
tree (level 0). Then each cell can be recursively subdivided further into
four children. In this example, the tree is ungraded since the difference
of level between cells exceeds one.\label{fig_quadtree_description}}
\end{figure}

In this work we choose to impose that the finest cells lie on the
interface, since it is the region of interest for the level set method. In
order to generate adaptive Cartesian grids, one can use the signed distance
function to the interface along with the Whitney decomposition, as first
proposed by Strain in \cite{strain:1999:tree}. Simply stated, one
\textit{"splits any cell whose edge length exceeds its distance to the
interface"}. For a general function $\phi:\mathbb{R}^n\to\mathbb{R}$ with
Lipschitz constant $Lip(\phi)$, the Whitney decomposition was extended in
Min \cite{min:2004:lls}: Starting from a root cell split any cell $C$ if
$$\min_{v\in \textrm{vertices}(C)}|\phi(v)| \le \textrm{Lip}(\phi)\cdot
\textrm{diag-size}(C),$$ where diag-size($C$) refers to the length of the
diagonal of the current cell $C$ and $v$ refers to a vertex (node) of the
current cell.

%-------------------------------------------------------------------------------
%
% T-junction Nodes
%
%-------------------------------------------------------------------------------
\section{Finite Difference Discretizations\label{sec_finite_differences}}
\begin{figure}
\begin{center}\includegraphics[scale=0.45]{neighbors_of_node_quadtree.eps}
\end{center}
\caption{Neighboring nodes of a T-junction node,
$v_0$.\label{fig_T_junction_quadtree}}
\end{figure}

In the case of non regular Cartesian grids, the main difficulty comes from
deriving discretizations at T-junction nodes, i.e. nodes for which there is
a missing neighboring node in one of the Cartesian directions. For example
figure \ref{fig_T_junction_quadtree} depicts a T-junction node $v_0$, with
three neighboring nodes $v_1$, $v_2$ and $v_3$ aligned in the Cartesian
directions and one ghost neighboring node $v_4$ replacing the missing grid
node in the positive Cartesian direction. The value of a node-sampled
function $\phi:\{v_i\}\to\mathbb{R}$ at the ghost node $v_4$ could for
example be define by linear interpolation:
\begin{equation}
\phi_4^G = \frac{\phi_5 s_6+\phi_6 s_5}{s_5+s_6}. \label{eq_bilinear_interpolation_ghost_node_in_quadtree}
\end{equation}
However, instead of using this second order accurate interpolation, one can
instead use the following third order accurate interpolation: First, note
that a simple Taylor expansion demonstrates that the interpolation error in
equation \ref{eq_bilinear_interpolation_ghost_node_in_quadtree} is given
by:
\begin{equation}
\phi_4^G = \frac{\phi_5 s_6+\phi_6
s_5}{s_5+s_6}=\phi(v_4)+\frac{s_5s_6}{2}\phi_{yy}(v_0) + O(\triangle
x_{\textrm{smallest}})^3, \label{Taylor_Quadtree}
\end{equation}
where $\triangle x_{\textrm{smallest}}$ is the size of the smallest grid
cell with vertex $v_0$. The term $\phi_{yy}(v_0)$ can be approximated using
the standard first order accurate discretization
$\frac{2}{s_2+s_3}\left(\frac{\phi_2-\phi_0}{s_2}
+\frac{\phi_3-\phi_0}{s_3}\right)$ and cancelled out in equation
\ref{Taylor_Quadtree} to give:
\begin{equation}
\phi_4^G = \frac{\phi_5 s_6+\phi_6 s_5}{s_5+s_6}-\frac{s_5s_6}{s_2+s_3}\left(\frac{\phi_2-\phi_0}{s_2} +\frac{\phi_3-\phi_0}{s_3}\right). \label{eq_quadratic_interpolation_ghost_node_in_quadtree}
\end{equation}
We also point out that this interpolation only uses the node values of the
cells adjacent to $v_0$, which is particularly beneficial since access to
cells not immediately adjacent to the current cell is more difficult and
could add on CPU time and/or memory requirement.

\begin{figure}
\begin{center}\includegraphics[
  scale=0.45]{neighbors_of_node_octree.eps}
\end{center}
\caption{Neighboring vertices of a vertex three spatial
dimensions.\label{fig_T_junction_octree}}
\end{figure}

In three spatial dimensions, similar interpolation procedures can be used
to define the value of $\phi$ at ghost nodes. Referring to figure
\ref{fig_T_junction_octree}, a T-junction node $v_0$ has four regular
neighboring nodes and two ghost nodes. The values of a node-sampled
function $\phi:\{v_i\}\to\mathbb{R}$ at the ghost nodes $v_4$ and $v_5$ can
be defined by second order linear and bilinear interpolations as:
\begin{equation}
\begin{split}
\phi_4^G &= \frac{s_7\phi_8+s_8\phi_7}{s_7+s_8}, \\
\phi_5^G &= \frac{ s_{11}s_{12}\phi_{11}+ s_{11}s_{ 9}\phi_{12}+
s_{10}s_{12}\phi_{ 9}+ s_{10}s_{9}\phi_{10} }
{(s_{10}+s_{11})(s_{9}+s_{12})}.
\end{split}
\label{eq_linear_interpolation_ghost_node_in_octree}
\end{equation}
As in the case of quadtrees, third order accurate interpolations can be
derived by cancelling out the second order derivatives in the error term to
arrive at:
\begin{equation}
\begin{split}
\phi_4^G &= \frac{s_7\phi_8+s_8\phi_7}{s_7+s_8}
           - \frac{s_7s_8}{s_3+s_6}
            \left( \frac{\phi_3-\phi_0}{s_3} +
                   \frac{\phi_6-\phi_0}{s_6}
            \right), \\
\phi_5^G &= \frac{
s_{11}s_{12}\phi_{11}+
s_{11}s_{ 9}\phi_{12}+
s_{10}s_{12}\phi_{ 9}+
s_{10}s_{9}\phi_{10}
}{(s_{10}+s_{11})(s_{9}+s_{12})} \\
          &\quad - \frac{s_{10}s_{11}}{s_3+s_6}
            \left( \frac{\phi_3-\phi_0}{s_3} +
                   \frac{\phi_6-\phi_0}{s_6}
            \right) \\
          &\quad - \frac{s_{9}s_{12}}{s_1+s_4}
            \left( \frac{\phi_1-\phi_0}{s_1} +
                   \frac{\phi_4^G-\phi_0}{s_4}
            \right).
\end{split}
\label{eq_quadratic_interpolation_ghost_node_in_octree}
\end{equation}
We emphasize that figure \ref{fig_T_junction_octree} represents the general
configuration of neighboring nodes in the case of an octree as described in
Min \etal \cite{Min::Poisson_AMR_Second_Order}.

The third order interpolations defined above allow us to treat T-junction
nodes in a same fashion as a regular node, up to third order accuracy.
Here, we refer to a regular node as a node for which all the neighboring
nodes in the Cartesian directions exist. Therefore, we can then define
finite differences for $\phi_x$, $\phi_y$, $\phi_z$, $\phi_{xx}$,
$\phi_{yy}$ and $\phi_{zz}$ at every nodes using standard finite difference
formulas in a dimension by dimension framework. For example, refering to
figure \ref{fig_one_dimensional_adaptive_grid}, we use the standard
discretization for $\phi_x$ and $\phi_{xx}$, namely the central difference
formulas:
\begin{equation}
\begin{split}D^0_x\phi_0=\frac{\phi_2-\phi_0}{s_2}\cdot\frac{s_1}{s_1+s_2} + \frac{\phi_0-\phi_1}{s_1}\cdot\frac{s_2}{s_1+s_2},\\
D^0_{xx}\phi_0=\frac{\phi_2-\phi_0}{s_2}\cdot\frac{2}{s_1+s_2} - \frac{\phi_0-\phi_1}{s_1}\cdot\frac{2}{s_1+s_2},
\end{split}
\label{Central_Differencing}
\end{equation}
the forward and backward first order accurate approximations of the first order derivatives:
\begin{equation}
\begin{split}
D^+_x\phi_0 &= \frac{\phi_2-\phi_0}{s_2},\\
D^-_x\phi_0 &= \frac{\phi_0-\phi_1}{s_1},
\end{split}
\label{one_sided_1_fdm}
\end{equation}
and the second order accurate approximations of the first order derivatives:
\begin{equation}
\begin{split}
D^+_x\phi_0 &= \frac{\phi_2-\phi_0}{s_2} - \frac{s_2}{2}\text{minmod}\left(D^0_{xx}\phi_0,D^0_{xx}\phi_2\right),\\
D^-_x\phi_0 &= \frac{\phi_0-\phi_1}{s_1} +
\frac{s_1}{2}\text{minmod}\left(D^0_{xx}\phi_0,D^0_{xx}\phi_1\right),
\end{split}
\label{one_sided_2_fdm}
\end{equation}
where we use the minmod slope limiter \cite{Shu:1988:ENO,liu:1996:WENO}
because it produces more stable results in region where $\phi$ might
present kinks. Similarly, approximations for first and second order
derivatives are obtained in the $y$ and $z$ directions.

\begin{figure}
\begin{center}
\includegraphics[scale=0.9]{one_dimensional_adaptive_grid.eps}
\end{center}
\caption{One dimensional adaptive grid}
\label{fig_one_dimensional_adaptive_grid}
\end{figure}
%-------------------------------------------------------------------------------
%
% Interpolations
%
%-------------------------------------------------------------------------------
\section{Interpolation Procedures\label{sec_interpolation}}
Some reserve must be provided to define data anywhere in a cell, for
example in order to use semi-Lagrangian methods (see section
\ref{sec_semi_lagrangian}). As pointed out in Strain
\cite{Strain:2000:FSL}, the most natural choice of interpolation in
quadtree (resp. octree) data structures is the piecewise bilinear (resp.
trilinear) interpolation: Consider a cell $C$ with dimensions $[0, 1]^2$,
the bilinear interpolation at a point $x \in C$ using the values at the
nodes reads:
\begin{equation}
\begin{split}
\phi(x,y) &= \phi(0,0)(1-x)(1-y) \\
          &+ \phi(0,1)(1-x)(\quad\,\,\,\,\,y) \\
          &+ \phi(1,0)(\quad\,\,\,\,\,x)(1-y) \\
          &+ \phi(1,1)(\quad\,\,\,\,\,x)(\quad\,\,\,\,\,y)
\end{split}
\label{Bilinear_Interpolation}
\end{equation}
Quadratic interpolation can also easily be constructed using the data from the parent cell: Since
the parent cell of any current cell of a quadtree (resp. octree) owns $2\times 2$ children cells
(resp. $2\times 2\times 2$) and $3\times 3$ nodes (resp. $3\times 3\times 3$), one can defined the
multidimensional Lagrange quadratic interpolation on the parent cell. For example in the case of a
cell $[-1, 1]^2$ in a quadtree, we can define the Lagrange interpolation as:
\begin{equation*}
\begin{aligned}
\phi(x,y) &= \phi(\textrm{-1},\textrm{-1})\frac{x(x-1)}2\frac{y(y-1)}2 \\
          &+ \phi(\textrm{-1},\,\,      0)\frac{x(x-1)}2(y^2-1)        \\
          &+ \phi(\textrm{-1},\,\,      1)\frac{x(x-1)}2\frac{y(y+1)}2
\end{aligned}
\begin{aligned}
+ \phi(0,\textrm{-1})(x^2-1)\frac{y(y-1)}2 &+ \phi(1,\textrm{-1})\frac{x(x+1)}2\frac{y(y-1)}2 \\
+ \phi(0,\,\,      0)(x^2-1)(y^2-1)        &+ \phi(1,\,\,     0)\frac{x(x+1)}2(y^2-1)        \\
+ \phi(0,\,\,      1)(x^2-1)\frac{y(y+1)}2 &+ \phi(1,\,\,
1)\frac{x(x+1)}2\frac{y(y+1)}2.
\end{aligned}
\end{equation*}
However, this interpolation procedure is sensitive to nearby
discontinuities, e.g. near kinks. We therefore prefer to define a quadratic
interpolation by correcting equation (\ref{Bilinear_Interpolation}) using
second order derivatives. For a cell $[0, 1]^2$, we have:
\begin{equation}
\begin{aligned}
\phi(x,y) &= \phi(0,0)(1-x)(1-y) \\
          &+ \phi(0,1)(1-x)(\quad\,\,\,\,\,y) \\
          &+ \phi(1,0)(\quad\,\,\,\,\,x)(1-y) \\
          &+ \phi(1,1)(\quad\,\,\,\,\,x)(\quad\,\,\,\,\,y)
\end{aligned}
\begin{aligned}
\\
\\
\\        &- \phi_{xx}\frac{x(1-x)}2
           - \phi_{yy}\frac{y(1-y)}2,
\end{aligned}\label{Stabilized_Interpolation1}
\end{equation}
where we define
\begin{equation}
\begin{aligned}
\phi_{xx}&=\min_{v\in \textrm{vertices}(C)}(|D^0_{xx}\phi_{v}|),\\
\phi_{yy}&=\min_{v\in \textrm{vertices}(C)}(|D^0_{yy}\phi_{v}|).
\end{aligned}\label{Stabilized_Interpolation2}
\end{equation}
Since a distant function is piecewise differentiable in general, the choice of the smallest in
absolute value enhances the numerical stability of the interpolation.

%-------------------------------------------------------------------------------
%
% Reinitialization
%
%-------------------------------------------------------------------------------
\section{Reinitialization Scheme\label{sec_reinitialization}}
In principle, the level function can be chosen as any Lipschitz continuous function. However, the
so-called signed distance function is known to produce more robust numerical results, to improve
mass conservation and to reduce errors in the computations of geometrical quantities such as the
interface curvatures. Sussman \etal proposed in \cite{sussmanfatemi:1998:redistancing} to evolve
the following partial differential equation to steady state in order to reinitialize a level set
function $\phi^0:\mathbb{R}^n\to\mathbb{R}$ into the signed distance function $\phi$:
\begin{equation}
\phi_{\tau} + \mathrm{sgn}(\phi^0)\left( |\nabla\phi|-1 \right)=0,
\label{eq_reinitialization}
\end{equation}
where $\tau$ represents the fictitious time. A standard discretization for
this equation in its semi-discrete form is given by:
\begin{equation}\frac{d\phi}{d\tau} + \mathrm{sgn}(\phi^0)
\left[ H_G\left(D_x^+\phi, D_x^-\phi,
                D_y^+\phi, D_y^-\phi\right)-1\right]=0,\label{Semi_Discrete}\end{equation}
where $\mathrm{sgn}(\phi^0)$ denotes the signum of $\phi^0$ and $H_G$ is the Godunov Hamiltonian
defined as:
\begin{equation*}
H_G(a,b,c,d)=
\begin{cases}
\sqrt{\max(|a^+|^2,|b^-|^2)+\max(|c^+|^2,|d^-|^2)} & \textrm{ if }
\mathrm{sgn}(\phi^0)\le 0 \\
\sqrt{\max(|a^-|^2,|b^+|^2)+\max(|c^-|^2,|d^+|^2)} & \textrm{ if } \mathrm{sgn}(\phi^0)>0
\end{cases},
\end{equation*}
with $a^+=\max(a,0)$ and $a^-=\min(a,0)$. The one-sided derivatives,
$D_x^{\pm}\phi$ and $D_y^{\pm}\phi$ are discretized by the second order
accurate one-sided finite differences defined in section
\ref{sec_finite_differences}. Equation (\ref{Semi_Discrete}) is evolved in
time with the TVD RK-2 method given in Shu and Osher \cite{Shu:1988:ENO}:
First define $\tilde{\phi}^{n+1}$ and $\tilde{\phi}^{n+2}$ by Euler steps
\begin{eqnarray*}
\frac{\tilde{\phi}^{n+1}-\phi^n}{\triangle \tau} &+& \mathrm{sgn}(\phi^0) \left[
H_G\left(D_x^+\phi^n, D_x^-\phi^n,
                D_y^+\phi^n, D_y^-\phi^n\right)-1\right]=0,\\
\frac{\tilde{\phi}^{n+2}-\tilde{\phi}^{n+1}}{\triangle \tau} &+& \mathrm{sgn}(\phi^0) \left[
H_G\left(D_x^+\tilde{\phi}^{n+1}, D_x^-\tilde{\phi}^{n+1},
                D_y^+\tilde{\phi}^{n+1}, D_y^-\tilde{\phi}^{n+1}\right)-1\right]=0,
\end{eqnarray*} and then define $\phi^{n+1}$ by averaging:
\begin{eqnarray*}
\phi^{n+1} &=& \frac{\phi^n + \tilde{\phi}^{n+2}}2.
\end{eqnarray*}

In order to preserve area/volume, the reinitialization procedure is
required not to move the original interface defined by $\phi_0$. In their
seminal work, Russo and Smereka \cite{Russo_Smereka} solved this problem by
simply including the initial interface location (given by $\phi_0$) in the
stencils of the one-sided derivatives. Consider the case depicted by figure
\ref{fig_one_dimensional_adaptive_grid} and suppose that
$\phi^0_0\cdot\phi^0_2<0$, i.e. the interface is located between the nodes
$v_0$ and $v_2$. The interface location $v_I$ can be calculated by finding
the root of the quadratic interpolation of $\phi^0$ on the interval
$\overline{v_0v_2}$ with the origin at the center of the interval:
\begin{equation*}
\phi^0(x)=c_2x^2+c_1x+c_2,
\text{  with }
\begin{cases}
c_2 &= \frac12\text{minmod}\left[D^0_{xx}\phi^0_0,D^0_{xx}\phi^0_2\right] \\
c_1 &= (\phi^0_2-\phi^0_0)/s_2 \\
c_0 &= (\phi^0_2+\phi^0_0)/  2 - c_2s_2^2/4
\end{cases}.
\end{equation*}
The distance $s_I$ between $v_0$ and the interface location is then defined by:
\begin{equation*}
s_I = \frac{s_2}2 +
\begin{cases}
-c_0/c_1                            & \text{ if $|c_2|<  \epsilon$ }\\
(-c_1+\sqrt{c_1^2-4c_2c_0})/(2c_2)  & \text{ if $|c_2|\ge\epsilon$ and $\phi^0_0<0$}\\
(-c_1-\sqrt{c_1^2-4c_2c_0})/(2c_2)  & \text{ if $|c_2|\ge\epsilon$ and $\phi^0_0>0$}
\end{cases}.
\end{equation*}
The calculation of $D_x^+\phi^n_0$ is then modified using the interface
location and the fact that $\phi=0$ at the interface:
\begin{equation*}
D^+_x\phi^n_0 = \frac{0-\phi_0^n}{s_I} -
\frac{s_I}{2}\text{minmod}\left(D^0_{xx}\phi^n_0,D^0_{xx}\phi^n_2\right).
\end{equation*}
We note that in the original work of Russo and Smereka
\cite{Russo_Smereka}, a cubic interpolation was employed to locate the
interface, but that the above quadratic interpolation with the minmod
operator acting on the second order derivatives proved to be more stable in
the case where the level set function presents a kink nearby. We also point
out that in the original work of \cite{Russo_Smereka}, the first order
derivative $D^+_x\phi^n_0$ was discretized as:
\begin{equation*}
D^+_x\phi^n_0 = \frac{0-\phi_0^n}{s_I} -
\frac{s_I}{2}\text{minmod}\left(D^0_{xx}\phi^n_0,D^0_{xx}\phi^n_I\right),
\end{equation*}
thus included $v_I$ in the discretization of $D^0_{xx}\phi^n_I$. However,
we found that this choice leads to unstable results when the interface is
close to grid nodes. We thus slightly changed the discretization by only
using the location of the interface in the first term in order to maintain
the location of $\phi^0$, not in the discretization of second order
derivatives. Likewise, in the case where $s_I$ is close to zero (hence
$\phi^0_0$ is close to zero) we simply set $\phi^n_0=0$ to guarantee
stability. This only introduces a negligible perturbation in the location
of the zero level set.

The same process is then applied to $D^-_x\phi$ if there is a sign change between $\phi^0_0$ and
$\phi^0_1$. The time step restriction for cells cut by the interface is then:
\begin{eqnarray}
\Delta \tau = \begin{cases}
\min(s_I,s_1,s_2)                   & \textrm{in $1D$}, \\
\min(s_I,s_1,s_2,s_3,s_4)/2         & \textrm{in $2D$}, \\
\min(s_I,s_1,s_2,s_3,s_4,s_5,s_6)/3 & \textrm{in $3D$}. \\
\end{cases}
\end{eqnarray}

\subsection{Adaptive time stepping}
We note that an adaptive time step is possible since only the steady state
of \eqref{eq_reinitialization} is sought. Since the time step restriction
is adapted for each cell, the reinitialization procedure is fast: small
cells with a stringent time step restriction are located near the interface
and therefore only a few iterations are required to reach the steady state
at those cells (characteristic information flow away \textit{from} the
interface); cells far away from the interface are large and therefore do
not require a small time step restriction. For example, consider the
example depicted in figure \ref{fig_ex_reinitialization_square}, for which
the level set function is defined initially as -1 inside a square domain
(not aligned with the grid cells) and +1 outside. This initial level set
function is therefore very far from the signed distance function that we
seek to define. However, on a grid where the smallest grid has size
$dx=1/2048$, the reinitialization procedure takes only 35 iterations to
fully converge to the signed distance function in the entire domain. In
practice, the initial level set is never that far to the signed distance
function and therefore only about 5 iterations are required regardless of
the resolution of the finest level. Figure
\ref{fig_reinitialization_error_decay} illustrates the difference in the
number of iterations required between uniform time stepping and adaptive
time stepping. In the case of a uniform time step, we take $\triangle
t=\triangle x_{\text{smallest}}/2$, with $\triangle x_{\text{smallest}}$
the size of the smallest cell.

\subsection{Third order accuracy}\label{subsec_Accuracy_Reinitialization}
We also computed the convergence rates of the reinitialization algorithm
for the test problem proposed in \cite{Russo_Smereka}: Consider the level
set function initially defined as:
$$\phi^0(x,y)=\left(0.1+(x-1)^2+(y-1)^2\right)\left(\sqrt{x^2+y^2}-1\right),$$ which defines the
interface as a circle with center the origin and radius $1$. In this case, $\phi^0$ is not a signed
distance function and its gradients vary widely. Figure \ref{fig_reinitialization} illustrates the
gradual deformation of the cross-sections of $\phi^0$ as it evolves to the signed distance
function. Table \ref{tab_reinitialization} illustrates that the method is third order accurate in
the $L^1$ and $L^\infty$ norms near the interface, where we use the standard formulas for the $L^1$
and $L^\infty$ norms:
\begin{align*}
||\phi||_{\infty}=\max_{v:|\phi(v)|<1.2\Delta x}|\phi(v)-\phi_{\textrm{exact}}(v)|, \\
||\phi||_{1}=\avg_{v:|\phi(v)|<1.2\Delta x}|\phi(v)-\phi_{\textrm{exact}}(v)|, \\
\end{align*}
where $\Delta x=\Delta x_{\textrm{smallest}}$. Note that after the reinitializing the level set
function, the choice of $v:|\phi(v)|<1.2\Delta x$ ensures the selection of all the nodes adjacent
to the interface.

In the entire domain, the method is second order accurate if we keep refining all the cells. In the
practical case where only cells near the interface are refined, the accuracy in regions far away
from the interface is meaningless. In the case where the interface presents sharp corners, the
accuracy is reduced to first order in the $L^{\infty}$ norm.

%\begin{figure}[t]
%\centering
%\includegraphics[scale=0.2]{initial_phi.eps}
%\includegraphics[scale=0.2]{final_phi.eps}
%\\
%\includegraphics[scale=0.2]{initial_cross_section.eps}
%\includegraphics[scale=0.2]{final_cross_section.eps}
%\caption{Reinitialization procedure. Left: Initial level set function (top) and its zero cross
%section (bottom) defining a square domain. Right: Reinitialized level set function (top) and its
%zero cross section (bottom). In particular, the difference in the zero level set between the
%initial and final stages is negligible. In this example, the level difference between adjacent
%cells is not restricted.} \label{fig_ex_reinitialization_square}
%\end{figure}

\begin{figure}[t]
\centering
\includegraphics[scale=0.2]{losange_step.eps}
\includegraphics[scale=0.2]{losange_reinitialized.eps}
\\
\includegraphics[scale=0.2]{losange_initial.eps}
\includegraphics[scale=0.2]{losange_final.eps}
\caption{Reinitialization procedure. Left: Initial level set function (top)
and its zero cross section (bottom) defining a square domain. Right:
Reinitialized level set function (top) and its zero cross section (bottom).
In particular, the difference in the zero level set between the initial and
final stages is negligible. In this example, the level difference between
adjacent cells is not restricted.} \label{fig_ex_reinitialization_square}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[scale=1.0]{reinitialization_error_decrease.eps}
\end{center}
\caption{$L_1$ errors of the reinitialization algorithm in the case of the
adaptive time step (solid line) and the uniform time step (dotted line).}
\label{fig_reinitialization_error_decay}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[scale=0.8]{reinitialization_initial.eps}
\includegraphics[scale=0.8]{reinitialization_5_iterations.eps} \\
\includegraphics[scale=0.8]{reinitialization_10_iterations.eps}
\includegraphics[scale=0.8]{reinitialization_20_iterations.eps}
\end{center}
\caption{From top-left to bottom-right: Contours of the reinitialized level
set function of example \ref{subsec_Accuracy_Reinitialization} after 0, 5,
10 and 20 iterations. The contours are evenly plotted from -1 to 1 with a
thick line representing the zero contour.} \label{fig_reinitialization}
\end{figure}

\begin{table}
\begin{center}
\begin{tabular}{|c|c|c|c|c|c|c|c|}\hline
\multicolumn{3}{|c|}{} & \multicolumn{5}{|c|}{Finest Resolution}   \\\cline{4-8}
\multicolumn{3}{|c|}{} & $128^2$ & rate & $256^2$ & rate & $512^2$ \\\hline \multirow{4}{0.7
in}{Uniform Refinement} & \multirow{2}{*}{Near Interface}
 & $L_1     $ & $4.36\times10^{-6}$ & 2.92 & $5.77\times10^{-7}$ & 3.02 & $7.12\times10^{-8}$ \\\cline{3-8}
&& $L_\infty$ & $2.16\times10^{-5}$ & 2.74 & $3.24\times10^{-6}$ & 3.26 & $3.38\times10^{-7}$
\\\cline{2-8}
                                       & \multirow{2}{*}{Whole Domain}
 & $L_1     $ & $3.27\times10^{-4}$ & 2.14 & $7.42\times10^{-5}$ & 2.11 & $1.71\times10^{-5}$ \\\cline{3-8}
&& $L_\infty$ & $4.20\times10^{-2}$ & 1.56 & $1.43\times10^{-2}$ & 1.87 & $3.89\times10^{-3}$
\\\hline

\multirow{4}{0.7 in}{Adaptive Refinement} & \multirow{2}{*}{Near Interface}
 & $L_1     $ & $4.36\times10^{-6}$ & 2.94 & $5.70\times10^{-7}$ & 3.00 & $7.14\times10^{-8}$ \\\cline{3-8}
&& $L_\infty$ & $2.16\times10^{-5}$ & 2.87 & $2.96\times10^{-6}$ & 3.09 & $3.48\times10^{-7}$
\\\cline{2-8}
                                         & \multirow{2}{*}{Whole Domain}
 & $L_1     $ & $3.27\times10^{-4}$ & 1.06 & $1.57\times10^{-4}$ & 1.01 & $7.82\times10^{-5}$ \\\cline{3-8}
&& $L_\infty$ & $4.20\times10^{-2}$ & 0.00 & $4.20\times10^{-2}$ & 0.00 & $4.20\times10^{-3}$
\\\hline
\end{tabular}
\end{center}
\caption{Convergence rates for the reinitialization for example
\ref{subsec_Accuracy_Reinitialization}. The initial grid is shown in figure
\ref{fig_reinitialization}. The condition for a node $v_i$ to be
'\textit{near interface}' is chosen as $|\phi(v_i)|<\sqrt{2}\triangle
x_{\text{smallest}}$, where $\triangle x_{\text{smallest}}$ is the size of
the smallest cell. The '\textit{whole domain}' excludes the region near the
kink located at the origin, where accuracy drops to first order.}
\label{tab_reinitialization}
\end{table}

%-------------------------------------------------------------------------------
%
% semi-Lagrangian Method
%
%-------------------------------------------------------------------------------
%\section{Semi-Lagrangian Method\label{sec_semiL}}
%-------------------------------------------------------------------------------
%
% Moving Interface with Given Velocity
%
%-------------------------------------------------------------------------------
\section{Motion Under an Externally Generated Velocity Field}\label{sec_semi_lagrangian}
\subsection{Second Order Accurate Semi-Lagrangian Method\label{Sec_midpoint_rule}}
In the case where the velocity field is externally generated, the level set
equation \eqref{eq_levelset} is linear. In this case, one can use
semi-Lagrangian methods. Semi-Lagrangian schemes are extensions of the
Courant-Isaacson-Rees \cite{Courant:1952:CIR_Method} method for hyperbolic
equations and are unconditionally stable thus avoiding standard CFL
condition of $\Delta t \approx \Delta x_{\rm smallest}$. The general idea
behind semi-Lagrangian methods is to reconstruct the solution by
integrating numerically the equation along characteristic curves, starting
from any grid point $x_i$ and tracing back the departure point $x_d$ in the
upwind direction. Interpolation formulas are then used to recover the value
of the solution at such points. In this work, we use a second order
accurate Semi-Lagrangian method.

Consider the linear advection equation:
\begin{eqnarray}
\phi_t + U\cdot\nabla\phi&=&0, \label{Level_Set_Equation}
\end{eqnarray}
where $U$ is an externally generated velocity field. Then
$\phi^{n+1}(x^{n+1})=\phi^n(x_d)$, where $x^{n+1}$ is any grid node and
$x_d$ is the corresponding departure point from which the characteristic
curve originates. In this work, we use the second order mid-point method
for locating the departure point, as in \cite{Xiu:2001:Semi_Lagrangian_BD}:
$$ \begin{array}{rcl}
\hat{x} &=& x^{n+1} - \frac{\Delta t}2\cdot U^n(x^{n+1}), \\
x_d     &=& x^{n+1} -       \Delta t  \cdot U^{n+\frac12}(\hat{x}),
\end{array}
$$
where we define the velocity at the mid-time step $t^{n+\frac12}$ by a
linear combination of the velocities at the two previous time steps, i.e.
$U^{n+\frac12}=\frac32U^n-\frac12U^{n-1}$. Since $\hat{x}$ and $x_d$ are
not on grid nodes in general, interpolation procedures must be applied to
define $U^{n+\frac12}(\hat{x})$ and $\phi^n(x_d)$. We note that it is
enough to define $U^{n+\frac12}(\hat{x})$ with a multilinear interpolation
\eqref{Bilinear_Interpolation} and $\phi^n(x_d)$ with the quadratic
interpolation described by equations \eqref{Stabilized_Interpolation1} and
\eqref{Stabilized_Interpolation2}: Since a distance function has
discontinuities in its derivative in general, the stabilized quadratic
interpolation is preferred to the Hermite quadratic interpolation.

%-------------------------------------------------------------------------------
% Example : 2D Rotation Test
%-------------------------------------------------------------------------------
\subsection{Test: Rotation in 2D \label{ex_rotation_test_2D}}

Consider a domain $\Omega=[-1, 1]^2$ and a disk of radius $R=.15$ and center initially at $(0,
.75)$, rotating under the divergence free velocity field
$$
\begin{array}{lcr}
u(x,y) &=& -y \\
v(x,y) &=& x \\
\end{array}
$$
The final time $t=2\pi$ is the time when the rotation completes one
revolution. In the simulation, the adaptive refinement is used, and the
time step restriction is $\Delta t= 5\Delta x$. Table
\ref{tab_ex_rotation_test_2D} demonstrates second order accuracy for the
level set as well as for the mass conservation.  We note that we only
consider the grid nodes neighboring the interface in our computation of the
accuracy for the level set function $\phi$ since only those points define
the location of the interface.

\begin{table}
\begin{center}
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline
Finest &$L^\infty$ error&rate&$L^1 error$&rate&Loss of&rate \\
Resolution& of $\phi$      &    & of $\phi$ &    &volume(\%)&    \\\hline
$32^2$&$7.24\times10^{-2}$&    &$3.11\times10^{-2}$&    &28.51&  \\\hline
$64^2$&$1.78\times10^{-2}$&2.02&$8.86\times10^{-3}$&1.81&7.21 & 1.98   \\\hline
$128^2$&$4.52\times10^{-3}$&1.97&$2.13\times10^{-4}$&2.05&1.78& 2.01   \\\hline
$256^2$&$1.13\times10^{-3}$&1.99&$5.56\times10^{-4}$&1.93&0.45& 1.98   \\\hline
$512^2$&$2.85\times10^{-4}$&2.00&$1.38\times10^{-4}$&2.01&0.11& 2.03   \\\hline
$1024^2$&$7.14\times10^{-5}$&2.00&$3.46\times10^{-5}$&2.00&0.03& 1.87   \\\hline
$2048^2$&$1.78\times10^{-5}$&2.00&$8.64\times10^{-6}$&2.00&0.007& 2.01   \\\hline
\end{tabular}
\end{center}
\caption{Convergence rates for example \ref{ex_rotation_test_2D}
\label{tab_ex_rotation_test_2D}}
\end{table}

%-------------------------------------------------------------------------------
% Example : Vortex Test Test
%-------------------------------------------------------------------------------
\subsection{Test: Vortex in 2D \label{ex_large_deformation_test_2D}}
In this example, we test our level set implementation on the more
challenging flow proposed by Bell \etal \cite{Bell_Colella_Glaz}: Consider
a domain $\Omega=[0, 1]^2$ and a disk of radius $.15$ and center $(.5,
.75)$ as the initial zero level set contour. The level set is then deformed
under the divergence free velocity field $U=(u,v)$ given by:
$$
\begin{array}{rcr}
u(x,y) &=& -\sin^2(\pi x)\sin(2\pi y) \\
v(x,y) &=&  \sin^2(\pi y)\sin(2\pi x) \\
\end{array}
$$
The disk is deformed forward until $t=1$ and then backward to the original
shape using the reverse velocity field with a time step restriction of
$\triangle t=5\cdot\triangle x_{\text{smallest}}$.

\begin{table}
\begin{center}
\begin{tabular}{|c|c|c|c|c|c|c|c|}\hline
\multirow{2}{*}{} & \multicolumn{7}{|c|}{Finest Resolution}   \\\cline{2-8}
                  & $64^2$ & rate & $128^2$ & rate & $256^2$ & rate & $512^2$ \\\hline
$L_1     $ error of $\phi$&$9.58\TT{-3}$&2.80&$1.38\TT{-3}$&2.02&$3.41\TT{-4}$&2.08&$8.09\TT{-5}$
\\\hline $L_\infty$ error of
$\phi$&$1.83\TT{-2}$&1.57&$6.17\TT{-3}$&1.08&$2.91\TT{-3}$&1.39&$1.11\TT{-3}$ \\\hline Volume Loss
&$4.48       $&2.36&$0.874      $&1.40&$0.331      $&1.79&$0.0954     $ \\\hline Max number of
nodes   &$1045       $&1.10&$2243       $&1.10&$4815       $&1.09&$10256      $ \\\hline Min number
of nodes   &$439        $&1.07&$924        $&0.99&$1831       $&1.01&$3679       $ \\\hline
time(sec) & 1.420 & 2.29 & 6.96 & 2.17 &31.4 & 2.13 & 138 \\
\hline minimum memory(MB) & 0.0448 & 1.07 & 0.0943 & .97 & 0.185 & 1.00 & 0.371 \\
\hline maximum memory(MB) & 0.101 & 1.30 & 0.248 & 1.06 & 0.517 & 1.03 & 1.06 \\ \hline
\end{tabular}
\end{center}
\caption{Convergence rates for example \ref{ex_large_deformation_test_2D}. The memory
requirement increases linearly with effective resolution since most of the computational resources
is focused near the one-dimensional interface, i.e. our method is an efficient implementation of
local level set methods. The computational time increases quadratically with effective resolution,
since the number of nodes is doubled and the time step is halved.} \label{tab_vortex}
\end{table}

Table \ref{tab_vortex} demonstrate second order accuracy for $L_1$ error of
$\phi$ and volume of loss, and linear increase in the maximum/minimum
number of nodes. Note that the uniform grid of resolution $512^2$ requires
about 25 times more nodes than the adaptive grid with the same resolution.
Although second order accuracy was achieved in both the maximum and average
norms in the previous example, the convergence rate of the maximum error is
oscillating between one and two. This is due to the fact that as the
interface deforms, some part of the interface are under resolved. Figure
\ref{fig_vortex_test} illustrates this at $t=1$: Here, the tail of the
interface is not resolved accurately. This deterioration in accuracy was
also reported in \cite{Olsson_Kreiss:2005}.

\begin{figure}
\begin{center}
\includegraphics[scale=0.7]{vortex_flow_at_one.eps}
\includegraphics[scale=0.7]{vortex_flow_at_two.eps}
%%includegraphics[scale=0.7]{volume_loss.eps}
\end{center}
\caption{Contours of the zero level sets for example
\ref{ex_large_deformation_test_2D} with effective resolutions of $64^2$,
$128^2$, $256^2$ and $512^2$ at $t=1$ (left) and $t=2$ (right). The colors
red, green, blue represent the difference in the interface location between
the resolutions $64^2$ and $128^2$, $128^2$ and $256^2$, $256^2$ and
$512^2$, respectively.} \label{fig_vortex_test}
\end{figure}

Figure \ref{large_deformation_test_2D} illustrates the evolution of the
interface location initially (left), at $t=6$ (center) and when the
interface is fully rewinded (right). This example illustrates the ability
of the present method to accurately capture the evolution of an interface
undergoing large deformations and the ability to preserve mass effectively
(mass loss $\approx.3\%$).

\begin{figure}
\begin{center}
\includegraphics[scale=0.25]{Large_Deformation_2D_T6_Initial.eps}
\includegraphics[scale=0.25]{Large_Deformation_2D_T6_Mid.eps}
\includegraphics[scale=0.25]{Large_Deformation_2D_T6_Final.eps}
\end{center}
\caption{Level set evolution at $t=0$ (left), $t=3$ (center) and $t=6$ (right). The effective
resolution is $2048^2$ and the mass is conserved within .3\%} \label{large_deformation_test_2D}
\end{figure}

%-------------------------------------------------------------------------------
% Example : 2D Zalesak's Disk
%-------------------------------------------------------------------------------
%\subsection{Zalesak's Disk in 2D \label{ex_zalesak_2d}}
%In this example, we test our method on the Zalesak's disk problem in 2D. Consider a domain $\Omega=[0, 100]^2$ and a disk of radius $15$ and center
%$(50, 75)$ that is slotted of width $5$ and height $25$ as the initial zero level set contour. The slotted disk is rotated around $(50,50)$ with velocity field $\frac\pi{314}(50-y,x-50)$. After one revolution that happens when $T=628$, the loss of volume was measured in Table \ref{tab_ex_zalesak_2d}. The volume losses of 0.8\% in $128^2$ and 0.26\% in $256^2$ are similiar to the results of Particle Level Set method in \cite{En} of 0.31\% in $100^2$ and 0.20\% in $200^2$.

%\begin{table}
%\begin{center}
%\begin{tabular}{|c|c|c|}
%\hline
%The most    & Loss of & rate \\
%refined grid& Volume  &      \\\hline
%$ 64^2$&4.60  \% & 2.51  \\\hline
%$128^2$&0.808 \% & 2.51  \\\hline
%$256^2$&0.259 \% & 1.64  \\\hline
%$512^2$&0.0468\% & 2.47  \\\hline
%\end{tabular}
%\end{center}
%\caption{Convergence rate for example \ref{ex_zalesak_2d} \label{tab_ex_zalesak_2d}}
%\end{table}

%-------------------------------------------------------------------------------
% Example : Reinitialization_3D
%-------------------------------------------------------------------------------
%\subsection{Reinitialization in 3D \label{ex_reinitialization_3d}}
%In this section, we test the convergence rate of the reinitialization algorithm in octree. In $[0,1]^3$, a signed distance function to the sphere of center $(0.35,0.35,0.35)$ and of radius $0.15$ is reinitialized in an octree depicted in Figure \ref{fig_ex_reinitialization_3d}. In the next refinement, the automatic grid generation algorithm is used. Table \ref{tab_ex_reinitialization_3d} shows that the reinitialization method is 3rd order accurate. As shown in figure \ref{fig_ex_reinitialization_3d}, the reinitialization in a coarse grid of $16^3$ could not be noticed visually.

%\begin{figure}[t]
%\centering
%\includegraphics[scale=0.4]{octree_before_reinitialization.eps}
%\includegraphics[scale=0.4]{octree_after_reinitialization.eps}
%\caption{Level sets on a octree before and after the reinitialization in $16^3$ grid for Example \ref{ex_reinitialization_3d}}
%\label{fig_ex_reinitialization_3d}
%\end{figure}

%\begin{table}
%\begin{center}\begin{tabular}{|c|c|c|c|c|c|c|c|}
%\hline
%Effective  &&&&&&&\\
%Resolution & Error in $L_\infty$ & Order & Error in $L_1$ & order & Volume Loss & \#nodes &\#cells %\tabularnewline \hline
%$  32^3$ & $3.63\times 10^{-4}$ &      & $1.26\times 10^{-4}$ &      & - 2.1\% & 3541 & 3137 \tabularnewline \hline
%$  64^3$ & $9.89\times 10^{-5}$ & 1.87 & $1.54\times 10^{-5}$ & 3.03 & -0.54\% & 12059 & 11257 \tabularnewline \hline
%$ 128^3$ & $3.18\times 10^{-5}$ & 1.63 & $2.22\times 10^{-6}$ & 2.79 & -0.13\% & 44955 & 42777 \tabularnewline \hline
%$ 256^3$ & $3.46\times 10^{-6}$ & 3.19 & $2.59\times 10^{-7}$ & 3.09 & -0.034\% & 176236 & 168881 \tabularnewline \hline
%$ 512^3$ & $5.05\times 10^{-7}$ & 2.77 & $3.36\times 10^{-8}$ & 2.94 & -0.0084\% & 701711 & 674089 \tabularnewline \hline
%\end{tabular}
%\end{center}
%\caption{Convergence rate of the reinitialization algorithm in 3D for  example \ref{ex_reinitialization_3d}
%\label{tab_ex_reinitialization_3d}}
%\end{table}

%-------------------------------------------------------------------------------
% Example : 3D Rotation Test
%-------------------------------------------------------------------------------
\subsection{Test: Rotation in 3D \label{ex_rotation_test_3D}}

Consider a domain $\Omega=[-2, 2]^3$ and a sphere of radius $R=.5$ and
center initially at $(0,1,0)$, rotating under the divergence free velocity
field
$$
\begin{array}{lcr}
u(x,y,z) &=& -y \\
v(x,y,z) &=& x  \\
w(x,y,z) &=& 0
\end{array}
$$
The simulation is run until $t=2\pi$, when the rotation completes one
revolution. In the simulation, the adaptive refinement is used, and the
time step restriction is $\Delta t=6\Delta x_{\rm smallest}$. Table
\ref{tab_ex_rotation_test_3D} demonstrates second order accuracy for the
level set as well as for the mass conservation.  We note that we only
consider the grid nodes neighboring the interface in our computation of the
accuracy for the level set function $\phi$. Figure
\ref{fig_rotating_sphere} shows the adaptive grid for the rotating sphere.

\begin{table}
\begin{center}
\begin{tabular}{|c|c|c|c|c|c|c|c|}\hline
\multirow{2}{*}{} & \multicolumn{7}{|c|}{Finest Resolution}\\\cline{2-8}
                  & $32^3$ & rate & $64^3$ & rate & $128^3$ & rate & $256^3$ \\\hline
$L_1$      error of $\phi$ & $6.86\TT{-2}$ & 1.88 & $1.87\TT{-2}$ & 1.99 & $4.70\TT{-3}$ & 1.99 &
$1.18\TT{-3}$ \\\hline $L_\infty$ error of $\phi$ & $1.76\TT{-1}$ & 2.02 & $4.35\TT{-2}$ & 2.02 &
$1.07\TT{-2}$ & 2.02 & $2.65\TT{-3}$ \\\hline Volume loss(\%)            & 23.1 & 2.16 & 5.14 &
2.12 & 1.18 & 2.07 & 0.282 \\\hline
\end{tabular}
\end{center}
\caption{Convergence rates for the interface's location for example \ref{ex_rotation_test_3D}.}
\label{tab_ex_rotation_test_3D}
\end{table}

\begin{figure}
\begin{center}
\includegraphics[scale=0.8]{rotating_sphere_initial.eps}
\includegraphics[scale=0.8]{rotating_sphere_a_quarter.eps}\\
\includegraphics[scale=0.8]{rotating_sphere_a_half.eps}
\includegraphics[scale=0.8]{rotating_sphere_final.eps}
\end{center}
\caption{Evolution of the interface for example \ref{ex_rotation_test_3D}: Initial data (top-left),
interface after a quarter turn (top-right), interface after a half turn (bottom-left) and final
location (bottom-right). The finest resolution is $128^3$.} \label{fig_rotating_sphere}
\end{figure}

%-------------------------------------------------------------------------------
% Example : Enright's Test in 3D
%-------------------------------------------------------------------------------
\subsection{Enright's Test in 3D \label{ex_enright_test}}
We consider the test proposed in Enright \etal \cite{Enright:2002:PLS}: A
sphere of center $(0.35,0.35,0.35)$ and radius $0.15$ in the domain of
$[0,1]^3$ is deformed under the following divergence free velocity field:
$$
\begin{array}{lcr}
u(x,y,z) &=& 2\sin^2(\pi x)\sin(2\pi y)\sin(2\pi z) \\
v(x,y,z) &=& -\sin^2(\pi y)\sin(2\pi x)\sin(2\pi z) \\
w(x,y,z) &=& -\sin^2(\pi z)\sin(2\pi x)\sin(2\pi y)
\end{array}
$$
forward in time and then backward to its original shape with the reversed velocity. Figure
\ref{fig_enrights_test_until_one} illustrates the interface motion with a time step restriction of
$\triangle t=5\cdot\triangle x_{\text{smallest}}$. We note that, in the simulation with an
effective resolution of $512^3$, minimum the number of nodes used was $685220$ and the maximum was
$2606710$. In contrast, the number of nodes in the case of a uniform grid with the same resolution,
the number of nodes would be about 50 times larger. The volume loss is 3.21\% for an effective
resolution of $256^3$ and 0.739\% for an effective resolution of $512^3$. Figure
\ref{fig_enrights_test_refinement} compares the interface evolution with an effective resolution of $128^3$, $256^3$ and $512^3$. Table \ref{Enright_Table1} describes the memory and CPU requirements, and table \ref{Enright_Table2} describes the volume loss and the accuracy of the interface location after reconvering the original shape.

The Enright's test is a canonical example to test the amount of numerical dissipation of level set
methods. The particle level set method reported 2.6\% volume loss on a $100^3$ uniform grid
together with Lagrangian particles \cite{Enright:2002:PLS}. Our results show that we obtain a loss
of mass of $.74\%$ in the case of a $512^3$ effective resolution, which corresponds to a $137^3$
uniform grid in term of number of nodes. The particle level set was further improved in
\cite{Enright:2002:AAR} using octree data structures in addition to particles. Although
\cite{Enright:2002:AAR} does not report any quantitative results, we find that our result for the
Enright's test is visually comparable to that obtained in \cite{Enright:2002:AAR} for the same
effective resolution and compares favorably with the results in \cite{Hieber:Particle_Level_Set}.

\begin{table}
\begin{center}\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|}
  \hline
  Finest     &\multirow{2}{*}{Time}
             &\multirow{2}{*}{Rate}
             & Min \#
             &\multirow{2}{*}{Rate}
             & Max \#
             &\multirow{2}{*}{Rate}
             & Min
             &\multirow{2}{*}{Rate}
             & Max
             &\multirow{2}{*}{Rate}  \\
  Resolution & & & nodes & & nodes & & Memory & & Memory &  \\  \hline
  $128^3$ & 237.5 &  & 44943 &  & 133308 &  & 4.54 &  & 13.7 &  \\  \hline
  $256^3$ & 2214 & 3.23 & 173637 & 1.95 & 598264 & 2.17 & 17.6 & 1.96 & 61.5 & 2.17 \\  \hline
  $512^3$ & 19521 & 3.14 & 685220 & 1.98 & 2606710 & 2.12 & 69.6 & 1.98 & 268 & 2.12 \\  \hline

  \end{tabular}\caption{Memory and CPU requirements for example \ref{ex_enright_test}. The memory
requirement increases quadratically with effective resolution since most of the computational
resources is focused near the two-dimensional interface, i.e. our method is an efficient
implementation of local level set methods. The computational time increases cubically with
effective resolution, since the number of nodes is multiplied by four and the time step is halved.}
  \label{Enright_Table1}
  \end{center}  \end{table}

\begin{table}
\begin{center}
\begin{tabular}{|c|c|c|c|c|c|c|}
  \hline
  Finest Resolution & Volume Loss & Rate & $L^1     $ error of $\phi$ & Rate
                                         & $L^\infty$ error of $\phi$ & Rate \\\hline
  $128^3$ & 16.02\% &  & $1.96 \times 10^{-2}$ &  & $1.54 \times 10^{-1}$&\\\hline
  $256^3$ & 3.21\% & 2.32 & $2.83 \times 10^{-3}$ & 2.79 & $1.06 \times 10^{-1}$& .88\\\hline
  $512^3$ & .739 \% & 2.12 & $4.38 \times 10^{-4}$ & 2.69 & $5.74 \times 10^{-3}$& 7.52\\\hline
\end{tabular}\caption{Convergence rates for the interface's location for example \ref{ex_enright_test}.}
\label{Enright_Table2}
\end{center}\end{table}

We note that the jump in rate in table \ref{Enright_Table2} can be explained by the lack of
resolution for describing the developing thin film. This is related to the Nyquist-Shannon sampling
theorem that states that in order to fully reconstruct a signal the sampling frequency should be at
least twice the signal bandwidth. In our case, the fact that the thin film is under-resolved
prevents subcell resolution of the reinitialization scheme. For higher resolutions, we would expect
second-order accuracy.

\begin{figure}
\begin{center}
\includegraphics[scale=0.3]{enrights_test_512_until_1.eps}
\end{center}
\caption{Evolution of the interface for the Enright's test with finest resolution of $512^3$.}
\label{fig_enrights_test_until_one}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[scale=0.3]{image_0001_resolution_128.eps}\includegraphics[scale=0.3]{image_0001_resolution_256.eps}\includegraphics[scale=0.3]{image_0001_resolution_512.eps}\\
\includegraphics[scale=0.3]{image_0002_resolution_128.eps}\includegraphics[scale=0.3]{image_0002_resolution_256.eps}\includegraphics[scale=0.3]{image_0002_resolution_512.eps}\\
\end{center}
\caption{Effect of refinement on the Enright's test: The top figures correspond to the interface
fully stretched and the bottom figures correspond to the interface rewinded to the original sphere.
The finest resolutions are $128^3$ (left), $256^3$ (center) and $512^3$ (right).}
\label{fig_enrights_test_refinement}
\end{figure}

%-------------------------------------------------------------------------------
%
% Moving Interface with Curvature Dependent Speed
%
%-------------------------------------------------------------------------------
\section{Motion in the Normal Direction and Curvature Driven Flow}
The equation describing an interface propagating in its normal direction and under its mean
curvature is given by \cite{osher:1988:originallevelset}:
\begin{equation}
\phi_t+(\alpha-\beta\kappa)|\nabla\phi|=0, \label{Motion_Normal_Curvature}
\end{equation}
where $\kappa$ is the mean curvature of the interface $\kappa=\nabla\cdot\left(
\nabla\phi/|\nabla\phi| \right)$. The coefficients $\alpha$ and $\beta\ge0$ control the magnitude
of the speed in the normal direction and the strength of the curvature dependence, respectively.
The case where $\beta<0$ is ill-posed and therefore we do not consider it here.

\subsection{Motion in the Normal Direction}
First, we discuss the case when $\beta=0$. Using the second order one-sided derivatives described
in section \ref{sec_finite_differences} and discretizing the Hamiltonian using a Godunov scheme, we
semi-discretize the equation as :
$$\frac{d\phi}{dt} + \alpha\cdot H_G(\phi)=0$$
where the Godunov Hamiltonian $H_G$ is defined as:
$$H_G(\phi) =
\begin{cases}
\sqrt{\max(|(D_x^+\phi)^-|^2,|(D_x^-\phi)^+|^2)+\max(|(D_y^+\phi)^-|^2,|(D_y^-\phi)^+|^2)}&\text{if $\alpha>0$}\\
\sqrt{\max(|(D_x^+\phi)^+|^2,|(D_x^-\phi)^-|^2)+\max(|(D_y^+\phi)^+|^2,|(D_y^-\phi)^-|^2)}&\text{otherwise}
\end{cases}$$
This equation is discretized in time using the second order TVD Runge-Kutta method (see
\cite{Shu:1988:ENO,liu:1996:WENO}):
\begin{align}
\frac{\tilde{\phi}^{n+1}-       \phi ^n}{\triangle t} &+ \alpha\cdot H_G\left(\phi^n\right)=0 \\
\frac{\tilde{\phi}^{n+2}-\tilde{\phi}^n}{\triangle t} &+ \alpha\cdot H_G\left(\tilde{\phi}^{n+1}\right)=0 \\
\phi^{n+1}&=\frac{\phi^n+\tilde{\phi}^{n+2}}2
\end{align}

Table \ref{tab_shrinking_circle} illustrates that the method described above is second order
accurate in both the maximum and the average norms for smooth data. In the case where the interface
presents sharp corners, figure \ref{fig_shrinking_and_expading_square} illustrates that the method
converges to the correct viscosity solution \cite{osher:1988:originallevelset}.

\begin{table}
\begin{center}
\begin{tabular}{|c|c|c|c|c|c|c|c|}\hline
\multirow{2}{*}{} & \multicolumn{7}{|c|}{Finest Resolution}\\\cline{2-8}
                  & $64^2$ & rate & $128^2$ & rate & $256^2$ &rate & $512^2$ \\\hline
$L_1$      error of
$\phi$&$1.46\TT{-3}$&2.02&$3.58\TT{-4}$&2.00&$8.56\TT{-4}$&1.98&$2.26\TT{-5}$\\\hline $L_\infty$
error of $\phi$&$2.77\TT{-3}$&2.04&$6.72\TT{-4}$&1.95&$1.73\TT{-4}$&1.99&$4.36\TT{-5}$\\\hline
\end{tabular}
\end{center}
\caption{Convergence rate for a circle shrinking with unit normal velocity. consider a domain of
$[-2,2]^2$ and an interface initially described by a circle centered at the origin with radius
$R=1$. The interface is evolved until $t=0.5$.} \label{tab_shrinking_circle}
\end{table}
\begin{figure}
\begin{center}
\includegraphics[scale=0.5]{initial_square_new.eps}
\includegraphics[scale=0.5]{shrinking_square_point_5.eps}
\includegraphics[scale=0.5]{shrinking_square_1.eps}\\
\includegraphics[scale=0.5]{initial_square_new.eps}
\includegraphics[scale=0.5]{expanding_square_point_5.eps}
\includegraphics[scale=0.5]{expanding_square_1.eps}
\end{center}
\caption{Shrinking square in the first row, and expanding square in the second row}
\label{fig_shrinking_and_expading_square}
\end{figure}

\subsection{Adding Motion by Mean Curvature}
Now we discuss the case when $\beta>0$. The curvature term can be
discretized explicitly or implicitly. In the case where the curvature term
is discretized explicitly, the corresponding time step restriction of
$\Delta t \approx \Delta x^2$ is too stringent to be practical since it
would be constrained by the size of the smallest grid cell in the grid. In
\cite{Smereka:2003:Curvature}, Smereka proposed an implicit discretization
of the curvature term in the case of uniform grids: Using the following
operator splitting:
$$\kappa|\nabla\phi|=\Delta \phi -
\frac{\nabla\phi}{|\nabla\phi|}\nabla(|\nabla\phi|),$$ equation \eqref{Motion_Normal_Curvature} is
discretized as:
$$\frac{\phi^{n+1}-\phi^n}{\triangle t} +\alpha H_G(\phi^n) = \beta\Delta \phi^{n+1}
- \beta\frac{\nabla\phi^n}{|\nabla\phi^n|}\cdot\nabla(|\nabla\phi^n|).$$

In this work, we used a Backward Euler step to treat the linear term, and a
Forward Euler step for the nonlinear term. The derivatives $\Delta$ and
$\nabla$ are discretized by the central finite differences described in
section \ref{sec_finite_differences}. Discretizing implicitly the Laplacian
requires a linear system that we solve using the supra convergent method
presented in Min, Gibou and Ceniceros \cite{Min::Poisson_AMR_Second_Order}.
As noted in \cite{Smereka:2003:Curvature}, the semi-implicit discretization
on the curvature term allows for a big time step, so that the time step
restriction is that of the convection part, i.e.
\begin{equation}
\Delta t=\frac{\Delta x_{\text{smallest}}}{\alpha\cdot\text{\# dimensions}},
\end{equation} where $\text{\# dimensions}$ is the number of dimensions.

Table \ref{tab_circle_motion_by_curvature} demonstrates that the method is
first order accurate in the average norm for smooth a interface. The
deterioration in the maximum norm probably comes from the Elliptic part of
the solver, which propagates the errors from the regions where the grid
cells are coarse and unrefined to the regions where the grid cells are
refined. Figure \ref{fig_curvature_flow_in_octree} illustrates the motion
of an interface under mean curvature for the example presented in
\cite{Smereka:2003:Curvature}.

%So it is a first order discretization. One can go further for a better accuracy. However there is
%the following issue in this discretization. $\phi^{n+1}$ is calculated by solving a Poisson-type
%equation. As discussed in \cite{Min::Poisson_AMR_Second_Order}, the Poisson solver is second order
%accurate with the uniform refinement, i.e. splitting every cell. In level set method, it is natural
%to put the most resource near interface. So that in solving the Poisson-type equation, the errors
%in the coarse grids, i.e. the region far away from interface contaminates the region near
%interface, since the solution of one point affects the whole domain in Poisson equation. The
%accuracy of the Poisson solver in \cite{Min::Poisson_AMR_Second_Order} has not been analyzed in the
%adaptive refinement.

%\begin{figure}
%\begin{center}
%\includegraphics[scale=0.25]{curvature_flow_quadtree.eps}
%\end{center}
%\caption{$\alpha=1,\beta=1$ in $256^3$ resolution. From top-left to bottom-right, the times are $0$, $0.039$, %$0.107$, $0.507$, $0.800$ and $1.03$. The CFL condition is $\Delta t=\frac{\Delta x_{\text{smallest}}}2$. %Note that the velocities of convection and curvature cancel out not to move the side parts of the two circles %for some small time. }
%\label{fig_curvature_flow_in_quadtree}
%\end{figure}

\begin{table}
\begin{center}
\begin{tabular}{|c|c|c|c|c|c|c|c|}\hline
\multirow{2}{*}{} & \multicolumn{7}{|c|}{Finest Resolution}\\\cline{2-8}
                  & $128^2$ & rate & $256^2$ &rate & $512^2$ & rate &$1024^2$ \\\hline
$L_1$      error of $\phi$&$5.22\TT{-3}$&1.00&$2.60\TT{-3}$&0.96&$1.33\TT{-3}$&0.95&$6.95\TT{-4}$\\\hline
$L_\infty$ error of $\phi$&$5.47\TT{-3}$&0.93&$2.86\TT{-3}$&0.86&$1.56\TT{-3}$&0.81&$8.91\TT{-4}$\\\hline
\end{tabular}
\end{center}
\caption{Convergence rate for a circle with curvature dependent speed of
$\alpha=1.5$ and $\beta=1$. Initially circle is centered at $(0,0)$ with
radius one in a domain of $[-2,2]^2$. Test was run until $0.5$.The radius
$r(t)$ of the circle satisfies $r'=\alpha-\frac\beta{r}$ with $r(0)=1$.
$r(0.5)$ is approximated as $1.3108122$ from the ordinary differential
equation within error bound of $10^{-7}$.}
\label{tab_circle_motion_by_curvature}
\end{table}

%\begin{figure}
%\begin{center}
%\includegraphics[scale=0.25]{curvature_flow_quadtree.eps}
%\end{center}
%\caption{$\alpha=1,\beta=1$ in $256^3$ resolution. From top-left to
%bottom-right, the times are $0$, $0.039$, $0.107$, $0.507$, $0.800$ and
%$1.03$. The time step restriction is $\Delta t=\Delta
%x_{\text{smallest}}/2$. Note that the velocities of convection and
%curvature cancel out not to move the side parts of the two circles for some
%small time. } \label{fig_curvature_flow_in_quadtree}
%\end{figure}

\begin{figure}
\begin{center}
\includegraphics[scale=0.25]{curvature_flow_in_octree.eps}
\end{center}
\caption{Motion with curvature flow for a barbel shape. $\alpha=0,\beta=1$ in $256^3$ resolution. From top-left to bottom-right, the times are $0$, $0.023$, $0.093$, $0.140$, $0.304$ and $0.323$. The CFL condition is $\Delta t=\Delta x_{\text{smallest}}$.}
\label{fig_curvature_flow_in_octree}
\end{figure}

%--------------------------------------------------------------------------
%
%
%
%--------------------------------------------------------------------------
\section{Adaptive Grid Generation}
As the interface deforms some provisions must be given to refine the grid
near the interface while coarsening in regions farther away. The grid is
constructed in such a way that the smallest grid cells lie on the interface
as described in section \ref{sec_spatial_discretization}. This construction
depends on an input function,
$\tilde{\phi}^{n+1}:\mathbb{R}^n\to\mathbb{R}$ that is close to the signed
distance function at any point in space. This function can be constructed
in two different ways: (1) In the case where semi-Lagrangian methods are
used, a function $\tilde{\phi}^{n+1}$ can be defined as
$\tilde{\phi}^{n+1}=\phi^n(x_d)$, where $x_d$ is found by tracing back the
characteristic curves and where $\phi^n(x_d)$ is interpolated from the node
values of $\phi^n$ as described in section \ref{sec_semi_lagrangian}. (2)
In the case where the velocity field is nonlinear, semi-Lagrangian methods
cannot be used. In this case the level set function is first evolved from
$\phi^{n}$ to $\phi^{n+1}$ on the same grid $G^n$. Then $\phi^{n+1}$ is
reinitialized into a signed distance function using the algorithm described
in section \ref{sec_reinitialization}. Now at every point in space, we can
define $\tilde{\phi}^{n+1}:\mathbb{R}^n\to\mathbb{R}$ by interpolation of
$\phi^{n+1}$. Once the function
$\tilde{\phi}^{n+1}:\mathbb{R}^n\to\mathbb{R}$ can be define anywhere in
space, the new grid $G^{n+1}$ is generated by simply splitting a cell if
the Lipschitz condition: $$\min_{v\in \textrm{vertices}(C)}|\phi(v)| \le
\textrm{Lip}(\phi)\cdot \textrm{diag-size}(C)$$ is satisfied. In practice,
instead of generating $G^{n+1}$ from the root cell, we start from $G^n$ and
apply the procedure detailed in Algorithm 1, i.e. starting the recursion
from the root cell of $G^{n+1}$, the cell is recursively split if the
refinement criteria is satisfied, otherwise all of its children are merged.

\begin{center}
\begin{tabular}{ll}\hline
\multicolumn{2}{c}{\textbf{Algorithm 1} : Grid Generation} \bigstrut\\\hline
\multicolumn{2}{l}{\textbf{Input} : $G^n$ and ${\tilde{\phi}}^{n+1}:\mathbb{R}^d \to \mathbb{R}$
}\bigstrut\\\hline
1. &$G^{n+1}=G^{n}$ \\
2. &$C= \text{the root cell of } G^{n+1}$ \\
3. &\textbf{if} the Lipschitz condition for  ${\tilde{\phi}}^{n+1}$ is satisfied at $C$ \\
4. &\qquad \textbf{if} $C$ is a leaf cell \\
5. &\qquad\qquad split $C$ \\
6. &\qquad \textbf{end if} \\
7. &\qquad \textbf{for} each child cell $C'$ of $C$ \\
8. &\qquad \qquad \textbf{go to} 3 with $C=C'$ \\
9. &\qquad \textbf{end for} \\
10.&\textbf{else} \\
11.&\qquad merge $C$ \\
12.& \textbf{end if} \\\hline
\multicolumn{2}{l}{\textbf{Output} : $G^{n+1}$}\bigstrut\\\hline
\end{tabular}
\end{center}

%-------------------------------------------------------------------------------
%
% Extrapolation
%
%-------------------------------------------------------------------------------
\section{High Order Extrapolation in the Normal Direction}
The ghost fluid method, introduced by Fedkiw \etal \cite{Fedkiw:1999:ANE}, is a technique for
imposing boundary conditions at the interface in a level set framework and has been successfully
applied to a wide range of applications (see e.g.
\cite{fedkiw:2001:VSO,fedkiw:2002:CEF,Fedkiw:Ghost_Fluid:Godunov_Methods,nguyen:2001:jcpflame,
Nguyen:2002:Conservative_Ghost_Fluid,Caiden:Incompressible_Compressible,Liu:2000:ABC,Gibou:2002:Second_Order_Accurate,Gibou:2003:Crystal_Growth_JSC,
Gibou:2003:Fourth_Order_Accurate,Gibou::Vaporization} and the references therein). One basic
component of this method is the extrapolation of some scalar quantities in the normal direction. In
some cases (see e.g. \cite{Gibou:2003:Fourth_Order_Accurate}), high order extrapolations in the
normal direction are needed. This can be performed in a series of steps, as proposed in Aslam
\cite{Aslam:Extrapolation}. For example, suppose that we seek to extrapolate $u$ quadratically from
the region where $\phi\leq 0$ to the region where $\phi>0$. We first compute
$u_{nn}=\vec{n}\cdot\nabla\left( \vec{n}\cdot\nabla u \right)$ in the region where $\phi \leq 0$
and extrapolate (constant extrapolation) this quantity across the interface by solving the
following partial differential equation:
\begin{equation*}
\frac{\partial u_{nn}}{\partial \tau} + H(\phi,u_{nn})(\vec{n}\cdot\nabla u_{nn}       )=0,
\end{equation*}
where $H(\phi,u_{nn})$ is the Heaviside function defined below and $\tau$ is the fictitious time
step. Then we define $u_n$ in the region where $\phi>0$ in such a way its normal derivative is
$u_{nn}$. This can be accomplished by solving the following PDE:
\begin{equation*}
\frac{\partial u_{n }}{\partial \tau} + H(\phi,u_n)(\vec{n}\cdot\nabla u_{n }-u_{nn})=0.
\end{equation*}
Finally we can define $u$ in such a way its normal derivative is $u_n$ by solving:
\begin{equation*}
\frac{\partial u_{  }}{\partial \tau} + H(\phi,u)(\vec{n}\cdot\nabla u     -u_{n })=0.
\end{equation*}
Numerically the Heaviside function $H(\phi,S)(v_i)$ associated with a quantity $S$ at the node
$v_i$ is set to zero if the nodes involved in the computation of $S$ are all in the region where
$\phi<0$. Otherwise, it is set to 1. Therefore we define the Heaviside functions $H(\phi,u)$,
$H(\phi,u_{n})$ and $H(\phi,u_{nn})$ as follows:
\begin{align*}
H(\phi,u)(v_i) &= \begin{cases}
                 0 & \text{ if $\phi(v_i)<0$ } \\
                 1 & \text{ otherwise}
                 \end{cases} ,\\
H(\phi,u_n)(v_i) &= \begin{cases}
                   0 & \text{ if $H(\phi,u)(v_j)=0$ for all $v_j\in$ ngbd$(v_i)$} \\
                   1 & \text{ otherwise}
                   \end{cases} ,\\
H(\phi,u_{nn})(v_i) &= \begin{cases}
                   0 & \text{ if $H(\phi,u_n)(v_j)=0$ for all $v_j\in$ ngbd$(v_i)$} \\
                   1 & \text{ otherwise}
                   \end{cases},
\end{align*}
where $\text{ngbd}(v_i)$ denotes the set of direct neighboring nodes of $v_i$. The quantity
$u_n=\vec{n}\cdot\nabla u$ is computed by the central finite differences described in section
\ref{sec_finite_differences} for all the nodes where $H(\phi,u_n)=0$. Likewise, using the values of
$u_{nn}$ is then computed by central differencing for all the nodes where $H(\phi,u_{nn})=0$. The
three partial differential equations above are discretized in a dimension by dimension framework
using the upwind schemes and the one-sided finite differences of section
\ref{sec_finite_differences}, i.e. the discretizations in a semi-discrete form read:
\begin{align*}
\frac{d}{d\tau}u_{nn}+H(\phi,u_{nn})\left(n_x^+D_x^-u_{nn}+n_x^-D_x^+u_{nn}\right)&=0,
\end{align*}
\begin{align*}
\frac{d}{d\tau}u_{n }+H(\phi,u_{n })\left(n_x^+D_x^-u_{n }+n_x^-D_x^+u_{n }\right)&=H(\phi,u_{n})
u_{nn},
\end{align*}
and
\begin{align*}
\frac{d}{d\tau}u     +H(\phi,u     )\left(n_x^+D_x^-u     + n_x^-D_x^+u    \right)&=H(\phi,u ) u_{n
}.
\end{align*}
These semi-discrete equations are then evolved in time using the same TVD RK-2 method of section
\ref{sec_reinitialization}. Since the equations are evolved in fictitious time, we can take the
same time step restriction as in the reinitialization procedure of section
\ref{sec_reinitialization}.

Figure \ref{extrapolation_aslam} illustrates the constant, linear and quadratic extrapolation
obtained with the algorithm described above: Consider a computational domain $\Omega=(-\pi,
\pi)\times(-\pi, \pi)$ separated into two regions: $\Omega^-$ defined as the interior of a disk
with center at the origin and radius two, and $\Omega^+=\Omega \setminus \Omega^-$. The function
$u$ to be extrapolated from $\Omega^-$ to $\Omega^+$ is defined as $u=\cos(x)\sin(y)$ for $x \in
\Omega^-$. We have extrapolated $u$ in the entire region in this example for the sake of
presentation but we emphasize that in practice the extrapolation is performed only in a
neighborhood of the interface. Tables \ref{table_constant_extrapolation},
\ref{table_linear_extrapolation} and \ref{table_quadratic_extrapolation} demonstrate the first
order accuracy for the constant extrapolation, the second order accuracy for the linear
extrapolation and the third order accuracy for the quadratic extrapolation. We note that it is
enough to discretize $D^{\pm}_xu_{nn}$ and $D^{\pm}_xu_{n}$ with the first order accurate finite
difference of section \ref{sec_finite_differences}, and $D^{\pm}_xu$ with the second order accurate
finite difference in section \ref{sec_finite_differences} to achieve third order accuracy in $u$ in
the case of a quadratic extrapolation. The same accuracy would be achieved in the case where the
second order accurate finite differences were used for $D^{\pm}_xu_{nn}$, $D^{\pm}_xu_{n}$, and
$D^{\pm}_xu_{n}$. However, using the first order accurate finite difference schemes for
$D^{\pm}_xu_{nn}$ and $D^{\pm}_xu_{n}$ yields more robust results since $u_{nn}$ and $u_{n}$ may be
noisy unless $u$ is a very smooth function.

\begin{table}
\begin{center}
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline Finest &$L^\infty$ error&rate&$L^1 error$&rate \\ Resolution& of $\phi$      &    & of
$\phi$ & \\\hline

$32^2$&$5.11\times10^{-1}$& &$1.11\times10^{-1}$&      \\\hline
$64^2$&$2.55\times10^{-1}$&1.01&$3.75\times10^{-2}$&1.56   \\\hline
$128^2$&$1.25\times10^{-1}$&1.02&$1.06\times10^{-2}$&1.82   \\\hline
$256^2$&$6.20\times10^{-2}$&1.01&$2.89\times10^{-3}$&1.87   \\\hline
$512^2$&$3.14\times10^{-2}$&.97&$7.59\times10^{-4}$&1.92   \\\hline
$1024^2$&$1.59\times10^{-2}$&.98&$2.01\times10^{-4}$&1.91   \\\hline
\end{tabular}
\end{center}
\caption{Convergence rate for the constant
extrapolation}\label{table_constant_extrapolation}
\end{table}

\begin{table}
\begin{center}
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline Effective &$L^\infty$ error&rate&$L^1 error$&rate \\ Resolution& of $\phi$      &    & of
$\phi$ & \\\hline

$32^2$&$2.02\times10^{-1}$& &$3.52\times10^{-2}$&      \\\hline
$64^2$&$7.21\times10^{-2}$&1.48&$6.09\times10^{-3}$&2.53   \\\hline
$128^2$&$1.78\times10^{-2}$&2.00&$8.79\times10^{-4}$&2.79   \\\hline
$256^2$&$5.27\times10^{-3}$&1.76&$1.19\times10^{-4}$&2.88   \\\hline
$512^2$&$1.12\times10^{-3}$&2.22&$1.54\times10^{-5}$&2.94   \\\hline
$1024^2$&$2.83\times10^{-4}$&1.99&$2.04\times10^{-6}$&2.91   \\\hline
\end{tabular}
\end{center}
\caption{Convergence rate for the linear
extrapolation}\label{table_linear_extrapolation}
\end{table}

\begin{table}
\begin{center}
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline Effective &$L^\infty$ error&rate&$L^1 error$&rate \\ Resolution& of $\phi$      &    & of
$\phi$ & \\\hline

$32^2$&$1.62\times10^{-1}$& &$2.26\times10^{-2}$&      \\\hline
$64^2$&$2.31\times10^{-2}$&2.82&$2.21\times10^{-3}$&3.36   \\\hline
$128^2$&$2.95\times10^{-3}$&2.96&$1.65\times10^{-4}$&3.73   \\\hline
$256^2$&$3.81\times10^{-4}$&2.95&$1.17\times10^{-5}$&3.81   \\\hline
$512^2$&$4.89\times10^{-5}$&2.96&$7.82\times10^{-7}$&3.91   \\\hline
$1024^2$&$6.19\times10^{-6}$&2.98&$5.31\times10^{-8}$&3.88   \\\hline
\end{tabular}
\end{center}
\caption{Convergence rate for the quadratic
extrapolation}\label{table_quadratic_extrapolation}
\end{table}

\begin{figure}
\begin{center}
\includegraphics[scale=0.25]{constant_extrapolation.eps}\includegraphics[scale=0.25]{linear_extrapolation.eps}\includegraphics[scale=0.25]{quadratic_extrapolation.eps}
\includegraphics[scale=0.2]{constant_extrapolation_zoom.eps}\hspace{1cm}\includegraphics[scale=0.2]{linear_extrapolation_zoom.eps}\hspace{1cm}\includegraphics[scale=0.2]{quadratic_extrapolation_zoom.eps}
\end{center}
\caption{Contours of the solution after it has been extrapolated across the
interface with a constant (left), linear (center) and quadratic (right)
extrapolations across an interface. The top row illustrates the
extrapolation on the entire domain and the bottom row is a zoom near the
interface. The exact solution is given inside the circle centered at the
origin and with radius 2 and is extrapolated outside in the normal
direction. We then plot the level curves of the solution.}
\label{extrapolation_aslam}
\end{figure}

\section{Conclusion}
We have presented a level set method on non-graded adaptive Cartesian
grids, i.e. grids for which the ratio between adjacent cells is not
constrained. We use quadtree and octree data structures to represent the
grid and a simple algorithm to generate a mesh with the finest resolution
at the interface. We have presented (1) a locally third order accurate
reinitialization scheme that transforms an arbitrary level set function
into a signed distance function, (2) a second order accurate
semi-Lagrangian methods to evolve the linear level set advection equation
under an externally generated velocity field, (3) a second order accurate
upwind method to evolve the nonlinear level set equation under a normal
velocity as well as to extrapolate scalar quantities across an interface in
the normal direction, and (4) a semi-implicit scheme to evolve the
interface under mean curvature. This method produces results with a
negligible amount of mass loss. We have proposed numerical examples in two
and three spatial dimensions to demonstrate the accuracy of the method.

\section{Acknowledgment}
The research of F. Gibou was supported in part by the Alfred P. Sloan Foundation through a research
fellowship in Mathematics.
%-------------------------------------------------------------------------------
%
% References : "reference.bib"
%
%-------------------------------------------------------------------------------
\bibliographystyle{plain}
\addcontentsline{toc}{section}{\refname}\bibliography{../references}

\end{document}

### 2012-11-19 doc/semilagr/semilagrangian.tex

\documentclass[11pt]{article}

\usepackage{algorithm,algorithmic}
\usepackage{amsmath}
\usepackage[margin=1in]{geometry}
\usepackage{xspace}

\newcommand{\pforest}{\texttt{p4est}\xspace}

\setlength{\emergencystretch}{20pt}

\title{A scalable semi-Lagrangian method on octree meshes}
\author{}

\begin{document}

\maketitle

\begin{abstract}
We present a semi-Lagrangian method for solving the advection equation
on nested non-conforming hexahedral meshes.
\end{abstract}

\section{Introduction}

\section{The semi-Lagrangian method}

\begin{algorithm}
  \caption{Basic semi-Lagrangian procedure}
  \begin{algorithmic}
    \STATE Do something.
  \end{algorithmic}
\end{algorithm}

\end{document}

### 2012-11-19 doc/semilagr/semilagrangian.tex
Some text
Complet basic sl algorithm

@@
 \begin{algorithm}
-  \caption{Basic semi-Lagrangian procedure}
+  \caption{Basic semi-Lagrangian time step}
   \begin{algorithmic}
-    \STATE Do something.
+    \FOR{$e$ in local cells}
+      \FOR{$n$ in nodes on $e$ that have not yet been processed}
+        \STATE Extrapolate location of $n$ backward in time as $x_n$
+	\IF{$x_n$ is located on a different processor $p_n$}
+	\STATE Append $x_n$ to send buffer for $p_n$
+	\ENDIF
+      \ENDFOR
+    \ENDFOR
+    \STATE Reverse communication pattern so the receivers know the senders
+    \STATE Send the non-local points $x_n$ to the target processors
+    \STATE Loop over local points $x_n$ and interpolate new values at $n$
+    \STATE Receive points $x_n$, interpolate, add to send buffer
+    \STATE Send back the interpolated values
+    \STATE Receive values for non-local points and update node values
   \end{algorithmic}

### 2012-11-19 src/ArrayV.h
Added the ArrayV class

@@ -1,0 +1,434 @@
+#ifndef _CHOHONG_ArrayV_____
+#define _CHOHONG_ArrayV_____
+
+#include <assert.h>
+#include <memory.h>
+#include <stdexcept>
+#include <iostream>
+#include <stdlib.h>
+#include <cstring>
+// CASL
+#include <lib/utilities/Macros.h>
+
+#define _DEBUG
+
+using namespace std;
+
+//----------------------------------------------------
+//
+// By Chohong Min, 2005
+//
+// ArrayV : [0]~[m_size-1]~[m_max-1]
+//
+//----------------------------------------------------
+template< class  T>
+class ArrayV
+{
+protected:
+    T*  m_data;
+    CaslInt	m_size;
+    CaslInt m_max;
+public:
+    std::string name;
+
+public:
+
+    void reallocate(CaslInt size) {
+      if (NULL != m_data) delete [] m_data; // free(m_data);
+        m_size = size; m_max = size;
+        if (0 != size) m_data = new T[m_max];//(T*)malloc(m_max*sizeof(T));
+    }
+
+    void resize_without_copy( CaslInt size )
+    {
+        if(m_max < size)
+        {
+            if(m_data!=NULL) {
+                delete [] m_data; //  free(m_data);
+                m_data=NULL;
+            }
+
+            m_max  = 2*size+1;
+            m_size =   size;
+
+            if(size !=0) m_data = new T[m_max]; //(T*)malloc(m_max*sizeof(T));
+            else         m_data = NULL;
+        }
+        else
+            m_size = size;
+    }
+
+    double dot( const ArrayV<T>& V ) const
+    {
+      double  ret=0;
+      const T* data = (const T*)V;
+      for( int i=0; i<m_size; i++ )
+        ret += m_data[i] * data[i];
+      return ret;
+    }
+
+    void resize_with_copy( CaslInt size )
+    {
+        if(m_max < size )
+        {
+            CaslInt copy_size = (m_size<size) ? m_size : size;
+            CaslInt i;
+            T* new_data = new T[copy_size]; //(T*)malloc(copy_size*sizeof(T));
+            for(i=0;i<copy_size;i++) {
+                new_data[i] =   m_data[i];
+            }
+            resize_without_copy(size);
+            for(i=0;i<copy_size;i++)   {
+                m_data[i] = new_data[i];
+            }
+            delete [] new_data; //free(new_data);
+        }
+        else
[…]

### 2012-11-19 examples/Poisson/main_poisson.cpp
Rename to my_p4est_vtk

@@ -4,8 +4,8 @@
 #include <p4est_nodes.h>
 
 // My files for this project
-#include "srcs/p4est_vtk.h"
-#include "srcs/poisson_solver.h"
+#include "src/my_p4est_vtk.h"
+#include "src/poisson_solver.h"
 
 // System
 #include <stdexcept>

### 2012-11-19 src/my_p4est_vtk.cpp

@@ -34,7 +34,7 @@
 #include <p8est_nodes.h>
 #define P4EST_VTK_CELL_TYPE     11      /* VTK_VOXEL */
 #else
-#include "p4est_vtk.h"
+#include "my_p4est_vtk.h"
 #include <p4est_nodes.h>
 #define P4EST_VTK_CELL_TYPE      8      /* VTK_PIXEL */
 #endif /* !P4_TO_P8 */

### 2012-11-20 examples/dummy/dummymain.c
Enhance dummy example

@@ -1,7 +1,15 @@
 
 #include <p4est.h>
 
-int main (int argc, char ** argv)
+static int
+simple_refine (p4est_t * p4est, p4est_topidx_t which_tree,
+        p4est_quadrant_t * quadrant)
+{
+        return which_tree == 0 && quadrant->level < 2;
+}
+
+int
+main (int argc, char ** argv)
 {
 	int mpiret;
 	int mpirank;
@@ -23,13 +31,19 @@ int main (int argc, char ** argv)
 	SC_CHECK_MPI (mpiret);
 
 	/* set log level */
-	lp = SC_LP_ESSENTIAL;
+	/* lp = SC_LP_ESSENTIAL; */
+	lp = SC_LP_DEFAULT;
 	sc_init (mpicomm, 1, 1, NULL, lp);
 	p4est_init (NULL, lp);
 
-	conn = p4est_connectivity_new_brick (1, 1, 0, 0);
+        /* create some p4est refinement */
+	conn = p4est_connectivity_new_brick (1, 2, 0, 0);
 	p4est = p4est_new (mpicomm, conn, 0, NULL, NULL);
+        p4est_refine (p4est, 2, simple_refine, NULL);
+        p4est_partition (p4est, NULL);
+        p4est_vtk_write_file (p4est, NULL, "twobrick");
 
+        /* clean up */
 	p4est_destroy (p4est);
 	p4est_connectivity_destroy (conn);
 

### 2012-11-20 examples/dummy/dummymain.c
Make and compile fixes

@@ -1,5 +1,6 @@
 
 #include <p4est.h>
+#include <p4est_vtk.h>
 
 static int
 simple_refine (p4est_t * p4est, p4est_topidx_t which_tree,

### 2014-08-27 paper/tex/abstract.tex
Started writing the draft

%!TEX root = draft.tex
\begin{abstract}
Abstract will be filled in after finalization of the paper.
\end{abstract}

\begin{keyword}
Level-set Method \sep Adaptive Grids \sep Parallel Computing \sep Oct-/Quadtree datastructures
\end{keyword}

\maketitle

### 2014-08-27 paper/tex/applications.tex
Started writing the draft

%!TEX root = draft.tex
\section{Application to Stefan Problems}

We now propose an application of the level-set method on parallel adaptive cartesian mesh to the well studied stefan problem.

\subsection{Presentation of the problem}

We study the problem of the phase transition of a liquid melt to a solid crystaline structure. For a liquid melt and a solid crystal with respective heat capacities $c_l$ and $c_s$ and uniform thermal diffusivities $k_l$ and $k_s$, the respective temperatures $T_l$ and $T_s$ diffuse as

\begin{align*}
c_l \pd{T_l}{t} & = k_l \nabla T_l \quad \mathrm{in} ~~ \Omega^- , \\
c_s \pd{T_s}{t} & = k_s \Delta T_s \quad \mathrm{in} ~~ \Omega^+ .
\end{align*}

At the interface between the solid and the liquid phases, the temperature is defined as

\begin{equation*}
T_s = T_l = T_{\Gamma} = -\epsilon_c \kappa - \epsilon_v (\mathbf{V} \cdot \mathbf{n}),
\end{equation*}

where $\kappa$ is the local interface curvature, $V$ is the velocity of the interface, $\mathbf{n}$ is the outward normal to the solidification front and $\epsilon_c$ and $\epsilon_v$ are the curvature and kinetic undercooling coefficients. The velocity $V$ is defined from the jump in the heat flux at the interface

\begin{equation}
L (\mathbf{V} \cdot \mathbf{n}) = - \left[ k_l \pd{T_l}{\mathbf{n}} - k_s \pd{T_s}{\mathbf{n}} \right],
\end{equation}

with $L$ the latent heat.

\subsection{Scalability}

\subsection{Numerical experiments}

\begin{figure}[ht!]
\begin{center}
\includegraphics[width=.8\textwidth]{pictures/crystal_grid_low.png}
\end{center}
\end{figure}

### 2014-08-27 paper/tex/introduction.tex
Started writing the draft

%!TEX root = draft.tex
\section{Introduction}
\begin{enumerate}
\item \textbf{What is level-set and why is it important?}
The level-set method, originally proposed by Sethian and Osher \fcite{classic LS paper}, is a popular and powerful framework for tracking arbitrary interfaces that undergo complicated topological changes. As a result, the level-set method has wide range of application such as multiphase flows, moving boundary problems, imgae segmentation, and computer graphics graphics, to only name a few \fcite{fedkiw/osher book, sethian book}.

\item \textbf{What are the main issues of using level-set method?}
What makes level-set method powerful and easy to use is that in this method location of the interface is defined implicitly on an underlying grid. This conveninvce, however, comes at a price. First, compared to an explicit method, e.g. front tracking \fcite{front trcking refs}, level-set method is typically less accurate and mass conservation could be a problem although progress has been made in resolving this issue \fcite{Enright/Fedkiw paper, other?}. Second, the level-set function has to be defined in a higher dimentional space compared to the interface. If only the location of interface is needed, the added dimension greatly increases the overall computational cost. This problem, however, could be handled by computing the level-set only close to the interface, e.g. as in the narrow-band level-set method \fcite{narrow band, the hash table paper?}.

\item \textbf{Why using AMR helps? How do we plan to extend that paper here? What are the challenges?}

\item \textbf{Parallel octrees review: p4est, octor, dendro, etc.}

\item \textbf{Parallel reinitialization methods? What are the challenges? Whats out there?}

\item \textbf{Parallel semi-lagrangian methods? What are the challenges? Whats out there?}

\item \textbf{How do we go about the parallization? Whats the organization of the paper}
\end{enumerate}

### 2014-08-27 paper/tex/levelset_method.tex
Started writing the draft

%!TEX root = draft.tex
\section{The level-set method}

\indent The level-set method is a method to capture interfaces implicitely by constructing a function $\phi$, called the level-set function, that is negative on on side of the domain $\Omega^-$ and positive on the other side $\Omega^+$, thus defining the interface $\Gamma$ as the zero contour of $\phi$, $\Gamma = \{\mathbf{x} \vert \phi(\mathbf{x}=0\}$.\\

If infinitely many functions satisfy these requirements, the most natural and most convenient representation is a signed distance function to the interface. The level-set function is therefore reinitialized to a signed distance function at every time step by solving the reinitialization equation

\begin{equation} \label{eq::reinitialization}
\pd{\phi}{\tau} + \mathrm{sign}(\phi) \left( \lvert \phi \rvert - 1 \right) = 0,
\end{equation}

where $\tau$ is a fictitious time. The discretization and implementation of this procedure will be discussed later in section \ref{section::reinitialization}. Geometrical quantities such as the normal to the interface and the local curvature can be computed straightforwardly from the level-set function as

\begin{equation*}
\mathbf{n} = \frac{\nabla \phi}{\lvert \nabla \phi \rvert},
\end{equation*}

where it should be noted that $\lvert \nabla \phi \rvert = 1$ when the level-set function is reinitialized, and

\begin{equation*}
\kappa = \nabla \cdot \mathbf{n} =  \nabla \cdot \left( \frac{\nabla \phi}{\lvert \nabla \phi \rvert} \right).
\end{equation*}

\section{The \texttt{p4est} library}

The work we present relies on the parallel octree library \texttt{p4est} \cite{burstedded:2011:p4est}. It is scalable implementation of the general octree structure in a massively parallel mpi environment. We limit our presentation of \texttt{p4est} to the relevant information for our application, for more details the reader is invited to read the original article \cite{burstedded:2011:p4est}.

The \texttt{p4est} library provides the geometrical information for the leaves of an octree which includes the neighboring information for the leaves (called quadrants in the original article) of the tree, a layer of ghost quadrants and ghost vertices for each process, the coarsening and refining procedures, and encapsulates the communication between processes. However, it does not provide the vertical structure of the octrees. If it is generally unnecessary, it is a desirable feature for a finite differences based code as it speeds up the access to arbitrary quadrants information dramatically. Thus, we choose to reconstruct the local vertical structure of the octree on every processes. Note that given the local and ghost layer information, this is an entirely local procedure and no communication is needed.

In addition to the owned and ghost categories, we choose to group the vertices in \textit{local} and \textit{layer} sets. We define the layer vertices as the vertices that are part of another process's ghost layer and the local vertices as the vertices that are not part of any other process's ghost layer. Figure \ref{fig::local_layer_vertices} gives a graphical representation of a boundary between two processes and the classification of the various quadrants and vertices. The idea behind this sorting is that layer vertices are needed by other processes for finite difference calculations and therefore require to be synchronized while the local vertices are used only by the process they belong to. We make use of this classification to implement scalable algorithms for the level-set method, as explained in the following sections.

\begin{figure}[ht!]
\begin{center}
\includegraphics[width=.4\textwidth]{pictures/layer_nodes.pdf}
\caption{Illustration of a possible layout at the interface between two processes. The red quarants and the orange and purples vertices belong to process 1 while the green quadrants and the green and black vertices belong to process 2. The green G denotes the ghost quadrants for process 2, the green dots are the ghost vertices for process 1, and the purple dots correspond to layer vertices for process 1 (and therefore ghost vertices for process 2).} \label{fig::local_layer_vertices}
\end{center}
\end{figure}

### 2014-08-27 paper/tex/parallel_algorithms.tex
Started writing the draft

%!TEX root = draft.tex
\section{Presentation of the algorithms}

We now propose a detailed description of the main algorithms for the level-set method.

\subsection{Interpolations}

The first important step to build a level-set based code is to provide an interpolation procedure on the vertices. Being able to evaluate quantities at arbitrary locations in the computational domain is mandatory for procedures such as advecting the level-set function or enforcing boundary conditions on the irregular interface. The interpolation procedure enables any process to evaluate a quantity anywhere in the domain.

\begin{algorithm}[ht!]
\begin{algorithmic}
\If {interpolation order $>$ 2}
	\State compute second order derivatives for local and ghost points if needed
\EndIf
\State begin send point buffer
\State do the interpolation for local points
\State do the interpolation for ghost points
\State begin receiving point buffer
\For {in $\in$ \{remote senders\} }
	\State finish receiving from send i
	\State interpolate all points from sender i
\EndFor
\State receive the point buffer
\end{algorithmic}
\end{algorithm}

\subsection{Semi lagrangian advection}

\subsection{Reinitialization and extensions} \label{section::reinitialization}

The reinitialization procedure transforms an arbitrary function into a signed distance function to its zero contour. One of the popular methods to achieve this results is to solve the reinitialization equation \ref{eq::reinitialization} \cite{Sussman:1994:levelsetincompressible, Min_Gibou:2007:Level_Set_Adaptive}. Other methods, such as the Fast Sweeping \cite{??} and the Fast Marching methods \cite{Sethian:1996:FMM, chopp:2001:fmm}, are based on the wave propagation nature of the reinitialization front and provide faster implementations of the reinitialization procedure in a serial environment. However, their parallelization is an involved process and we opt for solving the reinitialization equation.

We base our approach on Min and Gibou \cite{Min_Gibou:2007:Level_Set_Adaptive} and discretize the reinitialization equation as

\begin{equation*}
\frac{d\phi}{d\tau} + \mathrm{sign}(\phi^0)[H_G(D^+_x\phi, D^-_x\phi, D^+_y\phi, D^-_y\phi)-1] = 0,
\end{equation*}

with $\mathrm{sign}(\phi^0)$ the signum of the initial level-set function and $H_G$ the Godunov Hamiltonian:

\begin{equation*}
H_G(a,b,c,d) =
\begin{cases}
\sqrt{\max(\lvert a^+ \rvert^2, \lvert b^- \rvert^2) + \max(\lvert c^+ \rvert^2, \lvert d^- \rvert^2)} \quad \mathrm{if~sgn}(\phi^0) \leq 0\\
\sqrt{\max(\lvert a^- \rvert^2, \lvert b^+ \rvert^2) + \max(\lvert c^- \rvert^2, \lvert d^+ \rvert^2)} \quad \mathrm{if~sgn}(\phi^0) > 0
\end{cases}
\end{equation*}

with $a^+=\max(a,0)$ and $a^-=\min(a,0)$. The one-sided derivatives $D^{+/-}_{x/y}\phi$ are discretized by second order accurate finite differences as explained by Min and Gibou in \cite{Min_Gibou:2007:Level_Set_Adaptive}. Following the presentation in \cite{Min_Gibou:2007:Level_Set_Adaptive} which incorporates the subcell resolution introduced by Russo and Smereka \cite{Russo_Smereka}, we include the initial position of the interface, given by $\phi^0$, in the computation of the one sided derivatives next to the interface in order to preserve the interface location. Finally, we choose a simple Euler step for the time discretization.

The general parallel algorithm for the reinitialization procedure is presented in \ref{algo::reinitialization}. For each fictitious time step, we start by computing the new value of the level-set function for the layer nodes. We then initiate the communication to update the ghost values while computing the new values for local nodes. After finishing the communication, we can then update $\phi^n$ to $\phi^{n+1}$ and move on to the next time step. The number of iterations needed for the reinitialization procedure to converge on the whole domain or on a band surrounding the interface depends on the size of the mesh, the geometry of the interface and how far from a signed distance function $\phi^0$ is. In practice, twenty iterations provide reasonable results for the applications we present.

\begin{algorithm}[ht!]
\begin{algorithmic}
\State Compute initial derivatives of $\phi$ for subcell resolution
\For {number of iterations}
	\State Compute the derivatives of $\phi$
	\State Solve reinitialization of layer nodes: $\phi_{layer} \leftarrow \phi_{layer}^{n+1}$
	\State Begin update for layer nodes
	\State Solve reinitialization on local nodes : $\phi_{local} \leftarrow \phi_{local}^{n+1}$
	\State End update for layer nodes
	\State Update $\phi$: $\phi \leftarrow \phi^{n+1}$
\EndFor
\end{algorithmic}
\caption{General structure of the parallel implementation of the reinitialization procedure presented in section \ref{section::reinitialization}.} \label{algo::reinitialization}
\end{algorithm}

### 2014-08-27 paper/tex/scaling_results.tex
Started writing the draft

%!TEX root = draft.tex
\section{scalability and accuracy}

\subsection{Interpolation}

\begin{figure}[ht!]
\begin{center}
\includegraphics[width=.7\textwidth]{pictures/scaling_interpolation_all.pdf}
\caption{Scalability of the interpolation procedure. blue: alpha=0.05, purple alpha=0.50, red alpha=0.95 (alpha = percent of non-local points)}
\end{center}
\end{figure}

\subsection{Semi lagrangian advection}

\subsection{Reinitialization} \label{section::scaling_reinitialization}

The reinitialization procedure presented in section \ref{section::reinitialization} necessitates the second order derivatives of the level-set function for the second order discretization in space. The finite difference schemes we implemented rely on the description of the neighborhood of each vertex in the adaptive tree environment. This information can either be computed on the fly or stored in a buffer. Fetching the neighbors information is rather time consuming and was observed to improve the overall scalability of the algorithm by increasing the local computation load. However, it slowed down the procedure significantly and therefore we choose to present the buffered version of the implementation. If one was to be limited in memory, constructing the neighborhood information on the fly would be a desirable option and would lead to better scaling, but in our case memory is not a limiting factor.

The scaling results are presented in figure \ref{fig::scaling_reinitialization} and demonstrate satisfying scalability. Taking for reference the timing with $128$ processes, we observe efficiencies of $0.877$ with $1024$ processes, $0.763$ with $2048$ processes and $0.609$ with $4096$ processes.

\begin{figure}[ht!]
\begin{center}
\includegraphics[width=.7\textwidth]{pictures/scaling_reinitialization_1st_time_2nd_space_with_buffer.pdf}
\caption{Scalability of the reinitialization procedure presented in section \ref{section::reinitialization} and analyzed in section \ref{section::scaling_reinitialization}.} \label{fig::scaling_reinitialization}
\end{center}
\end{figure}

\subsection{Poisson solver}

### 2014-08-27 paper/tex/p4est.tex
updating citations with new key format

\documentclass{elsarticle}
%\usepackage[pdftex,breaklinks,linktocpage,pagebackref,hyperindex,hyperfigures]
\usepackage{hyperref}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage[tmargin=1in,bmargin=1in]{geometry}
\usepackage{subfigure}
\usepackage{graphicx}
\usepackage{latexsym}
\usepackage{pdfsync}
\usepackage[boxed]{algorithm}
\usepackage{algpseudocode}
\usepackage{algorithm}
%\usepackage{algorithmic}
\usepackage{multirow}
\usepackage{rotating}
\usepackage{color}
\usepackage{caption}
\usepackage{url}

\hypersetup{
    colorlinks=true,      		% false: boxed links; true: colored links
    linkcolor=blue,       		% color of internal links
    citecolor=blue,       		% color of links to bibliography
    filecolor=black,      		% color of file links
    urlcolor=red,       		% color of external links
    bookmarks=false,
    pdffitwindow=true,
    pdfpagelayout=SinglePage
}

\newcommand{\etal}{\textit{et al.\ }}
\newcommand{\pd}[2]{\frac{\partial #1}{\partial #2}}
\newcommand{\mbf}[1]{\mathbf{#1}}
\newcommand{\comment}[1]{\textcolor{red}{#1}}
\newcommand{\tensor}[1]{\underline{\underline{\boldsymbol{#1}}}}

\newtheorem{property}{Property}

\begin{document}

\begin{frontmatter}
\title{Level-Set method on \texttt{p4est}}

\cortext[cor]{Corresponding author: arthur.guittet@gmail.com}

\address[MECHE]{Department of Mechanical Engineering, University of California, Santa Barbara, CA 93106-5070}
\address[CS]{Department of Computer Science, University of California, Santa Barbara, CA 93106-5110}

 \author[MECHE]{Mohammad Mirzadeh} \author[MECHE]{Arthur Guittet} \author[MECHE,CS]{Fr\'ed\'eric Gibou}

\begin{abstract}
cool abstract
\end{abstract}
\begin{keyword}
p4est \sep Level-Set
\end{keyword}

\end{frontmatter}

\section{Introduction}

\section{The level-set method}

\indent The level-set method is a method to capture interfaces implicitly by constructing a function $\phi$, called the level-set function, that is negative on on side of the domain $\Omega^-$ and positive on the other side $\Omega^+$, thus defining the interface $\Gamma$ as the zero contour of $\phi$, $\Gamma = \{\mathbf{x} \vert \phi(\mathbf{x}=0\}$.\\

If infinitely many functions satisfy these requirements, the most natural and most convenient representation is a signed distance function to the interface. The level-set function is therefore reinitialized to a signed distance function at every time step by solving the reinitialization equation

\begin{equation} \label{eq::reinitialization}
\pd{\phi}{\tau} + \mathrm{sign}(\phi) \left( \lvert \phi \rvert - 1 \right) = 0,
\end{equation}

where $\tau$ is a fictitious time. The discretization and implementation of this procedure will be discussed later in section \ref{section::reinitialization}. Geometrical quantities such as the normal to the interface and the local curvature can be computed straightforwardly from the level-set function as

\begin{equation*}
\mathbf{n} = \frac{\nabla \phi}{\lvert \nabla \phi \rvert},
\end{equation*}

where it should be noted that $\lvert \nabla \phi \rvert = 1$ when the level-set function is reinitialized, and

\begin{equation*}
\kappa = \nabla \cdot \mathbf{n} =  \nabla \cdot \left( \frac{\nabla \phi}{\lvert \nabla \phi \rvert} \right).
\end{equation*}

\section{The \texttt{p4est} library}

The work we present relies on the parallel octree library \texttt{p4est} \cite{Burstedde;Wilcox;Ghattas:11:p4est:-Scalable-Algo}. It is scalable implementation of the general octree structure in a massively parallel mpi environment. We limit our presentation of \texttt{p4est} to the relevant information for our application, for more details the reader is invited to read the original article \cite{Burstedde;Wilcox;Ghattas:11:p4est:-Scalable-Algo}.

The \texttt{p4est} library provides the geometrical information for the leaves of an octree which includes the neighboring information for the leaves (called quadrants in the original article) of the tree, a layer of ghost quadrants and ghost vertices for each process, the coarsening and refining procedures, and encapsulates the communication between processes. However, it does not provide the vertical structure of the octrees. If it is generally unnecessary, it is a desirable feature for a finite differences based code as it speeds up the access to arbitrary quadrants information dramatically. Thus, we choose to reconstruct the local vertical structure of the octree on every processes. Note that given the local and ghost layer information, this is an entirely local procedure and no communication is needed.

In addition to the owned and ghost categories, we choose to group the vertices in \textit{local} and \textit{layer} sets. We define the layer vertices as the vertices that are part of another process's ghost layer and the local vertices as the vertices that are not part of any other process's ghost layer. Figure \ref{fig::local_layer_vertices} gives a graphical representation of a boundary between two processes and the classification of the various quadrants and vertices. The idea behind this sorting is that layer vertices are needed by other processes for finite difference calculations and therefore require to be synchronized while the local vertices are used only by the process they belong to. We make use of this classification to implement scalable algorithms for the level-set method, as explained in the following sections.

\begin{figure}[ht!]
\begin{center}
\includegraphics[width=.4\textwidth]{pictures/layer_nodes.pdf}
\caption{Illustration of a possible layout at the interface between two processes. The red quarants and the orange and purples vertices belong to process 1 while the green quadrants and the green and black vertices belong to process 2. The green G denotes the ghost quadrants for process 2, the green dots are the ghost vertices for process 1, and the purple dots correspond to layer vertices for process 1 (and therefore ghost vertices for process 2).} \label{fig::local_layer_vertices}
\end{center}
\end{figure}

\section{Presentation of the algorithms}

We now propose a detailed description of the main algorithms for the level-set method.

\subsection{Interpolations}

The first important step to build a level-set based code is to provide an interpolation procedure on the vertices. Being able to evaluate quantities at arbitrary locations in the computational domain is mandatory for procedures such as advecting the level-set function or enforcing boundary conditions on the irregular interface. The interpolation procedure enables any process to evaluate a quantity anywhere in the domain.

\begin{algorithm}[ht!]
\begin{algorithmic}
\If {interpolation order $>$ 2}
	\State compute second order derivatives for local and ghost points if needed
\EndIf
\State begin send point buffer
\State do the interpolation for local points
\State do the interpolation for ghost points
\State begin receiving point buffer
\For {in $\in$ \{remote senders\} }
	\State finish receiving from send i
	\State interpolate all points from sender i
\EndFor
\State receive the point buffer
\end{algorithmic}
\end{algorithm}

\subsection{Semi lagrangian advection}

\subsection{Reinitialization and extensions} \label{section::reinitialization}

The reinitialization procedure transforms an arbitrary function into a signed distance function to its zero contour. One of the popular methods to achieve this results is to solve the reinitialization equation \ref{eq::reinitialization} \cite{Sussman;Smereka;Osher:94:A-Level-Set-Approach, Min;Gibou:07:A-second-order-accur}. Other methods, such as the Fast Sweeping \cite{Zhao:04:A-Fast-Sweeping-Meth} and the Fast Marching methods \cite{Sethian:96:A-Fast-Marching-Leve, Chopp:01:Some-improvements-of}, are based on the wave propagation nature of the reinitialization front and provide faster implementations of the reinitialization procedure in a serial environment. However, their parallelization is an involved process and we opt for solving the reinitialization equation.

We base our approach on Min and Gibou \cite{Min;Gibou:07:A-second-order-accur} and discretize the reinitialization equation as

\begin{equation*}
\frac{d\phi}{d\tau} + \mathrm{sign}(\phi^0)[H_G(D^+_x\phi, D^-_x\phi, D^+_y\phi, D^-_y\phi)-1] = 0,
\end{equation*}

with $\mathrm{sign}(\phi^0)$ the signum of the initial level-set function and $H_G$ the Godunov Hamiltonian:

\begin{equation*}
H_G(a,b,c,d) =
\begin{cases}
\sqrt{\max(\lvert a^+ \rvert^2, \lvert b^- \rvert^2) + \max(\lvert c^+ \rvert^2, \lvert d^- \rvert^2)} \quad \mathrm{if~sgn}(\phi^0) \leq 0\\
\sqrt{\max(\lvert a^- \rvert^2, \lvert b^+ \rvert^2) + \max(\lvert c^- \rvert^2, \lvert d^+ \rvert^2)} \quad \mathrm{if~sgn}(\phi^0) > 0
\end{cases}
\end{equation*}

with $a^+=\max(a,0)$ and $a^-=\min(a,0)$. The one-sided derivatives $D^{+/-}_{x/y}\phi$ are discretized by second order accurate finite differences as explained by Min and Gibou in \cite{Min;Gibou:07:A-second-order-accur}. Following the presentation in \cite{Min;Gibou:07:A-second-order-accur} which incorporates the subcell resolution introduced by Russo and Smereka \cite{Russo;Smereka:00:A-Remark-on-Computin}, we include the initial position of the interface, given by $\phi^0$, in the computation of the one sided derivatives next to the interface in order to preserve the interface location. Finally, we choose a simple Euler step for the time discretization.

The general parallel algorithm for the reinitialization procedure is presented in \ref{algo::reinitialization}. For each fictitious time step, we start by computing the new value of the level-set function for the layer nodes. We then initiate the communication to update the ghost values while computing the new values for local nodes. After finishing the communication, we can then update $\phi^n$ to $\phi^{n+1}$ and move on to the next time step. The number of iterations needed for the reinitialization procedure to converge on the whole domain or on a band surrounding the interface depends on the size of the mesh, the geometry of the interface and how far from a signed distance function $\phi^0$ is. In practice, twenty iterations provide reasonable results for the applications we present.

\begin{algorithm}[ht!]
\begin{algorithmic}
\State Compute initial derivatives of $\phi$ for subcell resolution
\For {number of iterations}
	\State Compute the derivatives of $\phi$
	\State Solve reinitialization of layer nodes: $\phi_{layer} \leftarrow \phi_{layer}^{n+1}$
	\State Begin update for layer nodes
	\State Solve reinitialization on local nodes : $\phi_{local} \leftarrow \phi_{local}^{n+1}$
	\State End update for layer nodes
	\State Update $\phi$: $\phi \leftarrow \phi^{n+1}$
\EndFor
\end{algorithmic}
\caption{General structure of the parallel implementation of the reinitialization procedure presented in section \ref{section::reinitialization}.} \label{algo::reinitialization}
\end{algorithm}

\section{scalability and accuracy}

\subsection{Interpolation}

\begin{figure}[ht!]
\begin{center}
\includegraphics[width=.7\textwidth]{pictures/scaling_interpolation_all.pdf}
\caption{Scalability of the interpolation procedure. blue: alpha=0.05, purple alpha=0.50, red alpha=0.95 (alpha = percent of non-local points)}
\end{center}
\end{figure}

\subsection{Semi lagrangian advection}

\subsection{Reinitialization} \label{section::scaling_reinitialization}

The reinitialization procedure presented in section \ref{section::reinitialization} necessitates the second order derivatives of the level-set function for the second order discretization in space. The finite difference schemes we implemented rely on the description of the neighborhood of each vertex in the adaptive tree environment. This information can either be computed on the fly or stored in a buffer. Fetching the neighbors information is rather time consuming and was observed to improve the overall scalability of the algorithm by increasing the local computation load. However, it slowed down the procedure significantly and therefore we choose to present the buffered version of the implementation. If one was to be limited in memory, constructing the neighborhood information on the fly would be a desirable option and would lead to better scaling, but in our case memory is not a limiting factor.

The scaling results are presented in figure \ref{fig::scaling_reinitialization} and demonstrate satisfying scalability. Taking for reference the timing with $128$ processes, we observe efficiencies of $0.877$ with $1024$ processes, $0.763$ with $2048$ processes and $0.609$ with $4096$ processes.

\begin{figure}[ht!]
\begin{center}
\includegraphics[width=.7\textwidth]{pictures/scaling_reinitialization_1st_time_2nd_space_with_buffer.pdf}
\caption{Scalability of the reinitialization procedure presented in section \ref{section::reinitialization} and analyzed in section \ref{section::scaling_reinitialization}.} \label{fig::scaling_reinitialization}
\end{center}
\end{figure}

\subsection{Poisson solver}

\section{Application to supercapacitor}

\section{Application to the Stefan problem}

We now propose an application of the level-set method on parallel adaptive cartesian mesh to the well studied stefan problem.

\subsection{Presentation of the problem}

We study the problem of the phase transition of a liquid melt to a solid crystaline structure. For a liquid melt and a solid crystal with respective heat capacities $c_l$ and $c_s$ and uniform thermal diffusivities $k_l$ and $k_s$, the respective temperatures $T_l$ and $T_s$ diffuse as

\begin{align*}
c_l \pd{T_l}{t} & = k_l \nabla T_l \quad \mathrm{in} ~~ \Omega^- , \\
c_s \pd{T_s}{t} & = k_s \Delta T_s \quad \mathrm{in} ~~ \Omega^+ .
\end{align*}

At the interface between the solid and the liquid phases, the temperature is defined as

\begin{equation*}
T_s = T_l = T_{\Gamma} = -\epsilon_c \kappa - \epsilon_v (\mathbf{V} \cdot \mathbf{n}),
\end{equation*}

where $\kappa$ is the local interface curvature, $V$ is the velocity of the interface, $\mathbf{n}$ is the outward normal to the solidification front and $\epsilon_c$ and $\epsilon_v$ are the curvature and kinetic undercooling coefficients. The velocity $V$ is defined from the jump in the heat flux at the interface

\begin{equation}
L (\mathbf{V} \cdot \mathbf{n}) = - \left[ k_l \pd{T_l}{\mathbf{n}} - k_s \pd{T_s}{\mathbf{n}} \right],
\end{equation}

with $L$ the latent heat.

\subsection{Scalability}

\subsection{Numerical experiments}

\begin{figure}[ht!]
\begin{center}
\includegraphics[width=.8\textwidth]{pictures/crystal_grid_low.png}
\end{center}
\end{figure}

\section{Acknowledgment}

\section*{References}
\bibliographystyle{plain}
\addcontentsline{toc}{section}{\refname}
\bibliography{/home/guittet/code/CASL_Library/references.bib}

\end{document}

### 2014-08-27 paper/tex/introduction.tex
Updated tex files

@@
 What makes level-set method powerful and easy to use is that in this method location of the interface is defined implicitly on an underlying grid.
-This conveninvce, however, comes at a price.
+This convenience, however, comes at a price.
 First, compared to an explicit method, e.g.
-front tracking \fcite{front trcking refs}, level-set method is typically less accurate and mass conservation could be a problem although progress has been made in resolving this issue \fcite{Enright/Fedkiw paper, other?}.
-Second, the level-set function has to be defined in a higher dimentional space compared to the interface.
+front tracking \fcite{front tracking refs}, level-set method is typically less accurate and mass conservation could be a problem although progress has been made in resolving this issue \fcite{Enright/Fedkiw paper, other?}.
+Second, the level-set function has to be defined in a higher dimensional space compared to the interface.
 If only the location of interface is needed, the added dimension greatly increases the overall computational cost.
@@
 \item \textbf{Why using AMR helps? How do we plan to extend that paper here? What are the challenges?}
+Another approach that can address both problems is the use of local grid refinement.
+In \fcite{Frederic paper}, authors present second-order accurate level-set methods on Cartesian Quadtree (two dimensions) and Octree (three dimensions) grids.
+Use of tree-based Cartesian adaptive grids in the context of level-set method is quite advantageous because 1) It gives fine-grain control over error, which typically occurs close to the interface, 2) It can effectively reduce the dimensionality of the problem by focusing most of the cells close to the interface, and 3) Construction and handling of the tree is quite simple in the presence of an interface which defines a an ideal metric for refinement.
 

### 2014-08-28 paper/tex/introduction.tex
Updated introduction

@@
 The level-set method, originally proposed by Sethian and Osher \fcite{classic LS paper}, is a popular and powerful framework for tracking arbitrary interfaces that undergo complicated topological changes.
-As a result, the level-set method has wide range of application such as multiphase flows, moving boundary problems, imgae segmentation, and computer graphics graphics, to only name a few \fcite{fedkiw/osher book, sethian book}.
+As a result, the level-set method has wide range of application such as multiphase flows, moving boundary problems, image segmentation, and computer graphics graphics, to only name a few \fcite{fedkiw/osher's book, sethian's book}.
 
@@
 
-\item \textbf{Why using AMR helps? How do we plan to extend that paper here? What are the challenges?}
+\item \textbf{Why using AMR helps? How do we plan to extend that paper here?}
 Another approach that can address both problems is the use of local grid refinement.
@@
 Use of tree-based Cartesian adaptive grids in the context of level-set method is quite advantageous because 1) It gives fine-grain control over error, which typically occurs close to the interface, 2) It can effectively reduce the dimensionality of the problem by focusing most of the cells close to the interface, and 3) Construction and handling of the tree is quite simple in the presence of an interface which defines a an ideal metric for refinement.
+Although use of adaptive grids dramatically reduces the computational time, performing high-resolution three dimensional calculations of complex interfacial problems, e.g.
+crystal growth in binary alloys \fcite{Maxime's paper}, is still challenging.
+In this paper we extend these algorithms by proposing parallel implementations on distributed memory machines using domain decomposition technique and Message Passing Interface (MPI) library.
 
-\item \textbf{Parallel octrees review: p4est, octor, dendro, etc.}
+\item \textbf{What is the main challenge for parallelization?}
+One of the main challenges in parallelizing level-set algorithms on adaptive grids, is handling the grid itself in parallel.
+One option is to replicate the entire grid on each processor and use off-the-self graph partitioners, e.g.
+Metis or ParMetis \fcite{metis/parmetis paper}, for load balancing and domain decomposition \fcite{origial deal.II paper}.
+This approach, however, is only scalable to a few hundred processor at best and is limited by the size of the grid itself.
+Moreover, use of a general-purpose graph partitioner adds extra overhead that limits the overall scalability even further.
+Interestingly, Quadtrees and Octrees have nice spatial ordering that naturally lead to the concept of Space Filling Curves (SFC) \fcite{SFC reference} which can be efficiently exploited for high quality load balancing.
+
+\item \textbf{Parallel octrees: p4est, dendro, octor, etc. How do we plan to use p4est?}
+The idea of using SFCs for parallel partitioning of Quad-/Octrees is not new in itself.
+In \fcite{octor} authors first used the idea of using Morton curve (also known as Z-curve) for traversing the leaves of an Octree which is then used to index and distribute them between processors.
+Dendro \fcite{dendro paper} follows similar ideas for developing a parallel geometric multigrid on balanced octrees.
+More recently, authors in \fcite{p4est} extended these ideas to a collection, or a ``forest'', of octrees that are connected through a common, potentially unstructured, hexahedral grid.
+They then use a global Morton curve to partition the entire forest in parallel.
+Implementation of these algorithms, which were shown to scale to more than 200,000 processors, are publicly available through simple API provided by the \texttt{p4est} library \fcite{p4est website?}.
+In fact the algorithms presented in this paper are directly implemented on top of \texttt{p4est} library and we do not discuss any algorithm that is already covered by \texttt{p4est} implementation.
+Instead make appropriate references to \fcite{p4st paper} whenever necessary.
 
@@
 
-\item \textbf{How do we go about the parallization? Whats the organization of the paper}
+\item \textbf{Whats the organization of the paper}
 \end{enumerate}

### 2014-08-29 paper/tex/introduction.tex
More update on the introduction

@@
 One option is to replicate the entire grid on each processor and use off-the-self graph partitioners, e.g.
-Metis or ParMetis \fcite{metis/parmetis paper}, for load balancing and domain decomposition \fcite{origial deal.II paper}.
+Metis or ParMetis \fcite{metis/parmetis paper} or Zoltan \fcite{zoltan paper}, for load balancing and domain decomposition \fcite{origial deal.II paper}.
 This approach, however, is only scalable to a few hundred processor at best and is limited by the size of the grid itself.
@@
 \item \textbf{Parallel octrees: p4est, dendro, octor, etc. How do we plan to use p4est?}
-The idea of using SFCs for parallel partitioning of Quad-/Octrees is not new in itself.
-In \fcite{octor} authors first used the idea of using Morton curve (also known as Z-curve) for traversing the leaves of an Octree which is then used to index and distribute them between processors.
-Dendro \fcite{dendro paper} follows similar ideas for developing a parallel geometric multigrid on balanced octrees.
+The idea of using SFCs for parallel partitioning of Quad-/Octrees is not new in itself and has been tried by many researchers.
+For instance, Octor \fcite{octor} uses a Morton curve (also known as Z-curve) for traversing the leaves of an Octree which is then used to index and distribute them between processors and has been shown to scale up to 6200 cores \fcite{mantle convection octor}.
+Dendro \fcite{original dendro paper} is another example of parallel octree code in which similar ideas are used for parallel partitioning and development of a parallel geometric multigrid that has been scaled up to about 32000 cores \fcite{dendro parallel multigrid}.
 More recently, authors in \fcite{p4est} extended these ideas to a collection, or a ``forest'', of octrees that are connected through a common, potentially unstructured, hexahedral grid.
@@
 Implementation of these algorithms, which were shown to scale to more than 200,000 processors, are publicly available through simple API provided by the \texttt{p4est} library \fcite{p4est website?}.
-In fact the algorithms presented in this paper are directly implemented on top of \texttt{p4est} library and we do not discuss any algorithm that is already covered by \texttt{p4est} implementation.
-Instead make appropriate references to \fcite{p4st paper} whenever necessary.
+In fact the algorithms presented in this paper are directly implemented on top of \texttt{p4est} library and we do not discuss any algorithm that is already covered in \texttt{p4est paper}.
+Instead we focus on parallelization of level-set method on Quad-/Octrees and make appropriate references to \fcite{p4st paper} whenever necessary.
+
+\item \textbf{Parallel semi-lagrangian methods? What are the challenges? Whats out there?}
+Parallel level-set algorithms can be categorized in three separate groups: 1) Parallel advection algorithms, 2) Parallel reinitialization algorithms, and 3) Parallel motion under curvature algorithms.
+In this article we focus on developing algorithms for the first two categories.
+We also make use of the popular \texttt{PETSc} \fcite{petsc} library for solution of the linear systems that arise from the third category.
+Moreover, \texttt{PETSc} provides primitives, such as parallel ghosted vector and scatter/gather operations, which simplifies the implementation.
+
+Eulerian advection schemes can easily be parallelized but unfortunately are limited by the CFL condition.
+For adaptive grids, this could be very restrictive as the tree could be locally refined very deeply to capture quantities of interest.
+Semi-Lagrangian methods combine the unconditional stability of Lagrangian methods and the ease of use of Eulerian grids and have been successfully used for advecting the level-set function on Quad-/Octree grids before \fcite{Frederic's paper}.
+However, parallelizing the Semi-Lagrangian algorithm in a domain decomposition context is not an easy task.
+The reason for this is twofold: First, depending on the CFL number, the departure points may end up outside the ghost region and in remote processors that are potentially far away.
+This requires a very dynamic and nonuniform communication pattern which is complicated to implement.
+For adaptive grids situation is even more complicated due to the asymmetric nature of communication, i.e.
+every processor knows the message destination but no processor knows which processor to expect message from and how many.
+Second, depending on the CFL number, the distribution of departure points could be entirely independent of the original parallel partitioning.
+As a result large CFL numbers could result in very load-unbalanced distribution of departure points which restricts the scalability of the algorithm.
+Both of these problems, of course, could be avoided by choosing $\text{CFL} \le 1$ but that would defeat the purpose of using Semi-Lagrangian algorithm in the first place.
 
@@
 
-\item \textbf{Parallel semi-lagrangian methods? What are the challenges? Whats out there?}
-
 \item \textbf{Whats the organization of the paper}

### 2014-08-29 paper/examples/main_poisson_2d.cpp
sync

@@ -476,8 +476,6 @@ int main (int argc, char* argv[]){
 		ierr = VecDuplicate(phi, &err); CHKERRXX(ierr);
 
     /* initialize the poisson solver */
-    PetscLogEvent log_PoissonSolverNodeBased_global;
-    ierr = PetscLogEventRegister("PoissonSolverNodeBased::global                           ", 0, &log_PoissonSolverNodeBased_global); CHKERRXX(ierr);
 		for(int i=0; i<repeat; ++i)
 		{
 			w2.start("solve the poisson equation");
@@ -488,9 +486,7 @@ int main (int argc, char* argv[]){
 
 			/* solve the system */
 			MPI_Barrier(p4est->mpicomm);
-			ierr = PetscLogEventBegin(log_PoissonSolverNodeBased_global, sol, 0, 0, 0); CHKERRXX(ierr);
 			solver.solve(sol);
-			ierr = PetscLogEventEnd  (log_PoissonSolverNodeBased_global, sol, 0, 0, 0); CHKERRXX(ierr);
 
 			if(cmd.contains("compute_error"))
 			{

### 2014-08-29 paper/tex/levelset_method.tex
updated citation format

@@
 
-The work we present relies on the parallel octree library \texttt{p4est} \cite{burstedded:2011:p4est}.
+The work we present relies on the parallel octree library \texttt{p4est} \cite{Burstedde;Wilcox;Ghattas:11:p4est:-Scalable-Algo}.
 It is scalable implementation of the general octree structure in a massively parallel mpi environment.
-We limit our presentation of \texttt{p4est} to the relevant information for our application, for more details the reader is invited to read the original article \cite{burstedded:2011:p4est}.
+We limit our presentation of \texttt{p4est} to the relevant information for our application, for more details the reader is invited to read the original article \cite{Burstedde;Wilcox;Ghattas:11:p4est:-Scalable-Algo}.
 

### 2014-08-29 paper/tex/parallel_algorithms.tex
updated citation format

@@
 The reinitialization procedure transforms an arbitrary function into a signed distance function to its zero contour.
-One of the popular methods to achieve this results is to solve the reinitialization equation \ref{eq::reinitialization} \cite{Sussman:1994:levelsetincompressible, Min_Gibou:2007:Level_Set_Adaptive}.
-Other methods, such as the Fast Sweeping \cite{??} and the Fast Marching methods \cite{Sethian:1996:FMM, chopp:2001:fmm}, are based on the wave propagation nature of the reinitialization front and provide faster implementations of the reinitialization procedure in a serial environment.
+One of the popular methods to achieve this results is to solve the reinitialization equation \ref{eq::reinitialization} \cite{Sussman;Smereka;Osher:94:A-Level-Set-Approach, Min;Gibou:07:A-second-order-accur}.
+Other methods, such as the Fast Sweeping \cite{Zhao:04:A-Fast-Sweeping-Meth} and the Fast Marching methods \cite{Sethian:96:A-Fast-Marching-Leve, Chopp:01:Some-improvements-of}, are based on the wave propagation nature of the reinitialization front and provide faster implementations of the reinitialization procedure in a serial environment.
 However, their parallelization is an involved process and we opt for solving the reinitialization equation.
 
-We base our approach on Min and Gibou \cite{Min_Gibou:2007:Level_Set_Adaptive} and discretize the reinitialization equation as
+We base our approach on Min and Gibou \cite{Min;Gibou:07:A-second-order-accur} and discretize the reinitialization equation as
 
@@
 with $a^+=\max(a,0)$ and $a^-=\min(a,0)$.
-The one-sided derivatives $D^{+/-}_{x/y}\phi$ are discretized by second order accurate finite differences as explained by Min and Gibou in \cite{Min_Gibou:2007:Level_Set_Adaptive}.
-Following the presentation in \cite{Min_Gibou:2007:Level_Set_Adaptive} which incorporates the subcell resolution introduced by Russo and Smereka \cite{Russo_Smereka}, we include the initial position of the interface, given by $\phi^0$, in the computation of the one sided derivatives next to the interface in order to preserve the interface location.
+The one-sided derivatives $D^{+/-}_{x/y}\phi$ are discretized by second order accurate finite differences as explained by Min and Gibou in \cite{Min;Gibou:07:A-second-order-accur}.
+Following the presentation in \cite{Min;Gibou:07:A-second-order-accur} which incorporates the subcell resolution introduced by Russo and Smereka \cite{Russo;Smereka:00:A-Remark-on-Computin}, we include the initial position of the interface, given by $\phi^0$, in the computation of the one sided derivatives next to the interface in order to preserve the interface location.
 Finally, we choose a simple Euler step for the time discretization.

### 2014-08-29 paper/examples/main_poisson_2d.cpp
small fix

@@ -317,8 +317,8 @@ int main (int argc, char* argv[]){
     cmd.add_option("lmax", "the max level of the tree");
     cmd.add_option("sp", "number of splits to apply to the min and max level");
     cmd.add_option("save_vtk", "give this flag to save a vtk file of the solution and error");
-		cmd.add_option("repeat", "number of repeat to perform");
-		cmd.add_option("compute_error", "compute the error on the computed solution");
+    cmd.add_option("repeat", "number of repeat to perform");
+    cmd.add_option("compute_error", "compute the error on the computed solution");
     cmd.parse(argc, argv);
     cmd.print();
 
@@ -330,7 +330,7 @@ int main (int argc, char* argv[]){
     nb_splits         = cmd.get("sp" , 0);
     min_level         = cmd.get("lmin"      , 3);
     max_level         = cmd.get("lmax"      , 8);
-		int repeat        = cmd.get("repeat"    , 1);
+    int repeat        = cmd.get("repeat"    , 1);
 
 #ifdef P4_TO_P8
     CF_3 *bc_wall_value, *bc_interface_value;
@@ -472,136 +472,136 @@ int main (int argc, char* argv[]){
     bc.setWallTypes(*wall_bc);
     bc.setWallValues(*bc_wall_value);
 
-		Vec err;
-		ierr = VecDuplicate(phi, &err); CHKERRXX(ierr);
+    Vec err;
+    ierr = VecDuplicate(phi, &err); CHKERRXX(ierr);
 
     /* initialize the poisson solver */
-		for(int i=0; i<repeat; ++i)
-		{
-			w2.start("solve the poisson equation");
-			PoissonSolverNodeBase solver(&ngbd);
-			solver.set_phi(phi);
-			solver.set_rhs(rhs);
-			solver.set_bc(bc);
+    for(int i=0; i<repeat; ++i)
+    {
+      w2.start("solve the poisson equation");
+      PoissonSolverNodeBase solver(&ngbd);
+      solver.set_phi(phi);
+      solver.set_rhs(rhs);
+      solver.set_bc(bc);
 
[…]

### 2014-08-29 paper/tex/p4est.tex

@@
 
-\indent The level-set method is a method to capture interfaces implicitly by constructing a function $\phi$, called the level-set function, that is negative on on side of the domain $\Omega^-$ and positive on the other side $\Omega^+$, thus defining the interface $\Gamma$ as the zero contour of $\phi$, $\Gamma = \{\mathbf{x} \vert \phi(\mathbf{x}=0\}$.\\
+\indent The level-set method is a method to capture interfaces implicitely by constructing a function $\phi$, called the level-set function, that is negative on on side of the domain $\Omega^-$ and positive on the other side $\Omega^+$, thus defining the interface $\Gamma$ as the zero contour of $\phi$, $\Gamma = \{\mathbf{x} \vert \phi(\mathbf{x}=0\}$.\\
 
@@
 
-The work we present relies on the parallel octree library \texttt{p4est} \cite{Burstedde;Wilcox;Ghattas:11:p4est:-Scalable-Algo}.
+The work we present relies on the parallel octree library \texttt{p4est} \cite{burstedded:2011:p4est}.
 It is scalable implementation of the general octree structure in a massively parallel mpi environment.
-We limit our presentation of \texttt{p4est} to the relevant information for our application, for more details the reader is invited to read the original article \cite{Burstedde;Wilcox;Ghattas:11:p4est:-Scalable-Algo}.
+We limit our presentation of \texttt{p4est} to the relevant information for our application, for more details the reader is invited to read the original article \cite{burstedded:2011:p4est}.
 
@@
 The reinitialization procedure transforms an arbitrary function into a signed distance function to its zero contour.
-One of the popular methods to achieve this results is to solve the reinitialization equation \ref{eq::reinitialization} \cite{Sussman;Smereka;Osher:94:A-Level-Set-Approach, Min;Gibou:07:A-second-order-accur}.
-Other methods, such as the Fast Sweeping \cite{Zhao:04:A-Fast-Sweeping-Meth} and the Fast Marching methods \cite{Sethian:96:A-Fast-Marching-Leve, Chopp:01:Some-improvements-of}, are based on the wave propagation nature of the reinitialization front and provide faster implementations of the reinitialization procedure in a serial environment.
+One of the popular methods to achieve this results is to solve the reinitialization equation \ref{eq::reinitialization} \cite{Sussman:1994:levelsetincompressible, Min_Gibou:2007:Level_Set_Adaptive}.
+Other methods, such as the Fast Sweeping \cite{??} and the Fast Marching methods \cite{Sethian:1996:FMM, chopp:2001:fmm}, are based on the wave propagation nature of the reinitialization front and provide faster implementations of the reinitialization procedure in a serial environment.
 However, their parallelization is an involved process and we opt for solving the reinitialization equation.
 
-We base our approach on Min and Gibou \cite{Min;Gibou:07:A-second-order-accur} and discretize the reinitialization equation as
+We base our approach on Min and Gibou \cite{Min_Gibou:2007:Level_Set_Adaptive} and discretize the reinitialization equation as
 
@@
 with $a^+=\max(a,0)$ and $a^-=\min(a,0)$.
-The one-sided derivatives $D^{+/-}_{x/y}\phi$ are discretized by second order accurate finite differences as explained by Min and Gibou in \cite{Min;Gibou:07:A-second-order-accur}.
-Following the presentation in \cite{Min;Gibou:07:A-second-order-accur} which incorporates the subcell resolution introduced by Russo and Smereka \cite{Russo;Smereka:00:A-Remark-on-Computin}, we include the initial position of the interface, given by $\phi^0$, in the computation of the one sided derivatives next to the interface in order to preserve the interface location.
+The one-sided derivatives $D^{+/-}_{x/y}\phi$ are discretized by second order accurate finite differences as explained by Min and Gibou in \cite{Min_Gibou:2007:Level_Set_Adaptive}.
+Following the presentation in \cite{Min_Gibou:2007:Level_Set_Adaptive} which incorporates the subcell resolution introduced by Russo and Smereka \cite{Russo_Smereka}, we include the initial position of the interface, given by $\phi^0$, in the computation of the one sided derivatives next to the interface in order to preserve the interface location.
 Finally, we choose a simple Euler step for the time discretization.
@@
 
-\section*{References}
+\newpage
 \bibliographystyle{plain}
-\addcontentsline{toc}{section}{\refname}
-\bibliography{/home/guittet/code/CASL_Library/references.bib}
+\addcontentsline{toc}{section}{\refname}\bibliography{references}
 

### 2014-08-29 paper/examples/main_poisson_2d.cpp
compatibility update

@@ -394,7 +394,7 @@ int main (int argc, char* argv[]){
     p4est_refine(p4est, P4EST_TRUE, refine_levelset_cf, NULL);
 
     /* partition the p4est */
-    p4est_partition(p4est, NULL);
+    p4est_partition(p4est, P4EST_FALSE, NULL);
 
     /* create the ghost layer */
     p4est_ghost_t* ghost = p4est_ghost_new(p4est, P4EST_CONNECT_FULL);
@@ -487,6 +487,13 @@ int main (int argc, char* argv[]){
       /* solve the system */
       MPI_Barrier(p4est->mpicomm);
       solver.solve(sol);
+
+      if(i==repeat-1 && bc_interface_type==NEUMANN && bc_wall_type==NEUMANN)
+      {
+        PetscPrintf(p4est->mpicomm, "Neumann BC only! Shifting solution\n\n");
+        solver.shift_to_exact_solution(sol, uex);
+      }
+
     }
 
     if(cmd.contains("compute_error"))
@@ -494,12 +501,6 @@ int main (int argc, char* argv[]){
       ierr = VecGhostUpdateBegin(sol, INSERT_VALUES, SCATTER_FORWARD); CHKERRXX(ierr);
       ierr = VecGhostUpdateEnd  (sol, INSERT_VALUES, SCATTER_FORWARD);   CHKERRXX(ierr);
       w2.stop(); w2.read_duration();
-
-      if(bc_interface_type==NEUMANN && bc_wall_type==NEUMANN)
-      {
-        PetscPrintf(p4est->mpicomm, "Neumann BC only! Shifting solution\n\n");
-        solver.shift_to_exact_solution(sol, uex);
-      }
 
       /* prepare for output */
       double *sol_p, *phi_p, *uex_p;

### 2014-09-03 paper/tex/introduction.tex
updated introduction and added refs (see details)
currently newer refs are added in the refs.bib until we finialize on
which ones we actually want to retain. For the older ones we use the
master ref for CASL.

To cope with the fact that the file could be in different locations on
different machines, please add a symlink casl.bib to point to the actual
reference file on you machine. This file was not added to the repo since
the address could be different.

A better approach is to put the reference file in a system-provided
location that can be accessed by latex without the need for any relative

@@
 
-\item \textbf{Parallel semi-lagrangian methods? What are the challenges? Whats out there?}
+\item \textbf{Parallel advection methods? What are the challenges?}
 Parallel level-set algorithms can be categorized in three separate groups: 1) Parallel advection algorithms, 2) Parallel reinitialization algorithms, and 3) Parallel motion under curvature algorithms.
@@
 
-\item \textbf{Parallel reinitialization methods? What are the challenges? Whats out there?}
+\item \textbf{Parallel reinitialization methods? What are the challenges?}
+In several applications of the level-set method, it is desirable for the level-set function to be a signed distance function, i.e.
+$|\nabla \phi| = 1$.
+There are generally two approaches to enforce this property: 1) Solving the pseudo-time transient reinitialization equation $\phi_\tau + S(\phi_0)\left(|\nabla \phi| - 1\right) = 0$ (c.f.
+section \ref{section::levelset method}) \fcite{osher's book, smerka's paper?} or 2) Solving the static Eikonal equation $F(x)|\nabla\phi| = 1$ with constant speed function $F(x) \equiv 1$.
+The transient reinitialization equation can be solved using explicit finite differences and thus can easily be parallelized with a domain decomposition approach.
+Moreover, only a few iterations may be needed if the signed-distance property is only required close to the interface \fcite{Frederic's levelset paper}.
+This is the approach we have chosen in this paper.
+If, however, the signed-distance property is required in the entire domain solving the static Eikonal equation is more computationally efficient.
+Unfortunately, however, the most popular algorithm for solving the Eikonal equation, i.e.
+the Fast Marching Method \fcite{sethian's book, original FMM paper}, is inherently sequential due to causal relationship between grid points and cannot be easily parallelized.
+The Fast Sweeping Method (FSM) \fcite{Zhao's paper} is an alternative method for solving the static Eikonal equation iteratively which can be more computationally efficient for simple choices of speed function, e.g.
+as in this context, and for simple interfaces.
+Moreover, FSM has more potential for parallelization compared to the FMM.
+
+\item \textbf{literature review on parallel levelset}
+Large body of research has been devoted to parallelization of the level-set methods and here we review some of the more relevant works.
 

### 2014-09-04 paper/tex/introduction.tex
Introduction almost finished -- needs cleaning.

@@
 \item \textbf{What is level-set and why is it important?}
-The level-set method, originally proposed by Sethian and Osher \fcite{classic LS paper}, is a popular and powerful framework for tracking arbitrary interfaces that undergo complicated topological changes.
-As a result, the level-set method has wide range of application such as multiphase flows, moving boundary problems, image segmentation, and computer graphics graphics, to only name a few \fcite{fedkiw/osher's book, sethian's book}.
+The level-set method, originally proposed by Sethian and Osher \cite{Osher;Sethian:88:Fronts-Propagating-w}, is a popular and powerful framework for tracking arbitrary interfaces that undergo complicated topological changes.
+As a result, the level-set method has wide range of application such as multiphase flows, moving boundary problems, image segmentation, and computer graphics graphics, to only name a few \cite{Osher;Fedkiw:01:Level-Set-Methods:-A,Sethian:99:Level-set-methods-an}.
 
@@
 First, compared to an explicit method, e.g.
-front tracking \fcite{front tracking refs}, level-set method is typically less accurate and mass conservation could be a problem although progress has been made in resolving this issue \fcite{Enright/Fedkiw paper, other?}.
+front tracking \fcite{front tracking refs}, level-set method is typically less accurate and mass conservation could be a problem although progress has been made in resolving this issue \cite{Enright;Fedkiw;Ferziger;etal:02:A-Hybrid-Particle-Le}\fcite{other?}.
 Second, the level-set function has to be defined in a higher dimensional space compared to the interface.
@@
 This problem, however, could be handled by computing the level-set only close to the interface, e.g.
-as in the narrow-band level-set method \fcite{narrow band, the hash table paper?}.
+as in the narrow-band level-set method \cite{Brun;Guittet;Gibou:12:A-local-level-set-me}\fcite{narrow band ?}.
 
@@
 Another approach that can address both problems is the use of local grid refinement.
-In \fcite{Frederic paper}, authors present second-order accurate level-set methods on Cartesian Quadtree (two dimensions) and Octree (three dimensions) grids.
+In \cite{Min;Gibou:07:A-second-order-accur}, authors present second-order accurate level-set methods on Cartesian Quadtree (two dimensions) and Octree (three dimensions) grids.
 Use of tree-based Cartesian adaptive grids in the context of level-set method is quite advantageous because 1) It gives fine-grain control over error, which typically occurs close to the interface, 2) It can effectively reduce the dimensionality of the problem by focusing most of the cells close to the interface, and 3) Construction and handling of the tree is quite simple in the presence of an interface which defines a an ideal metric for refinement.
@@
 Dendro \fcite{original dendro paper} is another example of parallel octree code in which similar ideas are used for parallel partitioning and development of a parallel geometric multigrid that has been scaled up to about 32000 cores \fcite{dendro parallel multigrid}.
-More recently, authors in \fcite{p4est} extended these ideas to a collection, or a ``forest'', of octrees that are connected through a common, potentially unstructured, hexahedral grid.
+More recently, authors in \cite{Burstedde;Wilcox;Ghattas:11:p4est:-Scalable-Algo} extended these ideas to a collection, or a ``forest'', of octrees that are connected through a common, potentially unstructured, hexahedral grid.
 They then use a global Morton curve to partition the entire forest in parallel.
@@
 In fact the algorithms presented in this paper are directly implemented on top of \texttt{p4est} library and we do not discuss any algorithm that is already covered in \texttt{p4est paper}.
-Instead we focus on parallelization of level-set method on Quad-/Octrees and make appropriate references to \fcite{p4st paper} whenever necessary.
+Instead we focus on parallelization of level-set method on Quad-/Octrees and make appropriate references to \cite{Burstedde;Wilcox;Ghattas:11:p4est:-Scalable-Algo} whenever necessary.
 
@@
 For adaptive grids, this could be very restrictive as the tree could be locally refined very deeply to capture quantities of interest.
-Semi-Lagrangian methods combine the unconditional stability of Lagrangian methods and the ease of use of Eulerian grids and have been successfully used for advecting the level-set function on Quad-/Octree grids before \fcite{Frederic's paper}.
+Semi-Lagrangian methods combine the unconditional stability of Lagrangian methods and the ease of use of Eulerian grids and have been successfully used for advecting the level-set function on Quad-/Octree grids before \cite{Min;Gibou:07:A-second-order-accur}.
 However, parallelizing the Semi-Lagrangian algorithm in a domain decomposition context is not an easy task.
@@
 There are generally two approaches to enforce this property: 1) Solving the pseudo-time transient reinitialization equation $\phi_\tau + S(\phi_0)\left(|\nabla \phi| - 1\right) = 0$ (c.f.
-section \ref{section::levelset method}) \fcite{osher's book, smerka's paper?} or 2) Solving the static Eikonal equation $F(x)|\nabla\phi| = 1$ with constant speed function $F(x) \equiv 1$.
+section \ref{section::levelset method}) \cite{Osher;Fedkiw:01:Level-Set-Methods:-A} \fcite{smerka's paper?} or 2) Solving the static Eikonal equation $F(x)|\nabla\phi| = 1$ with constant speed function $F(x) \equiv 1$.
 The transient reinitialization equation can be solved using explicit finite differences and thus can easily be parallelized with a domain decomposition approach.
-Moreover, only a few iterations may be needed if the signed-distance property is only required close to the interface \fcite{Frederic's levelset paper}.
+Moreover, only a few iterations may be needed if the signed-distance property is only required close to the interface \cite{Min;Gibou:07:A-second-order-accur}.
 This is the approach we have chosen in this paper.
@@
 Unfortunately, however, the most popular algorithm for solving the Eikonal equation, i.e.
-the Fast Marching Method \fcite{sethian's book, original FMM paper}, is inherently sequential due to causal relationship between grid points and cannot be easily parallelized.
-The Fast Sweeping Method (FSM) \fcite{Zhao's paper} is an alternative method for solving the static Eikonal equation iteratively which can be more computationally efficient for simple choices of speed function, e.g.
+the Fast Marching Method \cite{Sethian:96:A-Fast-Marching-Leve,Sethian:99:Level-set-methods-an}, is inherently sequential due to causal relationship between grid points and cannot be easily parallelized.
+The Fast Sweeping Method (FSM) \cite{Zhao:05:A-fast-sweeping-meth} is an alternative method for solving the static Eikonal equation iteratively which can be more computationally efficient for simple choices of speed function, e.g.
 as in this context, and for simple interfaces.
@@
 \item \textbf{literature review on parallel levelset}
-Large body of research has been devoted to parallelization of the level-set methods and here we review some of the more relevant works.
+Parallel semi-Lagrangian algorithms have been of interest for atmospheric studies in the past.
+A simple domain decomposition technique was used in \cite{Thomas;Cote:95:Massively-parallel-s} where the width of ghost layer is fixed based on the maximum CFL number.
+At large CFL numbers, this technique potentially leads to large communication volume which can limit the scalability.
+Nonetheless good scaling was report for small CFL numbers ($\text{CFL} \le 2$).
+In \cite{Drake;Foster;Michalakes;etal:95:Design-and-performan} authors propose a more sophisticated domain decomposition approach which uses a ``dynamic ghost layer''.
+Here width of the ghost layer is dynamically determined at runtime based on information from previous time steps.
+Nonetheless this approach seems to suffer from excessive communication overhead at larger number of processors.
+More recently, authors in \cite{White-III;Dongarra:11:High-performance-hig} used a domain decomposition strategy on a cubed sphere but with a single layer of ghost nodes.
+Interpolation on remote processors is handled by sending query points to the corresponding processor and asking for the interpolated result.
+This approach seems to provide good scalability for transporting a single tracer up to about 1000 cores for $\text{CFL} \sim 10$.
+At higher CFL numbers the method begins to loose scalability due an increase in communication volume.
+Moreover this strategy is susceptible to load imbalances at higher CFL numbers and for non-uniform velocity fields.
+Although in this article we are mainly interested in parallel semi-Lagrangian methods, one could alternatively resort to finite difference or finite element discretization methods which are easier to parallelize.
+Indeed several algorithms has been proposed with applications in dendritic crystal growth \cite{Wang;Chang;Kale;etal:06:Parallelization-of-a}, multiphase flows \cite{Sussman:05:A-parallelized-adapt, Fortmeier;Bucker:11:A-parallel-strategy-, Rodriguez;Sahni;Lahey-Jr;etal:13:A-parallel-adaptive-}, atomization process \cite{Herrmann:10:A-parallel-Eulerian-}, and image segmentation on GPUs \cite{Lefohn;Cates;Whitaker:03:Interactive-GPU-base,Cates;Lefohn;Whitaker:04:GIST:-an-interactive,Roberts;Packer;Sousa;etal:10:A-work-efficient-GPU}.
+
+As mentioned earlier, in this article we opt for the parallel implementation of the transient reinitialization equation which is considerably easier to parallelize.
+Nonetheless it is important to review some of the recent progress in parallelizing solvers for the static Eikonal equation.
+One of the earliest attempt in parallelizing the FMM is reported in \cite{Herrmann:03:A-domain-decompositi} where a domain decomposition algorithm was introduced.
+Unlike the serial FMM, however, parallel FMM potentially requires multiple iteration or ``roolback'' operations to enforce the causality property across processors.
+Similar ideas are described in details in \cite{Tugurlan:08:Fast-marching-method}.
+It should be noted that number of iterations for the parallel FMM to converge greatly depends on the complexity of the interface and also the parallel partitioning and, in general, fewer iterations are required if domains are more or less parallel to the interface normals.
+Due to the nature of Eikonal equation, shared memory machines might be a better environment for parallelization.
+For instance, in \cite{Breus;Cristiani;Gwosdek;etal:11:An-adaptive-domain-d} authors use an ``adaptive'' technique in which individual threads implicitly define a domain decomposition at runtime.
+Unfortunately, however, this approach does not seem to be more effective than a simple static domain decomposition and is not portable to a distributed memory machine.
+In \cite{Zhao:07:Parallel-implementat} a parallel FSM method was presented for the fist time.
+However, a more scalable FSM was more recently proposed in \cite{Detrixhe;Gibou;Min:13:A-parallel-fast-swee} where the Cuthill-McKee node numbering was utilize to expose more parallelism.
+More recently a two-scale, hybrid FMM-FSM was presented in \cite{Chacon;Vladimirsky:13:A-parallel-Heap-Cell} which, albeit being more complicated to implement, promises even better scalability.
+Finally, a parallel Fast Iterative Method (FIM) was proposed in \cite{Jeong;Whitaker:08:A-fast-iterative-met}.
+The FIM is similar to FMM in that it also maintains a list of ``active nodes''.
+Unlike FMM, however, FIM avoids sorting the list and allows for concurrent updating of all nodes in an iterative fashion.
 

### 2014-09-11 examples/biomol/bio_molecule.cpp
switching to work on paper

@@ -15,7 +15,7 @@
 
 using namespace std;
 
-#define MAX_BLOCK_SIZE 16
+#define MAX_BLOCK_SIZE 32
 #define FAST_SURFACE_COMPUTATION
 
 istream& operator >> (istream& is, Atom& atom) {
@@ -282,12 +282,20 @@ void BioMolecule::construct_SES_by_reinitialization(p4est_t* &p4est, p4est_nodes
   const p4est_connectivity_t *connectivity = p4est->connectivity;
 
   // split based on the SAS distance
-  splitting_criteria_threshold_cf_t sp_thr(sp->min_lvl, sp->max_lvl, -1.5*rp_, 1.5*rp_, this, sp->lip);
+  parStopWatch w;
+  w.start("making intial tree");
+  splitting_criteria_threshold_cf_t sp_thr(sp->min_lvl, sp->max_lvl, -1.0*rp_, 1.5*rp_, this, sp->lip);
   p4est->user_pointer = &sp_thr;
-  my_p4est_refine(p4est, P4EST_TRUE, refine_threshold_cf, NULL);
 
-  // partition the p4est
-  my_p4est_partition(p4est, NULL);
+	// refine and partition
+	// Note: Using recursive on large molecules causes the whole thing to be build in serial on 
+	// one processor which is obviously not scalable 
+	for (int l = 0; l<sp->max_lvl; l++){
+	  my_p4est_refine(p4est, P4EST_FALSE, refine_threshold_cf, NULL);
+		my_p4est_partition(p4est, NULL);
+	}
+
+  PetscPrintf(p4est->mpicomm, "num of global quadrants = %ld\n", p4est->global_num_quadrants);
 
   // create the ghost layer
   ghost = p4est_ghost_new(p4est, P4EST_CONNECT_FULL);
@@ -298,6 +306,8 @@ void BioMolecule::construct_SES_by_reinitialization(p4est_t* &p4est, p4est_nodes
   // calculate the SAS on
   PetscErrorCode ierr = VecCreateGhostNodes(p4est, nodes, &phi); CHKERRXX(ierr);
   sample_cf_on_nodes(p4est, nodes, *this, phi);
+  w.stop(); w.read_duration();
+
 
   // subtract off the probe radius to get SES
   my_p4est_hierarchy_t hierarchy(p4est, ghost, &brick);

### 2014-09-11 examples/biomol/main_3d.cpp

@@ -81,7 +81,6 @@ int main(int argc, char *argv[]) {
     p4est_gloidx_t global_num_quadrants = p4est->global_num_quadrants;
     p4est_gloidx_t global_num_nodes = 0;
     for (int i = 0; i<p4est->mpisize; i++){
-      PetscPrintf(p4est->mpicomm, "%ld\n", nodes->global_owned_indeps[i]);
       global_num_nodes += nodes->global_owned_indeps[i];
     }
 

### 2014-09-11 paper/tex/introduction.tex

--- 2014-08-29 paper/tex/introduction.tex
@@
 \section{Introduction}
-\begin{enumerate}
-\item \textbf{What is level-set and why is it important?}
-The level-set method, originally proposed by Sethian and Osher \fcite{classic LS paper}, is a popular and powerful framework for tracking arbitrary interfaces that undergo complicated topological changes.
-As a result, the level-set method has wide range of application such as multiphase flows, moving boundary problems, image segmentation, and computer graphics graphics, to only name a few \fcite{fedkiw/osher's book, sethian's book}.
-
-\item \textbf{What are the main issues of using level-set method?}
-What makes level-set method powerful and easy to use is that in this method location of the interface is defined implicitly on an underlying grid.
+The level-set method, originally proposed by Sethian and Osher \cite{Osher;Sethian:88:Fronts-Propagating-w}, is a popular and powerful framework for tracking arbitrary interfaces that undergo complicated topological changes.
+As a result, the level-set method has wide range of application such as multiphase flows, moving boundary problems, image segmentation, and computer graphics graphics \cite{Osher;Fedkiw:01:Level-Set-Methods:-A,Sethian:99:Level-set-methods-an}.
+An important feature which makes this method powerful and easy to use is that the location of interface is defined implicitly on an underlying grid.
 This convenience, however, comes at a price.
 First, compared to an explicit method, e.g.
-front tracking \fcite{front tracking refs}, level-set method is typically less accurate and mass conservation could be a problem although progress has been made in resolving this issue \fcite{Enright/Fedkiw paper, other?}.
+front tracking \cite{Juric:96:A-Front-Tracking-Met, Tryggvason;Bunner;Esmaeeli;etal:01:A-Front-Tracking-Met}, level-set method is typically less accurate and mass conservation could be a problem; although progress has been made in resolving this issue \cite{Enright;Fedkiw;Ferziger;etal:02:A-Hybrid-Particle-Le}.
 Second, the level-set function has to be defined in a higher dimensional space compared to the interface.
 If only the location of interface is needed, the added dimension greatly increases the overall computational cost.
-This problem, however, could be handled by computing the level-set only close to the interface, e.g.
-as in the narrow-band level-set method \fcite{narrow band, the hash table paper?}.
+One way to avoid this problem is by computing the level-set only close to the interface, e.g.
+as in the narrow-band level-set method \cite{Adalsteinsson;Sethian:95:A-Fast-Level-Set-Met} or, more recently, by using a hash table to restrict both computation and storage requirements \cite{Brun;Guittet;Gibou:12:A-local-level-set-me}.
 
-\item \textbf{Why using AMR helps? How do we plan to extend that paper here?}
 Another approach that can address both problems is the use of local grid refinement.
-In \fcite{Frederic paper}, authors present second-order accurate level-set methods on Cartesian Quadtree (two dimensions) and Octree (three dimensions) grids.
-Use of tree-based Cartesian adaptive grids in the context of level-set method is quite advantageous because 1) It gives fine-grain control over error, which typically occurs close to the interface, 2) It can effectively reduce the dimensionality of the problem by focusing most of the cells close to the interface, and 3) Construction and handling of the tree is quite simple in the presence of an interface which defines a an ideal metric for refinement.
-Although use of adaptive grids dramatically reduces the computational time, performing high-resolution three dimensional calculations of complex interfacial problems, e.g.
-crystal growth in binary alloys \fcite{Maxime's paper}, is still challenging.
-In this paper we extend these algorithms by proposing parallel implementations on distributed memory machines using domain decomposition technique and Message Passing Interface (MPI) library.
+In \cite{Strain:99:Tree-Methods-for-Mov} the idea of using tree-based grids for level-set calculations was first introduced and later extended in \cite{Losasso;Gibou;Fedkiw:04:Simulating-Water-and} for graphics application.
+More recently, authors in \cite{Min;Gibou:07:A-second-order-accur} proposed second-order accurate level-set methods on Quadtree (two dimensions) and Octree (three dimensions) grids.
+Use of tree-based adaptive grids in the context of level-set method is quite advantageous because: 1) It gives fine-grain control over error, which typically occurs close to the interface and 2) It can effectively reduce the dimensionality of the problem by focusing most of the grid cells close to the interface.
+Fortunately, construction of the tree is quite simple in the presence of an interface that naturally defines an ideal metric for refinement.
+Although use of adaptive grids can dramatically reduce the computational cost, performing high-resolution three dimensional calculations of complex interfacial problems, e.g.
+crystal growth in binary alloys \cite{Theillard;Gibou;Pollock:14:A-Sharp-Computationa}, could take a very long time in serial.
+In this paper we extend these algorithms by proposing parallel algorithms for distributed memory machines using domain decomposition technique.
 
-\item \textbf{What is the main challenge for parallelization?}
-One of the main challenges in parallelizing level-set algorithms on adaptive grids, is handling the grid itself in parallel.
-One option is to replicate the entire grid on each processor and use off-the-self graph partitioners, e.g.
-Metis or ParMetis \fcite{metis/parmetis paper} or Zoltan \fcite{zoltan paper}, for load balancing and domain decomposition \fcite{origial deal.II paper}.
+One of the main challenges in parallelizing level-set algorithms on adaptive grids is handling the grid itself.
+One option is to replicate the entire grid on each processor and use off-the-shelf graph partitioners, e.g.
+ParMetis \cite{Karypis;Kumar:98:A-parallel-algorithm} or Zoltan \cite{Boman;Catalyurek;Chevalier;etal:12:The-Zoltan-and-Isorr}, for load balancing and domain decomposition.
+For instance, this was the approach originally taken by the \texttt{deal.II} library \cite{Bangerth;Hartmann;Kanschat:07:deal.II----a-General}.
 This approach, however, is only scalable to a few hundred processor at best and is limited by the size of the grid itself.
-Moreover, use of a general-purpose graph partitioner adds extra overhead that limits the overall scalability even further.
-Interestingly, Quadtrees and Octrees have nice spatial ordering that naturally lead to the concept of Space Filling Curves (SFC) \fcite{SFC reference} which can be efficiently exploited for high quality load balancing.
+Moreover, use of a general-purpose graph partitioner adds extra overhead that can limit the overall scalability even further.
+Interestingly, tree-based grids have nice spatial ordering that naturally lead to the concept of space-filling curves (SFCs) which can be efficiently exploited for parallel load balancing \cite{Aluru;Sevilgen:97:Parallel-domain-deco,Campbell;Devine;Flaherty;etal:03:Dynamic-octree-load-}.
 
-\item \textbf{Parallel octrees: p4est, dendro, octor, etc. How do we plan to use p4est?}
 The idea of using SFCs for parallel partitioning of Quad-/Octrees is not new in itself and has been tried by many researchers.
-For instance, Octor \fcite{octor} uses a Morton curve (also known as Z-curve) for traversing the leaves of an Octree which is then used to index and distribute them between processors and has been shown to scale up to 6200 cores \fcite{mantle convection octor}.
-Dendro \fcite{original dendro paper} is another example of parallel octree code in which similar ideas are used for parallel partitioning and development of a parallel geometric multigrid that has been scaled up to about 32000 cores \fcite{dendro parallel multigrid}.
-More recently, authors in \fcite{p4est} extended these ideas to a collection, or a ``forest'', of octrees that are connected through a common, potentially unstructured, hexahedral grid.
-They then use a global Morton curve to partition the entire forest in parallel.
-Implementation of these algorithms, which were shown to scale to more than 200,000 processors, are publicly available through simple API provided by the \texttt{p4est} library \fcite{p4est website?}.
-In fact the algorithms presented in this paper are directly implemented on top of \texttt{p4est} library and we do not discuss any algorithm that is already covered in \texttt{p4est paper}.
-Instead we focus on parallelization of level-set method on Quad-/Octrees and make appropriate references to \fcite{p4st paper} whenever necessary.
+For instance, \texttt{Octor} \cite{Tu;OHallaron;Ghattas:05:Scalable-parallel-oc} uses a Morton curve (also known as Z-curve) for traversing the leaves of an Octree for indexing and load balancing and has been scaled up to 62,000 cores \cite{Burstedde;Ghattas;Gurnis;etal:08:Scalable-adaptive-ma}.
+\texttt{Dendro} \cite{Sampath;Adavani;Sundar;etal:08:Dendro:-parallel-alg} is another example of an octree code in which similar ideas are used for parallel partitioning and development of a parallel geometric multigrid that has been scaled up to about 32,000 cores \cite{Sampath;Biros:10:A-parallel-geometric}.
+More recently, authors in \cite{Burstedde;Wilcox;Ghattas:11:p4est:-Scalable-Algo} extended these ideas to a collection, or a ``forest'', of octrees that are connected through a common, potentially unstructured, hexahedral grid.
+This forest is then partitioned in parallel using a global Morton curve.
+Implementation of these algorithms, which were shown to scale to more than 200,000 processors, are publicly available through simple API provided by the \texttt{p4est} library \cite{p4est-github}.
+In fact the algorithms presented in this paper are directly implemented on top of \texttt{p4est} library and we do not discuss any algorithm that is already covered in \cite{Burstedde;Wilcox;Ghattas:11:p4est:-Scalable-Algo}.
+We also make use of the popular \texttt{PETSc} \cite{Balay;Abhyankar;Adams;etal:14:PETSc-Web-page} library for linear algebra and its parallel primitives, such as parallel ghosted vector and scatter/gather operations, which simplifies the implementation.
 
-\item \textbf{Parallel semi-lagrangian methods? What are the challenges? Whats out there?}
-Parallel level-set algorithms can be categorized in three separate groups: 1) Parallel advection algorithms, 2) Parallel reinitialization algorithms, and 3) Parallel motion under curvature algorithms.
-In this article we focus on developing algorithms for the first two categories.
-We also make use of the popular \texttt{PETSc} \fcite{petsc} library for solution of the linear systems that arise from the third category.
-Moreover, \texttt{PETSc} provides primitives, such as parallel ghosted vector and scatter/gather operations, which simplifies the implementation.
-
-Eulerian advection schemes can easily be parallelized but unfortunately are limited by the CFL condition.
-For adaptive grids, this could be very restrictive as the tree could be locally refined very deeply to capture quantities of interest.
-Semi-Lagrangian methods combine the unconditional stability of Lagrangian methods and the ease of use of Eulerian grids and have been successfully used for advecting the level-set function on Quad-/Octree grids before \fcite{Frederic's paper}.
-However, parallelizing the Semi-Lagrangian algorithm in a domain decomposition context is not an easy task.
+Parallel level-set algorithms can be categorized in two groups: 1) Parallel advection algorithms and 2) Parallel reinitialization algorithms.
+Eulerian advection schemes can easily be parallelized but unfortunately are limited by the CFL condition which could be very restrictive for adaptive grids.
+Semi-Lagrangian methods combine the unconditional stability of Lagrangian methods and the ease of use of Eulerian grids and have been successfully used for advecting the level-set function on tree-based grids before \cite{Min;Gibou:07:A-second-order-accur}.
+However, parallelizing the semi-Lagrangian algorithm in a domain decomposition context is not an easy task.
 The reason for this is twofold: First, depending on the CFL number, the departure points may end up outside the ghost region and in remote processors that are potentially far away.
 This requires a very dynamic and nonuniform communication pattern which is complicated to implement.
-For adaptive grids situation is even more complicated due to the asymmetric nature of communication, i.e.
-every processor knows the message destination but no processor knows which processor to expect message from and how many.
-Second, depending on the CFL number, the distribution of departure points could be entirely independent of the original parallel partitioning.
-As a result large CFL numbers could result in very load-unbalanced distribution of departure points which restricts the scalability of the algorithm.
-Both of these problems, of course, could be avoided by choosing $\text{CFL} \le 1$ but that would defeat the purpose of using Semi-Lagrangian algorithm in the first place.
+For adaptive grids situation is even more complicated due to the asymmetric nature of communication (c.f.
+\ref{section::parallel algorithms}).
+Second, load balancing could be an issue for large CFL numbers and nonuniform velocity fields due to clustering of departure points and can considerably restrict the scalability of the algorithm.
+Both of these problems, of course, could be avoided by choosing $\text{CFL} \le 1$ but that would defeat the purpose of using semi-Lagrangian algorithm in the first place.
 
-\item \textbf{Parallel reinitialization methods? What are the challenges? Whats out there?}
+Nonetheless, several parallel semi-Lagrangian algorithms have been proposed in the past.
+A simple domain decomposition technique was used in \cite{Thomas;Cote:95:Massively-parallel-s} where the width of ghost layer is fixed based on the maximum CFL number to ensure all departure points are covered by the ghost layer.
+At large CFL numbers, this leads to large communication volume that can limit the scalability.
+Nonetheless good scaling was report for small CFL numbers ($\text{CFL} \le 2$).
+In \cite{Drake;Foster;Michalakes;etal:95:Design-and-performan} authors propose a more sophisticated domain decomposition approach which uses a ``dynamic ghost layer''.
+Here the width of the ghost layer is dynamically determined at runtime based on information from previous time steps.
+Unfortunately, however, this approach seems to suffer from excessive communication overhead at larger number of processors as well.
+More recently, authors in \cite{White-III;Dongarra:11:High-performance-hig} used a domain decomposition strategy on a cubed sphere but with a single layer of ghost nodes.
+Interpolation on remote processors is then handled by sending query points to the corresponding processor and asking for the interpolated result.
+This approach seems to provide good scalability for transporting a single tracer up to about 1000 cores for $\text{CFL} \sim 10$.
+At higher CFL numbers the method begins to loose scalability due to an increase in communication volume.
+Although in this article we are mainly interested in parallel semi-Lagrangian methods, one could resort to finite difference or finite element discretization methods if small CFL numbers are acceptable.
+Indeed several algorithms of this type has been proposed with applications in modeling dendritic crystal growth \cite{Wang;Chang;Kale;etal:06:Parallelization-of-a}, multiphase flows \cite{Sussman:05:A-parallelized-adapt, Fortmeier;Bucker:11:A-parallel-strategy-, Rodriguez;Sahni;Lahey-Jr;etal:13:A-parallel-adaptive-}, and atomization process \cite{Herrmann:10:A-parallel-Eulerian-}.
+%, and image segmentation on GPUs \cite{Lefohn;Cates;Whitaker:03:Interactive-GPU-base,Cates;Lefohn;Whitaker:04:GIST:-an-interactive,Roberts;Packer;Sousa;etal:10:A-work-efficient-GPU}.
 
-\item \textbf{Whats the organization of the paper}
-\end{enumerate}
+In many applications of the level-set method, it is desirable for the level-set function to be a signed distance function, i.e.
+$|\nabla \phi| = 1$.
+There are generally two approaches to enforce this property: 1) Solving the pseudo-time transient reinitialization equation \cite{Sussman;Smereka;Osher:94:A-Level-Set-Approach, Osher;Fedkiw:01:Level-Set-Methods:-A}:
+\ben
+\phi_\tau + S(\phi_0)\left(|\nabla \phi| - 1\right) = 0,
+\een
+or 2) Solving the Eikonal equation:
+\ben
+F(x)|\nabla\phi| = 1,
+\een 
+with constant speed function $F(x) \equiv 1$.
+The transient reinitialization equation can be solved using explicit finite differences and thus can easily be parallelized in a domain decomposition approach.
+Moreover, only a few iterations may be needed if the signed-distance property is only required close to the interface \cite{Min;Gibou:07:A-second-order-accur}.
+This is the approach we have chosen in this paper.
+If, however, the signed-distance property is required in the entire domain solving the Eikonal equation is more computationally efficient.
+Unfortunately, however, the most popular algorithm for solving the Eikonal equation, i.e.
+the Fast Marching Method \cite{Sethian:96:A-Fast-Marching-Leve,Sethian:99:Level-set-methods-an}, is inherently sequential due to causal relationship between grid points and cannot be easily parallelized.
+The Fast Sweeping Method (FSM) \cite{Zhao:05:A-fast-sweeping-meth} is an alternative method for solving the Eikonal equation iteratively.
+The FSM can be more computationally efficient for simple choices of speed function, e.g.
+as in this context, and for simple interfaces.
+Moreover, FSM has more potential for parallelization compared to the FMM.
+
+One of the earliest attempt in parallelizing the FMM is reported in \cite{Herrmann:03:A-domain-decompositi} where a domain decomposition algorithm was introduced.
+Unlike the serial FMM, however, parallel FMM potentially requires multiple iteration or ``rollback operations'' to enforce the causality across processors.
+Similar ideas are described in details in \cite{Tugurlan:08:Fast-marching-method}.
+It should be noted that number of iterations for the parallel FMM to converge greatly depends on the complexity of the interface and also the parallel partitioning and, in general, fewer iterations are required if domains are aligned with the normals to the interface.
+Due to the nature of Eikonal equation, shared memory machines might be a better environment for parallelization.
+For instance, in \cite{Breus;Cristiani;Gwosdek;etal:11:An-adaptive-domain-d} authors use an ``adaptive'' technique where individual threads implicitly define a domain decomposition at runtime.
+Unfortunately, however, this approach does not seem to be more effective than a simple static decomposition.
+In \cite{Zhao:07:Parallel-implementat} a parallel FSM method was presented for the fist time.
+However, a more scalable FSM was more recently proposed in \cite{Detrixhe;Gibou;Min:13:A-parallel-fast-swee} where the Cuthill-McKee numbering was utilize to expose more parallelism.
+More recently a two-scale, hybrid FMM-FSM was presented in \cite{Chacon;Vladimirsky:13:A-parallel-Heap-Cell} which, albeit being more complicated to implement, promises even better scalability.
+Finally, a parallel Fast Iterative Method (FIM) was proposed in \cite{Jeong;Whitaker:08:A-fast-iterative-met}.
+The FIM is similar to FMM in that it also maintains a list of ``active nodes''.
+Unlike FMM, however, FIM avoids sorting the list and allows for concurrent updating of all nodes in an iterative fashion.
+
+The rest of this article is organized as follows: In section \ref{section::levelset method} we briefly review the sequential algorithms and discretization methods for the level-set equation on adaptive tree-based grids.
+These ideas are then extended in section \ref{section::parallel algorithms} to parallel using a domain decomposition method.
+In section \ref{section::scaling results} we provide several examples that illustrate the scalability of our algorithms.
+Finally, we close by providing an important application of our algorithms in modeling the solidification process by solving a Stefan problem in \ref{section::applications}.

### 2014-11-12 paper/tex/introduction.tex
Finished with the algorithm section of paper

@@
 %!TEX root = draft.tex
-\section{Introduction}
+\section{Introduction}\label{sec:introduction}
 The level-set method, originally proposed by Sethian and Osher \cite{Osher;Sethian:88:Fronts-Propagating-w}, is a popular and powerful framework for tracking arbitrary interfaces that undergo complicated topological changes.
@@
 For adaptive grids situation is even more complicated due to the asymmetric nature of communication (c.f.
-\ref{section::parallel algorithms}).
+\ref{sec:parallel algorithms}).
 Second, load balancing could be an issue for large CFL numbers and nonuniform velocity fields due to clustering of departure points and can considerably restrict the scalability of the algorithm.
@@
 
-The rest of this article is organized as follows: In section \ref{section::levelset method} we briefly review the sequential algorithms and discretization methods for the level-set equation on adaptive tree-based grids.
-These ideas are then extended in section \ref{section::parallel algorithms} to parallel using a domain decomposition method.
+The rest of this article is organized as follows: In section \ref{sec:levelset method} we briefly review the sequential algorithms and discretization methods for the level-set equation on adaptive tree-based grids.
+These ideas are then extended in section \ref{sec:parallel algorithms} to parallel using a domain decomposition method.
 In section \ref{section::scaling results} we provide several examples that illustrate the scalability of our algorithms.

### 2014-11-12 paper/tex/levelset_method.tex
Finished with the algorithm section of paper

--- 2014-08-27 paper/tex/levelset_method.tex
@@
 %!TEX root = draft.tex
-\section{The level-set method}
+\section{The level-set method}\label{sec:levelset method}
+The level-set method, introduced in \cite{Osher;Sethian:88:Fronts-Propagating-w}, is an implicit framework for tracking interfaces that undergo complicated topological changes.
+In this framework an interface is represented by the zero contour of a higher dimensional function, e.g.
+a curve in two spatial dimension can be described as $\Gamma = \{(x,y) | \phi(x,y) = 0\}$ where $\phi(x,y)$ is the level-set function.
+Evolution of the curve under a velocity field $\mathbf{V}$ is then obtained by solving the level-set equation:
+\be
+\phi_t + \mathbf{V} \cdot \nabla \phi = 0.
+\label{eq:ls}
+\ee
+When velocity field does not depend on the level set function itself, equation \eqref{eq:ls} can be solved using the Semi-Lagrangian method.
+An important advantage of the Semi-Lagrangian method over regular finite difference method is its unconditional stability which allows for arbitrarily large time steps.
+This is specially of interest when using adaptive grids that allow for higher grid resolutions.
 
-\indent The level-set method is a method to capture interfaces implicitely by constructing a function $\phi$, called the level-set function, that is negative on on side of the domain $\Omega^-$ and positive on the other side $\Omega^+$, thus defining the interface $\Gamma$ as the zero contour of $\phi$, $\Gamma = \{\mathbf{x} \vert \phi(\mathbf{x}=0\}$.\\
+In general, an infinite number of level-set functions can describe the same zero contour and thus interface.
+However, it is desirable to chose functions with signed distance property, i.e.
+$|\nabla \phi| = 1$.
+As detailed in section \ref{sec:introduction}, we solve the pseudo-time transient reinitialization equation \cite{Sussman;Smereka;Osher:94:A-Level-Set-Approach, Osher;Fedkiw:01:Level-Set-Methods:-A} to achieve this property,
+\be
+\phi_\tau + S(\phi_0)\left(|\nabla \phi| - 1\right) = 0,
+\label{eq:reinitialization}
+\ee
+where $\phi_0$ is any level-set function that correctly describes the interface location and $S(\phi_0)$ is an appropriate approximation of the sign function \cite{Osher;Fedkiw:02:Level-Set-Methods-an}.
+Here, we do not go into the details of the sequential algorithms for solving equations \eqref{eq:ls} and \eqref{eq:reinitialization}.
+Instead, we note that the algorithms presented in section \ref{sec:parallel algorithms} are based on the sequential methods presented in \cite{Min;Gibou:07:A-second-order-accur} and refer the interested reader to the aforementioned article and references therein for more details.
 
-If infinitely many functions satisfy these requirements, the most natural and most convenient representation is a signed distance function to the interface.
+% \indent The level-set method is a method to capture interfaces implicitely by constructing a function $\phi$, called the level-set function, that is negative on on side of the domain $\Omega^-$ and positive on the other side $\Omega^+$, thus defining the interface $\Gamma$ as the zero contour of $\phi$, $\Gamma = \{\mathbf{x} \vert \phi(\mathbf{x}=0\}$.\\
+
+% If infinitely many functions satisfy these requirements, the most natural and most convenient representation is a signed distance function to the interface.
 The level-set function is therefore reinitialized to a signed distance function at every time step by solving the reinitialization equation
 
-\begin{equation} \label{eq::reinitialization}
-\pd{\phi}{\tau} + \mathrm{sign}(\phi) \left( \lvert \phi \rvert - 1 \right) = 0,
-\end{equation}
+% \begin{equation} \label{eq::reinitialization}
+% \pd{\phi}{\tau} + \mathrm{sign}(\phi) \left( \lvert \phi \rvert - 1 \right) = 0,
+% \end{equation}
 
-where $\tau$ is a fictitious time.
+% where $\tau$ is a fictitious time.
 The discretization and implementation of this procedure will be discussed later in section \ref{section::reinitialization}.
@@
 
-\begin{equation*}
-\mathbf{n} = \frac{\nabla \phi}{\lvert \nabla \phi \rvert},
-\end{equation*}
+% \begin{equation*}
+% \mathbf{n} = \frac{\nabla \phi}{\lvert \nabla \phi \rvert},
+% \end{equation*}
 
-where it should be noted that $\lvert \nabla \phi \rvert = 1$ when the level-set function is reinitialized, and
+% where it should be noted that $\lvert \nabla \phi \rvert = 1$ when the level-set function is reinitialized, and
 
-\begin{equation*}
-\kappa = \nabla \cdot \mathbf{n} =  \nabla \cdot \left( \frac{\nabla \phi}{\lvert \nabla \phi \rvert} \right).
-\end{equation*}
+% \begin{equation*}
+% \kappa = \nabla \cdot \mathbf{n} =  \nabla \cdot \left( \frac{\nabla \phi}{\lvert \nabla \phi \rvert} \right).
+% \end{equation*}
 
-\section{The \texttt{p4est} library}
+% \section{The \texttt{p4est} library}
 
-The work we present relies on the parallel octree library \texttt{p4est} \cite{burstedded:2011:p4est}.
+% The work we present relies on the parallel octree library \texttt{p4est} \cite{Burstedde;Wilcox;Ghattas:11:p4est:-Scalable-Algo}.
 It is scalable implementation of the general octree structure in a massively parallel mpi environment.
-We limit our presentation of \texttt{p4est} to the relevant information for our application, for more details the reader is invited to read the original article \cite{burstedded:2011:p4est}.
+We limit our presentation of \texttt{p4est} to the relevant information for our application, for more details the reader is invited to read the original article \cite{Burstedde;Wilcox;Ghattas:11:p4est:-Scalable-Algo}.
 
-The \texttt{p4est} library provides the geometrical information for the leaves of an octree which includes the neighboring information for the leaves (called quadrants in the original article) of the tree, a layer of ghost quadrants and ghost vertices for each process, the coarsening and refining procedures, and encapsulates the communication between processes.
+% The \texttt{p4est} library provides the geometrical information for the leaves of an octree which includes the neighboring information for the leaves (called quadrants in the original article) of the tree, a layer of ghost quadrants and ghost vertices for each process, the coarsening and refining procedures, and encapsulates the communication between processes.
 However, it does not provide the vertical structure of the octrees.
@@
 
-In addition to the owned and ghost categories, we choose to group the vertices in \textit{local} and \textit{layer} sets.
+% In addition to the owned and ghost categories, we choose to group the vertices in \textit{local} and \textit{layer} sets.
 We define the layer vertices as the vertices that are part of another process's ghost layer and the local vertices as the vertices that are not part of any other process's ghost layer.
@@
 
-\begin{figure}[ht!]
-\begin{center}
-\includegraphics[width=.4\textwidth]{pictures/layer_nodes.pdf}
-\caption{Illustration of a possible layout at the interface between two processes.
+% \begin{figure}[ht!]
+% \begin{center}
+% \includegraphics[width=.4\textwidth]{pictures/layer_nodes.pdf}
+% \caption{Illustration of a possible layout at the interface between two processes.
 The red quarants and the orange and purples vertices belong to process 1 while the green quadrants and the green and black vertices belong to process 2.
 The green G denotes the ghost quadrants for process 2, the green dots are the ghost vertices for process 1, and the purple dots correspond to layer vertices for process 1 (and therefore ghost vertices for process 2).} \label{fig::local_layer_vertices}
-\end{center}
-\end{figure}
+% \end{center}
+% \end{figure}

### 2014-11-12 paper/tex/parallel_algorithms.tex
Finished with the algorithm section of paper

--- 2014-08-27 paper/tex/parallel_algorithms.tex
@@
 %!TEX root = draft.tex
-\section{Presentation of the algorithms}
+\section{Parallel algorithms}\label{sec:parallel algorithms}
+\subsection{Grid management}
+Adaptive tree-based grids can significantly reduce the computational cost of level-set methods by restricting the fine grid close to the interface where it is most needed \cite{Strain:99:Tree-Methods-for-Mov}.
+Moreover, adaptive tree-based grids are easy to generate in the presence of a signed-distance level-set function \cite{Min;Gibou:07:A-second-order-accur} and can efficiently be encoded using a tree datastructure \cite{Samet:90:Applications-of-Spat}.
+To develope scalable parallel algorithms on these grids, it is necessary to parallelize the datastructure and grid manipulations method such as refinement and coarsening of cells as well as provide with a fast method for grid partitioning and load balancing.
+\texttt{p4est} library \cite{p4est-github} is a collection of such parallel algorithms that has recently emerged and shown to scale up to 200,000 processors \cite{Burstedde;Wilcox;Ghattas:11:p4est:-Scalable-Algo}.
 
-We now propose a detailed description of the main algorithms for the level-set method.
+In \texttt{p4est} the adaptive grid is represented as a non-overlapping collection of trees that are rooted in individual cells of a common coarse grid (c.f.
+figure \ref{fig:p4est_zcurve}).
+This common coarse grid, which we will refer to as the ``macromesh'', can in general be an unstructured quadrilateral, in two spatial dimensions, or hexahedral mesh, in three spatial dimensions.
+In this article, however, we shall limit ourselves to simple uniform Cartesian macromeshes.
+Moreover it is implicitly assumed that the macromesh is small enough that can be entirely replicated on all processors.
+For instance, in many of the applications that we are interested in this paper the macromesh simply a single cell.
+\texttt{p4est} allows for arbitrary refinement and coarsening criteria through defining callback functions.
+In this article the refinement criteria is chosen based on the distance of individual cells to the interface.
+Specifically a cell $C$ is marked for refinement if:
+\be
+\min_{\forall v \in V(C)} |\phi| \le \frac{L D}{2}.
+\label{eq:refine}
+\ee
+Conversely an existing cell is marked for coarsening if:
+\be
+\min_{\forall v \in V(C)} |\phi| > L D.
+\label{eq:coarsen}
+\ee
+Here $V(C)$ denotes the set of all vertices of cell $C$, $L$ denotes the Lipschitz constant of the level-set function, and $D$ denotes the diagonal size of cell $C$.
+We refer to section 3.2 of \cite{Burstedde;Wilcox;Ghattas:11:p4est:-Scalable-Algo} for details on parallel refinement and coarsening algorithms used in \texttt{p4est}.
 
-\subsection{Interpolations}
+Once the gird is adapted to the interface, it must be partitioned to ensure load balancing across processors.
+This is achieve by constructing a Z-curve that traverses the leaves of all trees in order of tree index (c.f.
+figure \ref{fig:p4est_zcurve}).
+A Z-curve is a Space Filling Curve (SFC) with the important property that cells with close Z-indecies are also geometrically close in the physical domain.
+This is beneficial since it both leads to reduction in MPI communication and improves the cache performance of several algorithms such as interpolation and finite difference calculations.
+For more details on parallel partitioning in \texttt{p4est} one may refer to section 3.3 of \cite{Burstedde;Wilcox;Ghattas:11:p4est:-Scalable-Algo}.
+\begin{figure}[hbtp]
+\begin{center}
+\includegraphics[width = \columnwidth]{figures/p4est_zcurve.pdf}
+\caption{Left: A ``forest'' made up of two trees $k_0$ and $k_1$.
+Parallel partitioning is achieved by first constructing a Z-curve which starts at cell ``S'' and ends at cell ``E''.
+Next this tree is split up among processors either uniformly or by assigning different weights to cells.
+Here processors are represented via different colors.
+Note how use of Z-curve naturally leads to clustering of most cells in each processor domain.
+Right: Schematic of a tree data structure representing the forest and its partitioning.}
+\label{fig:p4est_zcurve}
+\end{center}
+\end{figure}
+Aside from grid manipulation and partitioning, we use two additional features of \texttt{p4est}, namely generation of ghost layer cells and creating a globally unique node indexing.
+These algorithms are detailed in sections 3.5 and 3.6 of \cite{Burstedde;Wilcox;Ghattas:11:p4est:-Scalable-Algo}.
+An important feature of our algorithms is that they are designed for non-graded trees.
+This is important because we can entirely skip tree balancing which was shown to be one of most time consuming parts of grid adaptation in \texttt{p4est} \cite{Burstedde;Wilcox;Ghattas:11:p4est:-Scalable-Algo}.
 
-The first important step to build a level-set based code is to provide an interpolation procedure on the vertices.
-Being able to evaluate quantities at arbitrary locations in the computational domain is mandatory for procedures such as advecting the level-set function or enforcing boundary conditions on the irregular interface.
-The interpolation procedure enables any process to evaluate a quantity anywhere in the domain.
+Finally, in \texttt{p4est} trees are linearized, i.e.
+only the leaves are explicitly stored.
+However, all of algorithms considered here require explicit knowledge of the entire tree.
+As a result we introduce a simple algorithm, called \texttt{Reconstruct}, which recreates a local representation of the entire ``forest'' that is only adapted to local cells and, potentially, the ghost layer.
+This approach is similar to the ideas introduced in \cite{Bangerth;Burstedde;Heister;etal:11:Algorithms-and-data-} and our tests show that in a typical application they usually amount to less that $1\%$ of the entire runtime.
+Algorithm \ref{alg:reconstruction} illustrates how this reconstruction is performed.
+Given a forest and a layer of ghost cells from \texttt{p4est}, the algorithm generates a local representation of forest by recursively refining from the root until reaching the same level and location of all leaves in \texttt{p4est} and \texttt{ghost}.
+Note that algorithm \ref{alg:reconstruction} does not involve any communication and is load balanced.
+Figure \ref{fig:reconstruction} illustrates an application of algorithm \ref{alg:reconstruction}.
+Note how each processor has independently generated a local representation of the forest that is refined to match the same leaves as in the global forest and ghost layer.
+% , and has an average runtime complexity of $\mathcal{O}\left((n_l + n_g)\log_{2^d}N\right)$ for well-balanced trees in $d$ dimensions.
+Here, $n_l$ and $n_g$ are the number of local and ghost cells, respectively, and $N$ is the global number of cells.
+\begin{algorithm}[htbp]
+\caption{$H \gets \texttt{Reconstruct (}G\texttt{)}$}\label{euclid}
+\begin{algorithmic}[1]
+\State $H \gets G.\texttt{macromesh()}$
+\For {$tr : G.\texttt{local\_trees()}$} \Comment{traverse local cells}
+	\For {$q : tr.\texttt{cells()}$}
+		\State $H.\texttt{update\_tree(} tr, q \texttt{)}$
+	\EndFor
+\EndFor
 
-\begin{algorithm}[ht!]
-\begin{algorithmic}
-\If {interpolation order $>$ 2}
-	\State compute second order derivatives for local and ghost points if needed
+\For {$q : G.\texttt{ghost\_cells()}$} \Comment{traverse ghost cells}
+	\State $H.\texttt{update\_tree(} q.\texttt{tree()}, q \texttt{)}$
+\EndFor
+
+\State \Return $H$
+\\
+\Function{$H.\texttt{update\_tree}$}{$tr, q$}   \Comment{recursive reconstruction}
+	\State $q_l \gets H.\texttt{root(}tr\texttt{)}$
+	\While{$q_l.\texttt{level()} \not= q.\texttt{level()} $}
+		\If {$q_l.$\texttt{is\_leaf()}} $q_l$\texttt{.split()}
 		\EndIf
-\State begin send point buffer
-\State do the interpolation for local points
-\State do the interpolation for ghost points
-\State begin receiving point buffer
-\For {in $\in$ \{remote senders\} }
-	\State finish receiving from send i
-	\State interpolate all points from sender i
-\EndFor
-\State receive the point buffer
+		\State $h \gets q_l.\texttt{length()} / 2$ \Comment{select the next child based on direction}
+		\State $i \gets q.x \ge q_l.x + h$
+		\State $j \gets q.y \ge q_l.y + h$
+		\State $k \gets q.z \ge q_l.z + h$
+
+		\State $q_l \gets q_l.\texttt{child(} i, j, k \texttt{)}$
+	\EndWhile	
+\EndFunction
 \end{algorithmic}
+\label{alg:reconstruction}
 \end{algorithm}
 
-\subsection{Semi lagrangian advection}
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[width = \textwidth]{figures/reconstruct.pdf}
+\end{center}
+\caption{Left: A forest refined close to an interface and partitioned among four processors as indicated by colors.
+Right: Each processor independently recreates a local forest which is refined to match the global grid and is as coarse as possible elsewhere.
+Note that empty cells are fictitious, i.e.
+they are only required to generate the hierarchal structure and are not matched by any corresponding cell in the global forest.}
+\label{fig:reconstruction}
+\end{figure}
 
-\subsection{Reinitialization and extensions} \label{section::reinitialization}
+\subsection{Semi-Lagrangian method}
+As indicated earlier, we use the Semi-Lagrangian method to solve equation \eqref{eq:ls} when the velocity field is externally generated, i.e.
+does not depend on level-set function itself.
+Let us rewrite equation \eqref{eq:ls} along the characteristic curve $\mathbf{X}(t)$ as:
+\be
+\left\{
+\begin{array}{rcl}
+\dfrac{\text{d} \mathbf{X}}{\text{d} t} &=& \mathbf{V}, \\ [3ex]
+\dfrac{\text{d} \phi(\mathbf{X}(t), t)}{\text{d} t} &=& 0.
+\end{array}
+\right.
+\label{eq:sl}
+\ee
+The Semi-Lagrangian method integrates equations \eqref{eq:sl} backward in time, i.e.
+starting from the grid $G^{n+1}$ \footnote{$G^{n+1}$ itself is computed in an iterative fashion as explained later on.}, we simply write $\phi^{n+1}(\mathbf{X}^{n+1}) = \phi(\mathbf{X}(t^{n+1}), t^{n+1}) = \phi(\mathbf{X}(t^n), t^n) = \phi^n(\mathbf{X}_d)$.
+Here the characteristic curves are chosen such that $\mathbf{X}(t^{n+1})$ are simply the coordinates of grids points of $G^{n+1}$.
+Moreover $\mathbf{X}_d$ are the departure points which are computed using the second-order midpoint method \cite{Min;Gibou:07:A-second-order-accur}:
+\bea
+\mathbf{X}^\star &=& \mathbf{X}^{n+1} - \frac{\Delta t}{2} \mathbf{V}^{n}(\mathbf{X}^n),	   \label{eq:xstar}      \\
+\mathbf{X}_d     &=& \mathbf{X}^{n+1} - \Delta t \mathbf{V}^{n+\frac{1}{2}}(\mathbf{X}^\star), \label{eq:xdeparture}
+\eea
+where $\mathbf{V}^{n+\frac{1}{2}}$ is obtained via extrapolation from previous times, i.e.:
+\be
+\mathbf{V}^{n+\frac{1}{2}} = \frac{3}{2} \mathbf{V}^n - \frac{1}{2}\mathbf{V}^{n-1}. \label{eq:vn_p_half}
+\ee
 
-The reinitialization procedure transforms an arbitrary function into a signed distance function to its zero contour.
-One of the popular methods to achieve this results is to solve the reinitialization equation \ref{eq::reinitialization} \cite{Sussman:1994:levelsetincompressible, Min_Gibou:2007:Level_Set_Adaptive}.
-Other methods, such as the Fast Sweeping \cite{??} and the Fast Marching methods \cite{Sethian:1996:FMM, chopp:2001:fmm}, are based on the wave propagation nature of the reinitialization front and provide faster implementations of the reinitialization procedure in a serial environment.
-However, their parallelization is an involved process and we opt for solving the reinitialization equation.
+Note that all values at the intermediate point, $\mathbf{X}^\star$, and departure point, $\mathbf{X}_d$, must be calculated via interpolation from previous grids, $G^{n}$ and $G^{n-1}$.
+Here we use the stabilized second-order interpolation for $\phi(\mathbf{X}_d)$ and the multi-linear interpolation for $\mathbf{V}^{n+\frac{1}{2}}(\mathbf{X}^\star)$ \cite{Min;Gibou:07:A-second-order-accur}.
+Although parallelization of the interpolation process on a shared-memory machine is trivial, the same cannot be said for distributed-memory machines.
+In fact, the parallel interpolation algorithm \ref{alg:interpolation} is probably the most important contribution of this article.
+Complication arises because not all calculated departure points will reside in the domain of current processor.
+Moreover, due to the irregular shape of partitions, we cannot be certain that they are entirely owned by neighboring processors.
+At best we can only expect that their location are bounded by a halo of width $w \le \text{CFL} \: \Delta x_{\min}$ around the local partition.
+Of course if one enforces $\text{CFL} \le 1$, one can ensure that the halo is bounded by the ghost layer which significantly simplifies the communication problem.
+This assumption, however, defeats the purpose of using Semi-Lagrangian in the first place since we are interested in large values of $\text{CFL}$ number.
 
-We base our approach on Min and Gibou \cite{Min_Gibou:2007:Level_Set_Adaptive} and discretize the reinitialization equation as
+One remedy to this problem, proposed in \cite{Thomas;Cote:95:Massively-parallel-s} for uniform grids, is to increase the size of ghost layer to $\lceil \text{CFL} \rceil$.
+For large values of $\text{CFL}$ number, however, this approach can substantially increase the communication volume.
+Moreover, this simple approach does not work in the process of generating $G^{n+1}$ due to repartitioning.
+An alternative approach would be to handle local and remote interpolations separately.
+Our remote interpolation algorithm is composed of three separate phases.
+In the first phase, which we call buffering, every processor searches for all departure points inside the local tree.
+If the point is owned by a local cell, it is added to a local buffer, otherwise we find the processor which owns the point and add the point to a separate buffer belonging to the found rank.
+Note that searching the point in the local tree is performed recursively, similar to the \texttt{UpdateLocalTree} function in algorithm \ref{alg:reconstruction}.
+The owner's rank is found by computing the Z-index of the point and then using binary search on the Z-curve.
+This is already implemented in \texttt{p4est} and explained in details in section 2.5 of \cite{Burstedde;Wilcox;Ghattas:11:p4est:-Scalable-Algo}.
 
-\begin{equation*}
-\frac{d\phi}{d\tau} + \mathrm{sign}(\phi^0)[H_G(D^+_x\phi, D^-_x\phi, D^+_y\phi, D^-_y\phi)-1] = 0,
-\end{equation*}
+Once buffering is done, every processor knows exactly how many messages it needs to send and to which processors.
+This also implicitly defines processors that will later on send a reply message to this processor.
+However, at this point no one knows which processors they should expect a message from.
+We solve this problem using a simple \textit{communication matrix} (see figure \ref{fig:communication}).
+Other approaches for solving this problem could include using non-blocking collective or one-sided communication operations introduced in the MPI-3 standard \cite{MPI_ref, sparse dynamic exchange protocol paper}.
+Although these algorithms have better theoretical communication complexities, we did not see any difference in the runtime or scalability of algorithm \ref{alg:interpolation} when using them.
 
-with $\mathrm{sign}(\phi^0)$ the signum of the initial level-set function and $H_G$ the Godunov Hamiltonian:
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[width = \textwidth] {figures/communication.pdf}
+\end{center}
+\caption{Left: Location of back-traced points depends on the magnitude of local velocity and time-step.
+Although departure distance is bounded by $\text{CFL} \: \Delta x_{\min}$, one cannot make any special assumption about receiving rank without explicit searching of the entire Z-curve.
+Moreover, the receiving processor has no prior knowledge about which processors to check for incoming messages nor does it know anything about the possible message length (i.e.
+number of points).
+Middle: A directed graph illustrating the communication pattern among processors with arrows representing the direction that messages are sent.
+Right: The adjacency matrix of communication graph.
+For each row, sum of all columns represents the number of sent messages.
+Conversely, for each column, sum of all rows represents number of messages that need to be received.
+As detailed in algorithm \ref{alg:interpolation} this information is enough to build a parallel interpolation scheme.}
+\label{fig:communication}
+\end{figure}
 
-\begin{equation*}
-H_G(a,b,c,d) =
-\begin{cases}
-\sqrt{\max(\lvert a^+ \rvert^2, \lvert b^- \rvert^2) + \max(\lvert c^+ \rvert^2, \lvert d^- \rvert^2)} \quad \mathrm{if~sgn}(\phi^0) \leq 0\\
-\sqrt{\max(\lvert a^- \rvert^2, \lvert b^+ \rvert^2) + \max(\lvert c^- \rvert^2, \lvert d^+ \rvert^2)} \quad \mathrm{if~sgn}(\phi^0) > 0
-\end{cases}
-\end{equation*}
+To solve the communication problem, we first compute the adjacency matrix of the communication pattern, i.e. we construct the matrix $A_{P \times P}$ such that:
+\ben
+a_{ij} = 
+\left\{
+\begin{array}{lr}
+1 \hspace{5 mm} \text{if processor `\textit{i}' sends a message to processor `\textit{j}',} \\
+0 \hspace{5 mm} \text{otherwise.}
+\end{array}
+\right.
+\een
+Note that this matrix is also distributed among processors, i.e. each row is owned by a separate rank. Next, we compute:
+\ben
+S_i = \sum_j a_{ij} \hspace{5 mm} \text {and} \hspace{5 mm} R_i = \sum_j a_{ji}
+\een
+where $S_i$ and $R_i$ denote the number of sent and received messages respectively.
+While $S_i$ can be computed trivially, a reduction operation is required for computing $R_i$.
+For instance, this can be achieved using a single \texttt{MPI\_Reduce\_scatter} function call.
+The last phase of interpolation involves overlapping the computation of interpolated values for local points with the communication of data between processors.
+This is done by alternating between local calculations and probing for incoming messages from other processors.
+Interpolation is finished once values for all local points have been calculated and all remote requests have been processed (see algorithm \ref{alg:interpolation}).
+\begin{algorithm}[htbp]
+\caption{$values \gets \texttt{Interpolate (}H, \mathbf{X}\texttt{)}$}\label{euclid}
+\begin{algorithmic}[1]
+\State $col \gets 0, buff \gets null$ \Comment{Phase I -- buffering}
+\For {$\mathbf{p} : \mathbf{X}$} 
+	\State $[r, cell] \gets H.\texttt{search(}\mathbf{p}\texttt{)}$ \Comment{search for the owner's rank and cell}
+	\If {$r = mpirank$}
+		\State $buff[r].\texttt{push\_back(}\mathbf{p}, cell\texttt{)}$
+	\Else
+		\State $buff[r].\texttt{push\_back(}\mathbf{p}\texttt{)}$
+		\State $col[r] \gets 1$
+	\EndIf
+\EndFor
+\For {$r: mpisize$} \Comment{Phase II -- initiate communication and compute number of messages}
+	\If {$col[r]$}
+		\State $\texttt{MP\_Isend(}r, buff[r]\texttt{)}$
+	\EndIf
+\EndFor
+\State  $S \gets \texttt{sum(}col\texttt{)}$ 
+\State  $R \gets \texttt{MPI\_Reduce\_scatter(}col,\texttt{MPI\_SUM)}$
+\State $done \gets false, it \gets buff[mpirank].\texttt{begin()}$
+\While {$!done$} \Comment{Phase III -- main loop}
+	\If {$it \not= buff[mpirank].\texttt{end()}$}
+		\State $values \gets \texttt{process\_local(}it\texttt{)}$ \Comment{process local interpolations}
+		\State $\texttt{++}it$
+	\EndIf
+	\If {$R > 0$} \Comment{process queries sent from remote processors}
+		\State $[msg, st] \gets \texttt{MPI\_Iprobe()}$
+		\If {$msg$}
+			% \State $values \gets \texttt{MPI\_Recv(}st.\texttt{MPI\_SOURCE)}$ \Comment{}
+			\State $val\_buff \gets \texttt{process\_queries(}st\texttt{)}$ \Comment{receive, search, and interpolate values}
+			\State $\texttt{MPI\_Isend(}st.\texttt{MPI\_SOURCE},val\_buff\texttt{)}$ \Comment{send back interpolated values}
+			\State $R\texttt{--}$
+		\EndIf
+	\EndIf
+	\If {$S > 0$} \Comment{process replies sent to our queries}
+		\State $[msg, st] \gets \texttt{MPI\_Iprobe()}$
+		\If {$msg$}
+			\State $values \gets \texttt{process\_replies(}st.\texttt{MPI\_SOURCE)}$ \Comment{receive remotely interpolated values}
+			\State $S\texttt{--}$
+		\EndIf
+	\EndIf
+	\State $done \gets S = 0 \: \And \: R = 0 \: \And \: it = buff[mpirank].\texttt{end()}$
+\EndWhile
+\State \Return $values$
+\end{algorithmic}
+\label{alg:interpolation}
+\end{algorithm}
 
-with $a^+=\max(a,0)$ and $a^-=\min(a,0)$.
-The one-sided derivatives $D^{+/-}_{x/y}\phi$ are discretized by second order accurate finite differences as explained by Min and Gibou in \cite{Min_Gibou:2007:Level_Set_Adaptive}.
-Following the presentation in \cite{Min_Gibou:2007:Level_Set_Adaptive} which incorporates the subcell resolution introduced by Russo and Smereka \cite{Russo_Smereka}, we include the initial position of the interface, given by $\phi^0$, in the computation of the one sided derivatives next to the interface in order to preserve the interface location.
-Finally, we choose a simple Euler step for the time discretization.
+Using the interpolation algorithm \ref{alg:interpolation}, we close this section by presenting the final Semi-Lagrangian algorithm \ref{alg:semi-lagrangian}.
+The basic idea here is to start from an initial guess $G^{n+1}_0$ and modify the grid using refinement and coarsening criteria in \eqref{eq:refine} and \eqref{eq:coarsen} until convergence is obtained.
+Various options are available for $G^{n+1}_0$.
+For instance it is possible to start from the macromesh and only perform refinement steps until convergence.
+This choice, however, is not suitable since first few iterations do not contain many cells and there is little work for parallelism.
+Here we simply take the previous grid as the starting point, i.e.
+$G^{n+1}_0 = G^n$.
+Note that this iterative process is essentially unavoidable since the grid is based on the value of the level-set function at $t^{n+1}$, which itself is unknown and is to be defined on $G^{n+1}$.
+Nonetheless the process converges to the final grid in at most $l_{\max}$ steps where $l_{\max}$ denotes the maximum final depth of trees across all processors.
+\begin{algorithm}[htbp]
+\caption{$[G^{n+1}, \phi^{n+1}] \gets \texttt{SemiLagrangian (}G^n, \phi^n, \mathbf{V}^n, \mathbf{V}^{n-1}, \text{CFL}\texttt{)}$}\label{euclid}
+\begin{algorithmic}[1]
+\State $\Delta t_l \gets \text{CFL} \times \max \{\mathbf{V}^n\} / G^n.\texttt{hmin()}$
+\State $\Delta t   \gets \texttt{MPI\_Allreduce(}\Delta t_l,\texttt{MPI\_MIN)}$
+\State $H^n \gets \texttt{Reconstruct(} G^n\texttt{)}$ 
+\State $G^{n+1}_0 \gets G^n$
+\While {$true$}
+	\State $\mathbf{X}_d  \gets \texttt{ComputeDeparturePoints(}G^{n+1}_0, \mathbf{V}^n, \mathbf{V}^{n-1}, \Delta t\texttt{)}$ \Comment{using equations \ref{eq:xstar} -- \ref{eq:vn_p_half}}
+	\State $\phi^{n+1} \gets \texttt{Interpolate(}H^n, \textbf{X}_d\texttt{)}$
+	\State $G^{n+1} \gets G^{n+1}_0.\texttt{refine\_and\_coarsen(}\phi^{n+1}\texttt{)}$ \Comment{using equations \ref{eq:refine} and \ref{eq:coarsen} as criteria}
+	\If {$G^{n+1} \not= G^{n+1}_0$}
+		\State $G^{n+1}.\texttt{partition()}$
+		\State $G^{n+1}_0 \gets G^{n+1}$
+	\Else
+		\State \textbf{break}
+	\EndIf
+\EndWhile
 
-The general parallel algorithm for the reinitialization procedure is presented in \ref{algo::reinitialization}.
-For each fictitious time step, we start by computing the new value of the level-set function for the layer nodes.
-We then initiate the communication to update the ghost values while computing the new values for local nodes.
-After finishing the communication, we can then update $\phi^n$ to $\phi^{n+1}$ and move on to the next time step.
-The number of iterations needed for the reinitialization procedure to converge on the whole domain or on a band surrounding the interface depends on the size of the mesh, the geometry of the interface and how far from a signed distance function $\phi^0$ is.
-In practice, twenty iterations provide reasonable results for the applications we present.
+\State \Return $[G^{n+1}, \phi^{n+1}]$
+\end{algorithmic}
+\label{alg:semi-lagrangian}
+\end{algorithm}
 
-\begin{algorithm}[ht!]
-\begin{algorithmic}
-\State Compute initial derivatives of $\phi$ for subcell resolution
-\For {number of iterations}
-	\State Compute the derivatives of $\phi$
-	\State Solve reinitialization of layer nodes: $\phi_{layer} \leftarrow \phi_{layer}^{n+1}$
-	\State Begin update for layer nodes
-	\State Solve reinitialization on local nodes : $\phi_{local} \leftarrow \phi_{local}^{n+1}$
-	\State End update for layer nodes
-	\State Update $\phi$: $\phi \leftarrow \phi^{n+1}$
+\subsection{Reinitialization}
+Successive application of algorithm \ref{alg:semi-lagrangian}, especially for large values of CFL number, potentially lead degrades the signed distance property of level-set function.
+It is thus important to reinitialize the level-set every few iterations, especially because the quality of generated grid heavily depends on the signed distance property.
+To achieve this property we solve the pseudo-time transient equation \eqref{eq:reinitialization} using the discretization scheme detailed in \cite{Min;Gibou:07:A-second-order-accur}.
+For completeness, we briefly review the scheme here.
+First equation \eqref{eq:reinitialization} is written in the following semi-discrete form:
+\be
+\frac{\text{d} \phi}{\text{d} \tau} + S(\phi_0) \left(\mathcal{H}_G(D^+_i \phi, D^-_i \phi) - 1 \right) = 0,
+\label{eq:semidiscrete_reinit}
+\ee
+where $D^+_i \phi$ and $D^-_i \phi$ are the forward and backward derivatives in the $x_i$ direction and $\mathcal{H}_G$ is the Godunov Hamiltonian defined as:\
+\ben
+\mathcal{H}_G(a_i, b_i) = 
+\left\{
+\begin{array}{lcr}
+	\sqrt{\sum_i \max\left(|a^+_i|^2, |b^-_i|^2\right)} & \hspace {5 mm} \text{if} & S(\phi_0) \le 0 \\
+	\\
+	\sqrt{\sum_i \max\left(|a^-_i|^2, |b^+_i|^2\right)} & \hspace {5 mm} \text{if} & S(\phi_0)  >  0 
+\end{array}
+\right.
+,
+\een
+where $a^+ = \max(a, 0)$ and $a^- = \min(a, 0)$.
+Similar to \cite{Min;Gibou:07:A-second-order-accur}, equation \eqref{eq:semidiscrete_reinit} is integrated in time using the TVD-RK2 scheme using adaptive time-stepping to accelerate convergence to steady state.
+Since all computation is grid based, parallel implementation of this scheme is mostly trivial.
+However, one minor point requires further explanation.
+As suggested in \cite{Min;Gibou:07:A-second-order-accur}, one-sided derivatives $D^+_i \phi$ and $D^-_i \phi$ are computed using second order discretizations which require computation of second derivatives.
+To enable overlap between computation and communications when computing second derivatives and also integrating \eqref{eq:semidiscrete_reinit}, we use the following common technique.
+First, we label all local points, $L_p$, as either private, $P_p$, or boundary, $B_p$.
+Here, boundary points are the collection of all local points that are regarded as ghost point, $G_p$, on all other processors, i.e.
+$B_p = \underset{r,\;r\neq p}{\bigcup} G_r$.
+Private points are fined as the collection of all local points that are not a boundary point, i.e.
+$L_p = P_p\cup B_p$.
+Algorithm \ref{alg:overlap} illustrates how this labeling can help with computation/computation overlap of an arbitrary local operation $\phi^{n+1} \gets \mathcal{F}(\phi^n)$.
+Note that \texttt{p4est} library already includes all the primitives required for labeling local points without any further communication.
+
+\begin{algorithm}[htbp]
+\caption{$\phi^{n+1} \gets \texttt{Overlap (}\phi^n, \mathcal{F}\texttt{)}$}\label{euclid}
+\begin{algorithmic}[1]
+\For {$i:B_p$} \Comment{I -- perform computation on boundary points}
+	\State $\phi^{n+1}_i \gets \mathcal{F}(\phi_i^n})$
 \EndFor
+\State $send\_req \gets \texttt{MPI\_Isend(}\phi^{n+1}_B\texttt{)}$ \Comment{II -- begin updating ghost values}
+\State $recv\_req \gets \texttt{MPI\_Irecv(}\phi^{n+1}_G\texttt{)}$
+\For {$i:P_p$} \Comment{III -- perform computation on private points}
+	\State $\phi^{n+1}_i \gets \mathcal{F}(\phi_i^n})$ 
+\EndFor
+\State $\texttt{MPI\_Waitall(}send\_req, recv\_req\texttt{)}$ \Comment{IV -- finish updating ghost values}
+\State \Return $\phi^{n+1}$
 \end{algorithmic}
-\caption{General structure of the parallel implementation of the reinitialization procedure presented in section \ref{section::reinitialization}.} \label{algo::reinitialization}
+\label{alg:overlap}
 \end{algorithm}

### 2014-11-13 paper/tex/parallel_algorithms.tex
Fixed minor issues in .tex. Added draft.pdf
more edits
added interpolation results

@@
 First, we label all local points, $L_p$, as either private, $P_p$, or boundary, $B_p$.
-Here, boundary points are the collection of all local points that are regarded as ghost point, $G_p$, on all other processors, i.e.
+Here, boundary points are the collection of all local points that are regarded as ghost point, $G_r$, on all other processors, i.e.
 $B_p = \underset{r,\;r\neq p}{\bigcup} G_r$.
 Private points are fined as the collection of all local points that are not a boundary point, i.e.
-$L_p = P_p\cup B_p$.
+$P_p = L_p \setminus B_p$.
 Algorithm \ref{alg:overlap} illustrates how this labeling can help with computation/computation overlap of an arbitrary local operation $\phi^{n+1} \gets \mathcal{F}(\phi^n)$.
@@
 \For {$i:B_p$} \Comment{I -- perform computation on boundary points}
-	\State $\phi^{n+1}_i \gets \mathcal{F}(\phi_i^n})$
+	\State $\phi^{n+1}_i \gets \mathcal{F}(\phi_i^n)$
 \EndFor
@@
 \For {$i:P_p$} \Comment{III -- perform computation on private points}
-	\State $\phi^{n+1}_i \gets \mathcal{F}(\phi_i^n})$ 
+	\State $\phi^{n+1}_i \gets \mathcal{F}(\phi_i^n)$ 
 \EndFor
-\State $\texttt{MPI\_Waitall(}send\_req, recv\_req\texttt{)}$ \Comment{IV -- finish updating ghost values}
+\State $\texttt{MPI\_Waitall(}send\_req, recv\_req\texttt{)}$ \Comment{IV -- wait for ghost update to finish}
 \State \Return $\phi^{n+1}$

### 2014-11-13 paper/tex/scaling_results.tex
added interpolation results

@@
 %!TEX root = draft.tex
-\section{scalability and accuracy}
+\section{Scaling} \label{sec:scaling}
 
 \subsection{Interpolation}
+In this section we show results for a simple test to measure the scalability of the interpolation algorithm \ref{alg:interpolation}.
 
-\begin{figure}[ht!]
+\begin{figure}[htbp]
 	\begin{center}
-\includegraphics[width=.7\textwidth]{pictures/scaling_interpolation_all.pdf}
-\caption{Scalability of the interpolation procedure. blue: alpha=0.05, purple alpha=0.50, red alpha=0.95 (alpha = percent of non-local points)}
+		\subfigure[]{\includegraphics[width = 0.45 \textwidth] {figures/interpolation_small_05.pdf}}
+		\subfigure[]{\includegraphics[width = 0.45 \textwidth] {figures/interpolation_small_95.pdf}}
+		\\
+		\subfigure[]{\includegraphics[width = 0.45 \textwidth] {figures/interpolation_large_05.pdf}}
+		\subfigure[]{\includegraphics[width = 0.45 \textwidth] {figures/interpolation_large_95.pdf}}
 	\end{center}
@@
 
-\subsection{Semi lagrangian advection}
+% \subsection{Semi lagrangian advection}
 
-\subsection{Reinitialization} \label{section::scaling_reinitialization}
+% \subsection{Reinitialization} \label{section::scaling_reinitialization}
 
-The reinitialization procedure presented in section \ref{section::reinitialization} necessitates the second order derivatives of the level-set function for the second order discretization in space.
+% The reinitialization procedure presented in section \ref{section::reinitialization} necessitates the second order derivatives of the level-set function for the second order discretization in space.
 The finite difference schemes we implemented rely on the description of the neighborhood of each vertex in the adaptive tree environment.
@@
 
-The scaling results are presented in figure \ref{fig::scaling_reinitialization} and demonstrate satisfying scalability.
+% The scaling results are presented in figure \ref{fig::scaling_reinitialization} and demonstrate satisfying scalability.
 Taking for reference the timing with $128$ processes, we observe efficiencies of $0.877$ with $1024$ processes, $0.763$ with $2048$ processes and $0.609$ with $4096$ processes.
 
-\begin{figure}[ht!]
-\begin{center}
-\includegraphics[width=.7\textwidth]{pictures/scaling_reinitialization_1st_time_2nd_space_with_buffer.pdf}
-\caption{Scalability of the reinitialization procedure presented in section \ref{section::reinitialization} and analyzed in section \ref{section::scaling_reinitialization}.} \label{fig::scaling_reinitialization}
-\end{center}
-\end{figure}
+% \begin{figure}[ht!]
+% \begin{center}
+% \includegraphics[width=.7\textwidth]{figures/scaling_reinitialization_1st_time_2nd_space_with_buffer.pdf}
+% \caption{Scalability of the reinitialization procedure presented in section \ref{section::reinitialization} and analyzed in section \ref{section::scaling_reinitialization}.} \label{fig::scaling_reinitialization}
+% \end{center}
+% \end{figure}
 
-\subsection{Poisson solver}
+% \subsection{Poisson solver}

### 2014-11-14 paper/tex/scaling_results.tex
added scaling results for interpolation

@@
 \section{Scaling} \label{sec:scaling}
+In this section we present results that show the scalability of our algorithms.
+All of our tests have been run on the Stampede cluster at Texas Advanced Computing Center (TACC).
+Each node of Stampede has 2 eight-core Xenon E5-2680 processors clocked at 2.7 GHz with 32 GB of DDR3-1600 MHz memory and interconnected using an InfiniBand network.
+Unless explicitly mentioned otherwise, in all tests we use all 16 cores of every node.
+Finally, in all cases we report the maximum wall time recorded using PETSc's logging interface which has a temporal resolution of roughly $0.1 \: \mu s$.
 
@@
 In this section we show results for a simple test to measure the scalability of the interpolation algorithm \ref{alg:interpolation}.
+The test consists of interpolating a function at a number of random points on a randomly refined octree in three dimensions.
+We consider two cases.
+A small test on a level 9 tree with roughtly 20M cells and 33M nodes and larger test on a level 13 tree with roughly 128M cells and 280M nodes.
+In both cases number of randomly generated points are chosen to be equal to the number of nodes and the stablized second-order interpolation of \cite{Min;Gibou:07:A-second-order-accur} was performed 10 times to smooth out possible timing fluctuations.
+
+To simulate the effect of different CFL numbers, we generate the random points such that on each processor $\alpha$ percentage of them are located outside the processor boundary and thus will initiate communication.
+Scaling results are presented for $\alpha = 5 \%$ and $\alpha = 95\%$ for both small and large problems in figure \ref{fig:interpolation}.
+Excellent scaling is obtained for the small problem for $P = 16-512$ even when $95\%$ of interpolation points are remote to processors.
+For the larger problem, communication overhead prevents further scaling of algorithm beyond $4096$ processors when $\alpha = 95\%$.
+Note that for $\alpha = 5\%$, even though some sections of algorithm stop scaling, the overall algorithm still scales since the local work dominates the timing.
+This illustrates the effectiveness of non-blocking communication pattern in algorithm \ref{alg:interpolation}.
 
@@
 	\begin{center}
-		\subfigure[]{\includegraphics[width = 0.45 \textwidth] {figures/interpolation_small_05.pdf}}
-		\subfigure[]{\includegraphics[width = 0.45 \textwidth] {figures/interpolation_small_95.pdf}}
+		\subfigure[$N_G = 33\times 10^6$, $\alpha = 5\%$]{\includegraphics[width = 0.45 \textwidth] {figures/interpolation_small_05.pdf}}
+		\subfigure[$N_G = 33\times 10^6$, $\alpha = 95\%$]{\includegraphics[width = 0.45 \textwidth] {figures/interpolation_small_95.pdf}}
 		\\
-		\subfigure[]{\includegraphics[width = 0.45 \textwidth] {figures/interpolation_large_05.pdf}}
-		\subfigure[]{\includegraphics[width = 0.45 \textwidth] {figures/interpolation_large_95.pdf}}
+		\subfigure[$N_G = 280\times 10^6$, $\alpha = 5\%$]{\includegraphics[width = 0.45 \textwidth] {figures/interpolation_large_05.pdf}}
+		\subfigure[$N_G = 280\times 10^6$, $\alpha = 95\%$]{\includegraphics[width = 0.45 \textwidth] {figures/interpolation_large_95.pdf}}
 	\end{center}
+	\caption{Strong scaling of algorithm \ref{alg:interpolation} for several tests where $N_G$ denotes the number of random interpolation points (same as number of nodes in the octree) and $\alpha$ denotes the percentage of these points that are remote to each processor.
+Here ``Total'' represents the total time spent in the interpolation while ``buffer'', ``local'', ``queries'', and ``replies'' represent the the timing for different sections (c.f.
+algorithm \ref{alg:interpolation}).
+Results indicate excellent scaling for the small test (a-b) and for the large test when $\alpha = 5\%$ (c).
+For the extreme case (d) the algorithm stops scaling at $4096$ processors due to communication overhead.}
+	\label{fig:interpolation}
 \end{figure}
 
-% \subsection{Semi lagrangian advection}
+\subsection{Semi-Lagrangian}
 

### 2014-11-25 paper/tex/scaling_results.tex
Updated scaling results

@@
 In this section we present results that show the scalability of our algorithms.
-All of our tests have been run on the Stampede cluster at Texas Advanced Computing Center (TACC).
-Each node of Stampede has 2 eight-core Xenon E5-2680 processors clocked at 2.7 GHz with 32 GB of DDR3-1600 MHz memory and interconnected using an InfiniBand network.
-Unless explicitly mentioned otherwise, in all tests we use all 16 cores of every node.
+All of our tests have been run on the Stampede cluster at Texas Advanced Computing Center (TACC) where we are limited to $4096$ cores at most.
+Each node of Stampede has 2 eight-core Xenon E5-2680 processors clocked at 2.7 GHz with 32 GB of DDR3-1600 MHz memory and interconnected using an InfiniBand network card.
+Unless mentioned otherwise, in all the tests we have used all 16 cores of every node.
 Finally, in all cases we report the maximum wall time recorded using PETSc's logging interface which has a temporal resolution of roughly $0.1 \: \mu s$.
@@
 	\begin{center}
-		\subfigure[$N_G = 33\times 10^6$, $\alpha = 5\%$]{\includegraphics[width = 0.45 \textwidth] {figures/interpolation_small_05.pdf}}
-		\subfigure[$N_G = 33\times 10^6$, $\alpha = 95\%$]{\includegraphics[width = 0.45 \textwidth] {figures/interpolation_small_95.pdf}}
+		\subfigure[$N_G = 33$M, $\alpha = 5\%$]{\includegraphics[width = 0.45 \textwidth] {figures/interpolation_small_05.pdf}}
+		\subfigure[$N_G = 33$M, $\alpha = 95\%$]{\includegraphics[width = 0.45 \textwidth] {figures/interpolation_small_95.pdf}}
 		\\
-		\subfigure[$N_G = 280\times 10^6$, $\alpha = 5\%$]{\includegraphics[width = 0.45 \textwidth] {figures/interpolation_large_05.pdf}}
-		\subfigure[$N_G = 280\times 10^6$, $\alpha = 95\%$]{\includegraphics[width = 0.45 \textwidth] {figures/interpolation_large_95.pdf}}
+		\subfigure[$N_G = 280$M, $\alpha = 5\%$]{\includegraphics[width = 0.45 \textwidth] {figures/interpolation_large_05.pdf}}
+		\subfigure[$N_G = 280$M, $\alpha = 95\%$]{\includegraphics[width = 0.45 \textwidth] {figures/interpolation_large_95.pdf}}
 	\end{center}
@@
 \subsection{Semi-Lagrangian}
+To test the scalability of the Semi-Lagrangian algorithm \ref{alg:semi-lagrangian} we consider a slightly modified version of the Enright's rotation test \cite{Enright;Fedkiw;Ferziger;etal:02:A-Hybrid-Particle-Le}, i.e.
+we advect a sphere of radius 0.35 located at $(0.4, 0.4, 0.4)$ with a divergent free velocity field given by:
+\bean
+u(x,y,z) &=& 2\sin(\pi x)^2 \sin(2\pi y) \sin(2 \pi z), \\
+v(x,y,z) &=& -\sin(\pi y)^2 \sin(2\pi x) \sin(2 \pi z),\\
+w(x,y,z) &=& -\sin(\pi z)^2 \sin(2\pi x) \sin(2 \pi y). 
+\eean
+To understand the effect of CFLnumber on the scalability of the algorithm we perform one step of the Semi-Lagrangian algorithm for $\text{CFL} = 1$, $\text{CFL} = 10$, and $\text{CFL} = 100$.
+We also perform the test for two set of initial girds, a small grid with $l_\max = 10$ and a larger grid with $l_\max = 12$.
+After one advection step, each of these grids have approximately 15M and 255M nodes, respectively.
+\begin{table}[htbp]
+	\begin{minipage}{.5\linewidth}
+		\begin{center}
+		\begin{tabular}{ccccccc}
+			\hline
+				CFL & 16 & 32 & 64 & 128 & 256 & 512 \\
+				\hline
+				1   & 2 & 2 & 3 & 3 & 3 & 3 \\
+				10  & 3 & 3 & 3 & 3 & 3 & 3 \\
+				100 & 6 & 6 & 6 & 6 & 6 & 6 \\
+			\hline
+		\end{tabular}
+		\caption*{(a) $l_\max = 10$}
+		\end{center}
+	\end{minipage}%
+	\begin{minipage}{.5\linewidth}
+		\begin{center}
+		\begin{tabular}{ccccccc}
+			\hline
+				CFL & 128 & 256 & 512 & 1024 & 2048 & 4096 \\
+			\hline
+				1   & 3 & 3 & 3 & 3 & 3 & 3 \\
+				10  & 3 & 3 & 4 & 4 & 4 & 4 \\
+				100 & 6 & 6 & 6 & 6 & 6 & 7 \\
+			\hline
+		\end{tabular}
+		\caption*{(b) $l_\max = 12$}
+		\end{center}
+	\end{minipage}	
+	\caption{Number of iterations required for grid construction in algorithm \ref{alg:semi-lagrangian} for the rotation test on a (a) level-10 and (b) level-12 octree with approximately 15M and 255M nodes, respectively.
+Note how the iteration count increases with the CFL number but is almost independent of number of processors.
+The slight dependence of iteration count on number of processors is most likely due to dependence of round-off errors on the number of processors.
+Nonetheless, close examination of generated octrees (data not shown) reveals that they are identical and independent of the number of processors used to perform the test.}
+	\label{tab:semilagrangian}
+\end{table}
+
+Table \ref{tab:semilagrangian} illustrates the dependence of number of iterations required to build the grid; as the CFL is increased, interface travels a farther distance which necessitates more iterations to generate the grid.
+Note that the apparent dependence on the number processors is most likely due to round-off errors during the interpolation and/or backtracking step.
+Figures \ref{fig:semilagrangian_small} and \ref{fig:semilagrangian_large} illustrate the scalability of the algorithm for small and large problems, respectively.
+To enable meaningful comparisons between different CFL numbers and number of processors, the maximum time has been scaled to the number of iterations required for grid construction as reported in table \ref{tab:semilagrangian}.
+For both problems excellent scalability is observed for $\text{CFL} = 1$ and $\text{CFL} = 10$.
+The algorithm even shows good scalability when taken to the extreme, i.e.
+for $\text{CFL} = 100$.
+
+An increase in the CFL number has two effects on the algorithm.
+First, a larger fraction of departure points land in domains of remote processors.
+Moreover, these points are potentially dispersed across a larger number of processors.
+This means both the number MPI messages and their volume increases for larger CFL numbers.
+Second, as more points are shipped to remote processors for interpolation, there is a greater chance that interpolation load is unbalanced across processors.
+This is specially true for regions of space in which streamlines cluster.
+Both factors contribute to decrease the scalability of the algorithm at larger CFL numbers.
+To shed more light on this, we have recorded a complete history of communication pattern in the interpolation step.
+
+\begin{figure}[htbp]
+	\begin{center}
+		\subfigure[Semi-Lagrangian, $\text{CFL} = 1$]{\includegraphics[width = 0.3 \textwidth] {figures/SL_Small_1_SemiLagrangian.pdf}}
+		\subfigure[Semi-Lagrangian, $\text{CFL} = 10$]{\includegraphics[width = 0.3 \textwidth] {figures/SL_Small_10_SemiLagrangian.pdf}}
+		\subfigure[Semi-Lagrangian, $\text{CFL} = 100$]{\includegraphics[width = 0.3 \textwidth] {figures/SL_Small_100_SemiLagrangian.pdf}}
+		\\
+		\subfigure[Interpolation, $\text{CFL} = 1$]{\includegraphics[width = 0.3 \textwidth] {figures/SL_Small_1_Interpolation.pdf}}
+		\subfigure[Interpolation, $\text{CFL} = 10$]{\includegraphics[width = 0.3 \textwidth] {figures/SL_Small_10_Interpolation.pdf}}
+		\subfigure[Interpolation, $\text{CFL} = 100$]{\includegraphics[width = 0.3 \textwidth] {figures/SL_Small_100_Interpolation.pdf}}
+	\end{center}
+	\caption{Strong scaling of a single iteration of algorithm \ref{alg:semi-lagrangian} for the rotation test on a level-10 octree with approximately 13M cells and 15M nodes.
+Top row: scaling of various components of the algorithm for (a) $\text{CFL} = 1$, (b) $\text{CFL} = 10$, and (c) $\text{CFL} = 100$.
+Bottom row: breakdown of various components of the interpolation phase for the same CFL numbers.
+Note that the maximum time has been scaled with the number of iterations required to build the tree (c.f.
+table \ref{tab:semilagrangian}).}
+	\label{fig:semilagrangian_small}
+\end{figure}
+
+\begin{figure}[htbp]
+	\begin{center}
+		\subfigure[Semi-Lagrangian, $\text{CFL} = 1$]{\includegraphics[width = 0.3 \textwidth] {figures/SL_Large_1_SemiLagrangian.pdf}}
+		\subfigure[Semi-Lagrangian, $\text{CFL} = 10$]{\includegraphics[width = 0.3 \textwidth] {figures/SL_Large_10_SemiLagrangian.pdf}}
+		\subfigure[Semi-Lagrangian, $\text{CFL} = 100$]{\includegraphics[width = 0.3 \textwidth] {figures/SL_Large_100_SemiLagrangian.pdf}}
+		\\
+		\subfigure[Interpolation, $\text{CFL} = 1$]{\includegraphics[width = 0.3 \textwidth] {figures/SL_Large_1_Interpolation.pdf}}
+		\subfigure[Interpolation, $\text{CFL} = 10$]{\includegraphics[width = 0.3 \textwidth] {figures/SL_Large_10_Interpolation.pdf}}
+		\subfigure[Interpolation, $\text{CFL} = 100$]{\includegraphics[width = 0.3 \textwidth] {figures/SL_Large_100_Interpolation.pdf}}
+	\end{center}
+	\caption{Strong scaling of a single iteration of algorithm \ref{alg:semi-lagrangian} for the rotation test on a level-12 octree with approximately 251M cells and 255M nodes.
+Top row: scaling of various components of the algorithm for (a) $\text{CFL} = 1$, (b) $\text{CFL} = 10$, and (c) $\text{CFL} = 100$.
+Bottom row: breakdown of various components of the interpolation phase for the same CFL numbers.
+Note that the maximum time has been scaled with the number of iterations required to build the tree (c.f.
+table \ref{tab:semilagrangian}).}
+	\label{fig:semilagrangian_large}
+\end{figure}
 

### 2014-12-02 paper/tex/scaling_results.tex
Updated scaling results

@@
 To understand the effect of CFLnumber on the scalability of the algorithm we perform one step of the Semi-Lagrangian algorithm for $\text{CFL} = 1$, $\text{CFL} = 10$, and $\text{CFL} = 100$.
-We also perform the test for two set of initial girds, a small grid with $l_\max = 10$ and a larger grid with $l_\max = 12$.
+We also perform the test for two set of initial girds, a small grid with $l_\text{max} = 10$ and a larger grid with $l_\text{max} = 12$.
 After one advection step, each of these grids have approximately 15M and 255M nodes, respectively.
@@
 		\end{tabular}
-		\caption*{(a) $l_\max = 10$}
+		\caption*{(a) $l_\text{max} = 10$}
 		\end{center}
@@
 		\end{tabular}
-		\caption*{(b) $l_\max = 12$}
+		\caption*{(b) $l_\text{max} = 12$}
 		\end{center}
@@
 To shed more light on this, we have recorded a complete history of communication pattern in the interpolation step.
-
+Figures \ref{fig:communication_128} and \ref{fig:communication_4096} illustrates the effect of CFL number on three different metrics for $p= 128$ and $p=4096$ processors in the first and last iterations of the Semi-Lagrangian algorithm.
 \begin{figure}[htbp]
@@
 
+\begin{figure}[htbp]
+	\begin{center}
+		\subfigure[]{\includegraphics[width = 0.3 \textwidth] {figures/point_first_128.pdf}}
+		\subfigure[]{\includegraphics[width = 0.3 \textwidth] {figures/message_first_128.pdf}}
+		\subfigure[]{\includegraphics[width = 0.3 \textwidth] {figures/volume_first_128.pdf}}
+		\\
+		\subfigure[]{\includegraphics[width = 0.3 \textwidth] {figures/point_last_128.pdf}}
+		\subfigure[]{\includegraphics[width = 0.3 \textwidth] {figures/message_last_128.pdf}}
+		\subfigure[]{\includegraphics[width = 0.3 \textwidth] {figures/volume_last_128.pdf}}		
+	\end{center}
+	\caption{Performance indicators of the first (top row) and last (bottom row) iterations of the Semi-Lagrangian algorithm for the advection test for $p=128$ processors.
+Increasing the CFL number greatly causes load imbalance in the amount of computation done during interpolation (a) and increases the communication volume (c).
+However the CFL number does not seem to appreciably affect the number of messages sent by processors (b).
+During the last Semi-Lagrangian iteration the initial grid, $G_0$ (c.f.
+algorithm \ref{alg:semi-lagrangian}), is very close to the final grid.
+As a result load imbalance is considerably improved (d).
+Curiously, however, communication pattern does not seem to be affected by much between first and last iterations (e-f).}
+	\label{fig:communication_128}
+\end{figure}
+
+\begin{figure}[htbp]
+	\begin{center}
+		\subfigure[]{\includegraphics[width = 0.3 \textwidth] {figures/point_first_4096.pdf}}
+		\subfigure[]{\includegraphics[width = 0.3 \textwidth] {figures/message_first_4096.pdf}}
+		\subfigure[]{\includegraphics[width = 0.3 \textwidth] {figures/volume_first_4096.pdf}}
+		\\
+		\subfigure[]{\includegraphics[width = 0.3 \textwidth] {figures/point_last_4096.pdf}}
+		\subfigure[]{\includegraphics[width = 0.3 \textwidth] {figures/message_last_4096.pdf}}
+		\subfigure[]{\includegraphics[width = 0.3 \textwidth] {figures/volume_last_4096.pdf}}		
+	\end{center}
+	\caption{Performance indicators of the first (top row) and last (bottom row) iterations of the Semi-Lagrangian algorithm for the advection test for $p=4096$ processors.
+Increasing the CFL number greatly causes load imbalance in the amount of computation done during interpolation (a) and increases the communication volume (c).
+However the CFL number does not seem to appreciably affect the number of messages sent by processors (b).
+During the last Semi-Lagrangian iteration the initial grid, $G_0$ (c.f.
+algorithm \ref{alg:semi-lagrangian}), is very close to the final grid.
+As a result load imbalance is considerably improved (d).
+Curiously, however, communication pattern does not seem to be affected by much between first and last iterations (e-f).}
+	\label{fig:communication_4096}
+\end{figure}
+
 % \subsection{Reinitialization} \label{section::scaling_reinitialization}
@@
 % \subsection{Poisson solver}
+% p = 128, folder =  host/large/100, mode = _first_ (min, max, avg, std) 
+% point  : 1.56e+06, 2.82e+06, 2.07e+06, 1.65e+05
+% message: 19, 131, 41.78, 19.73
+% volume : 5.81e+00, 3.45e+01, 1.97e+01, 6.46e+00
+
+% p = 128, folder =  host/large/10, mode = _first_ (min, max, avg, std) 
+% point  : 1.98e+06, 2.21e+06, 2.07e+06, 3.52e+04
+% message: 22, 131, 42.92, 19.56
+% volume : 2.58e+00, 9.64e+00, 4.92e+00, 1.24e+00
+
+% p = 128, folder =  host/large/1, mode = _first_ (min, max, avg, std) 
+% point  : 2.02e+06, 2.20e+06, 2.07e+06, 2.97e+04
+% message: 22, 131, 42.97, 19.53
+% volume : 2.45e+00, 9.03e+00, 4.21e+00, 1.17e+00
+
+% p = 4096, folder =  host/large/100, mode = _first_ (min, max, avg, std) 
+% point  : 8.28e+02, 4.98e+05, 7.57e+04, 3.14e+04
+% message: 11, 373, 41.18, 30.85
+% volume : 5.28e-01, 1.01e+01, 2.65e+00, 9.24e-01
+
+% p = 4096, folder =  host/large/10, mode = _first_ (min, max, avg, std) 
+% point  : 5.56e+04, 1.63e+05, 7.57e+04, 6.07e+03
+% message: 17, 387, 46.73, 31.78
+% volume : 4.01e-01, 3.06e+00, 8.40e-01, 2.21e-01
+
+% p = 4096, folder =  host/large/1, mode = _first_ (min, max, avg, std) 
+% point  : 6.67e+04, 1.59e+05, 7.57e+04, 4.86e+03
+% message: 18, 387, 47.55, 31.72
+% volume : 4.01e-01, 2.93e+00, 7.25e-01, 1.95e-01
+
+% p = 128, folder =  host/large/100, mode = _last_ (min, max, avg, std) 
+% point  : 1.96e+06, 2.17e+06, 2.06e+06, 3.37e+04
+% message: 22, 133, 41.86, 18.62
+% volume : 1.04e+01, 6.63e+01, 2.98e+01, 1.16e+01
+
+% p = 128, folder =  host/large/10, mode = _last_ (min, max, avg, std) 
+% point  : 2.02e+06, 2.16e+06, 2.07e+06, 2.76e+04
+% message: 22, 131, 42.88, 19.53
+% volume : 3.22e+00, 1.26e+01, 6.24e+00, 1.43e+00
+
+% p = 128, folder =  host/large/1, mode = _last_ (min, max, avg, std) 
+% point  : 2.02e+06, 2.19e+06, 2.07e+06, 2.93e+04
+% message: 22, 131, 42.89, 19.54
+% volume : 2.72e+00, 9.54e+00, 4.46e+00, 1.12e+00
+
+% p = 4096, folder =  host/large/100, mode = _last_ (min, max, avg, std) 
+% point  : 6.30e+04, 1.10e+05, 7.57e+04, 4.20e+03
+% message: 16, 357, 41.77, 22.66
+% volume : 9.76e-01, 4.86e+00, 3.78e+00, 5.69e-01
+
+% p = 4096, folder =  host/large/10, mode = _last_ (min, max, avg, std) 
+% point  : 6.65e+04, 1.33e+05, 7.57e+04, 4.50e+03
+% message: 19, 393, 46.68, 27.93
+% volume : 4.40e-01, 4.80e+00, 2.14e+00, 9.88e-01
+
+% p = 4096, folder =  host/large/1, mode = _last_ (min, max, avg, std) 
+% point  : 6.69e+04, 1.55e+05, 7.57e+04, 4.73e+03
+% message: 18, 392, 47.55, 31.19
+% volume : 4.13e-01, 3.91e+00, 9.68e-01, 2.62e-01

### 2014-12-04 paper/tex/scaling_results.tex
Started with reinitialization scaling

--- 2014-11-25 paper/tex/scaling_results.tex
@@
 		\begin{center}
-		\begin{tabular}{ccccccc}
+		\begin{tabular}{|c|cccccc|}
 			\hline
@@
 Both factors contribute to decrease the scalability of the algorithm at larger CFL numbers.
-To shed more light on this, we have recorded a complete history of communication pattern in the interpolation step.
-
 \begin{figure}[htbp]
@@
 
-% \subsection{Reinitialization} \label{section::scaling_reinitialization}
+To better understand the importance of CFL number on the scalability, we have recorded a complete history of communication pattern in the interpolation step.
+Figure \ref{fig:communication_4096} illustrates the effects of the CFL number on three different metrics, namely number of interpolation points\footnote{Note that this includes both local points and those queried by other processors}, $N_p$, number of sent and received messages, $N_m = S + R$, and the total communication volume, $V_m$ in mega bytes (MB), for $p=4096$ processors.
+Moreover, these values are reported for the first (top row) and last (bottom row) iterations of the Semi-Lagrangian algorithm.
+There are several points to make.
+First, increasing the CFL number greatly increases the load imbalance as evident in the spread of data in figure \ref{fig:communication_4096_pf}.
+This is because at higher CFL numbers, it is more likely that some processors receive larger portions of backtracked points.
+Second, increasing the CFL number increases both the communication volume and its spread across processors (c.f.
+figure \ref{fig:communication_4096_vf}.
+Interestingly, however, the number of sent and received messages seems to remain unaffected by the CFL number.
+The bottom row of figure \ref{fig:communication_4096} illustrates better balance both in the computation and communication volume in the last iteration of the Semi-Lagrangian algorithm.
+This can be explained by noting that as the algorithm converges to the final grid, partitioning of the grid essentially becomes equivalent to partitioning of the backtracked points.
+Detailed information about the load balancing and communication pattern is listed in table \ref{tab:communication}.
+\begin{figure}[htbp]
+	\begin{center}
+		\subfigure[]{\includegraphics[width = 0.3 \textwidth] {figures/point_first_4096.pdf}  \label{fig:communication_4096_pf}}
+		\subfigure[]{\includegraphics[width = 0.3 \textwidth] {figures/message_first_4096.pdf}\label{fig:communication_4096_mf}}
+		\subfigure[]{\includegraphics[width = 0.3 \textwidth] {figures/volume_first_4096.pdf} \label{fig:communication_4096_vf}}
+		\\
+		\subfigure[]{\includegraphics[width = 0.3 \textwidth] {figures/point_last_4096.pdf}   \label{fig:communication_4096_pl}}
+		\subfigure[]{\includegraphics[width = 0.3 \textwidth] {figures/message_last_4096.pdf} \label{fig:communication_4096_ml}}
+		\subfigure[]{\includegraphics[width = 0.3 \textwidth] {figures/volume_last_4096.pdf}  \label{fig:communication_4096_vl}}		
+	\end{center}
+	\caption{Performance indicators of the first (top row) and last (bottom row) iterations of the Semi-Lagrangian algorithm for the advection on 4096 processors for $\text{CFL} = 1$ (\drawdiamond{black}), $\text{CFL} = 10$ (\drawsquare{red}), and $\text{CFL} = 100$ (\drawcircle{blue}).
+Increasing the CFL number greatly causes load imbalance in the amount of computation done during interpolation (a) and increases the communication volume (c).
+However the CFL number does not seem to appreciably affect the number of messages sent by processors (b).
+During the last Semi-Lagrangian iteration the initial grid, $G_0$ (c.f.
+algorithm \ref{alg:semi-lagrangian}), is very close to the final grid.
+As a result load imbalance is considerably improved (d).
+Curiously, however, communication pattern does not seem to be affected by much between first and last iterations (e-f).}
+	\label{fig:communication_4096}
+\end{figure}
 
-% The reinitialization procedure presented in section \ref{section::reinitialization} necessitates the second order derivatives of the level-set function for the second order discretization in space.
-The finite difference schemes we implemented rely on the description of the neighborhood of each vertex in the adaptive tree environment.
-This information can either be computed on the fly or stored in a buffer.
-Fetching the neighbors information is rather time consuming and was observed to improve the overall scalability of the algorithm by increasing the local computation load.
-However, it slowed down the procedure significantly and therefore we choose to present the buffered version of the implementation.
-If one was to be limited in memory, constructing the neighborhood information on the fly would be a desirable option and would lead to better scaling, but in our case memory is not a limiting factor.
+\begin{table}[htbp]
+	\centering
+	\resizebox{\columnwidth}{!}{%
+	\begin{tabular}{|c|c|cccc|cccc|}
+	\hline
+	\multirow{2}{*}{CFL} & \multirow{2}{*}{Metric} & \multicolumn{4}{|c|}{First Iteration} & \multicolumn{4}{|c|}{Last Iteration} \\
+	\cline{3-10}
+											 & 												 & min & max & avg & stddev 					   & min & max & avg & stddev 						\\
+	\hline
+	\multirow{3}{*}{1}   & $N_p$        & 6.67E+04 & 1.59E+05 & 7.57E+04 & 4.86E+03 & 6.69E+04 & 1.55E+05 & 7.57E+04 & 4.73E+03 \\
+										   & $N_m$ 			  & 18 			 & 387 		  & 47.55 	 & 31.72 		& 18 			 & 392 			& 47.55 	 & 31.19 		\\
+										   & $V_m$ (MB)   & 4.01E-01 & 2.93E+00 & 7.25E-01 & 1.95E-01 & 4.13E-01 & 3.91E+00 & 9.68E-01 & 2.62E-01 \\
+										   & $T_\max$ (s) & \multicolumn{4}{|c|}{6.96E-01}						& \multicolumn{4}{|c|}{3.67E-01}						\\
+	\hline
+	\multirow{3}{*}{10}  & $N_p$        & 5.56E+04 & 1.63E+05 & 7.57E+04 & 6.07E+03 & 6.65E+04 & 1.33E+05 & 7.57E+04 & 4.50E+03 \\
+										   & $N_m$ 			  & 17 			 & 387 		  & 46.73 	 & 31.78 		& 19 			 & 393 			& 46.68 	 & 27.93 		\\
+										   & $V_m$ (MB)   & 4.01E-01 & 3.06E+00 & 8.40E-01 & 2.21E-01 & 4.40E-01 & 4.80E+00 & 2.14E+00 & 9.88E-01 \\
+											 & $T_\max$ (s) & \multicolumn{4}{|c|}{7.55E-01}						& \multicolumn{4}{|c|}{3.30E-01}						\\
+	\hline
+	\multirow{3}{*}{100} & $N_p$        & 8.28E+02 & 4.98E+05 & 7.57E+04 & 3.14E+04 & 6.30E+04 & 1.10E+05 & 7.57E+04 & 4.20E+03 \\
+										   & $N_m$ 			  & 11 			 & 373 		  & 41.18 	 & 30.85 		& 16 			 & 357 			& 41.77 	 & 22.66 		\\
+										   & $V_m$ (MB)   & 5.28E-01 & 1.01E+01 & 2.65E+00 & 9.24E-01 & 9.76E-01 & 4.86E+00 & 3.78E+00 & 5.69E-01 \\
+											 & $T_\max$ (s) & \multicolumn{4}{|c|}{9.25E-01}						& \multicolumn{4}{|c|}{3.55E-01}						\\
+	\hline
+	\end{tabular}
+	}
+	\caption{Detailed load balancing and communication information for the advection test problem for $\text{CFL} = 1$, $\text{CFL} = 10$, and $\text{CFL} = 100$.
+Note how increasing the CFL number causes load imbalance and increases the communication volume while it does not affect the number of sent and received messages during an iteration of Semi-Lagrangian step.}
+	\label{tab:communication}
+\end{table}
 
-% The scaling results are presented in figure \ref{fig::scaling_reinitialization} and demonstrate satisfying scalability.
-Taking for reference the timing with $128$ processes, we observe efficiencies of $0.877$ with $1024$ processes, $0.763$ with $2048$ processes and $0.609$ with $4096$ processes.
+\subsection{Reinitialization} \label{section::scaling_reinitialization}
 
-% \begin{figure}[ht!]
-% \begin{center}
-% \includegraphics[width=.7\textwidth]{figures/scaling_reinitialization_1st_time_2nd_space_with_buffer.pdf}
-% \caption{Scalability of the reinitialization procedure presented in section \ref{section::reinitialization} and analyzed in section \ref{section::scaling_reinitialization}.} \label{fig::scaling_reinitialization}
-% \end{center}
-% \end{figure}
-
-% \subsection{Poisson solver}
+Finally we present scaling results of our parallel reinitialization algorithm where we extensively make use of algorithm \ref{alg:overlap} for overlapping computation with communication when computing spatial derivatives.
+Our test is consisted of computing the signed distance to a collection of 100 sphere, whose radii and centers have been selected randomly (c.f.
+\ref{fig:reinit}).
+Computation is conducted on a forest with a $3\times3\times3$ macromesh and level-10 octree per macro-cell which results in a final grid with roughly $350$M nodes.
+\begin{figure}
+\centering
+\includegraphics[width=0.5\textwidth]{figures/reinit.png}
+\label{fig:reinit}
+\caption{Computational domain for the reinitialization test}
+\end{figure}

### 2014-12-05 paper/tex/scaling_results.tex
Finished the scaling results. Now we need to add in Stefan application

@@
 	\centering
-	\resizebox{\columnwidth}{!}{%
+	\resizebox{\columnwidth}{!}{
 	\begin{tabular}{|c|c|cccc|cccc|}
@@
 Finally we present scaling results of our parallel reinitialization algorithm where we extensively make use of algorithm \ref{alg:overlap} for overlapping computation with communication when computing spatial derivatives.
-Our test is consisted of computing the signed distance to a collection of 100 sphere, whose radii and centers have been selected randomly (c.f.
-\ref{fig:reinit}).
-Computation is conducted on a forest with a $3\times3\times3$ macromesh and level-10 octree per macro-cell which results in a final grid with roughly $350$M nodes.
+Our test is consisted of computing the signed distance to a collection of 100 sphere, whose radii and centers have been chosen randomly.
+Test is perform on a small, level-8 octree with nearly 21M and a larger, level-10 octree with nearly 337M grid points.
+In both cases the forest is built on a $3\times3\times3$ macro-mesh.
+Figures \ref{fig:reinit} illustrates that our reinitialization algorithm and, in particular the overlapping strategy presented in algorithm \ref{alg:overlap}, scales very well.
+In general we expect similar scaling results for any local, finite-difference based calculations on octrees that can efficiently utilize algorithm \ref{alg:overlap}.
 \begin{figure}
 \centering
-\includegraphics[width=0.5\textwidth]{figures/reinit.png}
+\subfigure[]{\includegraphics[width=0.45\textwidth]{figures/Reinit_Small.pdf}}
+\subfigure[]{\includegraphics[width=0.45\textwidth]{figures/Reinit_Large.pdf}}
+\caption{Scalability of the reinitialization test for a small (left) and large octrees with roughtly 21M and 337M grid points, respectively.
+Excellent results are obtained in both cases, illustrating the scalability of the overlapping strategy (c.f.
+algorithm \ref{alg:overlap}).}
 \label{fig:reinit}
-\caption{Computational domain for the reinitialization test}
 \end{figure}

### 2015-01-01 paper/tex/applications.tex
added scaling figures for stefan, starting to work on that part

@@
 %!TEX root = draft.tex
-\section{Application to Stefan Problems}
+\section{Application to Stefan Problems} \label{sec:application}
 
@@
 
+\begin{figure}
+\centering
+\subfigure[]{\includegraphics[width=0.45\textwidth]{figures/stefan_scaling_small.pdf}}
+\subfigure[]{\includegraphics[width=0.45\textwidth]{figures/stefan_scaling_large.pdf}}
+\caption{Scalability of the Stefan problem for small (left) and large (right) octrees with roughly 7M and 105M grid points, respectively.
+As expected from the scalability analysis of the individual components, we observe excellent results, illustrating the potential of our solver.}
+\label{fig:stefan_scaling}
+\end{figure}
+
 \subsection{Numerical experiments}
@@
 \begin{center}
-\includegraphics[width=.8\textwidth]{pictures/crystal_grid_low.png}
+\subfigure[]{\includegraphics[width=0.45\textwidth]{figures/stefan_grid.png}}
+\subfigure[]{\includegraphics[width=0.45\textwidth]{figures/stefan_grid_zoom.png}}
+\caption{} \label{fig:stefan_grid}
 \end{center}

### 2015-01-01 paper/tex/introduction.tex
added interpolation results
added scaling figures for stefan, starting to work on that part

@@
 These ideas are then extended in section \ref{sec:parallel algorithms} to parallel using a domain decomposition method.
-In section \ref{section::scaling results} we provide several examples that illustrate the scalability of our algorithms.
-Finally, we close by providing an important application of our algorithms in modeling the solidification process by solving a Stefan problem in \ref{section::applications}.
+In section \ref{sec:scaling} we provide several examples that illustrate the scalability of our algorithms.
+Finally, we close by providing an important application of our algorithms in modeling the solidification process by solving a Stefan problem in \ref{sec:application}.

### 2015-01-02 paper/tex/applications.tex
done with first pass for the stefan problem

@@
 %!TEX root = draft.tex
-\section{Application to Stefan Problems} \label{sec:application}
-
-We now propose an application of the level-set method on parallel adaptive cartesian mesh to the well studied stefan problem.
+\section{Application to Stefan Problem} \label{sec:application}
 
@@
 
-We study the problem of the phase transition of a liquid melt to a solid crystaline structure.
-For a liquid melt and a solid crystal with respective heat capacities $c_l$ and $c_s$ and uniform thermal diffusivities $k_l$ and $k_s$, the respective temperatures $T_l$ and $T_s$ diffuse as
+We propose to apply our approach to the study of the phase transition of a liquid melt to a solid crystaline structure.
+In the case of a single component melt and in the absence of convection, the process is dominated by diffusion and can be modeled as a Stefan problem.
+We decompose the computational domain $\Omega$ into two subdomains $\Omega^-$ and $\Omega^+$, separated by an interface $\Gamma$.
+The Stefan problem describes the evolution of the temperature $T$, decomposed into $T_s$ in the solid phase $\Omega^+$ and $T_l$ in the liquid phase $\Omega^-$, as
+\begin{align}
+\pd{T_l}{t} & = D_l \Delta T_l \quad \mathrm{in} ~~ \Omega^-, \label{eq:stefan_heat_equation_l} \\
+\pd{T_s}{t} & = D_s \Delta T_s \quad \mathrm{in} ~~ \Omega^+. \label{eq:stefan_heat_equation_s}
+\end{align}
+The diffusion constants $D_l$ and $D_s$ can be discontinuous across the interface.
+We prescribe homogeneous Neumann boundary conditions on the edge of the computational domain, $\nabla T \cdot \mathbf{n}\vert_{\partial \Omega}=0$.
+At the interface between the solid and the liquid phases, the temperature is given by the Gibbs-Tompson boundary condition \cite{Alexiades;Solomon;Wilson:88:The-formation-of-a-s, Alexiades;Solomon:93:Mathematical-Modelin}
+\begin{equation} \label{eq:stefan_gibbs_tompson}
+T_s = T_l = T_{\Gamma} = -\epsilon_c \kappa - \epsilon_v (\mathbf{V} \cdot \mathbf{n}),
+\end{equation}
+where $\kappa$ is the local interface curvature, $V$ is the velocity of the interface, $\mathbf{n}$ is the outward normal to the solidification front and $\epsilon_c$ and $\epsilon_v$ are the curvature and kinetic undercooling coefficients.
+The interface velocity $V$ is defined from the jump in the heat flux at the interface,
+\begin{equation} \label{eq:stefan_velocity}
+(\mathbf{V} \cdot \mathbf{n}) = - \left[ D_l \pd{T_l}{\mathbf{n}} - D_s \pd{T_s}{\mathbf{n}} \right].
+\end{equation}
+We choose to use an adaptive time step with a $\text{CFL} = 5$, i.e.
+\begin{equation} \label{eq:stefan_dt}
+\Delta t = 5 ~ \Delta x_{min} ~ \min(1,1/\max \lVert \mathbf{V} \rVert),
+\end{equation}
+where $\Delta x_{min}$ is the size of the smallest cell of the forest. The general procedure to solve the Stefan problem is presented in algorithm \ref{alg:stefan}.
 
-\begin{align*}
-c_l \pd{T_l}{t} & = k_l \nabla T_l \quad \mathrm{in} ~~ \Omega^- , \\
-c_s \pd{T_s}{t} & = k_s \Delta T_s \quad \mathrm{in} ~~ \Omega^+ .
-\end{align*}
-
-At the interface between the solid and the liquid phases, the temperature is defined as
-
-\begin{equation*}
-T_s = T_l = T_{\Gamma} = -\epsilon_c \kappa - \epsilon_v (\mathbf{V} \cdot \mathbf{n}),
-\end{equation*}
-
-where $\kappa$ is the local interface curvature, $V$ is the velocity of the interface, $\mathbf{n}$ is the outward normal to the solidification front and $\epsilon_c$ and $\epsilon_v$ are the curvature and kinetic undercooling coefficients.
-The velocity $V$ is defined from the jump in the heat flux at the interface
-
-\begin{equation}
-L (\mathbf{V} \cdot \mathbf{n}) = - \left[ k_l \pd{T_l}{\mathbf{n}} - k_s \pd{T_s}{\mathbf{n}} \right],
-\end{equation}
-
-with $L$ the latent heat.
+\begin{algorithm}[htbp]
+\caption{$\texttt{General procedure for solving the Stefan problem}$}
+\begin{algorithmic}[1]
+\State $\texttt{Initialize the forest and }\phi\texttt{ given the initial geometry}.$
+\State $\texttt{Initialize }T_s\texttt{ in }\Omega^+\texttt{ and }T_l\texttt{ in }\Omega^-.$
+\State $\texttt{Reinitialize }\phi\texttt{ and compute the local interface curvature }\kappa.$
+\State $\texttt{Compute }T^{n+1}_l\texttt{ and }T^{n+1}_s\texttt{ by solving the heat equations \eqref{eq:stefan_heat_equation_l} and \eqref{eq:stefan_heat_equation_s}}.$
+\State $\texttt{Extrapolate }T^{n+1}_s\texttt{ from }\Omega^+\texttt{ to }\Omega^-\texttt{ and }T^{n+1}_l\texttt{ from }\Omega^-\texttt{ to }\Omega^+.$
+\State $\texttt{Compute the velocity field }\mathbf{V}\texttt{ according to \eqref{eq:stefan_velocity}}.$
+\State $\texttt{Compute the time step dt following \eqref{eq:stefan_dt}.}$
+\State $\texttt{Evolve the interface and contruct the new forest using the Semi-Lagrangian procedure}.$
+\State $\texttt{Interpolate }T^{n+1}_s\texttt{ and }T^{n+1}_l\texttt{ from the old forest to the new forest}.$
+\State $\texttt{Go to 3 with n=n+1}.$
+\end{algorithmic}
+\label{alg:stefan}
+\end{algorithm}
 
 \subsection{Scalability}
+
+The implementation of the Stefan problem relies on the components described in the previous sections, and it is therefore a good summary of the performance of the various algorithms.
+We monitored the performance of the code over five time iterations, as presented in algorithm \ref{alg:stefan}, for two different maximum resolutions.
+In both cases, the forest is built on a $20\times20\times20$ macro-mesh.
+The maximum tree resolution for the small test is 9, leading to approximately 7M grid points, and the maximum resolution for the large test is 11, corresponding to 105M grid points.
+The results are presented in figure \ref{fig:stefan_scaling}.
+As expected from the results obtained for each component in the previous sections, our implementation of the Stefan problem exhibits very satisfactory scaling.
 
@@
 
+We now present the results from a large simulation of the Stefan problem.
+We choose to work with a $20\times20\times20$ macro-mesh and with level-10 octrees.
+We choose to define the Gibbs-Tompson anisotropy undercooling coefficients \eqref{eq:stefan_gibbs_tompson} as
+\begin{align*}
+\epsilon_c & = \left[ \epsilon_1 \left( 1+\alpha_1 \cos(3\theta_1) \right) + \epsilon_2 \left( 1+\alpha_2 \cos(3\theta_2) \right) \right] \kappa,\\
+\epsilon_v & = 0,
+\end{align*}
+with $\theta_1$ the angle between the normal to the interface $\mathbf{n}$ and the x-axis in the $(x,y)$ plane, $\theta_2$ the angle between $\mathbf{n}$ and the x-axis in the $(x,z)$ plane, and
+\begin{align*}
+\epsilon_1 & = 2~(\sin(x)+\cos(y)+2)\cdot10^{-6}, & \epsilon_2 & = 2~(\sin(x)+\cos(z)+2)\cdot10^{-6}, \\
+\alpha_1 & = \frac{1}{4}(\cos(x)+\sin(y)+2), & \alpha_2 & = \frac{1}{4}(\cos(x)+\sin(z)+2).
+\end{align*}
+Note that these expressions are arbitrary and lead to very varied crystal shapes.
+The computation is initialized with twenty spherical seeds of radius $0.0015$ placed randomly in the domain.
+We choose $D_s=D_l=1$ and set the initial temperatures $T^0_l=-0.25$ and $T^0_s=0$.
+
+We let the simulation run with $256$ mpi processes for $6$ hours and $30$ minutes, resulting in $396$ time iterations.
+Visualizations of the final iteration are presented in figures \ref{fig:stefan_grid} and \ref{fig:stefan_crystals}.
+The final iteration of the simulation consisted of $167$M grid points when a uniform grid with the equivalent finest resolution would lead to $8.59\cdot10^{12}$ grid points, i.e.
+over eight trillion grid points !
+Our simulation used only $0.001948\%$ of the number of grid points needed for the same simulation on a uniform grid.
+This application demonstrates the potential of our approach.
+
 \begin{figure}[ht!]
 \begin{center}
+\subfigure[]{\includegraphics[width=0.45\textwidth]{figures/stefan_grid_overview.png}}
+\subfigure[]{\includegraphics[width=0.45\textwidth]{figures/stefan_temperature.png}}
 \subfigure[]{\includegraphics[width=0.45\textwidth]{figures/stefan_grid.png}}
 \subfigure[]{\includegraphics[width=0.45\textwidth]{figures/stefan_grid_zoom.png}}
-\caption{} \label{fig:stefan_grid}
+\caption{Visualization of the computational mesh (a,b,c) and the temperature field (b) for the Stefan problem simulation.} \label{fig:stefan_grid}
 \end{center}
 \end{figure}
+
+\begin{figure}[ht!]
+\begin{center}
+\subfigure[]{\includegraphics[width=0.45\textwidth]{figures/stefan_crystal_1.png}}
+\subfigure[]{\includegraphics[width=0.45\textwidth]{figures/stefan_crystal_2.png}}
+\subfigure[]{\includegraphics[width=0.45\textwidth]{figures/stefan_crystal_4.png}}
+\subfigure[]{\includegraphics[width=0.45\textwidth]{figures/stefan_crystal_3.png}}
+\caption{Representation of the computational domain (a) and of various crystal observed for the Stefan problem simulation.} \label{fig:stefan_crystals}
+\end{center}
+\end{figure}

### 2015-01-07 paper/tex/introduction.tex
finished first pass

@@
 The level-set method, originally proposed by Sethian and Osher \cite{Osher;Sethian:88:Fronts-Propagating-w}, is a popular and powerful framework for tracking arbitrary interfaces that undergo complicated topological changes.
-As a result, the level-set method has wide range of application such as multiphase flows, moving boundary problems, image segmentation, and computer graphics graphics \cite{Osher;Fedkiw:01:Level-Set-Methods:-A,Sethian:99:Level-set-methods-an}.
+As a result, the level-set method has wide range of application such as multiphase flows, moving boundary problems, image segmentation, and computer graphics \cite{Osher;Fedkiw:01:Level-Set-Methods:-A,Sethian:99:Level-set-methods-an}.
 An important feature which makes this method powerful and easy to use is that the location of interface is defined implicitly on an underlying grid.
@@
 First, compared to an explicit method, e.g.
-front tracking \cite{Juric:96:A-Front-Tracking-Met, Tryggvason;Bunner;Esmaeeli;etal:01:A-Front-Tracking-Met}, level-set method is typically less accurate and mass conservation could be a problem; although progress has been made in resolving this issue \cite{Enright;Fedkiw;Ferziger;etal:02:A-Hybrid-Particle-Le}.
+front tracking \cite{Juric:96:A-Front-Tracking-Met, Tryggvason;Bunner;Esmaeeli;etal:01:A-Front-Tracking-Met}, the level-set method is typically less accurate and mass conservation could be a problem; although progress has been made in resolving this issue \cite{Enright;Fedkiw;Ferziger;etal:02:A-Hybrid-Particle-Le}.
 Second, the level-set function has to be defined in a higher dimensional space compared to the interface.
-If only the location of interface is needed, the added dimension greatly increases the overall computational cost.
+If only the location of the interface is needed, the added dimension greatly increases the overall computational cost.
 One way to avoid this problem is by computing the level-set only close to the interface, e.g.
@@
 In \cite{Strain:99:Tree-Methods-for-Mov} the idea of using tree-based grids for level-set calculations was first introduced and later extended in \cite{Losasso;Gibou;Fedkiw:04:Simulating-Water-and} for graphics application.
-More recently, authors in \cite{Min;Gibou:07:A-second-order-accur} proposed second-order accurate level-set methods on Quadtree (two dimensions) and Octree (three dimensions) grids.
-Use of tree-based adaptive grids in the context of level-set method is quite advantageous because: 1) It gives fine-grain control over error, which typically occurs close to the interface and 2) It can effectively reduce the dimensionality of the problem by focusing most of the grid cells close to the interface.
-Fortunately, construction of the tree is quite simple in the presence of an interface that naturally defines an ideal metric for refinement.
-Although use of adaptive grids can dramatically reduce the computational cost, performing high-resolution three dimensional calculations of complex interfacial problems, e.g.
+More recently, authors in \cite{Min;Gibou:07:A-second-order-accur} proposed second-order accurate level-set methods on Quadtree (two spatial dimensions) and Octree (three spatial dimensions) grids.
+The use of tree-based adaptive grids in the context of the level-set method is quite advantageous because: 1) It gives fine-grain control over errors, which typically occur close to the interface and 2) It can effectively reduce the dimensionality of the problem by focusing most of the grid cells close to the interface.
+Fortunately, constructing the tree is quite simple in the presence of an interface that naturally defines an ideal metric for refinement.
+Although the use of adaptive grids can dramatically reduce the computational cost, performing high-resolution three dimensional calculations of complex interfacial problems, e.g.
 crystal growth in binary alloys \cite{Theillard;Gibou;Pollock:14:A-Sharp-Computationa}, could take a very long time in serial.
-In this paper we extend these algorithms by proposing parallel algorithms for distributed memory machines using domain decomposition technique.
+In this paper we extend these algorithms by proposing parallel algorithms for distributed memory machines using a domain decomposition technique.
 
@@
 For instance, this was the approach originally taken by the \texttt{deal.II} library \cite{Bangerth;Hartmann;Kanschat:07:deal.II----a-General}.
-This approach, however, is only scalable to a few hundred processor at best and is limited by the size of the grid itself.
-Moreover, use of a general-purpose graph partitioner adds extra overhead that can limit the overall scalability even further.
-Interestingly, tree-based grids have nice spatial ordering that naturally lead to the concept of space-filling curves (SFCs) which can be efficiently exploited for parallel load balancing \cite{Aluru;Sevilgen:97:Parallel-domain-deco,Campbell;Devine;Flaherty;etal:03:Dynamic-octree-load-}.
+This approach, however, is only scalable to a few hundred processors at best and is limited by the size of the grid itself.
+Moreover, the use of a general-purpose graph partitioner adds extra overhead that can limit the overall scalability even further.
+Interestingly, tree-based grids have nice spatial ordering that naturally leads to the concept of space-filling curves (SFCs) which can be efficiently exploited for parallel load balancing \cite{Aluru;Sevilgen:97:Parallel-domain-deco,Campbell;Devine;Flaherty;etal:03:Dynamic-octree-load-}.
 
@@
 \texttt{Dendro} \cite{Sampath;Adavani;Sundar;etal:08:Dendro:-parallel-alg} is another example of an octree code in which similar ideas are used for parallel partitioning and development of a parallel geometric multigrid that has been scaled up to about 32,000 cores \cite{Sampath;Biros:10:A-parallel-geometric}.
-More recently, authors in \cite{Burstedde;Wilcox;Ghattas:11:p4est:-Scalable-Algo} extended these ideas to a collection, or a ``forest'', of octrees that are connected through a common, potentially unstructured, hexahedral grid.
+More recently, authors in \cite{Burstedde;Wilcox;Ghattas:11:p4est:-Scalable-Algo} extended these ideas to a collection, or a ``forest'', of Octrees that are connected through a common, potentially unstructured, hexahedral grid.
 This forest is then partitioned in parallel using a global Morton curve.
-Implementation of these algorithms, which were shown to scale to more than 200,000 processors, are publicly available through simple API provided by the \texttt{p4est} library \cite{p4est-github}.
-In fact the algorithms presented in this paper are directly implemented on top of \texttt{p4est} library and we do not discuss any algorithm that is already covered in \cite{Burstedde;Wilcox;Ghattas:11:p4est:-Scalable-Algo}.
+The implementation of these algorithms, which were shown to scale to more than 200,000 processors, is publicly available through a simple API provided by the \texttt{p4est} library \cite{p4est-github}.
+In fact the algorithms presented in this paper are directly implemented on top of the \texttt{p4est} library and we do not discuss any algorithm that is already covered in \cite{Burstedde;Wilcox;Ghattas:11:p4est:-Scalable-Algo}.
 We also make use of the popular \texttt{PETSc} \cite{Balay;Abhyankar;Adams;etal:14:PETSc-Web-page} library for linear algebra and its parallel primitives, such as parallel ghosted vector and scatter/gather operations, which simplifies the implementation.
@@
 Eulerian advection schemes can easily be parallelized but unfortunately are limited by the CFL condition which could be very restrictive for adaptive grids.
-Semi-Lagrangian methods combine the unconditional stability of Lagrangian methods and the ease of use of Eulerian grids and have been successfully used for advecting the level-set function on tree-based grids before \cite{Min;Gibou:07:A-second-order-accur}.
+Semi-Lagrangian methods combine the unconditional stability of Lagrangian methods and the ease of use of Eulerian grids and have been successfully used for advecting the level-set function on tree-based grids in the past \cite{Min;Gibou:07:A-second-order-accur}.
 However, parallelizing the semi-Lagrangian algorithm in a domain decomposition context is not an easy task.
-The reason for this is twofold: First, depending on the CFL number, the departure points may end up outside the ghost region and in remote processors that are potentially far away.
+The reason for this is twofold.
+First, depending on the CFL number, the departure points may end up outside the ghost region and in remote processors that are potentially far away.
 This requires a very dynamic and nonuniform communication pattern which is complicated to implement.
-For adaptive grids situation is even more complicated due to the asymmetric nature of communication (c.f.
+For an adaptive grid situation, it is even more complicated due to the asymmetric nature of the communication (c.f.
 \ref{sec:parallel algorithms}).
@@
 Nonetheless, several parallel semi-Lagrangian algorithms have been proposed in the past.
-A simple domain decomposition technique was used in \cite{Thomas;Cote:95:Massively-parallel-s} where the width of ghost layer is fixed based on the maximum CFL number to ensure all departure points are covered by the ghost layer.
-At large CFL numbers, this leads to large communication volume that can limit the scalability.
-Nonetheless good scaling was report for small CFL numbers ($\text{CFL} \le 2$).
-In \cite{Drake;Foster;Michalakes;etal:95:Design-and-performan} authors propose a more sophisticated domain decomposition approach which uses a ``dynamic ghost layer''.
+A simple domain decomposition technique was used in \cite{Thomas;Cote:95:Massively-parallel-s} where the width of the ghost layer is fixed based on the maximum CFL number to ensure that all the departure points are covered by the ghost layer.
+At large CFL numbers, this leads to a large volume of communication that can limit the scalability.
+Nonetheless good scaling was reported for small CFL numbers ($\text{CFL} \le 2$).
+In \cite{Drake;Foster;Michalakes;etal:95:Design-and-performan} the authors propose a more sophisticated domain decomposition approach which uses a ``dynamic ghost layer''.
 Here the width of the ghost layer is dynamically determined at runtime based on information from previous time steps.
-Unfortunately, however, this approach seems to suffer from excessive communication overhead at larger number of processors as well.
-More recently, authors in \cite{White-III;Dongarra:11:High-performance-hig} used a domain decomposition strategy on a cubed sphere but with a single layer of ghost nodes.
+Unfortunately, however, this approach seems to suffer from excessive communication overhead at larger numbers of processors as well \comment{as well ?}.
+More recently, the authors in \cite{White-III;Dongarra:11:High-performance-hig} used a domain decomposition strategy on a cubed sphere but with a single layer of ghost nodes.
 Interpolation on remote processors is then handled by sending query points to the corresponding processor and asking for the interpolated result.
@@
 Although in this article we are mainly interested in parallel semi-Lagrangian methods, one could resort to finite difference or finite element discretization methods if small CFL numbers are acceptable.
-Indeed several algorithms of this type has been proposed with applications in modeling dendritic crystal growth \cite{Wang;Chang;Kale;etal:06:Parallelization-of-a}, multiphase flows \cite{Sussman:05:A-parallelized-adapt, Fortmeier;Bucker:11:A-parallel-strategy-, Rodriguez;Sahni;Lahey-Jr;etal:13:A-parallel-adaptive-}, and atomization process \cite{Herrmann:10:A-parallel-Eulerian-}.
+Indeed several algorithms of this type have been proposed with applications in modeling dendritic crystal growth \cite{Wang;Chang;Kale;etal:06:Parallelization-of-a}, multiphase flows \cite{Sussman:05:A-parallelized-adapt, Fortmeier;Bucker:11:A-parallel-strategy-, Rodriguez;Sahni;Lahey-Jr;etal:13:A-parallel-adaptive-}, and atomization process \cite{Herrmann:10:A-parallel-Eulerian-}.
 %, and image segmentation on GPUs \cite{Lefohn;Cates;Whitaker:03:Interactive-GPU-base,Cates;Lefohn;Whitaker:04:GIST:-an-interactive,Roberts;Packer;Sousa;etal:10:A-work-efficient-GPU}.
@@
 This is the approach we have chosen in this paper.
-If, however, the signed-distance property is required in the entire domain solving the Eikonal equation is more computationally efficient.
-Unfortunately, however, the most popular algorithm for solving the Eikonal equation, i.e.
+However, if the signed-distance property is required in the entire domain solving the Eikonal equation is more computationally efficient.
+Unfortunately, the most popular algorithm for solving the Eikonal equation, i.e.
 the Fast Marching Method \cite{Sethian:96:A-Fast-Marching-Leve,Sethian:99:Level-set-methods-an}, is inherently sequential due to causal relationship between grid points and cannot be easily parallelized.
@@
 One of the earliest attempt in parallelizing the FMM is reported in \cite{Herrmann:03:A-domain-decompositi} where a domain decomposition algorithm was introduced.
-Unlike the serial FMM, however, parallel FMM potentially requires multiple iteration or ``rollback operations'' to enforce the causality across processors.
+Unlike the serial FMM, however, parallel FMM potentially requires multiple iterations or ``rollback operations'' to enforce the causality across processors.
 Similar ideas are described in details in \cite{Tugurlan:08:Fast-marching-method}.
-It should be noted that number of iterations for the parallel FMM to converge greatly depends on the complexity of the interface and also the parallel partitioning and, in general, fewer iterations are required if domains are aligned with the normals to the interface.
-Due to the nature of Eikonal equation, shared memory machines might be a better environment for parallelization.
-For instance, in \cite{Breus;Cristiani;Gwosdek;etal:11:An-adaptive-domain-d} authors use an ``adaptive'' technique where individual threads implicitly define a domain decomposition at runtime.
-Unfortunately, however, this approach does not seem to be more effective than a simple static decomposition.
+It should be noted that the number of iterations needed for the parallel FMM to converge greatly depends on the complexity of the interface and on the parallel partitioning and, in general, fewer iterations are required if the domains are aligned with the normals to the interface.
+Due to the nature of the Eikonal equation, shared memory machines might be a better environment for parallelization.
+For instance, in \cite{Breus;Cristiani;Gwosdek;etal:11:An-adaptive-domain-d} the authors use an ``adaptive'' technique where individual threads implicitly define a domain decomposition at runtime.
+Unfortunately, this approach does not seem to be more effective than a simple static decomposition.
 In \cite{Zhao:07:Parallel-implementat} a parallel FSM method was presented for the fist time.
-However, a more scalable FSM was more recently proposed in \cite{Detrixhe;Gibou;Min:13:A-parallel-fast-swee} where the Cuthill-McKee numbering was utilize to expose more parallelism.
+However, a more scalable FSM was more recently proposed in \cite{Detrixhe;Gibou;Min:13:A-parallel-fast-swee} where the Cuthill-McKee numbering was utilized to expose more parallelism.
 More recently a two-scale, hybrid FMM-FSM was presented in \cite{Chacon;Vladimirsky:13:A-parallel-Heap-Cell} which, albeit being more complicated to implement, promises even better scalability.
@@
 The FIM is similar to FMM in that it also maintains a list of ``active nodes''.
-Unlike FMM, however, FIM avoids sorting the list and allows for concurrent updating of all nodes in an iterative fashion.
+However, unlike FMM, FIM avoids sorting the list and allows for concurrent updating of all nodes in an iterative fashion.
+\comment{Shouldn't we say that concluding from this analysis, we choose to use iterative reinitialization ?}
 

### 2015-01-07 paper/tex/levelset_method.tex
finished first pass

@@
 a curve in two spatial dimension can be described as $\Gamma = \{(x,y) | \phi(x,y) = 0\}$ where $\phi(x,y)$ is the level-set function.
-Evolution of the curve under a velocity field $\mathbf{V}$ is then obtained by solving the level-set equation:
+The evolution of the curve under a velocity field $\mathbf{V}$ is then obtained by solving the level-set equation:
 \be
@@
 \ee
-When velocity field does not depend on the level set function itself, equation \eqref{eq:ls} can be solved using the Semi-Lagrangian method.
-An important advantage of the Semi-Lagrangian method over regular finite difference method is its unconditional stability which allows for arbitrarily large time steps.
-This is specially of interest when using adaptive grids that allow for higher grid resolutions.
+When the velocity field does not depend on the level-set function itself, equation \eqref{eq:ls} can be solved using the Semi-Lagrangian method.
+An important advantage of the Semi-Lagrangian method over the regular finite difference method is its unconditional stability which allows for arbitrarily large time steps.
+This is particularly important when using adaptive grids that allow for higher grid resolutions.
 
 In general, an infinite number of level-set functions can describe the same zero contour and thus interface.
-However, it is desirable to chose functions with signed distance property, i.e.
+However, it is desirable to chose a function with the signed distance property, i.e.
 $|\nabla \phi| = 1$.

### 2015-01-07 paper/tex/parallel_algorithms.tex
added scaling figures for stefan, starting to work on that part
finished first pass

@@
 Moreover, adaptive tree-based grids are easy to generate in the presence of a signed-distance level-set function \cite{Min;Gibou:07:A-second-order-accur} and can efficiently be encoded using a tree datastructure \cite{Samet:90:Applications-of-Spat}.
-To develope scalable parallel algorithms on these grids, it is necessary to parallelize the datastructure and grid manipulations method such as refinement and coarsening of cells as well as provide with a fast method for grid partitioning and load balancing.
-\texttt{p4est} library \cite{p4est-github} is a collection of such parallel algorithms that has recently emerged and shown to scale up to 200,000 processors \cite{Burstedde;Wilcox;Ghattas:11:p4est:-Scalable-Algo}.
+In order to develop scalable parallel algorithms on these grids, it is necessary to parallelize the datastructure and grid manipulations methods such as refinement and coarsening of cells as well as to provide a fast method for grid partitioning and load balancing.
+The \texttt{p4est} library \cite{p4est-github} is a collection of such parallel algorithms that has recently emerged and shown to scale up to 200,000 processors \cite{Burstedde;Wilcox;Ghattas:11:p4est:-Scalable-Algo}.
 
@@
 figure \ref{fig:p4est_zcurve}).
-This common coarse grid, which we will refer to as the ``macromesh'', can in general be an unstructured quadrilateral, in two spatial dimensions, or hexahedral mesh, in three spatial dimensions.
+This common coarse grid, which we will refer to as the ``macromesh'', can in general be an unstructured quadrilateral mesh, in two spatial dimensions, or hexahedral mesh, in three spatial dimensions.
 In this article, however, we shall limit ourselves to simple uniform Cartesian macromeshes.
-Moreover it is implicitly assumed that the macromesh is small enough that can be entirely replicated on all processors.
-For instance, in many of the applications that we are interested in this paper the macromesh simply a single cell.
+Moreover it is implicitly assumed that the macromesh is small enough that it can be entirely replicated on all processors.
+For instance, in many of the applications that we are interested in in this paper the macromesh is simply a single cell.
 \texttt{p4est} allows for arbitrary refinement and coarsening criteria through defining callback functions.
@@
 Here $V(C)$ denotes the set of all vertices of cell $C$, $L$ denotes the Lipschitz constant of the level-set function, and $D$ denotes the diagonal size of cell $C$.
-We refer to section 3.2 of \cite{Burstedde;Wilcox;Ghattas:11:p4est:-Scalable-Algo} for details on parallel refinement and coarsening algorithms used in \texttt{p4est}.
+We refer to section 3.2 of \cite{Burstedde;Wilcox;Ghattas:11:p4est:-Scalable-Algo} for details on the parallel refinement and coarsening algorithms used in \texttt{p4est}.
 
 Once the gird is adapted to the interface, it must be partitioned to ensure load balancing across processors.
-This is achieve by constructing a Z-curve that traverses the leaves of all trees in order of tree index (c.f.
+This is achieved by constructing a Z-curve that traverses the leaves of all trees in order of tree index (c.f.
 figure \ref{fig:p4est_zcurve}).
-A Z-curve is a Space Filling Curve (SFC) with the important property that cells with close Z-indecies are also geometrically close in the physical domain.
-This is beneficial since it both leads to reduction in MPI communication and improves the cache performance of several algorithms such as interpolation and finite difference calculations.
+A Z-curve is a Space Filling Curve (SFC) with the important property that cells with close Z-indices are also geometrically close in the physical domain.
+This is beneficial since it leads to both reduction in MPI communication and improvements of the cache performance of several algorithms such as interpolation and finite difference calculations.
 For more details on parallel partitioning in \texttt{p4est} one may refer to section 3.3 of \cite{Burstedde;Wilcox;Ghattas:11:p4est:-Scalable-Algo}.
@@
 Here processors are represented via different colors.
-Note how use of Z-curve naturally leads to clustering of most cells in each processor domain.
+Note how using the Z-curve naturally leads to clustering of most cells in each processor domain.
 Right: Schematic of a tree data structure representing the forest and its partitioning.}
@@
 \end{figure}
-Aside from grid manipulation and partitioning, we use two additional features of \texttt{p4est}, namely generation of ghost layer cells and creating a globally unique node indexing.
+Aside from grid manipulation and partitioning, we use two additional features of \texttt{p4est}, namely the generation of ghost layer cells and the creation of a globally unique node indexing.
 These algorithms are detailed in sections 3.5 and 3.6 of \cite{Burstedde;Wilcox;Ghattas:11:p4est:-Scalable-Algo}.
 An important feature of our algorithms is that they are designed for non-graded trees.
-This is important because we can entirely skip tree balancing which was shown to be one of most time consuming parts of grid adaptation in \texttt{p4est} \cite{Burstedde;Wilcox;Ghattas:11:p4est:-Scalable-Algo}.
+This is important because we can entirely skip tree balancing which was shown to be one of the most time consuming parts of grid adaptation in \texttt{p4est} \cite{Burstedde;Wilcox;Ghattas:11:p4est:-Scalable-Algo}.
 
@@
 only the leaves are explicitly stored.
-However, all of algorithms considered here require explicit knowledge of the entire tree.
+However, all the algorithms considered here require explicit knowledge of the entire tree \comment{that's not ture.
+We need the hierarchy to speed up point lookup}.
 As a result we introduce a simple algorithm, called \texttt{Reconstruct}, which recreates a local representation of the entire ``forest'' that is only adapted to local cells and, potentially, the ghost layer.
@@
 Algorithm \ref{alg:reconstruction} illustrates how this reconstruction is performed.
-Given a forest and a layer of ghost cells from \texttt{p4est}, the algorithm generates a local representation of forest by recursively refining from the root until reaching the same level and location of all leaves in \texttt{p4est} and \texttt{ghost}.
-Note that algorithm \ref{alg:reconstruction} does not involve any communication and is load balanced.
+Given a forest and a layer of ghost cells from \texttt{p4est}, the algorithm generates a local representation of the forest by recursively refining from the root until reaching the same level and location of all leaves in \texttt{p4est} and \texttt{ghost}.
+Note that algorithm \ref{alg:reconstruction} does not involve any communication and is load balanced \comment{as long as the forest is balanced}.
 Figure \ref{fig:reconstruction} illustrates an application of algorithm \ref{alg:reconstruction}.
@@
 \begin{algorithm}[htbp]
-\caption{$H \gets \texttt{Reconstruct (}G\texttt{)}$}\label{euclid}
+\caption{$H \gets \texttt{Reconstruct (}G\texttt{)}$} %\label{euclid}
 \begin{algorithmic}[1]
@@
 
-\subsection{Semi-Lagrangian method}
+\subsection{Semi-Lagrangian method \comment{and interpolations}}
 As indicated earlier, we use the Semi-Lagrangian method to solve equation \eqref{eq:ls} when the velocity field is externally generated, i.e.
-does not depend on level-set function itself.
+when it does not depend on the level-set function itself.
 Let us rewrite equation \eqref{eq:ls} along the characteristic curve $\mathbf{X}(t)$ as:
@@
 starting from the grid $G^{n+1}$ \footnote{$G^{n+1}$ itself is computed in an iterative fashion as explained later on.}, we simply write $\phi^{n+1}(\mathbf{X}^{n+1}) = \phi(\mathbf{X}(t^{n+1}), t^{n+1}) = \phi(\mathbf{X}(t^n), t^n) = \phi^n(\mathbf{X}_d)$.
-Here the characteristic curves are chosen such that $\mathbf{X}(t^{n+1})$ are simply the coordinates of grids points of $G^{n+1}$.
-Moreover $\mathbf{X}_d$ are the departure points which are computed using the second-order midpoint method \cite{Min;Gibou:07:A-second-order-accur}:
+Here the characteristic curves are chosen such that $\mathbf{X}(t^{n+1})$ are simply the coordinates of grids points of $G^{n+1}$, and $\mathbf{X}_d$ are the departure points which are computed using the second-order midpoint method \cite{Min;Gibou:07:A-second-order-accur}:
 \bea
@@
 
-Note that all values at the intermediate point, $\mathbf{X}^\star$, and departure point, $\mathbf{X}_d$, must be calculated via interpolation from previous grids, $G^{n}$ and $G^{n-1}$.
+Note that all values at the intermediate point, $\mathbf{X}^\star$, and departure point, $\mathbf{X}_d$, must be calculated via interpolation from previous the grids $G^{n}$ and $G^{n-1}$.
 Here we use the stabilized second-order interpolation for $\phi(\mathbf{X}_d)$ and the multi-linear interpolation for $\mathbf{V}^{n+\frac{1}{2}}(\mathbf{X}^\star)$ \cite{Min;Gibou:07:A-second-order-accur}.
 Although parallelization of the interpolation process on a shared-memory machine is trivial, the same cannot be said for distributed-memory machines.
-In fact, the parallel interpolation algorithm \ref{alg:interpolation} is probably the most important contribution of this article.
+In fact, the parallel interpolation algorithm \ref{alg:interpolation} is probably the most important contribution of this article \comment{but hasn't it been done before ?}.
 Complication arises because not all calculated departure points will reside in the domain of current processor.
@@
 Of course if one enforces $\text{CFL} \le 1$, one can ensure that the halo is bounded by the ghost layer which significantly simplifies the communication problem.
-This assumption, however, defeats the purpose of using Semi-Lagrangian in the first place since we are interested in large values of $\text{CFL}$ number.
+This assumption, however, defeats the purpose of using a Semi-Lagrangian approach in the first place since we are interested in large values of $\text{CFL}$ number.
 
 One remedy to this problem, proposed in \cite{Thomas;Cote:95:Massively-parallel-s} for uniform grids, is to increase the size of ghost layer to $\lceil \text{CFL} \rceil$.
-For large values of $\text{CFL}$ number, however, this approach can substantially increase the communication volume.
+For large values of the $\text{CFL}$ number, however, this approach can substantially increase the communication volume.
 Moreover, this simple approach does not work in the process of generating $G^{n+1}$ due to repartitioning.
@@
 Our remote interpolation algorithm is composed of three separate phases.
-In the first phase, which we call buffering, every processor searches for all departure points inside the local tree.
+In the first phase, which we call buffering, every processor searches for all departure points inside the local trees.
 If the point is owned by a local cell, it is added to a local buffer, otherwise we find the processor which owns the point and add the point to a separate buffer belonging to the found rank.
-Note that searching the point in the local tree is performed recursively, similar to the \texttt{UpdateLocalTree} function in algorithm \ref{alg:reconstruction}.
+Note that searching the point in the local tree is performed recursively \comment{using the hierarchal reconstruction}, similarly to the \texttt{UpdateLocalTree} function in algorithm \ref{alg:reconstruction}.
 The owner's rank is found by computing the Z-index of the point and then using binary search on the Z-curve.
@@
 
-Once buffering is done, every processor knows exactly how many messages it needs to send and to which processors.
+Once buffering is done, every processor knows exactly how many messages it needs to send and to which processors \comment{throughout the whole article, shouldn't we talk of ``processes'' ?}.
 This also implicitly defines processors that will later on send a reply message to this processor.
@@
 \end{center}
-\caption{Left: Location of back-traced points depends on the magnitude of local velocity and time-step.
-Although departure distance is bounded by $\text{CFL} \: \Delta x_{\min}$, one cannot make any special assumption about receiving rank without explicit searching of the entire Z-curve.
+\caption{Left: Location of back-traced points depends on the magnitude of the local velocity and on the time-step.
+Although the distance to the departure point is bounded by $\text{CFL} \: \Delta x_{\min}$, one cannot make any special assumption about receiving rank without explicitly searching the entire Z-curve.
 Moreover, the receiving processor has no prior knowledge about which processors to check for incoming messages nor does it know anything about the possible message length (i.e.
 number of points).
-Middle: A directed graph illustrating the communication pattern among processors with arrows representing the direction that messages are sent.
-Right: The adjacency matrix of communication graph.
-For each row, sum of all columns represents the number of sent messages.
-Conversely, for each column, sum of all rows represents number of messages that need to be received.
+Middle: A directed graph illustrating the communication pattern among processors with arrows representing the direction in which messages are sent.
+Right: The adjacency matrix of the communication graph.
+For each row, the sum of all columns represents the number of messages sent.
+Conversely, for each column, the sum of all rows represents the number of messages that need to be received.
 As detailed in algorithm \ref{alg:interpolation} this information is enough to build a parallel interpolation scheme.}
@@
 
-To solve the communication problem, we first compute the adjacency matrix of the communication pattern, i.e. we construct the matrix $A_{P \times P}$ such that:
+To solve the communication problem, we first compute the adjacency matrix of the communication pattern, i.e. we construct the matrix $A_{P \times P}$, where $P$ is the number of processors, such that
 \ben
@@
 \een
-Note that this matrix is also distributed among processors, i.e. each row is owned by a separate rank. Next, we compute:
+Note that this matrix is also distributed among processors, i.e. each row is owned by a separate rank. Next, we compute
 \ben
@@
 where $S_i$ and $R_i$ denote the number of sent and received messages respectively.
-While $S_i$ can be computed trivially, a reduction operation is required for computing $R_i$.
+While $S_i$ can be computed trivially, a reduction operation is required to compute $R_i$.
 For instance, this can be achieved using a single \texttt{MPI\_Reduce\_scatter} function call.
-The last phase of interpolation involves overlapping the computation of interpolated values for local points with the communication of data between processors.
+The last phase of the interpolation procedure involves overlapping the computation of interpolated values for local points with the communication of data between processors.
 This is done by alternating between local calculations and probing for incoming messages from other processors.
-Interpolation is finished once values for all local points have been calculated and all remote requests have been processed (see algorithm \ref{alg:interpolation}).
+The interpolation is finished once the values for all local the points have been calculated and all the remote requests have been processed (see algorithm \ref{alg:interpolation}).
 \begin{algorithm}[htbp]
-\caption{$values \gets \texttt{Interpolate (}H, \mathbf{X}\texttt{)}$}\label{euclid}
+\caption{$values \gets \texttt{Interpolate (}H, \mathbf{X}\texttt{)}$} %\label{euclid}
 \begin{algorithmic}[1]
@@
 \For {$r: mpisize$} \Comment{Phase II -- initiate communication and compute number of messages}
-	\If {$col[r]$}
+	\If {$\textcolor{red}{r!=mpirank~and~} col[r]$}
 		\State $\texttt{MP\_Isend(}r, buff[r]\texttt{)}$
@@
 Using the interpolation algorithm \ref{alg:interpolation}, we close this section by presenting the final Semi-Lagrangian algorithm \ref{alg:semi-lagrangian}.
-The basic idea here is to start from an initial guess $G^{n+1}_0$ and modify the grid using refinement and coarsening criteria in \eqref{eq:refine} and \eqref{eq:coarsen} until convergence is obtained.
+The basic idea here is to start from an initial guess $G^{n+1}_0$ and modify the grid using the refinement \eqref{eq:refine} and coarsening \eqref{eq:coarsen} criteria until convergence is obtained.
 Various options are available for $G^{n+1}_0$.
 For instance it is possible to start from the macromesh and only perform refinement steps until convergence.
-This choice, however, is not suitable since first few iterations do not contain many cells and there is little work for parallelism.
+This choice, however, is not suitable since the first few iterations do not contain many cells and there is little work for parallelism.
 Here we simply take the previous grid as the starting point, i.e.
 $G^{n+1}_0 = G^n$.
-Note that this iterative process is essentially unavoidable since the grid is based on the value of the level-set function at $t^{n+1}$, which itself is unknown and is to be defined on $G^{n+1}$.
-Nonetheless the process converges to the final grid in at most $l_{\max}$ steps where $l_{\max}$ denotes the maximum final depth of trees across all processors.
+Note that this iterative process is essentially unavoidable since the grid is based on the values of the level-set function at $t^{n+1}$, which itself is unknown and is to be defined on $G^{n+1}$.
+Nonetheless the process converges to the final grid in at most $l_{\max}$ steps \comment{shouldn't it be $l_{max}-l_{min}$ ?} where $l_{\max}$ denotes the maximum final depth of the trees across all processors.
 \begin{algorithm}[htbp]
-\caption{$[G^{n+1}, \phi^{n+1}] \gets \texttt{SemiLagrangian (}G^n, \phi^n, \mathbf{V}^n, \mathbf{V}^{n-1}, \text{CFL}\texttt{)}$}\label{euclid}
+\caption{$[G^{n+1}, \phi^{n+1}] \gets \texttt{SemiLagrangian (}G^n, \phi^n, \mathbf{V}^n, \mathbf{V}^{n-1}, \text{CFL}\texttt{)}$} %\label{euclid}
 \begin{algorithmic}[1]
-\State $\Delta t_l \gets \text{CFL} \times \max \{\mathbf{V}^n\} / G^n.\texttt{hmin()}$
+\State $\Delta t_l \gets \text{CFL} \times G^n.\texttt{hmin()} / \max \{\mathbf{V}^n \} $
 \State $\Delta t   \gets \texttt{MPI\_Allreduce(}\Delta t_l,\texttt{MPI\_MIN)}$
@@
 \subsection{Reinitialization}
-Successive application of algorithm \ref{alg:semi-lagrangian}, especially for large values of CFL number, potentially lead degrades the signed distance property of level-set function.
-It is thus important to reinitialize the level-set every few iterations, especially because the quality of generated grid heavily depends on the signed distance property.
+Successive application of algorithm \ref{alg:semi-lagrangian}, especially for large values of the CFL number, can potentially degrades the signed distance property of the level-set function.
+It is thus important to reinitialize the level-set function every few iterations, especially because the quality of grid generated depends heavily on the signed distance property.
 To achieve this property we solve the pseudo-time transient equation \eqref{eq:reinitialization} using the discretization scheme detailed in \cite{Min;Gibou:07:A-second-order-accur}.
-For completeness, we briefly review the scheme here.
-First equation \eqref{eq:reinitialization} is written in the following semi-discrete form:
+For completeness, we briefly review the scheme.
+First, we write equation \eqref{eq:reinitialization} in the following semi-discrete form:
 \be
@@
 where $a^+ = \max(a, 0)$ and $a^- = \min(a, 0)$.
-Similar to \cite{Min;Gibou:07:A-second-order-accur}, equation \eqref{eq:semidiscrete_reinit} is integrated in time using the TVD-RK2 scheme using adaptive time-stepping to accelerate convergence to steady state.
-Since all computation is grid based, parallel implementation of this scheme is mostly trivial.
+Similar to \cite{Min;Gibou:07:A-second-order-accur}, equation \eqref{eq:semidiscrete_reinit} is integrated in time using the TVD-RK2 scheme and with adaptive time-stepping in order to accelerate the convergence to the steady state.
+Since all the computation is based on a local stencil, the parallel implementation of this scheme is mostly trivial.
 However, one minor point requires further explanation.
-As suggested in \cite{Min;Gibou:07:A-second-order-accur}, one-sided derivatives $D^+_i \phi$ and $D^-_i \phi$ are computed using second order discretizations which require computation of second derivatives.
-To enable overlap between computation and communications when computing second derivatives and also integrating \eqref{eq:semidiscrete_reinit}, we use the following common technique.
+As suggested in \cite{Min;Gibou:07:A-second-order-accur}, one-sided derivatives $D^+_i \phi$ and $D^-_i \phi$ are computed using second order discretizations which require to compute the second derivatives.
+To enable overlap between computation and communications when computing second order derivatives and also integrating \eqref{eq:semidiscrete_reinit}, we use the following common technique.
 First, we label all local points, $L_p$, as either private, $P_p$, or boundary, $B_p$.
-Here, boundary points are the collection of all local points that are regarded as ghost point, $G_r$, on all other processors, i.e.
+Here, the boundary points are the collection of all local points that are regarded as a ghost point, $G_r$, on at least one other processor, i.e.
 $B_p = \underset{r,\;r\neq p}{\bigcup} G_r$.
-Private points are fined as the collection of all local points that are not a boundary point, i.e.
+Private points are defined as the collection of all local points that are not a boundary point, i.e.
 $P_p = L_p \setminus B_p$.
-Algorithm \ref{alg:overlap} illustrates how this labeling can help with computation/computation overlap of an arbitrary local operation $\phi^{n+1} \gets \mathcal{F}(\phi^n)$.
-Note that \texttt{p4est} library already includes all the primitives required for labeling local points without any further communication.
+Algorithm \ref{alg:overlap} illustrates how this labelling can help with overlapping the computation and the communication associated to an arbitrary local operation $\phi^{n+1} \gets \mathcal{F}(\phi^n)$.
+Note that \texttt{p4est} library already includes all the primitives required for labelling local points without any further communication.
 
 \begin{algorithm}[htbp]
-\caption{$\phi^{n+1} \gets \texttt{Overlap (}\phi^n, \mathcal{F}\texttt{)}$}\label{euclid}
+\caption{$\phi^{n+1} \gets \texttt{Overlap (}\phi^n, \mathcal{F}\texttt{)}$} %\label{euclid}
 \begin{algorithmic}[1]

### 2015-01-07 paper/tex/scaling_results.tex
added scaling figures for stefan, starting to work on that part
finished first pass

@@
 \section{Scaling} \label{sec:scaling}
-In this section we present results that show the scalability of our algorithms.
-All of our tests have been run on the Stampede cluster at Texas Advanced Computing Center (TACC) where we are limited to $4096$ cores at most.
+In this section we present some results that demonstrate the scalability of our algorithms.
+All of our tests were ran on the Stampede cluster at Texas Advanced Computing Center (TACC) where we are limited to $4096$ cores at most.
 Each node of Stampede has 2 eight-core Xenon E5-2680 processors clocked at 2.7 GHz with 32 GB of DDR3-1600 MHz memory and interconnected using an InfiniBand network card.
@@
 \subsection{Interpolation}
-In this section we show results for a simple test to measure the scalability of the interpolation algorithm \ref{alg:interpolation}.
-The test consists of interpolating a function at a number of random points on a randomly refined octree in three dimensions.
-We consider two cases.
-A small test on a level 9 tree with roughtly 20M cells and 33M nodes and larger test on a level 13 tree with roughly 128M cells and 280M nodes.
-In both cases number of randomly generated points are chosen to be equal to the number of nodes and the stablized second-order interpolation of \cite{Min;Gibou:07:A-second-order-accur} was performed 10 times to smooth out possible timing fluctuations.
+In this section we show the results for a simple test to measure the scalability of the interpolation algorithm \ref{alg:interpolation}.
+The test consists in interpolating a function at a number of random points on a randomly refined Octree in three spatial dimensions.
+We consider two cases, a small test on a level 9 tree with roughly 20M cells and 33M nodes and larger test on a level 13 tree with roughly 128M cells and 280M nodes.
+In both cases the number of randomly generated points is chosen to be equal to the number of nodes and the stablized second-order interpolation of \cite{Min;Gibou:07:A-second-order-accur} is performed 10 times to smooth out possible timing fluctuations.
 
-To simulate the effect of different CFL numbers, we generate the random points such that on each processor $\alpha$ percentage of them are located outside the processor boundary and thus will initiate communication.
-Scaling results are presented for $\alpha = 5 \%$ and $\alpha = 95\%$ for both small and large problems in figure \ref{fig:interpolation}.
-Excellent scaling is obtained for the small problem for $P = 16-512$ even when $95\%$ of interpolation points are remote to processors.
-For the larger problem, communication overhead prevents further scaling of algorithm beyond $4096$ processors when $\alpha = 95\%$.
-Note that for $\alpha = 5\%$, even though some sections of algorithm stop scaling, the overall algorithm still scales since the local work dominates the timing.
-This illustrates the effectiveness of non-blocking communication pattern in algorithm \ref{alg:interpolation}.
+To simulate the effect of different CFL numbers, we generate the random points such that on each processor $\alpha\%$ of them are located outside the processor boundary and will thus initiate communication.
+Scaling results are presented for $\alpha = 5 \%$ and $\alpha = 95\%$ for both the small and large problems in figure \ref{fig:interpolation}.
+Excellent scaling is obtained for the small problem for $P = 16-512$ even when $95\%$ of the interpolation points belong to a remote processor.
+For the larger problem, the communication overhead prevents the algorithm from scaling beyond $4096$ processors when $\alpha = 95\%$.
+Note that for $\alpha = 5\%$, even though some sections of the algorithm stop scaling, the overall algorithm still scales since the local work dominates the timing.
+This illustrates the effectiveness of the non-blocking communication pattern of algorithm \ref{alg:interpolation}.
 
@@
 	\end{center}
-	\caption{Strong scaling of algorithm \ref{alg:interpolation} for several tests where $N_G$ denotes the number of random interpolation points (same as number of nodes in the octree) and $\alpha$ denotes the percentage of these points that are remote to each processor.
+	\caption{Strong scaling of algorithm \ref{alg:interpolation} for several tests where $N_G$ denotes the number of random interpolation points (which is the same as the number of nodes in the Octree) and $\alpha$ denotes the percentage of these points that are remote for each processor.
 Here ``Total'' represents the total time spent in the interpolation while ``buffer'', ``local'', ``queries'', and ``replies'' represent the the timing for different sections (c.f.
 algorithm \ref{alg:interpolation}).
-Results indicate excellent scaling for the small test (a-b) and for the large test when $\alpha = 5\%$ (c).
+The results indicate excellent scaling for the small test (a-b) and for the large test when $\alpha = 5\%$ (c).
 For the extreme case (d) the algorithm stops scaling at $4096$ processors due to communication overhead.}
@@
 To test the scalability of the Semi-Lagrangian algorithm \ref{alg:semi-lagrangian} we consider a slightly modified version of the Enright's rotation test \cite{Enright;Fedkiw;Ferziger;etal:02:A-Hybrid-Particle-Le}, i.e.
-we advect a sphere of radius 0.35 located at $(0.4, 0.4, 0.4)$ with a divergent free velocity field given by:
+we advect a sphere of radius 0.35 located at $(0.4, 0.4, 0.4)$ with a divergence free velocity field given by:
 \bean
@@
 \eean
-To understand the effect of CFLnumber on the scalability of the algorithm we perform one step of the Semi-Lagrangian algorithm for $\text{CFL} = 1$, $\text{CFL} = 10$, and $\text{CFL} = 100$.
-We also perform the test for two set of initial girds, a small grid with $l_\text{max} = 10$ and a larger grid with $l_\text{max} = 12$.
-After one advection step, each of these grids have approximately 15M and 255M nodes, respectively.
+To understand the effect of the CFL number on the scalability of the algorithm we perform one step of the Semi-Lagrangian algorithm for $\text{CFL} = 1$, $\text{CFL} = 10$, and $\text{CFL} = 100$.
+We also perform the test for two different initial girds, a small grid with maximum level $l_\text{max} = 10$ and a large grid with maximum level $l_\text{max} = 12$.
+After one advection step, these grids have approximately 15M and 255M nodes, respectively.
 \begin{table}[htbp]
@@
 	\end{minipage}	
-	\caption{Number of iterations required for grid construction in algorithm \ref{alg:semi-lagrangian} for the rotation test on a (a) level-10 and (b) level-12 octree with approximately 15M and 255M nodes, respectively.
-Note how the iteration count increases with the CFL number but is almost independent of number of processors.
-The slight dependence of iteration count on number of processors is most likely due to dependence of round-off errors on the number of processors.
-Nonetheless, close examination of generated octrees (data not shown) reveals that they are identical and independent of the number of processors used to perform the test.}
+	\caption{Number of iterations required for the grid construction in algorithm \ref{alg:semi-lagrangian} for the rotation test on a (a) level-10 and (b) level-12 Octree with approximately 15M and 255M nodes, respectively.
+Note how the iteration count increases with the CFL number but is almost independent of the number of processors.
+The slight dependence between the number of iterations and the number of processors is most likely due to the dependence of round-off errors on the number of processors.
+Nonetheless, close examination of the Octrees generated (data not shown) reveals that they are identical and independent of the number of processors used to perform the test.}
 	\label{tab:semilagrangian}
@@
 
-Table \ref{tab:semilagrangian} illustrates the dependence of number of iterations required to build the grid; as the CFL is increased, interface travels a farther distance which necessitates more iterations to generate the grid.
-Note that the apparent dependence on the number processors is most likely due to round-off errors during the interpolation and/or backtracking step.
-Figures \ref{fig:semilagrangian_small} and \ref{fig:semilagrangian_large} illustrate the scalability of the algorithm for small and large problems, respectively.
-To enable meaningful comparisons between different CFL numbers and number of processors, the maximum time has been scaled to the number of iterations required for grid construction as reported in table \ref{tab:semilagrangian}.
-For both problems excellent scalability is observed for $\text{CFL} = 1$ and $\text{CFL} = 10$.
+Table \ref{tab:semilagrangian} illustrates the dependence of the number of iterations required to build the grid on the CFL number; as the CFL is increased, the interface travels a farther distance which necessitates more iterations to generate the grid.
+Note that the apparent dependence on the number of processors is most likely due to round-off errors during the interpolation and/or backtracking steps.
+Figures \ref{fig:semilagrangian_small} and \ref{fig:semilagrangian_large} illustrate the scalability of the algorithm for the small and large problems, respectively.
+To enable meaningful comparisons between different CFL numbers and number of processors, the maximum time has been scaled by the number of iterations required for the grid construction as reported in table \ref{tab:semilagrangian}.
+For both problems, excellent scalability is observed for $\text{CFL} = 1$ and $\text{CFL} = 10$.
 The algorithm even shows good scalability when taken to the extreme, i.e.
@@
 An increase in the CFL number has two effects on the algorithm.
-First, a larger fraction of departure points land in domains of remote processors.
+First, a larger fraction of the departure points lands in the domains of remote processors.
 Moreover, these points are potentially dispersed across a larger number of processors.
-This means both the number MPI messages and their volume increases for larger CFL numbers.
-Second, as more points are shipped to remote processors for interpolation, there is a greater chance that interpolation load is unbalanced across processors.
-This is specially true for regions of space in which streamlines cluster.
-Both factors contribute to decrease the scalability of the algorithm at larger CFL numbers.
+This means that both the number of MPI messages and their volume increase as the CFL number increases.
+Second, as more points are shipped to remote processors for interpolation, there is a greater chance that the interpolation load is imbalanced across processors.
+This is especially true for regions of space in which the streamlines cluster.
+Both factors contribute to decrease the scalability of the algorithm at large CFL numbers.
 \begin{figure}[htbp]
@@
 	\end{center}
-	\caption{Strong scaling of a single iteration of algorithm \ref{alg:semi-lagrangian} for the rotation test on a level-10 octree with approximately 13M cells and 15M nodes.
-Top row: scaling of various components of the algorithm for (a) $\text{CFL} = 1$, (b) $\text{CFL} = 10$, and (c) $\text{CFL} = 100$.
-Bottom row: breakdown of various components of the interpolation phase for the same CFL numbers.
-Note that the maximum time has been scaled with the number of iterations required to build the tree (c.f.
+	\caption{Strong scaling of a single iteration of algorithm \ref{alg:semi-lagrangian} for the rotation test on a level-10 Octree with approximately 13M cells and 15M nodes.
+Top row: scaling of the various components of the algorithm for (a) $\text{CFL} = 1$, (b) $\text{CFL} = 10$, and (c) $\text{CFL} = 100$.
+Bottom row: breakdown of the various components of the interpolation phase for the same CFL numbers.
+Note that the maximum time has been scaled by the number of iterations required to build the tree (c.f.
 table \ref{tab:semilagrangian}).}
@@
 	\end{center}
-	\caption{Strong scaling of a single iteration of algorithm \ref{alg:semi-lagrangian} for the rotation test on a level-12 octree with approximately 251M cells and 255M nodes.
-Top row: scaling of various components of the algorithm for (a) $\text{CFL} = 1$, (b) $\text{CFL} = 10$, and (c) $\text{CFL} = 100$.
-Bottom row: breakdown of various components of the interpolation phase for the same CFL numbers.
-Note that the maximum time has been scaled with the number of iterations required to build the tree (c.f.
+	\caption{Strong scaling of a single iteration of algorithm \ref{alg:semi-lagrangian} for the rotation test on a level-12 Octree with approximately 251M cells and 255M nodes.
+Top row: scaling of the various components of the algorithm for (a) $\text{CFL} = 1$, (b) $\text{CFL} = 10$, and (c) $\text{CFL} = 100$.
+Bottom row: breakdown of the various components of the interpolation phase for the same CFL numbers.
+Note that the maximum time has been scaled by the number of iterations required to build the tree (c.f.
 table \ref{tab:semilagrangian}).}
@@
 
-To better understand the importance of CFL number on the scalability, we have recorded a complete history of communication pattern in the interpolation step.
-Figure \ref{fig:communication_4096} illustrates the effects of the CFL number on three different metrics, namely number of interpolation points\footnote{Note that this includes both local points and those queried by other processors}, $N_p$, number of sent and received messages, $N_m = S + R$, and the total communication volume, $V_m$ in mega bytes (MB), for $p=4096$ processors.
-Moreover, these values are reported for the first (top row) and last (bottom row) iterations of the Semi-Lagrangian algorithm.
+To better understand the importance of the CFL number on the scalability, we have recorded a complete history of the communication pattern in the interpolation step.
+Figure \ref{fig:communication_4096} illustrates the effects of the CFL number on three different metrics, namely the number of interpolation points\footnote{Note that this includes both the local points and the points queried by other processors}, $N_p$, the number of sent and received messages, $N_m = S + R$, and the total communication volume, $V_m$ in mega bytes (MB), for $p=4096$ processors.
+Furthermore, these values are reported for the first (top row) and last (bottom row) iterations of the Semi-Lagrangian algorithm.
 There are several points to make.
-First, increasing the CFL number greatly increases the load imbalance as evident in the spread of data in figure \ref{fig:communication_4096_pf}.
-This is because at higher CFL numbers, it is more likely that some processors receive larger portions of backtracked points.
+First, increasing the CFL number greatly increases the load imbalance, as shown by the spread of the data in figure \ref{fig:communication_4096_pf}.
+This is because at higher CFL numbers, and with the velocity field prescribed, it is more likely that some processors will receive a larger portion of the backtracked points.
 Second, increasing the CFL number increases both the communication volume and its spread across processors (c.f.
-figure \ref{fig:communication_4096_vf}.
-Interestingly, however, the number of sent and received messages seems to remain unaffected by the CFL number.
-The bottom row of figure \ref{fig:communication_4096} illustrates better balance both in the computation and communication volume in the last iteration of the Semi-Lagrangian algorithm.
-This can be explained by noting that as the algorithm converges to the final grid, partitioning of the grid essentially becomes equivalent to partitioning of the backtracked points.
-Detailed information about the load balancing and communication pattern is listed in table \ref{tab:communication}.
+figure \ref{fig:communication_4096_vf}).
+Interestingly, however, the number of messages sent and received seems to be unaffected by the CFL number.
+The bottom row of figure \ref{fig:communication_4096} exhibits a better balance both in the computation and communication volume in the last iteration of the Semi-Lagrangian algorithm.
+This can be explained by noting that as the algorithm converges to the final grid, partitioning of the grid essentially becomes equivalent to partitioning of the backtracked points \comment{awkward ...
+maybe: the partitioning of the the grids $G^n$ and $G^{n+1}$ become very similar ?}.
+Detailed information about the load balancing and the communication patterns is listed in table \ref{tab:communication}.
 \begin{figure}[htbp]
@@
 	\end{center}
-	\caption{Performance indicators of the first (top row) and last (bottom row) iterations of the Semi-Lagrangian algorithm for the advection on 4096 processors for $\text{CFL} = 1$ (\drawdiamond{black}), $\text{CFL} = 10$ (\drawsquare{red}), and $\text{CFL} = 100$ (\drawcircle{blue}).
-Increasing the CFL number greatly causes load imbalance in the amount of computation done during interpolation (a) and increases the communication volume (c).
-However the CFL number does not seem to appreciably affect the number of messages sent by processors (b).
-During the last Semi-Lagrangian iteration the initial grid, $G_0$ (c.f.
-algorithm \ref{alg:semi-lagrangian}), is very close to the final grid.
-As a result load imbalance is considerably improved (d).
-Curiously, however, communication pattern does not seem to be affected by much between first and last iterations (e-f).}
+	\caption{Performance indicators of the first (top row) and last (bottom row) iterations of the Semi-Lagrangian algorithm for the advection on 4096 processors with $\text{CFL} = 1$ (\drawdiamond{black} \comment{this does not work on my computer, I see a W}), $\text{CFL} = 10$ (\drawsquare{red}), and $\text{CFL} = 100$ (\drawcircle{blue}).
+Increasing the CFL number greatly causes load imbalance during interpolation (a) and increases the communication volume (c).
+However, the CFL number does not seem to appreciably affect the number of messages sent by the processors (b).
+During the last Semi-Lagrangian iteration, the initial grid $G_0$ (c.f.
+algorithm \ref{alg:semi-lagrangian}) is very close to the final grid.
+As a result, the load imbalance is considerably improved (d).
+Curiously, however, the communication pattern does not seem to be change much between first and last iterations (e-f) \comment{e-f ?}.}
 	\label{fig:communication_4096}
@@
 										   & $V_m$ (MB)   & 4.01E-01 & 2.93E+00 & 7.25E-01 & 1.95E-01 & 4.13E-01 & 3.91E+00 & 9.68E-01 & 2.62E-01 \\
-										   & $T_\max$ (s) & \multicolumn{4}{|c|}{6.96E-01}						& \multicolumn{4}{|c|}{3.67E-01}						\\
+										   & $T_{\max}$ (s) & \multicolumn{4}{|c|}{6.96E-01}						& \multicolumn{4}{|c|}{3.67E-01}						\\
 	\hline
@@
 										   & $V_m$ (MB)   & 4.01E-01 & 3.06E+00 & 8.40E-01 & 2.21E-01 & 4.40E-01 & 4.80E+00 & 2.14E+00 & 9.88E-01 \\
-											 & $T_\max$ (s) & \multicolumn{4}{|c|}{7.55E-01}						& \multicolumn{4}{|c|}{3.30E-01}						\\
+											 & $T_{\max}$ (s) & \multicolumn{4}{|c|}{7.55E-01}						& \multicolumn{4}{|c|}{3.30E-01}						\\
 	\hline
@@
 										   & $V_m$ (MB)   & 5.28E-01 & 1.01E+01 & 2.65E+00 & 9.24E-01 & 9.76E-01 & 4.86E+00 & 3.78E+00 & 5.69E-01 \\
-											 & $T_\max$ (s) & \multicolumn{4}{|c|}{9.25E-01}						& \multicolumn{4}{|c|}{3.55E-01}						\\
+											 & $T_{\max}$ (s) & \multicolumn{4}{|c|}{9.25E-01}						& \multicolumn{4}{|c|}{3.55E-01}						\\
 	\hline
@@
 	\caption{Detailed load balancing and communication information for the advection test problem for $\text{CFL} = 1$, $\text{CFL} = 10$, and $\text{CFL} = 100$.
-Note how increasing the CFL number causes load imbalance and increases the communication volume while it does not affect the number of sent and received messages during an iteration of Semi-Lagrangian step.}
+Note how increasing the CFL number causes load imbalance and increases the communication volume while it does not affect the number of messages sent and received during an iteration of the Semi-Lagrangian step.}
 	\label{tab:communication}
@@
 
-Finally we present scaling results of our parallel reinitialization algorithm where we extensively make use of algorithm \ref{alg:overlap} for overlapping computation with communication when computing spatial derivatives.
-Our test is consisted of computing the signed distance to a collection of 100 sphere, whose radii and centers have been chosen randomly.
-Test is perform on a small, level-8 octree with nearly 21M and a larger, level-10 octree with nearly 337M grid points.
+Finally we present the scaling results of our parallel reinitialization algorithm where we extensively make use of algorithm \ref{alg:overlap} for overlapping computation with communication when computing spatial derivatives.
+Our test consists in computing the signed distance to a collection of 100 spheres, whose radii and centers have been chosen randomly.
+The test is performed on a small, level-8 Octree with nearly 21M and a larger, level-10 Octree with nearly 337M grid points.
 In both cases the forest is built on a $3\times3\times3$ macro-mesh.
-Figures \ref{fig:reinit} illustrates that our reinitialization algorithm and, in particular the overlapping strategy presented in algorithm \ref{alg:overlap}, scales very well.
-In general we expect similar scaling results for any local, finite-difference based calculations on octrees that can efficiently utilize algorithm \ref{alg:overlap}.
+Figure \ref{fig:reinit} illustrates that our reinitialization algorithm, and in particular the overlapping strategy presented in algorithm \ref{alg:overlap}, scales very well.
+In general we expect similar scaling results for any local, finite-difference based calculations on Octrees that can efficiently utilize algorithm \ref{alg:overlap}.
 \begin{figure}
@@
 \subfigure[]{\includegraphics[width=0.45\textwidth]{figures/Reinit_Large.pdf}}
-\caption{Scalability of the reinitialization test for a small (left) and large octrees with roughtly 21M and 337M grid points, respectively.
+\caption{Scalability of the reinitialization test for a small (left) and large (right) Octree with roughly 21M and 337M grid points, respectively.
 Excellent results are obtained in both cases, illustrating the scalability of the overlapping strategy (c.f.

### 2015-01-07 paper/tex/abstract.tex
Added abstract

@@
 \begin{abstract}
-Abstract will be filled in after finalization of the paper.
+In this article we present scalable parallel algorithms for the advection and reinitialization of level-set functions on adaptive Quadtree and Octree grids.
+Our algorithms are based on domain decomposition technique and implemented using \texttt{MPI} and the open-source \texttt{p4est} library.
+An important contribution of the paper is a parallel Semi-Lagrangian method which, similar to its serial implementation, is free of any time-step restriction due to CFL condition.
+This is achieved by introducing a scalable parallel interpolation scheme on adaptive Quadtree and Octree grids.
+Moreover, we present a simple parallel reinitialization scheme using the pseudo-time transient formulation.
+Both parallel algorithms scale nicely on the Stampede supercomputer where we are limited to 4096 cores at most.
+Finally a relevant application of the algorithms is presented in modeling the crystallization phenomena by solving a Stefan problem, illustrating a level of detailed calculation that would be impossible on uniform grids.
+We believe that the algorithms presented in this article will be of interest and useful to researchers working with the level-set framework and modeling multi-scale physics in general.
 \end{abstract}
-
 \begin{keyword}
-Level-set Method \sep Adaptive Grids \sep Parallel Computing \sep Oct-/Quadtree datastructures
 \end{keyword}

### 2015-01-08 paper/tex/conclusion.tex
Added conclusion and acknowledgment

%!TEX root = draft.tex
\section{Conclusions}
In this article we have presented parallel algorithms for the advection and reinitialization of the level-set functions on adaptive Quadtree and Octree grids using domain decomposition approach. These algorithms are implemented using a combination of \texttt{MPI} standard and the open-source \texttt{p4est} library. An important feature of the Semi-Lagrangian method is its unconditional stability property which must be preserved in parallel, requiring a parallel interpolation scheme. However, due to the non-uniform distribution of departure points, scalable implementation of such algorithm is not trivial on adaptive Quadtree and Octree grids. An asynchronous interpolation algorithm is presented using non-blocking point-to-point communications which proves good scalability.

The scalability of the Semi-Lagrangian algorithm, however, depends on the CFL number. Great scalability is observed for intermediate CFL numbers while, e.g. $\text{CFL}~10$. At higher CFL numbers, however, the departure points are potentially further dispersed across processors which limits the scalability. This is due to the fact the domain decomposition technique used here is based on the Z-ordering of quadrants and does not take the velocity field information into account. A possible remedy for this problem could be assigning weights to quadrants based on some estimate of what the grid should look like after one step of the advection algorithm, e.g. by using a forward-in-time integration of grid points. Such an estimate could also reduce the number of Semi-Lagrangian iterations. These ideas are postponed for further investigations. We have also presented a simple parallelization technique for the reinitialization algorithm based on the pseudo-time transient formulation. Both the Semi-Lagrangian and the reinitialization algorithms show great scalability up to 4096 processors.

Finally, an application of these algorithms is presented in modeling the solidification process by solving the Stefan problem. This application clearly illustrates the applicability of our algorithms to complex multi-scale problems that cannot be treated using the normal domain decomposition techniques on uniform grids. We believe that our findings would be of interest to other researchers using the level-set framework for other complex and multi-scale problems.

\section*{Acknowledgment} 
The funding for this project was provided by \comment{the funding agency}. We would also like to thank the technical staff at the Texas Advanced Computing Center (TACC) and the developer community of PETSc library for their valuable suggestions.

### 2015-01-08 paper/tex/abstract.tex
proof-read up to sec. 3.2

@@
 Our algorithms are based on domain decomposition technique and implemented using \texttt{MPI} and the open-source \texttt{p4est} library.
-An important contribution of the paper is a parallel Semi-Lagrangian method which, similar to its serial implementation, is free of any time-step restriction due to CFL condition.
-This is achieved by introducing a scalable parallel interpolation scheme on adaptive Quadtree and Octree grids.
+An important contribution of the paper is a parallel semi-Lagrangian method which, similar to its serial implementation, is free of any time-step restrictions.
+This is achieved by introducing a scalable global interpolation scheme on adaptive Quadtree and Octree grids.
 Moreover, we present a simple parallel reinitialization scheme using the pseudo-time transient formulation.
@@
 \begin{keyword}
+Quadtree/Octree Grids \sep Parallel Computing \sep Space Filling Curves \sep Semi-Lagrangian Method \sep Level-set Method
 \end{keyword}
-
 \maketitle

### 2015-01-08 paper/tex/introduction.tex
proof-read up to sec. 3.2

@@
 The level-set method, originally proposed by Sethian and Osher \cite{Osher;Sethian:88:Fronts-Propagating-w}, is a popular and powerful framework for tracking arbitrary interfaces that undergo complicated topological changes.
-As a result, the level-set method has wide range of application such as multiphase flows, moving boundary problems, image segmentation, and computer graphics \cite{Osher;Fedkiw:01:Level-Set-Methods:-A,Sethian:99:Level-set-methods-an}.
-An important feature which makes this method powerful and easy to use is that the location of interface is defined implicitly on an underlying grid.
+As a result, the level-set method has wide range of application such as in multiphase flows, moving boundary problems, image segmentation, and computer graphics \cite{Osher;Fedkiw:01:Level-Set-Methods:-A,Sethian:99:Level-set-methods-an}.
+An important feature of this method is that the location of interface is defined implicitly on an underlying grid.
 This convenience, however, comes at a price.
-First, compared to an explicit method, e.g.
-front tracking \cite{Juric:96:A-Front-Tracking-Met, Tryggvason;Bunner;Esmaeeli;etal:01:A-Front-Tracking-Met}, the level-set method is typically less accurate and mass conservation could be a problem; although progress has been made in resolving this issue \cite{Enright;Fedkiw;Ferziger;etal:02:A-Hybrid-Particle-Le}.
+First, compared to an explicit method, e.g.\ front tracking \cite{Juric:96:A-Front-Tracking-Met, Tryggvason;Bunner;Esmaeeli;etal:01:A-Front-Tracking-Met}, the level-set method is typically less accurate and mass conservation could be a problem; although progress has been made in resolving this issue \cite{Enright;Fedkiw;Ferziger;etal:02:A-Hybrid-Particle-Le}.
 Second, the level-set function has to be defined in a higher dimensional space compared to the interface.
 If only the location of the interface is needed, the added dimension greatly increases the overall computational cost.
-One way to avoid this problem is by computing the level-set only close to the interface, e.g.
-as in the narrow-band level-set method \cite{Adalsteinsson;Sethian:95:A-Fast-Level-Set-Met} or, more recently, by using a hash table to restrict both computation and storage requirements \cite{Brun;Guittet;Gibou:12:A-local-level-set-me}.
+One way to avoid this problem is by computing the level-set only close to the interface, e.g.\ as in the narrow-band level-set method \cite{Adalsteinsson;Sethian:95:A-Fast-Level-Set-Met} or, more recently, by using a hash table to restrict both computation and storage requirements \cite{Brun;Guittet;Gibou:12:A-local-level-set-me}.
 
@@
 More recently, authors in \cite{Min;Gibou:07:A-second-order-accur} proposed second-order accurate level-set methods on Quadtree (two spatial dimensions) and Octree (three spatial dimensions) grids.
-The use of tree-based adaptive grids in the context of the level-set method is quite advantageous because: 1) It gives fine-grain control over errors, which typically occur close to the interface and 2) It can effectively reduce the dimensionality of the problem by focusing most of the grid cells close to the interface.
+The use of adaptive tree-base grids in the context of the level-set method is quite advantageous because: 1) It gives fine-grain control over errors, which typically occur close to the interface and 2) It can effectively reduce the dimensionality of the problem by focusing most of the grid cells close to the interface.
 Fortunately, constructing the tree is quite simple in the presence of an interface that naturally defines an ideal metric for refinement.
-Although the use of adaptive grids can dramatically reduce the computational cost, performing high-resolution three dimensional calculations of complex interfacial problems, e.g.
-crystal growth in binary alloys \cite{Theillard;Gibou;Pollock:14:A-Sharp-Computationa}, could take a very long time in serial.
+Although the use of adaptive grids can dramatically reduce the computational cost, performing high-resolution three dimensional calculations of complex interfacial problems, e.g.\ crystal growth in binary alloys \cite{Theillard;Gibou;Pollock:14:A-Sharp-Computationa}, could take a very long time in serial.
 In this paper we extend these algorithms by proposing parallel algorithms for distributed memory machines using a domain decomposition technique.
@@
 One of the main challenges in parallelizing level-set algorithms on adaptive grids is handling the grid itself.
-One option is to replicate the entire grid on each processor and use off-the-shelf graph partitioners, e.g.
-ParMetis \cite{Karypis;Kumar:98:A-parallel-algorithm} or Zoltan \cite{Boman;Catalyurek;Chevalier;etal:12:The-Zoltan-and-Isorr}, for load balancing and domain decomposition.
+One option is to replicate the entire grid on each process and use off-the-shelf graph partitioners, e.g.\ ParMetis \cite{Karypis;Kumar:98:A-parallel-algorithm} or Zoltan \cite{Boman;Catalyurek;Chevalier;etal:12:The-Zoltan-and-Isorr}, for load balancing and domain decomposition.
 For instance, this was the approach originally taken by the \texttt{deal.II} library \cite{Bangerth;Hartmann;Kanschat:07:deal.II----a-General}.
-This approach, however, is only scalable to a few hundred processors at best and is limited by the size of the grid itself.
+This approach, however, is only scalable to a few hundred processes at best and is limited by the size of the grid itself.
 Moreover, the use of a general-purpose graph partitioner adds extra overhead that can limit the overall scalability even further.
@@
 For instance, \texttt{Octor} \cite{Tu;OHallaron;Ghattas:05:Scalable-parallel-oc} uses a Morton curve (also known as Z-curve) for traversing the leaves of an Octree for indexing and load balancing and has been scaled up to 62,000 cores \cite{Burstedde;Ghattas;Gurnis;etal:08:Scalable-adaptive-ma}.
-\texttt{Dendro} \cite{Sampath;Adavani;Sundar;etal:08:Dendro:-parallel-alg} is another example of an octree code in which similar ideas are used for parallel partitioning and development of a parallel geometric multigrid that has been scaled up to about 32,000 cores \cite{Sampath;Biros:10:A-parallel-geometric}.
+\texttt{Dendro} \cite{Sampath;Adavani;Sundar;etal:08:Dendro:-parallel-alg} is another example of an Octree code in which similar ideas are used for parallel partitioning and development of a parallel geometric multigrid that has been scaled up to about 32,000 cores \cite{Sampath;Biros:10:A-parallel-geometric}.
 More recently, authors in \cite{Burstedde;Wilcox;Ghattas:11:p4est:-Scalable-Algo} extended these ideas to a collection, or a ``forest'', of Octrees that are connected through a common, potentially unstructured, hexahedral grid.
 This forest is then partitioned in parallel using a global Morton curve.
-The implementation of these algorithms, which were shown to scale to more than 200,000 processors, is publicly available through a simple API provided by the \texttt{p4est} library \cite{p4est-github}.
+The implementation of these algorithms, which were shown to scale to more than 200,000 cores, is publicly available through a simple API provided by the \texttt{p4est} library \cite{p4est-github}.
 In fact the algorithms presented in this paper are directly implemented on top of the \texttt{p4est} library and we do not discuss any algorithm that is already covered in \cite{Burstedde;Wilcox;Ghattas:11:p4est:-Scalable-Algo}.
@@
 Eulerian advection schemes can easily be parallelized but unfortunately are limited by the CFL condition which could be very restrictive for adaptive grids.
-Semi-Lagrangian methods combine the unconditional stability of Lagrangian methods and the ease of use of Eulerian grids and have been successfully used for advecting the level-set function on tree-based grids in the past \cite{Min;Gibou:07:A-second-order-accur}.
+semi-Lagrangian methods combine the unconditional stability of Lagrangian methods and the ease of use of Eulerian grids and have been successfully used for advecting the level-set function on tree-based grids in the past \cite{Min;Gibou:07:A-second-order-accur}.
 However, parallelizing the semi-Lagrangian algorithm in a domain decomposition context is not an easy task.
 The reason for this is twofold.
-First, depending on the CFL number, the departure points may end up outside the ghost region and in remote processors that are potentially far away.
+First, depending on the CFL number, the departure points may end up outside the ghost region and in remote processes that are potentially far away.
 This requires a very dynamic and nonuniform communication pattern which is complicated to implement.
-For an adaptive grid situation, it is even more complicated due to the asymmetric nature of the communication (c.f.
-\ref{sec:parallel algorithms}).
+For an adaptive grid, the situation is even more complicated due to the asymmetric nature of the communication (c.f.
+section \ref{sec:parallel algorithms}).
 Second, load balancing could be an issue for large CFL numbers and nonuniform velocity fields due to clustering of departure points and can considerably restrict the scalability of the algorithm.
-Both of these problems, of course, could be avoided by choosing $\text{CFL} \le 1$ but that would defeat the purpose of using semi-Lagrangian algorithm in the first place.
+Both of these problems, of course, could be avoided by choosing $\text{CFL} \le 1$ but that would defeat the purpose of using the semi-Lagrangian algorithm in the first place.
 
@@
 Here the width of the ghost layer is dynamically determined at runtime based on information from previous time steps.
-Unfortunately, however, this approach seems to suffer from excessive communication overhead at larger numbers of processors as well \comment{as well ?}.
+Unfortunately, this approach also suffers from excessive communication overhead at large numbers of processes.
 More recently, the authors in \cite{White-III;Dongarra:11:High-performance-hig} used a domain decomposition strategy on a cubed sphere but with a single layer of ghost nodes.
-Interpolation on remote processors is then handled by sending query points to the corresponding processor and asking for the interpolated result.
+Interpolation on remote processes is then handled by sending query points to the corresponding process and asking for the interpolated result.
 This approach seems to provide good scalability for transporting a single tracer up to about 1000 cores for $\text{CFL} \sim 10$.
 At higher CFL numbers the method begins to loose scalability due to an increase in communication volume.
-Although in this article we are mainly interested in parallel semi-Lagrangian methods, one could resort to finite difference or finite element discretization methods if small CFL numbers are acceptable.
+Finally, note that although we are mainly interested in parallel semi-Lagrangian methods, one could resort to finite difference or finite element discretization methods if small CFL numbers are acceptable.
 Indeed several algorithms of this type have been proposed with applications in modeling dendritic crystal growth \cite{Wang;Chang;Kale;etal:06:Parallelization-of-a}, multiphase flows \cite{Sussman:05:A-parallelized-adapt, Fortmeier;Bucker:11:A-parallel-strategy-, Rodriguez;Sahni;Lahey-Jr;etal:13:A-parallel-adaptive-}, and atomization process \cite{Herrmann:10:A-parallel-Eulerian-}.
@@
 
-In many applications of the level-set method, it is desirable for the level-set function to be a signed distance function, i.e.
-$|\nabla \phi| = 1$.
-There are generally two approaches to enforce this property: 1) Solving the pseudo-time transient reinitialization equation \cite{Sussman;Smereka;Osher:94:A-Level-Set-Approach, Osher;Fedkiw:01:Level-Set-Methods:-A}:
+In many applications, it is desirable that the level-set function has signed-distance property, i.e $|\nabla \phi| = 1$.
+Generally, there are two approaches to enforce this property: 1) by solving the pseudo-time transient reinitialization equation \cite{Sussman;Smereka;Osher:94:A-Level-Set-Approach, Osher;Fedkiw:01:Level-Set-Methods:-A}:
 \ben
@@
 \een
-or 2) Solving the Eikonal equation:
+or 2) by solving the Eikonal equation:
 \ben
@@
 This is the approach we have chosen in this paper.
-However, if the signed-distance property is required in the entire domain solving the Eikonal equation is more computationally efficient.
-Unfortunately, the most popular algorithm for solving the Eikonal equation, i.e.
-the Fast Marching Method \cite{Sethian:96:A-Fast-Marching-Leve,Sethian:99:Level-set-methods-an}, is inherently sequential due to causal relationship between grid points and cannot be easily parallelized.
+However, if the signed-distance property is required in the entire domain, solving the Eikonal equation is more computationally efficient.
+Unfortunately, the most popular algorithm for solving the Eikonal equation, i.e.\ the Fast Marching Method \cite{Sethian:96:A-Fast-Marching-Leve,Sethian:99:Level-set-methods-an}, is inherently sequential due to causal relationship between grid points and cannot be easily parallelized.
 The Fast Sweeping Method (FSM) \cite{Zhao:05:A-fast-sweeping-meth} is an alternative method for solving the Eikonal equation iteratively.
-The FSM can be more computationally efficient for simple choices of speed function, e.g.
-as in this context, and for simple interfaces.
+The FSM can be more computationally efficient for simple choices of speed function, e.g.\ as in this context, and for simple interfaces.
 Moreover, FSM has more potential for parallelization compared to the FMM.
@@
 One of the earliest attempt in parallelizing the FMM is reported in \cite{Herrmann:03:A-domain-decompositi} where a domain decomposition algorithm was introduced.
-Unlike the serial FMM, however, parallel FMM potentially requires multiple iterations or ``rollback operations'' to enforce the causality across processors.
+Unlike the serial FMM, however, parallel FMM potentially requires multiple iterations or ``rollback operations'' to enforce causality across processes.
 Similar ideas are described in details in \cite{Tugurlan:08:Fast-marching-method}.
@@
 However, unlike FMM, FIM avoids sorting the list and allows for concurrent updating of all nodes in an iterative fashion.
-\comment{Shouldn't we say that concluding from this analysis, we choose to use iterative reinitialization ?}
+In this article we simply choose the pseudo-time transient formulation for two reasons: 1) it is considerably easier to parallelize on Quadtrees and Octrees and 2) we are merely interested in the sign-distance property close to the interface which only requires a few iterations.
 

### 2015-01-08 paper/tex/levelset_method.tex
proof-read up to sec. 3.2

@@
 The level-set method, introduced in \cite{Osher;Sethian:88:Fronts-Propagating-w}, is an implicit framework for tracking interfaces that undergo complicated topological changes.
-In this framework an interface is represented by the zero contour of a higher dimensional function, e.g.
-a curve in two spatial dimension can be described as $\Gamma = \{(x,y) | \phi(x,y) = 0\}$ where $\phi(x,y)$ is the level-set function.
+In this framework an interface is represented by the zero contour of a higher dimensional function, e.g.\ a curve in two spatial dimension can be described as $\Gamma = \{(x,y) | \phi(x,y) = 0\}$ where $\phi(x,y)$ is the level-set function.
 The evolution of the curve under a velocity field $\mathbf{V}$ is then obtained by solving the level-set equation:
@@
 \ee
-When the velocity field does not depend on the level-set function itself, equation \eqref{eq:ls} can be solved using the Semi-Lagrangian method.
-An important advantage of the Semi-Lagrangian method over the regular finite difference method is its unconditional stability which allows for arbitrarily large time steps.
+When the velocity field does not depend on the level-set function itself, equation \eqref{eq:ls} can be solved using the semi-Lagrangian method.
+An important advantage of the semi-Lagrangian method over the regular finite difference method is its unconditional stability which allows for arbitrarily large time steps.
 This is particularly important when using adaptive grids that allow for higher grid resolutions.
@@
 In general, an infinite number of level-set functions can describe the same zero contour and thus interface.
-However, it is desirable to chose a function with the signed distance property, i.e.
-$|\nabla \phi| = 1$.
+However, it is desirable to chose a function with the signed distance property, i.e.\ $|\nabla \phi| = 1$.
 As detailed in section \ref{sec:introduction}, we solve the pseudo-time transient reinitialization equation \cite{Sussman;Smereka;Osher:94:A-Level-Set-Approach, Osher;Fedkiw:01:Level-Set-Methods:-A} to achieve this property,

### 2015-01-08 paper/tex/parallel_algorithms.tex
proof-read up to sec. 3.2

@@
 In order to develop scalable parallel algorithms on these grids, it is necessary to parallelize the datastructure and grid manipulations methods such as refinement and coarsening of cells as well as to provide a fast method for grid partitioning and load balancing.
-The \texttt{p4est} library \cite{p4est-github} is a collection of such parallel algorithms that has recently emerged and shown to scale up to 200,000 processors \cite{Burstedde;Wilcox;Ghattas:11:p4est:-Scalable-Algo}.
+The \texttt{p4est} library \cite{p4est-github} is a collection of such parallel algorithms that has recently emerged and shown to scale up to 200,000 cores \cite{Burstedde;Wilcox;Ghattas:11:p4est:-Scalable-Algo}.
 
@@
 This common coarse grid, which we will refer to as the ``macromesh'', can in general be an unstructured quadrilateral mesh, in two spatial dimensions, or hexahedral mesh, in three spatial dimensions.
-In this article, however, we shall limit ourselves to simple uniform Cartesian macromeshes.
-Moreover it is implicitly assumed that the macromesh is small enough that it can be entirely replicated on all processors.
+In this article, however, we limit discussions to simple uniform Cartesian macromeshshes.
+Moreover it is implicitly assumed that the macromesh is small enough that it can be entirely replicated on all processes.
 For instance, in many of the applications that we are interested in in this paper the macromesh is simply a single cell.
@@
 
-Once the gird is adapted to the interface, it must be partitioned to ensure load balancing across processors.
+Once the gird is adapted to the interface, it must be partitioned to ensure load balancing across processes.
 This is achieved by constructing a Z-curve that traverses the leaves of all trees in order of tree index (c.f.
@@
 \caption{Left: A ``forest'' made up of two trees $k_0$ and $k_1$.
-Parallel partitioning is achieved by first constructing a Z-curve which starts at cell ``S'' and ends at cell ``E''.
-Next this tree is split up among processors either uniformly or by assigning different weights to cells.
-Here processors are represented via different colors.
-Note how using the Z-curve naturally leads to clustering of most cells in each processor domain.
+Parallel partitioning is achieved by first constructing a Z-curve starting at cell ``S'' and ending at cell ``E''.
+Next the curve is split up among processes either uniformly or by assigning different weights to cells.
+Here processes are represented via different colors.
+Note how using the Z-curve naturally leads to clustering of most cells in each domain.
 Right: Schematic of a tree data structure representing the forest and its partitioning.}
@@
 only the leaves are explicitly stored.
-However, all the algorithms considered here require explicit knowledge of the entire tree \comment{that's not ture.
-We need the hierarchy to speed up point lookup}.
-As a result we introduce a simple algorithm, called \texttt{Reconstruct}, which recreates a local representation of the entire ``forest'' that is only adapted to local cells and, potentially, the ghost layer.
+However, \textcolor{blue}{explicit knowledge of the hierarchal structure of tree is greatly beneficial in several algorithms, e.g.\ in search operations needed for the interpolation algorithm.} Thus, we introduce a simple reconstruction algorithm which recreates a local representation of the entire ``forest'' that is only adapted to local cells and, potentially, the ghost layer.
 This approach is similar to the ideas introduced in \cite{Bangerth;Burstedde;Heister;etal:11:Algorithms-and-data-} and our tests show that in a typical application they usually amount to less that $1\%$ of the entire runtime.
@@
 Given a forest and a layer of ghost cells from \texttt{p4est}, the algorithm generates a local representation of the forest by recursively refining from the root until reaching the same level and location of all leaves in \texttt{p4est} and \texttt{ghost}.
-Note that algorithm \ref{alg:reconstruction} does not involve any communication and is load balanced \comment{as long as the forest is balanced}.
+Note that algorithm \ref{alg:reconstruction} does not involve any communication and is load balanced provided that the initial forest is balanced.
 Figure \ref{fig:reconstruction} illustrates an application of algorithm \ref{alg:reconstruction}.
-Note how each processor has independently generated a local representation of the forest that is refined to match the same leaves as in the global forest and ghost layer.
-% , and has an average runtime complexity of $\mathcal{O}\left((n_l + n_g)\log_{2^d}N\right)$ for well-balanced trees in $d$ dimensions.
-Here, $n_l$ and $n_g$ are the number of local and ghost cells, respectively, and $N$ is the global number of cells.
+Note how each process has independently generated a local representation of the forest that is refined to match the same leaves as in the global forest and ghost layer.
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[width = \textwidth]{figures/reconstruct.pdf}
+\end{center}
+\caption{Left: A forest refined close to an interface and partitioned among four processes as indicated by colors.
+Right: Each process independently recreates a local forest which is refined to match the global grid and is as coarse as possible elsewhere.
+Note that empty cells are fictitious, i.e.
+they are only required to generate the hierarchal structure and are not matched by any corresponding cell in the global forest.}
+\label{fig:reconstruction}
+\end{figure}
 \begin{algorithm}[htbp]
@@
 
-\begin{figure}[htbp]
-\begin{center}
-\includegraphics[width = \textwidth]{figures/reconstruct.pdf}
-\end{center}
-\caption{Left: A forest refined close to an interface and partitioned among four processors as indicated by colors.
-Right: Each processor independently recreates a local forest which is refined to match the global grid and is as coarse as possible elsewhere.
-Note that empty cells are fictitious, i.e.
-they are only required to generate the hierarchal structure and are not matched by any corresponding cell in the global forest.}
-\label{fig:reconstruction}
-\end{figure}
-
-\subsection{Semi-Lagrangian method \comment{and interpolations}}
+\subsection{Interpolation and semi-Lagrangian method}
 As indicated earlier, we use the Semi-Lagrangian method to solve equation \eqref{eq:ls} when the velocity field is externally generated, i.e.
@@
 In fact, the parallel interpolation algorithm \ref{alg:interpolation} is probably the most important contribution of this article \comment{but hasn't it been done before ?}.
-Complication arises because not all calculated departure points will reside in the domain of current processor.
-Moreover, due to the irregular shape of partitions, we cannot be certain that they are entirely owned by neighboring processors.
+Complication arises because not all calculated departure points will reside in the domain of current process.
+Moreover, due to the irregular shape of partitions, we cannot be certain that they are entirely owned by neighboring processes.
 At best we can only expect that their location are bounded by a halo of width $w \le \text{CFL} \: \Delta x_{\min}$ around the local partition.
@@
 Our remote interpolation algorithm is composed of three separate phases.
-In the first phase, which we call buffering, every processor searches for all departure points inside the local trees.
-If the point is owned by a local cell, it is added to a local buffer, otherwise we find the processor which owns the point and add the point to a separate buffer belonging to the found rank.
+In the first phase, which we call buffering, every process searches for all departure points inside the local trees.
+If the point is owned by a local cell, it is added to a local buffer, otherwise we find the process which owns the point and add the point to a separate buffer belonging to the found rank.
 Note that searching the point in the local tree is performed recursively \comment{using the hierarchal reconstruction}, similarly to the \texttt{UpdateLocalTree} function in algorithm \ref{alg:reconstruction}.
@@
 
-Once buffering is done, every processor knows exactly how many messages it needs to send and to which processors \comment{throughout the whole article, shouldn't we talk of ``processes'' ?}.
-This also implicitly defines processors that will later on send a reply message to this processor.
-However, at this point no one knows which processors they should expect a message from.
+Once buffering is done, every process knows exactly how many messages it needs to send and to which processes \comment{throughout the whole article, shouldn't we talk of ``processes'' ?}.
+This also implicitly defines processes that will later on send a reply message to this process.
+However, at this point no one knows which processes they should expect a message from.
 We solve this problem using a simple \textit{communication matrix} (see figure \ref{fig:communication}).
@@
 Although the distance to the departure point is bounded by $\text{CFL} \: \Delta x_{\min}$, one cannot make any special assumption about receiving rank without explicitly searching the entire Z-curve.
-Moreover, the receiving processor has no prior knowledge about which processors to check for incoming messages nor does it know anything about the possible message length (i.e.
+Moreover, the receiving process has no prior knowledge about which processes to check for incoming messages nor does it know anything about the possible message length (i.e.
 number of points).
-Middle: A directed graph illustrating the communication pattern among processors with arrows representing the direction in which messages are sent.
+Middle: A directed graph illustrating the communication pattern among processes with arrows representing the direction in which messages are sent.
 Right: The adjacency matrix of the communication graph.
@@
 
-To solve the communication problem, we first compute the adjacency matrix of the communication pattern, i.e. we construct the matrix $A_{P \times P}$, where $P$ is the number of processors, such that
+To solve the communication problem, we first compute the adjacency matrix of the communication pattern, i.e. we construct the matrix $A_{P \times P}$, where $P$ is the number of processes, such that
 \ben
@@
 \begin{array}{lr}
-1 \hspace{5 mm} \text{if processor `\textit{i}' sends a message to processor `\textit{j}',} \\
+1 \hspace{5 mm} \text{if process `\textit{i}' sends a message to process `\textit{j}',} \\
 0 \hspace{5 mm} \text{otherwise.}
@@
 \een
-Note that this matrix is also distributed among processors, i.e. each row is owned by a separate rank. Next, we compute
+Note that this matrix is also distributed among processes, i.e. each row is owned by a separate rank. Next, we compute
 \ben
@@
 For instance, this can be achieved using a single \texttt{MPI\_Reduce\_scatter} function call.
-The last phase of the interpolation procedure involves overlapping the computation of interpolated values for local points with the communication of data between processors.
-This is done by alternating between local calculations and probing for incoming messages from other processors.
+The last phase of the interpolation procedure involves overlapping the computation of interpolated values for local points with the communication of data between processes.
+This is done by alternating between local calculations and probing for incoming messages from other processes.
 The interpolation is finished once the values for all local the points have been calculated and all the remote requests have been processed (see algorithm \ref{alg:interpolation}).
@@
 \For {$r: mpisize$} \Comment{Phase II -- initiate communication and compute number of messages}
-	\If {$\textcolor{red}{r!=mpirank~and~} col[r]$}
+	\If {$col[r]$ \textcolor{blue}{col[mpirank] is always zero ...}}
 		\State $\texttt{MP\_Isend(}r, buff[r]\texttt{)}$
@@
 	\EndIf
-	\If {$R > 0$} \Comment{process queries sent from remote processors}
+	\If {$R > 0$} \Comment{process queries sent from remote processes}
 		\State $[msg, st] \gets \texttt{MPI\_Iprobe()}$
@@
 Note that this iterative process is essentially unavoidable since the grid is based on the values of the level-set function at $t^{n+1}$, which itself is unknown and is to be defined on $G^{n+1}$.
-Nonetheless the process converges to the final grid in at most $l_{\max}$ steps \comment{shouldn't it be $l_{max}-l_{min}$ ?} where $l_{\max}$ denotes the maximum final depth of the trees across all processors.
+Nonetheless the process converges to the final grid in at most $l_{\max}$ steps \comment{shouldn't it be $l_{max}-l_{min}$ ?} where $l_{\max}$ denotes the maximum final depth of the trees across all processes.
 \begin{algorithm}[htbp]
@@
 First, we label all local points, $L_p$, as either private, $P_p$, or boundary, $B_p$.
-Here, the boundary points are the collection of all local points that are regarded as a ghost point, $G_r$, on at least one other processor, i.e.
+Here, the boundary points are the collection of all local points that are regarded as a ghost point, $G_r$, on at least one other process, i.e.
 $B_p = \underset{r,\;r\neq p}{\bigcup} G_r$.

### 2015-01-09 paper/tex/applications.tex
Finished my pass

@@
 
-We propose to apply our approach to the study of the phase transition of a liquid melt to a solid crystaline structure.
-In the case of a single component melt and in the absence of convection, the process is dominated by diffusion and can be modeled as a Stefan problem.
-We decompose the computational domain $\Omega$ into two subdomains $\Omega^-$ and $\Omega^+$, separated by an interface $\Gamma$.
+We propose to apply \mohammad{what about: In this section we apply ...} our approach to the study of the phase transition of a liquid melt to a solid crystaline structure.
+In the case of a single component melt, and in the absence of convection, the process is dominated by diffusion and can be modeled as a Stefan problem.
+We decompose the computational domain $\Omega$ into \mohammad{How about: Consider a computational domain $\Omega$ decomposed into ...} two subdomains $\Omega^-$ and $\Omega^+$, separated by an interface $\Gamma$.
 The Stefan problem describes the evolution of the temperature $T$, decomposed into $T_s$ in the solid phase $\Omega^+$ and $T_l$ in the liquid phase $\Omega^-$, as
@@
 \caption{$\texttt{General procedure for solving the Stefan problem}$}
+\mohammad{Is there a reason all of the algorithm is in type-writer font? Usually its reserved for code segments, etc.}
 \begin{algorithmic}[1]
@@
 \caption{Scalability of the Stefan problem for small (left) and large (right) octrees with roughly 7M and 105M grid points, respectively.
-As expected from the scalability analysis of the individual components, we observe excellent results, illustrating the potential of our solver.}
+As expected from the scalability analysis of the individual components, we observe excellent results, illustrating the potential \mohammad{applicability?} of our solver \mohammad{algorithms?}.}
 \label{fig:stefan_scaling}
@@
 \subsection{Numerical experiments}
-
+\mohammad{too many transitive statements and keyword ``choose''. Can we rephrase this?}
 We now present the results from a large simulation of the Stefan problem.
@@
 \end{align*}
-Note that these expressions are arbitrary and lead to very varied crystal shapes.
+Note that these expressions are arbitrary and lead to very varied crystal shapes \mohammad{How about: ...
+lead to a variety of shapes?}.
 The computation is initialized with twenty spherical seeds of radius $0.0015$ placed randomly in the domain.
@@
 
-We let the simulation run with $256$ mpi processes for $6$ hours and $30$ minutes, resulting in $396$ time iterations.
+We let the simulation run with \mohammad{How about: The simulation is ran on ...} $256$ mpi processes for $6$ hours and $30$ minutes, resulting in $396$ time iterations.
 Visualizations of the final iteration are presented in figures \ref{fig:stefan_grid} and \ref{fig:stefan_crystals}.
@@
 over eight trillion grid points !
-Our simulation used only $0.001948\%$ of the number of grid points needed for the same simulation on a uniform grid.
+Our simulation used only $0.001948\%$ \mohammad{I think if we are going to report percentage, we should trim it to 1st digit} of the number of grid points needed for the same simulation on a uniform grid.
 This application demonstrates the potential of our approach.

### 2015-01-09 paper/tex/conclusion.tex
Finished my pass

@@
 In this article we have presented parallel algorithms for the advection and reinitialization of the level-set functions on adaptive Quadtree and Octree grids using domain decomposition approach.
-These algorithms are implemented using a combination of \texttt{MPI} standard and the open-source \texttt{p4est} library.
-An important feature of the Semi-Lagrangian method is its unconditional stability property which must be preserved in parallel, requiring a parallel interpolation scheme.
+These algorithms are implemented using a combination of \texttt{MPI} and the open-source \texttt{p4est} library.
+An important feature of the semi-Lagrangian method is its unconditional stability property which must be preserved in parallel, requiring a parallel interpolation scheme.
 However, due to the non-uniform distribution of departure points, scalable implementation of such algorithm is not trivial on adaptive Quadtree and Octree grids.
@@
 
-The scalability of the Semi-Lagrangian algorithm, however, depends on the CFL number.
+The scalability of the semi-Lagrangian algorithm, however, depends on the CFL number.
 Great scalability is observed for intermediate CFL numbers while, e.g.
-$\text{CFL}~10$.
+$\text{CFL}\sim10$.
 At higher CFL numbers, however, the departure points are potentially further dispersed across processors which limits the scalability.
-This is due to the fact the domain decomposition technique used here is based on the Z-ordering of quadrants and does not take the velocity field information into account.
-A possible remedy for this problem could be assigning weights to quadrants based on some estimate of what the grid should look like after one step of the advection algorithm, e.g.
+This is because the domain decomposition technique used here is based on the Z-ordering of cells and does not take the velocity field information into account.
+A possible remedy for this problem could be assigning weights to cells based on some estimate of what the grid should look like after one step of the advection algorithm, e.g.
 by using a forward-in-time integration of grid points.
-Such an estimate could also reduce the number of Semi-Lagrangian iterations.
+Such an estimate could also reduce the number of semi-Lagrangian iterations.
 These ideas are postponed for further investigations.
 We have also presented a simple parallelization technique for the reinitialization algorithm based on the pseudo-time transient formulation.
-Both the Semi-Lagrangian and the reinitialization algorithms show great scalability up to 4096 processors.
+Both the semi-Lagrangian and the reinitialization algorithms show good scalability up to 4096 processors.
 
-Finally, an application of these algorithms is presented in modeling the solidification process by solving the Stefan problem.
+Finally, an application of these algorithms is presented in modeling the solidification process by solving a Stefan problem.
 This application clearly illustrates the applicability of our algorithms to complex multi-scale problems that cannot be treated using the normal domain decomposition techniques on uniform grids.
-We believe that our findings would be of interest to other researchers using the level-set framework for other complex and multi-scale problems.
+We believe that our findings would be of interest to other researchers who use the level-set framework for similar complex and multi-scale problems.
 
 \section*{Acknowledgment} 
-The funding for this project was provided by \comment{the funding agency}.
-We would also like to thank the technical staff at the Texas Advanced Computing Center (TACC) and the developer community of PETSc library for their valuable suggestions.
+The funding for this project was provided by \mohammad{the funding agency}.
+We would also like to thank the technical staff at the Texas Advanced Computing Center (TACC) and the developer community of PETSc library for their valuable suggestions during this project.

### 2015-01-09 paper/tex/parallel_algorithms.tex
Finished my pass

@@
 
-\subsection{Interpolation and semi-Lagrangian method}
-As indicated earlier, we use the Semi-Lagrangian method to solve equation \eqref{eq:ls} when the velocity field is externally generated, i.e.
+\subsection{Interpolation and semi-Lagrangian methods}
+As indicated earlier, we use the semi-Lagrangian method to solve equation \eqref{eq:ls} when the velocity field is externally generated, i.e.
 when it does not depend on the level-set function itself.
@@
 \ee
-The Semi-Lagrangian method integrates equations \eqref{eq:sl} backward in time, i.e.
+The semi-Lagrangian method integrates equations \eqref{eq:sl} backward in time, i.e.
 starting from the grid $G^{n+1}$ \footnote{$G^{n+1}$ itself is computed in an iterative fashion as explained later on.}, we simply write $\phi^{n+1}(\mathbf{X}^{n+1}) = \phi(\mathbf{X}(t^{n+1}), t^{n+1}) = \phi(\mathbf{X}(t^n), t^n) = \phi^n(\mathbf{X}_d)$.
@@
 Although parallelization of the interpolation process on a shared-memory machine is trivial, the same cannot be said for distributed-memory machines.
-In fact, the parallel interpolation algorithm \ref{alg:interpolation} is probably the most important contribution of this article \comment{but hasn't it been done before ?}.
-Complication arises because not all calculated departure points will reside in the domain of current process.
-Moreover, due to the irregular shape of partitions, we cannot be certain that they are entirely owned by neighboring processes.
-At best we can only expect that their location are bounded by a halo of width $w \le \text{CFL} \: \Delta x_{\min}$ around the local partition.
+In fact, the parallel interpolation algorithm \ref{alg:interpolation} is probably the most important contribution of this article \comment{but hasn't it been done before ?} \textcolor{blue}{Mohammad: not on adaptive since that's what makes it complicated}.
+Complication arises because not all departure points will reside in the domain of current process.
+Moreover, due to the irregular shape of partitions, one cannot even ensure they are entirely owned by neighboring processes.
+At best we can only expect that their locations are bounded by a halo of width $w \le \text{CFL} \: \Delta x_{\min}$ around the local partition.
 Of course if one enforces $\text{CFL} \le 1$, one can ensure that the halo is bounded by the ghost layer which significantly simplifies the communication problem.
-This assumption, however, defeats the purpose of using a Semi-Lagrangian approach in the first place since we are interested in large values of $\text{CFL}$ number.
+This assumption, however, defeats the purpose of using a semi-Lagrangian approach in the first place since we are interested in large values of $\text{CFL}$ number.
 
@@
 If the point is owned by a local cell, it is added to a local buffer, otherwise we find the process which owns the point and add the point to a separate buffer belonging to the found rank.
-Note that searching the point in the local tree is performed recursively \comment{using the hierarchal reconstruction}, similarly to the \texttt{UpdateLocalTree} function in algorithm \ref{alg:reconstruction}.
-The owner's rank is found by computing the Z-index of the point and then using binary search on the Z-curve.
+Note that searching the point in the local tree is performed recursively using the hierarchal reconstruction (c.f.\ algorithm \ref{alg:reconstruction}).
+Moreover, the owner's rank is found by computing the Z-index of the point and then using binary search on the Z-curve.
 This is already implemented in \texttt{p4est} and explained in details in section 2.5 of \cite{Burstedde;Wilcox;Ghattas:11:p4est:-Scalable-Algo}.
 
-Once buffering is done, every process knows exactly how many messages it needs to send and to which processes \comment{throughout the whole article, shouldn't we talk of ``processes'' ?}.
+Once buffering is done, every process knows exactly how many messages it needs to send and to which processes.
 This also implicitly defines processes that will later on send a reply message to this process.
-However, at this point no one knows which processes they should expect a message from.
+However, at this point no process knows which processes to expect a message from.
 We solve this problem using a simple \textit{communication matrix} (see figure \ref{fig:communication}).
-Other approaches for solving this problem could include using non-blocking collective or one-sided communication operations introduced in the MPI-3 standard \cite{MPI_ref, sparse dynamic exchange protocol paper}.
+Other approaches for solving this problem could include using non-blocking collective or one-sided communication operations introduced in the MPI-3 standard \cite{Hoefler;Siebert;Lumsdaine:10:Scalable-communication}.
 Although these algorithms have better theoretical communication complexities, we did not see any difference in the runtime or scalability of algorithm \ref{alg:interpolation} when using them.
@@
 \caption{Left: Location of back-traced points depends on the magnitude of the local velocity and on the time-step.
-Although the distance to the departure point is bounded by $\text{CFL} \: \Delta x_{\min}$, one cannot make any special assumption about receiving rank without explicitly searching the entire Z-curve.
+Although the distance to the departure point is bounded by $\text{CFL} \: \Delta x_{\min}$, one cannot make any assumption about receiving rank without explicitly searching the entire Z-curve.
 Moreover, the receiving process has no prior knowledge about which processes to check for incoming messages nor does it know anything about the possible message length (i.e.
@@
 Right: The adjacency matrix of the communication graph.
-For each row, the sum of all columns represents the number of messages sent.
+For each row, the sum of all columns represents the number of messages that need to be sent.
 Conversely, for each column, the sum of all rows represents the number of messages that need to be received.
@@
 \een
-where $S_i$ and $R_i$ denote the number of sent and received messages respectively.
+where $S_i$ and $R_i$ denote the number of sent and received messages, respectively.
 While $S_i$ can be computed trivially, a reduction operation is required to compute $R_i$.
@@
 \For {$r: mpisize$} \Comment{Phase II -- initiate communication and compute number of messages}
-	\If {$col[r]$ \textcolor{blue}{col[mpirank] is always zero ...}}
+	\If {$col[r]$}
 		\State $\texttt{MP\_Isend(}r, buff[r]\texttt{)}$
@@
 
-Using the interpolation algorithm \ref{alg:interpolation}, we close this section by presenting the final Semi-Lagrangian algorithm \ref{alg:semi-lagrangian}.
+Using the interpolation algorithm \ref{alg:interpolation}, we close this section by presenting the final semi-Lagrangian algorithm \ref{alg:semi-lagrangian}.
 The basic idea here is to start from an initial guess $G^{n+1}_0$ and modify the grid using the refinement \eqref{eq:refine} and coarsening \eqref{eq:coarsen} criteria until convergence is obtained.
@@
 Note that this iterative process is essentially unavoidable since the grid is based on the values of the level-set function at $t^{n+1}$, which itself is unknown and is to be defined on $G^{n+1}$.
-Nonetheless the process converges to the final grid in at most $l_{\max}$ steps \comment{shouldn't it be $l_{max}-l_{min}$ ?} where $l_{\max}$ denotes the maximum final depth of the trees across all processes.
+Nonetheless the process converges to the final grid in at most $l_{\max}-l_{\min}$ steps where $l_{\min}$ and $l_{\max}$ denote the maximum and minimum depth of all trees in the forest, receptively.
 \begin{algorithm}[htbp]
@@
 \subsection{Reinitialization}
-Successive application of algorithm \ref{alg:semi-lagrangian}, especially for large values of the CFL number, can potentially degrades the signed distance property of the level-set function.
-It is thus important to reinitialize the level-set function every few iterations, especially because the quality of grid generated depends heavily on the signed distance property.
+Successive application of algorithm \ref{alg:semi-lagrangian}, especially for large values of the CFL number, eventually degrades the signed distance property of the level-set function.
+Thus, it is important to reinitialize the level-set function every few iterations, especially because the quality of generated grid heavily depends on the signed distance property.
 To achieve this property we solve the pseudo-time transient equation \eqref{eq:reinitialization} using the discretization scheme detailed in \cite{Min;Gibou:07:A-second-order-accur}.
@@
 Similar to \cite{Min;Gibou:07:A-second-order-accur}, equation \eqref{eq:semidiscrete_reinit} is integrated in time using the TVD-RK2 scheme and with adaptive time-stepping in order to accelerate the convergence to the steady state.
-Since all the computation is based on a local stencil, the parallel implementation of this scheme is mostly trivial.
+Since the computation is based on a local stencil, the parallel implementation of this scheme is mostly trivial.
 However, one minor point requires further explanation.
-As suggested in \cite{Min;Gibou:07:A-second-order-accur}, one-sided derivatives $D^+_i \phi$ and $D^-_i \phi$ are computed using second order discretizations which require to compute the second derivatives.
+As suggested in \cite{Min;Gibou:07:A-second-order-accur}, one-sided derivatives $D^+_i \phi$ and $D^-_i \phi$ are computed using second order discretization which require to compute the second derivatives.
 To enable overlap between computation and communications when computing second order derivatives and also integrating \eqref{eq:semidiscrete_reinit}, we use the following common technique.
@@
 $P_p = L_p \setminus B_p$.
-Algorithm \ref{alg:overlap} illustrates how this labelling can help with overlapping the computation and the communication associated to an arbitrary local operation $\phi^{n+1} \gets \mathcal{F}(\phi^n)$.
-Note that \texttt{p4est} library already includes all the primitives required for labelling local points without any further communication.
+Algorithm \ref{alg:overlap} illustrates how this labeling can help with overlapping the computation and the communication associated to an arbitrary local operation $\phi^{n+1} \gets \mathcal{F}(\phi^n)$.
+Note that \texttt{p4est} library already includes all the primitives required for labeling local points without any further communication.
 

### 2015-01-09 paper/tex/scaling_results.tex
Finished my pass

@@
 %!TEX root = draft.tex
-\section{Scaling} \label{sec:scaling}
+\section{Scaling results} \label{sec:scaling}
 In this section we present some results that demonstrate the scalability of our algorithms.
 All of our tests were ran on the Stampede cluster at Texas Advanced Computing Center (TACC) where we are limited to $4096$ cores at most.
-Each node of Stampede has 2 eight-core Xenon E5-2680 processors clocked at 2.7 GHz with 32 GB of DDR3-1600 MHz memory and interconnected using an InfiniBand network card.
+Each node of Stampede has 2 eight-core Xenon E5-2680 processes clocked at 2.7 GHz with 32 GB of DDR3-1600 MHz memory and interconnected using an InfiniBand network card.
 Unless mentioned otherwise, in all the tests we have used all 16 cores of every node.
@@
 In this section we show the results for a simple test to measure the scalability of the interpolation algorithm \ref{alg:interpolation}.
-The test consists in interpolating a function at a number of random points on a randomly refined Octree in three spatial dimensions.
-We consider two cases, a small test on a level 9 tree with roughly 20M cells and 33M nodes and larger test on a level 13 tree with roughly 128M cells and 280M nodes.
-In both cases the number of randomly generated points is chosen to be equal to the number of nodes and the stablized second-order interpolation of \cite{Min;Gibou:07:A-second-order-accur} is performed 10 times to smooth out possible timing fluctuations.
+The test consists of interpolating a function at a number of random points on a randomly refined Octree in three spatial dimensions.
+We consider two cases, a small test on a level 9 tree with roughly 33M nodes and larger test on a level 13 tree with roughly 280M nodes.
+In both cases the number of randomly generated points is chosen to be equal to the number of nodes and the stabilized second-order interpolation of \cite{Min;Gibou:07:A-second-order-accur} is performed 10 times to smooth out possible timing fluctuations.
 
-To simulate the effect of different CFL numbers, we generate the random points such that on each processor $\alpha\%$ of them are located outside the processor boundary and will thus initiate communication.
+To simulate the effect of different CFL numbers, we generate the random points such that on each process $\alpha$ percentage of them are located outside the process boundary and thus will initiate communication.
 Scaling results are presented for $\alpha = 5 \%$ and $\alpha = 95\%$ for both the small and large problems in figure \ref{fig:interpolation}.
-Excellent scaling is obtained for the small problem for $P = 16-512$ even when $95\%$ of the interpolation points belong to a remote processor.
-For the larger problem, the communication overhead prevents the algorithm from scaling beyond $4096$ processors when $\alpha = 95\%$.
-Note that for $\alpha = 5\%$, even though some sections of the algorithm stop scaling, the overall algorithm still scales since the local work dominates the timing.
-This illustrates the effectiveness of the non-blocking communication pattern of algorithm \ref{alg:interpolation}.
-
+Excellent scaling is obtained for the small problem for $P = 16-512$ even when $95\%$ of the interpolation points belong to a remote process.
+For the larger problem, however, the communication overhead prevents the algorithm from scaling beyond $4096$ processes when $\alpha = 95\%$.
 \begin{figure}[htbp]
@@
 	\end{center}
-	\caption{Strong scaling of algorithm \ref{alg:interpolation} for several tests where $N_G$ denotes the number of random interpolation points (which is the same as the number of nodes in the Octree) and $\alpha$ denotes the percentage of these points that are remote for each processor.
+	\caption{Strong scaling of algorithm \ref{alg:interpolation} for several tests where $N_G$ denotes the number of random interpolation points (which is the same as the number of nodes in the Octree) and $\alpha$ denotes the percentage of these points that are remote for each process.
 Here ``Total'' represents the total time spent in the interpolation while ``buffer'', ``local'', ``queries'', and ``replies'' represent the the timing for different sections (c.f.
@@
 The results indicate excellent scaling for the small test (a-b) and for the large test when $\alpha = 5\%$ (c).
-For the extreme case (d) the algorithm stops scaling at $4096$ processors due to communication overhead.}
+For the extreme case (d) the algorithm stops scaling at $4096$ processes due to communication overhead.}
 	\label{fig:interpolation}
@@
 \subsection{Semi-Lagrangian}
-To test the scalability of the Semi-Lagrangian algorithm \ref{alg:semi-lagrangian} we consider a slightly modified version of the Enright's rotation test \cite{Enright;Fedkiw;Ferziger;etal:02:A-Hybrid-Particle-Le}, i.e.
+To test the scalability of the semi-Lagrangian algorithm \ref{alg:semi-lagrangian}, we consider a slightly modified version of the Enright's rotation test \cite{Enright;Fedkiw;Ferziger;etal:02:A-Hybrid-Particle-Le}, i.e.
 we advect a sphere of radius 0.35 located at $(0.4, 0.4, 0.4)$ with a divergence free velocity field given by:
@@
 \eean
-To understand the effect of the CFL number on the scalability of the algorithm we perform one step of the Semi-Lagrangian algorithm for $\text{CFL} = 1$, $\text{CFL} = 10$, and $\text{CFL} = 100$.
+To understand the effect of the CFL number on the scalability of the algorithm we perform one step of the semi-Lagrangian algorithm for $\text{CFL} = 1$, $\text{CFL} = 10$, and $\text{CFL} = 100$.
 We also perform the test for two different initial girds, a small grid with maximum level $l_\text{max} = 10$ and a large grid with maximum level $l_\text{max} = 12$.
@@
 	\caption{Number of iterations required for the grid construction in algorithm \ref{alg:semi-lagrangian} for the rotation test on a (a) level-10 and (b) level-12 Octree with approximately 15M and 255M nodes, respectively.
-Note how the iteration count increases with the CFL number but is almost independent of the number of processors.
-The slight dependence between the number of iterations and the number of processors is most likely due to the dependence of round-off errors on the number of processors.
-Nonetheless, close examination of the Octrees generated (data not shown) reveals that they are identical and independent of the number of processors used to perform the test.}
+Note how the iteration count increases with the CFL number but is almost independent of the number of processes.
+The slight dependence between the number of iterations and the number of processes is most likely due to the dependence of round-off errors on the number of processes.
+Nonetheless, close examination of the Octrees generated (data not shown) reveals that they are identical and independent of the number of processes used to perform the test.}
 	\label{tab:semilagrangian}
@@
 Table \ref{tab:semilagrangian} illustrates the dependence of the number of iterations required to build the grid on the CFL number; as the CFL is increased, the interface travels a farther distance which necessitates more iterations to generate the grid.
-Note that the apparent dependence on the number of processors is most likely due to round-off errors during the interpolation and/or backtracking steps.
+Note that the apparent dependence on the number of processes is most likely due to round-off errors during the interpolation and/or backtracking steps.
 Figures \ref{fig:semilagrangian_small} and \ref{fig:semilagrangian_large} illustrate the scalability of the algorithm for the small and large problems, respectively.
-To enable meaningful comparisons between different CFL numbers and number of processors, the maximum time has been scaled by the number of iterations required for the grid construction as reported in table \ref{tab:semilagrangian}.
+To enable meaningful comparisons between different CFL numbers and number of processes, the maximum time has been scaled by the number of iterations required for the grid construction as reported in table \ref{tab:semilagrangian}.
 For both problems, excellent scalability is observed for $\text{CFL} = 1$ and $\text{CFL} = 10$.
@@
 An increase in the CFL number has two effects on the algorithm.
-First, a larger fraction of the departure points lands in the domains of remote processors.
-Moreover, these points are potentially dispersed across a larger number of processors.
-This means that both the number of MPI messages and their volume increase as the CFL number increases.
-Second, as more points are shipped to remote processors for interpolation, there is a greater chance that the interpolation load is imbalanced across processors.
+First, a larger fraction of the departure points lands in the domains of remote processes.
+Moreover, these points are potentially dispersed across a larger number of processes.
+This means that the communication volume should increase with the CFL number.
+Second, as more points are shipped to remote processes for interpolation, there is a greater chance that the interpolation load is imbalanced across processes.
 This is especially true for regions of space in which the streamlines cluster.
-Both factors contribute to decrease the scalability of the algorithm at large CFL numbers.
+Both factors can contribute to reduce the scalability of the algorithm at large CFL numbers.
 \begin{figure}[htbp]
 	\begin{center}
-		\subfigure[Semi-Lagrangian, $\text{CFL} = 1$]{\includegraphics[width = 0.3 \textwidth] {figures/SL_Small_1_SemiLagrangian.pdf}}
-		\subfigure[Semi-Lagrangian, $\text{CFL} = 10$]{\includegraphics[width = 0.3 \textwidth] {figures/SL_Small_10_SemiLagrangian.pdf}}
-		\subfigure[Semi-Lagrangian, $\text{CFL} = 100$]{\includegraphics[width = 0.3 \textwidth] {figures/SL_Small_100_SemiLagrangian.pdf}}
+		\subfigure[semi-Lagrangian, $\text{CFL} = 1$]{\includegraphics[width = 0.3 \textwidth] {figures/SL_Small_1_SemiLagrangian.pdf}}
+		\subfigure[semi-Lagrangian, $\text{CFL} = 10$]{\includegraphics[width = 0.3 \textwidth] {figures/SL_Small_10_SemiLagrangian.pdf}}
+		\subfigure[semi-Lagrangian, $\text{CFL} = 100$]{\includegraphics[width = 0.3 \textwidth] {figures/SL_Small_100_SemiLagrangian.pdf}}
 		\\
@@
 	\end{center}
-	\caption{Strong scaling of a single iteration of algorithm \ref{alg:semi-lagrangian} for the rotation test on a level-10 Octree with approximately 13M cells and 15M nodes.
+	\caption{Strong scaling of a single iteration of algorithm \ref{alg:semi-lagrangian} for the rotation test on a level-10 Octree with approximately 15M nodes.
 Top row: scaling of the various components of the algorithm for (a) $\text{CFL} = 1$, (b) $\text{CFL} = 10$, and (c) $\text{CFL} = 100$.
@@
 	\begin{center}
-		\subfigure[Semi-Lagrangian, $\text{CFL} = 1$]{\includegraphics[width = 0.3 \textwidth] {figures/SL_Large_1_SemiLagrangian.pdf}}
-		\subfigure[Semi-Lagrangian, $\text{CFL} = 10$]{\includegraphics[width = 0.3 \textwidth] {figures/SL_Large_10_SemiLagrangian.pdf}}
-		\subfigure[Semi-Lagrangian, $\text{CFL} = 100$]{\includegraphics[width = 0.3 \textwidth] {figures/SL_Large_100_SemiLagrangian.pdf}}
+		\subfigure[semi-Lagrangian, $\text{CFL} = 1$]{\includegraphics[width = 0.3 \textwidth] {figures/SL_Large_1_SemiLagrangian.pdf}}
+		\subfigure[semi-Lagrangian, $\text{CFL} = 10$]{\includegraphics[width = 0.3 \textwidth] {figures/SL_Large_10_SemiLagrangian.pdf}}
+		\subfigure[semi-Lagrangian, $\text{CFL} = 100$]{\includegraphics[width = 0.3 \textwidth] {figures/SL_Large_100_SemiLagrangian.pdf}}
 		\\
@@
 	\end{center}
-	\caption{Strong scaling of a single iteration of algorithm \ref{alg:semi-lagrangian} for the rotation test on a level-12 Octree with approximately 251M cells and 255M nodes.
+	\caption{Strong scaling of a single iteration of algorithm \ref{alg:semi-lagrangian} for the rotation test on a level-12 Octree with approximately 255M nodes.
 Top row: scaling of the various components of the algorithm for (a) $\text{CFL} = 1$, (b) $\text{CFL} = 10$, and (c) $\text{CFL} = 100$.
@@
 To better understand the importance of the CFL number on the scalability, we have recorded a complete history of the communication pattern in the interpolation step.
-Figure \ref{fig:communication_4096} illustrates the effects of the CFL number on three different metrics, namely the number of interpolation points\footnote{Note that this includes both the local points and the points queried by other processors}, $N_p$, the number of sent and received messages, $N_m = S + R$, and the total communication volume, $V_m$ in mega bytes (MB), for $p=4096$ processors.
-Furthermore, these values are reported for the first (top row) and last (bottom row) iterations of the Semi-Lagrangian algorithm.
+Figure \ref{fig:communication_4096} illustrates the effects of the CFL number on different metrics, namely the number of interpolation points\footnote{Note that this includes both the local points and the points queried by other processes}, $N_p$, the number of sent and received messages, $N_m = S + R$, and the total communication volume, $V_m$ in megabytes (MB), for $p=4096$ processes.
+Furthermore, these values are reported for the first (top row) and last (bottom row) iterations of the semi-Lagrangian algorithm.
 There are several points to make.
 First, increasing the CFL number greatly increases the load imbalance, as shown by the spread of the data in figure \ref{fig:communication_4096_pf}.
-This is because at higher CFL numbers, and with the velocity field prescribed, it is more likely that some processors will receive a larger portion of the backtracked points.
-Second, increasing the CFL number increases both the communication volume and its spread across processors (c.f.
+This is because at higher CFL numbers, it is more likely that some processes will receive a larger portion of the backtracked points.
+Second, increasing the CFL number increases both the communication volume and its spread across processes (c.f.
 figure \ref{fig:communication_4096_vf}).
-Interestingly, however, the number of messages sent and received seems to be unaffected by the CFL number.
-The bottom row of figure \ref{fig:communication_4096} exhibits a better balance both in the computation and communication volume in the last iteration of the Semi-Lagrangian algorithm.
-This can be explained by noting that as the algorithm converges to the final grid, partitioning of the grid essentially becomes equivalent to partitioning of the backtracked points \comment{awkward ...
-maybe: the partitioning of the the grids $G^n$ and $G^{n+1}$ become very similar ?}.
+Interestingly, however, the number of sent and received messages do not seem to affected by the CFL number.
+The bottom row of figure \ref{fig:communication_4096} exhibits a better balance both in the computation and communication volume in the last iteration of the semi-Lagrangian algorithm.
+This can be explained by noting that for the final iteration, partitioning of $G^{n+1}$ is a good approximation of partitioning of departure points on $G^n$.
+\comment{awkward ...
+maybe: the partitioning of the grids $G^n$ and $G^{n+1}$ become very similar ?} \mohammad{not quite ...
+I am not sure what would be the best to phrase it but what i want to say is load balancing really requires you to partition the departure points on $G^n$ and not the $G^{n+1}$ which, of course, you cannot since you do not have that information ...
+the situation becomes better at the final iteration because both partitions are proportional to (Surface area of interface)/(Number of processors) ...} .
 Detailed information about the load balancing and the communication patterns is listed in table \ref{tab:communication}.
@@
 	\end{center}
-	\caption{Performance indicators of the first (top row) and last (bottom row) iterations of the Semi-Lagrangian algorithm for the advection on 4096 processors with $\text{CFL} = 1$ (\drawdiamond{black} \comment{this does not work on my computer, I see a W}), $\text{CFL} = 10$ (\drawsquare{red}), and $\text{CFL} = 100$ (\drawcircle{blue}).
+	\caption{Performance indicators of the first (top row) and last (bottom row) iterations of the semi-Lagrangian algorithm for the advection on 4096 processes with $\text{CFL} = 1$ (\drawdiamond{black} \comment{this does not work on my computer, I see a W} \mohammad{hum ...
+not sure what to do for this one.
+it should be part of latex...}), $\text{CFL} = 10$ (\drawsquare{red}), and $\text{CFL} = 100$ (\drawcircle{blue}).
 Increasing the CFL number greatly causes load imbalance during interpolation (a) and increases the communication volume (c).
-However, the CFL number does not seem to appreciably affect the number of messages sent by the processors (b).
-During the last Semi-Lagrangian iteration, the initial grid $G_0$ (c.f.
+However, the CFL number does not seem to appreciably affect the number of messages sent by the processes (b).
+During the last semi-Lagrangian iteration, the initial grid $G_0$ (c.f.
 algorithm \ref{alg:semi-lagrangian}) is very close to the final grid.
 As a result, the load imbalance is considerably improved (d).
-Curiously, however, the communication pattern does not seem to be change much between first and last iterations (e-f) \comment{e-f ?}.}
+Curiously, however, the communication pattern does not seem to be change much between first and last iterations (e-f) \comment{e-f ?} \mohammad{e and f figures.
+Should we rephrase?}.}
 	\label{fig:communication_4096}
@@
 	}
-	\caption{Detailed load balancing and communication information for the advection test problem for $\text{CFL} = 1$, $\text{CFL} = 10$, and $\text{CFL} = 100$.
-Note how increasing the CFL number causes load imbalance and increases the communication volume while it does not affect the number of messages sent and received during an iteration of the Semi-Lagrangian step.}
+	\caption{Detailed load balancing and communication information for the advection test for $\text{CFL} = 1$, $\text{CFL} = 10$, and $\text{CFL} = 100$.
+Note how increasing the CFL number causes load imbalance and increases the communication volume while it does not affect the number of messages sent and received during an iteration of the semi-Lagrangian step.}
 	\label{tab:communication}

### 2015-01-13 paper/tex/applications.tex
increased teh fonts in all figures; added explicit efficiency tables

@@
 The results are presented in figure \ref{fig:stefan_scaling}.
-As expected from the results obtained for each component in the previous sections, our implementation of the Stefan problem exhibits very satisfactory scaling.
+As expected from the results obtained for each component in the previous sections, our implementation of the Stefan problem exhibits very satisfactory scaling (c.f.\ table \ref{tab:scaling_stefan}).
 
@@
 \centering
-\subfigure[]{\includegraphics[width=0.45\textwidth]{figures/stefan_scaling_small.pdf}}
-\subfigure[]{\includegraphics[width=0.45\textwidth]{figures/stefan_scaling_large.pdf}}
+\subfigure[]{\includegraphics[width=0.48\textwidth]{figures/Stefan_small.pdf}}
+\subfigure[]{\includegraphics[width=0.48\textwidth]{figures/Stefan_large.pdf}}
 \caption{Scalability of the Stefan problem for small (left) and large (right) octrees with roughly 7M and 105M grid points, respectively.
+The solid dashed line represents perfect scaling.
 As expected from the scalability analysis of the individual components, we observe excellent results, illustrating the potential \mohammad{applicability?} of our solver \mohammad{algorithms?}.}
@@
 \end{figure}
+
+\begin{table}
+\centering
+	\begin{tabular}{|l|c|cccccc|}
+	\hline
+	\multirow{2}{*}{Small Test} & $P$ & 16      & 32      & 64      & 128     & 256    & 512 \\ 	                            
+	                            & $e$ & $100\%$ & $95\%$  & $90\%$  & $82\%$  & $73\%$ & $64\%$ \\
+	\hline
+	\multirow{2}{*}{Large Test} & $P$ & 128     & 256     & 512     & 1024    & 2048   & 4096 \\ 	                            
+	                            & $e$ & $100\%$ & $98\%$  & $94\%$  & $89\%$  & $73\%$ & $61\%$ \\
+	\hline
+	\end{tabular}
+	\caption{Parallel efficiency of the total runtime for the stefan test based on the lowest number of processes for each test.}
+	\label{tab:scaling_stefan} 
+\end{table}
 

### 2015-01-13 paper/tex/scaling_results.tex
increased teh fonts in all figures; added explicit efficiency tables

@@
 Excellent scaling is obtained for the small problem for $P = 16-512$ even when $95\%$ of the interpolation points belong to a remote process.
-For the larger problem, however, the communication overhead prevents the algorithm from scaling beyond $4096$ processes when $\alpha = 95\%$.
+For the larger problem, however, the communication overhead prevents the algorithm from scaling beyond $2048$ processes when $\alpha = 95\%$ (c.f\ table \ref{tab:scaling_interpolation}).
 \begin{figure}[htbp]
 	\begin{center}
-		\subfigure[$N_G = 33$M, $\alpha = 5\%$]{\includegraphics[width = 0.45 \textwidth] {figures/interpolation_small_05.pdf}}
-		\subfigure[$N_G = 33$M, $\alpha = 95\%$]{\includegraphics[width = 0.45 \textwidth] {figures/interpolation_small_95.pdf}}
+		\subfigure[$N_G = 33$M, $\alpha = 5\%$]{\includegraphics[width = 0.48 \textwidth] {figures/host_Interpolation_small_alpha_5.pdf}}
+		\subfigure[$N_G = 33$M, $\alpha = 95\%$]{\includegraphics[width = 0.48 \textwidth] {figures/host_Interpolation_small_alpha_95.pdf}}
 		\\
-		\subfigure[$N_G = 280$M, $\alpha = 5\%$]{\includegraphics[width = 0.45 \textwidth] {figures/interpolation_large_05.pdf}}
-		\subfigure[$N_G = 280$M, $\alpha = 95\%$]{\includegraphics[width = 0.45 \textwidth] {figures/interpolation_large_95.pdf}}
+		\subfigure[$N_G = 280$M, $\alpha = 5\%$]{\includegraphics[width = 0.48 \textwidth] {figures/host_Interpolation_large_alpha_5.pdf}}
+		\subfigure[$N_G = 280$M, $\alpha = 95\%$]{\includegraphics[width = 0.48 \textwidth] {figures/host_Interpolation_large_alpha_95.pdf}}
 	\end{center}
@@
 algorithm \ref{alg:interpolation}).
+The black dashed line represents the ideal scaling.
 The results indicate excellent scaling for the small test (a-b) and for the large test when $\alpha = 5\%$ (c).
-For the extreme case (d) the algorithm stops scaling at $4096$ processes due to communication overhead.}
+For the extreme case (d) the algorithm stops scaling at $2048$ processes due to communication overhead.}
 	\label{fig:interpolation}
 \end{figure}
+
+\begin{table}
+\centering
+	\begin{tabular}{|l|l|cccccc|}
+	\hline
+	\multirow{3}{*}{Small Test} & \multicolumn{1}{|c|}{$P$} & 16      & 32      & 64      & 128     & 256    & 512 \\
+	\cline{2-8} 	                            
+	                            & $\alpha = 5\%$   & $100\%$ & $101\%$ & $102\%$ & $106\%$ & $127\%$& $165\%$ \\
+	                            & $\alpha = 95\%$  & $100\%$ & $104\%$ & $110\%$ & $125\%$ & $127\%$& $106\%$ \\ 	                            
+	\hline
+	\multirow{3}{*}{Large Test} & \multicolumn{1}{|c|}{$P$} & 128     & 256     & 512     & 1024    & 2048   & 4096 \\
+	\cline{2-8} 	                            
+	                            & $\alpha = 5\%$   & $100\%$ & $89\%$  & $95\%$  & $101\%$ & $109\%$& $114\%$ \\
+	                            & $\alpha = 95\%$  & $100\%$ & $103\%$ & $108\%$ & $99\%$  & $67\%$ & $34\%$ \\
+	\hline
+	\end{tabular}
+	\caption{Parallel efficiency of the total runtime the interpolation test. Reported efficiencies are based on the lowest number of processes for each test.}
+	\label{tab:scaling_interpolation}
+\end{table}
 
@@
 	\begin{center}
-		\subfigure[semi-Lagrangian, $\text{CFL} = 1$]{\includegraphics[width = 0.3 \textwidth] {figures/SL_Small_1_SemiLagrangian.pdf}}
-		\subfigure[semi-Lagrangian, $\text{CFL} = 10$]{\includegraphics[width = 0.3 \textwidth] {figures/SL_Small_10_SemiLagrangian.pdf}}
-		\subfigure[semi-Lagrangian, $\text{CFL} = 100$]{\includegraphics[width = 0.3 \textwidth] {figures/SL_Small_100_SemiLagrangian.pdf}}
+		\subfigure[semi-Lagrangian, $\text{CFL} = 1$]{\includegraphics[width = 0.32 \textwidth] {figures/host_SemiLagrangian_small_CFL_1.pdf}}
+		\subfigure[semi-Lagrangian, $\text{CFL} = 10$]{\includegraphics[width = 0.32 \textwidth] {figures/host_SemiLagrangian_small_CFL_10.pdf}}
+		\subfigure[semi-Lagrangian, $\text{CFL} = 100$]{\includegraphics[width = 0.32 \textwidth] {figures/host_SemiLagrangian_small_CFL_100.pdf}}
 		\\
-		\subfigure[Interpolation, $\text{CFL} = 1$]{\includegraphics[width = 0.3 \textwidth] {figures/SL_Small_1_Interpolation.pdf}}
-		\subfigure[Interpolation, $\text{CFL} = 10$]{\includegraphics[width = 0.3 \textwidth] {figures/SL_Small_10_Interpolation.pdf}}
-		\subfigure[Interpolation, $\text{CFL} = 100$]{\includegraphics[width = 0.3 \textwidth] {figures/SL_Small_100_Interpolation.pdf}}
+		\subfigure[Interpolation, $\text{CFL} = 1$]{\includegraphics[width = 0.32 \textwidth] {figures/host_SemiLagrangian_Interpolation_small_CFL_1.pdf}}
+		\subfigure[Interpolation, $\text{CFL} = 10$]{\includegraphics[width = 0.32 \textwidth] {figures/host_SemiLagrangian_Interpolation_small_CFL_1.pdf}}
+		\subfigure[Interpolation, $\text{CFL} = 100$]{\includegraphics[width = 0.32 \textwidth] {figures/host_SemiLagrangian_Interpolation_small_CFL_1.pdf}}
 	\end{center}
@@
 Bottom row: breakdown of the various components of the interpolation phase for the same CFL numbers.
+The solid dashed line represents the ideal scaling.
 Note that the maximum time has been scaled by the number of iterations required to build the tree (c.f.
@@
 	\begin{center}
-		\subfigure[semi-Lagrangian, $\text{CFL} = 1$]{\includegraphics[width = 0.3 \textwidth] {figures/SL_Large_1_SemiLagrangian.pdf}}
-		\subfigure[semi-Lagrangian, $\text{CFL} = 10$]{\includegraphics[width = 0.3 \textwidth] {figures/SL_Large_10_SemiLagrangian.pdf}}
-		\subfigure[semi-Lagrangian, $\text{CFL} = 100$]{\includegraphics[width = 0.3 \textwidth] {figures/SL_Large_100_SemiLagrangian.pdf}}
+		\subfigure[semi-Lagrangian, $\text{CFL} = 1$]{\includegraphics[width = 0.32 \textwidth] {figures/host_SemiLagrangian_large_CFL_1.pdf}}
+		\subfigure[semi-Lagrangian, $\text{CFL} = 10$]{\includegraphics[width = 0.32 \textwidth] {figures/host_SemiLagrangian_large_CFL_10.pdf}}
+		\subfigure[semi-Lagrangian, $\text{CFL} = 100$]{\includegraphics[width = 0.32 \textwidth] {figures/host_SemiLagrangian_large_CFL_100.pdf}}
 		\\
-		\subfigure[Interpolation, $\text{CFL} = 1$]{\includegraphics[width = 0.3 \textwidth] {figures/SL_Large_1_Interpolation.pdf}}
-		\subfigure[Interpolation, $\text{CFL} = 10$]{\includegraphics[width = 0.3 \textwidth] {figures/SL_Large_10_Interpolation.pdf}}
-		\subfigure[Interpolation, $\text{CFL} = 100$]{\includegraphics[width = 0.3 \textwidth] {figures/SL_Large_100_Interpolation.pdf}}
+		\subfigure[Interpolation, $\text{CFL} = 1$]{\includegraphics[width = 0.32 \textwidth] {figures/host_SemiLagrangian_Interpolation_large_CFL_1.pdf}}
+		\subfigure[Interpolation, $\text{CFL} = 10$]{\includegraphics[width = 0.32 \textwidth] {figures/host_SemiLagrangian_Interpolation_large_CFL_10.pdf}}
+		\subfigure[Interpolation, $\text{CFL} = 100$]{\includegraphics[width = 0.32 \textwidth] {figures/host_SemiLagrangian_Interpolation_large_CFL_100.pdf}}
 	\end{center}
@@
 \end{figure}
+
+\begin{table}
+\centering
+	\begin{tabular}{|l|l|cccccc|}
+	\hline
+	\multirow{4}{*}{Small Test} & \multicolumn{1}{|c|}{$P$} & 16      & 32      & 64      & 128     & 256    & 512 \\
+	\cline{2-8} 	                            
+	                            & $\text{CFL} = 1$   & $100\%$ & $88\%$  & $84\%$  & $82\%$  & $74\%$ & $67\%$ \\
+	                            & $\text{CFL} = 10$  & $100\%$ & $99\%$  & $104\%$ & $88\%$  & $80\%$ & $71\%$ \\ 	                            
+	                            & $\text{CFL} = 100$ & $100\%$ & $95\%$  & $90\%$  & $84\%$  & $67\%$ & $49\%$ \\
+	\hline
+	\multirow{4}{*}{Large Test} & \multicolumn{1}{|c|}{$P$} & 128     & 256     & 512     & 1024    & 2048   & 4096 \\
+	\cline{2-8} 	                            
+	                            & $\text{CFL} = 1$   & $100\%$ & $94\%$  & $87\%$  & $82\%$  & $75\%$ & $65\%$ \\
+	                            & $\text{CFL} = 10$  & $100\%$ & $90\%$  & $92\%$  & $86\%$  & $79\%$ & $63\%$ \\
+	                            & $\text{CFL} = 100$ & $100\%$ & $94\%$  & $90\%$  & $84\%$  & $70\%$ & $57\%$ \\
+	\hline
+	\end{tabular}
+	\caption{Parallel efficiency of the runtime of a single semi-Lagrangian step. Reported efficiencies are based on the lowest number of processes for each test.}
+	\label{tab:scaling_semilagrangian}
+\end{table}
 
@@
 	\begin{center}
-		\subfigure[]{\includegraphics[width = 0.3 \textwidth] {figures/point_first_4096.pdf}  \label{fig:communication_4096_pf}}
-		\subfigure[]{\includegraphics[width = 0.3 \textwidth] {figures/message_first_4096.pdf}\label{fig:communication_4096_mf}}
-		\subfigure[]{\includegraphics[width = 0.3 \textwidth] {figures/volume_first_4096.pdf} \label{fig:communication_4096_vf}}
+		\subfigure[]{\includegraphics[width = 0.3 \textwidth] {figures/host_large_point_first_4096.pdf}  \label{fig:communication_4096_pf}}
+		\subfigure[]{\includegraphics[width = 0.3 \textwidth] {figures/host_large_message_first_4096.pdf}\label{fig:communication_4096_mf}}
+		\subfigure[]{\includegraphics[width = 0.3 \textwidth] {figures/host_large_volume_first_4096.pdf} \label{fig:communication_4096_vf}}
 		\\
-		\subfigure[]{\includegraphics[width = 0.3 \textwidth] {figures/point_last_4096.pdf}   \label{fig:communication_4096_pl}}
-		\subfigure[]{\includegraphics[width = 0.3 \textwidth] {figures/message_last_4096.pdf} \label{fig:communication_4096_ml}}
-		\subfigure[]{\includegraphics[width = 0.3 \textwidth] {figures/volume_last_4096.pdf}  \label{fig:communication_4096_vl}}		
+		\subfigure[]{\includegraphics[width = 0.3 \textwidth] {figures/host_large_point_last_4096.pdf}   \label{fig:communication_4096_pl}}
+		\subfigure[]{\includegraphics[width = 0.3 \textwidth] {figures/host_large_message_last_4096.pdf} \label{fig:communication_4096_ml}}
+		\subfigure[]{\includegraphics[width = 0.3 \textwidth] {figures/host_large_volume_last_4096.pdf}  \label{fig:communication_4096_vl}}		
 	\end{center}
@@
 In both cases the forest is built on a $3\times3\times3$ macro-mesh.
-Figure \ref{fig:reinit} illustrates that our reinitialization algorithm, and in particular the overlapping strategy presented in algorithm \ref{alg:overlap}, scales very well.
+Figure \ref{fig:reinit} illustrates that our reinitialization algorithm, and in particular the overlapping strategy presented in algorithm \ref{alg:overlap}, scales very well (c.f.\ table \ref{tab:scaling_reinit}).
 In general we expect similar scaling results for any local, finite-difference based calculations on Octrees that can efficiently utilize algorithm \ref{alg:overlap}.
@@
 \centering
-\subfigure[]{\includegraphics[width=0.45\textwidth]{figures/Reinit_Small.pdf}}
-\subfigure[]{\includegraphics[width=0.45\textwidth]{figures/Reinit_Large.pdf}}
+\subfigure[]{\includegraphics[width=0.48\textwidth]{figures/Reinit_Small.pdf}}
+\subfigure[]{\includegraphics[width=0.48\textwidth]{figures/Reinit_Large.pdf}}
 \caption{Scalability of the reinitialization test for a small (left) and large (right) Octree with roughly 21M and 337M grid points, respectively.
+The black dashed line represents ideal scaling.
 Excellent results are obtained in both cases, illustrating the scalability of the overlapping strategy (c.f.
@@
 \end{figure}
+
+\begin{table}
+\centering
+	\begin{tabular}{|l|c|cccccc|}
+	\hline
+	\multirow{2}{*}{Small Test} & $P$ & 16      & 32      & 64      & 128     & 256    & 512 \\ 	                            
+	                            & $e$ & $100\%$ & $115\%$ & $110\%$ & $105\%$ & $96\%$ & $80\%$ \\
+	\hline
+	\multirow{2}{*}{Large Test} & $P$ & 128     & 256     & 512     & 1024    & 2048   & 4096 \\ 	                            
+	                            & $e$ & $100\%$ & $95\%$ & $95\%$ & $89\%$ & $82\%$ & $67\%$ \\
+	\hline
+	\end{tabular}
+	\caption{Parallel efficiency of the total runtime for the reinitialization test based on the lowest number of processes for each test.}
+	\label{tab:scaling_reinit}
+\end{table}

### 2015-01-13 paper/tex/data/3d_interpolation/parse_interpolation.m
Added stdout and parsers for all scaling graphs and tables

@@ -1,0 +1,143 @@
+set(0,'defaulttextinterpreter','latex')
+set(0,'DefaultTextFontname', 'CMU Serif')
+set(0,'DefaultAxesFontName', 'CMU Serif')
+
+type = 'host';
+run  = 'large';
+alpha = 5;
+
+folder = strcat(type, '/', run, '/alpha_', num2str(alpha/100));
+files = {
+%     'stdout.n_16';
+%     'stdout.n_32';
+%     'stdout.n_64';
+    'stdout.n_128';
+    'stdout.n_256';
+    'stdout.n_512';
+    'stdout.n_1024';
+    'stdout.n_2048';
+    'stdout.n_4096';
+    };
+
+modes = {'ro-', 'bs-', 'k>-', 'md-', 'ro--', 'bs--', 'k>--', 'md--'};
+faces = {'r', 'b', 'k', 'm', 'r', 'b', 'k', 'm'};
+events = {
+    'log_interpolation_all'
+    'log_interpolation_add_points'
+    'InterpolatingFunctionHost::process_local'  
+    'InterpolatingFunctionHost::process_queries'
+    'InterpolatingFunctionHost::process_replies'
+};
+
+%% Total time:
+t = zeros(1, length(files));
+p = zeros(1, length(files));
+
+log_begin = '---------------------------------------------- PETSc Performance Summary: ----------------------------------------------';
+flag = 0;
+
+for i=1:length(files)
+    fid = fopen(strcat(folder, '/', files{i}));
+    
+    while ~feof(fid)
+        line = fgetl(fid);
+        if flag
+            line = fgetl(fid);
+            p(i) = sscanf(line, '%*s %*s %*s %*s %*s %*s %*s %d %*[^\n]');
+            flag = 0;
+        end
+        if strfind(line, log_begin)
+            flag = 1;
+        end
+        if strfind(line, 'Time (sec):');
+            t(i) = sscanf(line, '%*s %*s %e %*[^\n]');
+            break
+        end
+    end
+    
+    fclose(fid);
+    
+end
+
+%% separate events
+t = zeros(length(events), length(files));
+% p = zeros(1, length(files));
+for j=1:length(events)
+    event = events{j};
+    
+    log_begin = '---------------------------------------------- PETSc Performance Summary: ----------------------------------------------';
+    flag = 0;
+    
[…]

### 2015-01-15 paper/tex/applications.tex
added time evolution figures for stefan problem

@@
 
-We propose to apply \mohammad{what about: In this section we apply ...} our approach to the study of the phase transition of a liquid melt to a solid crystaline structure.
+In this section we apply our approach to the study of the phase transition of a liquid melt to a solid crystaline structure.
 In the case of a single component melt, and in the absence of convection, the process is dominated by diffusion and can be modeled as a Stefan problem.
-We decompose the computational domain $\Omega$ into \mohammad{How about: Consider a computational domain $\Omega$ decomposed into ...} two subdomains $\Omega^-$ and $\Omega^+$, separated by an interface $\Gamma$.
+We decompose the computational domain $\Omega$ into \mohammad{How about: Consider a computational domain $\Omega$ decomposed into ...} \comment{this would be the exact phrasing of another article} two subdomains $\Omega^-$ and $\Omega^+$, separated by an interface $\Gamma$.
 The Stefan problem describes the evolution of the temperature $T$, decomposed into $T_s$ in the solid phase $\Omega^+$ and $T_l$ in the liquid phase $\Omega^-$, as
@@
 The solid dashed line represents perfect scaling.
-As expected from the scalability analysis of the individual components, we observe excellent results, illustrating the potential \mohammad{applicability?} of our solver \mohammad{algorithms?}.}
+As expected from the scalability analysis of the individual components, we observe excellent results, illustrating the potential of our algorithms.}
 \label{fig:stefan_scaling}
@@
 	\end{tabular}
-	\caption{Parallel efficiency of the total runtime for the stefan test based on the lowest number of processes for each test.}
+	\caption{Parallel efficiency of the total runtime for the Stefan test based on the lowest number of processes for each test. \comment{frederic, what do you think about this table ?}}
 	\label{tab:scaling_stefan} 
@@
 \mohammad{too many transitive statements and keyword ``choose''. Can we rephrase this?}
-We now present the results from a large simulation of the Stefan problem.
-We choose to work with a $20\times20\times20$ macro-mesh and with level-10 octrees.
+We now present the results from a large simulation of the Stefan problem on a $20\times20\times20$ macro-mesh and with level-10 Octrees.
 We choose to define the Gibbs-Tompson anisotropy undercooling coefficients \eqref{eq:stefan_gibbs_tompson} as
@@
 \end{align*}
-Note that these expressions are arbitrary and lead to very varied crystal shapes \mohammad{How about: ...
-lead to a variety of shapes?}.
+Note that these expressions are arbitrary and lead to a variety of shapes.
 The computation is initialized with twenty spherical seeds of radius $0.0015$ placed randomly in the domain.
@@
 
-We let the simulation run with \mohammad{How about: The simulation is ran on ...} $256$ mpi processes for $6$ hours and $30$ minutes, resulting in $396$ time iterations.
-Visualizations of the final iteration are presented in figures \ref{fig:stefan_grid} and \ref{fig:stefan_crystals}.
+The simulation was ran on $256$ mpi processes for $6$ hours and $30$ minutes, resulting in $396$ time iterations.
+Visualizations of the final iteration are presented in figures \ref{fig:stefan_grid}, \ref{fig:stefan_crystals} and \ref{fig:stefan_evolution}.
 The final iteration of the simulation consisted of $167$M grid points when a uniform grid with the equivalent finest resolution would lead to $8.59\cdot10^{12}$ grid points, i.e.
@@
 \subfigure[]{\includegraphics[width=0.45\textwidth]{figures/stefan_grid_zoom.png}}
-\caption{Visualization of the computational mesh (a,b,c) and the temperature field (b) for the Stefan problem simulation.} \label{fig:stefan_grid}
+\caption{Visualization of the computational mesh (a, c, d) and the temperature field (b) for the Stefan problem simulation.} \label{fig:stefan_grid}
 \end{center}
@@
 \end{figure}
+
+\begin{figure}[ht!]
+\begin{center}
+\subfigure[]{
+\includegraphics[width=.2\textwidth]{stefan/nb2_iter19.png}
+\includegraphics[width=.2\textwidth]{stefan/nb2_iter39.png}
+\includegraphics[width=.2\textwidth]{stefan/nb2_iter59.png}
+\includegraphics[width=.2\textwidth]{stefan/nb2_iter79.png}
+}
+\subfigure[]{
+\includegraphics[width=.2\textwidth]{stefan/nb5_iter19.png}
+\includegraphics[width=.2\textwidth]{stefan/nb5_iter39.png}
+\includegraphics[width=.2\textwidth]{stefan/nb5_iter59.png}
+\includegraphics[width=.2\textwidth]{stefan/nb5_iter79.png}
+}
+\subfigure[]{
+\includegraphics[width=.2\textwidth]{stefan/nb8_iter19.png}
+\includegraphics[width=.2\textwidth]{stefan/nb8_iter39.png}
+\includegraphics[width=.2\textwidth]{stefan/nb8_iter59.png}
+\includegraphics[width=.2\textwidth]{stefan/nb8_iter79.png}
+}
+\subfigure[]{
+\includegraphics[width=.2\textwidth]{stefan/nb12_iter19.png}
+\includegraphics[width=.2\textwidth]{stefan/nb12_iter39.png}
+\includegraphics[width=.2\textwidth]{stefan/nb12_iter59.png}
+\includegraphics[width=.2\textwidth]{stefan/nb12_iter79.png}
+}
+\caption{Time evolution of four of the crystals obtained for the Stefan problem simulation. The snapshots represent, from left to right, iterations 96, 196, 296 and 396.} \label{fig:stefan_evolution}
+\end{center}
+\end{figure}

### 2015-01-15 paper/tex/parallel_algorithms.tex
added time evolution figures for stefan problem

@@
 only the leaves are explicitly stored.
-However, \textcolor{blue}{explicit knowledge of the hierarchal structure of tree is greatly beneficial in several algorithms, e.g.\ in search operations needed for the interpolation algorithm.} Thus, we introduce a simple reconstruction algorithm which recreates a local representation of the entire ``forest'' that is only adapted to local cells and, potentially, the ghost layer.
+However, \textcolor{blue}{explicit knowledge of the hierarchal structure of the tree is greatly beneficial in several algorithms, e.g.\ in search operations needed for the interpolation algorithm.} Thus, we introduce a simple reconstruction algorithm which recreates a local representation of the entire ``forest'' that is only adapted to local cells and, potentially, the ghost layer.
 This approach is similar to the ideas introduced in \cite{Bangerth;Burstedde;Heister;etal:11:Algorithms-and-data-} and our tests show that in a typical application they usually amount to less that $1\%$ of the entire runtime.
@@
 Although parallelization of the interpolation process on a shared-memory machine is trivial, the same cannot be said for distributed-memory machines.
-In fact, the parallel interpolation algorithm \ref{alg:interpolation} is probably the most important contribution of this article \comment{but hasn't it been done before ?} \textcolor{blue}{Mohammad: not on adaptive since that's what makes it complicated}.
+In fact, the parallel interpolation algorithm \ref{alg:interpolation} is probably the most important contribution of this article \comment{but hasn't it been done before ?} \textcolor{blue}{Mohammad: not on adaptive since that's what makes it complicated} \comment{then maybe explain in a sentence that it's easy on uniform because it's easy to tell which process the departure point belongs to ?}.
 Complication arises because not all departure points will reside in the domain of current process.

### 2015-01-28 paper/tex/abstract.tex
added time evolution figures for stefan problem
added frederic's corrections. Some points still need to be discussed

--- 2014-08-27 paper/tex/abstract.tex
@@
 \begin{abstract}
-Abstract will be filled in after finalization of the paper.
+We present scalable parallel algorithms for the level-set technology on adaptive Quadtree and Octree Cartesian grids.
+The algorithms are based on a domain decomposition technique and implemented using \texttt{MPI} and the open-source \texttt{p4est} library.
+An important contribution is a scalable parallel semi-Lagrangian method which, similar to its serial implementation, is free of any time-step restrictions.
+This is achieved by introducing a scalable global interpolation scheme on adaptive Quadtree and Octree grids.
+Moreover, we present a simple parallel reinitialization scheme using the pseudo-time transient formulation.
+Both parallel algorithms scale on the Stampede supercomputer, where we are limited to 4096 cores at most.
+Finally a relevant application of the algorithms is presented in modeling the crystallization phenomena by solving a Stefan problem, illustrating a level of detailed calculation that would be impossible without a parallel adaptive strategy.
+We believe that the algorithms presented in this article will be of interest and useful to researchers working with the level-set framework and modeling multi-scale physics in general.
 \end{abstract}
-
 \begin{keyword}
-Level-set Method \sep Adaptive Grids \sep Parallel Computing \sep Oct-/Quadtree datastructures
+Quadtree/Octree Grids \sep Parallel Computing \sep Space Filling Curves \sep Semi-Lagrangian Method \sep Level-set Method
 \end{keyword}
-
 \maketitle

### 2015-01-28 paper/tex/applications.tex
added frederic's corrections. Some points still need to be discussed

@@
 %!TEX root = draft.tex
-\section{Application to Stefan Problem} \label{sec:application}
+\section{Application to the Stefan Problem} \label{sec:application}
 
@@
 In the case of a single component melt, and in the absence of convection, the process is dominated by diffusion and can be modeled as a Stefan problem.
-We decompose the computational domain $\Omega$ into \mohammad{How about: Consider a computational domain $\Omega$ decomposed into ...} \comment{this would be the exact phrasing of another article} two subdomains $\Omega^-$ and $\Omega^+$, separated by an interface $\Gamma$.
-The Stefan problem describes the evolution of the temperature $T$, decomposed into $T_s$ in the solid phase $\Omega^+$ and $T_l$ in the liquid phase $\Omega^-$, as
+We decompose the computational domain $\Omega$ into \mohammad{How about: Consider a computational domain $\Omega$ decomposed into ...} \comment{this would be the exact phrasing of another article} two subdomains $\Omega_l$ and $\Omega_s$, separated by an interface $\Gamma$.
+The Stefan problem describes the evolution of the temperature $T$, decomposed into $T_s$ in the solid phase $\Omega_s$ and $T_l$ in the liquid phase $\Omega_l$, as
 \begin{align}
-\pd{T_l}{t} & = D_l \Delta T_l \quad \mathrm{in} ~~ \Omega^-, \label{eq:stefan_heat_equation_l} \\
-\pd{T_s}{t} & = D_s \Delta T_s \quad \mathrm{in} ~~ \Omega^+. \label{eq:stefan_heat_equation_s}
+\pd{T_l}{t} & = D_l \Delta T_l \quad \mathrm{in} ~~ \Omega_l, \label{eq:stefan_heat_equation_l} \\
+\pd{T_s}{t} & = D_s \Delta T_s \quad \mathrm{in} ~~ \Omega_s. \label{eq:stefan_heat_equation_s}
 \end{align}
 The diffusion constants $D_l$ and $D_s$ can be discontinuous across the interface.
-We prescribe homogeneous Neumann boundary conditions on the edge of the computational domain, $\nabla T \cdot \mathbf{n}\vert_{\partial \Omega}=0$.
-At the interface between the solid and the liquid phases, the temperature is given by the Gibbs-Tompson boundary condition \cite{Alexiades;Solomon;Wilson:88:The-formation-of-a-s, Alexiades;Solomon:93:Mathematical-Modelin}
+We prescribe homogeneous Neumann boundary conditions on the edge of the computational domain, $\nabla T \cdot \underline{\mathbf{n}}\vert_{\partial \Omega}=0$.
+At the interface between the solid and the liquid phases, the temperature is given by the Gibbs-Tompson boundary condition \cite{Alexiades;Solomon;Wilson:88:The-formation-of-a-s, Alexiades;Solomon:93:Mathematical-Modelin}:
 \begin{equation} \label{eq:stefan_gibbs_tompson}
-T_s = T_l = T_{\Gamma} = -\epsilon_c \kappa - \epsilon_v (\mathbf{V} \cdot \mathbf{n}),
+T_s = T_l = T_{\Gamma} = -\epsilon_c \kappa - \epsilon_v (\underline{\mathbf{V}} \cdot \underline{\mathbf{n}}),
 \end{equation}
-where $\kappa$ is the local interface curvature, $V$ is the velocity of the interface, $\mathbf{n}$ is the outward normal to the solidification front and $\epsilon_c$ and $\epsilon_v$ are the curvature and kinetic undercooling coefficients.
-The interface velocity $V$ is defined from the jump in the heat flux at the interface,
+where $\kappa$ is the local interface curvature, $\underline{\mbf{V}}$ is the velocity of the interface, $\underline{\mathbf{n}}$ is the outward normal to the solidification front and $\epsilon_c$ and $\epsilon_v$ are the surface tension and kinetic undercooling coefficients.
+The interface velocity $\underline{\mbf{V}}$ is defined from the jump in the heat flux at the interface,
 \begin{equation} \label{eq:stefan_velocity}
-(\mathbf{V} \cdot \mathbf{n}) = - \left[ D_l \pd{T_l}{\mathbf{n}} - D_s \pd{T_s}{\mathbf{n}} \right].
+(\underline{\mathbf{V}} \cdot \underline{\mathbf{n}}) = - \left[ D_l \pd{T_l}{\underline{\mathbf{n}}} - D_s \pd{T_s}{\underline{\mathbf{n}}} \right].
 \end{equation}
@@
 \begin{equation} \label{eq:stefan_dt}
-\Delta t = 5 ~ \Delta x_{min} ~ \min(1,1/\max \lVert \mathbf{V} \rVert),
+\Delta t = 5 ~ \Delta x_{min} ~ \min(1,1/\max \lVert \underline{\mathbf{V}} \rVert),
 \end{equation}
-where $\Delta x_{min}$ is the size of the smallest cell of the forest. The general procedure to solve the Stefan problem is presented in algorithm \ref{alg:stefan}.
+where $\Delta x_{min}$ is the size of the smallest cell of the forest. The general procedure to solve the Stefan problem is presented in Algorithm \ref{alg:stefan}.
 
@@
 
-The implementation of the Stefan problem relies on the components described in the previous sections, and it is therefore a good summary of the performance of the various algorithms.
-We monitored the performance of the code over five time iterations, as presented in algorithm \ref{alg:stefan}, for two different maximum resolutions.
+The implementation of the Stefan problem relies on the components described in the previous sections, and it is therefore a good synthesis of the performance of the various algorithms.
+We monitored the performance of the code over five time iterations, as presented in Algorithm \ref{alg:stefan}, for two different maximum resolutions.
 In both cases, the forest is built on a $20\times20\times20$ macro-mesh.
@@
 \subfigure[]{\includegraphics[width=0.48\textwidth]{figures/Stefan_large.pdf}}
-\caption{Scalability of the Stefan problem for small (left) and large (right) octrees with roughly 7M and 105M grid points, respectively.
+\caption{Scalability of the Stefan problem for small (left) and large (right) Octrees with roughly 7M and 105M grid points, respectively.
 The solid dashed line represents perfect scaling.
@@
 \subsection{Numerical experiments}
-\mohammad{too many transitive statements and keyword ``choose''. Can we rephrase this?}
 We now present the results from a large simulation of the Stefan problem on a $20\times20\times20$ macro-mesh and with level-10 Octrees.
-We choose to define the Gibbs-Tompson anisotropy undercooling coefficients \eqref{eq:stefan_gibbs_tompson} as
+The Gibbs-Tompson anisotropy undercooling coefficients in equation \eqref{eq:stefan_gibbs_tompson} are defined as
 \begin{align*}
@@
 \end{align*}
-with $\theta_1$ the angle between the normal to the interface $\mathbf{n}$ and the x-axis in the $(x,y)$ plane, $\theta_2$ the angle between $\mathbf{n}$ and the x-axis in the $(x,z)$ plane, and
+with $\theta_1$ the angle between the normal to the interface $\underline{\mathbf{n}}$ and the x-axis in the $(x,y)$ plane and $\theta_2$ the angle between $\underline{\mathbf{n}}$ and the x-axis in the $(x,z)$ plane.
+The coefficients
 \begin{align*}
 \epsilon_1 & = 2~(\sin(x)+\cos(y)+2)\cdot10^{-6}, & \epsilon_2 & = 2~(\sin(x)+\cos(z)+2)\cdot10^{-6}, \\
-\alpha_1 & = \frac{1}{4}(\cos(x)+\sin(y)+2), & \alpha_2 & = \frac{1}{4}(\cos(x)+\sin(z)+2).
+\alpha_1 & = \frac{1}{4}(\cos(x)+\sin(y)+2), & \alpha_2 & = \frac{1}{4}(\cos(x)+\sin(z)+2),
 \end{align*}
-Note that these expressions are arbitrary and lead to a variety of shapes.
-The computation is initialized with twenty spherical seeds of radius $0.0015$ placed randomly in the domain.
-We choose $D_s=D_l=1$ and set the initial temperatures $T^0_l=-0.25$ and $T^0_s=0$.
+are used to enforce a variety of crystal shapes.
+The computation is initialized with twenty spherical seeds of radius $1.5 \cdot 10^{-3}$ placed randomly in the domain.
+We take the diffusion coefficients $D_s=D_l=1$ and set the initial temperatures $T^0_l=-0.25$ and $T^0_s=0$.
 
-The simulation was ran on $256$ mpi processes for $6$ hours and $30$ minutes, resulting in $396$ time iterations.
+The simulation was ran on $256$ MPI processes for $6$ hours and $30$ minutes, resulting in $396$ time iterations.
 Visualizations of the final iteration are presented in figures \ref{fig:stefan_grid}, \ref{fig:stefan_crystals} and \ref{fig:stefan_evolution}.
-The final iteration of the simulation consisted of $167$M grid points when a uniform grid with the equivalent finest resolution would lead to $8.59\cdot10^{12}$ grid points, i.e.
-over eight trillion grid points !
-Our simulation used only $0.001948\%$ \mohammad{I think if we are going to report percentage, we should trim it to 1st digit} of the number of grid points needed for the same simulation on a uniform grid.
-This application demonstrates the potential of our approach.
+The final iteration of the simulation consisted of $167$M grid points whereas a uniform grid with the equivalent finest resolution would lead to $8.59\cdot10^{12}$ grid points, i.e.
+over eight trillion grid points.
+Our simulation used only $0.002\%$ of the number of grid points needed for the same simulation on a uniform grid.
+This application demonstrates the ability of our approach to resolve small scale details, while accounting for long range interactions.
 
@@
 \subfigure[]{\includegraphics[width=0.45\textwidth]{figures/stefan_crystal_3.png}}
-\caption{Representation of the computational domain (a) and of various crystal observed for the Stefan problem simulation.} \label{fig:stefan_crystals}
+\caption{Representation of the computational domain (a) as well as various developped crystal structures obtained from solving the Stefan problem simulation (b,c,d).} \label{fig:stefan_crystals}
 \end{center}

### 2015-01-28 paper/tex/conclusion.tex
added frederic's corrections. Some points still need to be discussed

@@
 \section{Conclusions}
-In this article we have presented parallel algorithms for the advection and reinitialization of the level-set functions on adaptive Quadtree and Octree grids using domain decomposition approach.
+In this article we have presented parallel algorithms related to the level-set technology on adaptive Quadtree and Octree grids using a domain decomposition approach.
 These algorithms are implemented using a combination of \texttt{MPI} and the open-source \texttt{p4est} library.
-An important feature of the semi-Lagrangian method is its unconditional stability property which must be preserved in parallel, requiring a parallel interpolation scheme.
-However, due to the non-uniform distribution of departure points, scalable implementation of such algorithm is not trivial on adaptive Quadtree and Octree grids.
-An asynchronous interpolation algorithm is presented using non-blocking point-to-point communications which proves good scalability.
+In order to preserve the unconditional stability property of the semi-Lagrangian scheme while enabling scalable computations, we introduced an asynchronous interpolation algorithm using non-blocking point-to-point communications, and demonstrated its scalability.
 
-The scalability of the semi-Lagrangian algorithm, however, depends on the CFL number.
-Great scalability is observed for intermediate CFL numbers while, e.g.
+In particular we showed that the scalability of the semi-Lagrangian algorithm, depends on the CFL number.
+Great scalability is observed for intermediate CFL numbers, e.g.
 $\text{CFL}\sim10$.
-At higher CFL numbers, however, the departure points are potentially further dispersed across processors which limits the scalability.
+At higher CFL numbers, however, the departure points are potentially further dispersed across processors, which limits the scalability.
 This is because the domain decomposition technique used here is based on the Z-ordering of cells and does not take the velocity field information into account.
-A possible remedy for this problem could be assigning weights to cells based on some estimate of what the grid should look like after one step of the advection algorithm, e.g.
+A possible remedy for this problem could be assigning weights to cells based on some estimate of the grid structure after one step of the advection algorithm, e.g.
 by using a forward-in-time integration of grid points.
@@
 Finally, an application of these algorithms is presented in modeling the solidification process by solving a Stefan problem.
-This application clearly illustrates the applicability of our algorithms to complex multi-scale problems that cannot be treated using the normal domain decomposition techniques on uniform grids.
-We believe that our findings would be of interest to other researchers who use the level-set framework for similar complex and multi-scale problems.
+This application clearly illustrates the applicability of our algorithms to complex multi-scale problems that cannot be treated practically using the domain decomposition techniques on uniform grids.
+We believe that our findings can serve as a basis to simulating a wide range of multi-scale problems and free body problems.
 

### 2015-01-28 paper/tex/introduction.tex
added frederic's corrections. Some points still need to be discussed

@@
 The level-set method, originally proposed by Sethian and Osher \cite{Osher;Sethian:88:Fronts-Propagating-w}, is a popular and powerful framework for tracking arbitrary interfaces that undergo complicated topological changes.
-As a result, the level-set method has wide range of application such as in multiphase flows, moving boundary problems, image segmentation, and computer graphics \cite{Osher;Fedkiw:01:Level-Set-Methods:-A,Sethian:99:Level-set-methods-an}.
-An important feature of this method is that the location of interface is defined implicitly on an underlying grid.
+As a result, the level-set method has been used to a wide range of applications such as multiphase flows, image segmentation, and computer graphics \cite{Osher;Fedkiw:02:Level-Set-Methods-an,Sethian:99:Level-set-methods-an}.
+An important feature of this method is that the location of the interface is defined implicitly on an underlying grid.
 This convenience, however, comes at a price.
 First, compared to an explicit method, e.g.\ front tracking \cite{Juric:96:A-Front-Tracking-Met, Tryggvason;Bunner;Esmaeeli;etal:01:A-Front-Tracking-Met}, the level-set method is typically less accurate and mass conservation could be a problem; although progress has been made in resolving this issue \cite{Enright;Fedkiw;Ferziger;etal:02:A-Hybrid-Particle-Le}.
-Second, the level-set function has to be defined in a higher dimensional space compared to the interface.
+Second, the level-set function has to be defined in a higher dimensional space that that of the interface.
 If only the location of the interface is needed, the added dimension greatly increases the overall computational cost.
@@
 Another approach that can address both problems is the use of local grid refinement.
-In \cite{Strain:99:Tree-Methods-for-Mov} the idea of using tree-based grids for level-set calculations was first introduced and later extended in \cite{Losasso;Gibou;Fedkiw:04:Simulating-Water-and} for graphics application.
+In \cite{Strain:99:Tree-Methods-for-Mov} the idea of using tree-based grids for level-set calculations was first introduced and later extended in \cite{Popinet:03:Gerris:-A-Tree-Based, Losasso;Gibou;Fedkiw:04:Simulating-Water-and} for fluid simulations.
 More recently, authors in \cite{Min;Gibou:07:A-second-order-accur} proposed second-order accurate level-set methods on Quadtree (two spatial dimensions) and Octree (three spatial dimensions) grids.
-The use of adaptive tree-base grids in the context of the level-set method is quite advantageous because: 1) It gives fine-grain control over errors, which typically occur close to the interface and 2) It can effectively reduce the dimensionality of the problem by focusing most of the grid cells close to the interface.
+The use of adaptive tree-base grids in the context of the level-set method is quite advantageous because (i) it gives fine-grain control over errors, which typically occur close to the interface and (ii) it can effectively reduce the dimensionality of the problem by focusing most of the grid cells close to the interface.
 Fortunately, constructing the tree is quite simple in the presence of an interface that naturally defines an ideal metric for refinement.
-Although the use of adaptive grids can dramatically reduce the computational cost, performing high-resolution three dimensional calculations of complex interfacial problems, e.g.\ crystal growth in binary alloys \cite{Theillard;Gibou;Pollock:14:A-Sharp-Computationa}, could take a very long time in serial.
-In this paper we extend these algorithms by proposing parallel algorithms for distributed memory machines using a domain decomposition technique.
+However, even though the use of adaptive grids can dramatically reduce the computational cost, performing high-resolution three dimensional calculations of complex interfacial problems, e.g.\ crystal growth in binary alloys \cite{Theillard;Gibou;Pollock:14:A-Sharp-Computationa}, could be prohibitively expensive or even impossible on a serial machine.
+In this paper we extend the level-set technology on Quad-/Oc-tree by proposing parallel algorithms for distributed memory machines using a domain decomposition technique.
 
-One of the main challenges in parallelizing level-set algorithms on adaptive grids is handling the grid itself.
+One of the main challenges in parallelizing algorithms on adaptive grids is handling the grid itself.
 One option is to replicate the entire grid on each process and use off-the-shelf graph partitioners, e.g.\ ParMetis \cite{Karypis;Kumar:98:A-parallel-algorithm} or Zoltan \cite{Boman;Catalyurek;Chevalier;etal:12:The-Zoltan-and-Isorr}, for load balancing and domain decomposition.
 For instance, this was the approach originally taken by the \texttt{deal.II} library \cite{Bangerth;Hartmann;Kanschat:07:deal.II----a-General}.
-This approach, however, is only scalable to a few hundred processes at best and is limited by the size of the grid itself.
+This approach, however, is only scalable to a few hundred processes at best and is limited by the size of the grid itself that can fit in memory.
 Moreover, the use of a general-purpose graph partitioner adds extra overhead that can limit the overall scalability even further.
-Interestingly, tree-based grids have nice spatial ordering that naturally leads to the concept of space-filling curves (SFCs) which can be efficiently exploited for parallel load balancing \cite{Aluru;Sevilgen:97:Parallel-domain-deco,Campbell;Devine;Flaherty;etal:03:Dynamic-octree-load-}.
+Interestingly, tree-based grids have nice spatial ordering that naturally leads to the concept of space-filling curves (SFCs) that can be efficiently exploited for parallel load balancing \cite{Aluru;Sevilgen:97:Parallel-domain-deco,Campbell;Devine;Flaherty;etal:03:Dynamic-octree-load-}.
 
-The idea of using SFCs for parallel partitioning of Quad-/Octrees is not new in itself and has been tried by many researchers.
+The idea of using SFCs for parallel partitioning of Quad-/Octrees is not new in itself and has been used by many researchers.
 For instance, \texttt{Octor} \cite{Tu;OHallaron;Ghattas:05:Scalable-parallel-oc} uses a Morton curve (also known as Z-curve) for traversing the leaves of an Octree for indexing and load balancing and has been scaled up to 62,000 cores \cite{Burstedde;Ghattas;Gurnis;etal:08:Scalable-adaptive-ma}.
@@
 The implementation of these algorithms, which were shown to scale to more than 200,000 cores, is publicly available through a simple API provided by the \texttt{p4est} library \cite{p4est-github}.
-In fact the algorithms presented in this paper are directly implemented on top of the \texttt{p4est} library and we do not discuss any algorithm that is already covered in \cite{Burstedde;Wilcox;Ghattas:11:p4est:-Scalable-Algo}.
+In fact the algorithms presented in this paper are implemented on top of the \texttt{p4est} library and we do not discuss any algorithm that is already covered in \cite{Burstedde;Wilcox;Ghattas:11:p4est:-Scalable-Algo}.
 We also make use of the popular \texttt{PETSc} \cite{Balay;Abhyankar;Adams;etal:14:PETSc-Web-page} library for linear algebra and its parallel primitives, such as parallel ghosted vector and scatter/gather operations, which simplifies the implementation.
 
-Parallel level-set algorithms can be categorized in two groups: 1) Parallel advection algorithms and 2) Parallel reinitialization algorithms.
-Eulerian advection schemes can easily be parallelized but unfortunately are limited by the CFL condition which could be very restrictive for adaptive grids.
-semi-Lagrangian methods combine the unconditional stability of Lagrangian methods and the ease of use of Eulerian grids and have been successfully used for advecting the level-set function on tree-based grids in the past \cite{Min;Gibou:07:A-second-order-accur}.
+Parallel level-set algorithms can be categorized into two groups: parallel advection algorithms and parallel reinitialization algorithms.
+Eulerian advection schemes can easily be parallelized but unfortunately are limited by the CFL condition, which could be very restrictive for adaptive grids.
+Semi-Lagrangian methods combine the unconditional stability of Lagrangian methods and the ease of use of Eulerian grids and have been successfully used for advecting the level-set function on tree-based grids \cite{Losasso;Gibou;Fedkiw:04:Simulating-Water-and, Losasso;Fedkiw;Osher:06:Spatially-Adaptive-T, Min;Gibou:07:A-second-order-accur}.
 However, parallelizing the semi-Lagrangian algorithm in a domain decomposition context is not an easy task.
@@
 section \ref{sec:parallel algorithms}).
-Second, load balancing could be an issue for large CFL numbers and nonuniform velocity fields due to clustering of departure points and can considerably restrict the scalability of the algorithm.
+Second, load balancing could be an issue for large CFL numbers and nonuniform velocity fields, due to clustering of departure points, which can thus considerably restrict the scalability of the algorithm.
 Both of these problems, of course, could be avoided by choosing $\text{CFL} \le 1$ but that would defeat the purpose of using the semi-Lagrangian algorithm in the first place.
 
-Nonetheless, several parallel semi-Lagrangian algorithms have been proposed in the past.
+Nonetheless, several parallel semi-Lagrangian algorithms have been proposed.
 A simple domain decomposition technique was used in \cite{Thomas;Cote:95:Massively-parallel-s} where the width of the ghost layer is fixed based on the maximum CFL number to ensure that all the departure points are covered by the ghost layer.
-At large CFL numbers, this leads to a large volume of communication that can limit the scalability.
-Nonetheless good scaling was reported for small CFL numbers ($\text{CFL} \le 2$).
+Good scalings were reported for small CFL numbers ($\text{CFL} \le 2$).
+However, for large CFL numbers, this leads to a large volume of communication that can limit the scalability.
 In \cite{Drake;Foster;Michalakes;etal:95:Design-and-performan} the authors propose a more sophisticated domain decomposition approach which uses a ``dynamic ghost layer''.
@@
 This approach seems to provide good scalability for transporting a single tracer up to about 1000 cores for $\text{CFL} \sim 10$.
-At higher CFL numbers the method begins to loose scalability due to an increase in communication volume.
+At higher CFL numbers, the method begins to loose scalability due to an increase in communication volume.
 Finally, note that although we are mainly interested in parallel semi-Lagrangian methods, one could resort to finite difference or finite element discretization methods if small CFL numbers are acceptable.
-Indeed several algorithms of this type have been proposed with applications in modeling dendritic crystal growth \cite{Wang;Chang;Kale;etal:06:Parallelization-of-a}, multiphase flows \cite{Sussman:05:A-parallelized-adapt, Fortmeier;Bucker:11:A-parallel-strategy-, Rodriguez;Sahni;Lahey-Jr;etal:13:A-parallel-adaptive-}, and atomization process \cite{Herrmann:10:A-parallel-Eulerian-}.
+Indeed several algorithms of this type have been proposed with applications to modeling dendritic crystal growth \cite{Wang;Chang;Kale;etal:06:Parallelization-of-a}, multiphase flows \cite{Sussman:05:A-parallelized-adapt, Fortmeier;Bucker:11:A-parallel-strategy-, Rodriguez;Sahni;Lahey-Jr;etal:13:A-parallel-adaptive-}, and atomization process \cite{Herrmann:10:A-parallel-Eulerian-}.
 %, and image segmentation on GPUs \cite{Lefohn;Cates;Whitaker:03:Interactive-GPU-base,Cates;Lefohn;Whitaker:04:GIST:-an-interactive,Roberts;Packer;Sousa;etal:10:A-work-efficient-GPU}.
@@
 In many applications, it is desirable that the level-set function has signed-distance property, i.e $|\nabla \phi| = 1$.
-Generally, there are two approaches to enforce this property: 1) by solving the pseudo-time transient reinitialization equation \cite{Sussman;Smereka;Osher:94:A-Level-Set-Approach, Osher;Fedkiw:01:Level-Set-Methods:-A}:
+Generally, there are two approaches to enforce this property, either by solving the pseudo-time transient reinitialization equation \cite{Sussman;Smereka;Osher:94:A-Level-Set-Approach, Osher;Fedkiw:01:Level-Set-Methods:-A}
 \ben
@@
 \een
-or 2) by solving the Eikonal equation:
+or by solving the Eikonal equation
 \ben
-F(x)|\nabla\phi| = 1,
+F(x)|\nabla\phi| = 1
 \een 
@@
 Unfortunately, this approach does not seem to be more effective than a simple static decomposition.
-In \cite{Zhao:07:Parallel-implementat} a parallel FSM method was presented for the fist time.
-However, a more scalable FSM was more recently proposed in \cite{Detrixhe;Gibou;Min:13:A-parallel-fast-swee} where the Cuthill-McKee numbering was utilized to expose more parallelism.
+In \cite{Zhao:07:Parallel-implementat} a parallel FSM method was presented for the first time, which suffered from a plateau in the speed up.
+A scalable FSM was more recently proposed in \cite{Detrixhe;Gibou;Min:13:A-parallel-fast-swee}, where the Cuthill-McKee numbering was utilized to exploit more parallelism.
 More recently a two-scale, hybrid FMM-FSM was presented in \cite{Chacon;Vladimirsky:13:A-parallel-Heap-Cell} which, albeit being more complicated to implement, promises even better scalability.
@@
 However, unlike FMM, FIM avoids sorting the list and allows for concurrent updating of all nodes in an iterative fashion.
-In this article we simply choose the pseudo-time transient formulation for two reasons: 1) it is considerably easier to parallelize on Quadtrees and Octrees and 2) we are merely interested in the sign-distance property close to the interface which only requires a few iterations.
+In this article we simply choose the pseudo-time transient formulation for two reasons: 1) it is considerably easier to parallelize on Quadtrees and Octrees and 2) we are merely interested in the sign-distance property close to the interface, which only requires a few iterations.
 
-The rest of this article is organized as follows: In section \ref{sec:levelset method} we briefly review the sequential algorithms and discretization methods for the level-set equation on adaptive tree-based grids.
-These ideas are then extended in section \ref{sec:parallel algorithms} to parallel using a domain decomposition method.
-In section \ref{sec:scaling} we provide several examples that illustrate the scalability of our algorithms.
-Finally, we close by providing an important application of our algorithms in modeling the solidification process by solving a Stefan problem in \ref{sec:application}.
+The rest of this article is organized as follows.
+In section \ref{sec:levelset method}, we briefly review the sequential algorithms and discretization methods for the level-set equation on adaptive tree-based grids.
+These ideas are then extended in section \ref{sec:parallel algorithms} to parallel environments using a domain decomposition method.
+In section \ref{sec:scaling}, we provide several examples that illustrate the scalability of our algorithms.
+Finally, we close by providing an important application of our strategy by considering the simulation of the solidification process by solving a Stefan problem in section \ref{sec:application}.

### 2015-01-28 paper/tex/levelset_method.tex
added frederic's corrections. Some points still need to be discussed

@@
 The level-set method, introduced in \cite{Osher;Sethian:88:Fronts-Propagating-w}, is an implicit framework for tracking interfaces that undergo complicated topological changes.
-In this framework an interface is represented by the zero contour of a higher dimensional function, e.g.\ a curve in two spatial dimension can be described as $\Gamma = \{(x,y) | \phi(x,y) = 0\}$ where $\phi(x,y)$ is the level-set function.
-The evolution of the curve under a velocity field $\mathbf{V}$ is then obtained by solving the level-set equation:
+In this framework, an interface is represented by the zero contour of a higher dimensional function, e.g.\ a curve in two spatial dimensions can be described as $\Gamma = \{(x,y) | \phi(x,y) = 0\}$, where $\phi(x,y)$ is the level-set function.
+The evolution of the curve under a velocity field $\underline{\mbf{u}}$ is then obtained by solving the level-set equation:
 \be
-\phi_t + \mathbf{V} \cdot \nabla \phi = 0.
+\phi_t + \underline{\mbf{u}} \cdot \nabla \phi = 0.
 \label{eq:ls}
@@
 When the velocity field does not depend on the level-set function itself, equation \eqref{eq:ls} can be solved using the semi-Lagrangian method.
-An important advantage of the semi-Lagrangian method over the regular finite difference method is its unconditional stability which allows for arbitrarily large time steps.
-This is particularly important when using adaptive grids that allow for higher grid resolutions.
+An important advantage of the semi-Lagrangian method over the regular finite difference method is its unconditional stability, which allows for arbitrarily large time steps.
+This is particularly important when using adaptive grids, as higher grid resolutions translate into unpractical smaller time steps.
 
-In general, an infinite number of level-set functions can describe the same zero contour and thus interface.
+In general, an infinite number of level-set functions can describe the same zero contour and thus an interface.
 However, it is desirable to chose a function with the signed distance property, i.e.\ $|\nabla \phi| = 1$.
@@
 \ee
-where $\phi_0$ is any level-set function that correctly describes the interface location and $S(\phi_0)$ is an appropriate approximation of the sign function \cite{Osher;Fedkiw:02:Level-Set-Methods-an}.
+where $\tau$ is a pseudo time step, $\phi_0$ is any level-set function that correctly describes the interface location and $S(\phi_0)$ is an appropriate approximation of the sign function.
 Here, we do not go into the details of the sequential algorithms for solving equations \eqref{eq:ls} and \eqref{eq:reinitialization}.
-Instead, we note that the algorithms presented in section \ref{sec:parallel algorithms} are based on the sequential methods presented in \cite{Min;Gibou:07:A-second-order-accur} and refer the interested reader to the aforementioned article and references therein for more details.
+Instead, we note that the algorithms presented in section \ref{sec:parallel algorithms} are based on the sequential methods presented in \cite{Min;Gibou:07:A-second-order-accur} and refer the interested reader to the aforementioned articles and references therein for more details.
 

### 2015-01-28 paper/tex/parallel_algorithms.tex
added frederic's corrections. Some points still need to be discussed

--- 2015-01-09 paper/tex/parallel_algorithms.tex
@@
 Moreover, adaptive tree-based grids are easy to generate in the presence of a signed-distance level-set function \cite{Min;Gibou:07:A-second-order-accur} and can efficiently be encoded using a tree datastructure \cite{Samet:90:Applications-of-Spat}.
-In order to develop scalable parallel algorithms on these grids, it is necessary to parallelize the datastructure and grid manipulations methods such as refinement and coarsening of cells as well as to provide a fast method for grid partitioning and load balancing.
+In order to develop scalable parallel algorithms on these grids, it is necessary to parallelize the datastructure and grid manipulation methods such as refinement and coarsening of cells as well as to provide a fast method for grid partitioning and load balancing.
 The \texttt{p4est} library \cite{p4est-github} is a collection of such parallel algorithms that has recently emerged and shown to scale up to 200,000 cores \cite{Burstedde;Wilcox;Ghattas:11:p4est:-Scalable-Algo}.
@@
 Moreover it is implicitly assumed that the macromesh is small enough that it can be entirely replicated on all processes.
-For instance, in many of the applications that we are interested in in this paper the macromesh is simply a single cell.
+For instance, in many of the applications that we are considering in this paper the macromesh is simply a single cell.
 \texttt{p4est} allows for arbitrary refinement and coarsening criteria through defining callback functions.
@@
 \be
-\min_{\forall v \in V(C)} |\phi| \le \frac{L D}{2}.
+\min_{\forall v \in V(C)} |\phi| \le \frac{L D}{2},
 \label{eq:refine}
 \ee
+where $V(C)$ denotes the set of all vertices of cell $C$, $L$ denotes the Lipschitz constant of the level-set function, and $D$ denotes the diagonal size of cell $C$.
 Conversely an existing cell is marked for coarsening if:
@@
 \ee
-Here $V(C)$ denotes the set of all vertices of cell $C$, $L$ denotes the Lipschitz constant of the level-set function, and $D$ denotes the diagonal size of cell $C$.
 We refer to section 3.2 of \cite{Burstedde;Wilcox;Ghattas:11:p4est:-Scalable-Algo} for details on the parallel refinement and coarsening algorithms used in \texttt{p4est}.
@@
 A Z-curve is a Space Filling Curve (SFC) with the important property that cells with close Z-indices are also geometrically close in the physical domain.
-This is beneficial since it leads to both reduction in MPI communication and improvements of the cache performance of several algorithms such as interpolation and finite difference calculations.
+This is beneficial since it leads to both reduction in MPI communications and improvements of the cache performance of several algorithms such as interpolation and finite difference calculations.
 For more details on parallel partitioning in \texttt{p4est} one may refer to section 3.3 of \cite{Burstedde;Wilcox;Ghattas:11:p4est:-Scalable-Algo}.
@@
 \includegraphics[width = \columnwidth]{figures/p4est_zcurve.pdf}
-\caption{Left: A ``forest'' made up of two trees $k_0$ and $k_1$.
+\caption{Left: a ``forest'' made up of two trees $k_0$ and $k_1$.
 Parallel partitioning is achieved by first constructing a Z-curve starting at cell ``S'' and ending at cell ``E''.
-Next the curve is split up among processes either uniformly or by assigning different weights to cells.
+Next, the one dimensional curve is split up among processes either uniformly or by assigning different weights to cells.
 Here processes are represented via different colors.
 Note how using the Z-curve naturally leads to clustering of most cells in each domain.
-Right: Schematic of a tree data structure representing the forest and its partitioning.}
+Right: schematic of a tree data structure representing the forest and its partitioning.}
 \label{fig:p4est_zcurve}
@@
 An important feature of our algorithms is that they are designed for non-graded trees.
-This is important because we can entirely skip tree balancing which was shown to be one of the most time consuming parts of grid adaptation in \texttt{p4est} \cite{Burstedde;Wilcox;Ghattas:11:p4est:-Scalable-Algo}.
+This is important because we can entirely skip tree balancing, which was shown to be one of the most time consuming parts of grid adaptation in \texttt{p4est} \cite{Burstedde;Wilcox;Ghattas:11:p4est:-Scalable-Algo}.
 
@@
 only the leaves are explicitly stored.
-However, \textcolor{blue}{explicit knowledge of the hierarchal structure of tree is greatly beneficial in several algorithms, e.g.\ in search operations needed for the interpolation algorithm.} Thus, we introduce a simple reconstruction algorithm which recreates a local representation of the entire ``forest'' that is only adapted to local cells and, potentially, the ghost layer.
-This approach is similar to the ideas introduced in \cite{Bangerth;Burstedde;Heister;etal:11:Algorithms-and-data-} and our tests show that in a typical application they usually amount to less that $1\%$ of the entire runtime.
+However, explicit knowledge of the hierarchal structure of the tree is greatly beneficial in several algorithms, e.g.
+in search operations needed for the interpolation algorithm.
+Thus, we introduce a simple reconstruction algorithm that recreates a local representation of the entire ``forest'' that is only adapted to local cells and, potentially, the ghost layer.
+This approach is similar to the ideas introduced in \cite{Bangerth;Burstedde;Heister;etal:11:Algorithms-and-data-} and our tests show that in a typical application they amount to less that $1\%$ of the entire runtime.
 Algorithm \ref{alg:reconstruction} illustrates how this reconstruction is performed.
-Given a forest and a layer of ghost cells from \texttt{p4est}, the algorithm generates a local representation of the forest by recursively refining from the root until reaching the same level and location of all leaves in \texttt{p4est} and \texttt{ghost}.
+Given a forest and a layer of ghost cells from \texttt{p4est}, the algorithm generates a local representation of the forest by recursively refining from the root until reaching the same level and location of all leaves in the local forest and the ghost layer.
 Note that algorithm \ref{alg:reconstruction} does not involve any communication and is load balanced provided that the initial forest is balanced.
-Figure \ref{fig:reconstruction} illustrates an application of algorithm \ref{alg:reconstruction}.
+Figure \ref{fig:reconstruction} illustrates an example where Algorithm \ref{alg:reconstruction} is applied.
 Note how each process has independently generated a local representation of the forest that is refined to match the same leaves as in the global forest and ghost layer.
@@
 \end{center}
-\caption{Left: A forest refined close to an interface and partitioned among four processes as indicated by colors.
-Right: Each process independently recreates a local forest which is refined to match the global grid and is as coarse as possible elsewhere.
+\caption{Left: a forest refined close to an interface and partitioned among four processes, as indicated by colors.
+Right: each process independently recreates a local forest that is refined to match the local grid and is as coarse as possible elsewhere.
 Note that empty cells are fictitious, i.e.
@@
 As indicated earlier, we use the semi-Lagrangian method to solve equation \eqref{eq:ls} when the velocity field is externally generated, i.e.
-when it does not depend on the level-set function itself.
+when it does not depend explicitly on the level-set function itself.
 Let us rewrite equation \eqref{eq:ls} along the characteristic curve $\mathbf{X}(t)$ as:
@@
 \begin{array}{rcl}
-\dfrac{\text{d} \mathbf{X}}{\text{d} t} &=& \mathbf{V}, \\ [3ex]
+\dfrac{\text{d} \mathbf{X}}{\text{d} t} &=& \underline{\mbf{u}}, \\ [3ex]
 \dfrac{\text{d} \phi(\mathbf{X}(t), t)}{\text{d} t} &=& 0.
@@
 starting from the grid $G^{n+1}$ \footnote{$G^{n+1}$ itself is computed in an iterative fashion as explained later on.}, we simply write $\phi^{n+1}(\mathbf{X}^{n+1}) = \phi(\mathbf{X}(t^{n+1}), t^{n+1}) = \phi(\mathbf{X}(t^n), t^n) = \phi^n(\mathbf{X}_d)$.
-Here the characteristic curves are chosen such that $\mathbf{X}(t^{n+1})$ are simply the coordinates of grids points of $G^{n+1}$, and $\mathbf{X}_d$ are the departure points which are computed using the second-order midpoint method \cite{Min;Gibou:07:A-second-order-accur}:
+Here, the characteristic curves are chosen such that $\mathbf{X}(t^{n+1})$ are simply the coordinates of grids points of $G^{n+1}$, and $\mathbf{X}_d$ are the departure points, which are computed using the second-order midpoint method \cite{Min;Gibou:07:A-second-order-accur}:
 \bea
-\mathbf{X}^\star &=& \mathbf{X}^{n+1} - \frac{\Delta t}{2} \mathbf{V}^{n}(\mathbf{X}^n),	   \label{eq:xstar}      \\
-\mathbf{X}_d     &=& \mathbf{X}^{n+1} - \Delta t \mathbf{V}^{n+\frac{1}{2}}(\mathbf{X}^\star), \label{eq:xdeparture}
+\mathbf{X}^\star &=& \mathbf{X}^{n+1} - \frac{\Delta t}{2} \underline{\mbf{u}}^{n}(\mathbf{X}^n),	   \label{eq:xstar}      \\
+\mathbf{X}_d     &=& \mathbf{X}^{n+1} - \Delta t \underline{\mbf{u}}^{n+\frac{1}{2}}(\mathbf{X}^\star), \label{eq:xdeparture}
 \eea
-where $\mathbf{V}^{n+\frac{1}{2}}$ is obtained via extrapolation from previous times, i.e.:
+where $\underline{\mbf{u}}^{n+\frac{1}{2}}$ is obtained via extrapolation from previous times, i.e.:
 \be
-\mathbf{V}^{n+\frac{1}{2}} = \frac{3}{2} \mathbf{V}^n - \frac{1}{2}\mathbf{V}^{n-1}. \label{eq:vn_p_half}
+\underline{\mbf{u}}^{n+\frac{1}{2}} = \frac{3}{2} \underline{\mbf{u}}^n - \frac{1}{2}\underline{\mbf{u}}^{n-1}. \label{eq:vn_p_half}
 \ee
-
-Note that all values at the intermediate point, $\mathbf{X}^\star$, and departure point, $\mathbf{X}_d$, must be calculated via interpolation from previous the grids $G^{n}$ and $G^{n-1}$.
-Here we use the stabilized second-order interpolation for $\phi(\mathbf{X}_d)$ and the multi-linear interpolation for $\mathbf{V}^{n+\frac{1}{2}}(\mathbf{X}^\star)$ \cite{Min;Gibou:07:A-second-order-accur}.
+Note that all values at the intermediate point, $\mathbf{X}^\star$, and departure point, $\mathbf{X}_d$, must be calculated via interpolation from the previous grids $G^{n}$ and $G^{n-1}$.
+Here, we use the stabilized second-order interpolation for $\phi(\mathbf{X}_d)$ and the multi-linear interpolation for $\underline{\mbf{u}}^{n+\frac{1}{2}}(\mathbf{X}^\star)$ \cite{Min;Gibou:07:A-second-order-accur}.
 Although parallelization of the interpolation process on a shared-memory machine is trivial, the same cannot be said for distributed-memory machines.
-In fact, the parallel interpolation algorithm \ref{alg:interpolation} is probably the most important contribution of this article \comment{but hasn't it been done before ?} \textcolor{blue}{Mohammad: not on adaptive since that's what makes it complicated}.
-Complication arises because not all departure points will reside in the domain of current process.
-Moreover, due to the irregular shape of partitions, one cannot even ensure they are entirely owned by neighboring processes.
-At best we can only expect that their locations are bounded by a halo of width $w \le \text{CFL} \: \Delta x_{\min}$ around the local partition.
-Of course if one enforces $\text{CFL} \le 1$, one can ensure that the halo is bounded by the ghost layer which significantly simplifies the communication problem.
-This assumption, however, defeats the purpose of using a semi-Lagrangian approach in the first place since we are interested in large values of $\text{CFL}$ number.
+In fact, the parallel interpolation procedure given in Algorithm \ref{alg:interpolation} is probably the most important contribution of this article since the procedure, which is trivial on uniform grids, is challenging in the case of trees because it is not straightforward to identify which processes owns the departure points.
+Indeed, complications arise because not all departure points will reside in the domain owned by the current process.
+Moreover, due to the irregular shapes of the partitions, one cannot even ensure they are entirely owned by neighboring processes.
+At best we can only expect that their locations are bounded by a halo of width $w \le \text{CFL} \: \Delta x_{\min}$ around the local partition, where $x_{\min}$ is the size of the smallest cell in the forest.
+Naturally, if one enforces $\text{CFL} \le 1$, one can ensure that the halo is bounded by the ghost layer, which significantly simplifies the communication problem.
+This assumption, however, defeats the purpose of using a semi-Lagrangian approach, whose purpose is to enable large CFL values.
 
@@
 Note that searching the point in the local tree is performed recursively using the hierarchal reconstruction (c.f.\ algorithm \ref{alg:reconstruction}).
-Moreover, the owner's rank is found by computing the Z-index of the point and then using binary search on the Z-curve.
+Moreover, the owner's rank is found by computing the Z-index of the point and then using a binary search on the Z-curve.
 This is already implemented in \texttt{p4est} and explained in details in section 2.5 of \cite{Burstedde;Wilcox;Ghattas:11:p4est:-Scalable-Algo}.
@@
 \end{center}
-\caption{Left: Location of back-traced points depends on the magnitude of the local velocity and on the time-step.
-Although the distance to the departure point is bounded by $\text{CFL} \: \Delta x_{\min}$, one cannot make any assumption about receiving rank without explicitly searching the entire Z-curve.
+\caption{Left: the location of back-traced points depends on the magnitude of the local velocity and on the time-step.
+Although the distance to the departure point is bounded by $\text{CFL} \: \Delta x_{\min}$, one cannot predict the receiving rank without explicitly searching the entire Z-curve.
 Moreover, the receiving process has no prior knowledge about which processes to check for incoming messages nor does it know anything about the possible message length (i.e.
 number of points).
-Middle: A directed graph illustrating the communication pattern among processes with arrows representing the direction in which messages are sent.
-Right: The adjacency matrix of the communication graph.
+Middle: a directed graph illustrating the communication pattern among processes with arrows representing the direction in which messages are sent.
+Right: the adjacency matrix of the communication graph.
 For each row, the sum of all columns represents the number of messages that need to be sent.
 Conversely, for each column, the sum of all rows represents the number of messages that need to be received.
-As detailed in algorithm \ref{alg:interpolation} this information is enough to build a parallel interpolation scheme.}
+As detailed in Algorithm \ref{alg:interpolation}, this information is enough to build a parallel interpolation scheme.}
 \label{fig:communication}
@@
 This is done by alternating between local calculations and probing for incoming messages from other processes.
-The interpolation is finished once the values for all local the points have been calculated and all the remote requests have been processed (see algorithm \ref{alg:interpolation}).
+The interpolation is finished once the values for all local the points have been calculated and all the remote requests have been processed (see Algorithm \ref{alg:interpolation}).
 \begin{algorithm}[htbp]
@@
 
-Using the interpolation algorithm \ref{alg:interpolation}, we close this section by presenting the final semi-Lagrangian algorithm \ref{alg:semi-lagrangian}.
-The basic idea here is to start from an initial guess $G^{n+1}_0$ and modify the grid using the refinement \eqref{eq:refine} and coarsening \eqref{eq:coarsen} criteria until convergence is obtained.
+Using the interpolation Algorithm \ref{alg:interpolation}, we close this section by presenting the final semi-Lagrangian Algorithm \ref{alg:semi-lagrangian}.
+The basic idea is to start from an initial guess $G^{n+1}_0$ for the grid and modify it using the refinement \eqref{eq:refine} and coarsening \eqref{eq:coarsen} criteria until convergence is obtained.
 Various options are available for $G^{n+1}_0$.
@@
 \subsection{Reinitialization}
-Successive application of algorithm \ref{alg:semi-lagrangian}, especially for large values of the CFL number, eventually degrades the signed distance property of the level-set function.
+Successive application of Algorithm \ref{alg:semi-lagrangian}, especially for large values of the CFL number, eventually degrades the signed distance property of the level-set function.
 Thus, it is important to reinitialize the level-set function every few iterations, especially because the quality of generated grid heavily depends on the signed distance property.
@@
 \be
-\frac{\text{d} \phi}{\text{d} \tau} + S(\phi_0) \left(\mathcal{H}_G(D^+_i \phi, D^-_i \phi) - 1 \right) = 0,
+\frac{\text{d} \phi}{\text{d} \tau} + S(\phi_0) \left( \mathcal{H}_G(D^+_i \phi, D^-_i \phi) - 1 \right) = 0,
 \label{eq:semidiscrete_reinit}
@@
 \begin{array}{lcr}
-	\sqrt{\sum_i \max\left(|a^+_i|^2, |b^-_i|^2\right)} & \hspace {5 mm} \text{if} & S(\phi_0) \le 0 \\
+	\sqrt{\sum_i \max \left(|a^+_i|^2, |b^-_i|^2\right)} & \hspace {5 mm} \text{if} & S(\phi_0) \le 0 \\
 	\\
@@
 where $a^+ = \max(a, 0)$ and $a^- = \min(a, 0)$.
-Similar to \cite{Min;Gibou:07:A-second-order-accur}, equation \eqref{eq:semidiscrete_reinit} is integrated in time using the TVD-RK2 scheme and with adaptive time-stepping in order to accelerate the convergence to the steady state.
+Similar to \cite{Min;Gibou:07:A-second-order-accur}, equation \eqref{eq:semidiscrete_reinit} is integrated in time using the TVD-RK2 scheme with adaptive time-stepping in order to accelerate the convergence to the steady state.
 Since the computation is based on a local stencil, the parallel implementation of this scheme is mostly trivial.
 However, one minor point requires further explanation.
-As suggested in \cite{Min;Gibou:07:A-second-order-accur}, one-sided derivatives $D^+_i \phi$ and $D^-_i \phi$ are computed using second order discretization which require to compute the second derivatives.
-To enable overlap between computation and communications when computing second order derivatives and also integrating \eqref{eq:semidiscrete_reinit}, we use the following common technique.
+As suggested in \cite{Min;Gibou:07:A-second-order-accur}, one-sided derivatives $D^+_i \phi$ and $D^-_i \phi$ are computed using second order discretization which require to compute the second-order derivatives.
+To enable overlap between computation and communications when computing second-order derivatives and also integrating equation \eqref{eq:semidiscrete_reinit}, we use the following common technique.
 First, we label all local points, $L_p$, as either private, $P_p$, or boundary, $B_p$.
@@
 Algorithm \ref{alg:overlap} illustrates how this labeling can help with overlapping the computation and the communication associated to an arbitrary local operation $\phi^{n+1} \gets \mathcal{F}(\phi^n)$.
-Note that \texttt{p4est} library already includes all the primitives required for labeling local points without any further communication.
+Note that the \texttt{p4est} library already includes all the primitives required for labeling local points without any further communication.
 

### 2015-01-28 paper/tex/scaling_results.tex
added frederic's corrections. Some points still need to be discussed

@@
 In this section we present some results that demonstrate the scalability of our algorithms.
-All of our tests were ran on the Stampede cluster at Texas Advanced Computing Center (TACC) where we are limited to $4096$ cores at most.
+All of our tests were ran on the Stampede cluster at the Texas Advanced Computing Center (TACC) where we are limited to $4096$ cores at most.
 Each node of Stampede has 2 eight-core Xenon E5-2680 processes clocked at 2.7 GHz with 32 GB of DDR3-1600 MHz memory and interconnected using an InfiniBand network card.
@@
 \subsection{Interpolation}
-In this section we show the results for a simple test to measure the scalability of the interpolation algorithm \ref{alg:interpolation}.
+In this section we show the results for a simple test to measure the scalability of the interpolation Algorithm \ref{alg:interpolation}.
 The test consists of interpolating a function at a number of random points on a randomly refined Octree in three spatial dimensions.
-We consider two cases, a small test on a level 9 tree with roughly 33M nodes and larger test on a level 13 tree with roughly 280M nodes.
+We consider two cases, a small test on a level 9 tree with roughly 33M nodes and a larger test on a level 13 tree with roughly 280M nodes.
 In both cases the number of randomly generated points is chosen to be equal to the number of nodes and the stabilized second-order interpolation of \cite{Min;Gibou:07:A-second-order-accur} is performed 10 times to smooth out possible timing fluctuations.
@@
 	\end{center}
-	\caption{Strong scaling of algorithm \ref{alg:interpolation} for several tests where $N_G$ denotes the number of random interpolation points (which is the same as the number of nodes in the Octree) and $\alpha$ denotes the percentage of these points that are remote for each process.
+	\caption{Strong scaling of Algorithm \ref{alg:interpolation} for several tests where $N_G$ denotes the number of random interpolation points (which is the same as the number of nodes in the Octree) and $\alpha$ denotes the percentage of these points that are remote for each process.
 Here ``Total'' represents the total time spent in the interpolation while ``buffer'', ``local'', ``queries'', and ``replies'' represent the the timing for different sections (c.f.
-algorithm \ref{alg:interpolation}).
+Algorithm \ref{alg:interpolation}).
 The black dashed line represents the ideal scaling.
@@
 \subsection{Semi-Lagrangian}
-To test the scalability of the semi-Lagrangian algorithm \ref{alg:semi-lagrangian}, we consider a slightly modified version of the Enright's rotation test \cite{Enright;Fedkiw;Ferziger;etal:02:A-Hybrid-Particle-Le}, i.e.
+To test the scalability of the semi-Lagrangian scheme of Algorithm \ref{alg:semi-lagrangian}, we consider a slightly modified version of the Enright's rotation test \cite{Enright;Fedkiw;Ferziger;etal:02:A-Hybrid-Particle-Le}, i.e.
 we advect a sphere of radius 0.35 located at $(0.4, 0.4, 0.4)$ with a divergence free velocity field given by:
@@
 We also perform the test for two different initial girds, a small grid with maximum level $l_\text{max} = 10$ and a large grid with maximum level $l_\text{max} = 12$.
+In both cases, the minimum level is $l_\text{min}=0$.
 After one advection step, these grids have approximately 15M and 255M nodes, respectively.
 \begin{table}[htbp]
-	\begin{minipage}{.5\linewidth}
+	\begin{minipage}{.48\linewidth}
 		\begin{center}
@@
 			\hline
-				CFL & 16 & 32 & 64 & 128 & 256 & 512 \\
+				\diagbox{CFL}{\#p} & 16 & 32 & 64 & 128 & 256 & 512 \\
 				\hline
@@
 	\end{minipage}%
-	\begin{minipage}{.5\linewidth}
+	\begin{minipage}{.48\linewidth}
 		\begin{center}
@@
 			\hline
-				CFL & 128 & 256 & 512 & 1024 & 2048 & 4096 \\
+				\diagbox{CFL}{\#p} & 128 & 256 & 512 & 1024 & 2048 & 4096 \\
 			\hline
@@
 	\end{minipage}	
-	\caption{Number of iterations required for the grid construction in algorithm \ref{alg:semi-lagrangian} for the rotation test on a (a) level-10 and (b) level-12 Octree with approximately 15M and 255M nodes, respectively.
+	\caption{Number of iterations required for the grid construction in Algorithm \ref{alg:semi-lagrangian} for the rotation test on a (a) level-10 and (b) level-12 Octree with approximately 15M and 255M nodes, respectively.
 Note how the iteration count increases with the CFL number but is almost independent of the number of processes.
@@
 
-Table \ref{tab:semilagrangian} illustrates the dependence of the number of iterations required to build the grid on the CFL number; as the CFL is increased, the interface travels a farther distance which necessitates more iterations to generate the grid.
+Table \ref{tab:semilagrangian} illustrates the dependence of the number of iterations required to build the grid on the CFL number; as the CFL is increased, the interface travels a farther distance, which necessitates more iterations to generate the grid.
 Note that the apparent dependence on the number of processes is most likely due to round-off errors during the interpolation and/or backtracking steps.
@@
 This is especially true for regions of space in which the streamlines cluster.
-Both factors can contribute to reduce the scalability of the algorithm at large CFL numbers.
+Both factors can contribute to reducing the scalability of the algorithm at large CFL numbers.
 \begin{figure}[htbp]
@@
 	\end{center}
-	\caption{Strong scaling of a single iteration of algorithm \ref{alg:semi-lagrangian} for the rotation test on a level-10 Octree with approximately 15M nodes.
+	\caption{Strong scaling of a single iteration of Algorithm \ref{alg:semi-lagrangian} for the rotation test on a level-10 Octree with approximately 15M nodes.
 Top row: scaling of the various components of the algorithm for (a) $\text{CFL} = 1$, (b) $\text{CFL} = 10$, and (c) $\text{CFL} = 100$.
@@
 	\end{center}
-	\caption{Strong scaling of a single iteration of algorithm \ref{alg:semi-lagrangian} for the rotation test on a level-12 Octree with approximately 255M nodes.
+	\caption{Strong scaling of a single iteration of Algorithm \ref{alg:semi-lagrangian} for the rotation test on a level-12 Octree with approximately 255M nodes.
 Top row: scaling of the various components of the algorithm for (a) $\text{CFL} = 1$, (b) $\text{CFL} = 10$, and (c) $\text{CFL} = 100$.
@@
 The bottom row of figure \ref{fig:communication_4096} exhibits a better balance both in the computation and communication volume in the last iteration of the semi-Lagrangian algorithm.
-This can be explained by noting that for the final iteration, partitioning of $G^{n+1}$ is a good approximation of partitioning of departure points on $G^n$.
+This can be explained by noting that for the final iteration, the partitioning of $G^{n+1}$ is a good approximation of the partitioning of the departure points on $G^n$.
 \comment{awkward ...
@@
 	\end{center}
-	\caption{Performance indicators of the first (top row) and last (bottom row) iterations of the semi-Lagrangian algorithm for the advection on 4096 processes with $\text{CFL} = 1$ (\drawdiamond{black} \comment{this does not work on my computer, I see a W} \mohammad{hum ...
+	\caption{Performance indicators of the first (top row) and last (bottom row) iterations of the semi-Lagrangian algorithm for the level-set advection on 4096 processes with $\text{CFL} = 1$ (\drawdiamond{black} \comment{this does not work on my computer, I see a W} \mohammad{hum ...
 not sure what to do for this one.
@@
 During the last semi-Lagrangian iteration, the initial grid $G_0$ (c.f.
-algorithm \ref{alg:semi-lagrangian}) is very close to the final grid.
+Algorithm \ref{alg:semi-lagrangian}) is very close to the final grid.
 As a result, the load imbalance is considerably improved (d).
-Curiously, however, the communication pattern does not seem to be change much between first and last iterations (e-f) \comment{e-f ?} \mohammad{e and f figures.
-Should we rephrase?}.}
+Curiously, however, the communication pattern does not seem to be change much between first and last iterations (e,f).}
 	\label{fig:communication_4096}
@@
 
-Finally we present the scaling results of our parallel reinitialization algorithm where we extensively make use of algorithm \ref{alg:overlap} for overlapping computation with communication when computing spatial derivatives.
-Our test consists in computing the signed distance to a collection of 100 spheres, whose radii and centers have been chosen randomly.
-The test is performed on a small, level-8 Octree with nearly 21M and a larger, level-10 Octree with nearly 337M grid points.
+Finally we present the scaling results of our parallel reinitialization algorithm where we extensively make use of Algorithm \ref{alg:overlap} for overlapping the computations with the communications when computing spatial derivatives.
+Our test consists in computing the signed distance function to a collection of 100 spheres, whose radii and centers are chosen randomly.
+The test is performed on a small, level-8 Octree with about 21M and a larger, level-10 Octree with about 337M grid points.
 In both cases the forest is built on a $3\times3\times3$ macro-mesh.
-Figure \ref{fig:reinit} illustrates that our reinitialization algorithm, and in particular the overlapping strategy presented in algorithm \ref{alg:overlap}, scales very well (c.f.\ table \ref{tab:scaling_reinit}).
-In general we expect similar scaling results for any local, finite-difference based calculations on Octrees that can efficiently utilize algorithm \ref{alg:overlap}.
+Figure \ref{fig:reinit} illustrates that our reinitialization algorithm, and in particular the overlapping strategy presented in Algorithm \ref{alg:overlap}, scales very well (c.f.\ table \ref{tab:scaling_reinit}).
+In general we expect similar scaling results for any local, finite-difference based calculations on Octrees that can efficiently utilize Algorithm \ref{alg:overlap}.
 \begin{figure}

### 2015-02-04 paper/tex/applications.tex
final pass (mohammad+frederic+arthur) done. One last thing remains, scaling with large number of processes and for a heavier load

@@
 %!TEX root = draft.tex
-\section{Application to the Stefan Problem} \label{sec:application}
+\section{Application to the Stefan problem} \label{sec:application}
 
@@
 In the case of a single component melt, and in the absence of convection, the process is dominated by diffusion and can be modeled as a Stefan problem.
-We decompose the computational domain $\Omega$ into \mohammad{How about: Consider a computational domain $\Omega$ decomposed into ...} \comment{this would be the exact phrasing of another article} two subdomains $\Omega_l$ and $\Omega_s$, separated by an interface $\Gamma$.
+We decompose the computational domain $\Omega$ into two subdomains $\Omega_l$ and $\Omega_s$, separated by an interface $\Gamma$.
 The Stefan problem describes the evolution of the temperature $T$, decomposed into $T_s$ in the solid phase $\Omega_s$ and $T_l$ in the liquid phase $\Omega_l$, as
@@
 \caption{$\texttt{General procedure for solving the Stefan problem}$}
-\mohammad{Is there a reason all of the algorithm is in type-writer font? Usually its reserved for code segments, etc.}
 \begin{algorithmic}[1]
-\State $\texttt{Initialize the forest and }\phi\texttt{ given the initial geometry}.$
-\State $\texttt{Initialize }T_s\texttt{ in }\Omega^+\texttt{ and }T_l\texttt{ in }\Omega^-.$
-\State $\texttt{Reinitialize }\phi\texttt{ and compute the local interface curvature }\kappa.$
-\State $\texttt{Compute }T^{n+1}_l\texttt{ and }T^{n+1}_s\texttt{ by solving the heat equations \eqref{eq:stefan_heat_equation_l} and \eqref{eq:stefan_heat_equation_s}}.$
-\State $\texttt{Extrapolate }T^{n+1}_s\texttt{ from }\Omega^+\texttt{ to }\Omega^-\texttt{ and }T^{n+1}_l\texttt{ from }\Omega^-\texttt{ to }\Omega^+.$
-\State $\texttt{Compute the velocity field }\mathbf{V}\texttt{ according to \eqref{eq:stefan_velocity}}.$
-\State $\texttt{Compute the time step dt following \eqref{eq:stefan_dt}.}$
-\State $\texttt{Evolve the interface and contruct the new forest using the Semi-Lagrangian procedure}.$
-\State $\texttt{Interpolate }T^{n+1}_s\texttt{ and }T^{n+1}_l\texttt{ from the old forest to the new forest}.$
-\State $\texttt{Go to 3 with n=n+1}.$
+\State Initialize the forest and $\phi$ given the initial geometry.
+\State Initialize $T_s$ in $\Omega^+$ and $T_l$ in $\Omega^-$.
+\State Reinitialize $\phi$ and compute the local interface curvature $\kappa$.
+\State Compute $T^{n+1}_l$ and $T^{n+1}_s$ by solving the heat equations \eqref{eq:stefan_heat_equation_l} and \eqref{eq:stefan_heat_equation_s}.
+\State Extrapolate $T^{n+1}_s$ from $\Omega^+$ to $\Omega^-$ and $T^{n+1}_l$ from $\Omega^-$ to $\Omega^+$.
+\State Compute the velocity field $\underline{\mathbf{V}}$ according to \eqref{eq:stefan_velocity}.
+\State Compute the time step dt following \eqref{eq:stefan_dt}.
+\State Evolve the interface and construct the new forest using the Semi-Lagrangian procedure.
+\State Interpolate $T^{n+1}_s$ and $T^{n+1}_l$ from the old forest to the new forest.
+\State Go to 3 with n=n+1.
 \end{algorithmic}
@@
 	\end{tabular}
-	\caption{Parallel efficiency of the total runtime for the Stefan test based on the lowest number of processes for each test. \comment{frederic, what do you think about this table ?}}
+	\caption{Parallel efficiency of the total runtime for the Stefan test based on the lowest number of processes for each test.}
 	\label{tab:scaling_stefan} 
@@
 The simulation was ran on $256$ MPI processes for $6$ hours and $30$ minutes, resulting in $396$ time iterations.
-Visualizations of the final iteration are presented in figures \ref{fig:stefan_grid}, \ref{fig:stefan_crystals} and \ref{fig:stefan_evolution}.
+Visualizations of the final iteration are presented in figures \ref{fig:stefan_grid} and \ref{fig:stefan_evolution}.
 The final iteration of the simulation consisted of $167$M grid points whereas a uniform grid with the equivalent finest resolution would lead to $8.59\cdot10^{12}$ grid points, i.e.
@@
 
-\begin{figure}[ht!]
-\begin{center}
-\subfigure[]{\includegraphics[width=0.45\textwidth]{figures/stefan_crystal_1.png}}
-\subfigure[]{\includegraphics[width=0.45\textwidth]{figures/stefan_crystal_2.png}}
-\subfigure[]{\includegraphics[width=0.45\textwidth]{figures/stefan_crystal_4.png}}
-\subfigure[]{\includegraphics[width=0.45\textwidth]{figures/stefan_crystal_3.png}}
-\caption{Representation of the computational domain (a) as well as various developped crystal structures obtained from solving the Stefan problem simulation (b,c,d).} \label{fig:stefan_crystals}
-\end{center}
-\end{figure}
+%\begin{figure}[ht!]
+%\begin{center}
+%\subfigure[]{\includegraphics[width=0.45\textwidth]{figures/stefan_crystal_1.png}}
+%\subfigure[]{\includegraphics[width=0.45\textwidth]{figures/stefan_crystal_2.png}}
+%\subfigure[]{\includegraphics[width=0.45\textwidth]{figures/stefan_crystal_4.png}}
+%\subfigure[]{\includegraphics[width=0.45\textwidth]{figures/stefan_crystal_3.png}}
+%\caption{Representation of the computational domain (a) as well as various developped crystal structures obtained from solving the Stefan problem simulation (b,c,d).} \label{fig:stefan_crystals}
+%\end{center}
+%\end{figure}
 
@@
 \begin{center}
+\begin{tabular}{c c c c c}
+(a) &
+\includegraphics[width=.2\textwidth]{stefan/nb2_iter19.png} &
+\includegraphics[width=.2\textwidth]{stefan/nb2_iter39.png} &
+\includegraphics[width=.2\textwidth]{stefan/nb2_iter59.png} &
+\includegraphics[width=.2\textwidth]{stefan/nb2_iter79.png}
+\end{tabular}
+
 \subfigure[]{

### 2015-02-04 paper/tex/parallel_algorithms.tex
final pass (mohammad+frederic+arthur) done. One last thing remains, scaling with large number of processes and for a heavier load

@@
 \includegraphics[width = \columnwidth]{figures/p4est_zcurve.pdf}
-\caption{Left: a ``forest'' made up of two trees $k_0$ and $k_1$.
+\caption{Left: a ``forest'' made up of two trees $T_0$ and $T_1$.
 Parallel partitioning is achieved by first constructing a Z-curve starting at cell ``S'' and ending at cell ``E''.
@@
 \begin{algorithm}[htbp]
-\caption{$H \gets \texttt{Reconstruct (}G\texttt{)}$} %\label{euclid}
+\caption{$H \gets \texttt{Reconstruct (}G\texttt{)}$: construction of the local tree hierarchy $H$ from the grid $G$}
 \begin{algorithmic}[1]
 \State $H \gets G.\texttt{macromesh()}$
-\For {$tr : G.\texttt{local\_trees()}$} \Comment{traverse local cells}
-	\For {$q : tr.\texttt{cells()}$}
-		\State $H.\texttt{update\_tree(} tr, q \texttt{)}$
+\For {$tr : G.\texttt{local\_trees()}$} \Comment{build hierarchy for local cells}
+	\For {$c : tr.\texttt{cells()}$}
+		\State $H.\texttt{update\_tree(} tr, c \texttt{)}$
 	\EndFor
@@
 
-\For {$q : G.\texttt{ghost\_cells()}$} \Comment{traverse ghost cells}
-	\State $H.\texttt{update\_tree(} q.\texttt{tree()}, q \texttt{)}$
+\For {$c : G.\texttt{ghost\_cells()}$} \Comment{build hierarchy for ghost cells}
+	\State $H.\texttt{update\_tree(} c.\texttt{tree()}, c \texttt{)}$
 \EndFor
@@
 \\
-\Function{$H.\texttt{update\_tree}$}{$tr, q$}   \Comment{recursive reconstruction}
-	\State $q_l \gets H.\texttt{root(}tr\texttt{)}$
-	\While{$q_l.\texttt{level()} \not= q.\texttt{level()} $}
-		\If {$q_l.$\texttt{is\_leaf()}} $q_l$\texttt{.split()}
+\Function{$H.\texttt{update\_tree}$}{$tr, c$}   \Comment{recursive tree reconstruction}
+	\State $c_l \gets H.\texttt{root(}tr\texttt{)}$
+	\While{$c_l.\texttt{level()} \not= c.\texttt{level()} $}
+		\If {$c_l.$\texttt{is\_leaf()}} $c_l$\texttt{.split()}
 		\EndIf
-		\State $h \gets q_l.\texttt{length()} / 2$ \Comment{select the next child based on direction}
-		\State $i \gets q.x \ge q_l.x + h$
-		\State $j \gets q.y \ge q_l.y + h$
-		\State $k \gets q.z \ge q_l.z + h$
+		\State $h \gets c_l.\texttt{length()} / 2$ \Comment{select the next child based on direction}
+		\State $i \gets c.x \ge c_l.x + h$
+		\State $j \gets c.y \ge c_l.y + h$
+		\State $k \gets c.z \ge c_l.z + h$
 
-		\State $q_l \gets q_l.\texttt{child(} i, j, k \texttt{)}$
+		\State $c_l \gets c_l.\texttt{child(} i, j, k \texttt{)}$
 	\EndWhile	
@@
 when it does not depend explicitly on the level-set function itself.
-Let us rewrite equation \eqref{eq:ls} along the characteristic curve $\mathbf{X}(t)$ as:
+Let us rewrite equation \eqref{eq:ls} along the characteristic curve $\underline{\mathbf{X}}(t)$ as:
 \be
@@
 \begin{array}{rcl}
-\dfrac{\text{d} \mathbf{X}}{\text{d} t} &=& \underline{\mbf{u}}, \\ [3ex]
-\dfrac{\text{d} \phi(\mathbf{X}(t), t)}{\text{d} t} &=& 0.
+\dfrac{\text{d} \underline{\mathbf{X}}}{\text{d} t} &=& \underline{\mbf{u}}, \\ [3ex]
+\dfrac{\text{d} \phi(\underline{\mathbf{X}}(t), t)}{\text{d} t} &=& 0.
 \end{array}
@@
 The semi-Lagrangian method integrates equations \eqref{eq:sl} backward in time, i.e.
-starting from the grid $G^{n+1}$ \footnote{$G^{n+1}$ itself is computed in an iterative fashion as explained later on.}, we simply write $\phi^{n+1}(\mathbf{X}^{n+1}) = \phi(\mathbf{X}(t^{n+1}), t^{n+1}) = \phi(\mathbf{X}(t^n), t^n) = \phi^n(\mathbf{X}_d)$.
-Here, the characteristic curves are chosen such that $\mathbf{X}(t^{n+1})$ are simply the coordinates of grids points of $G^{n+1}$, and $\mathbf{X}_d$ are the departure points, which are computed using the second-order midpoint method \cite{Min;Gibou:07:A-second-order-accur}:
+starting from the grid $G^{n+1}$ (computed in iteratively as explained later on), we simply write $\phi^{n+1}(\underline{\mathbf{X}}^{n+1}) = \phi(\underline{\mathbf{X}}(t^{n+1}), t^{n+1}) = \phi(\underline{\mathbf{X}}(t^n), t^n) = \phi^n(\underline{\mathbf{X}}_d)$.
+Here, the characteristic curves are chosen such that $\underline{\mathbf{X}}(t^{n+1})$ are simply the coordinates of grids points of $G^{n+1}$, and $\underline{\mathbf{X}}_d$ are the departure points, which are computed using the second-order midpoint method \cite{Min;Gibou:07:A-second-order-accur}:
 \bea
-\mathbf{X}^\star &=& \mathbf{X}^{n+1} - \frac{\Delta t}{2} \underline{\mbf{u}}^{n}(\mathbf{X}^n),	   \label{eq:xstar}      \\
-\mathbf{X}_d     &=& \mathbf{X}^{n+1} - \Delta t \underline{\mbf{u}}^{n+\frac{1}{2}}(\mathbf{X}^\star), \label{eq:xdeparture}
+\underline{\mathbf{X}}^\star &=& \underline{\mathbf{X}}^{n+1} - \frac{\Delta t}{2} \underline{\mbf{u}}^{n}(\underline{\mathbf{X}}^n),	   \label{eq:xstar}      \\
+\underline{\mathbf{X}}_d     &=& \underline{\mathbf{X}}^{n+1} - \Delta t \underline{\mbf{u}}^{n+\frac{1}{2}}(\underline{\mathbf{X}}^\star), \label{eq:xdeparture}
 \eea
@@
 \ee
-Note that all values at the intermediate point, $\mathbf{X}^\star$, and departure point, $\mathbf{X}_d$, must be calculated via interpolation from the previous grids $G^{n}$ and $G^{n-1}$.
-Here, we use the stabilized second-order interpolation for $\phi(\mathbf{X}_d)$ and the multi-linear interpolation for $\underline{\mbf{u}}^{n+\frac{1}{2}}(\mathbf{X}^\star)$ \cite{Min;Gibou:07:A-second-order-accur}.
+Note that all values at the intermediate point, $\underline{\mathbf{X}}^\star$, and departure point, $\underline{\mathbf{X}}_d$, must be calculated via interpolation from the previous grids $G^{n}$ and $G^{n-1}$.
+Here, we use the stabilized second-order interpolation for $\phi(\underline{\mathbf{X}}_d)$ and the multi-linear interpolation for $\underline{\mbf{u}}^{n+\frac{1}{2}}(\underline{\mathbf{X}}^\star)$ \cite{Min;Gibou:07:A-second-order-accur}.
 Although parallelization of the interpolation process on a shared-memory machine is trivial, the same cannot be said for distributed-memory machines.
@@
 Moreover, this simple approach does not work in the process of generating $G^{n+1}$ due to repartitioning.
+Indeed, $G^{n+1}$ is built iteratively and load balancing is enforced by repartitioning at each sub-iteration.
+Therefore, after one such sub-iteration, the backtracked points can end up outside of the initial ghost layer.
 An alternative approach would be to handle local and remote interpolations separately.
@@
 \begin{algorithm}[htbp]
-\caption{$values \gets \texttt{Interpolate (}H, \mathbf{X}\texttt{)}$} %\label{euclid}
+\caption{$values \gets \texttt{Interpolate (}H, F, \underline{\mathbf{X}}\texttt{)}$: interpolate the value of $F$, defined on the local tree hierarchy $H$, at coordinates $\underline{\mathbf{X}}$}
 \begin{algorithmic}[1]
 \State $col \gets 0, buff \gets null$ \Comment{Phase I -- buffering}
-\For {$\mathbf{p} : \mathbf{X}$} 
-	\State $[r, cell] \gets H.\texttt{search(}\mathbf{p}\texttt{)}$ \Comment{search for the owner's rank and cell}
+\For {$\underline{\mathbf{p}} : \underline{\mathbf{X}}$} 
+	\State $[r, cell] \gets H.\texttt{search(}\underline{\mathbf{p}}\texttt{)}$ \Comment{search for the owner's rank and cell}
 	\If {$r = mpirank$}
-		\State $buff[r].\texttt{push\_back(}\mathbf{p}, cell\texttt{)}$
+		\State $buff[r].\texttt{push\_back(}\underline{\mathbf{p}}, cell\texttt{)}$
 	\Else
-		\State $buff[r].\texttt{push\_back(}\mathbf{p}\texttt{)}$
+		\State $buff[r].\texttt{push\_back(}\underline{\mathbf{p}}\texttt{)}$
 		\State $col[r] \gets 1$
@@
 \begin{algorithm}[htbp]
-\caption{$[G^{n+1}, \phi^{n+1}] \gets \texttt{SemiLagrangian (}G^n, \phi^n, \mathbf{V}^n, \mathbf{V}^{n-1}, \text{CFL}\texttt{)}$} %\label{euclid}
+\caption{$[G^{n+1}, \phi^{n+1}] \gets \texttt{SemiLagrangian (}G^n, \phi^n, \underline{\mathbf{u}}^n, \underline{\mathbf{u}}^{n-1}, \text{CFL}\texttt{)}$: update $\phi^{n+1}$ from $\phi^n$ using a semi-Lagrangian scheme and construct the new forest $G^{n+1}$ that is consistent with the zero level-set of $\phi^{n+1}$}
 \begin{algorithmic}[1]
-\State $\Delta t_l \gets \text{CFL} \times G^n.\texttt{hmin()} / \max \{\mathbf{V}^n \} $
+\State $\Delta t_l \gets \text{CFL} \times G^n.\texttt{hmin()} / \max \{\underline{\mathbf{u}}^n \} $
 \State $\Delta t   \gets \texttt{MPI\_Allreduce(}\Delta t_l,\texttt{MPI\_MIN)}$
@@
 \While {$true$}
-	\State $\mathbf{X}_d  \gets \texttt{ComputeDeparturePoints(}G^{n+1}_0, \mathbf{V}^n, \mathbf{V}^{n-1}, \Delta t\texttt{)}$ \Comment{using equations \ref{eq:xstar} -- \ref{eq:vn_p_half}}
-	\State $\phi^{n+1} \gets \texttt{Interpolate(}H^n, \textbf{X}_d\texttt{)}$
+	\State $\underline{\mathbf{X}}_d  \gets \texttt{ComputeDeparturePoints(}G^{n+1}_0, \underline{\mathbf{u}}^n, \underline{\mathbf{u}}^{n-1}, \Delta t\texttt{)}$ \Comment{using equations \ref{eq:xstar} -- \ref{eq:vn_p_half}}
+	\State $\phi^{n+1} \gets \texttt{Interpolate(}H^n, \phi^n, \underline{\textbf{X}}_d\texttt{)}$
 	\State $G^{n+1} \gets G^{n+1}_0.\texttt{refine\_and\_coarsen(}\phi^{n+1}\texttt{)}$ \Comment{using equations \ref{eq:refine} and \ref{eq:coarsen} as criteria}
@@
 $P_p = L_p \setminus B_p$.
-Algorithm \ref{alg:overlap} illustrates how this labeling can help with overlapping the computation and the communication associated to an arbitrary local operation $\phi^{n+1} \gets \mathcal{F}(\phi^n)$.
+Algorithm \ref{alg:overlap} illustrates how this labeling can help with overlapping the computation and the communication associated to an arbitrary local operation $y \gets \mathcal{F}(x)$.
 Note that the \texttt{p4est} library already includes all the primitives required for labeling local points without any further communication.
@@
 \begin{algorithm}[htbp]
-\caption{$\phi^{n+1} \gets \texttt{Overlap (}\phi^n, \mathcal{F}\texttt{)}$} %\label{euclid}
+\caption{$y \gets \texttt{Overlap (}x, \mathcal{F}\texttt{)}$: compute $y_i=\mathcal{F}(x_i)$ for all nodes $i$, where $\mathcal{F}$ is a local operation, while hiding the communication to update the ghost layer}
 \begin{algorithmic}[1]
 \For {$i:B_p$} \Comment{I -- perform computation on boundary points}
-	\State $\phi^{n+1}_i \gets \mathcal{F}(\phi_i^n)$
+	\State $y_i \gets \mathcal{F}(x_i)$
 \EndFor
-\State $send\_req \gets \texttt{MPI\_Isend(}\phi^{n+1}_B\texttt{)}$ \Comment{II -- begin updating ghost values}
-\State $recv\_req \gets \texttt{MPI\_Irecv(}\phi^{n+1}_G\texttt{)}$
+\State $send\_req \gets \texttt{MPI\_Isend(}y_B\texttt{)}$ \Comment{II -- begin updating ghost values}
+\State $recv\_req \gets \texttt{MPI\_Irecv(}y_G\texttt{)}$
 \For {$i:P_p$} \Comment{III -- perform computation on private points}
-	\State $\phi^{n+1}_i \gets \mathcal{F}(\phi_i^n)$ 
+	\State $y_i \gets \mathcal{F}(x_i)$ 
 \EndFor
 \State $\texttt{MPI\_Waitall(}send\_req, recv\_req\texttt{)}$ \Comment{IV -- wait for ghost update to finish}
-\State \Return $\phi^{n+1}$
+\State \Return $y$
 \end{algorithmic}

### 2015-02-04 paper/tex/scaling_results.tex
final pass (mohammad+frederic+arthur) done. One last thing remains, scaling with large number of processes and for a heavier load

@@
 
+We define parallel efficiency as $e=s\cdot P_1 / P$ where $s=t_1/t_P$ is the speed-up, $P_1$ is the smallest number of processes for which the test was run, $t_1$ is the time to run the problem on $P_1$ processes, $P$ is the number of processes and $t_P$ is the time to run the problem on $P$ processes.
+We note that efficiencies larger than $100\%$ are reported for some cases.
+This is common and can be hardware related, for instance linked to the problem being locally smaller for larger number of processes and thus exploiting the cache better.
+
 \subsection{Interpolation}
@@
 The test consists of interpolating a function at a number of random points on a randomly refined Octree in three spatial dimensions.
-We consider two cases, a small test on a level 9 tree with roughly 33M nodes and a larger test on a level 13 tree with roughly 280M nodes.
+We consider two cases, a small test on a level\footnote{the level is the number of recursive splits allowed for each tree.} 9 tree with roughly 33M nodes and a larger test on a level 13 tree with roughly 280M nodes.
 In both cases the number of randomly generated points is chosen to be equal to the number of nodes and the stabilized second-order interpolation of \cite{Min;Gibou:07:A-second-order-accur} is performed 10 times to smooth out possible timing fluctuations.
@@
 
-\begin{table}
+\begin{table}[ht!]
 \centering
@@
 	\end{minipage}	
-	\caption{Number of iterations required for the grid construction in Algorithm \ref{alg:semi-lagrangian} for the rotation test on a (a) level-10 and (b) level-12 Octree with approximately 15M and 255M nodes, respectively.
-Note how the iteration count increases with the CFL number but is almost independent of the number of processes.
-The slight dependence between the number of iterations and the number of processes is most likely due to the dependence of round-off errors on the number of processes.
+	\caption{Number of sub-iterations required for the grid construction in Algorithm \ref{alg:semi-lagrangian} for the rotation test on a (a) level-10 and (b) level-12 Octree with approximately 15M and 255M nodes, respectively.
+Note how the sub-iteration count increases with the CFL number but is almost independent of the number of processes.
+The slight dependence between the number of sub-iterations and the number of processes is most likely due to the dependence of round-off errors on the number of processes.
 Nonetheless, close examination of the Octrees generated (data not shown) reveals that they are identical and independent of the number of processes used to perform the test.}
@@
 
-Table \ref{tab:semilagrangian} illustrates the dependence of the number of iterations required to build the grid on the CFL number; as the CFL is increased, the interface travels a farther distance, which necessitates more iterations to generate the grid.
+Table \ref{tab:semilagrangian} illustrates the dependence of the number of sub-iterations required to build the grid on the CFL number; as the CFL is increased, the interface travels a farther distance, which necessitates more sub-iterations to generate the grid.
 Note that the apparent dependence on the number of processes is most likely due to round-off errors during the interpolation and/or backtracking steps.
 Figures \ref{fig:semilagrangian_small} and \ref{fig:semilagrangian_large} illustrate the scalability of the algorithm for the small and large problems, respectively.
-To enable meaningful comparisons between different CFL numbers and number of processes, the maximum time has been scaled by the number of iterations required for the grid construction as reported in table \ref{tab:semilagrangian}.
+To enable meaningful comparisons between different CFL numbers and number of processes, the maximum time has been scaled by the number of sub-iterations required for the grid construction as reported in table \ref{tab:semilagrangian}.
 For both problems, excellent scalability is observed for $\text{CFL} = 1$ and $\text{CFL} = 10$.
@@
 	\end{center}
-	\caption{Strong scaling of a single iteration of Algorithm \ref{alg:semi-lagrangian} for the rotation test on a level-10 Octree with approximately 15M nodes.
+	\caption{Strong scaling of a single time step of Algorithm \ref{alg:semi-lagrangian} for the rotation test on a level-10 Octree with approximately 15M nodes.
 Top row: scaling of the various components of the algorithm for (a) $\text{CFL} = 1$, (b) $\text{CFL} = 10$, and (c) $\text{CFL} = 100$.
@@
 The solid dashed line represents the ideal scaling.
-Note that the maximum time has been scaled by the number of iterations required to build the tree (c.f.
-table \ref{tab:semilagrangian}).}
+Note that the maximum time has been scaled by the number of sub-iterations required to build the tree (c.f.
+table \ref{tab:semilagrangian}).
+Here \texttt{p4est\_nodes\_new}, \texttt{p4est\_partition} and \texttt{p4est\_refine\_coarsen} refer to constructing the global indexing for nodes, partitioning the forest, and the refining/coarsening operation, respectively \cite{Burstedde;Wilcox;Ghattas:11:p4est:-Scalable-Algo}.}
 	\label{fig:semilagrangian_small}
@@
 	\end{center}
-	\caption{Strong scaling of a single iteration of Algorithm \ref{alg:semi-lagrangian} for the rotation test on a level-12 Octree with approximately 255M nodes.
+	\caption{Strong scaling of a single time step of Algorithm \ref{alg:semi-lagrangian} for the rotation test on a level-12 Octree with approximately 255M nodes.
 Top row: scaling of the various components of the algorithm for (a) $\text{CFL} = 1$, (b) $\text{CFL} = 10$, and (c) $\text{CFL} = 100$.
@@
 The solid dashed line represents the ideal scaling.
-Note that the maximum time has been scaled by the number of iterations required to build the tree (c.f.
+Note that the maximum time has been scaled by the number of sub-iterations required to build the tree (c.f.
 table \ref{tab:semilagrangian}).}
@@
 Figure \ref{fig:communication_4096} illustrates the effects of the CFL number on different metrics, namely the number of interpolation points\footnote{Note that this includes both the local points and the points queried by other processes}, $N_p$, the number of sent and received messages, $N_m = S + R$, and the total communication volume, $V_m$ in megabytes (MB), for $p=4096$ processes.
-Furthermore, these values are reported for the first (top row) and last (bottom row) iterations of the semi-Lagrangian algorithm.
+Furthermore, these values are reported for the first (top row) and last (bottom row) sub-iterations of the semi-Lagrangian algorithm.
 There are several points to make.
@@
 Interestingly, however, the number of sent and received messages do not seem to affected by the CFL number.
-The bottom row of figure \ref{fig:communication_4096} exhibits a better balance both in the computation and communication volume in the last iteration of the semi-Lagrangian algorithm.
-This can be explained by noting that for the final iteration, the partitioning of $G^{n+1}$ is a good approximation of the partitioning of the departure points on $G^n$.
-\comment{awkward ...
-maybe: the partitioning of the grids $G^n$ and $G^{n+1}$ become very similar ?} \mohammad{not quite ...
-I am not sure what would be the best to phrase it but what i want to say is load balancing really requires you to partition the departure points on $G^n$ and not the $G^{n+1}$ which, of course, you cannot since you do not have that information ...
-the situation becomes better at the final iteration because both partitions are proportional to (Surface area of interface)/(Number of processors) ...} .
+The bottom row of figure \ref{fig:communication_4096} exhibits a better balance both in the computation and communication volume in the last sub-iteration of the semi-Lagrangian algorithm.
+This can be justified by noting that for the final sub-iteration, the partitioning of $G^{n+1}$ is more consistent with the partitioning of the departure points on $G^n$.
 Detailed information about the load balancing and the communication patterns is listed in table \ref{tab:communication}.
@@
 	\end{center}
-	\caption{Performance indicators of the first (top row) and last (bottom row) iterations of the semi-Lagrangian algorithm for the level-set advection on 4096 processes with $\text{CFL} = 1$ (\drawdiamond{black} \comment{this does not work on my computer, I see a W} \mohammad{hum ...
-not sure what to do for this one.
-it should be part of latex...}), $\text{CFL} = 10$ (\drawsquare{red}), and $\text{CFL} = 100$ (\drawcircle{blue}).
+	\caption{Performance indicators of the first (top row) and last (bottom row) sub-iterations of the semi-Lagrangian algorithm for the level-set advection on 4096 processes with $\text{CFL} = 1$ (\drawdiamond{black}), $\text{CFL} = 10$ (\drawsquare{red}), and $\text{CFL} = 100$ (\drawcircle{blue}).
 Increasing the CFL number greatly causes load imbalance during interpolation (a) and increases the communication volume (c).
 However, the CFL number does not seem to appreciably affect the number of messages sent by the processes (b).
-During the last semi-Lagrangian iteration, the initial grid $G_0$ (c.f.
+During the last semi-Lagrangian sub-iteration, the initial grid $G_0$ (c.f.
 Algorithm \ref{alg:semi-lagrangian}) is very close to the final grid.
 As a result, the load imbalance is considerably improved (d).
-Curiously, however, the communication pattern does not seem to be change much between first and last iterations (e,f).}
+Curiously, however, the communication pattern does not seem to be change much between first and last sub-iterations (e,f).}
 	\label{fig:communication_4096}
@@
 	\hline
-	\multirow{2}{*}{CFL} & \multirow{2}{*}{Metric} & \multicolumn{4}{|c|}{First Iteration} & \multicolumn{4}{|c|}{Last Iteration} \\
+	\multirow{2}{*}{CFL} & \multirow{2}{*}{Metric} & \multicolumn{4}{|c|}{First sub-iteration} & \multicolumn{4}{|c|}{Last sub-iteration} \\
 	\cline{3-10}
@@
 	\caption{Detailed load balancing and communication information for the advection test for $\text{CFL} = 1$, $\text{CFL} = 10$, and $\text{CFL} = 100$.
-Note how increasing the CFL number causes load imbalance and increases the communication volume while it does not affect the number of messages sent and received during an iteration of the semi-Lagrangian step.}
+Note how increasing the CFL number causes load imbalance and increases the communication volume while it does not affect the number of messages sent and received during a sub-iteration of the semi-Lagrangian step.}
 	\label{tab:communication}

### 2015-02-07 paper/tex/scaling_results.tex
Added figures with a bigger run for interpolation

@@
 Scaling results are presented for $\alpha = 5 \%$ and $\alpha = 95\%$ for both the small and large problems in figure \ref{fig:interpolation}.
+We also present a third row of results for a much larger problem with roughly 1.66B nodes on a level 14 tree.
 Excellent scaling is obtained for the small problem for $P = 16-512$ even when $95\%$ of the interpolation points belong to a remote process.
 For the larger problem, however, the communication overhead prevents the algorithm from scaling beyond $2048$ processes when $\alpha = 95\%$ (c.f\ table \ref{tab:scaling_interpolation}).
+Note, however, that this is expected since the total time is dominated by communication for $\alpha = 95\%$ and there is very little local work in this case.
+Indeed, the last row of figure \ref{fig:interpolation} show much better scaling behavior on a larger problem size, e.g.
+efficiencies are increased from $e=34\%$ to $e=68\%$ for $\alpha = 95\%$ on 4096 processes.
+This is a typical result with strong scaling and simply implies that our algorithms are scalablable for sufficiently large problems.
 \begin{figure}[htbp]
 	\begin{center}
-		\subfigure[$N_G = 33$M, $\alpha = 5\%$]{\includegraphics[width = 0.48 \textwidth] {figures/host_Interpolation_small_alpha_5.pdf}}
-		\subfigure[$N_G = 33$M, $\alpha = 95\%$]{\includegraphics[width = 0.48 \textwidth] {figures/host_Interpolation_small_alpha_95.pdf}}
+		\subfigure[$N_G = 33$M, $\alpha = 5\%$]{\includegraphics[width = 0.38 \textwidth] {figures/host_Interpolation_small_alpha_5.pdf}} \hspace{0.5 in}
+		\subfigure[$N_G = 33$M, $\alpha = 95\%$]{\includegraphics[width = 0.38 \textwidth] {figures/host_Interpolation_small_alpha_95.pdf}}
 		\\
-		\subfigure[$N_G = 280$M, $\alpha = 5\%$]{\includegraphics[width = 0.48 \textwidth] {figures/host_Interpolation_large_alpha_5.pdf}}
-		\subfigure[$N_G = 280$M, $\alpha = 95\%$]{\includegraphics[width = 0.48 \textwidth] {figures/host_Interpolation_large_alpha_95.pdf}}
+		\subfigure[$N_G = 280$M, $\alpha = 5\%$]{\includegraphics[width = 0.38 \textwidth] {figures/host_Interpolation_large_alpha_5.pdf}} \hspace{0.5 in}
+		\subfigure[$N_G = 280$M, $\alpha = 95\%$]{\includegraphics[width = 0.38 \textwidth] {figures/host_Interpolation_large_alpha_95.pdf}}
+		\\
+		\subfigure[$N_G = 1.66$B, $\alpha = 50\%$]{\includegraphics[width = 0.38 \textwidth] {figures/host_Interpolation_super_large_alpha_50.pdf}} \hspace{0.5 in}
+		\subfigure[$N_G = 1.66$B, $\alpha = 95\%$]{\includegraphics[width = 0.38 \textwidth] {figures/host_Interpolation_super_large_alpha_95.pdf}}
 	\end{center}
@@
 The results indicate excellent scaling for the small test (a-b) and for the large test when $\alpha = 5\%$ (c).
-For the extreme case (d) the algorithm stops scaling at $2048$ processes due to communication overhead.}
+For the extreme case (d) the algorithm stops scaling at $2048$ processes due to communication overhead.
+Note, however, that this merely indicates that the problem size is not large enough for this test case.
+Indeed much better scaling is obtained when the problem size is increased to $N_G = 1.66$B points (e-f).}
 	\label{fig:interpolation}
@@
 	\hline
+	\multirow{3}{*}{Very Large Test} & \multicolumn{1}{|c|}{$P$} & 128     & 256     & 512     & 1024    & 2048   & 4096 \\
+	\cline{2-8} 	                            
+	                            & $\alpha = 50\%$  & -- & -- & $100\%$ & $108\%$ & $102\%$& $78\%$ \\
+	                            & $\alpha = 95\%$  & -- & -- & $100\%$ & $103\%$ & $94\%$ & $64\%$ \\
+	\hline
 	\end{tabular}
-	\caption{Parallel efficiency of the total runtime the interpolation test. Reported efficiencies are based on the lowest number of processes for each test.}
+	\caption{Parallel efficiency of the total runtime of the interpolation algorithm for the small (33M nodes), large (280M nodes), and very large (1.66B nodes) tests.
+Reported efficiencies are based on the lowest number of processes for each test.}
 	\label{tab:scaling_interpolation}

### 2015-02-07 paper/tex/data/3d_interpolation/host/very_large/alpha_0.95/parse_comm.m

@@ -1,0 +1,105 @@
+clear all;
+close all;
+
+folder = { 
+    '.';
+    };
+
+p = [512; 4096];
+xticks = [0, 120, 240, 360, 480;
+          0, 1000, 2000, 3000, 4000];
+it = [6 7;
+      3 4;
+      3 3];
+modes = {'o', 's', 'd'};
+faces = {'b', 'r', 'k'};
+edges = faces;
+
+fs = 16;
+ms = [4 5 5] - 1;
+for i=1:length(p)
+    close all;
+    for k =1:length(folder)    
+%         comm = zeros(p(i), 5);
+%         for j=1:it(k, i)
+%             comm = comm + load(strcat(folder{k}, '/n_', num2str(p(i)), '/interpolation_log_0_', num2str(j-1), '.dat'));
+%         end
+%         comm = comm / it(k, i);
+
+%         comm = load(strcat(folder{k}, '/n_', num2str(p(i)), '/interpolation_log_0_', num2str(it(k,i)-1), '.dat')); postfix='_last_';
+        comm = load(strcat(folder{k}, '/n_', num2str(p(i)), '/interpolation_log_0.dat')); postfix='_first_';
+
+        num_points   = comm(:,1) + comm(:,4);
+        num_messages = (comm(:,3) + comm(:,5));
+        volume       = (4*comm(:,2) + 3*comm(:,4)) * 8 / 1024^2; % recv: 3 doubles (xyz of point) -- send: 4 doubles (1 func + 3 derv)
+
+        h1 = figure(1); hold on
+        plot(num_points, modes{k}, 'markersize', ms(k), 'markeredgecolor', edges{k}, 'markerfacecolor', faces{k});
+        set(gca,'DefaultTextFontname', 'CMU Serif', 'fontsize', fs);
+%         h = legend('$\text{CFL} = 100$', '$\text{CFL} = 10$', '$\text{CFL} = 1$');
+%         set(h, 'fontsize', fs, 'interpreter', 'latex');
+        xlabel('$p$', 'fontsize', fs, 'interpreter', 'latex');
+        ylabel('$N_p$', 'fontsize', fs, 'interpreter', 'latex');    
+        xlim([0 p(i)]); box on;
+        set(gca, 'fontsize', fs);
+        set(gca,'DefaultTextFontname', 'CMU Serif')
+        set(gca, 'XTick', xticks(i,:))
+        set(gca, 'color', 'none');
+        
+        if i == 1
[…]

### 2015-02-07 paper/tex/data/3d_interpolation/parse_interpolation.m

@@ -3,16 +3,16 @@ set(0,'DefaultTextFontname', 'CMU Serif')
 set(0,'DefaultAxesFontName', 'CMU Serif')
 
 type = 'host';
-run  = 'large';
-alpha = 5;
+run  = 'super_large';
+alpha = 50;
 
-folder = strcat(type, '/', run, '/alpha_', num2str(alpha/100));
+folder = strcat(type, '/', run, '/alpha_0.50');%, num2str(alpha/100));
 files = {
 %     'stdout.n_16';
 %     'stdout.n_32';
 %     'stdout.n_64';
-    'stdout.n_128';
-    'stdout.n_256';
+%     'stdout.n_128';
+%     'stdout.n_256';
     'stdout.n_512';
     'stdout.n_1024';
     'stdout.n_2048';
@@ -82,6 +82,7 @@ for j=1:length(events)
                 flag = 1;
             end
             if strfind(line, event);
+%                 line
                 t(j, i) = sscanf(line, '%*s %*d %*f %e %*[^\n]') / 10;
                 break
             end
@@ -113,7 +114,7 @@ close all;
 hfig = figure(1);
 fs = 16;
 ls = 15;
-
+t(:,:) = 2*t(:,:);
 for i=1:length(events)
     loglog(p, t(idx(i),:), modes{i}, 'markersize', 7, 'MarkerfaceColor',faces{i}, 'MarkerEdgeColor','k', 'linewidth', 1 ); hold on;           
     set(gca, 'fontsize', fs);

### 2015-06-04 examples/stefan/main_2d.cpp
adding control on CFL through command line for stefan

@@ -701,6 +701,7 @@ int main (int argc, char* argv[])
   cmd.add_option("save_vtk", "1 to export vtu images, 0 otherwise");
   cmd.add_option("save_every_n", "export images every n iterations");
   cmd.add_option("max_iter", "maximum number of iterations");
+  cmd.add_option("n_times_dt", "CFL restriction, dt = n_times_dt * dx/umax");
   cmd.add_option("test", "the test to run. Available options are\
                  \t 0 - frank sphere\n\
                  \t 1 - a single seed\n\
@@ -714,6 +715,7 @@ int main (int argc, char* argv[])
   bool save_vtk = cmd.get("save_vtk", 1);
   int save_every_n = cmd.get("save_every_n", 1);
   int max_iter = cmd.get("max_iter", INT_MAX);
+  double n_times_dt = cmd.get("n_times_dt", 1);
 
   splitting_criteria_cf_t data(lmin, lmax, &level_set, 1.2);
 
@@ -844,9 +846,9 @@ int main (int argc, char* argv[])
     MPI_Allreduce(MPI_IN_PLACE, &max_norm_u, 1, MPI_DOUBLE, MPI_MAX, p4est->mpicomm);
 
 #ifdef P4_TO_P8
-    dt = min(1.,1/max_norm_u) * .5 * MIN(dxyz_min[0], dxyz_min[1], dxyz_min[2]);
+    dt = min(1.,1/max_norm_u) * n_times_dt * MIN(dxyz_min[0], dxyz_min[1], dxyz_min[2]);
 #else
-    dt = min(1.,1/max_norm_u) * .5 * MIN(dxyz_min[0], dxyz_min[1]);
+    dt = min(1.,1/max_norm_u) * n_times_dt * MIN(dxyz_min[0], dxyz_min[1]);
 #endif
 
     if(tn+dt>tf) dt = tf-tn;

### 2015-06-04 paper/tex/abstract.tex
starting revisions ...

@@
 \begin{abstract}
-We present scalable parallel algorithms for the level-set technology on adaptive Quadtree and Octree Cartesian grids.
+We present scalable parallel algorithms for the level-set method on adaptive Quadtree and Octree Cartesian grids.
 The algorithms are based on a domain decomposition technique and implemented using \texttt{MPI} and the open-source \texttt{p4est} library.
 An important contribution is a scalable parallel semi-Lagrangian method which, similar to its serial implementation, is free of any time-step restrictions.
-This is achieved by introducing a scalable global interpolation scheme on adaptive Quadtree and Octree grids.
+This is achieved by introducing a scalable global interpolation scheme on adaptive tree-based grids.
 Moreover, we present a simple parallel reinitialization scheme using the pseudo-time transient formulation.
-Both parallel algorithms scale on the Stampede supercomputer, where we are limited to 4096 cores at most.
-Finally a relevant application of the algorithms is presented in modeling the crystallization phenomena by solving a Stefan problem, illustrating a level of detailed calculation that would be impossible without a parallel adaptive strategy.
+Both parallel algorithms scale on the Stampede supercomputer, where we are limited to 4096 cores currently.
+Finally a relevant application of the algorithms is presented in modeling the crystallization phenomena by solving a Stefan problem, illustrating a level of detail that would be impossible to achieve without a parallel adaptive strategy.
 We believe that the algorithms presented in this article will be of interest and useful to researchers working with the level-set framework and modeling multi-scale physics in general.

### 2015-06-04 paper/tex/applications.tex
Added figures with a bigger run for interpolation
starting revisions ...

@@
 \begin{equation} \label{eq:stefan_gibbs_tompson}
-T_s = T_l = T_{\Gamma} = -\epsilon_c \kappa - \epsilon_v (\underline{\mathbf{V}} \cdot \underline{\mathbf{n}}),
+T_s = T_l = T_{\Gamma} = -\epsilon_c \kappa - \epsilon_v (\underline{\mathbf{u}} \cdot \underline{\mathbf{n}}),
 \end{equation}
-where $\kappa$ is the local interface curvature, $\underline{\mbf{V}}$ is the velocity of the interface, $\underline{\mathbf{n}}$ is the outward normal to the solidification front and $\epsilon_c$ and $\epsilon_v$ are the surface tension and kinetic undercooling coefficients.
-The interface velocity $\underline{\mbf{V}}$ is defined from the jump in the heat flux at the interface,
+where $\kappa$ is the local interface curvature, $\underline{\mbf{u}}$ is the velocity of the interface, $\underline{\mathbf{n}}$ is the outward normal to the solidification front and $\epsilon_c$ and $\epsilon_v$ are the surface tension and kinetic undercooling coefficients.
+The interface velocity $\underline{\mbf{u}}$ is defined from the jump in the heat flux at the interface,
 \begin{equation} \label{eq:stefan_velocity}
-(\underline{\mathbf{V}} \cdot \underline{\mathbf{n}}) = - \left[ D_l \pd{T_l}{\underline{\mathbf{n}}} - D_s \pd{T_s}{\underline{\mathbf{n}}} \right].
+(\underline{\mathbf{u}} \cdot \underline{\mathbf{n}}) = - \left[ D_l \pd{T_l}{\underline{\mathbf{n}}} - D_s \pd{T_s}{\underline{\mathbf{n}}} \right].
 \end{equation}
@@
 \begin{equation} \label{eq:stefan_dt}
-\Delta t = 5 ~ \Delta x_{min} ~ \min(1,1/\max \lVert \underline{\mathbf{V}} \rVert),
+\Delta t = 5 ~ \Delta x_{min} ~ \min(1,1/\max \lVert \underline{\mathbf{u}} \rVert),
 \end{equation}
@@
 \State Extrapolate $T^{n+1}_s$ from $\Omega^+$ to $\Omega^-$ and $T^{n+1}_l$ from $\Omega^-$ to $\Omega^+$.
-\State Compute the velocity field $\underline{\mathbf{V}}$ according to \eqref{eq:stefan_velocity}.
+\State Compute the velocity field $\underline{\mathbf{u}}$ according to \eqref{eq:stefan_velocity}.
 \State Compute the time step dt following \eqref{eq:stefan_dt}.
@@
 \begin{center}
-\begin{tabular}{c c c c c}
-(a) &
-\includegraphics[width=.2\textwidth]{stefan/nb2_iter19.png} &
-\includegraphics[width=.2\textwidth]{stefan/nb2_iter39.png} &
-\includegraphics[width=.2\textwidth]{stefan/nb2_iter59.png} &
-\includegraphics[width=.2\textwidth]{stefan/nb2_iter79.png}
-\end{tabular}
-
-\subfigure[]{
-\includegraphics[width=.2\textwidth]{stefan/nb2_iter19.png}
-\includegraphics[width=.2\textwidth]{stefan/nb2_iter39.png}
-\includegraphics[width=.2\textwidth]{stefan/nb2_iter59.png}
-\includegraphics[width=.2\textwidth]{stefan/nb2_iter79.png}
+\subfigure{
+\includegraphics[width=.24\textwidth]{stefan/nb2_iter19.png}
+\includegraphics[width=.24\textwidth]{stefan/nb2_iter39.png}
+\includegraphics[width=.24\textwidth]{stefan/nb2_iter59.png}
+\includegraphics[width=.24\textwidth]{stefan/nb2_iter79.png}
 }
-\subfigure[]{
-\includegraphics[width=.2\textwidth]{stefan/nb5_iter19.png}
-\includegraphics[width=.2\textwidth]{stefan/nb5_iter39.png}
-\includegraphics[width=.2\textwidth]{stefan/nb5_iter59.png}
-\includegraphics[width=.2\textwidth]{stefan/nb5_iter79.png}
+\subfigure{
+\includegraphics[width=.24\textwidth]{stefan/nb5_iter19.png}
+\includegraphics[width=.24\textwidth]{stefan/nb5_iter39.png}
+\includegraphics[width=.24\textwidth]{stefan/nb5_iter59.png}
+\includegraphics[width=.24\textwidth]{stefan/nb5_iter79.png}
 }
-\subfigure[]{
-\includegraphics[width=.2\textwidth]{stefan/nb8_iter19.png}
-\includegraphics[width=.2\textwidth]{stefan/nb8_iter39.png}
-\includegraphics[width=.2\textwidth]{stefan/nb8_iter59.png}
-\includegraphics[width=.2\textwidth]{stefan/nb8_iter79.png}
+\subfigure{
+\includegraphics[width=.24\textwidth]{stefan/nb8_iter19.png}
+\includegraphics[width=.24\textwidth]{stefan/nb8_iter39.png}
+\includegraphics[width=.24\textwidth]{stefan/nb8_iter59.png}
+\includegraphics[width=.24\textwidth]{stefan/nb8_iter79.png}
 }
-\subfigure[]{
-\includegraphics[width=.2\textwidth]{stefan/nb12_iter19.png}
-\includegraphics[width=.2\textwidth]{stefan/nb12_iter39.png}
-\includegraphics[width=.2\textwidth]{stefan/nb12_iter59.png}
-\includegraphics[width=.2\textwidth]{stefan/nb12_iter79.png}
+\subfigure{
+\includegraphics[width=.24\textwidth]{stefan/nb12_iter19.png}
+\includegraphics[width=.24\textwidth]{stefan/nb12_iter39.png}
+\includegraphics[width=.24\textwidth]{stefan/nb12_iter59.png}
+\includegraphics[width=.24\textwidth]{stefan/nb12_iter79.png}
 }

### 2015-06-04 paper/tex/introduction.tex
starting revisions ...

@@
 This convenience, however, comes at a price.
-First, compared to an explicit method, e.g.\ front tracking \cite{Juric:96:A-Front-Tracking-Met, Tryggvason;Bunner;Esmaeeli;etal:01:A-Front-Tracking-Met}, the level-set method is typically less accurate and mass conservation could be a problem; although progress has been made in resolving this issue \cite{Enright;Fedkiw;Ferziger;etal:02:A-Hybrid-Particle-Le}.
-Second, the level-set function has to be defined in a higher dimensional space that that of the interface.
+First, compared to an explicit method, e.g.\ front tracking \cite{Juric:96:A-Front-Tracking-Met, Tryggvason;Bunner;Esmaeeli;etal:01:A-Front-Tracking-Met}, the level-set method is typically less accurate and mass conservation could be a problem, although progress has been made in resolving this issue \cite{Enright;Fedkiw;Ferziger;etal:02:A-Hybrid-Particle-Le}.
+Second, the level-set function has to be defined in a higher dimensional space than that of the interface.
 If only the location of the interface is needed, the added dimension greatly increases the overall computational cost.
@@
 Moreover, the use of a general-purpose graph partitioner adds extra overhead that can limit the overall scalability even further.
-Interestingly, tree-based grids have nice spatial ordering that naturally leads to the concept of space-filling curves (SFCs) that can be efficiently exploited for parallel load balancing \cite{Aluru;Sevilgen:97:Parallel-domain-deco,Campbell;Devine;Flaherty;etal:03:Dynamic-octree-load-}.
+Interestingly, tree-based grids have nice spatial ordering that naturally leads to the concept of space-filling curves (SFCs) and can be efficiently exploited for parallel load balancing \cite{Aluru;Sevilgen:97:Parallel-domain-deco,Campbell;Devine;Flaherty;etal:03:Dynamic-octree-load-}.
 
@@
 
-In many applications, it is desirable that the level-set function has signed-distance property, i.e $|\nabla \phi| = 1$.
+In many applications, it is desirable that the level-set function has the signed-distance property, i.e $|\nabla \phi| = 1$.
 Generally, there are two approaches to enforce this property, either by solving the pseudo-time transient reinitialization equation \cite{Sussman;Smereka;Osher:94:A-Level-Set-Approach, Osher;Fedkiw:01:Level-Set-Methods:-A}
@@
 In \cite{Zhao:07:Parallel-implementat} a parallel FSM method was presented for the first time, which suffered from a plateau in the speed up.
-A scalable FSM was more recently proposed in \cite{Detrixhe;Gibou;Min:13:A-parallel-fast-swee}, where the Cuthill-McKee numbering was utilized to exploit more parallelism.
+A scalable FSM was more recently proposed in \cite{Detrixhe;Gibou;Min:13:A-parallel-fast-swee}, where the Cuthill-McKee numbering was utilized to improve scalability.
 More recently a two-scale, hybrid FMM-FSM was presented in \cite{Chacon;Vladimirsky:13:A-parallel-Heap-Cell} which, albeit being more complicated to implement, promises even better scalability.
@@
 However, unlike FMM, FIM avoids sorting the list and allows for concurrent updating of all nodes in an iterative fashion.
-In this article we simply choose the pseudo-time transient formulation for two reasons: 1) it is considerably easier to parallelize on Quadtrees and Octrees and 2) we are merely interested in the sign-distance property close to the interface, which only requires a few iterations.
+In this article we choose the pseudo-time transient formulation for two reasons: 1) it is considerably easier to parallelize on Quadtrees and Octrees and 2) we are merely interested in the sign-distance property close to the interface, which only requires a few iterations.
 
@@
 In section \ref{sec:scaling}, we provide several examples that illustrate the scalability of our algorithms.
-Finally, we close by providing an important application of our strategy by considering the simulation of the solidification process by solving a Stefan problem in section \ref{sec:application}.
+Finally, we close by providing an application of our method by considering the simulation of the solidification process by solving a Stefan problem in section \ref{sec:application}.

### 2015-06-04 paper/tex/levelset_method.tex
final pass (mohammad+frederic+arthur) done. One last thing remains, scaling with large number of processes and for a heavier load
starting revisions ...

@@
 \be
-\phi_t + \underline{\mbf{u}} \cdot \nabla \phi = 0.
+\phi_t + \underline{\mbf{u}} \cdot \underline{\nabla} \phi = 0.
 \label{eq:ls}
@@
 When the velocity field does not depend on the level-set function itself, equation \eqref{eq:ls} can be solved using the semi-Lagrangian method.
-An important advantage of the semi-Lagrangian method over the regular finite difference method is its unconditional stability, which allows for arbitrarily large time steps.
-This is particularly important when using adaptive grids, as higher grid resolutions translate into unpractical smaller time steps.
+An important advantage of the semi-Lagrangian method over the regular finite difference method is its unconditional stability that allows for arbitrarily large time steps.
+This is particularly important when using adaptive grids since higher grid resolutions translate into impractically small time steps.
 
 In general, an infinite number of level-set functions can describe the same zero contour and thus an interface.
-However, it is desirable to chose a function with the signed distance property, i.e.\ $|\nabla \phi| = 1$.
+However, it is desirable to chose a function with the signed distance property $|\underline{\nabla} \phi| = 1$.
 As detailed in section \ref{sec:introduction}, we solve the pseudo-time transient reinitialization equation \cite{Sussman;Smereka;Osher:94:A-Level-Set-Approach, Osher;Fedkiw:01:Level-Set-Methods:-A} to achieve this property,
 \be
-\phi_\tau + S(\phi_0)\left(|\nabla \phi| - 1\right) = 0,
+\phi_\tau + S(\phi_0)\left(\lvert \underline{\nabla} \phi \rvert - 1\right) = 0,
 \label{eq:reinitialization}
 \ee
-where $\tau$ is a pseudo time step, $\phi_0$ is any level-set function that correctly describes the interface location and $S(\phi_0)$ is an appropriate approximation of the sign function.
+where $\tau$ is a pseudo time step, $\phi_0$ is any level-set function that correctly describes the interface location and $S(\phi_0)$ is an approximation of the sign function.
 Here, we do not go into the details of the sequential algorithms for solving equations \eqref{eq:ls} and \eqref{eq:reinitialization}.
 Instead, we note that the algorithms presented in section \ref{sec:parallel algorithms} are based on the sequential methods presented in \cite{Min;Gibou:07:A-second-order-accur} and refer the interested reader to the aforementioned articles and references therein for more details.
-
-% \indent The level-set method is a method to capture interfaces implicitely by constructing a function $\phi$, called the level-set function, that is negative on on side of the domain $\Omega^-$ and positive on the other side $\Omega^+$, thus defining the interface $\Gamma$ as the zero contour of $\phi$, $\Gamma = \{\mathbf{x} \vert \phi(\mathbf{x}=0\}$.\\
-
-% If infinitely many functions satisfy these requirements, the most natural and most convenient representation is a signed distance function to the interface.
-The level-set function is therefore reinitialized to a signed distance function at every time step by solving the reinitialization equation
-
-% \begin{equation} \label{eq::reinitialization}
-% \pd{\phi}{\tau} + \mathrm{sign}(\phi) \left( \lvert \phi \rvert - 1 \right) = 0,
-% \end{equation}
-
-% where $\tau$ is a fictitious time.
-The discretization and implementation of this procedure will be discussed later in section \ref{section::reinitialization}.
-Geometrical quantities such as the normal to the interface and the local curvature can be computed straightforwardly from the level-set function as
-
-% \begin{equation*}
-% \mathbf{n} = \frac{\nabla \phi}{\lvert \nabla \phi \rvert},
-% \end{equation*}
-
-% where it should be noted that $\lvert \nabla \phi \rvert = 1$ when the level-set function is reinitialized, and
-
-% \begin{equation*}
-% \kappa = \nabla \cdot \mathbf{n} =  \nabla \cdot \left( \frac{\nabla \phi}{\lvert \nabla \phi \rvert} \right).
-% \end{equation*}
-
-% \section{The \texttt{p4est} library}
-
-% The work we present relies on the parallel octree library \texttt{p4est} \cite{Burstedde;Wilcox;Ghattas:11:p4est:-Scalable-Algo}.
-It is scalable implementation of the general octree structure in a massively parallel mpi environment.
-We limit our presentation of \texttt{p4est} to the relevant information for our application, for more details the reader is invited to read the original article \cite{Burstedde;Wilcox;Ghattas:11:p4est:-Scalable-Algo}.
-
-% The \texttt{p4est} library provides the geometrical information for the leaves of an octree which includes the neighboring information for the leaves (called quadrants in the original article) of the tree, a layer of ghost quadrants and ghost vertices for each process, the coarsening and refining procedures, and encapsulates the communication between processes.
-However, it does not provide the vertical structure of the octrees.
-If it is generally unnecessary, it is a desirable feature for a finite differences based code as it speeds up the access to arbitrary quadrants information dramatically.
-Thus, we choose to reconstruct the local vertical structure of the octree on every processes.
-Note that given the local and ghost layer information, this is an entirely local procedure and no communication is needed.
-
-% In addition to the owned and ghost categories, we choose to group the vertices in \textit{local} and \textit{layer} sets.
-We define the layer vertices as the vertices that are part of another process's ghost layer and the local vertices as the vertices that are not part of any other process's ghost layer.
-Figure \ref{fig::local_layer_vertices} gives a graphical representation of a boundary between two processes and the classification of the various quadrants and vertices.
-The idea behind this sorting is that layer vertices are needed by other processes for finite difference calculations and therefore require to be synchronized while the local vertices are used only by the process they belong to.
-We make use of this classification to implement scalable algorithms for the level-set method, as explained in the following sections.
-
-% \begin{figure}[ht!]
-% \begin{center}
-% \includegraphics[width=.4\textwidth]{pictures/layer_nodes.pdf}
-% \caption{Illustration of a possible layout at the interface between two processes.
-The red quarants and the orange and purples vertices belong to process 1 while the green quadrants and the green and black vertices belong to process 2.
-The green G denotes the ghost quadrants for process 2, the green dots are the ghost vertices for process 1, and the purple dots correspond to layer vertices for process 1 (and therefore ghost vertices for process 2).} \label{fig::local_layer_vertices}
-% \end{center}
-% \end{figure}

### 2015-06-04 paper/tex/abstract.tex
adding placeholder for carsten

@@
 We present scalable parallel algorithms for the level-set method on adaptive Quadtree and Octree Cartesian grids.
-The algorithms are based on a domain decomposition technique and implemented using \texttt{MPI} and the open-source \texttt{p4est} library.
-An important contribution is a scalable parallel semi-Lagrangian method which, similar to its serial implementation, is free of any time-step restrictions.
+The algorithms are based on a \textcolor{red}{domain decomposition technique} and implemented using \texttt{MPI} and the open-source \texttt{p4est} library.
+We present a scalable parallel semi-Lagrangian method which, similar to its serial implementation, is free of any time-step restrictions.
 This is achieved by introducing a scalable global interpolation scheme on adaptive tree-based grids.

### 2015-06-04 paper/tex/introduction.tex
adding placeholder for carsten

@@
 First, compared to an explicit method, e.g.\ front tracking \cite{Juric:96:A-Front-Tracking-Met, Tryggvason;Bunner;Esmaeeli;etal:01:A-Front-Tracking-Met}, the level-set method is typically less accurate and mass conservation could be a problem, although progress has been made in resolving this issue \cite{Enright;Fedkiw;Ferziger;etal:02:A-Hybrid-Particle-Le}.
-Second, the level-set function has to be defined in a higher dimensional space than that of the interface.
-If only the location of the interface is needed, the added dimension greatly increases the overall computational cost.
+Second, the level-set function has to be defined in a one dimension higher space than that of the interface.
+If only the location of the interface is needed, the added dimension greatly increases the overall computational cost for uniform grids.
 One way to avoid this problem is by computing the level-set only close to the interface, e.g.\ as in the narrow-band level-set method \cite{Adalsteinsson;Sethian:95:A-Fast-Level-Set-Met} or, more recently, by using a hash table to restrict both computation and storage requirements \cite{Brun;Guittet;Gibou:12:A-local-level-set-me}.
@@
 More recently, authors in \cite{Burstedde;Wilcox;Ghattas:11:p4est:-Scalable-Algo} extended these ideas to a collection, or a ``forest'', of Octrees that are connected through a common, potentially unstructured, hexahedral grid.
-This forest is then partitioned in parallel using a global Morton curve.
+\textcolor{red}{How about adding the p4est comments here ?} This forest is then partitioned in parallel using a global Morton curve.
 The implementation of these algorithms, which were shown to scale to more than 200,000 cores, is publicly available through a simple API provided by the \texttt{p4est} library \cite{p4est-github}.

### 2015-06-19 paper/tex/abstract.tex
Edit abstract

@@
 \begin{abstract}
-We present scalable parallel algorithms for the level-set method on adaptive Quadtree and Octree Cartesian grids.
-The algorithms are based on a \textcolor{red}{domain decomposition technique} and implemented using \texttt{MPI} and the open-source \texttt{p4est} library.
-We present a scalable parallel semi-Lagrangian method which, similar to its serial implementation, is free of any time-step restrictions.
-This is achieved by introducing a scalable global interpolation scheme on adaptive tree-based grids.
-Moreover, we present a simple parallel reinitialization scheme using the pseudo-time transient formulation.
-Both parallel algorithms scale on the Stampede supercomputer, where we are limited to 4096 cores currently.
-Finally a relevant application of the algorithms is presented in modeling the crystallization phenomena by solving a Stefan problem, illustrating a level of detail that would be impossible to achieve without a parallel adaptive strategy.
-We believe that the algorithms presented in this article will be of interest and useful to researchers working with the level-set framework and modeling multi-scale physics in general.
+We present scalable
+% parallel
+algorithms for the level-set method on dynamic-adaptive Quadtree and Octree
+Cartesian grids. The algorithms are fully parallelized
+% on a \textcolor{red}{domain decomposition technique}
+and implemented using the \texttt{MPI} standard and the open-source
+\texttt{p4est} library.
+We solve the level set equation with a semi-Lagrangian method which, similar to
+its serial implementation, is free of any time-step restrictions.
+This is achieved by introducing a scalable global interpolation scheme on
+adaptive tree-based grids.
+Moreover, we present a simple parallel reinitialization scheme using the
+pseudo-time transient formulation. Both parallel algorithms scale on the
+Stampede supercomputer,
+% where we are limited to 4096 cores currently.
+where we are currently using up to 4096 CPU cores.
+Finally, a relevant application of the algorithms is presented in modeling a
+crystallization phenomenon by solving a Stefan problem, illustrating a level of
+detail that would be impossible to achieve without a parallel adaptive
+strategy.
+We believe that the algorithms presented in this article will be of interest
+and useful to researchers working with the level-set framework and modeling
+multi-scale physics in general.
 \end{abstract}
 \begin{keyword}
-Quadtree/Octree Grids \sep Parallel Computing \sep Space Filling Curves \sep Semi-Lagrangian Method \sep Level-set Method
+Quadtree/Octree Grids \sep Parallel Computing \sep Space Filling Curves \sep
+Semi-Lagrangian Method \sep Level-Set Method
 \end{keyword}

### 2015-06-19 paper/tex/abstract.tex
Go through introduction

@@
 \begin{keyword}
-Quadtree/Octree Grids \sep Parallel Computing \sep Space Filling Curves \sep
-Semi-Lagrangian Method \sep Level-Set Method
+Quadtree/Octree Grids \sep Parallel Computing \sep Semi-Lagrangian Method \sep
+Level-Set Method \sep Adaptive Mesh Refinement \sep Space Filling Curves
+\sep Stefan Problem
 \end{keyword}

### 2015-06-19 paper/tex/introduction.tex
Go through introduction

@@
 However, even though the use of adaptive grids can dramatically reduce the computational cost, performing high-resolution three dimensional calculations of complex interfacial problems, e.g.\ crystal growth in binary alloys \cite{Theillard;Gibou;Pollock:14:A-Sharp-Computationa}, could be prohibitively expensive or even impossible on a serial machine.
-In this paper we extend the level-set technology on Quad-/Oc-tree by proposing parallel algorithms for distributed memory machines using a domain decomposition technique.
+In this paper we extend the level-set technology on Quad-/Octrees by proposing
+parallel algorithms for distributed memory machines using a domain
+decomposition technique.
 
-One of the main challenges in parallelizing algorithms on adaptive grids is handling the grid itself.
-One option is to replicate the entire grid on each process and use off-the-shelf graph partitioners, e.g.\ ParMetis \cite{Karypis;Kumar:98:A-parallel-algorithm} or Zoltan \cite{Boman;Catalyurek;Chevalier;etal:12:The-Zoltan-and-Isorr}, for load balancing and domain decomposition.
-For instance, this was the approach originally taken by the \texttt{deal.II} library \cite{Bangerth;Hartmann;Kanschat:07:deal.II----a-General}.
-This approach, however, is only scalable to a few hundred processes at best and is limited by the size of the grid itself that can fit in memory.
-Moreover, the use of a general-purpose graph partitioner adds extra overhead that can limit the overall scalability even further.
-Interestingly, tree-based grids have nice spatial ordering that naturally leads to the concept of space-filling curves (SFCs) and can be efficiently exploited for parallel load balancing \cite{Aluru;Sevilgen:97:Parallel-domain-deco,Campbell;Devine;Flaherty;etal:03:Dynamic-octree-load-}.
+One of the main challenges in parallelizing algorithms on adaptive grids is
+handling the grid itself.
+One option is to replicate the entire grid on each process and to employ serial
+ordering techniques, as originally implemented by the \texttt{deal.II} library
+\cite{Bangerth;Hartmann;Kanschat:07:deal.II----a-General}, or serial
+graph partitioners such as METIS \cite{KarypisKumar95}.
+%
+% off-the-shelf graph partitioners, e.g.\ ParMetis
+% \cite{Karypis;Kumar:98:A-parallel-algorithm} or Zoltan
+% \cite{Boman;Catalyurek;Chevalier;etal:12:The-Zoltan-and-Isorr}, for load
+% balancing and domain decomposition.
+% For instance, this was the approach originally taken by the \texttt{deal.II}
+% library \cite{Bangerth;Hartmann;Kanschat:07:deal.II----a-General}.
+%
+This approach, however, is only scalable to a few hundred processes at best and
+is limited by the size of the grid itself that can fit in memory.
+Even though parallel general-purpose partitioners have since been popularized
+\cite{Karypis;Kumar:98:A-parallel-algorithm,
+Boman;Catalyurek;Chevalier;etal:12:The-Zoltan-and-Isorr} and
+the scalability of partitioning algorithms for unstructured grids
+has been improved (see e.g.\ \cite{SahniZhouShephardEtAl09}),
+% the use of a general-purpose graph partitioner
+their use
+adds extra overhead that can limit the overall scalability.
+Interestingly, tree-based grids have a nice spatial ordering that naturally
+leads to the concept of space-filling curves (SFCs) and can be efficiently
+exploited for parallel load balancing
+\cite{Aluru;Sevilgen:97:Parallel-domain-deco, GriebelZumbusch99,
+      Campbell;Devine;Flaherty;etal:03:Dynamic-octree-load-}.
 
@@
 For instance, \texttt{Octor} \cite{Tu;OHallaron;Ghattas:05:Scalable-parallel-oc} uses a Morton curve (also known as Z-curve) for traversing the leaves of an Octree for indexing and load balancing and has been scaled up to 62,000 cores \cite{Burstedde;Ghattas;Gurnis;etal:08:Scalable-adaptive-ma}.
-\texttt{Dendro} \cite{Sampath;Adavani;Sundar;etal:08:Dendro:-parallel-alg} is another example of an Octree code in which similar ideas are used for parallel partitioning and development of a parallel geometric multigrid that has been scaled up to about 32,000 cores \cite{Sampath;Biros:10:A-parallel-geometric}.
-More recently, authors in \cite{Burstedde;Wilcox;Ghattas:11:p4est:-Scalable-Algo} extended these ideas to a collection, or a ``forest'', of Octrees that are connected through a common, potentially unstructured, hexahedral grid.
-\textcolor{red}{How about adding the p4est comments here ?} This forest is then partitioned in parallel using a global Morton curve.
-The implementation of these algorithms, which were shown to scale to more than 200,000 cores, is publicly available through a simple API provided by the \texttt{p4est} library \cite{p4est-github}.
-In fact the algorithms presented in this paper are implemented on top of the \texttt{p4est} library and we do not discuss any algorithm that is already covered in \cite{Burstedde;Wilcox;Ghattas:11:p4est:-Scalable-Algo}.
-We also make use of the popular \texttt{PETSc} \cite{Balay;Abhyankar;Adams;etal:14:PETSc-Web-page} library for linear algebra and its parallel primitives, such as parallel ghosted vector and scatter/gather operations, which simplifies the implementation.
+\texttt{Dendro} \cite{Sampath;Adavani;Sundar;etal:08:Dendro:-parallel-alg} is
+an example of a so-called linear Octree code in which new algorithms are
+introduced for parallel partitioning and the development of a parallel
+geometric multigrid that has been scaled up to about 32,000 cores
+\cite{Sampath;Biros:10:A-parallel-geometric}.
+More recently, authors in
+\cite{Burstedde;Wilcox;Ghattas:11:p4est:-Scalable-Algo} extended these ideas by
+optionally allowing for a collection, or a ``forest'', of connected Octrees,
+% that are connected through a common, potentially unstructured, hexahedral grid.
+which is partitioned in parallel using a global Morton curve.
+%
+% \textcolor{red}{How about adding the p4est comments here ?}
+%
+% An implementation of these algorithms is publicly available
+%
+% provides a simple API for handling the grid.
+The \texttt{p4est} library \cite{p4est-github} provides a publicly available
+implementation of these algorithms that is equally efficient for a single tree
+as well as multiple trees and has been shown to scale to more than
+% 200,000 cores,
+450,000 CPU cores \cite{IsaacBursteddeWilcoxEtAl15}.
+In fact, the algorithms presented in this paper are implemented on top of
+the \texttt{p4est} API using a single octree to represent the unit cube.
+% In fact, the algorithms presented in this paper are implemented on top of the
+% \texttt{p4est} library \cite{p4est-github} that is publicly available and
+% provides a simple API for handling the grid.
+%
+% and we do not discuss any algorithm that is already covered in
+% \cite{Burstedde;Wilcox;Ghattas:11:p4est:-Scalable-Algo}.
+We also make use of the popular \texttt{PETSc}
+\cite{Balay;Abhyankar;Adams;etal:14:PETSc-Web-page} library for linear algebra
+and its parallel primitives, such as parallel ghosted vector and scatter/gather
+operations, which simplifies the implementation.
 
@@
 The reason for this is twofold.
-First, depending on the CFL number, the departure points may end up outside the ghost region and in remote processes that are potentially far away.
-This requires a very dynamic and nonuniform communication pattern which is complicated to implement.
-For an adaptive grid, the situation is even more complicated due to the asymmetric nature of the communication (c.f.
-section \ref{sec:parallel algorithms}).
-Second, load balancing could be an issue for large CFL numbers and nonuniform velocity fields, due to clustering of departure points, which can thus considerably restrict the scalability of the algorithm.
-Both of these problems, of course, could be avoided by choosing $\text{CFL} \le 1$ but that would defeat the purpose of using the semi-Lagrangian algorithm in the first place.
+First, depending on the CFL number, the departure points may end up outside the
+ghost region and in remote processes that are potentially far away.
+This requires a very dynamic and nonuniform communication pattern that is
+challenging to implement.
+For an adaptive grid, the situation is even more complex due to the asymmetric
+nature of the communication pattern (cf.\ section \ref{sec:parallel
+algorithms}).
+Second, load balancing could be an issue for large CFL numbers and nonuniform
+velocity fields, due to clustering of departure points, which can thus
+considerably restrict the scalability of the algorithm. Both of these problems,
+of course, could be avoided by choosing $\text{CFL} \le 1$ but that would
+defeat the purpose of using the semi-Lagrangian algorithm in the first place.
 
 Nonetheless, several parallel semi-Lagrangian algorithms have been proposed.
-A simple domain decomposition technique was used in \cite{Thomas;Cote:95:Massively-parallel-s} where the width of the ghost layer is fixed based on the maximum CFL number to ensure that all the departure points are covered by the ghost layer.
+A simple domain decomposition technique was used in
+\cite{Thomas;Cote:95:Massively-parallel-s} where the width of the ghost layer
+is fixed based on the maximum CFL number to ensure that all departure points
+are covered by the ghost layer.
 Good scalings were reported for small CFL numbers ($\text{CFL} \le 2$).
@@
 
-In many applications, it is desirable that the level-set function has the signed-distance property, i.e $|\nabla \phi| = 1$.
+In many applications, it is desirable that the level-set function has the signed-distance property, i.e., $|\nabla \phi| = 1$.
 Generally, there are two approaches to enforce this property, either by solving the pseudo-time transient reinitialization equation \cite{Sussman;Smereka;Osher:94:A-Level-Set-Approach, Osher;Fedkiw:01:Level-Set-Methods:-A}
@@
 Unfortunately, the most popular algorithm for solving the Eikonal equation, i.e.\ the Fast Marching Method \cite{Sethian:96:A-Fast-Marching-Leve,Sethian:99:Level-set-methods-an}, is inherently sequential due to causal relationship between grid points and cannot be easily parallelized.
-The Fast Sweeping Method (FSM) \cite{Zhao:05:A-fast-sweeping-meth} is an alternative method for solving the Eikonal equation iteratively.
-The FSM can be more computationally efficient for simple choices of speed function, e.g.\ as in this context, and for simple interfaces.
+The Fast Sweeping Method (FSM) \cite{Zhao:05:A-fast-sweeping-meth} is an
+alternative for solving the Eikonal equation iteratively.
+The FSM can be more computationally efficient for simple choices of speed
+function, e.g.\ as in this context, and for simple interfaces.
 Moreover, FSM has more potential for parallelization compared to the FMM.
@@
 Unlike the serial FMM, however, parallel FMM potentially requires multiple iterations or ``rollback operations'' to enforce causality across processes.
-Similar ideas are described in details in \cite{Tugurlan:08:Fast-marching-method}.
+Similar ideas are described in detail in \cite{Tugurlan:08:Fast-marching-method}.
 It should be noted that the number of iterations needed for the parallel FMM to converge greatly depends on the complexity of the interface and on the parallel partitioning and, in general, fewer iterations are required if the domains are aligned with the normals to the interface.
@@
 Unfortunately, this approach does not seem to be more effective than a simple static decomposition.
-In \cite{Zhao:07:Parallel-implementat} a parallel FSM method was presented for the first time, which suffered from a plateau in the speed up.
+In \cite{Zhao:07:Parallel-implementat} a parallel FSM method was presented for the first time, which suffered from a plateau in the speedup.
 A scalable FSM was more recently proposed in \cite{Detrixhe;Gibou;Min:13:A-parallel-fast-swee}, where the Cuthill-McKee numbering was utilized to improve scalability.
-More recently a two-scale, hybrid FMM-FSM was presented in \cite{Chacon;Vladimirsky:13:A-parallel-Heap-Cell} which, albeit being more complicated to implement, promises even better scalability.
+A two-scale, hybrid FMM-FSM was presented in \cite{Chacon;Vladimirsky:13:A-parallel-Heap-Cell} which, albeit being more complicated to implement, promises even better scalability.
 Finally, a parallel Fast Iterative Method (FIM) was proposed in \cite{Jeong;Whitaker:08:A-fast-iterative-met}.
@@
 However, unlike FMM, FIM avoids sorting the list and allows for concurrent updating of all nodes in an iterative fashion.
-In this article we choose the pseudo-time transient formulation for two reasons: 1) it is considerably easier to parallelize on Quadtrees and Octrees and 2) we are merely interested in the sign-distance property close to the interface, which only requires a few iterations.
+In this article we choose the pseudo-time transient formulation for two reasons: 1) it is considerably easier to parallelize on Quadtrees and Octrees and 2) we are merely interested in the signed-distance property close to the interface, which only requires a few iterations.
 
-The rest of this article is organized as follows.
-In section \ref{sec:levelset method}, we briefly review the sequential algorithms and discretization methods for the level-set equation on adaptive tree-based grids.
+This article is organized as follows: In section \ref{sec:levelset method}, we briefly review the sequential algorithms and discretization methods for the level-set equation on adaptive tree-based grids.
 These ideas are then extended in section \ref{sec:parallel algorithms} to parallel environments using a domain decomposition method.

### 2015-06-19 paper/tex/conclusion.tex
final pass (mohammad+frederic+arthur) done. One last thing remains, scaling with large number of processes and for a heavier load
Level set method and acknowledgement

@@
 \section*{Acknowledgment} 
-The funding for this project was provided by \mohammad{the funding agency}.
-We would also like to thank the technical staff at the Texas Advanced Computing Center (TACC) and the developer community of PETSc library for their valuable suggestions during this project.
+The funding for this project was provided by ONR N00014-11-1-0027.
+We would also like to thank the technical staff at the Texas Advanced Computing
+Center (TACC) and the developer community of PETSc library for their valuable
+suggestions during this project.
+The third author acknowledges support by the Hausdorff Center for Mathematics
+(HCM) at Bonn University,
+% and the Transregio 32 research collaborative, both
+an initiative funded by the German Research Foundation (DFG).

### 2015-06-19 paper/tex/levelset_method.tex
Level set method and acknowledgement

@@
 
-In general, an infinite number of level-set functions can describe the same zero contour and thus an interface.
-However, it is desirable to chose a function with the signed distance property $|\underline{\nabla} \phi| = 1$.
-As detailed in section \ref{sec:introduction}, we solve the pseudo-time transient reinitialization equation \cite{Sussman;Smereka;Osher:94:A-Level-Set-Approach, Osher;Fedkiw:01:Level-Set-Methods:-A} to achieve this property,
+In general, an infinite number of level-set functions can describe the same
+zero contour and thus the same interface.
+However, it is desirable to choose a function with the signed distance property
+$|\underline{\nabla} \phi| = 1$. As detailed in section \ref{sec:introduction},
+we solve the pseudo-time transient reinitialization equation
+\cite{Sussman;Smereka;Osher:94:A-Level-Set-Approach,
+Osher;Fedkiw:01:Level-Set-Methods:-A} to achieve this property,
 \be
@@
 Here, we do not go into the details of the sequential algorithms for solving equations \eqref{eq:ls} and \eqref{eq:reinitialization}.
-Instead, we note that the algorithms presented in section \ref{sec:parallel algorithms} are based on the sequential methods presented in \cite{Min;Gibou:07:A-second-order-accur} and refer the interested reader to the aforementioned articles and references therein for more details.
+Instead, we note that the parallel algorithms presented in section \ref{sec:parallel
+algorithms} are based on the sequential methods presented earlier in
+\cite{Min;Gibou:07:A-second-order-accur} and refer the interested reader to the
+aforementioned articles and references therein for more details.

### 2015-06-19 paper/tex/parallel_algorithms.tex
starting revisions ...
Go through section 3

@@
 \section{Parallel algorithms}\label{sec:parallel algorithms}
+
+To achieve a good parallel performance of our numerical solver, we must ensure
+the sufficient scalability of all components.  This observation prompted the
+dedicated development and optimization of several techniques, which we discuss
+in the present section.
+
 \subsection{Grid management}
-Adaptive tree-based grids can significantly reduce the computational cost of level-set methods by restricting the fine grid close to the interface where it is most needed \cite{Strain:99:Tree-Methods-for-Mov}.
-Moreover, adaptive tree-based grids are easy to generate in the presence of a signed-distance level-set function \cite{Min;Gibou:07:A-second-order-accur} and can efficiently be encoded using a tree datastructure \cite{Samet:90:Applications-of-Spat}.
-In order to develop scalable parallel algorithms on these grids, it is necessary to parallelize the datastructure and grid manipulation methods such as refinement and coarsening of cells as well as to provide a fast method for grid partitioning and load balancing.
-The \texttt{p4est} library \cite{p4est-github} is a collection of such parallel algorithms that has recently emerged and shown to scale up to 200,000 cores \cite{Burstedde;Wilcox;Ghattas:11:p4est:-Scalable-Algo}.
 
-In \texttt{p4est} the adaptive grid is represented as a non-overlapping collection of trees that are rooted in individual cells of a common coarse grid (c.f.
-figure \ref{fig:p4est_zcurve}).
+Adaptive tree-based grids can significantly reduce the computational cost of
+level-set methods by restricting the fine grid close to the interface where it
+is most needed \cite{Strain:99:Tree-Methods-for-Mov}.
+Moreover, adaptive tree-based grids are easy to generate in the presence of a
+signed-distance level-set function \cite{Min;Gibou:07:A-second-order-accur} and
+can efficiently be encoded using a tree data structure
+\cite{Samet:90:Applications-of-Spat}.
+In order to develop scalable parallel algorithms on these grids, it is
+necessary to parallelize the data structure and grid manipulation methods such
+as refinement and coarsening of cells as well as to provide a fast method for
+grid partitioning and load balancing.
+The \texttt{p4est} library \cite{p4est-github} is a collection of such parallel
+algorithms that has recently emerged and shown to scale up to 458,752 cores
+%\cite{Burstedde;Wilcox;Ghattas:11:p4est:-Scalable-Algo}.
+\cite{IsaacBursteddeWilcoxEtAl15}.
+
+In \texttt{p4est} the adaptive grid is represented as a non-overlapping collection of trees that are rooted in individual cells of a common coarse grid (cf.\ Figure \ref{fig:p4est_zcurve}).
 This common coarse grid, which we will refer to as the ``macromesh'', can in general be an unstructured quadrilateral mesh, in two spatial dimensions, or hexahedral mesh, in three spatial dimensions.
-In this article, however, we limit discussions to simple uniform Cartesian macromeshshes.
-Moreover it is implicitly assumed that the macromesh is small enough that it can be entirely replicated on all processes.
+In this article, however, it is sufficient to limit discussions to simple
+uniform Cartesian macromeshes.
+Moreover, it is implicitly assumed that the macromesh is small enough that it
+can be entirely replicated on all processes.
 For instance, in many of the applications that we are considering in this paper the macromesh is simply a single cell.
@@
 In this article the refinement criteria is chosen based on the distance of individual cells to the interface.
-Specifically a cell $C$ is marked for refinement if:
+Specifically, a cell $C$ is marked for refinement if
 \be
-\min_{\forall v \in V(C)} |\phi| \le \frac{L D}{2},
+\min_{v \in V(C)} |\phi (v)| \le \frac{L D}{2},
 \label{eq:refine}
 \ee
-where $V(C)$ denotes the set of all vertices of cell $C$, $L$ denotes the Lipschitz constant of the level-set function, and $D$ denotes the diagonal size of cell $C$.
-Conversely an existing cell is marked for coarsening if:
+where $V(C)$ denotes the set of all vertices of cell $C$, $L$ denotes the
+Lipschitz constant of the level-set function, and $D$ denotes the diagonal size
+of cell $C$.
+Conversely, an existing cell is marked for coarsening if
 \be
-\min_{\forall v \in V(C)} |\phi| > L D.
+\min_{v \in V(C)} |\phi (v)| > L D.
 \label{eq:coarsen}
 \ee
-We refer to section 3.2 of \cite{Burstedde;Wilcox;Ghattas:11:p4est:-Scalable-Algo} for details on the parallel refinement and coarsening algorithms used in \texttt{p4est}.
+We refer to section 3.2 of
+\cite{Burstedde;Wilcox;Ghattas:11:p4est:-Scalable-Algo} for details on the
+parallel refinement and coarsening algorithms implemented in \texttt{p4est}.
 
-Once the gird is adapted to the interface, it must be partitioned to ensure load balancing across processes.
-This is achieved by constructing a Z-curve that traverses the leaves of all trees in order of tree index (c.f.
-figure \ref{fig:p4est_zcurve}).
-A Z-curve is a Space Filling Curve (SFC) with the important property that cells with close Z-indices are also geometrically close in the physical domain.
-This is beneficial since it leads to both reduction in MPI communications and improvements of the cache performance of several algorithms such as interpolation and finite difference calculations.
-For more details on parallel partitioning in \texttt{p4est} one may refer to section 3.3 of \cite{Burstedde;Wilcox;Ghattas:11:p4est:-Scalable-Algo}.
+Once the grid is adapted to the interface, it must be partitioned to ensure
+load balancing across processes.
+This is achieved by constructing a Z-curve that traverses the leaves of all
+trees in order of the tree index (cf.\ Figure \ref{fig:p4est_zcurve}).
+A Z-curve is a Space Filling Curve (SFC) with the important property that cells
+with close Z-indices are also geometrically close (on average) in the physical
+domain.
+This is beneficial since it leads to both a reduction in MPI communications and
+improvements of the cache performance of several algorithms such as
+interpolation and finite difference calculations.
+For more details on parallel partitioning in \texttt{p4est} one may refer to
+section 3.3 of \cite{Burstedde;Wilcox;Ghattas:11:p4est:-Scalable-Algo}.
 \begin{figure}[hbtp]
@@
 \includegraphics[width = \columnwidth]{figures/p4est_zcurve.pdf}
-\caption{Left: a ``forest'' made up of two trees $T_0$ and $T_1$.
-Parallel partitioning is achieved by first constructing a Z-curve starting at cell ``S'' and ending at cell ``E''.
-Next, the one dimensional curve is split up among processes either uniformly or by assigning different weights to cells.
-Here processes are represented via different colors.
-Note how using the Z-curve naturally leads to clustering of most cells in each domain.
-Right: schematic of a tree data structure representing the forest and its partitioning.}
+\caption{Left: a ``forest'' made up of two trees $T_0$ and $T_1$. Parallel
+partitioning is achieved by first constructing a Z-curve starting at cell ``S''
+and ending at cell ``E''.  Next, the one dimensional curve is split up among
+processes either uniformly or by assigning different weights to cells.  Here
+processes are represented via different colors.  Note how using the Z-curve
+naturally leads to clustering of most cells in each domain.  Right: schematic of
+a tree data structure representing this forest and its partitioning.}
 \label{fig:p4est_zcurve}
@@
 \end{figure}
-Aside from grid manipulation and partitioning, we use two additional features of \texttt{p4est}, namely the generation of ghost layer cells and the creation of a globally unique node indexing.
-These algorithms are detailed in sections 3.5 and 3.6 of \cite{Burstedde;Wilcox;Ghattas:11:p4est:-Scalable-Algo}.
-An important feature of our algorithms is that they are designed for non-graded trees.
-This is important because we can entirely skip tree balancing, which was shown to be one of the most time consuming parts of grid adaptation in \texttt{p4est} \cite{Burstedde;Wilcox;Ghattas:11:p4est:-Scalable-Algo}.
+Aside from grid manipulation and partitioning, we use two additional features
+of \texttt{p4est}, namely the generation of ghost layer cells and the creation
+of a globally unique node indexing.
+These algorithms are detailed in sections 3.5 and 3.6 of
+\cite{Burstedde;Wilcox;Ghattas:11:p4est:-Scalable-Algo}.
+We have specifically extended the latter algorithm such that it can be applied
+to a non-graded refinement pattern.
+This is important because we can entirely skip the 2:1 balance function, which
+was shown to be one of the most time consuming parts of grid adaptation
+in \texttt{p4est}
+\cite{Burstedde;Wilcox;Ghattas:11:p4est:-Scalable-Algo}.
+% (it has since been optimized \cite{IsaacBursteddeGhattas12}).
 
-Finally, in \texttt{p4est} trees are linearized, i.e.
-only the leaves are explicitly stored.
+Finally, in \texttt{p4est} trees are linearized, i.e.\ only the leaves are explicitly stored.
 However, explicit knowledge of the hierarchal structure of the tree is greatly beneficial in several algorithms, e.g.
@@
 Right: each process independently recreates a local forest that is refined to match the local grid and is as coarse as possible elsewhere.
-Note that empty cells are fictitious, i.e.
-they are only required to generate the hierarchal structure and are not matched by any corresponding cell in the global forest.}
+Note that empty cells are fictitious, i.e.\ they are only required to generate the hierarchal structure and are not matched by any corresponding cell in the global forest.}
 \label{fig:reconstruction}
@@
 \subsection{Interpolation and semi-Lagrangian methods}
-As indicated earlier, we use the semi-Lagrangian method to solve equation \eqref{eq:ls} when the velocity field is externally generated, i.e.
-when it does not depend explicitly on the level-set function itself.
+As indicated earlier, we use the semi-Lagrangian method to solve equation \eqref{eq:ls} when the velocity field is externally generated, i.e.\ when it does not depend explicitly on the level-set function itself.
 Let us rewrite equation \eqref{eq:ls} along the characteristic curve $\underline{\mathbf{X}}(t)$ as:
@@
 \ee
-The semi-Lagrangian method integrates equations \eqref{eq:sl} backward in time, i.e.
-starting from the grid $G^{n+1}$ (computed in iteratively as explained later on), we simply write $\phi^{n+1}(\underline{\mathbf{X}}^{n+1}) = \phi(\underline{\mathbf{X}}(t^{n+1}), t^{n+1}) = \phi(\underline{\mathbf{X}}(t^n), t^n) = \phi^n(\underline{\mathbf{X}}_d)$.
+The semi-Lagrangian method integrates equations \eqref{eq:sl} backward in time,
+i.e.\ starting from the grid $G^{n+1}$ (computed in iteratively as explained
+later on), we simply write $\phi^{n+1}(\underline{\mathbf{X}}^{n+1}) =
+\phi(\underline{\mathbf{X}}(t^{n+1}), t^{n+1}) =
+\phi(\underline{\mathbf{X}}(t^n), t^n) = \phi^n(\underline{\mathbf{X}}_d)$.
 Here, the characteristic curves are chosen such that $\underline{\mathbf{X}}(t^{n+1})$ are simply the coordinates of grids points of $G^{n+1}$, and $\underline{\mathbf{X}}_d$ are the departure points, which are computed using the second-order midpoint method \cite{Min;Gibou:07:A-second-order-accur}:
@@
 \ee
-Note that all values at the intermediate point, $\underline{\mathbf{X}}^\star$, and departure point, $\underline{\mathbf{X}}_d$, must be calculated via interpolation from the previous grids $G^{n}$ and $G^{n-1}$.
-Here, we use the stabilized second-order interpolation for $\phi(\underline{\mathbf{X}}_d)$ and the multi-linear interpolation for $\underline{\mbf{u}}^{n+\frac{1}{2}}(\underline{\mathbf{X}}^\star)$ \cite{Min;Gibou:07:A-second-order-accur}.
-Although parallelization of the interpolation process on a shared-memory machine is trivial, the same cannot be said for distributed-memory machines.
-In fact, the parallel interpolation procedure given in Algorithm \ref{alg:interpolation} is probably the most important contribution of this article since the procedure, which is trivial on uniform grids, is challenging in the case of trees because it is not straightforward to identify which processes owns the departure points.
-Indeed, complications arise because not all departure points will reside in the domain owned by the current process.
-Moreover, due to the irregular shapes of the partitions, one cannot even ensure they are entirely owned by neighboring processes.
-At best we can only expect that their locations are bounded by a halo of width $w \le \text{CFL} \: \Delta x_{\min}$ around the local partition, where $x_{\min}$ is the size of the smallest cell in the forest.
-Naturally, if one enforces $\text{CFL} \le 1$, one can ensure that the halo is bounded by the ghost layer, which significantly simplifies the communication problem.
-This assumption, however, defeats the purpose of using a semi-Lagrangian approach, whose purpose is to enable large CFL values.
+Note that all values at the intermediate point, $\underline{\mathbf{X}}^\star$,
+and departure point, $\underline{\mathbf{X}}_d$, must be calculated via
+interpolation from the previous grids $G^{n}$ and $G^{n-1}$.
+Here, we use the stabilized second-order interpolation for
+$\phi(\underline{\mathbf{X}}_d)$ and the multi-linear interpolation for
+$\underline{\mbf{u}}^{n+\frac{1}{2}}(\underline{\mathbf{X}}^\star)$
+\cite{Min;Gibou:07:A-second-order-accur}.
+Although parallelization of the interpolation process on a shared-memory
+machine is trivial, the same cannot be said for distributed-memory machines.
+In fact, the parallel interpolation given in Algorithm \ref{alg:interpolation}
+is probably the most important contribution of this article since this
+procedure, which is trivial on uniform grids, is challenging in the case of
+trees because it is not straightforward to identify which processes owns the
+departure points.
+Indeed, complications arise because not all departure points will reside in the
+domain owned by the current process. Moreover, due to the irregular shapes of
+the partitions, one cannot even ensure they are entirely owned by neighboring
+processes.
+At best we can only expect that their locations are bounded by a halo of width
+$w \le \text{CFL} \: \Delta x_{\min}$ around the local partition, where
+$x_{\min}$ is the size of the smallest cell in the forest.
+Naturally, if one enforces $\text{CFL} \le 1$, one can ensure that the halo is
+bounded by the ghost layer, which significantly simplifies the communication
+problem.
+This assumption, however, defeats the purpose of using a semi-Lagrangian
+approach, whose purpose is to enable large CFL values.
 
@@
 If the point is owned by a local cell, it is added to a local buffer, otherwise we find the process which owns the point and add the point to a separate buffer belonging to the found rank.
-Note that searching the point in the local tree is performed recursively using the hierarchal reconstruction (c.f.\ algorithm \ref{alg:reconstruction}).
+Note that searching the point in the local tree is performed recursively using the hierarchal reconstruction (cf.\ algorithm \ref{alg:reconstruction}).
 Moreover, the owner's rank is found by computing the Z-index of the point and then using a binary search on the Z-curve.
@@
 However, at this point no process knows which processes to expect a message from.
-We solve this problem using a simple \textit{communication matrix} (see figure \ref{fig:communication}).
+We solve this problem using a simple \textit{communication matrix} (see Figure \ref{fig:communication}).
 Other approaches for solving this problem could include using non-blocking collective or one-sided communication operations introduced in the MPI-3 standard \cite{Hoefler;Siebert;Lumsdaine:10:Scalable-communication}.
@@
 Although the distance to the departure point is bounded by $\text{CFL} \: \Delta x_{\min}$, one cannot predict the receiving rank without explicitly searching the entire Z-curve.
-Moreover, the receiving process has no prior knowledge about which processes to check for incoming messages nor does it know anything about the possible message length (i.e.
-number of points).
+Moreover, the receiving process has no prior knowledge about which processes to check for incoming messages nor does it know anything about the possible message length (i.e.\ number of points).
 Middle: a directed graph illustrating the communication pattern among processes with arrows representing the direction in which messages are sent.
@@
 
-To solve the communication problem, we first compute the adjacency matrix of the communication pattern, i.e. we construct the matrix $A_{P \times P}$, where $P$ is the number of processes, such that
+To solve the communication problem, we first compute the adjacency matrix of the communication pattern, i.e.\ we construct the matrix $A_{P \times P}$, where $P$ is the number of processes, such that
 \ben
@@
 \een
-Note that this matrix is also distributed among processes, i.e. each row is owned by a separate rank. Next, we compute
+Note that this matrix is also distributed among processes, i.e.\ each row is owned by a separate rank. Next, we compute
 \ben
-S_i = \sum_j a_{ij} \hspace{5 mm} \text {and} \hspace{5 mm} R_i = \sum_j a_{ji}
+S_i = \sum_j a_{ij} \hspace{5 mm} \text {and} \hspace{5 mm} R_i = \sum_j a_{ji} ,
 \een
-where $S_i$ and $R_i$ denote the number of sent and received messages, respectively.
+where $S_i$ and $R_i$ denote the number of messages sent and received, respectively.
 While $S_i$ can be computed trivially, a reduction operation is required to compute $R_i$.
@@
 This choice, however, is not suitable since the first few iterations do not contain many cells and there is little work for parallelism.
-Here we simply take the previous grid as the starting point, i.e.
-$G^{n+1}_0 = G^n$.
+Here we simply take the previous grid as the starting point, i.e.\ $G^{n+1}_0 = G^n$.
 Note that this iterative process is essentially unavoidable since the grid is based on the values of the level-set function at $t^{n+1}$, which itself is unknown and is to be defined on $G^{n+1}$.
@@
 \ee
-where $D^+_i \phi$ and $D^-_i \phi$ are the forward and backward derivatives in the $x_i$ direction and $\mathcal{H}_G$ is the Godunov Hamiltonian defined as:\
+where $D^+_i \phi$ and $D^-_i \phi$ are the forward and backward derivatives in the $x_i$ direction and $\mathcal{H}_G$ is the Godunov Hamiltonian defined as:
 \ben
@@
 \begin{array}{lcr}
-	\sqrt{\sum_i \max \left(|a^+_i|^2, |b^-_i|^2\right)} & \hspace {5 mm} \text{if} & S(\phi_0) \le 0 \\
+	\sqrt{\sum_i \max \left(|a^+_i|^2, |b^-_i|^2\right)} & \hspace {5 mm} \text{if} & S(\phi_0) \le 0 ,\\
 	\\
-	\sqrt{\sum_i \max\left(|a^-_i|^2, |b^+_i|^2\right)} & \hspace {5 mm} \text{if} & S(\phi_0)  >  0 
+	\sqrt{\sum_i \max\left(|a^-_i|^2, |b^+_i|^2\right)} & \hspace {5 mm} \text{if} & S(\phi_0)  >  0 ,
 \end{array}
 \right.
-,
 \een
@@
 First, we label all local points, $L_p$, as either private, $P_p$, or boundary, $B_p$.
-Here, the boundary points are the collection of all local points that are regarded as a ghost point, $G_r$, on at least one other process, i.e.
-$B_p = \underset{r,\;r\neq p}{\bigcup} G_r$.
-Private points are defined as the collection of all local points that are not a boundary point, i.e.
-$P_p = L_p \setminus B_p$.
+Here, the boundary points are the collection of all local points that are regarded as a ghost point, $G_r$, on at least one other process, i.e.\ $B_p = \underset{r,\;r\neq p}{\bigcup} G_r$.
+Private points are defined as the collection of all local points that are not a boundary point, i.e.\ $P_p = L_p \setminus B_p$.
 Algorithm \ref{alg:overlap} illustrates how this labeling can help with overlapping the computation and the communication associated to an arbitrary local operation $y \gets \mathcal{F}(x)$.

### 2015-06-19 paper/tex/applications.tex
Last edits for this time

@@
 The results are presented in figure \ref{fig:stefan_scaling}.
-As expected from the results obtained for each component in the previous sections, our implementation of the Stefan problem exhibits very satisfactory scaling (c.f.\ table \ref{tab:scaling_stefan}).
+As expected from the results obtained for each component in the previous sections, our implementation of the Stefan problem exhibits very satisfactory scaling (cf.\ Table \ref{tab:scaling_stefan}).
 

### 2015-06-19 paper/tex/conclusion.tex
Last edits for this time

@@
 This is because the domain decomposition technique used here is based on the Z-ordering of cells and does not take the velocity field information into account.
-A possible remedy for this problem could be assigning weights to cells based on some estimate of the grid structure after one step of the advection algorithm, e.g.
-by using a forward-in-time integration of grid points.
+A possible remedy for this problem could be assigning weights to cells based on some estimate of the grid structure after one step of the advection algorithm, e.g.\ by using a forward-in-time integration of grid points.
 Such an estimate could also reduce the number of semi-Lagrangian iterations.

### 2015-06-19 paper/tex/scaling_results.tex
Last edits for this time

@@
 The test consists of interpolating a function at a number of random points on a randomly refined Octree in three spatial dimensions.
-We consider two cases, a small test on a level\footnote{the level is the number of recursive splits allowed for each tree.} 9 tree with roughly 33M nodes and a larger test on a level 13 tree with roughly 280M nodes.
+We consider two cases, a small test on a level\footnote{The level is the number of recursive splits allowed for each tree.} 9 tree with roughly 33M nodes and a larger test on a level 13 tree with roughly 280M nodes.
 In both cases the number of randomly generated points is chosen to be equal to the number of nodes and the stabilized second-order interpolation of \cite{Min;Gibou:07:A-second-order-accur} is performed 10 times to smooth out possible timing fluctuations.
@@
 To simulate the effect of different CFL numbers, we generate the random points such that on each process $\alpha$ percentage of them are located outside the process boundary and thus will initiate communication.
-Scaling results are presented for $\alpha = 5 \%$ and $\alpha = 95\%$ for both the small and large problems in figure \ref{fig:interpolation}.
+Scaling results are presented for $\alpha = 5 \%$ and $\alpha = 95\%$ for both the small and large problems in Figure \ref{fig:interpolation}.
 We also present a third row of results for a much larger problem with roughly 1.66B nodes on a level 14 tree.
 Excellent scaling is obtained for the small problem for $P = 16-512$ even when $95\%$ of the interpolation points belong to a remote process.
-For the larger problem, however, the communication overhead prevents the algorithm from scaling beyond $2048$ processes when $\alpha = 95\%$ (c.f\ table \ref{tab:scaling_interpolation}).
+For the larger problem, however, the communication overhead prevents the algorithm from scaling beyond $2048$ processes when $\alpha = 95\%$ (cf.\ Table \ref{tab:scaling_interpolation}).
 Note, however, that this is expected since the total time is dominated by communication for $\alpha = 95\%$ and there is very little local work in this case.
-Indeed, the last row of figure \ref{fig:interpolation} show much better scaling behavior on a larger problem size, e.g.
+Indeed, the last row of Figure \ref{fig:interpolation} show much better scaling behavior on a larger problem size, e.g.
 efficiencies are increased from $e=34\%$ to $e=68\%$ for $\alpha = 95\%$ on 4096 processes.
@@
 	\caption{Strong scaling of Algorithm \ref{alg:interpolation} for several tests where $N_G$ denotes the number of random interpolation points (which is the same as the number of nodes in the Octree) and $\alpha$ denotes the percentage of these points that are remote for each process.
-Here ``Total'' represents the total time spent in the interpolation while ``buffer'', ``local'', ``queries'', and ``replies'' represent the the timing for different sections (c.f.
-Algorithm \ref{alg:interpolation}).
+Here ``Total'' represents the total time spent in the interpolation while ``buffer'', ``local'', ``queries'', and ``replies'' represent the the timing for different sections (cf.\ Algorithm \ref{alg:interpolation}).
 The black dashed line represents the ideal scaling.
@@
 Figures \ref{fig:semilagrangian_small} and \ref{fig:semilagrangian_large} illustrate the scalability of the algorithm for the small and large problems, respectively.
-To enable meaningful comparisons between different CFL numbers and number of processes, the maximum time has been scaled by the number of sub-iterations required for the grid construction as reported in table \ref{tab:semilagrangian}.
+To enable meaningful comparisons between different CFL numbers and number of processes, the maximum time has been scaled by the number of sub-iterations required for the grid construction as reported in Table \ref{tab:semilagrangian}.
 For both problems, excellent scalability is observed for $\text{CFL} = 1$ and $\text{CFL} = 10$.
@@
 The solid dashed line represents the ideal scaling.
-Note that the maximum time has been scaled by the number of sub-iterations required to build the tree (c.f.
-table \ref{tab:semilagrangian}).
+Note that the maximum time has been scaled by the number of sub-iterations required to build the tree (cf.\ Table \ref{tab:semilagrangian}).
 Here \texttt{p4est\_nodes\_new}, \texttt{p4est\_partition} and \texttt{p4est\_refine\_coarsen} refer to constructing the global indexing for nodes, partitioning the forest, and the refining/coarsening operation, respectively \cite{Burstedde;Wilcox;Ghattas:11:p4est:-Scalable-Algo}.}
@@
 The solid dashed line represents the ideal scaling.
-Note that the maximum time has been scaled by the number of sub-iterations required to build the tree (c.f.
-table \ref{tab:semilagrangian}).}
+Note that the maximum time has been scaled by the number of sub-iterations required to build the tree (cf.\ Table \ref{tab:semilagrangian}).}
 	\label{fig:semilagrangian_large}
@@
 To better understand the importance of the CFL number on the scalability, we have recorded a complete history of the communication pattern in the interpolation step.
-Figure \ref{fig:communication_4096} illustrates the effects of the CFL number on different metrics, namely the number of interpolation points\footnote{Note that this includes both the local points and the points queried by other processes}, $N_p$, the number of sent and received messages, $N_m = S + R$, and the total communication volume, $V_m$ in megabytes (MB), for $p=4096$ processes.
+Figure \ref{fig:communication_4096} illustrates the effects of the CFL number on different metrics, namely the number of interpolation points\footnote{Note that this includes both the local points and the points queried by other processes.}, $N_p$, the number of sent and received messages, $N_m = S + R$, and the total communication volume, $V_m$ in megabytes (MB), for $p=4096$ processes.
 Furthermore, these values are reported for the first (top row) and last (bottom row) sub-iterations of the semi-Lagrangian algorithm.
 There are several points to make.
-First, increasing the CFL number greatly increases the load imbalance, as shown by the spread of the data in figure \ref{fig:communication_4096_pf}.
+First, increasing the CFL number greatly increases the load imbalance, as shown by the spread of the data in Figure \ref{fig:communication_4096_pf}.
 This is because at higher CFL numbers, it is more likely that some processes will receive a larger portion of the backtracked points.
-Second, increasing the CFL number increases both the communication volume and its spread across processes (c.f.
-figure \ref{fig:communication_4096_vf}).
-Interestingly, however, the number of sent and received messages do not seem to affected by the CFL number.
-The bottom row of figure \ref{fig:communication_4096} exhibits a better balance both in the computation and communication volume in the last sub-iteration of the semi-Lagrangian algorithm.
+Second, increasing the CFL number increases both the communication volume and its spread across processes (cf.\ Figure \ref{fig:communication_4096_vf}).
+Interestingly, however, the number of sent and received messages do not seem to be affected by the CFL number.
+The bottom row of Figure \ref{fig:communication_4096} exhibits a better balance both in the computation and communication volume in the last sub-iteration of the semi-Lagrangian algorithm.
 This can be justified by noting that for the final sub-iteration, the partitioning of $G^{n+1}$ is more consistent with the partitioning of the departure points on $G^n$.
-Detailed information about the load balancing and the communication patterns is listed in table \ref{tab:communication}.
+Detailed information about the load balancing and the communication patterns is listed in Table \ref{tab:communication}.
 \begin{figure}[htbp]
@@
 	\caption{Performance indicators of the first (top row) and last (bottom row) sub-iterations of the semi-Lagrangian algorithm for the level-set advection on 4096 processes with $\text{CFL} = 1$ (\drawdiamond{black}), $\text{CFL} = 10$ (\drawsquare{red}), and $\text{CFL} = 100$ (\drawcircle{blue}).
-Increasing the CFL number greatly causes load imbalance during interpolation (a) and increases the communication volume (c).
+Increasing the CFL number causes load imbalance during interpolation (a) and increases the communication volume (c).
 However, the CFL number does not seem to appreciably affect the number of messages sent by the processes (b).
-During the last semi-Lagrangian sub-iteration, the initial grid $G_0$ (c.f.
-Algorithm \ref{alg:semi-lagrangian}) is very close to the final grid.
+During the last semi-Lagrangian sub-iteration, the initial grid $G_0$ (cf.\ Algorithm \ref{alg:semi-lagrangian}) is very close to the final grid.
 As a result, the load imbalance is considerably improved (d).
@@
 In both cases the forest is built on a $3\times3\times3$ macro-mesh.
-Figure \ref{fig:reinit} illustrates that our reinitialization algorithm, and in particular the overlapping strategy presented in Algorithm \ref{alg:overlap}, scales very well (c.f.\ table \ref{tab:scaling_reinit}).
+Figure \ref{fig:reinit} illustrates that our reinitialization algorithm, and in particular the overlapping strategy presented in Algorithm \ref{alg:overlap}, scales very well (cf.\ Table \ref{tab:scaling_reinit}).
 In general we expect similar scaling results for any local, finite-difference based calculations on Octrees that can efficiently utilize Algorithm \ref{alg:overlap}.
@@
 The black dashed line represents ideal scaling.
-Excellent results are obtained in both cases, illustrating the scalability of the overlapping strategy (c.f.
-algorithm \ref{alg:overlap}).}
+Excellent results are obtained in both cases, illustrating the scalability of the overlapping strategy (cf.\ Algorithm \ref{alg:overlap}).}
 \label{fig:reinit}

### 2015-06-19 paper/tex/introduction.tex
Small edits

@@
 In fact, the algorithms presented in this paper are implemented on top of
-the \texttt{p4est} API using a single octree to represent the unit cube.
+the \texttt{p4est} API.
+Due to the need for multiple adaptation and partitioning operations in each
+time step, the semi-Lagrangian method we describe below presents a stringent
+test of the algorithms and implementation both in terms of scalability and
+absolute run time.
+% using a single octree to represent the unit cube.
 % In fact, the algorithms presented in this paper are implemented on top of the
@@
 % \cite{Burstedde;Wilcox;Ghattas:11:p4est:-Scalable-Algo}.
-We also make use of the popular \texttt{PETSc}
+
+In implementing the numerical solver, we make use of the popular \texttt{PETSc}
 \cite{Balay;Abhyankar;Adams;etal:14:PETSc-Web-page} library for linear algebra

### 2015-06-23 paper/tex/applications.tex
Carsten's comments addressed.

@@
 \end{equation}
-where $\Delta x_{min}$ is the size of the smallest cell of the forest. The general procedure to solve the Stefan problem is presented in Algorithm \ref{alg:stefan}.
+where $\Delta x_{min}$ is the size of the smallest cell of the forest.
+The general procedure to solve the Stefan problem is presented in Algorithm \ref{alg:stefan} and we refer the interested reader to \cite{Chen;Min;Gibou:09:A-numerical-scheme-f} for the details of implementation.
+In implementing the numerical solver, we make use of the popular \texttt{PETSc}
+\cite{Balay;Abhyankar;Adams;etal:14:PETSc-Web-page} library for linear algebra and its parallel primitives, such as parallel ghosted vector and scatter/gather operations, which simplifies the implementation.
 
@@
 The maximum tree resolution for the small test is 9, leading to approximately 7M grid points, and the maximum resolution for the large test is 11, corresponding to 105M grid points.
-The results are presented in figure \ref{fig:stefan_scaling}.
+The results are presented in figure \ref{fig:stefan_scaling}, where ``\verb|Solution_Extension|'' refers to extrapolation procedure (see Algorithm \ref{alg:stefan} step 5).
 As expected from the results obtained for each component in the previous sections, our implementation of the Stefan problem exhibits very satisfactory scaling (cf.\ Table \ref{tab:scaling_stefan}).

### 2015-06-23 paper/tex/parallel_algorithms.tex
Exemplarily use mathit in some places
Carsten's comments addressed.

@@
 \State $H \gets G.\texttt{macromesh()}$
-\For {$tr : G.\texttt{local\_trees()}$} \Comment{build hierarchy for local cells}
-	\For {$c : tr.\texttt{cells()}$}
-		\State $H.\texttt{update\_tree(} tr, c \texttt{)}$
+\For {$\mathit{tr} : G.\texttt{local\_trees()}$} \Comment{build hierarchy for local cells}
+	\For {$c : \mathit{tr}.\texttt{cells()}$}
+		\State $H.\texttt{update\_tree(} \mathit{tr}, c \texttt{)}$
 	\EndFor
@@
 \\
-\Function{$H.\texttt{update\_tree}$}{$tr, c$}   \Comment{recursive tree reconstruction}
-	\State $c_l \gets H.\texttt{root(}tr\texttt{)}$
+\Function{$H.\texttt{update\_tree}$}{$\mathit{tr}, c$}   \Comment{recursive tree reconstruction}
+	\State $c_l \gets H.\texttt{root(}\mathit{tr}\texttt{)}$
 	\While{$c_l.\texttt{level()} \not= c.\texttt{level()} $}
@@
 \begin{algorithmic}[1]
-\State $col \gets 0, buff \gets null$ \Comment{Phase I -- buffering}
+\State $col \gets 0, \mathit{buff} \gets \mathit{null}$ \Comment{Phase I -- buffering}
 \For {$\underline{\mathbf{p}} : \underline{\mathbf{X}}$} 
-	\State $[r, cell] \gets H.\texttt{search(}\underline{\mathbf{p}}\texttt{)}$ \Comment{search for the owner's rank and cell}
-	\If {$r = mpirank$}
-		\State $buff[r].\texttt{push\_back(}\underline{\mathbf{p}}, cell\texttt{)}$
+	\State $[r, \mathit{cell})] \gets H.\texttt{search(}\underline{\mathbf{p}}\texttt{)}$ \Comment{search for the owner's rank and cell}
+	\If {$r = \mathit{mpirank}$}
+        \State $\mathit{buff}[r].\texttt{push\_back(}\underline{\mathbf{p}}, cell\texttt{)}$
 	\Else
-		\State $buff[r].\texttt{push\_back(}\underline{\mathbf{p}}\texttt{)}$
-		\State $col[r] \gets 1$
+        \State $\mathit{buff}[r].\texttt{push\_back(}\underline{\mathbf{p}}\texttt{)}$
+        \State $\mathit{col}[r] \gets 1$
 	\EndIf
 \EndFor
-\For {$r: mpisize$} \Comment{Phase II -- initiate communication and compute number of messages}
-	\If {$col[r]$}
-		\State $\texttt{MP\_Isend(}r, buff[r]\texttt{)}$
+\For {$r:\mathit{mpisize}$} \Comment{Phase II -- initiate communication and compute number of messages}
+	\If {$\mathit{col}[r]$}
+        \State $\texttt{MP\_Isend(}r, \mathit{buff}[r]\texttt{)}$
 	\EndIf
@@
 \State  $S \gets \texttt{sum(}col\texttt{)}$ 
-\State  $R \gets \texttt{MPI\_Reduce\_scatter(}col,\texttt{MPI\_SUM)}$
-\State $done \gets false, it \gets buff[mpirank].\texttt{begin()}$
-\While {$!done$} \Comment{Phase III -- main loop}
-	\If {$it \not= buff[mpirank].\texttt{end()}$}
-		\State $values \gets \texttt{process\_local(}it\texttt{)}$ \Comment{process local interpolations}
-		\State $\texttt{++}it$
+\State  $R \gets \texttt{MPI\_Reduce\_scatter(}\mathit{col},\texttt{MPI\_SUM)}$
+\State $\mathit{done} \gets \mathit{false}, it \gets \mathit{buff}[mpirank].\texttt{begin()}$
+\While {$\mathit{!done}$} \Comment{Phase III -- main loop}
+	\If {$\mathit{it} \not= \mathit{buff[mpirank]}.\texttt{end()}$}
+		\State $\mathit{values} \gets \texttt{process\_local(}it\texttt{)}$ \Comment{process local interpolations}
+		\State $\texttt{++}\mathit{it}$
 	\EndIf
 	\If {$R > 0$} \Comment{process queries sent from remote processes}
-		\State $[msg, st] \gets \texttt{MPI\_Iprobe()}$
-		\If {$msg$}
+		\State $[\mathit{msg, st}] \gets \texttt{MPI\_Iprobe()}$
+		\If {$\mathit{msg}$}
 			% \State $values \gets \texttt{MPI\_Recv(}st.\texttt{MPI\_SOURCE)}$ \Comment{}
-			\State $val\_buff \gets \texttt{process\_queries(}st\texttt{)}$ \Comment{receive, search, and interpolate values}
-			\State $\texttt{MPI\_Isend(}st.\texttt{MPI\_SOURCE},val\_buff\texttt{)}$ \Comment{send back interpolated values}
+			\State $\mathit{val\_buff} \gets \texttt{process\_queries(}\mathit{st}\texttt{)}$ \Comment{receive, search, and interpolate values}
+			\State $\texttt{MPI\_Isend(}\mathit{st}.\texttt{MPI\_SOURCE},\mathit{val\_buff}\texttt{)}$ \Comment{send back interpolated values}
 			\State $R\texttt{--}$
@@
 	\If {$S > 0$} \Comment{process replies sent to our queries}
-		\State $[msg, st] \gets \texttt{MPI\_Iprobe()}$
-		\If {$msg$}
-			\State $values \gets \texttt{process\_replies(}st.\texttt{MPI\_SOURCE)}$ \Comment{receive remotely interpolated values}
+		\State $\mathit{[msg, st]} \gets \texttt{MPI\_Iprobe()}$
+		\If {$\mathit{msg}$}
+			\State $\mathit{values} \gets \texttt{process\_replies(}\mathit{st}.\texttt{MPI\_SOURCE)}$ \Comment{receive remotely interpolated values}
 			\State $S\texttt{--}$
@@
 	\EndIf
-	\State $done \gets S = 0 \: \And \: R = 0 \: \And \: it = buff[mpirank].\texttt{end()}$
+	\State $\mathit{done} \gets S = 0 \: \And \: R = 0 \: \And \: \mathit{it = buff[mpirank]}.\texttt{end()}$
 \EndWhile
-\State \Return $values$
+\State \Return $\mathit{values}$
 \end{algorithmic}
@@
 \State $G^{n+1}_0 \gets G^n$
-\While {$true$}
+\While {$\mathit{true}$}
 	\State $\underline{\mathbf{X}}_d  \gets \texttt{ComputeDeparturePoints(}G^{n+1}_0, \underline{\mathbf{u}}^n, \underline{\mathbf{u}}^{n-1}, \Delta t\texttt{)}$ \Comment{using equations \ref{eq:xstar} -- \ref{eq:vn_p_half}}
@@
 \EndFor
-\State $send\_req \gets \texttt{MPI\_Isend(}y_B\texttt{)}$ \Comment{II -- begin updating ghost values}
-\State $recv\_req \gets \texttt{MPI\_Irecv(}y_G\texttt{)}$
+\State $\mathit{send\_req} \gets \texttt{MPI\_Isend(}y_B\texttt{)}$ \Comment{II -- begin updating ghost values}
+\State $\mathit{recv\_req} \gets \texttt{MPI\_Irecv(}y_G\texttt{)}$
 \For {$i:P_p$} \Comment{III -- perform computation on private points}
@@
 \EndFor
-\State $\texttt{MPI\_Waitall(}send\_req, recv\_req\texttt{)}$ \Comment{IV -- wait for ghost update to finish}
+\State $\texttt{MPI\_Waitall(}\mathit{send\_req, recv\_req}\texttt{)}$ \Comment{IV -- wait for ghost update to finish}
 \State \Return $y$

### 2015-06-23 paper/tex/scaling_results.tex
Carsten's comments addressed.

@@
 efficiencies are increased from $e=34\%$ to $e=68\%$ for $\alpha = 95\%$ on 4096 processes.
-This is a typical result with strong scaling and simply implies that our algorithms are scalablable for sufficiently large problems.
+This is a typical result with strong scaling and simply implies that our algorithms are scalable for sufficiently large problems.
 \begin{figure}[htbp]
@@
 	\caption{Detailed load balancing and communication information for the advection test for $\text{CFL} = 1$, $\text{CFL} = 10$, and $\text{CFL} = 100$.
+Here $N_p$ is the number of interpolation points, $N_m = S + R$ is the number of sent ($S$) and received ($R$) messages, and $V_m$ is the total communication volume in megabytes (MB).
 Note how increasing the CFL number causes load imbalance and increases the communication volume while it does not affect the number of messages sent and received during a sub-iteration of the semi-Lagrangian step.}

### 2015-06-26 paper/tex/scaling_results.tex
Added my modifications - Ready for submission

@@
 
+Unlike many existing applications where the mesh is changed infrequently, our semi-Lagrangian algorithm requires several sub-iterations of the refinement and coarsening operations.
+As a result, it is expected that refinement and coarsening steps constitute a significant portion of the total runtime which puts stringent scalability requirements on these algorithms.
+We refer the interested reader to section 3.2 of \cite{Burstedde;Wilcox;Ghattas:11:p4est:-Scalable-Algo} for detailed description of scalable refinement and coarsening algorithms in \texttt{p4est}.
+
 Table \ref{tab:semilagrangian} illustrates the dependence of the number of sub-iterations required to build the grid on the CFL number; as the CFL is increased, the interface travels a farther distance, which necessitates more sub-iterations to generate the grid.
-Note that the apparent dependence on the number of processes is most likely due to round-off errors during the interpolation and/or backtracking steps.
 Figures \ref{fig:semilagrangian_small} and \ref{fig:semilagrangian_large} illustrate the scalability of the algorithm for the small and large problems, respectively.

### 2015-06-27 paper/tex/novelty.tex
Paper submitted

\documentclass[11pt]{article}
%\usepackage[pdftex,breaklinks,linktocpage,pagebackref,hyperindex,hyperfigures]
\usepackage{hyperref}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage[tmargin=1in,bmargin=1in,lmargin=1in,rmargin=1in]{geometry}
\usepackage{subfigure}
\usepackage{graphicx}
\usepackage{latexsym}
\usepackage{pdfsync}
% \usepackage[boxed]{algorithm}
\usepackage{algpseudocode}
\usepackage{algorithm}
% \usepackage{algorithmic}
\usepackage{multirow}
\usepackage{rotating}
\usepackage{color}
\usepackage{caption}
\usepackage{url}
\usepackage{tikz}
\usepackage{diagbox}

\begin{document}

\title{Parallel Level-Set Methods on Adaptive Tree-Based Grids}

\author{Mohammad Mirzadeh, Arthur Guittet, Carsten Burstedde, and Frederic Gibou}
\maketitle
\section*{Significance and Novely of the Paper}
In this article we present scalable parallel level-set algorithms on adaptive Quadtree and Octree grids which allow efficient advection of the level-set function by automatically confining the fine grid close to the interface. These algorithms are suitable for massively parallel applications of the level-set methods on machines with distributed memory architecture. The parallelism is based on the partitioning of the adaptive grid using the idea of Space Filling Curves (SFCs). For this purpose we use the open-source \texttt{p4est} library which implements a collection of scalable refinement, coarsening, and partitioning algorithms and has recently been scaled to more than 450,000 CPU cores.

To advect the level-set function on such distributed adaptive grids, we present a parallel semi-Lagrangian algorithm which, similar to its serial counterpart, is unconditionally stable and allows for very large time-steps. Parallelizing the semi-Lagrangian algorithm on distributed adaptive grids, however, proves to be a nontrivial task. This has to do with the inconsistencies between the distribution of departure points across the processes and the parallel partitioning of the tree, which results in an asymmetric communication pattern and can potentially lead to load imbalances. We present an scalable interpolation algorithm that addresses these issues.

We also present a parallel reinitialization algorithm based on the pseudo-time formulation. To ensure maximum scalability, this algorithm is presented in a way that overlaps local calculation with remote communication. Finally we present an application of these algorithms to the solidification phenomena by solving a Stefan problem. Our algorithms show good scalability up to 4096 CPU cores, which is the current limit of our research allocation on the Stampede supercomputer.

To the best of our knowledge, this is the first large-scale application of the distributed parallelism to the level-set problem on adaptive tree-based grids. We believe that our findings are of interest to a diverse group of researchers in computational physics. We also believe that there are still many interesting directions that could be pursued to further improve upon our algorithms and we hope this paper serves as a starting point for such investigations.
\end{document}

### 2015-12-18 examples/advection/main_2d.cpp
Cleaned out the examples quite a bit (see details)
1) poisson_cell based is working
2) the bialloy example throws an exception related the semilagrangian
method. @bousouf is this expected? Also, I ahem changed the absolute
path you have been using in your examples. Not sure what the best option
is here. you might wanna have a look at it.
3) Added an script to make creating new projects easier.
4) Most of the rest is just filename changes

@@ -32,8 +32,8 @@
 #include <src/my_p4est_level_set.h>
 #endif
 
-#include <src/petsc_compatibility.h>
 #include <src/Parser.h>
+#include <src/petsc_compatibility.h>
 
 #undef MIN
 #undef MAX
@@ -193,14 +193,14 @@ int main (int argc, char* argv[])
 #endif
   tf = cmd.get("tf", tf);
 
+  const int n_xyz [] = {nx, ny};
+  const double xyz_min [] = {xmin, ymin};
+  const double xyz_max [] = {xmax, ymax};
+
   // Create the connectivity object
   p4est_connectivity_t *connectivity;
   my_p4est_brick_t brick;
-#ifdef P4_TO_P8
-  connectivity = my_p4est_brick_new(nx, ny, nz, xmin, xmax, ymin, ymax, zmin, zmax, &brick);
-#else
-  connectivity = my_p4est_brick_new(nx, ny, xmin, xmax, ymin, ymax, &brick);
-#endif
+  connectivity = my_p4est_brick_new(n_xyz, xyz_min, xyz_max, &brick);
 
   double dxyz_min[P4EST_DIM];
   dxyz_min[0] = (xmax-xmin)/nx/(1<<lmax);

### 2015-12-18 examples/basics/main_3d.cpp

@@ -1,2 +1,3 @@
 #include <src/my_p4est_to_p8est.h>
-#include "main_p4est.cpp"
+#include "main_2d.cpp"
+

### 2015-12-18 examples/bialloy/main_3d.cpp

@@ -1,2 +1,2 @@
 #include <src/my_p4est_to_p8est.h>
-#include "main_vtk_2d.cpp"
+#include "main_2d.cpp"

### 2015-12-18 examples/create_example.sh

@@ -1,0 +1,151 @@
+#!/bin/sh
+# Use this script to create a sample example which creats a tree 
+# and saves it into a vtk file.
+# 
+# Usage:
+#   ./new_example example_name
+#
+
+# creates a .pro file 
+function create_pro() {
+cat > $1/$1.pro << EOF
+CONFIG += 2d
+
+# --------------------------------------------------------------
+# qmake config file for the poisson example
+# --------------------------------------------------------------
+include(../../qt_config/$2.pri)
+include(../../qt_config/libparcasl.pri)
+
+main = main
+CONFIG(2d, 2d|3d): {
+  target = \$\${TARGET}_2d
+  TARGET = \$\${target}
+  SOURCES += \$\${main}_2d.cpp
+}
+
+CONFIG(3d, 2d|3d): {
+  target = \$\${TARGET}_3d
+  TARGET = \$\${target}
+  SOURCES += \$\${main}_3d.cpp
+}
+
+EOF
+}
+
+# creates a sample main_2d.cpp and main_3d.cpp
+function create_main() {
+cat > $1/main_2d.cpp << EOF
+/* 
+ *
+ * Title: $1
+ * Description:
+ * Author: 
+ * Date Created: 
+ *
+*/
+
+#ifndef P4_TO_P8
+#include <src/my_p4est_utils.h>
+#include <src/my_p4est_vtk.h>
+#include <src/my_p4est_nodes.h>
+#include <src/my_p4est_tools.h>
+#include <src/my_p4est_refine_coarsen.h>
+#include <src/my_p4est_log_wrappers.h>
+#include <src/my_p4est_node_neighbors.h>
+#else
+#include <src/my_p8est_utils.h>
+#include <src/my_p8est_vtk.h>
+#include <src/my_p8est_nodes.h>
+#include <src/my_p8est_tools.h>
+#include <src/my_p8est_refine_coarsen.h>
+#include <src/my_p8est_log_wrappers.h>
+#include <src/my_p8est_node_neighbors.h>
+#endif
+
+#include <src/Parser.h>
+#include <src/CASL_math.h>
+
+using namespace std;
+
+int main(int argc, char** argv) {
+  
+  // prepare parallel enviroment
+  mpi_enviroment_t mpi;
+  mpi.init(argc, argv);
+
+  // stopwatch
+  parStopWatch w;
+  w.start("Running example: $1");
+
+  // p4est variables
+  p4est_t*              p4est;
+  p4est_nodes_t*        nodes;
[…]

### 2015-12-18 examples/deprecated/levelset/main_reinitialization.cpp

@@ -1,0 +1,193 @@
+// System
+#include <stdexcept>
+#include <iostream>
+#include <sys/stat.h>
+#include <vector>
+#include <algorithm>
+#include <fstream>
+
+// p4est Library
+#include <p4est_bits.h>
+#include <p4est_extended.h>
+#include <p4est_vtk.h>
+#include <p4est_mesh.h>
+
+// casl_p4est
+#include <src/utils.h>
+#include <src/my_p4est_vtk.h>
+#include <src/my_p4est_nodes.h>
+#include <src/my_p4est_tools.h>
+#include <src/refine_coarsen.h>
+#include <src/petsc_compatibility.h>
+#include <src/interpolating_function.h>
+
+#include <src/neighbors.h>
+
+#undef MIN
+#undef MAX
+#include <src/CASL_math.h>
+
+#include <src/my_p4est_hierarchy.h>
+#include <src/my_p4est_cell_neighbors.h>
+#include <src/my_p4est_node_neighbors.h>
+#include <src/my_p4est_reinitialize.h>
+
+using namespace std;
+
+struct circle:CF_2{
+  circle(double x0_, double y0_, double r_): x0(x0_), y0(y0_), r(r_) {}
+  void update (double x0_, double y0_, double r_) {x0 = x0_; y0 = y0_; r = r_; }
+  double operator()(double x, double y) const {
+    return r - sqrt(SQR(x-x0) + SQR(y-y0));
+  }
+private:
+  double x0, y0, r;
+};
+
+struct parabola:CF_2{
+  parabola(double x0_, double y0_, double r_): x0(x0_), y0(y0_), r(r_) {}
+  void update (double x0_, double y0_, double r_) {x0 = x0_; y0 = y0_; r = r_; }
+  double operator()(double x, double y) const {
+    return r*r - (SQR(x-x0) + SQR(y-y0));
+  }
+private:
+  double x0, y0, r;
+};
+
+
+struct splitting_criteria_custom_t {
+    int max_lvl, min_lvl;
+};
+
+
+p4est_bool_t
+refine_custom(p4est_t *p4est, p4est_topidx_t which_tree, p4est_quadrant_t *quad)
+{
+    splitting_criteria_custom_t *data = (splitting_criteria_custom_t*)p4est->user_pointer;
+
+    if (quad->level < data->min_lvl)
+        return P4EST_TRUE;
+    else if (quad->level >= data->max_lvl)
+        return P4EST_FALSE;
[…]

### 2015-12-18 examples/dummy/dummytest.hpp

@@ -1,2 +1,0 @@
-
-void dummyio (void);

### 2015-12-18 examples/interpolation/main_3d.cpp

@@ -1,2 +1,2 @@
 #include <src/my_p4est_to_p8est.h>
-#include "main_test_nodes_2d.cpp"
+#include "main_2d.cpp"

### 2015-12-18 examples/levelset/main_Godunov_3d.cpp

@@ -1,2 +1,2 @@
 #include <src/my_p4est_to_p8est.h>
-#include "main_bialloy_2d.cpp"
+#include "main_Godunov_2d.cpp"

### 2015-12-18 examples/navier_stokes/main_navier_stokes_scaling_3d.cpp

@@ -1,2 +1,0 @@
-#include <src/my_p4est_to_p8est.h>
-#include "main_navier_stokes_scaling_2d.cpp"

### 2015-12-18 examples/navier_stokes/main_scaling_3d.cpp

@@ -1,2 +1,2 @@
 #include <src/my_p4est_to_p8est.h>
-#include "main_p4est.cpp"
+#include "main_scaling_2d.cpp"

### 2015-12-18 src/my_p4est_bialloy.h

@@ -1,8 +1,8 @@
 #ifndef MY_P4EST_BIALLOY_H
 #define MY_P4EST_BIALLOY_H
 
-#include <src/casl_types.h>
 #include <src/CASL_math.h>
+#include <src/casl_types.h>
 
 #ifdef P4_TO_P8
 #include <src/my_p8est_tools.h>

### 2015-12-18 src/my_p4est_tools.c

@@ -34,15 +34,9 @@
 #include <assert.h>
 
 p4est_connectivity_t *
-#ifdef P4_TO_P8
-my_p4est_brick_new (int nxtrees, int nytrees, int nztrees,
-                    double xmin, double xmax, double ymin, double ymax, double zmin, double zmax,
+my_p4est_brick_new (const int* n_xyz,
+                    const double* xyz_min, const double* xyz_max,
                     my_p4est_brick_t * myb)
-#else
-my_p4est_brick_new (int nxtrees, int nytrees,
-                    double xmin, double xmax, double ymin, double ymax,
-                    my_p4est_brick_t * myb)
-#endif
 {
   int                 i, j;
   int                 ii, jj;
@@ -57,25 +51,25 @@ my_p4est_brick_new (int nxtrees, int nytrees,
   p4est_topidx_t      tt, vindex;
   p4est_connectivity_t *conn;
 
-  P4EST_ASSERT (0 < nxtrees && 0 < nytrees);
+  P4EST_ASSERT (0 < n_xyz[0] && 0 < n_xyz[1]);
 #ifdef P4_TO_P8
-  P4EST_ASSERT (0 < nztrees);
+  P4EST_ASSERT (0 < n_xyz[2]);
 #endif
 #ifdef P4_TO_P8
-  conn = p4est_connectivity_new_brick (nxtrees, nytrees, nztrees, 0, 0, 0);
+  conn = p4est_connectivity_new_brick (n_xyz[0], n_xyz[1], n_xyz[2], 0, 0, 0);
 #else
-  conn = p4est_connectivity_new_brick (nxtrees, nytrees, 0, 0);
+  conn = p4est_connectivity_new_brick (n_xyz[0], n_xyz[1], 0, 0);
 #endif
   vv = conn->vertices;
   P4EST_ASSERT (vv != NULL);
 
 #ifdef P4_TO_P8
-  myb->nxyztrees[0] = nxtrees;
-  myb->nxyztrees[1] = nytrees;
-  myb->nxyztrees[2] = nztrees;
+  myb->nxyztrees[0] = n_xyz[0];
+  myb->nxyztrees[1] = n_xyz[1];
+  myb->nxyztrees[2] = n_xyz[2];
 #else
[…]

### 2015-12-18 src/my_p4est_tools.h

@@ -46,21 +46,18 @@ typedef struct
 }
 my_p4est_brick_t;
 
-/** Create a brick connectivity and tree lookup structure.
- * \param [in] nxtrees  Number of trees in x dimension.
- * \param [in] nytrees  Number of trees in y dimension.
- * \param [in,out] myb  Additional brick information will be populated.
- * \return              The brick connectivity structure.
+/*!
+ * \brief my_p4est_brick_new creates a new connectivity structure based on the brick type
+ * \param n_xyz pointer to an array of type int[P4EST_DIM], setting the number of trees in each direction
+ * \param xyz_min pointer to an array of type double[P4EST_DIM], setting the coordinate of the lower, left, back point
+ * \param xyz_max pointer to an array of type double[P4EST_DIM], setting the coordinate of the upper, right, front point
+ * \param myb pointer to the brick data structure
+ * \return
  */
-#ifdef P4_TO_P8
-p4est_connectivity_t *my_p4est_brick_new (int nxtrees, int nytrees, int nztrees,
-                                          double xmin, double xmax, double ymin, double ymax, double zmin, double zmax,
+p4est_connectivity_t *my_p4est_brick_new (const int* n_xyz,
+                                          const double* xyz_min, const double* xyz_max,
                                           my_p4est_brick_t * myb);
-#else
-p4est_connectivity_t *my_p4est_brick_new (int nxtrees, int nytrees,
-                                          double xmin, double xmax, double ymin, double ymax,
-                                          my_p4est_brick_t * myb);
-#endif
+
 
 /** Free a brick connectivity and tree lookup structure.
  * \param [in] conn     The connectivity will be destroyed.

### 2015-12-18 examples/create_example.sh
Deleted a lot of unnecessary files and cleaned existing ones

@@ -14,8 +14,8 @@ CONFIG += 2d
 # --------------------------------------------------------------
 # qmake config file for the poisson example
 # --------------------------------------------------------------
-include(../../qt_config/$2.pri)
-include(../../qt_config/libparcasl.pri)
+include(../../config/$2.pri)
+include(../../config/libparcasl.pri)
 
 main = main
 CONFIG(2d, 2d|3d): {

### 2015-12-18 tests/lookup/main_test_lookup.cpp

@@ -1,367 +1,0 @@
-// p4est Library
-#include <p4est_bits.h>
-#include <p4est_extended.h>
-#include <p4est_vtk.h>
-
-// My files for this project
-
-// System
-#include <stdexcept>
-#include <iostream>
-#include <sys/stat.h>
-#include <vector>
-#include <list>
-
-// casl_p4est
-#include <src/utilities.h>
-#include <src/utils.h>
-#include <src/my_p4est_vtk.h>
-#include <src/my_p4est_nodes.h>
-#include <src/my_p4est_tools.h>
-#include <src/semi_lagrangian.h>
-#include <src/refine_coarsen.h>
-
-// function prototype
-void run_test_1(p4est_t *p4est, const my_p4est_brick_t *brick);
-void run_test_2(p4est_t *p4est, const my_p4est_brick_t *brick);
-void run_test_3(p4est_t *p4est, const my_p4est_brick_t *brick);
-
-struct Point{
-  double x,y;
-  p4est_locidx_t owner;
-};
-
-struct Cell{
-  p4est_quadrant_t *quad;
-  p4est_locidx_t idx;
-};
-
-using namespace std;
-int main (int argc, char* argv[]){
-
-  mpi_context_t       mpi_context, *mpi = &mpi_context;
-  mpi->mpicomm = MPI_COMM_WORLD;
-  p4est_connectivity_t *connectivity;
-  p4est_t              *p4est;
-
-  //    srand(time(0));
-  rand_grid_data_t data = {9, 1, 5000, 100};
-
-  Session::init(argc, argv, mpi->mpicomm);
-
-  parStopWatch w1, w2;
-  w1.start("total time");
-
-  MPI_Comm_size (mpi->mpicomm, &mpi->mpisize);
-  MPI_Comm_rank (mpi->mpicomm, &mpi->mpirank);
-
-  // Create the connectivity object
-  w2.start("connectivity");
-  my_p4est_brick_t brick;
-  connectivity = my_p4est_brick_new(2, 2, &brick);
-  w2.stop(); w2.read_duration();
-
-  // Now create the forest
-  w2.start("p4est generation");
-  p4est = p4est_new(mpi->mpicomm, connectivity, 0, NULL, NULL);
-  w2.stop(); w2.read_duration();
-  p4est->user_pointer = (void*)(&data);
-
-  w2.start("refinement/coarsening");
-  for (int i=0; i<data.max_lvl; i++){
[…]

### 2015-12-18 tests/lookup/main_test_lookup_parallel.cpp

@@ -1,320 +1,0 @@
-// p4est Library
-#include <p4est.h>
-#include <p4est_bits.h>
-#include <p4est_extended.h>
-#include <p4est_vtk.h>
-
-// System
-#include <stdexcept>
-#include <iostream>
-#include <sys/stat.h>
-#include <vector>
-#include <list>
-
-// casl_p4est
-#include <src/utilities.h>
-#include <src/utils.h>
-#include <src/my_p4est_vtk.h>
-#include <src/my_p4est_nodes.h>
-#include <src/my_p4est_tools.h>
-#include <src/semi_lagrangian.h>
-#include <src/refine_coarsen.h>
-
-#include <mpi/mpi.h>
-
-// function prototype
-void run_test(p4est_t *p4est, const my_p4est_brick_t *brick);
-
-struct Point{
-  double x,y;
-  p4est_locidx_t owner;
-};
-
-struct Cell{
-  p4est_quadrant_t *quad;
-  p4est_gloidx_t idx;
-};
-
-using namespace std;
-int main (int argc, char* argv[]){
-
-  mpi_context_t       mpi_context, *mpi = &mpi_context;
-  mpi->mpicomm = MPI_COMM_WORLD;
-  p4est_connectivity_t *connectivity;
-  p4est_t              *p4est;
-
-  //  srand(time(0));
-  rand_grid_data_t data = {9, 1, 5000, 100};
-
-  Session::init(argc, argv, mpi->mpicomm);
-
-  parStopWatch w1, w2;
-  w1.start("total time");
-
-  MPI_Comm_size (mpi->mpicomm, &mpi->mpisize);
-  MPI_Comm_rank (mpi->mpicomm, &mpi->mpirank);
-
-  // Create the connectivity object
-  w2.start("connectivity");
-  my_p4est_brick_t brick;
-  connectivity = my_p4est_brick_new(2, 2, &brick);
-  w2.stop(); w2.read_duration();
-
-  // Now create the forest
-  w2.start("p4est generation");
-  p4est = p4est_new(mpi->mpicomm, connectivity, 0, NULL, NULL);
-  w2.stop(); w2.read_duration();
-  p4est->user_pointer = (void*)(&data);
-
-  w2.start("refinement/coarsening");
-  for (int i=0; i<data.max_lvl; i++){
-    data.counter = 0;
-    p4est_refine (p4est, P4EST_FALSE, refine_random , NULL);
-  }
[…]

### 2015-12-18 examples/extrapolation/main_constant_extrapolation_from_interface_2d.cpp
extrapolation example works

@@ -16,7 +16,7 @@
 #include <src/my_p8est_refine_coarsen.h>
 #include <src/my_p8est_log_wrappers.h>
 #include <src/my_p8est_node_neighbors.h>
-#include <src/my_p8est_levelset.h>
+#include <src/my_p8est_level_set.h>
 #else
 #include <p4est_bits.h>
 #include <p4est_extended.h>
@@ -26,7 +26,7 @@
 #include <src/my_p4est_tools.h>
 #include <src/my_p4est_refine_coarsen.h>
 #include <src/my_p4est_node_neighbors.h>
-#include <src/my_p4est_levelset.h>
+#include <src/my_p4est_level_set.h>
 #include <src/my_p4est_refine_coarsen.h>
 #endif
 
@@ -43,17 +43,18 @@ int nb_splits = 0;
 #define LINEAR
 //#define CONSTANT
 
-int brick_nx = 2;
-int brick_ny = 2;
-double x_center = (double) brick_nx / 2.;
-double y_center = (double) brick_ny / 2.;
-double r = .512092;
+int n_xyz [] = {2, 2, 2};
+double xyz_min [] = {0, 0, 0};
+double xyz_max [] = {2, 2, 2};
+
+double x_center = (xyz_max[0]-xyz_min[0]) / 2.;
+double y_center = (xyz_max[1]-xyz_min[1]) / 2.;
 
 #ifdef P4_TO_P8
-int brick_nz = 2;
-double z_center = (double) brick_nz / 2.;
+double z_center = (xyz_max[2]-xyz_min[2]) / 2.;
 #endif
 
+double r = .512092;
 #undef MIN
 #undef MAX
 
@@ -151,11 +152,11 @@ void check_accuracy(p4est_t *p4est, p4est_nodes_t *nodes, Vec phi, Vec q, CF_2 *
 
       double tree_xmin = p4est->connectivity->vertices[3*v_mm + 0];
       double tree_ymin = p4est->connectivity->vertices[3*v_mm + 1];
-      double x = node_x_fr_i(node) + tree_xmin;
-      double y = node_y_fr_j(node) + tree_ymin;
+      double x = node_x_fr_n(node) + tree_xmin;
+      double y = node_y_fr_n(node) + tree_ymin;
 #ifdef P4_TO_P8
       double tree_zmin = p4est->connectivity->vertices[3*v_mm + 2];
-      double z = node_z_fr_k(node) + tree_zmin;
+      double z = node_z_fr_n(node) + tree_zmin;
 #endif
 
[…]

### 2015-12-18 examples/extrapolation/main_extension_2d.cpp

@@ -16,7 +16,7 @@
 #include <src/my_p8est_refine_coarsen.h>
 #include <src/my_p8est_log_wrappers.h>
 #include <src/my_p8est_node_neighbors.h>
-#include <src/my_p8est_levelset.h>
+#include <src/my_p8est_level_set.h>
 #else
 #include <p4est_bits.h>
 #include <p4est_extended.h>
@@ -26,7 +26,7 @@
 #include <src/my_p4est_tools.h>
 #include <src/my_p4est_refine_coarsen.h>
 #include <src/my_p4est_node_neighbors.h>
-#include <src/my_p4est_levelset.h>
+#include <src/my_p4est_level_set.h>
 #endif
 
 #include <src/petsc_compatibility.h>
@@ -52,15 +52,16 @@ int order = 2;
 //double y_center = .5;
 //double r = .25112;
 
-int brick_nx = 2;
-int brick_ny = 2;
-double x_center = (double) brick_nx / 2.;
-double y_center = (double) brick_ny / 2.;
+int n_xyz [] = {2, 2, 2};
+double xyz_min [] = {0, 0, 0};
+double xyz_max [] = {2, 2, 2};
+
+double x_center = (xyz_max[0]-xyz_min[0]) / 2.;
+double y_center = (xyz_max[1]-xyz_min[1]) / 2.;
 double r = .512092;
 
 #ifdef P4_TO_P8
-int brick_nz = 2;
-double z_center = (double) brick_nz / 2.;
+double z_center = (xyz_max[2]-xyz_min[2]) / 2.;
 #endif
 
 //int brick_nx = 3;
@@ -206,8 +207,9 @@ public:
 
 int main (int argc, char* argv[])
 {
-  mpi_context_t mpi_context, *mpi = &mpi_context;
-  mpi->mpicomm  = MPI_COMM_WORLD;
+  mpi_enviroment_t mpi;
+  mpi.init(argc, argv);
+
   p4est_t            *p4est;
   p4est_nodes_t      *nodes;
   PetscErrorCode ierr;
@@ -230,33 +232,23 @@ int main (int argc, char* argv[])
   circle circ;
   splitting_criteria_cf_t data(min_level+nb_splits, max_level+nb_splits, &circ, 1.2);
 
-  Session mpi_session;
-  mpi_session.init(argc, argv, mpi->mpicomm);
-
   parStopWatch w1;
   w1.start("total time");
-
-  MPI_Comm_size (mpi->mpicomm, &mpi->mpisize);
[…]

### 2015-12-18 examples/interpolation/main_2d.cpp
interpolation example works

@@ -115,12 +115,11 @@ int main (int argc, char* argv[]){
     w2.start("connectivity");
     p4est_connectivity_t *connectivity;
     my_p4est_brick_t my_brick, *brick = &my_brick;
-#ifdef P4_TO_P8
-    connectivity = my_p4est_brick_new(2, 2, 2,
-                                      0, 2, 0, 2, 0, 2, brick);
-#else
-    connectivity = my_p4est_brick_new(2, 2, 0, 2, 0, 2, brick);
-#endif
+    int n_xyz [] = {2, 2, 2};
+    double xyz_min [] = {0, 0, 0};
+    double xyz_max [] = {2, 2, 2};
+
+    connectivity = my_p4est_brick_new(n_xyz, xyz_min, xyz_max, brick);
     w2.stop(); w2.read_duration();
 
     // Now create the forest

### 2015-12-18 examples/levelset/main_Godunov_2d.cpp
levelest example working

@@ -13,9 +13,9 @@
 #include <src/my_p8est_nodes.h>
 #include <src/my_p8est_tools.h>
 #include <src/my_p8est_refine_coarsen.h>
-#include <src/my_p8est_interpolating_function.h>
+#include <src/my_p8est_interpolation_nodes.h>
 #include <src/my_p8est_semi_lagrangian.h>
-#include <src/my_p8est_levelset.h>
+#include <src/my_p8est_level_set.h>
 #include <src/my_p8est_log_wrappers.h>
 #else
 #include <p4est_bits.h>
@@ -25,9 +25,9 @@
 #include <src/my_p4est_nodes.h>
 #include <src/my_p4est_tools.h>
 #include <src/my_p4est_refine_coarsen.h>
-#include <src/my_p4est_interpolating_function.h>
+#include <src/my_p4est_interpolation_nodes.h>
 #include <src/my_p4est_semi_lagrangian.h>
-#include <src/my_p4est_levelset.h>
+#include <src/my_p4est_level_set.h>
 #include <src/my_p4est_log_wrappers.h>
 #endif
 
@@ -104,8 +104,9 @@ private:
 
 int main (int argc, char* argv[]){
 
-  mpi_context_t mpi_context, *mpi = &mpi_context;
-  mpi->mpicomm  = MPI_COMM_WORLD;
+  mpi_enviroment_t mpi;
+  mpi.init(argc, argv);
+
   p4est_t            *p4est;
   p4est_nodes_t      *nodes;
   p4est_ghost_t      *ghost;
@@ -123,29 +124,23 @@ int main (int argc, char* argv[]){
 #endif
   splitting_criteria_cf_t data(cmd.get("lmin", 0), cmd.get("lmax", 5), &sq, 1.3);
 
-  Session mpi_session;
-  mpi_session.init(argc, argv, mpi->mpicomm);
 
   parStopWatch w1, w2;
   w1.start("total time");
-
-  MPI_Comm_size (mpi->mpicomm, &mpi->mpisize);
-  MPI_Comm_rank (mpi->mpicomm, &mpi->mpirank);
 
   // Create the connectivity object
   w2.start("connectivity");
   p4est_connectivity_t *connectivity;
   my_p4est_brick_t brick;
-#ifdef P4_TO_P8
-  connectivity = my_p4est_brick_new(2, 2, 2, &brick);
-#else
[…]

### 2015-12-18 examples/levelset/main_reinitialization_2d.cpp

@@ -6,133 +6,106 @@
 #include <algorithm>
 #include <fstream>
 
-// p4est Library
-#include <p4est_bits.h>
-#include <p4est_extended.h>
-#include <p4est_vtk.h>
-#include <p4est_mesh.h>
-
 // casl_p4est
-#include <src/utils.h>
+#ifdef P4_TO_P8
+#include <src/my_p8est_utils.h>
+#include <src/my_p8est_vtk.h>
+#include <src/my_p8est_nodes.h>
+#include <src/my_p8est_tools.h>
+#include <src/my_p8est_refine_coarsen.h>
+#include <src/my_p8est_node_neighbors.h>
+#include <src/my_p8est_hierarchy.h>
+#include <src/my_p8est_level_set.h>
+#else
+#include <src/my_p4est_utils.h>
 #include <src/my_p4est_vtk.h>
 #include <src/my_p4est_nodes.h>
 #include <src/my_p4est_tools.h>
-#include <src/refine_coarsen.h>
-#include <src/petsc_compatibility.h>
-#include <src/interpolating_function.h>
-
-#include <src/neighbors.h>
-
-#undef MIN
-#undef MAX
-#include <src/CASL_math.h>
-
-#include <src/my_p4est_hierarchy.h>
-#include <src/my_p4est_cell_neighbors.h>
+#include <src/my_p4est_refine_coarsen.h>
 #include <src/my_p4est_node_neighbors.h>
-#include <src/my_p4est_reinitialize.h>
+#include <src/my_p4est_hierarchy.h>
+#include <src/my_p4est_level_set.h>
+#endif
+
+#include <src/petsc_compatibility.h>
+#include <src/Parser.h>
+#include <src/CASL_math.h>
 
 using namespace std;
 
+#ifdef P4_TO_P8
+struct circle:CF_3{
+  circle(double x0_, double y0_, double z0_, double r_): x0(x0_), y0(y0_), z0(z0_), r(r_) {}
+  double operator()(double x, double y, double z) const {
+    return r - sqrt(SQR(x-x0) + SQR(y-y0) + SQR(z-z0));
+  }
+private:
+  double x0, y0, z0, r;
+};
+#else
 struct circle:CF_2{
   circle(double x0_, double y0_, double r_): x0(x0_), y0(y0_), r(r_) {}
-  void update (double x0_, double y0_, double r_) {x0 = x0_; y0 = y0_; r = r_; }
   double operator()(double x, double y) const {
[…]

### 2015-12-18 src/Parser.cpp

@@ -51,6 +51,8 @@ void cmdParser::parse(int argc, char* argv[])
 
     exit(EXIT_SUCCESS);
   }
+
+  print();
 }
 
 void cmdParser::print(FILE *f){

### 2016-02-12 examples/create_example.sh
fixed linux.pri and create_example script

@@ -1,4 +1,5 @@
-#!/bin/sh
+#!/usr/bin/env bash
+
 # Use this script to create a sample example which creats a tree 
 # and saves it into a vtk file.
 # 

### 2016-02-12 paper/tex/parallel_algorithms.tex
added accuracy section with enright test

@@
 We solve this problem using a simple \textit{communication matrix} (see Figure \ref{fig:communication}).
-Other approaches for solving this problem could include using non-blocking collective or one-sided communication operations introduced in the MPI-3 standard \cite{Hoefler;Siebert;Lumsdaine:10:Scalable-communication}.
-Although these algorithms have better theoretical communication complexities, we did not see any difference in the runtime or scalability of algorithm \ref{alg:interpolation} when using them.
+Our approach is very similar to the ``Personalized Census ($\mathcal{PCX}$)'' algorithm described in \cite{Hoefler;Siebert;Lumsdaine:10:Scalable-communication}.
+Another similar approach is the ``Notify'' algorithm introduced in \cite{Isaac;Burstedde;Ghattas:12:Low-cost-parallel-al}.
+Furthermore, the MPI-3 standard introduces non-blocking collectives and Remote Memory Access (RMA) operations which enable new ways of solving the communication problem.
+For instance, authors in \cite{Hoefler;Siebert;Lumsdaine:10:Scalable-communication} descried the ``Non-blocking Consensus ($\mathcal{NBX}$)'' and ``Remote Summation ($\mathcal{RSX}$)'' algorithms which make use of such operations and have better theoretical communication complexities.
+With the exception of $\mathcal{RSX}$ algorithm, which was not tested in this study, all remaining algorithms produced similar timing and scaling.
+Thus we have decided to describe our algorithm based on the idea of the \textit{communication matrix}.
 
@@
 \end{algorithm}
+
+\subsection{Accuracy} \label{sec:accuracy}
+
+The numerical methods detailed in the previous sections are widely used in the level-set community and their accuracy is studied for example in \cite{Min;Gibou:07:A-second-order-accur}.
+However, in order to validate our implementation, we present a brief convergence analysis.
+
+The advection of an irregular boundary using the semi-Lagrangian method is the perfect candidate to demonstrate the accuracy of our implementation as it makes use of the interpolation routine as well as the reinitialization procedure.
+We select the benchmark problem proposed in \cite{Enright;Fedkiw;Ferziger;etal:02:A-Hybrid-Particle-Le}.
+Consider a sphere centered at $(0.35,0.35,0.35)$ and with radius $0.15$ in a domain $[0,1]^3$ and deformed under the divergence free velocity field
+\be \label{eq:velo_enright}
+\begin{cases}
+u(x,y,z) = 2 \sin^2(\pi x)\sin(2 \pi y)\sin(2 \pi z), \\
+v(x,y,z) = - \sin^2(\pi y)\sin(2 \pi x)\sin(2 \pi z), \\
+w(x,y,z) = - \sin^2(\pi z)\sin(2 \pi x)\sin(2 \pi y).
+\end{cases}
+\ee
+forward in time until $t_{1/2}=1$ and then backward to its original state at $t_f=2$.
+With set the time step to $\Delta t = 5 \Delta x_{min}$, where $\Delta x_{min}$ is the size of the smallest cell in the forest.
+We monitor the volume loss and the error in the interface location at the final time, when the original shape should be recovered, as the finest resolution of the forest increases.
+The results are reported in table \ref{tab:accuracy} and figure \ref{fig:accuracy} shows a visualization of the sphere at the initial time, at $t_{1/2}$ when the deformation is maximal, and at the final time.
+The results are consistent with those reported previously in \cite{Min;Gibou:07:A-second-order-accur} and indicate second order accuracy for the mass loss.
+
+\begin{table}[htbp]
+\begin{center}
+\begin{tabular}{|c|c|c|c|c|}
+\hline
+Finest resolution & $L^\infty$ error on $\phi$ & Rate & Volume loss (\%) & Rate \\
+\hline
+$128^3$  & 1.53E-01 & -    & 1.85E-01 & -    \\
+$256^3$  & 1.18E-01 & 0.38 & 3.73E-02 & 2.31 \\
+$512^3$  & 9.82E-03 & 3.58 & 6.87E-03 & 2.44 \\
+$1024^3$ & 2.56E-03 & 1.94 & 1.92E-03 & 1.84 \\
+$2048^3$ & 1.60E-03 & 0.67 & 5.17E-04 & 1.89 \\
+\hline
+\end{tabular}
+\caption{Study of the convergence of the level-set algorithm using Enright's test \cite{Enright;Fedkiw;Ferziger;etal:02:A-Hybrid-Particle-Le}.
+The $L^\infty$ error on $\phi$ is computed only close to the interface as this is the relevant observable for an advection procedure.} \label{tab:accuracy}
+\end{center}
+\end{table}
+
+\begin{figure}[htbp]
+\begin{center}
+\begin{tabular}{ccccc}
+\includegraphics[width=.28\textwidth]{figures/enright_0.png}
+&
+\hspace{.1cm}
+&
+\includegraphics[width=.28\textwidth]{figures/enright_1.png}
+&
+\hspace{.1cm}
+&
+\includegraphics[width=.28\textwidth]{figures/enright_2.png}
+\end{tabular}
+\caption{Visualization of the deformation undergone by the sphere for the Enright's test at time 0, $t_{1/2}$ and $t_f$ and colored with the processes ranks.
+The forest's finest resolution is equivalent to a uniform grid with $2048^3$ cells and contains between 17 (at initial and final times) and 63 (at $t_{1/2}$) million nodes.
+These results were obtained on the Comet supercomputer with 960 processes (40 compute nodes).} \label{fig:accuracy}
+\end{center}
+\end{figure}

### 2016-02-12 paper/tex/scaling_results.tex
added accuracy section with enright test

@@
 \subsection{Semi-Lagrangian}
-To test the scalability of the semi-Lagrangian scheme of Algorithm \ref{alg:semi-lagrangian}, we consider a slightly modified version of the Enright's rotation test \cite{Enright;Fedkiw;Ferziger;etal:02:A-Hybrid-Particle-Le}, i.e.
-we advect a sphere of radius 0.35 located at $(0.4, 0.4, 0.4)$ with a divergence free velocity field given by:
-\bean
-u(x,y,z) &=& 2\sin(\pi x)^2 \sin(2\pi y) \sin(2 \pi z), \\
-v(x,y,z) &=& -\sin(\pi y)^2 \sin(2\pi x) \sin(2 \pi z),\\
-w(x,y,z) &=& -\sin(\pi z)^2 \sin(2\pi x) \sin(2 \pi y). 
-\eean
+To test the scalability of the semi-Lagrangian scheme of Algorithm \ref{alg:semi-lagrangian}, we consider a slightly modified version of the Enright's rotation test \cite{Enright;Fedkiw;Ferziger;etal:02:A-Hybrid-Particle-Le} presented in section \ref{sec:accuracy}, i.e.
+we advect a sphere of radius 0.35 located at $(0.4, 0.4, 0.4)$ with a divergence free velocity field given by equation \eqref{eq:velo_enright}.
+
 To understand the effect of the CFL number on the scalability of the algorithm we perform one step of the semi-Lagrangian algorithm for $\text{CFL} = 1$, $\text{CFL} = 10$, and $\text{CFL} = 100$.

### 2016-02-12 src/my_p4est_refine_coarsen.h
Fixed compilation issues on server

@@ -15,6 +15,7 @@
 
 #include <set>
 #include <vector>
+#include <stdexcept>
 
 #define SKIP_QUADRANT		 0
 #define REFINE_QUADRANT  1

### 2016-02-13 examples/viscous_fingering/one_fluid_solver_2d.cpp
added potential solver to viscous fingering

@@ -23,10 +23,18 @@ one_fluid_solver_t::one_fluid_solver_t(p4est_t* &p4est, p4est_ghost_t* &ghost, p
 
 one_fluid_solver_t::~one_fluid_solver_t() {}
 
-void one_fluid_solver_t::set_properties(cf_t &K_D, cf_t &gamma)
-{
-  this->K_D = &K_D;
+void one_fluid_solver_t::set_properties(cf_t &K_D, cf_t &K_EO, cf_t &gamma)
+{  
+  this->K_D   = &K_D;
+  this->K_EO  = &K_EO;
   this->gamma = &gamma;
+}
+
+void one_fluid_solver_t::set_injection_rates(cf_t &Q, cf_t &I, double alpha)
+{
+  this->Q     = &Q;
+  this->I     = &I;
+  this->alpha = alpha;
 }
 
 void one_fluid_solver_t::set_bc_wall(bc_wall_t &bc_wall_type, cf_t &bc_wall_value)
@@ -35,7 +43,7 @@ void one_fluid_solver_t::set_bc_wall(bc_wall_t &bc_wall_type, cf_t &bc_wall_valu
   this->bc_wall_value = &bc_wall_value;
 }
 
-double one_fluid_solver_t::advect_interface_semi_lagrangian(Vec &phi, Vec &pressure, double cfl, double dtmax)
+double one_fluid_solver_t::advect_interface_semi_lagrangian(Vec &phi, Vec &pressure, Vec& potential, double cfl, double dtmax)
 {
   // compute neighborhood information
   my_p4est_hierarchy_t hierarchy(p4est, ghost, brick);
@@ -44,12 +52,13 @@ double one_fluid_solver_t::advect_interface_semi_lagrangian(Vec &phi, Vec &press
 
   // compute interface velocity
   Vec vx_tmp[P4EST_DIM];
-  double *vx_p[P4EST_DIM], *pressure_p, *phi_p;
+  double *vx_p[P4EST_DIM], *pressure_p, *phi_p, *potential_p;
   foreach_dimension(dim) {
     VecCreateGhostNodes(p4est, nodes, &vx_tmp[dim]);
     VecGetArray(vx_tmp[dim], &vx_p[dim]);
   }
   VecGetArray(pressure, &pressure_p);
+  VecGetArray(potential, &potential_p);
   VecGetArray(phi, &phi_p);
 
   // compute on the layer nodes
[…]

### 2016-02-13 examples/viscous_fingering/one_fluid_solver_2d.h

@@ -24,20 +24,24 @@ class one_fluid_solver_t
   typedef CF_2 cf_t;
   typedef WallBC2D bc_wall_t;
 #endif
-  cf_t *K_D, *gamma, *bc_wall_value;
+  double alpha;
+  cf_t *Q, *I;
+  cf_t *K_D, *K_EO, *gamma;
+  cf_t *bc_wall_value;
   bc_wall_t *bc_wall_type;
 
-  double advect_interface_semi_lagrangian(Vec& phi, Vec& pressure, double cfl, double dtmax);
-  double advect_interface_godunov(Vec& phi, Vec& pressure, double cfl, double dtmax);
-  void solve_pressure(double t, Vec phi, Vec pressure);
+  double advect_interface_semi_lagrangian(Vec& phi, Vec& pressure, Vec &potential, double cfl, double dtmax);
+  double advect_interface_godunov(Vec& phi, Vec& pressure, Vec& potential, double cfl, double dtmax);
+  void solve_fields(double t, Vec phi, Vec pressure, Vec potential);
 
 public:
   one_fluid_solver_t(p4est_t* &p4est, p4est_ghost_t* &ghost, p4est_nodes_t* &nodes, my_p4est_brick_t& brick);
   ~one_fluid_solver_t();
 
-  void set_properties(cf_t &K_D, cf_t &gamma);
+  void set_properties(cf_t &K_D, cf_t &K_EO, cf_t &gamma);
+  void set_injection_rates(cf_t& Q, cf_t& I, double alpha);
   void set_bc_wall(bc_wall_t& bc_wall_type, cf_t& bc_wall_value);
-  double solve_one_step(double t, Vec &phi, Vec &pressure, const std::string& method = "semi_lagrangian", double cfl = 5.0, double dtmax = DBL_MAX);
+  double solve_one_step(double t, Vec &phi, Vec &pressure, Vec &potential, const std::string& method = "semi_lagrangian", double cfl = 5.0, double dtmax = DBL_MAX);
 };
 
 #endif // ONE_FLUID_SOLVER_2D_H

### 2016-02-14 paper/peer-reviews/response_letter.tex
Preparing response letter

\documentclass{scrlttr2}

\begin{document}
\begin{letter} {Editorial Office, \\ Journal of Computational Physics}
\setkomavar{subject}{Re: JCOMP-D-15-00846}
\opening{Dear Editor,}
We would like to appreciate the invaluable comments made by the reviewers. We have made adjustments to our article, titled `\textbf{Parallel level-set methods on adaptive tree-based grids}', to address the issues raised by the reviewers as detailed below.

\begin{itemize}
\item \textbf{Reviewer \#1} \\
The first reviewer has suggested to include ``\textit{some basic library codes in appendix, each matching its level set algorithm}''. Indeed it is important for the scientific community to share knowledge and facilitate ways for reproducing scientific work for others. In the case of computational science, unfortunately, this is a very daunting task that is not always possible. Many scientific codes are highly specialized and are often too technical to be fitted into a paper.

Specifically, in our case, our library amounts to nearly 30,000 lines of code, spread across more than 50 files and with complex dependencies. As such, it is practically not viable to include sample codes in the paper. Moreover, we believe inclusion of actual code, which involves many technical details, e.g. error checking, profiling tools, comments, etc., is not scientifically educational and may further alienate the reader from main message of the paper.

Instead, and in accordance to the second reviewer who has raised related issues, we have opted to revise the pseudo-codes in the algorithms to make them more clear. We believe that the modifications should make the algorithms easier to understand and useful to the readers.

\item \textbf{Reviewer \#2} \\
The second reviewer has raised several issues, mostly related to the pseudo-codes in the paper, as well as suggesting additional numerical tests. These issues are addressed as detailed below:
\begin{itemize}
\item ``\textit{It is recommended that the authors' pseudocode be closer to a language that many people are familiar with like C++ or Fortran 90}.''

Our library is coded in `\texttt{c++}' (with a mix of little `\texttt{c}' code). We have made adjustments to the to algorithms to make them easier to understand. However, we believe it is best not to directly use a `\texttt{c++}' syntax to keep the pseudo-code language-agnostic and accessible to all readers.

\item ``\textit{In Algorithm 2, what is the definition of “rank”? “owner’s rank”? “mpirank”? “st”?}''

\item ``\textit{}''

\item ``\textit{}''

\item ``\textit{}''
\end{itemize}
\end{itemize}

In light of these adjustments, we respectfully ask that our article be reconsidered for publication in the Journal of Computational Physics.
\closing{Sincerely,}
Mohammad Mirzadeh (on behalf of all authors)
\end{letter}
\end{document}

### 2016-02-15 paper/peer-reviews/response_letter.tex
reformatted tex file
Made changes to paper and prepared the response letter

@@
 \documentclass{scrlttr2}
+\usepackage{color}
 
 \begin{document} 
-\begin{letter} {Editorial Office, \\ Journal of Computational Physics}
+\begin{letter} {Editorial Office, \\ Journal of Computational
+Physics} 
 \setkomavar{subject}{Re: JCOMP-D-15-00846}
+
 \opening{Dear Editor,} 
-We would like to appreciate the invaluable comments made by the reviewers.
-We have made adjustments to our article, titled `\textbf{Parallel level-set methods on adaptive tree-based grids}', to address the issues raised by the reviewers as detailed below.
+We would like to appreciate the invaluable comments made by the reviewers. We
+have made adjustments to our article, titled `\textbf{Parallel level-set methods
+on adaptive tree-based grids}', to address the issues raised by the reviewers as
+detailed below.
 
@@
 \item \textbf{Reviewer \#1} \\ 
-The first reviewer has suggested to include ``\textit{some basic library codes in appendix, each matching its level set algorithm}''.
-Indeed it is important for the scientific community to share knowledge and facilitate ways for reproducing scientific work for others.
-In the case of computational science, unfortunately, this is a very daunting task that is not always possible.
-Many scientific codes are highly specialized and are often too technical to be fitted into a paper.
+The first reviewer has suggested to include ``\textit{some basic library codes
+in appendix, each matching its level set algorithm}''. Indeed it is important
+for the scientific community to share knowledge and facilitate ways for
+reproducing scientific work for others. In the case of computational science,
+unfortunately, this is a very daunting task that is not always possible.  Many
+scientific codes are highly specialized and are often too technical to be fitted
+into a paper.
 
-Specifically, in our case, our library amounts to nearly 30,000 lines of code, spread across more than 50 files and with complex dependencies.
-As such, it is practically not viable to include sample codes in the paper.
-Moreover, we believe inclusion of actual code, which involves many technical details, e.g.
-error checking, profiling tools, comments, etc., is not scientifically educational and may further alienate the reader from main message of the paper.
+Specifically, in our case, our library amounts to nearly 30,000 lines of code,
+spread across more than 50 files and with complex dependencies. As such, it is
+practically not viable to include sample codes in the paper. Moreover, we
+believe inclusion of actual code, which involves many technical details, e.g.
+error checking, profiling tools, comments, etc., is not scientifically
+educational and may further alienate the reader from main message of the paper.
 
-Instead, and in accordance to the second reviewer who has raised related issues, we have opted to revise the pseudo-codes in the algorithms to make them more clear.
-We believe that the modifications should make the algorithms easier to understand and useful to the readers.
+Instead, and in accordance to the second reviewer who has raised related issues,
+we have opted to revise the pseudo-codes in the algorithms to make them more
+clear. We believe that the modifications should make the algorithms easier to
+understand and useful to the readers.
 
 \item \textbf{Reviewer \#2} \\ 
-The second reviewer has raised several issues, mostly related to the pseudo-codes in the paper, as well as suggesting additional numerical tests. These issues are addressed as detailed below:
+The second reviewer has raised several issues, mostly related to the
+pseudo-codes in the paper, as well as suggesting additional numerical tests.
+These issues are addressed as detailed below:
+
 \begin{itemize} 
-\item ``\textit{It is recommended that the authors' pseudocode be closer to a language that many people are familiar with like C++ or Fortran 90}.''
+\item ``\textit{It took me a while to understand Algorithms 1 and 2. It is
+  recom- mended that the authors’ pseudocode be closer to a language that many
+people are familiar with like C++ or Fortran 90 (or perhaps, the authors’ are
+using a new language that this referee is not aware of?).}''
 
@@
 We have made adjustments to the to algorithms to make them easier to understand.
-However, we believe it is best not to directly use a `\texttt{c++}' syntax to keep the pseudo-code language-agnostic and accessible to all readers.
+However, we believe it is best not to directly use a `\texttt{c++}' syntax to
+keep the pseudo-code language-agnostic and accessible to all readers.
+\\
+\item ``\textit{In Algorithm 2, what is the definition of “rank”? “owner’s
+  rank”? “mpirank”? “st”?}''
 
-\item ``\textit{In Algorithm 2, what is the definition of “rank”? “owner’s rank”? “mpirank”? “st”?}''
+There were multiple references to the ``\textit{rank}'' in the pseudo-code that 
+would have caused confusion. We have made things clear by explicitly referring 
+to ``\textit{owners\_rank}'' and ``\textit{mpirank}''. Also in the previous
+pseudo-code ``\textit{st}'' was meant to indicate ``\texttt{MPI\_STATUS}''. Again
+we have reorganized the pseudo-code to make this point more clear. We have also
+added more comments and text to the caption to explain things in more details.
+\\
+\item ``\textit{In algorithm 2, if X is contained within the mesh on a given
+  processor, does that mean all the points needed in order to derive the value
+at X are also on the same processor? What if the points in the interpolatory
+stencil of X belong to multiple processors? Which processor does the actual
+interpolation? Is it the processor that needs the value at X? or is it the
+processor that owns X?}''
 
-\item ``\textit{}''
+Since the Z-curve defines non-overlapping partition of the grid in parallel, it
+is guarantied that the rank found by the search operation in line 3 of algorithm
+2, actually owns the interpolation point, i.e. there exists a \textit{local
+cell} \texttt{c} belonging to process \texttt{owners\_rank} which contains the
+interpolation point. This is because the search operation uses a binary search
+on the Z-curve to find the remote \texttt{owners\_rank}.
 
-\item ``\textit{}''
+Our method for interpolation only requires that the values of the function (and
+optionally its second derivatives if using multi-quadratic interpolation) be
+defined on all vertices of cell \texttt{c}. This, in turn is guarantied by the
+fact that the boundary of each process is covered by a layer of ghost cells that
+have updated the ghost values at all their vertices before performing the
+interpolation. As a result, all the vertices of cell \texttt{c}, whether local
+or ghost, have valid values that can be used for interpolation.
+\\
+\item ``\textit{In section 4.2, the authors should do a standard 3D reversible
+  test when checking speed-up. A standard test allows researchers to compare
+the new method to previous schemes. Also the authors should report the
+numerical error (e.g. the symmetric difference error or the level set error in
+proximity to the zero level set). For a reversible problem, the exact solution
+is known at the end time. So in Table 2, an extra column should be added to
+both the (a) and (b) parts which has the error (it is assumed that the error is
+independent of p?). For the reversible problem, the zero level set isosurface
+should be displayed at maximum deformation t = T/2 and at the end time t = T
+for some representative values of “CFL” and lmax.}''
 
-\item ``\textit{}''
+The idea behind the test in section 4.2 was to look at the individual
+components that make the semi-Lagrangian algorithm and check their scaling and
+see how different parameters, most notably the CFL number, would affect the
+scaling results. In doing so we decided to look at the performance of a single
+iteration of semi-Lagrangian algorithm in isolation without including other
+ingredients, such as reinitialization, that would otherwise affect the timing
+and scaling. In fact the scaling results for a more comprehensive test are
+included toward the end of article and in the Stefan test.
+
+Also, our main focus in this article has been the scalability of several level
+set algorithm in parallel. The accuracy of many of these algorithms, albeit in
+serial, have been studied before in relevant articles that we have cited in
+details. However, we agree with the reviewer that it is important to show that
+the presented methods are in fact accurate and convergent. As a result we have
+added a new section, 3.4, that checks the accuracy of the standard rotation
+test as was suggested by the reviewer.  
+\\
+\item ``\textit{Also in section 4.2, how frequently is reinitialization carried
+  out? What is the pseudo time step $\Delta \tau$? What is $\tau_{stop}$? If
+there is no reinitialization for this test, it is requested that the
+semi-Lagrangian algorithm together with reinitialization be analyzed in this
+section too}''
+
+Previous studies have shown that the adaptive time-stepping method that is used
+in this article greatly improves the convergence to steady state solution of
+the reinitialization equation (c.f. ref 33 in main text). We have generally
+observed that a few, typically 20, iterations of the algorithm is enough for
+the solution to converge within a small band (e.g. one smallest cell in the
+diagonal direction, i.e. $|\phi| < \sqrt{2}\Delta x_{min}$) around the
+interface. As a result we solve the reinitialization equation not to a final
+time (since each cell is using a different time-step) but to a fixed number of
+iterations. We have added a new paragraph in section 3.3 that clarifies this
+approach.
+
+As for the test in section 4.2, there is no reinitialization involved. This is
+because we are testing the effects of various parameters on the scalability of
+a single iteration of the semi-Lagrangian. We believe, however, that the
+reviewer is absolutely correct about the need to have a test of the entire
+algorithm. However, we argue that this role is already fulfilled by the Stefan
+test toward the end of the article.
+
+The Stefan test of section 5 is a complete and non-trivial example and
+represents a typical scenario that one should expect when using our algorithms
+in practice. This example serves, not only as a demonstration of what our
+algorithms can achieve, but also as a scalability test of all different
+components, such as interpolation, semi-Lagrangian advection, reinitialization,
+solution of implicit linear system, and extension of this solution over the
+interface. It also compares these costs with costs associated with grid
+operations in its entirety which includes generation, refinement and
+coarsening, and partitioning of the forest. As figure 10, and table 7,
+illustrate, both the semi-Lagrangian and the reinitialization scale well in a
+real scenario \textcolor{green}{@Arthur: How often do we call reinitialization
+and how many iterations each time? I think we should include these
+information}. Thus we hope the results of this test is satisfactory enough for
+the reviewer.
 \end{itemize} 
@@
 
-In light of these adjustments, we respectfully ask that our article be reconsidered for publication in the Journal of Computational Physics.
+\textcolor{red}{@Frederic: What should we do about the scaling of the full
+rotation test? Should we do it? I have the feeling that the reviewer is
+probably not going to be satisfied ... Deadline is Feb 21}
+
+In light of these adjustments, we respectfully ask that our article be
+reconsidered for publication in the Journal of Computational Physics.
 \closing{Sincerely,} 

### 2016-02-15 paper/tex/parallel_algorithms.tex
Made changes to paper and prepared the response letter

@@
 \begin{algorithm}[htbp]
-\caption{$H \gets \texttt{Reconstruct (}G\texttt{)}$: construction of the local tree hierarchy $H$ from the grid $G$}
+\caption{$H \gets \texttt{Reconstruct (}G\texttt{)}$: 
+Construction of the local tree hierarchy, $H$, from the parallel grid, $G$, supplied by \texttt{p4est}.
+The algorithm starts at the coarsest level, i.e.
+the macromesh, and recursively splits the each cell to match the finest local and ghost cells generated by \texttt{p4est}.}
 \begin{algorithmic}[1]
-\State $H \gets G.\texttt{macromesh()}$
-\For {$\mathit{tr} : G.\texttt{local\_trees()}$} \Comment{build hierarchy for local cells}
-	\For {$c : \mathit{tr}.\texttt{cells()}$}
+\State $H \gets G.\texttt{macromesh()}$ \Comment{start with the coarsest  grid}
+\For {$\mathit{tr} \ein G.\texttt{local\_trees()}$} \Comment{update the hierarchy to match local cells}
+	\For {$c \ein \mathit{tr}.\texttt{local\_cells()}$}
 		\State $H.\texttt{update\_tree(} \mathit{tr}, c \texttt{)}$
@@
 
-\For {$c : G.\texttt{ghost\_cells()}$} \Comment{build hierarchy for ghost cells}
-	\State $H.\texttt{update\_tree(} c.\texttt{tree()}, c \texttt{)}$
+\For {$c \ein G.\texttt{ghost\_cells()}$} \Comment{update the hierarchy to mach local cells}
+	\State $H.\texttt{update\_tree(} c.\texttt{get\_parent\_tree()}, c \texttt{)}$
 \EndFor
@@
 \Function{$H.\texttt{update\_tree}$}{$\mathit{tr}, c$}   \Comment{recursive tree reconstruction}
-	\State $c_l \gets H.\texttt{root(}\mathit{tr}\texttt{)}$
-	\While{$c_l.\texttt{level()} \not= c.\texttt{level()} $}
-		\If {$c_l.$\texttt{is\_leaf()}} $c_l$\texttt{.split()}
+	\State $c_l \gets H.\texttt{get\_root\_cell(}\mathit{tr}\texttt{)}$ 
+	\While{$c_l.\texttt{level()} \not= c.\texttt{level()} $} \Comment{recursively search for a cell of the same size as \textit{c}}
+		\If {$c_l.$\texttt{is\_leaf()}} $c_l$\texttt{.split()} \Comment{if the current cell is leaf, split the cell and continue search}
 		\EndIf
-		\State $h \gets c_l.\texttt{length()} / 2$ \Comment{select the next child based on direction}
-		\State $i \gets c.x \ge c_l.x + h$
-		\State $j \gets c.y \ge c_l.y + h$
-		\State $k \gets c.z \ge c_l.z + h$
+		\State $h \gets c_l.\texttt{length()}$ \Comment{select the search path based on cell coordinate.}
+		\State $i \gets c.x \ge c_l.x + h/2$	
+		\State $j \gets c.y \ge c_l.y + h/2$
+		\State $k \gets c.z \ge c_l.z + h/2$
 
-		\State $c_l \gets c_l.\texttt{child(} i, j, k \texttt{)}$
+		\State $c_l \gets c_l.\texttt{get\_child(} i, j, k \texttt{)}$
 	\EndWhile	
@@
 
-\subsection{Interpolation and semi-Lagrangian methods}
+\subsection{Interpolation and semi-Lagrangian methods} \label{sec:interpolation_semilagrangian}
 As indicated earlier, we use the semi-Lagrangian method to solve equation \eqref{eq:ls} when the velocity field is externally generated, i.e.\ when it does not depend explicitly on the level-set function itself.
@@
 \begin{algorithm}[htbp]
-\caption{$values \gets \texttt{Interpolate (}H, F, \underline{\mathbf{X}}\texttt{)}$: interpolate the value of $F$, defined on the local tree hierarchy $H$, at coordinates $\underline{\mathbf{X}}$}
+\caption{$values \gets \texttt{Interpolate (}H, F, \underline{\mathbf{X}}\texttt{)}$: interpolate the value of $F$, defined on the local tree hierarchy $H$, at coordinates $\underline{\mathbf{X}}$.
+This is achieved in three phases as detailed in the text.
+I) First, interpolation points are buffered in two arrays, those than can be interpolated locally and those that should be sent to other processes.
+II) In the second phase, each process initiates the remote buffer exchange and computes the number of messages it should expect to receive using the communication matrix idea described in section \ref{sec:interpolation_semilagrangian}.
+III) Finally, each process alternates between computing local interpolation and receiving and performing remote interpolation requests from other processes.}
 \begin{algorithmic}[1]
-\State $col \gets 0, \mathit{buff} \gets \mathit{null}$ \Comment{Phase I -- buffering}
-\For {$\underline{\mathbf{p}} : \underline{\mathbf{X}}$} 
-	\State $[r, \mathit{cell})] \gets H.\texttt{search(}\underline{\mathbf{p}}\texttt{)}$ \Comment{search for the owner's rank and cell}
-	\If {$r = \mathit{mpirank}$}
-        \State $\mathit{buff}[r].\texttt{push\_back(}\underline{\mathbf{p}}, cell\texttt{)}$
-	\Else
-        \State $\mathit{buff}[r].\texttt{push\_back(}\underline{\mathbf{p}}\texttt{)}$
-        \State $\mathit{col}[r] \gets 1$
+\State $col \gets 0, \mathit{local\_buff} \gets \mathit{null}, \mathit{remote\_buff} \gets \mathit{null}$ \Comment{Phase I -- buffering}
+\For {$\underline{\mathbf{p}} \ein \underline{\mathbf{X}}$} 
+	\State $[owners\_rank, \mathit{cell}] \gets H.\texttt{search(}\underline{\mathbf{p}}\texttt{)}$ \Comment{search for the owner's rank and cell}
+	\If {$owners\_rank = \mathit{mpirank}$} \Comment{can interpolate locally}
+        \State $\mathit{local\_buff}[owners\_rank].\texttt{push\_back(}\underline{\mathbf{p}}, cell\texttt{)}$ 
+	\Else \Comment{requires remote interpolation}
+        \State $\mathit{remote\_buff}[owners\_rank].\texttt{push\_back(}\underline{\mathbf{p}}\texttt{)}$ 
+        \State $\mathit{col}[owners\_rank] \gets 1$
 	\EndIf
 \EndFor
-\For {$r:\mathit{mpisize}$} \Comment{Phase II -- initiate communication and compute number of messages}
-	\If {$\mathit{col}[r]$}
-        \State $\texttt{MP\_Isend(}r, \mathit{buff}[r]\texttt{)}$
+\\
+\For {$r \ein \mathit{[0,mpisize)}$} \Comment{Phase II -- initiate communication and compute number of messages}
+	\If {$\mathit{col}[r] \deq 1$} 
+        \State $req \gets \texttt{MP\_Isend(}r, \texttt{QUERY\_TAG}, \mathit{remote\_buff}[r]\texttt{)}$ \Comment{initiate a non-blocking send to process \textit{r}}
+        	\State $\textit{query\_requests}.\texttt{push\_back(}req\texttt{)}$
 	\EndIf
 \EndFor
-\State  $S \gets \texttt{sum(}col\texttt{)}$ 
-\State  $R \gets \texttt{MPI\_Reduce\_scatter(}\mathit{col},\texttt{MPI\_SUM)}$
-\State $\mathit{done} \gets \mathit{false}, it \gets \mathit{buff}[mpirank].\texttt{begin()}$
-\While {$\mathit{!done}$} \Comment{Phase III -- main loop}
+\State  $S \gets \texttt{sum(}col\texttt{)}$ \Comment{compute number of messages to send}
+\State  $R \gets \texttt{MPI\_Reduce\_scatter(}\mathit{col},\texttt{MPI\_SUM)}$ \Comment{compute number of messages to receive}
+\\
+\State $\mathit{done} \gets \mathit{false}$ \Comment{Phase III -- interpolation}
+\State $it \gets \mathit{local\_buff}[mpirank].\texttt{begin()}$ 
+\While {$\mathit{!done}$} 
 	\If {$\mathit{it} \not= \mathit{buff[mpirank]}.\texttt{end()}$}
-		\State $\mathit{values} \gets \texttt{process\_local(}it\texttt{)}$ \Comment{process local interpolations}
+		\State $\mathit{values} \gets \texttt{process\_local\_interpolation(}it\texttt{)}$ \Comment{process local interpolations}
 		\State $\texttt{++}\mathit{it}$
 	\EndIf
-	\If {$R > 0$} \Comment{process queries sent from remote processes}
-		\State $[\mathit{msg, st}] \gets \texttt{MPI\_Iprobe()}$
-		\If {$\mathit{msg}$}
-			% \State $values \gets \texttt{MPI\_Recv(}st.\texttt{MPI\_SOURCE)}$ \Comment{}
-			\State $\mathit{val\_buff} \gets \texttt{process\_queries(}\mathit{st}\texttt{)}$ \Comment{receive, search, and interpolate values}
-			\State $\texttt{MPI\_Isend(}\mathit{st}.\texttt{MPI\_SOURCE},\mathit{val\_buff}\texttt{)}$ \Comment{send back interpolated values}
+	\If {$R > 0$} \Comment{search for interpolation query in the message queue}
+		\State $\mathit{message} \gets \texttt{MPI\_Iprobe(\texttt{MPI\_ANY\_SOURCE}, \texttt{QUERY\_TAG})}$
+		\If {$\mathit{message}.\texttt{is\_pending()}$}
+			\State $\mathit{values} \gets \texttt{process\_remote\_queries(}\mathit{message}\texttt{)}$ \Comment{receive, search, and interpolate values}
+			\State $req \gets \texttt{MPI\_Isend(}\mathit{message}.\texttt{MPI\_SOURCE}, \texttt{REPLY\_TAG}, \mathit{values}\texttt{)}$ \Comment{send back interpolated values}
+			\State $\textit{reply\_requests}.\texttt{push\_back(}req\texttt{)}$
 			\State $R\texttt{--}$
@@
 	\EndIf
-	\If {$S > 0$} \Comment{process replies sent to our queries}
-		\State $\mathit{[msg, st]} \gets \texttt{MPI\_Iprobe()}$
-		\If {$\mathit{msg}$}
-			\State $\mathit{values} \gets \texttt{process\_replies(}\mathit{st}.\texttt{MPI\_SOURCE)}$ \Comment{receive remotely interpolated values}
+	\If {$S > 0$} \Comment{search for interpolation reply in the message queue}
+		\State $\mathit{message} \gets \texttt{MPI\_Iprobe(\texttt{MPI\_ANY\_SOURCE}, \texttt{REPLY\_TAG})}$
+		\If {$\mathit{message}.\texttt{is\_pending()}$}
+			\State $\mathit{values} \gets \texttt{process\_replies(}\mathit{message}\texttt{)}$ \Comment{receive remotely interpolated values}
 			\State $S\texttt{--}$
@@
 	\EndIf
-	\State $\mathit{done} \gets S = 0 \: \And \: R = 0 \: \And \: \mathit{it = buff[mpirank]}.\texttt{end()}$
+	\State $\mathit{done} \gets S \deq 0 \And R \deq 0 \And \mathit{it \deq local\_buff[mpirank]}.\texttt{end()}$
 \EndWhile
+\State $\texttt{MPI\_Waitall(}query\_requests, reply\_requests\texttt{)}$ \Comment{make sure all messages have been received}
 \State \Return $\mathit{values}$
@@
 \State $\Delta t_l \gets \text{CFL} \times G^n.\texttt{hmin()} / \max \{\underline{\mathbf{u}}^n \} $
-\State $\Delta t   \gets \texttt{MPI\_Allreduce(}\Delta t_l,\texttt{MPI\_MIN)}$
-\State $H^n \gets \texttt{Reconstruct(} G^n\texttt{)}$ 
+\State $\Delta t   \gets \texttt{MPI\_Allreduce(}\Delta t_l,\texttt{MPI\_MIN)}$ \Comment{compute $\Delta t$ based on CFL condition across all processes}
+\State $H^n \gets \texttt{Reconstruct(} G^n\texttt{)}$  \Comment{using algorithm \ref{alg:reconstruction}}
 \State $G^{n+1}_0 \gets G^n$
@@
 	\State $\underline{\mathbf{X}}_d  \gets \texttt{ComputeDeparturePoints(}G^{n+1}_0, \underline{\mathbf{u}}^n, \underline{\mathbf{u}}^{n-1}, \Delta t\texttt{)}$ \Comment{using equations \ref{eq:xstar} -- \ref{eq:vn_p_half}}
-	\State $\phi^{n+1} \gets \texttt{Interpolate(}H^n, \phi^n, \underline{\textbf{X}}_d\texttt{)}$
+	\State $\phi^{n+1} \gets \texttt{Interpolate(}H^n, \phi^n, \underline{\textbf{X}}_d\texttt{)}$ \Comment{using algorithm \ref{alg:interpolation}}
 	\State $G^{n+1} \gets G^{n+1}_0.\texttt{refine\_and\_coarsen(}\phi^{n+1}\texttt{)}$ \Comment{using equations \ref{eq:refine} and \ref{eq:coarsen} as criteria}
@@
 Similar to \cite{Min;Gibou:07:A-second-order-accur}, equation \eqref{eq:semidiscrete_reinit} is integrated in time using the TVD-RK2 scheme with adaptive time-stepping in order to accelerate the convergence to the steady state.
+Generally it has been observed that adaptive time-stepping considerably improves the convergence rate and only a few iterations are needed if the signed-distance property is desired in a small band around the interface \cite{Min;Gibou:07:A-second-order-accur}.
+In this work, and based on previous findings, we use a fixed number of 20 iterations for the reinitialization equation.
+Of course, it should be noted that it is quite easy to define a custom tolerance as the termination criteria.
+
 Since the computation is based on a local stencil, the parallel implementation of this scheme is mostly trivial.

### 2016-02-15 paper/peer-reviews/response_letter.tex
spelling fix

@@
 Since the Z-curve defines non-overlapping partition of the grid in parallel, it
-is guarantied that the rank found by the search operation in line 3 of algorithm
+is guaranteed that the rank found by the search operation in line 3 of algorithm
 2, actually owns the interpolation point, i.e. there exists a \textit{local
@@
 optionally its second derivatives if using multi-quadratic interpolation) be
-defined on all vertices of cell \texttt{c}. This, in turn is guarantied by the
+defined on all vertices of cell \texttt{c}. This, in turn is guaranteed by the
 fact that the boundary of each process is covered by a layer of ghost cells that

### 2016-02-15 paper/tex/conclusion.tex
Small edits
update enright figure with blue color, add number of iterations for reinitilization, add citations for XSEDE, stampede and comet

@@
 We have also presented a simple parallelization technique for the reinitialization algorithm based on the pseudo-time transient formulation.
-Both the semi-Lagrangian and the reinitialization algorithms show good scalability up to 4096 processors.
+Both the semi-Lagrangian and the reinitialization algorithms show good scalability up to 4096 processors, the current limit of our account.
 
@@
 This application clearly illustrates the applicability of our algorithms to complex multi-scale problems that cannot be treated practically using the domain decomposition techniques on uniform grids.
-We believe that our findings can serve as a basis to simulating a wide range of multi-scale problems and free body problems.
+We believe that our findings can serve as a basis to simulating a wide range of multi-scale and free boundary problems.
 
@@
 The funding for this project was provided by ONR N00014-11-1-0027.
-We would also like to thank the technical staff at the Texas Advanced Computing
-Center (TACC) and the developer community of PETSc library for their valuable
+This work used the Extreme Science and Engineering Discovery Environment (XSEDE), which is supported by National Science Foundation grant number
+ACI-1053575. The authors acknowledge the Texas Advanced Computing
+Center (TACC) at the University of Texas at Austin for providing HPC resources that have contributed to the research results reported within this paper, URL: http://www.tacc.utexas.edu.
+We would also like to thank the developer community of PETSc library for their valuable
 suggestions during this project.

### 2016-02-15 paper/tex/parallel_algorithms.tex
update enright figure with blue color, add number of iterations for reinitilization, add citations for XSEDE, stampede and comet

@@
 forward in time until $t_{1/2}=1$ and then backward to its original state at $t_f=2$.
-With set the time step to $\Delta t = 5 \Delta x_{min}$, where $\Delta x_{min}$ is the size of the smallest cell in the forest.
+We set the time step to $\Delta t = 5 \Delta x_{min}$, where $\Delta x_{min}$ is the size of the smallest cell in the forest.
+The level-set function is reinitialized at every time step by applying 20 iterations of the reinitialization procedure.
 We monitor the volume loss and the error in the interface location at the final time, when the original shape should be recovered, as the finest resolution of the forest increases.
@@
 \begin{tabular}{ccccc}
-\includegraphics[width=.28\textwidth]{figures/enright_0.png}
+\includegraphics[width=.28\textwidth]{figures/enright_blue_0.png}
 &
@@
 &
-\includegraphics[width=.28\textwidth]{figures/enright_1.png}
+\includegraphics[width=.28\textwidth]{figures/enright_blue_1.png}
 &
@@
 &
-\includegraphics[width=.28\textwidth]{figures/enright_2.png}
+\includegraphics[width=.28\textwidth]{figures/enright_blue_2.png}
 \end{tabular}

### 2016-02-15 paper/tex/scaling_results.tex
update enright figure with blue color, add number of iterations for reinitilization, add citations for XSEDE, stampede and comet

@@
 In this section we present some results that demonstrate the scalability of our algorithms.
-All of our tests were ran on the Stampede cluster at the Texas Advanced Computing Center (TACC) where we are limited to $4096$ cores at most.
+All of our tests were ran on the Stampede cluster at the Texas Advanced Computing Center (TACC), a resource accessible through the Extreme Science and Engineering Discovery Environment (XSEDE) \cite{TownsCockerillDahanEtAl:14:XSEDE:AcceleratingScientific}, where we are limited to $4096$ cores at most.
 Each node of Stampede has 2 eight-core Xenon E5-2680 processes clocked at 2.7 GHz with 32 GB of DDR3-1600 MHz memory and interconnected using an InfiniBand network card.

### 2016-02-17 paper/tex/introduction.tex
Some edits on paper

@@
 One option is to replicate the entire grid on each process and to employ serial
-ordering techniques, as originally implemented by the \texttt{deal.II} library
-\cite{Bangerth;Hartmann;Kanschat:07:deal.II----a-General}, or serial
+ordering techniques, as implemented in an earlier version of the \texttt{deal.II} library
+\cite{Bangerth;Hartmann;Kanschat:07:deal.II----a-General}, or to use serial
 graph partitioners such as METIS \cite{KarypisKumar95}.
@@
 adds extra overhead that can limit the overall scalability.
+Refining a grid consisting of multiple trees using recursive coordinate
+bisection has been implemented in the SIERRA framework \cite{StewartEdwards04}.
 Interestingly, tree-based grids have a nice spatial ordering that naturally
@@
 % \cite{Burstedde;Wilcox;Ghattas:11:p4est:-Scalable-Algo}.
-
-In implementing the numerical solver, we make use of the popular \texttt{PETSc}
-\cite{Balay;Abhyankar;Adams;etal:14:PETSc-Web-page} library for linear algebra
-and its parallel primitives, such as parallel ghosted vector and scatter/gather
-operations, which simplifies the implementation.
 
@@
 However, if the signed-distance property is required in the entire domain, solving the Eikonal equation is more computationally efficient.
-Unfortunately, the most popular algorithm for solving the Eikonal equation, i.e.\ the Fast Marching Method \cite{Sethian:96:A-Fast-Marching-Leve,Sethian:99:Level-set-methods-an}, is inherently sequential due to causal relationship between grid points and cannot be easily parallelized.
+Unfortunately, the most popular algorithm for solving the Eikonal equation, i.e.\ the Fast Marching Method (FMM) \cite{Sethian:96:A-Fast-Marching-Leve,Sethian:99:Level-set-methods-an}, is inherently sequential due to causal relationship between grid points and cannot be easily parallelized.
 The Fast Sweeping Method (FSM) \cite{Zhao:05:A-fast-sweeping-meth} is an

### 2016-02-17 paper/tex/parallel_algorithms.tex
fixed a minor error in figure caption
Some edits on paper

@@
 In \texttt{p4est} the adaptive grid is represented as a non-overlapping collection of trees that are rooted in individual cells of a common coarse grid (cf.\ Figure \ref{fig:p4est_zcurve}).
-This common coarse grid, which we will refer to as the ``macromesh'', can in general be an unstructured quadrilateral mesh, in two spatial dimensions, or hexahedral mesh, in three spatial dimensions.
+This common coarse grid, which we will refer to as the ``macromesh'', can in
+general be an unstructured quadrilateral mesh, in two spatial dimensions, or
+hexahedral mesh, in three spatial dimensions.
+(In \cite{StewartEdwards04} such a coarsest possible grid has been introduced
+as the ``genesis'' mesh.)
+%
 In this article, however, it is sufficient to limit discussions to simple
@@
 can be entirely replicated on all processes.
-For instance, in many of the applications that we are considering in this paper the macromesh is simply a single cell.
+For instance, in many of the applications that we are considering in this paper the macromesh is simply a single cubical cell.
 \texttt{p4est} allows for arbitrary refinement and coarsening criteria through defining callback functions.
@@
 naturally leads to clustering of most cells in each domain.  Right: schematic of
-a tree data structure representing this forest and its partitioning.}
+a tree data structure representing this forest and its partitioning.%
+% (The \pforest code actually supports both 2D and 3D meshes whose trees are
+% rotated against each other.)
+}
 \label{fig:p4est_zcurve}
@@
 This is important because we can entirely skip the 2:1 balance function, which
-was shown to be one of the most time consuming parts of grid adaptation
+was shown to be one of the more time consuming parts of grid adaptation
 in \texttt{p4est}
 \cite{Burstedde;Wilcox;Ghattas:11:p4est:-Scalable-Algo}.
+% \cite{IsaacBursteddeGhattas12}.
 % (it has since been optimized \cite{IsaacBursteddeGhattas12}).
@@
 Finally, in \texttt{p4est} trees are linearized, i.e.\ only the leaves are explicitly stored.
-However, explicit knowledge of the hierarchal structure of the tree is greatly beneficial in several algorithms, e.g.
-in search operations needed for the interpolation algorithm.
+However, explicit knowledge of the hierarchal structure of the tree is greatly beneficial in several algorithms, e.g.\ in search operations needed for the interpolation algorithm.
 Thus, we introduce a simple reconstruction algorithm that recreates a local representation of the entire ``forest'' that is only adapted to local cells and, potentially, the ghost layer.
@@
 The semi-Lagrangian method integrates equations \eqref{eq:sl} backward in time,
-i.e.\ starting from the grid $G^{n+1}$ (computed in iteratively as explained
+i.e.\ starting from the grid $G^{n+1}$ (computed iteratively as explained
 later on), we simply write $\phi^{n+1}(\underline{\mathbf{X}}^{n+1}) =
@@
 \phi(\underline{\mathbf{X}}(t^n), t^n) = \phi^n(\underline{\mathbf{X}}_d)$.
-Here, the characteristic curves are chosen such that $\underline{\mathbf{X}}(t^{n+1})$ are simply the coordinates of grids points of $G^{n+1}$, and $\underline{\mathbf{X}}_d$ are the departure points, which are computed using the second-order midpoint method \cite{Min;Gibou:07:A-second-order-accur}:
+Here, the characteristic curves are chosen such that $\underline{\mathbf{X}}(t^{n+1})$ are the coordinates of grid points of $G^{n+1}$, and $\underline{\mathbf{X}}_d$ are the departure points, which are computed using the second-order midpoint method \cite{Min;Gibou:07:A-second-order-accur}:
 \bea
@@
 If the point is owned by a local cell, it is added to a local buffer, otherwise we find the process which owns the point and add the point to a separate buffer belonging to the found rank.
-Note that searching the point in the local tree is performed recursively using the hierarchal reconstruction (cf.\ algorithm \ref{alg:reconstruction}).
+Note that searching the point in the local tree is performed recursively using the hierarchal reconstruction (cf.\ Algorithm~\ref{alg:reconstruction}).
 Moreover, the owner's rank is found by computing the Z-index of the point and then using a binary search on the Z-curve.
@@
 \While {$\mathit{!done}$} 
-	\If {$\mathit{it} \not= \mathit{buff[mpirank]}.\texttt{end()}$}
+	\If {$\mathit{it} \not= \mathit{local\_buff[mpirank]}.\texttt{end()}$}
 		\State $\mathit{values} \gets \texttt{process\_local\_interpolation(}it\texttt{)}$ \Comment{process local interpolations}
@@
 \end{tabular}
-\caption{Visualization of the deformation undergone by the sphere for the Enright's test at time 0, $t_{1/2}$ and $t_f$ and colored with the processes ranks.
+\caption{Visualization of the deformation undergone by the sphere for the Enright's test at time 0, $t_{1/2}$ and $t_f$.
 The forest's finest resolution is equivalent to a uniform grid with $2048^3$ cells and contains between 17 (at initial and final times) and 63 (at $t_{1/2}$) million nodes.

### 2016-02-17 paper/tex/introduction.tex
Final literature updates

@@
 The idea of using SFCs for parallel partitioning of Quad-/Octrees is not new in itself and has been used by many researchers.
-For instance, \texttt{Octor} \cite{Tu;OHallaron;Ghattas:05:Scalable-parallel-oc} uses a Morton curve (also known as Z-curve) for traversing the leaves of an Octree for indexing and load balancing and has been scaled up to 62,000 cores \cite{Burstedde;Ghattas;Gurnis;etal:08:Scalable-adaptive-ma}.
+For instance, \texttt{Octor} \cite{Tu;OHallaron;Ghattas:05:Scalable-parallel-oc} uses a Morton curve (also known as Z-curve) for traversing the leaves of an Octree for indexing and load balancing and has been scaled up to 62,000 CPU cores \cite{Burstedde;Ghattas;Gurnis;etal:08:Scalable-adaptive-ma}.
 \texttt{Dendro} \cite{Sampath;Adavani;Sundar;etal:08:Dendro:-parallel-alg} is
@@
 % 200,000 cores,
-450,000 CPU cores \cite{IsaacBursteddeWilcoxEtAl15}.
+450k CPU cores \cite{IsaacBursteddeWilcoxEtAl15}.
+Applications built with \pforest have scaled to 1.6M cores
+\cite{RudiMalossiIsaacEtAl15}
+and 2.9M cores \cite{MullerKoperaMarrasEtAl15}.
 In fact, the algorithms presented in this paper are implemented on top of

### 2016-02-18 paper/tex/applications.tex
update enright figure with blue color, add number of iterations for reinitilization, add citations for XSEDE, stampede and comet
first revision submitted to JCP.

@@
 The maximum tree resolution for the small test is 9, leading to approximately 7M grid points, and the maximum resolution for the large test is 11, corresponding to 105M grid points.
+The level-set function is reinitialized at every time-step by applying 20 iterations of the reinitialization procedure.
 The results are presented in figure \ref{fig:stefan_scaling}, where ``\verb|Solution_Extension|'' refers to extrapolation procedure (see Algorithm \ref{alg:stefan} step 5).
@@
 \subfigure{
-\includegraphics[width=.24\textwidth]{stefan/nb2_iter19.png}
-\includegraphics[width=.24\textwidth]{stefan/nb2_iter39.png}
-\includegraphics[width=.24\textwidth]{stefan/nb2_iter59.png}
-\includegraphics[width=.24\textwidth]{stefan/nb2_iter79.png}
+\includegraphics[width=.24\textwidth]{figures/stefan_nb2_iter19.png}
+\includegraphics[width=.24\textwidth]{figures/stefan_nb2_iter39.png}
+\includegraphics[width=.24\textwidth]{figures/stefan_nb2_iter59.png}
+\includegraphics[width=.24\textwidth]{figures/stefan_nb2_iter79.png}
 }
 \subfigure{
-\includegraphics[width=.24\textwidth]{stefan/nb5_iter19.png}
-\includegraphics[width=.24\textwidth]{stefan/nb5_iter39.png}
-\includegraphics[width=.24\textwidth]{stefan/nb5_iter59.png}
-\includegraphics[width=.24\textwidth]{stefan/nb5_iter79.png}
+\includegraphics[width=.24\textwidth]{figures/stefan_nb5_iter19.png}
+\includegraphics[width=.24\textwidth]{figures/stefan_nb5_iter39.png}
+\includegraphics[width=.24\textwidth]{figures/stefan_nb5_iter59.png}
+\includegraphics[width=.24\textwidth]{figures/stefan_nb5_iter79.png}
 }
 \subfigure{
-\includegraphics[width=.24\textwidth]{stefan/nb8_iter19.png}
-\includegraphics[width=.24\textwidth]{stefan/nb8_iter39.png}
-\includegraphics[width=.24\textwidth]{stefan/nb8_iter59.png}
-\includegraphics[width=.24\textwidth]{stefan/nb8_iter79.png}
+\includegraphics[width=.24\textwidth]{figures/stefan_nb8_iter19.png}
+\includegraphics[width=.24\textwidth]{figures/stefan_nb8_iter39.png}
+\includegraphics[width=.24\textwidth]{figures/stefan_nb8_iter59.png}
+\includegraphics[width=.24\textwidth]{figures/stefan_nb8_iter79.png}
 }
 \subfigure{
-\includegraphics[width=.24\textwidth]{stefan/nb12_iter19.png}
-\includegraphics[width=.24\textwidth]{stefan/nb12_iter39.png}
-\includegraphics[width=.24\textwidth]{stefan/nb12_iter59.png}
-\includegraphics[width=.24\textwidth]{stefan/nb12_iter79.png}
+\includegraphics[width=.24\textwidth]{figures/stefan_nb12_iter19.png}
+\includegraphics[width=.24\textwidth]{figures/stefan_nb12_iter39.png}
+\includegraphics[width=.24\textwidth]{figures/stefan_nb12_iter59.png}
+\includegraphics[width=.24\textwidth]{figures/stefan_nb12_iter79.png}
 }

### 2016-03-17 doc/casl_p4est_manual/casl_p4est_manual.tex
adding beginning of casl_p4est manual

\documentclass{article}
\usepackage{hyperref}
\usepackage[utf8]{inputenc}
\usepackage{wrapfig}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage[tmargin=1in,bmargin=1in]{geometry}
\usepackage{subcaption}
\usepackage{graphicx}
\usepackage{latexsym}
\usepackage{pdfsync}
\usepackage[boxed]{algorithm}
\usepackage{algpseudocode}
\usepackage{multirow}
\usepackage{rotating}
\usepackage{color}
\usepackage{caption}
\usepackage{url}
\usepackage{diagbox}

\usepackage{fancyvrb}
\usepackage{listings}

\hypersetup{
    colorlinks=true,      		% false: boxed links; true: colored links
    linkcolor=blue,       		% color of internal links
    citecolor=blue,       		% color of links to bibliography
    filecolor=black,      		% color of file links
    urlcolor=red,       		% color of external links
    bookmarks=false,
    pdffitwindow=true,
    pdfpagelayout=SinglePage
}

\newcommand{\etal}{\textit{et al.\ }}
\newcommand{\pd}[2]{\frac{\partial #1}{\partial #2}}
\newcommand{\mbf}[1]{\mathbf{#1}}
\newcommand{\vect}[1]{\underline{\boldsymbol{#1}}}
\newcommand{\tensor}[1]{\underline{\underline{\boldsymbol{#1}}}}
\newcommand{\pest}{\texttt{p4est}}
\newcommand{\caslpest}{\texttt{casl\_p4est}}
\newcommand{\ttt}[1]{\texttt{#1}}
\newcommand{\figb}{\begin{figure}[htbp]\begin{center}}
\newcommand{\fige}{\end{center}\end{figure}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\makeindex

\begin{document}

\title{User manual for \texttt{casl\_p4est}}

\author{arthur guittet}

\maketitle

\tableofcontents

\section{Installation}

\subsection{Access to Stampede and Comet}

The first step is to create an account on the XSEDE website (www.xsede.org). After that, the account can be added to an active XSEDE grant by the corresponding PI and the resources available through that allocation can be activated. They need to be activated one by one, for instance if both Stampede and Comet are part of the allocation, each of them needs to be activated separately for the specific user.

Once the access is granted, the login nodes of the clusters can be accessed through ssh with the following commands,
\begin{Verbatim}[frame=single]
ssh username@stampede.tacc.utexas.edu
ssh username@comet.sdsc.edu
\end{Verbatim}
The password is the XSEDE password. Note that when updating the password on XSEDE, several minutes can be required for the update to be propagated to the supercomputers. For ease of access, the following information about an ssh server can be added in \texttt{.ssh/config} (for unix systems)
\begin{Verbatim}[frame=single]
Host stampede
Hostname stampede.tacc.utexas.edu
User username
\end{Verbatim}
A different name can be chosen for the \texttt{Host} part. The supercomputer can then be accessed with
\begin{Verbatim}[frame=single]
ssh stampede
\end{Verbatim}

To transfer files from and to a supercomputer, use the ``scp" or ``rsync" commands for small file, and ``globus" for big transfers between supercomputers (checkout www.globus.org). The documentation for Stampede can be found online on the Texas Advanced Computing Center website, at https://portal.tacc.utexas.edu/user-guides/stampede, and is very well written. The documentation for Comet is not as good ... but they work very similarly.

\subsection{Installing \texttt{p4est}}

The \texttt{p4est} library can be installed from a tarball or directly from the git repository. Some bugs have been fixed and numerous features added since the 1.1 version, and using the git repository is recommended. Start by cloning the repository with
\begin{Verbatim}[frame=single]
git clone https://github.com/cburstedde/p4est.git path_to_local_git_folder
\end{Verbatim}
In the folder created (replace ``path\_to\_local\_git\_folder" with you choice), run
\begin{Verbatim}[frame=single]
git submodule init && git submodule update
\end{Verbatim}
to initialize the \texttt{sc} submodule on which \texttt{p4est} depends. A memory alignment bug was fixed recently, and the branch with the corresponding fix must be selected by
\begin{Verbatim}[frame=single]
git checkout fix-memalign
\end{Verbatim}
in both the main git folder and the ``sc" sub-folder. You then need to run
\begin{Verbatim}[frame=single]
./bootstrap
\end{Verbatim}
from the git folder, followed by the configure command with your choice of option. For a release version, use
\begin{Verbatim}[frame=single]
./configure --prefix=/path/to/install/folder --enable-mpi --without-blas
--enable-shared --enable-memalign=16 CFLAGS=-O2 CPPFLAGS=-O2 FCFLAGS=-O2
\end{Verbatim}
If building for debug, the run
\begin{Verbatim}[frame=single]
./configure --prefix=/path/to/install/folder --enable-mpi --without-blas
--enable-shared --enable-debug --enable-memalign=16 CFLAGS="-O0 -g"
CPPFLAGS="-O0 -g" FCFLAGS="-O0 -g"
\end{Verbatim}
Note that you must have mpi installed and in the PATH. The default install folder (it should be something like ``/usr/lib/p4est") can also be selected by omitting the ``prefix" option. The \texttt{p4est} library can no be built and installed by running
\begin{Verbatim}[frame=single]
./make && ./make install
\end{Verbatim}

The \texttt{casl\_p4est} library also requires \texttt{Petsc}, which is available on the public repositories if running Linux Mint or Ubuntu. Otherwise, you can install it by following the instructions from the \texttt{Petsc} website (https://www.mcs.anl.gov/petsc/). Here are the flags I use if the package is not available on a public repository
\begin{Verbatim}[frame=single]
./configure --download-fblaslapack --download-hypre=1
--prefix=/path/to/install/dir --with-debugging=0 --with-mpi-dir=/path/to/mpi/dir
--with-shared-libraries=1 COPTFLAGS="-O2" CXXOPTFLAGS="-O2" FOPTFLAGS="-O2"
\end{Verbatim}

On Stampede and Comet, you can load the \texttt{Petsc} module with the module manager, by running
\begin{Verbatim}[frame=single]
module load petsc
\end{Verbatim}
The \texttt{p4est} library is also available as a module on Stampede, however it's the 1.1 version that does not have some of the new features and bug fixes that are needed by \texttt{casl\_p4est}.

\subsection{Installing \texttt{casl\_p4est}}

The first step is to create a bitbucket account (at bitbucket.org) and ask someone with admin rights to add you to the git repository. Once this is done, you can clone the repository by running
\begin{Verbatim}[frame=single]
git clone https://username@bitbucket.org/cburstedde/casl_p4est.git local_folder
\end{Verbatim}
where ``username" is your bitbucket username and ``local\_folder" is the path to the folder where the library is to be installed. You can use git through the command line or with a Graphical User Interface. I personally recommend SmartGit.

\subsection{Compiling against the \texttt{casl\_p4est} library}

A sample \texttt{Makefile} and a sample \texttt{project.pro} file are located in the ``doc/casl\_p4est\_manual" folder. The general things you need in you \texttt{Makefile} are
\begin{itemize}
\item include the headers for the \texttt{p4est} library and path to the compiled library for the linker
\begin{Verbatim}[frame=single]
INCPATH += -I/path/to/the/p4est/install/include
LIBS += -Wl,-rpath,/path/to/the/p4est/install/lib
	-L/path/to/the/p4est/install/lib -lp4est -lsc
\end{Verbatim}
\item include the headers for the \texttt{Petsc} library and path to the compiled library for the linker. Depending on the system, the PETSC\_DIR might already be set as an environment variable, it is for instance the case on Stampede.
\begin{Verbatim}[frame=single]
INCPATH += -I/path/to/the/petsc/install/include
LIBS += -Wl,-rpath,/path/to/the/petsc/install/lib
	-L/path/to/the/petsc/install/lib -lpetsc
\end{Verbatim}
\item include the headers for the \texttt{Petsc} library and path to the compiled library for the linker. Depending on the system, the PETSC\_DIR might already be set as an environment variable, it is for instance the case on Stampede.
\begin{Verbatim}[frame=single]
INCPATH += -I/path/to/the/casl/p4est/library
\end{Verbatim}
\end{itemize}

\section{Using the \texttt{p4est} library}

The following sections contain information about the \texttt{p4est} data structures and how to use them, including samples of code.

\subsection{Forests of Octrees and parallelization}

The \pest library works with forests of Octrees, that is a collection of Octrees rooted in a shared macromesh. The macromesh can be complicated, but for the \caslpest we limit ourselves to Cartesian macromeshes. This can be changed in the future if needed. The first step is therefore to declare a macromesh, with
\begin{Verbatim}[frame=single]
my_p4est_brick_t brick;
p4est_connectivity_t *connectivity = my_p4est_brick_new(nx, ny, nz,
				xm, xM, ym, yM, zm, zM, &brick, px, py, pz);
\end{Verbatim}
with \ttt{nx}, \ttt{ny} and \ttt{nz} are the dimensions of the macromesh. For instance, setting those to 2 will generate a 2x2x2 macromesh, i.e. the framework for a forest of 8 Octrees. The other parameters are the physical dimensions of the domain and the periodicity information. Every process knows about the macromesh, it is a shared information.

The forest can now be created and the Octrees refined, for example with a level-set function \ttt{LS} defined as a CF\_3 (Continuous Function in $\mathbb{R}^3$, or CF\_2 in 2d), with
\begin{Verbatim}[frame=single]
p4est_t *p4est = my_p4est_new(mpicomm, connectivity, 0, NULL, NULL);
splitting_criteria_cf_t criteria(lmin, lmax, &LS, 1.2);
p4est->user_pointer = (void*)(&criteria);
for(int lvl=0; lvl<lmax; ++lvl)
{
	my_p4est_refine(p4est, P4EST_FALSE, refine_levelset_cf, NULL);
	my_p4est_partition(p4est, P4EST_FALSE, NULL);
}
\end{Verbatim}
with \ttt{mpicomm} the mpi communicator to use, \ttt{lmin} and \ttt{lmax} the min and max level of each Octree, and \ttt{1.2} the Lipschitz constant for the level-set function (choosen conservatively to be larger than 1 here). The refinement is done within a loop because initially the macromesh is distributed across the processes evenly. This means that a 1x1x1 macromesh will be attributed to a single process, even if running 1024 mpi tasks. Conversely, a 2x2x2 macromesh with 4 processes will result in 2 blocks per process.

The forest is partitioned across processes by linearizing the trees, i.e. building an array of the leaves ordered by a particular procedure based on a space-filling curve. \ttt{p4est} uses the Z-ordering, or Morton ordering. Though alternatives exist (Hilbert curve for instance), their performances are similar to the Z-curve. The array is then distributed evenly across the processes. Figure \ref{fig::zcurve} illustrates the process.
\figb
\includegraphics[width=.9\textwidth]{p4est_zcurve.pdf}
\caption{Illustration of the Z-curve ordering and of the partitioning process for a forest of 2 Octrees (macromesh 2x1) and for 3 mpi processes.} \label{fig::zcurve}
\fige

\subsection{Organization of the quadrants}

The forest is now ready for use if the algorithm is based on cell-centered data. Each process knows about its local chunk of the quadrants array, stored inside the local tree structures. The number of quadrants owned by a process is given by
\begin{Verbatim}[frame=single]
p4est->local_num_quadrants
\end{Verbatim}
while the entire forest contains
\begin{Verbatim}[frame=single]
p4est->global_num_quadrants
\end{Verbatim}
Each chunk of the local quadrants belongs to a tree, and thanks to the Z-curve property the values are contiguous. Each tree stores the index at which its quadrants start in the local ordering in
\begin{Verbatim}[frame=single]
tree->quadrants_offset
\end{Verbatim}
\figb
\includegraphics[width=\textwidth]{figures/quadrants.pdf}
\caption{Representation of the quadrants data structure. $T_f$ is the first local tree for the current mpi process, $T_l$ is the last local tree, and $N_{T_f}$ is the number of local quadrants located in tree $T_f$, i.e.\ $T_f \rightarrow \texttt{quadrants} \rightarrow  \texttt{elem\_count}$.} \label{fig::quadrants}
\fige
Figure \ref{fig::quadrants} summarizes the \ttt{quadrants} structure, and the code structure for accessing the quadrants information is
\begin{Verbatim}[frame=single]
for(p4est_topidx_t tree_idx=p4est->first_local_tree;
	tree_idx<=p4est->last_local_tree;
	++tree_idx)
{
  p4est_tree_t *tree = (p4est_tree_t*)sc_array_index(p4est->trees, tree_idx);
  for(size_t q=0; q<tree->quadrants.elem_count; ++q)
  {
    p4est_locidx_t quad_idx = q+tree->quadrants_offset;
    p4est_quadrant_t *quad =
    		(p4est_quadrant_t*)sc_array_index(&tree->quadrants, q);
  }
}
\end{Verbatim}
Here \ttt{quad\_idx} is the index of the quadrant in the local ordering, and \ttt{q} is the index of the quadrant in the local tree. The global index of a local quadrant can be obtained by adding the global offset of its owning process to its index
\begin{Verbatim}[frame=single]
p4est_gloidx_t gloidx = quad_idx + p4est->global_first_quadrant[p4est->mpirank];
\end{Verbatim}

\subsection{The ghost quadrants data structure}

So far, each process only knows about its local information. The \texttt{p4est\_ghost\_t} gives access to a layer of ghost cells around a process. The structure can be initialized with
\begin{Verbatim}[frame=single]
p4est_ghost_t *ghost = my_p4est_ghost_new(p4est, P4EST_CONNECT_FULL);
\end{Verbatim}
\ttt{P4EST\_CONNECT\_FULL} indicates that the ghost neighbors are gathered across faces, edges and corners. This constructs the structure for a ghost layer of depth one. The ghost layer can be expanded with
\begin{Verbatim}[frame=single]
my_p4est_ghost_expand(p4est, ghost);
\end{Verbatim}
This results in a ghost layer of depth two. The expand function can be called the desired number of times to build a ghost layer of larger depth.

The ghost data structures contains the array of quadrants in the ghost layer, stored in
\begin{Verbatim}[frame=single]
ghost->ghosts
\end{Verbatim}
and of size
\begin{Verbatim}[frame=single]
ghost->ghosts.elem_count
\end{Verbatim}
One can loop over the ghost directly with
\begin{Verbatim}[frame=single]
for (g = 0; g < ghost->ghosts.elem_count; ++g)
{
      q = (p4est_quadrant_t*)sc_array_index(&ghost->ghosts, g);
}
\end{Verbatim}
The ghost structure also contains the information about the owner of a ghost quadrant and the tree it belongs to. Figure \ref{fig::ghost_quadrants} illustrate the structure of this information.
\figb
\includegraphics[width=\textwidth]{figures/ghosts.pdf}
\caption{Representation of the ghost layer information. The \ttt{tree\_offsets} arrays and \ttt{proc\_offsets} are members of the \ttt{ghost\_t} data structure. $T_k$ corresponds to the tree of index $k$ and $p_m$ to mpi process of index $m$. For this particular example, the trees of index strictly smaller than $k$ or larger or equal to $l$ are not represented in the ghost layer, and similarly none of the quadrants in the ghost layer belong to a mpi processes with a rank strictly smaller than $m$ or larger or equal to $n$.} \label{fig::ghost_quadrants}
\fige

\subsection{The nodes data structure}

\subsection{Communications with \texttt{Petsc}}

\newpage
\bibliographystyle{abbrv}
\addcontentsline{toc}{section}{\refname}
\bibliography{references}

\end{document}

### 2016-03-17 examples/nodes_mls/main_test_nodes_mls_2d.cpp
1

@@ -1,0 +1,846 @@
+
+/*
+ * Test the cell based multi level-set p4est.
+ * Intersection of two circles
+ *
+ * run the program with the -help flag to see the available options
+ */
+
+// System
+#include <stdexcept>
+#include <iostream>
+#include <sys/stat.h>
+#include <vector>
+#include <algorithm>
+#include <iterator>
+#include <set>
+#include <time.h>
+#include <stdio.h>
+
+// p4est Library
+#ifdef P4_TO_P8
+#include <p8est_bits.h>
+#include <p8est_extended.h>
+#include <src/my_p8est_utils.h>
+#include <src/my_p8est_vtk.h>
+#include <src/my_p8est_nodes.h>
+#include <src/my_p8est_tools.h>
+#include <src/my_p8est_refine_coarsen.h>
+#include <src/my_p8est_log_wrappers.h>
+#include <src/my_p8est_node_neighbors.h>
+#include <src/my_p8est_level_set.h>
+#include <src/my_p8est_poisson_nodes.h>
+#include <src/my_p8est_interpolation_nodes.h>
+#else
+#include <p4est_bits.h>
+#include <p4est_extended.h>
+#include <src/my_p4est_utils.h>
+#include <src/my_p4est_vtk.h>
+#include <src/my_p4est_nodes.h>
+#include <src/my_p4est_tools.h>
+#include <src/my_p4est_refine_coarsen.h>
+#include <src/my_p4est_log_wrappers.h>
+#include <src/my_p4est_node_neighbors.h>
+#include <src/my_p4est_level_set.h>
+#include <src/my_p4est_poisson_nodes.h>
+#include <src/my_p4est_interpolation_nodes.h>
+#endif
+
+#include <src/point3.h>
+
+#include <src/petsc_compatibility.h>
+#include <src/Parser.h>
+
+#undef MIN
+#undef MAX
+
+double xmin = -1;
+double xmax =  1;
+double ymin = -1;
+double ymax =  1;
+#ifdef P4_TO_P8
+double zmin = -1;
+double zmax =  1;
+#endif
+
+using namespace std;
+
+int lmin = 5; // uniform grid for time being
+int lmax = 5;
+int nb_splits = 1;
+
+int nx = 1;
+int ny = 1;
+#ifdef P4_TO_P8
+int nz = 1;
+#endif
+
+bool save_vtk = true;
+
+double mu = 1;
[…]

### 2016-03-17 examples/nodes_mls/main_test_nodes_mls_3d.cpp

@@ -1,0 +1,2 @@
+#include <src/my_p4est_to_p8est.h>
+#include "main_test_nodes_2d.cpp"

### 2016-03-17 src/my_p4est_poisson_nodes_mls.h

@@ -1,0 +1,142 @@
+#ifndef MY_P4EST_POISSON_NODES_MLS_H
+#define MY_P4EST_POISSON_NODES_MLS_H
+
+#include <petsc.h>
+
+#ifdef P4_TO_P8
+#include <src/my_p8est_node_neighbors.h>
+#include <src/my_p8est_tools.h>
+#include <src/my_p8est_interpolation_nodes.h>
+#include <src/my_p8est_utils.h>
+#else
+#include <src/my_p4est_node_neighbors.h>
+#include <src/my_p4est_tools.h>
+#include <src/my_p4est_interpolation_nodes.h>
+#include <src/my_p4est_utils.h>
+#endif
+
+enum {
+  INSIDE,
+  DIRICHLET,
+  MIXED_IN,
+  MIXED_OUT,
+  NEUMANN,
+  OUTSIDE
+} NodeLoc;
+
+class my_p4est_poisson_nodes_mls_t
+{
+  const my_p4est_node_neighbors_t *node_neighbors_;
+
+  // p4est objects
+  p4est_t *p4est;
+  p4est_nodes_t *nodes;
+  p4est_ghost_t *ghost;
+  my_p4est_brick_t *myb_;
+  std::vector<my_p4est_interpolation_nodes_t> phi_interp;
+
+  bool neumann_wall_first_order;
+  double mu_, diag_add_;
+  bool is_matrix_computed;
+  int matrix_has_nullspace;
+  double dx_min, dy_min, d_min, diag_min;
+#ifdef P4_TO_P8
+  double dz_min;
+#endif
+#ifdef P4_TO_P8
+  std::vector<BoundaryConditions3D> *bc_;
+#else
+  std::vector<BoundaryConditions2D> *bc_;
+#endif
+  std::vector<PetscInt> global_node_offset;
+  std::vector<PetscInt> petsc_gloidx;
+
+  std::vector<int> *color_;
+  std::vector<Action> *action_;
+  std::vector<NodeLoc> node_loc;
+
+  // PETSc objects
+  Mat A;
+  p4est_gloidx_t fixed_value_idx_g;
+  p4est_gloidx_t fixed_value_idx_l;
+  bool is_phi_dd_owned, is_mue_dd_owned;
+  Vec rhs_, add_, mue_, mue_xx_, mue_yy_;
+  std::vector<Vec> *phi_, *phi_xx_, *phi_yy_;
[…]

### 2016-03-17 src/my_p4est_utils.h

@@ -127,6 +127,7 @@ private:
 
   const CF_2 *p_WallValue;
   const CF_2 *p_InterfaceValue;
+  const CF_2 *p_InterfaceCoeff; // constant in Robin BC
 
 public:
   BoundaryConditions2D()
@@ -135,6 +136,7 @@ public:
     p_WallValue = NULL;
     InterfaceType_ = NOINTERFACE;
     p_InterfaceValue = NULL;
+    p_InterfaceCoeff = NULL;
   }
 
   inline void setWallTypes( const WallBC2D& w )
@@ -161,6 +163,14 @@ public:
 
   inline const CF_2& getInterfaceValue(){
     return *p_InterfaceValue;
+  }
+
+  inline void setInterfaceCoeff(const CF_2& in){
+    p_InterfaceCoeff = &in;
+  }
+
+  inline const CF_2& getInterfaceCoeff(){
+    return *p_InterfaceCoeff;
   }
 
   inline BoundaryConditionType wallType( double x, double y ) const
@@ -198,6 +208,7 @@ private:
 
   const CF_3 *p_WallValue;
   const CF_3 *p_InterfaceValue;
+  const CF_3 *p_InterfaceCoeff; // constant in Robin BC
 
 public:
   BoundaryConditions3D()
@@ -206,6 +217,7 @@ public:
     p_WallValue = NULL;
     InterfaceType_ = NOINTERFACE;
     p_InterfaceValue = NULL;
+    p_InterfaceCoeff = NULL;
   }
 
   inline void setWallTypes( const WallBC3D& w )
@@ -232,6 +244,14 @@ public:
 
   inline const CF_3& getInterfaceValue(){
     return *p_InterfaceValue;
+  }
+
+  inline void setInterfaceCoeff(const CF_3& in){
+    p_InterfaceCoeff = &in;
+  }
+
+  inline const CF_3& getInterfaceCoeff(){
+    return *p_InterfaceCoeff;
   }
 
   inline BoundaryConditionType wallType( double x, double y, double z ) const

### 2016-03-18 src/my_p4est_interpolation.cpp
working on the manual for casl_p4est, in progress

@@ -19,7 +19,7 @@
 
 // Logging variables -- defined in src/petsc_logging.cpp
 #ifndef CASL_LOG_EVENTS
-#undef  PetscLogEventBegintig
+#undef  PetscLogEventBegin
 #undef  PetscLogEventEnd
 #define PetscLogEventBegin(e, o1, o2, o3, o4) 0
 #define PetscLogEventEnd(e, o1, o2, o3, o4) 0

### 2016-03-18 doc/casl_p4est_manual/casl_p4est_manual.tex

@@
 \figb
-\includegraphics[width=.9\textwidth]{p4est_zcurve.pdf}
+\includegraphics[width=.9\textwidth]{figures/p4est_zcurve.pdf}
 \caption{Illustration of the Z-curve ordering and of the partitioning process for a forest of 2 Octrees (macromesh 2x1) and for 3 mpi processes.} \label{fig::zcurve}
@@
 
+If working on node (i.e.\ vertices) based data, the nodes structure must be initialized.
+It contains the basic information about the nodes, such as the index of the nodes and the remote information for ghost layer nodes.
+The nodes structure is initialized by calling
+\begin{Verbatim}[frame=single]
+p4est_nodes_t *nodes = my_p4est_nodes_new(p4est, ghost);
+\end{Verbatim}
+The ghost information is optional and can be replaced by \ttt{NULL} if the ghost layer is not used.
+In the original \ttt{p4est} library, the nodes are organized differently than in \ttt{casl\_p4est}.
+That is, the ghost nodes for lower ranks are located at the beginning of the nodes array, the ghost nodes for higher ranks are at the end, and the local nodes are in between, starting at \ttt{nodes$\rightarrow$offset\_owned\_indeps}.
+Furthermore, the hanging nodes are indexed separately.
+
+In \ttt{casl\_p4est}, the nodes are reorganized to match the quadrants organization, thus simplifying the coupling with \ttt{Petsc}.
+
 \subsection{Communications with \texttt{Petsc}}
+
+\subsection{Finite differences and local tree reconstruction}
 

### 2016-04-05 doc/casl_p4est_manual/casl_p4est_manual.tex
updating manual

@@
 
+Most of the communication can be handled with the \ttt{Vec} data type provided by Petsc.
+The \ttt{Petsc} library is thoroughly documented online.
+A \ttt{Vec} is an array that can have a ghost chunk.
+A ghosted array can be created with the data located at the quadrants' center or at the nodes.
+The nodes code is
+\begin{Verbatim}[frame=single]
+Vec phi;
+VecCreateGhostNodes(p4est, nodes, &phi);
+\end{Verbatim}
+The actual data stored in the \ttt{Vec} can be accessed by
+\begin{Verbatim}[frame=single]
+double *phi;
+VecGetArray(phi, &phi_p);
+for (p4est_locidx_t n = 0; n<nodes->num_owned_indeps; ++n)
+{
+  (do something with phi_p)
+}
+VecRestoreArray(phi, &phi_p);
+\end{Verbatim}
+In the previous code, the pointer \ttt{phi\_p} points to a contiguous array of the local part of the data stored in the \ttt{Vec}.
+\ttt{Petsc} does not guarantee that the local part and the ghosted part are stored in the same location, and therefore accessing the entire data - including the ghosted part - should be done after getting the local form of the \ttt{Vec}
+\begin{Verbatim}[frame=single]
+Vec phi_loc;
+double *phi_p;
+VecGhostGetLocalForm(phi, &phi_loc);
+VecGetArray(phi_loc, &phi_p);
+for(size_t n=0; n<nodes->indep_nodes.elem_count; ++n)
+{
+  (do something with phi_p)
+}
+VecRestoreArray(phi_loc, &phi_p);
+VecGhostRestoreLocalForm(phi, &phi_loc);
+\end{Verbatim}
+In many places in the library, the ghosted values are accessed without first accessing the local form of the \ttt{Vec}, this is a dangerous practice that has worked ... so far.
+
+The ghost values can be synchronized with the remote processes through the \ttt{Petsc} API, thus leading to the general communication layout
+\begin{Verbatim}[frame=single]
+{ do work for local values shared with other processes }
+VecGhostUpdateBegin(phi, INSERT_VALUES, SCATTER_FORWARD);
+{ do work for local values not shared with other processes }
+VecGhostUpdateEnd(phi, INSERT_VALUES, SCATTER_FORWARD);
+\end{Verbatim}
+
 \subsection{Finite differences and local tree reconstruction}

### 2016-04-05 examples/poisson_jump_voronoi/main_jump_2d.cpp

@@ -48,9 +48,9 @@
 
 using namespace std;
 
-int lmin = 0;
-int lmax = 0;
-int nb_splits = 1;
+int lmin = 4;
+int lmax = 6;
+int nb_splits = 4;
 
 int nx = 2;
 int ny = 2;
@@ -58,10 +58,10 @@ int ny = 2;
 int nz = 2;
 #endif
 
-double xmin = 0;
+double xmin = -1;
 double xmax = 3;
-double ymin = 0;
-double ymax = 3;
+double ymin = -1;
+double ymax = 4;
 #ifdef P4_TO_P8
 double zmin = 0;
 double zmax = 3;

### 2016-04-05 src/my_p4est_poisson_node_base_jump.cpp
comet c++11 compatibility

@@ -1139,13 +1139,7 @@ void PoissonSolverNodeBaseJump::setup_linear_system()
   double *rhs_p;
   ierr = VecGetArray(rhs, &rhs_p); CHKERRXX(ierr);
 
-  typedef struct entry
-  {
-    double val;
-    PetscInt n;
-  } entry_t;
-
-  std::vector< std::vector<entry_t> > matrix_entries(num_local_voro);
+  std::vector< std::vector<mat_entry_t> > matrix_entries(num_local_voro);
   std::vector<PetscInt> d_nnz(num_local_voro, 1), o_nnz(num_local_voro, 0);
 
   for(unsigned int n=0; n<num_local_voro; ++n)
@@ -1167,7 +1161,7 @@ void PoissonSolverNodeBaseJump::setup_linear_system()
 #endif
     {
       matrix_has_nullspace = false;
-      entry_t ent; ent.n = global_n_idx; ent.val = 1;
+      mat_entry_t ent; ent.n = global_n_idx; ent.val = 1;
       matrix_entries[n].push_back(ent);
 #ifdef P4_TO_P8
       rhs_p[n] = bc->wallValue(pc.x, pc.y, pc.z);
@@ -1231,7 +1225,7 @@ void PoissonSolverNodeBaseJump::setup_linear_system()
 #endif
     if(add_n>EPS) matrix_has_nullspace = false;
 
-    entry_t ent; ent.n = global_n_idx; ent.val = voro.get_volume()*add_n;
+    mat_entry_t ent; ent.n = global_n_idx; ent.val = voro.get_volume()*add_n;
     matrix_entries[n].push_back(ent);
 
     for(unsigned int l=0; l<points->size(); ++l)
@@ -1278,7 +1272,7 @@ void PoissonSolverNodeBaseJump::setup_linear_system()
           o_nnz[n]++;
         }
 
-        entry_t ent; ent.n = global_l_idx; ent.val = -s*mu_harmonic/d;
+        mat_entry_t ent; ent.n = global_l_idx; ent.val = -s*mu_harmonic/d;
         matrix_entries[n][0].val += s*mu_harmonic/d;
         matrix_entries[n].push_back(ent);
 

### 2016-04-05 src/my_p4est_poisson_node_base_jump.h

@@ -21,13 +21,19 @@
 
 class PoissonSolverNodeBaseJump
 {
-  typedef struct check_comm
+  typedef struct
+  {
+    double val;
+    PetscInt n;
+  } mat_entry_t;
+
+  typedef struct
   {
     unsigned int n;
     unsigned int k;
   } check_comm_t;
 
-  typedef struct voro_comm
+  typedef struct
   {
     p4est_locidx_t local_num;
     double x;
@@ -37,7 +43,7 @@ class PoissonSolverNodeBaseJump
 #endif
   } voro_comm_t;
 
-  typedef struct added_point
+  typedef struct
   {
     double x;
     double y;

### 2016-04-28 src/my_p4est_bialloy.cpp
compatibility with latest changes / merges

@@ -44,7 +44,7 @@ my_p4est_bialloy_t::my_p4est_bialloy_t(my_p4est_node_neighbors_t *ngbd)
   epsilon_c            = 2.7207e-5; /* cm.K     */
   epsilon_v            = 2.27e-2;   /* s.K.cm-1 */
 
-  dx_dy_dz(p4est, dxyz);
+  ::dxyz_min(p4est, dxyz);
 #ifdef P4_TO_P8
   dxyz_min = MIN(dxyz[0],dxyz[1],dxyz[2]);
   dxyz_max = MAX(dxyz[0],dxyz[1],dxyz[2]);

### 2016-04-28 src/my_p4est_epitaxy.cpp

@@ -67,7 +67,7 @@ my_p4est_epitaxy_t::my_p4est_epitaxy_t(my_p4est_node_neighbors_t *ngbd)
   Nuc = 0;
   new_island = 0;
   alpha = 1.05;
-  dx_dy_dz(p4est, dxyz);
+  dxyz_min(p4est, dxyz);
 //  dt_n = MIN(dxyz[0],dxyz[1]);
   dt_n = 1e-3;
 

### 2016-04-28 src/my_p4est_node_neighbors.h
starting on multialloy ... first step is to have a poisson solver for neumann/robin with voronoi strategy to avoid small control volumes

@@ -25,6 +25,7 @@
 class my_p4est_node_neighbors_t {
   friend class my_p4est_poisson_nodes_t;
   friend class my_p4est_poisson_cells_t;
+  friend class my_p4est_poisson_nodes_voronoi_t;
   friend class PoissonSolverNodeBaseJump;
   friend class my_p4est_interpolation_t;
   friend class my_p4est_interpolation_nodes_t;

### 2016-06-20 examples/poisson_jump_voronoi/main_jump_2d.cpp
voronoi solver seems to work in parallel. Merging into master for v0.1

@@ -648,8 +648,7 @@ void save_VTK(p4est_t *p4est, p4est_ghost_t *ghost, p4est_nodes_t *nodes, my_p4e
 
   std::ostringstream oss;
 
-  oss << out_dir
-      << "/vtu/jump_"
+  oss << out_dir << "/jump_"
       << p4est->mpisize << "_"
       << brick->nxyztrees[0] << "x"
       << brick->nxyztrees[1] <<
@@ -889,7 +888,7 @@ void solve_Poisson_Jump( p4est_t *p4est, p4est_nodes_t *nodes,
 
     if(save_voro)
     {
-      snprintf(out_path,1000, "%s/vtu/voronoi", out_dir);
+      snprintf(out_path,1000, "%s/voronoi", out_dir);
       solver.print_voronoi_VTK(out_path);
     }
   }

### 2016-06-20 src/my_p4est_poisson_jump_nodes_voronoi.cpp

@@ -1626,7 +1626,8 @@ double my_p4est_poisson_jump_nodes_voronoi_t::interpolate_solution_from_voronoi_
     }
   }
 #endif
-      std::cerr << "my_p4est_poisson_jump_nodes_voronoi_t->interpolate_solution_from_voronoi_to_tree: point is double !" << std::endl;
+  // FIXME: What is this?
+  // std::cerr << "my_p4est_poisson_jump_nodes_voronoi_t->interpolate_solution_from_voronoi_to_tree: point is double !" << std::endl;
 
   /* now find the two voronoi points closest to the node */
   double phi_n = phi_p[n];

### 2016-06-20 doc/paper/tex/draft.tex

\documentclass[final]{elsarticle}
%\usepackage[pdftex,breaklinks,linktocpage,pagebackref,hyperindex,hyperfigures]
\usepackage{hyperref}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage[tmargin=1in,bmargin=1in,lmargin=1in,rmargin=1in]{geometry}
\usepackage{subfigure}
\usepackage{graphicx}
\usepackage{latexsym}
\usepackage{pdfsync}
% \usepackage[boxed]{algorithm}
\usepackage{algpseudocode}
\usepackage{algorithm}
% \usepackage{algorithmic}
\usepackage{multirow}
\usepackage{rotating}
\usepackage{color}
\usepackage{caption}
\usepackage{url}
\usepackage{tikz}
\usepackage{diagbox}
\usepackage{kpfonts}

\DeclareSymbolFont{extraup}{U}{zavm}{m}{n}
\DeclareMathSymbol{\varheart}{\mathalpha}{extraup}{86}
\DeclareMathSymbol{\vardiamond}{\mathalpha}{extraup}{87}

\hypersetup{
    colorlinks=true,      		% false: boxed links; true: colored links
    linkcolor=blue,       		% color of internal links
    citecolor=blue,       		% color of links to bibliography
    filecolor=black,      		% color of file links
    urlcolor=red,       		% color of external links
    pdffitwindow=true,
    pdfpagelayout=SinglePage
}

% defining some new commands
\newcommand{\etal}{\textit{et al.\ }}
\newcommand{\mbf}[1]{\mathbf{#1}}
\newcommand{\comment}[1]{\textcolor{red}{{#1}}}
\newcommand{\mohammad}[1]{\textcolor{blue}{{Mohammad: #1}}}
\newcommand{\tensor}[1]{\underline{\underline{\boldsymbol{#1}}}}
\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\ben}{\begin{equation*}}
\newcommand{\een}{\end{equation*}}
\newcommand{\bea}{\begin{eqnarray}}
\newcommand{\eea}{\end{eqnarray}}
\newcommand{\bean}{\begin{eqnarray*}}
\newcommand{\eean}{\end{eqnarray*}}
\newcommand{\pd}[2]{\frac{\partial #1}{\partial #2}}
\newcommand{\pdn}[3]{\frac{\partial^#1 #2}{\partial #3^#1}}
\newcommand{\noi}{\noindent}
\newcommand{\ind}{\indent}
\newcommand{\divg}[1]{\nabla \cdot \left(#1\right)}
\newcommand{\ddivg}[1]{\nabla \cdot #1}
\newcommand{\non}{\nonumber}
\newcommand{\fcite}[1]{(\textcolor{blue}{cite: #1})}
% \newcommand{\drawcircle}[2]{\protect  \tikz \protect \node[circle,draw,scale=#1,fill=#2] {};}
% \newcommand{\drawsquare}[2]{\protect  \tikz \protect \node[rectangle,draw,scale=#1,fill=#2] {};}
% \newcommand{\drawdiamond}[2]{\protect \tikz \protect \node[rectangle,draw,scale=#1,fill=#2,rotate=45] {};}
\newcommand{\drawcircle}[1]{\textcolor{#1}{$\bullet$}}
\newcommand{\drawsquare}[1]{\textcolor{#1}{$\blacksquare$}}
\newcommand{\drawdiamond}[1]{\textcolor{#1}{$\vardiamond$}}
\newtheorem{property}{Property}

\begin{document}

\title{Parallel Level-Set Methods on Adaptive Tree-Based Grids}

\cortext[cor]{Corresponding author: m.mirzadeh@engineering.ucsb.edu}
\address[MECHE]{Department of Mechanical Engineering, University of California, Santa Barbara, CA 93106-5070, United States.}
\address[CS]{Department of Computer Science, University of California, Santa Barbara, CA 93106-5110, United States.}
\address[UBONN]{Institute for Numerical Simulation, University of Bonn, 53115 Bonn, Germany.}

\author[MECHE]{Mohammad Mirzadeh\corref{cor}} \author[MECHE]{Arthur Guittet} \author[UBONN] {Carsten Burstedde} \author[MECHE,CS]{Frederic Gibou}

\input{abstract}
\input{introduction}
\input{levelset_method}
\input{parallel_algorithms}
\input{scaling_results}
\input{applications}
\input{conclusion}

\section*{References}
\bibliographystyle{plain}
\addcontentsline{toc}{section}{\refname}
\bibliography{casl,refs,bonn}

\end{document}

### 2016-06-20 doc/paper/tex/parallel_algorithms.tex

--- 2016-02-15 paper/tex/parallel_algorithms.tex
@@
 \begin{algorithm}[htbp]
-\caption{$H \gets \texttt{Reconstruct (}G\texttt{)}$: 
-Construction of the local tree hierarchy, $H$, from the parallel grid, $G$, supplied by \texttt{p4est}.
-The algorithm starts at the coarsest level, i.e.
-the macromesh, and recursively splits the each cell to match the finest local and ghost cells generated by \texttt{p4est}.}
+\caption{$H \gets \texttt{Reconstruct (}G\texttt{)}$: construction of the local tree hierarchy $H$ from the grid $G$}
 \begin{algorithmic}[1]
-\State $H \gets G.\texttt{macromesh()}$ \Comment{start with the coarsest  grid}
-\For {$\mathit{tr} \ein G.\texttt{local\_trees()}$} \Comment{update the hierarchy to match local cells}
-	\For {$c \ein \mathit{tr}.\texttt{local\_cells()}$}
+\State $H \gets G.\texttt{macromesh()}$
+\For {$\mathit{tr} : G.\texttt{local\_trees()}$} \Comment{build hierarchy for local cells}
+	\For {$c : \mathit{tr}.\texttt{cells()}$}
 		\State $H.\texttt{update\_tree(} \mathit{tr}, c \texttt{)}$
@@
 
-\For {$c \ein G.\texttt{ghost\_cells()}$} \Comment{update the hierarchy to mach local cells}
-	\State $H.\texttt{update\_tree(} c.\texttt{get\_parent\_tree()}, c \texttt{)}$
+\For {$c : G.\texttt{ghost\_cells()}$} \Comment{build hierarchy for ghost cells}
+	\State $H.\texttt{update\_tree(} c.\texttt{tree()}, c \texttt{)}$
 \EndFor
@@
 \Function{$H.\texttt{update\_tree}$}{$\mathit{tr}, c$}   \Comment{recursive tree reconstruction}
-	\State $c_l \gets H.\texttt{get\_root\_cell(}\mathit{tr}\texttt{)}$ 
-	\While{$c_l.\texttt{level()} \not= c.\texttt{level()} $} \Comment{recursively search for a cell of the same size as \textit{c}}
-		\If {$c_l.$\texttt{is\_leaf()}} $c_l$\texttt{.split()} \Comment{if the current cell is leaf, split the cell and continue search}
+	\State $c_l \gets H.\texttt{root(}\mathit{tr}\texttt{)}$
+	\While{$c_l.\texttt{level()} \not= c.\texttt{level()} $}
+		\If {$c_l.$\texttt{is\_leaf()}} $c_l$\texttt{.split()}
 		\EndIf
-		\State $h \gets c_l.\texttt{length()}$ \Comment{select the search path based on cell coordinate.}
-		\State $i \gets c.x \ge c_l.x + h/2$	
-		\State $j \gets c.y \ge c_l.y + h/2$
-		\State $k \gets c.z \ge c_l.z + h/2$
+		\State $h \gets c_l.\texttt{length()} / 2$ \Comment{select the next child based on direction}
+		\State $i \gets c.x \ge c_l.x + h$
+		\State $j \gets c.y \ge c_l.y + h$
+		\State $k \gets c.z \ge c_l.z + h$
 
-		\State $c_l \gets c_l.\texttt{get\_child(} i, j, k \texttt{)}$
+		\State $c_l \gets c_l.\texttt{child(} i, j, k \texttt{)}$
 	\EndWhile	
@@
 
-\subsection{Interpolation and semi-Lagrangian methods} \label{sec:interpolation_semilagrangian}
+\subsection{Interpolation and semi-Lagrangian methods}
 As indicated earlier, we use the semi-Lagrangian method to solve equation \eqref{eq:ls} when the velocity field is externally generated, i.e.\ when it does not depend explicitly on the level-set function itself.
@@
 \begin{algorithm}[htbp]
-\caption{$values \gets \texttt{Interpolate (}H, F, \underline{\mathbf{X}}\texttt{)}$: interpolate the value of $F$, defined on the local tree hierarchy $H$, at coordinates $\underline{\mathbf{X}}$.
-This is achieved in three phases as detailed in the text.
-I) First, interpolation points are buffered in two arrays, those than can be interpolated locally and those that should be sent to other processes.
-II) In the second phase, each process initiates the remote buffer exchange and computes the number of messages it should expect to receive using the communication matrix idea described in section \ref{sec:interpolation_semilagrangian}.
-III) Finally, each process alternates between computing local interpolation and receiving and performing remote interpolation requests from other processes.}
+\caption{$values \gets \texttt{Interpolate (}H, F, \underline{\mathbf{X}}\texttt{)}$: interpolate the value of $F$, defined on the local tree hierarchy $H$, at coordinates $\underline{\mathbf{X}}$}
 \begin{algorithmic}[1]
-\State $col \gets 0, \mathit{local\_buff} \gets \mathit{null}, \mathit{remote\_buff} \gets \mathit{null}$ \Comment{Phase I -- buffering}
-\For {$\underline{\mathbf{p}} \ein \underline{\mathbf{X}}$} 
-	\State $[owners\_rank, \mathit{cell}] \gets H.\texttt{search(}\underline{\mathbf{p}}\texttt{)}$ \Comment{search for the owner's rank and cell}
-	\If {$owners\_rank = \mathit{mpirank}$} \Comment{can interpolate locally}
-        \State $\mathit{local\_buff}[owners\_rank].\texttt{push\_back(}\underline{\mathbf{p}}, cell\texttt{)}$ 
-	\Else \Comment{requires remote interpolation}
-        \State $\mathit{remote\_buff}[owners\_rank].\texttt{push\_back(}\underline{\mathbf{p}}\texttt{)}$ 
-        \State $\mathit{col}[owners\_rank] \gets 1$
+\State $col \gets 0, \mathit{buff} \gets \mathit{null}$ \Comment{Phase I -- buffering}
+\For {$\underline{\mathbf{p}} : \underline{\mathbf{X}}$} 
+	\State $[r, \mathit{cell})] \gets H.\texttt{search(}\underline{\mathbf{p}}\texttt{)}$ \Comment{search for the owner's rank and cell}
+	\If {$r = \mathit{mpirank}$}
+        \State $\mathit{buff}[r].\texttt{push\_back(}\underline{\mathbf{p}}, cell\texttt{)}$
+	\Else
+        \State $\mathit{buff}[r].\texttt{push\_back(}\underline{\mathbf{p}}\texttt{)}$
+        \State $\mathit{col}[r] \gets 1$
 	\EndIf
 \EndFor
-\\
-\For {$r \ein \mathit{[0,mpisize)}$} \Comment{Phase II -- initiate communication and compute number of messages}
-	\If {$\mathit{col}[r] \deq 1$} 
-        \State $req \gets \texttt{MP\_Isend(}r, \texttt{QUERY\_TAG}, \mathit{remote\_buff}[r]\texttt{)}$ \Comment{initiate a non-blocking send to process \textit{r}}
-        	\State $\textit{query\_requests}.\texttt{push\_back(}req\texttt{)}$
+\For {$r:\mathit{mpisize}$} \Comment{Phase II -- initiate communication and compute number of messages}
+	\If {$\mathit{col}[r]$}
+        \State $\texttt{MP\_Isend(}r, \mathit{buff}[r]\texttt{)}$
 	\EndIf
 \EndFor
-\State  $S \gets \texttt{sum(}col\texttt{)}$ \Comment{compute number of messages to send}
-\State  $R \gets \texttt{MPI\_Reduce\_scatter(}\mathit{col},\texttt{MPI\_SUM)}$ \Comment{compute number of messages to receive}
-\\
-\State $\mathit{done} \gets \mathit{false}$ \Comment{Phase III -- interpolation}
-\State $it \gets \mathit{local\_buff}[mpirank].\texttt{begin()}$ 
-\While {$\mathit{!done}$} 
+\State  $S \gets \texttt{sum(}col\texttt{)}$ 
+\State  $R \gets \texttt{MPI\_Reduce\_scatter(}\mathit{col},\texttt{MPI\_SUM)}$
+\State $\mathit{done} \gets \mathit{false}, it \gets \mathit{buff}[mpirank].\texttt{begin()}$
+\While {$\mathit{!done}$} \Comment{Phase III -- main loop}
 	\If {$\mathit{it} \not= \mathit{buff[mpirank]}.\texttt{end()}$}
-		\State $\mathit{values} \gets \texttt{process\_local\_interpolation(}it\texttt{)}$ \Comment{process local interpolations}
+		\State $\mathit{values} \gets \texttt{process\_local(}it\texttt{)}$ \Comment{process local interpolations}
 		\State $\texttt{++}\mathit{it}$
 	\EndIf
-	\If {$R > 0$} \Comment{search for interpolation query in the message queue}
-		\State $\mathit{message} \gets \texttt{MPI\_Iprobe(\texttt{MPI\_ANY\_SOURCE}, \texttt{QUERY\_TAG})}$
-		\If {$\mathit{message}.\texttt{is\_pending()}$}
-			\State $\mathit{values} \gets \texttt{process\_remote\_queries(}\mathit{message}\texttt{)}$ \Comment{receive, search, and interpolate values}
-			\State $req \gets \texttt{MPI\_Isend(}\mathit{message}.\texttt{MPI\_SOURCE}, \texttt{REPLY\_TAG}, \mathit{values}\texttt{)}$ \Comment{send back interpolated values}
-			\State $\textit{reply\_requests}.\texttt{push\_back(}req\texttt{)}$
+	\If {$R > 0$} \Comment{process queries sent from remote processes}
+		\State $[\mathit{msg, st}] \gets \texttt{MPI\_Iprobe()}$
+		\If {$\mathit{msg}$}
+			% \State $values \gets \texttt{MPI\_Recv(}st.\texttt{MPI\_SOURCE)}$ \Comment{}
+			\State $\mathit{val\_buff} \gets \texttt{process\_queries(}\mathit{st}\texttt{)}$ \Comment{receive, search, and interpolate values}
+			\State $\texttt{MPI\_Isend(}\mathit{st}.\texttt{MPI\_SOURCE},\mathit{val\_buff}\texttt{)}$ \Comment{send back interpolated values}
 			\State $R\texttt{--}$
@@
 	\EndIf
-	\If {$S > 0$} \Comment{search for interpolation reply in the message queue}
-		\State $\mathit{message} \gets \texttt{MPI\_Iprobe(\texttt{MPI\_ANY\_SOURCE}, \texttt{REPLY\_TAG})}$
-		\If {$\mathit{message}.\texttt{is\_pending()}$}
-			\State $\mathit{values} \gets \texttt{process\_replies(}\mathit{message}\texttt{)}$ \Comment{receive remotely interpolated values}
+	\If {$S > 0$} \Comment{process replies sent to our queries}
+		\State $\mathit{[msg, st]} \gets \texttt{MPI\_Iprobe()}$
+		\If {$\mathit{msg}$}
+			\State $\mathit{values} \gets \texttt{process\_replies(}\mathit{st}.\texttt{MPI\_SOURCE)}$ \Comment{receive remotely interpolated values}
 			\State $S\texttt{--}$
@@
 	\EndIf
-	\State $\mathit{done} \gets S \deq 0 \And R \deq 0 \And \mathit{it \deq local\_buff[mpirank]}.\texttt{end()}$
+	\State $\mathit{done} \gets S = 0 \: \And \: R = 0 \: \And \: \mathit{it = buff[mpirank]}.\texttt{end()}$
 \EndWhile
-\State $\texttt{MPI\_Waitall(}query\_requests, reply\_requests\texttt{)}$ \Comment{make sure all messages have been received}
 \State \Return $\mathit{values}$
@@
 \State $\Delta t_l \gets \text{CFL} \times G^n.\texttt{hmin()} / \max \{\underline{\mathbf{u}}^n \} $
-\State $\Delta t   \gets \texttt{MPI\_Allreduce(}\Delta t_l,\texttt{MPI\_MIN)}$ \Comment{compute $\Delta t$ based on CFL condition across all processes}
-\State $H^n \gets \texttt{Reconstruct(} G^n\texttt{)}$  \Comment{using algorithm \ref{alg:reconstruction}}
+\State $\Delta t   \gets \texttt{MPI\_Allreduce(}\Delta t_l,\texttt{MPI\_MIN)}$
+\State $H^n \gets \texttt{Reconstruct(} G^n\texttt{)}$ 
 \State $G^{n+1}_0 \gets G^n$
@@
 	\State $\underline{\mathbf{X}}_d  \gets \texttt{ComputeDeparturePoints(}G^{n+1}_0, \underline{\mathbf{u}}^n, \underline{\mathbf{u}}^{n-1}, \Delta t\texttt{)}$ \Comment{using equations \ref{eq:xstar} -- \ref{eq:vn_p_half}}
-	\State $\phi^{n+1} \gets \texttt{Interpolate(}H^n, \phi^n, \underline{\textbf{X}}_d\texttt{)}$ \Comment{using algorithm \ref{alg:interpolation}}
+	\State $\phi^{n+1} \gets \texttt{Interpolate(}H^n, \phi^n, \underline{\textbf{X}}_d\texttt{)}$
 	\State $G^{n+1} \gets G^{n+1}_0.\texttt{refine\_and\_coarsen(}\phi^{n+1}\texttt{)}$ \Comment{using equations \ref{eq:refine} and \ref{eq:coarsen} as criteria}
@@
 Similar to \cite{Min;Gibou:07:A-second-order-accur}, equation \eqref{eq:semidiscrete_reinit} is integrated in time using the TVD-RK2 scheme with adaptive time-stepping in order to accelerate the convergence to the steady state.
-Generally it has been observed that adaptive time-stepping considerably improves the convergence rate and only a few iterations are needed if the signed-distance property is desired in a small band around the interface \cite{Min;Gibou:07:A-second-order-accur}.
-In this work, and based on previous findings, we use a fixed number of 20 iterations for the reinitialization equation.
-Of course, it should be noted that it is quite easy to define a custom tolerance as the termination criteria.
-
 Since the computation is based on a local stencil, the parallel implementation of this scheme is mostly trivial.
@@
 \end{algorithm}
-
-\subsection{Accuracy} \label{sec:accuracy}
-
-The numerical methods detailed in the previous sections are widely used in the level-set community and their accuracy is studied for example in \cite{Min;Gibou:07:A-second-order-accur}.
-However, in order to validate our implementation, we present a brief convergence analysis.
-
-The advection of an irregular boundary using the semi-Lagrangian method is the perfect candidate to demonstrate the accuracy of our implementation as it makes use of the interpolation routine as well as the reinitialization procedure.
-We select the benchmark problem proposed in \cite{Enright;Fedkiw;Ferziger;etal:02:A-Hybrid-Particle-Le}.
-Consider a sphere centered at $(0.35,0.35,0.35)$ and with radius $0.15$ in a domain $[0,1]^3$ and deformed under the divergence free velocity field
-\be \label{eq:velo_enright}
-\begin{cases}
-u(x,y,z) = 2 \sin^2(\pi x)\sin(2 \pi y)\sin(2 \pi z), \\
-v(x,y,z) = - \sin^2(\pi y)\sin(2 \pi x)\sin(2 \pi z), \\
-w(x,y,z) = - \sin^2(\pi z)\sin(2 \pi x)\sin(2 \pi y).
-\end{cases}
-\ee
-forward in time until $t_{1/2}=1$ and then backward to its original state at $t_f=2$.
-With set the time step to $\Delta t = 5 \Delta x_{min}$, where $\Delta x_{min}$ is the size of the smallest cell in the forest.
-We monitor the volume loss and the error in the interface location at the final time, when the original shape should be recovered, as the finest resolution of the forest increases.
-The results are reported in table \ref{tab:accuracy} and figure \ref{fig:accuracy} shows a visualization of the sphere at the initial time, at $t_{1/2}$ when the deformation is maximal, and at the final time.
-The results are consistent with those reported previously in \cite{Min;Gibou:07:A-second-order-accur} and indicate second order accuracy for the mass loss.
-
-\begin{table}[htbp]
-\begin{center}
-\begin{tabular}{|c|c|c|c|c|}
-\hline
-Finest resolution & $L^\infty$ error on $\phi$ & Rate & Volume loss (\%) & Rate \\
-\hline
-$128^3$  & 1.53E-01 & -    & 1.85E-01 & -    \\
-$256^3$  & 1.18E-01 & 0.38 & 3.73E-02 & 2.31 \\
-$512^3$  & 9.82E-03 & 3.58 & 6.87E-03 & 2.44 \\
-$1024^3$ & 2.56E-03 & 1.94 & 1.92E-03 & 1.84 \\
-$2048^3$ & 1.60E-03 & 0.67 & 5.17E-04 & 1.89 \\
-\hline
-\end{tabular}
-\caption{Study of the convergence of the level-set algorithm using Enright's test \cite{Enright;Fedkiw;Ferziger;etal:02:A-Hybrid-Particle-Le}.
-The $L^\infty$ error on $\phi$ is computed only close to the interface as this is the relevant observable for an advection procedure.} \label{tab:accuracy}
-\end{center}
-\end{table}
-
-\begin{figure}[htbp]
-\begin{center}
-\begin{tabular}{ccccc}
-\includegraphics[width=.28\textwidth]{figures/enright_0.png}
-&
-\hspace{.1cm}
-&
-\includegraphics[width=.28\textwidth]{figures/enright_1.png}
-&
-\hspace{.1cm}
-&
-\includegraphics[width=.28\textwidth]{figures/enright_2.png}
-\end{tabular}
-\caption{Visualization of the deformation undergone by the sphere for the Enright's test at time 0, $t_{1/2}$ and $t_f$ and colored with the processes ranks.
-The forest's finest resolution is equivalent to a uniform grid with $2048^3$ cells and contains between 17 (at initial and final times) and 63 (at $t_{1/2}$) million nodes.
-These results were obtained on the Comet supercomputer with 960 processes (40 compute nodes).} \label{fig:accuracy}
-\end{center}
-\end{figure}

### 2016-12-15 examples/biomol/AtomTree.cpp
commit before merge. Minor change.

@@ -381,7 +381,7 @@ void AtomTree::print_VTK_format( std::string file_name, double time ) const
     fprintf(outFile,"CELL_TYPES %d \n",num_of_leaf_cells);
     for (int n=0; n<num_of_leaf_cells; n++)
         fprintf(outFile,"%d \n",12);
-    fprintf(outFile,"POINT_DATA %d \n",nodes.size());
+    //fprintf(outFile,"POINT_DATA %d \n",nodes.size());
     fprintf(outFile,"CELL_DATA %d \n",num_of_leaf_cells);
     fclose (outFile);
 }

### 2016-12-15 examples/biomol/AtomTree.h

@@ -102,6 +102,7 @@ class AtomTree{
      }
     void set_probe_radius(double rp_);
     void set_max_atom_radius(double rmax_);
+    void print_atom_count_per_cell( std::string file_name) const;
 
 
 
@@ -163,7 +164,7 @@ private:
     void set_atoms_belonging_to_cell(ATCell &cell, const std::vector<Atom> &atoms);
     //double dist_from_surface(double x, double y, double z);
 
-    void print_atom_count_per_cell( std::string file_name) const;
+
 
     void print_VTK_format( std::string file_name, double time=DBL_MIN) const;
 

### 2016-12-15 surf_gen_paper/tex/fast_surf_gen_paper.tex
Really rough first draft of surface gen paper

\documentclass{elsarticle}
%\usepackage[pdftex,breaklinks,linktocpage,pagebackref,hyperindex,hyperfigures]
\usepackage{hyperref}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage[tmargin=1in,bmargin=1in]{geometry}
\usepackage{subfigure}
\usepackage{graphicx}
\usepackage{latexsym}
\usepackage{pdfsync}
\usepackage{caption}

\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{multirow}
\usepackage{rotating}
\usepackage{color}
\usepackage{ulem}
\usepackage{wrapfig}
\usepackage{siunitx}

\journal{Journal of Computational Physics}

\hypersetup{
    colorlinks=true,      		% false: boxed links; true: colored links
    linkcolor=blue,       		% color of internal links
    citecolor=blue,       		% color of links to bibliography
    filecolor=black,      		% color of file links
    urlcolor=red,       		% color of external links
    %bookmarks=false,
    pdffitwindow=true,
    pdfpagelayout=SinglePage
}

\bibliographystyle{plain}

\newtheoremstyle{own}%
    {1pt}% Space above
    {1pt}% Space below
    {}% Body font
    {}% Indent amount
    {\color{black}\bfseries}% Theorem head font
    {:}% Punctuation after theorem head
    {1pt}% Space after theorem head
    {}% Theorem head spec

\theoremstyle{own}
\newtheorem{example}{Example}

\newtheorem*{remark}{Remark}

\algblock[parallel for]{ParFor}{EndParFor}

\newcommand{\giga}{\si{\giga}}
\newcommand{\second}{\si{\second}}
\newcommand{\kilo}{\si{\kilo}}
\newcommand{\meter}{\si{\meter}}
\newcommand{\comment}[1]{\textcolor{red}{#1}}
\newcommand{\milescomment}[1]{\textcolor{blue}{#1}}
\newcommand{\chohongcomment}[1]{\textcolor{green}{#1}}

\begin{document}

% declaration of the new block
\algblock{ParFor}{EndParFor}
% customising the new block
\algnewcommand\algorithmicparfor{\textbf{parallel\ for}}
\algnewcommand\algorithmicpardo{\textbf{do}}
\algnewcommand\algorithmicendparfor{\textbf{end\ parallel\ for}}
\algrenewtext{ParFor}[1]{\algorithmicparfor\ #1\ \algorithmicpardo}
\algrenewtext{EndParFor}{\algorithmicendparfor}

\title{Fast surface generation for biomolecular computations with the Voronoi Interface Method}	
%\author{Miles Detrixhe, Chohong Min, Fr\'ed\'eric Gibou}

\author{Miles Detrixhe}
\ead{mdetrixhe@engineering.ucsb.edu}
\address{Department of Mechanical Engineering \\
University of California Santa Barbara, Santa Barbara, CA, 93106}

\author{Mohammad Mirzadeh}
\ead{}
\address{MIT
}

\author{Arthur Guittet}
\ead{}
\address{
}

\author{Fr\'ed\'eric Gibou}
\ead{fgibou@engineering.ucsb.edu}
\address{Department of Mechanical Engineering \\
Department of Computer Science \\
Department of Mathematics \\
University of California Santa Barbara, Santa Barbara, CA, 93106}

\begin{abstract}
We improve upon the state-of-the-art level set surface generation technique for biomolecules by employing a fixed radius nearest neighbor search algorithm. By choosing an appropriate data structure we show that the time complexity of this algorithm is independent of the number of atoms in a particular biomolecule.
\end{abstract}

\begin{keyword}
  Parallel computing\sep 
  Poisson-Boltzmann equation \sep
  octree data structure \sep
  level set
\end{keyword}

\maketitle

\section{Introduction}

\section{Poisson-Boltzmann equations}

\section{Biomolecular surface description}

\section{fast surface generation}

The method of surface generation used in \cite{MirzadehTheillardHelgadottirEtAl:12:AdaptiveFiniteDifference} was a brute force implementation of a modified version of Nearest Neighbor Search.

In the classic nearest neighbor search, given a query point, one aims to find the nearest element of a set of points. In computing the value of the level set function at a point, one aims to find the maximum value of the following function for all atoms in the molecule.

\begin{equation*}
\phi(x,y,z) = \max_i\left\lbrace r_i - \sqrt{(x-x_i)^2 + (y-y_i)^2 + (z-z_i)^2} \right\rbrace
\end{equation*}

Because the reinitialization procedure is used to propagate the solution of $\phi$ away from the interface, it is sufficient to have an accurate method of computing $\phi$ near the interface.

This reduces it to the fixed-radius nearest neighbor problem. This is an old problem (even for molecular applications \cite{Levinthal:66:Molecular-model-buil}) and there exist efficient ways of solving it e.g. \cite{Bentley:75:A-Survey-of-techniqu,Bentley;Stanat;Williams:77:The-complexity-of-fi}.

We use an octree data structure with each cell containing those atoms within a fixed radius of the center of the cell. The radius is related to the maximum atom radius plus the probe radius.

For instance, consider the 3J6D protien from the PDB database. It has 131664 individual atoms. If we store those atoms in an octree such that each leaf cell contains all the atoms within a fixed radius from the cell center, then we get the figure below. This is for an atom tree with max level 7. The maximum number of atoms to be queried is 622. If no atoms are with the fixed radius of a leaf cell, then we populate that cell with the nearest atom.

\begin{figure}[ht]
\begin{center}
\includegraphics[width=0.945\textwidth]{../figs/atom_tree_3J6D}
\end{center}
\caption{Figure showing the atom tree data structure. It is a quadtree with max level 7. } \label{fig:atom_tree_3J6D}
\end{figure}

\section{solving PB on octree - Voronoi Interface method}

\section{Numerical examples}

\begin{figure}[ht]
\begin{center}
\includegraphics[width=0.945\textwidth]{../figs/fast_gen_timing}
\end{center}
\caption{Surface generated using 10 different molecules from the PDB database. Exact, brute-force method used as a benchmark.} \label{fig:fast_gen_timing}
\end{figure}

\begin{figure}[ht]
\begin{center}
\includegraphics[width=0.945\textwidth]{../figs/timing_msms_compare}
\end{center}
\caption{Time to generate surface on various grid levels. MSMS timings used as a benchmark. This shows that for large atoms, this method is faster than MSMS even on relatively fine octrees (max level 11).} \label{fig:timing_msms_compare}
\end{figure}

\begin{figure}[ht]
\begin{center}
\includegraphics[width=0.945\textwidth]{../figs/error_convergence_1MAH}
\end{center}
\caption{Error in total computed surface area for the 1MAH protien level set method compared to MSMS. This shows that the surface area accuracy is competitive with MSMS for moderately refined grids (max level 10).} \label{fig:error_convergence_1MAH}
\end{figure}

\subsection{Fast surface generation}

MSMS \cite{Sanner;Olson;Spehner;etal:96:REDUCED-SURFACE:-an-}

%\begin{table}[h]
%\begin{center}
%   \begin{tabular}{| c |  c | c | c | c |}
%  \hline
%     PDB ID & \# Atoms & Analytical & Triangulated & Time msms \\ \hline
%	 1d65 &  760 & 3573.80 & 3405.72 & 0.07 \\ \hline
%	 6rxn & 1030 & 3827.86  & 4065.94  & 0.07 \\ \hline
%	 2err & 1638  & 4917.44  & 5189.28  & 0.10  \\ \hline
%	 1aa2 & 2145 & 5714.00 & 5389.36 & 0.10     \\ \hline
 %    2x6a & 4297  & 12969.68  & 13605.47  & 0.28  \\ \hline
%     3M3D &  8976 & 16626.72  & 17493.79   & 0.38  \\ \hline
%  	 1mah & 9133  & 19431.54  & 18509.91   & 0.83  \\ \hline
%	 1fss & 9360  & 19495.40  & 18581.67  & 0.41  \\ \hline
%     1OED & 10245   & 31013.07  & 32413.19  & 1.95 \\ \hline
%	 3ecd & 24749  & 47360.27  & 49791.05  & 2.89  \\ \hline
%	 1JB0 & 34720  &	93297.92  & 89219.35 & 5.70 \\ \hline
%	 3J6D & 131664   & 384569.69  & 403729.06  & 29.74 \\ \hline
%    \end{tabular}
%\end{center}
%        \caption{Molecule data for MSMS software.}
%    \label{tab:total}
%\end{table}

\subsection{Parallel scaling}

There is a bug causing crashes when I run this on large numbers of processors. I haven't been able to deduce it yet.

\subsection{Large example}

This is a preliminary example because of the crashing issue.

Preliminary results 3J6D molecule has 131664 atoms.

\begin{figure}[ht]
\begin{center}
\includegraphics[width=0.945\textwidth]{../figs/3J6D_cutaway_7_10}
\end{center}
\caption{A cutaway of the computational domain (colored by bulk potential) and the computed SES surface of the 3J6D protien colored by the electric potential on its surface.} \label{fig:3J6D_cutaway_7_!0}
\end{figure}

\begin{figure}[ht]
\begin{center}
\includegraphics[width=0.945\textwidth]{../figs/3J6D_zoom_7_10}
\end{center}
\caption{Zoomed in version} \label{fig:3J6D_zoom_7_10}
\end{figure}

\section{Conclusion}\label{sec::conclusion}

\bibliography{references}

\end{document}

### 2016-12-15 surf_gen_paper/MATLAB/msms_compare_timing.m

@@ -1,0 +1,91 @@
+clear all;
+
+
+lmin=[4,4,5,5,5];
+lmax=[7,8,9,10,11];
+mols=['6rxn'; '1d65'; '2err'; '1aa2'; '2x6a'; '3M3D'; '1mah'; '1fss'; '1OED'; '3ecd'; '1JB0'; '3J6D'];
+atoms=[667, 760, 1638, 1755, 4294, 8346, 9133, 9252, 10245, 23468, 34720, 131664];
+  
+
+times_fast = zeros(length(atoms), length(lmin));
+
+
+%%Data from msms_compare.7921000.txt
+
+times_fast(:,1)=[0.342, 0.518, 0.295, 0.473, 0.291, 0.406, 0.323, 0.419, 0.413, 0.358, 0.414, 0.327];
+nodes_fast(:,1)=[17510, 28855, 14701, 28491, 14174, 23451, 18263, 25324, 22540, 20233, 23486, 18438];
+times_fast(:,2)=[0.808, 1.619, 0.702, 1.432, 0.658, 1.119, 0.827, 1.229, 1.098, 0.937, 1.110, 0.815];
+nodes_fast(:,2)=[51880, 100808, 39530, 100898, 36538, 81040, 57253, 89915, 78741, 67744, 85724, 59042];
+times_fast(:,3)=[3.272	6.25	2.611	6.286	2.458	4.835	3.403	5.287	4.681	3.828	4.773	3.288];
+nodes_fast(:,3)=[232228	447596	174219	452635	161512	369077	260287	408499	395127	320408	445281	294663];
+times_fast(:,4)=[10.982	20.836	9.072	22.4	8.297	21.024	14.081	23.685	21.029	16.163	21.687	13.764];
+nodes_fast(:,4)=[879936	1735988	644392	2001660	580330	1582491	1060807	1771092	1813669	1361327	2117694	1465829];
+times_fast(:,5)=[39.998	77.044	32.771	81.446	32.793	79.384	59.217	89.297	88.526	78.001	115.716	73.956];
+nodes_fast(:,5)=[3493670 6806782	2771095	7648041	2575580	7998187	5610226	8985196	9257182	6514929	10240970	7766836];
+
+area_fast(:,1) = [3186.766	2194.445	4254.071	4293.911	10212.342	13562.505	14238.882	15040.804	21841.541	34223.003	61530.231	216033.342];
[…]

### 2019-05-21 examples/navier_stokes/main_shs_2d.cpp
beginning of accuracy_check

@@ -730,6 +730,50 @@ double Re_b(const double& mass_flow, const double &width, const double& rho, con
   return rho*mass_flow*1.0/(mu*2.0*width); // delta = 1.0 --> height = 2.0
 }
 
+#ifdef P4_TO_P8
+void check_accuracy_of_solution(my_p4est_navier_stokes_t* ns, const external_force_u_t& force_per_unit_mass_x, const double& pitch, const double& GF, const bool spanwise, double my_errors[])
+#else
+void check_accuracy_of_solution(my_p4est_navier_stokes_t* ns, const external_force_u_t& force_per_unit_mass_x, const double& pitch, const double& GF, double my_errors[])
+#endif
+{
+  const double rho        = ns->get_rho();
+  const double mu         = ns->get_mu();
+  const double length     = ns->get_length_of_domain();
+  const double height     = ns->get_height_of_domain();
+#ifdef P4_TO_P8
+  const double width      = ns->get_width_of_domain();
+#endif
+  const double abs_dp_dx  = force_per_unit_mass_x.get_value();
+  double my_infty_norm_error_u = 0.0;
+  double my_infty_norm_error_v = 0.0;
+#ifdef P4_TO_P8
+  double my_infty_norm_error_w = 0.0;
+#endif
+  // calculate the face errors here;
+
+  my_errors[0] = my_infty_norm_error_u;
+  my_errors[1] = my_infty_norm_error_v;
+#ifdef P4_TO_P8
+  my_errors[2] = my_infty_norm_error_w;
+#endif
+
+  my_infty_norm_error_u = 0.0;
+  my_infty_norm_error_v = 0.0;
+#ifdef P4_TO_P8
+  my_infty_norm_error_w = 0.0;
+#endif
+  // calculate the node errors here;
+
+
+  my_errors[P4EST_DIM+0] = my_infty_norm_error_u;
+  my_errors[P4EST_DIM+1] = my_infty_norm_error_v;
+#ifdef P4_TO_P8
+  my_errors[P4EST_DIM+2] = my_infty_norm_error_w;
+#endif
+  int mpiret = MPI_Allreduce(MPI_IN_PLACE, my_errors, 2*P4EST_DIM, MPI_DOUBLE, MPI_MAX, ns->get_p4est()->mpicomm); SC_CHECK_MPI(mpiret);
+}
+
+
 int main (int argc, char* argv[])
 {
   mpi_environment_t mpi;
@@ -783,6 +827,7 @@ int main (int argc, char* argv[])
[…]

### 2019-05-22 examples/navier_stokes/optimization/navier_stokes_optimization.tex

\documentclass[11pt, a4paper]{article}
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{placeins}
\usepackage{empheq}
\usepackage{hhline}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{amsthm}
\usepackage[binary-units=true]{siunitx}
\usepackage{mathrsfs}
\usepackage[multiple]{footmisc}
\usepackage{bbold}
\usepackage{lscape}
\usepackage{setspace}
\usepackage{a4wide}
\usepackage{color}
\usepackage{tikz, pgfplots}
\pgfplotsset{compat=1.9}
\usetikzlibrary{arrows.meta}
\usetikzlibrary{calc}
\usetikzlibrary{shapes,arrows,backgrounds}
\usepackage{algcompatible}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{varwidth}
\usepackage[inline]{enumitem}
\usepackage{dsfont}
\usepackage{fullpage}
\usepackage{multirow}
\usepackage[Symbol]{upgreek}
\usepackage{wrapfig}
\usepackage[pdfborder={0 0 0}]{hyperref}
\usepackage{url}
\usepackage{authblk}
\usepackage[normalem]{ulem}
\usepackage{hyperref}
\hypersetup{
    colorlinks=true,       % false: boxed links; true: colored links
    linkcolor=blue,       % color of internal links
    citecolor=blue,       % color of links to bibliography
    filecolor=black,       % color of file links
    urlcolor=black,         % color of external links
    pdffitwindow=true,
    pdfpagelayout=SinglePage
}
\usepackage{cite}

\newcommand{\Sha}{\mbox{\usefont{T2A}{\rmdefault}{m}{n}\CYRSH}}
\newcommand{\fig}{\figurename \ }
\makeatletter
\newcommand{\algo}{\ALG@name \ }
\makeatother
\newcommand{\tableau}{\tablename \ }
\newcommand{\red}[1]{\textcolor{red}{#1}}
\renewcommand{\vec}[1]{\boldsymbol{#1}}
\newcommand{\secondOrderTensor}[1]{\underline{\boldsymbol{#1}}}
\newcommand{\tensor}[1]{\underline{\underline{\boldsymbol{#1}}}}
\newcommand{\green}[1]{\textcolor{green}{#1}}
\newcommand{\blue}[1]{\textcolor{blue}{#1}}
\newcommand{\im}{\text{i}}
\newcommand{\Der}[2]{\frac{\mathrm{d} #1}{\mathrm{d} #2}}
\newcommand{\DDer}[2]{\frac{\mathrm{d^2} #1}{\mathrm{d} #2^2}}
\newcommand{\partDer}[2]{\frac{\partial #1}{\partial #2}}
\newcommand{\partder}[2]{\partial_{#2} #1}
\newcommand{\dist}[2]{\text{dist}\left( #1, #2 \right)}
\newcommand{\diag}[1]{\text{diag}\left( #1\right)}
\newcommand{\Reynolds}{\text{Re}}
\newcommand{\delete}[1]{\textcolor{red}{\sout{#1}}}
\newcommand{\bigOhNothation}[1]{\mathcal{O}\left( #1 \right)}
\newcommand{\realSet}{\mathds{R}}
\newcommand{\spaceOfRealVectors}[1]{\realSet^{#1 \times 1}} 
\newcommand{\spaceOfRealMatrices}[2]{\realSet^{#1 \times #2}}

\theoremstyle{remark}
\newtheorem{remark}{Remark}
\newcommand{\rem}{Remark \ }
\newtheorem{theorem}{Theorem}[section]
\newcommand{\thm}{Theorem \ }
\newtheorem{lemma}[theorem]{Lemma}
\newcommand{\lem}{Lemma \ }

\definecolor{vert}{rgb}{0.3,0.7,0}
\renewcommand{\green}[1]{\textcolor{vert}{#1}}
\DeclareMathOperator{\cotan}{cotan}
\DeclareMathOperator{\sinc}{sinc}
\DeclareMathOperator*{\argmax}{argmax}

\newcommand{\Mymp}{\mathbin{\smash{%
\raisebox{0.35ex}{%
            $\underset{\raisebox{0.5ex}{$\textcolor{vert}{\smash +}$}}{\smash-}$%
            }%
        }%
    } %
}

\newcommand{\Mypm}{\mathbin{\smash{%
\raisebox{0.35ex}{%
            $\underset{\raisebox{0.5ex}{$\textcolor{vert}{\smash -}$}}{\smash+}$%
            }%
        }%
    } %
}
\newcommand{\ohm}{\Upomega}
\newcommand{\rot}[1]{\nabla \times #1}
\newcommand{\divergence}[1]{\nabla \cdot #1}
\newcommand*{\I}{\imath}

\begin{document}
\renewcommand{\arraystretch}{1.1}

\title{Optimization of the Navier-Stokes Solver on Distributed Adaptive Quad-/Oc-tree Grids}

\author{
Raphael Egan
}

\maketitle

\begin{abstract}
The optimization-based modifications brought to the Navier-Stokes solver on distributed adaptive Quad-/Oc-tree are presented, summarized and explained. The modifications are validated by comparing the original and optimized solvers on the flow past a sphere/disk in 3D/2D with advection of a passive scalar. The gain of performance is reported, discussed and illustrated with respect to user-defined numerical parameters for some applications.
\end{abstract}

\section{Summary of the modifications}
\label{sec:summary}
The typical usage of the Navier-Stokes solver follows the flow chart from \fig \ref{fig:flow_chart}. The modifications brought to the solver target different parts of such a typical flow chart but can be divided into various categories depending on their scope of interest and/or on the typical subpart of the flowchart that they aim to optimize.

\begin{figure}
\centering
% Define block styles
\tikzstyle{decision} = [diamond, draw, fill=blue!20, 
    text width=4.5em, text badly centered, node distance=3cm, inner sep=0pt]
\tikzstyle{block} = [rectangle, draw, fill=blue!20, 
    text width=10em, text centered, rounded corners, minimum height=3em]
\tikzstyle{line} = [draw, -latex']
\tikzstyle{cloud} = [draw, ellipse,fill=red!20, node distance=5.5cm,
    text width=12em, text centered, minimum height=3em]
\begin{tikzpicture}[node distance = 2.5cm, auto]
    % Place nodes
    \node [block] (input) {read user's input(s)};
    \node [block, below of=input, text width=16em] (init) {initialize all solver's parameters to input(s) or default values.\\ Set $n = 0$, $t_{0} = t_{\text{start}}$, and $\Delta t_{0} = t_{1}-t_{0}$};
    \node [cloud, right of=init, node distance=7.5cm] (flags) {set exportation flags, exportation parameters, grid update flags, grid update parameters, etc.};
    \node [right of=flags, node distance=4.8cm] (top_right_anchor){};
    \node [decision, text width=6.em, below of=init] (loop_start) {is $t_{n}$ equal to $t_{\text{end}}$?};
    \node [decision, right of=loop_start, node distance=3.8cm] (first_step) {is $n>0$?};
    \node [block, right of=first_step, node distance=4.5cm] (get_new_dt) {get $\Delta t_{n} = t_{n+1}-t_{n}$};
    \node [block, below of=get_new_dt, node distance=2.cm] (grid_update) {update grid if desired.};
    \node [block, below of=first_step, node distance=3.5cm] (export_state) {export solver's state if desired and if requried};
    \node [block, below of=export_state, node distance=2.5cm] (viscosity_step) {solve for $\mathbf{v}^{\star}$ (BC $\sim \nabla \Phi_{n+1}$)};
    \node [block, below of=viscosity_step, node distance=2.5cm] (projection_step) {solve for $\Phi_{n+1}$ ($-\Delta \Phi_{n+1} = \divergence{\mathbf{v}^{\star}}$)};
    \node [decision, right of=projection_step, node distance=5.5cm, text width = 6em] (exit_sub_loop) {$\Phi_{n+1}$ converged or max \# subiterations reached?};
    \node [below of=projection_step, node distance=2.4cm](under_projection_step){};
    \node [left of=under_projection_step, node distance=8em, label=above right:{\red{Inner loop for convergence of $\Phi_{n+1}$}}] (lower_left_subloop_corner){};
    \node [above of=exit_sub_loop, node distance=3.6cm](above_exit_subloop){};
    \node [right of=above_exit_subloop, node distance=2.6cm] (upper_right_subloop_corner){};
    \begin{scope}[on background layer]
     \draw[black,fill=yellow!40,rounded corners=1ex] (lower_left_subloop_corner) rectangle (upper_right_subloop_corner);
    \end{scope}
    \node [block, below of=projection_step, node distance=3.5cm] (interp_velocity) {interpolate $\mathbf{v}_{n+1}$ at nodes,\\ calculate $p_{n+1}$};
    \node [block, below of=interp_velocity, node distance=2.8cm, text width=14em] (exportation) {if desired and if required, export relevant data. \\ Update $t_{n} \leftarrow t_{n} + \Delta t_{n}$ and $n \leftarrow n+1$};
    \node [block, left of=exportation, node distance=6.cm, text width=6em] (stop) {clear memory and stop};
    % Draw edges
    \path [line] (init) -- (loop_start);
    \path [line] (loop_start) -- node {no}(first_step);
    \path [line] (loop_start) -| node [above left] {yes}(stop);
    \path [line] (first_step) -- node {yes}(get_new_dt);
    \path [line] (get_new_dt) -- (grid_update);
    \path [line] (first_step) -- node {no}(export_state);
    \path [line] (grid_update) |- (export_state);
    \path [line] (export_state) -- (viscosity_step);
    \path [line] (viscosity_step) -- (projection_step);
    \path [line] (projection_step) -- (exit_sub_loop);
    \path [line] (exit_sub_loop) |- node [above left] {no}(viscosity_step);
    \path [line] (exit_sub_loop) |- node [below right] {yes}(interp_velocity);
    \path [line] (interp_velocity) -- (exportation);
    \path [line] (exportation) -| (loop_start);
    \path [line] (input) -- (init);
    \path [line,dashed] (input) -- (flags);
    \path [line,dashed] (flags) -- (top_right_anchor) |- (grid_update);
    \path [line,dashed] (flags) -- (top_right_anchor) |- (export_state);
    \path [line,dashed] (flags) -- (top_right_anchor) |- (exportation);
\end{tikzpicture}
\caption{\label{fig:flow_chart}typical flow chart for the usage of the one-phase Navier-Stokes solver.}
\end{figure}

\subsection{Modifications or additions of general interest}
\label{subsec:general_interest}
The following modifications and additions were made to the \verb|casl_p4est| library. They may be of general interest to all users.
\begin{itemize}
 \item Restart tools were developped to enable the possibility to export/load an exact given state defined as the association of a given \verb|p4est| object and a set of relevant node-, cell- or even face-sampled data fields. The state can be reloaded with a different number of MPI processes than when written on disk, and the relevant data is re-scattered to the new global orderings in a way that is consistent with the data layout when exported.
 \item The interpolations tools (from nodes, cell-centers or faces) were generalized to allow the interpolation of several fields \emph{simultaneously}, under the same desired interpolation scheme. This is especially relevant when interpolating several node-sampled fields from one grid to a unique set of desired nodes, with the same interpolation procedure: for instance, when setting node values of a newly updated grid based on node values of the former grid. Another relevant example is the interpolation of node-sampled vector fields, for which this approach can cut down the costs by the number of vector components (\textit{e.g.,} calculations of semi-lagrangian trajectories).\\
 Such an approach reduces the number of lookups in the hierarchy and, more importantly, cuts down the number of more costly operations per point in the set (like the calculations of non-initialized direct neighbors in quadratic node interpolation, or the resolution of least-square systems in cell or face cases) by the number of fields to be interpolated simultaneously. \\
 The user must be \textbf{careful} when using such a feature with cell-centered or face-centered fields as boundary conditions are required and queried in such a case: \emph{currently, this generalization assumes that all fields share the exact same boundary conditions for interpolation of cell- or face-sampled data}.
 \item A similar generalization was done for the calculation of second derivatives of several node-sampled fields simultaneously, in \verb|my_p4est_node_neighbors|.
 \item Basic core function of the \verb|matrix_t| class (solving least-square systems, calculations of Choleski factorizations, etc.) were modified accordingly to allow the resolution of linear systems involving possibly more than one right-hand-side vector.
 \item The classes \verb|my_p4est_faces_t|, \verb|my_p4est_hierarchy_t|, \verb|my_p4est_node_neighbors_t|, \break \verb|my_p4est_navier_stokes_t| were augmented with \verb|memory_estimate| functions that return an estimate of the raw number of bytes that the corresponding objects require. Similarly, a function was added to \verb|my_p4est_tools.c| in order to estimate the memory requirement of \verb|my_p4est_brick_t| objects. Except for \verb|my_p4est_navier_stokes_t| objects, these functions assess the memory requirement associated with every single MPI process and an \verb|MPI_allreduce| call is required afterwards to estimate the global memory requirement. The estimates measure the raw data requirements, hence they are not exact but still very representative. They were designed to make a distinction between the growing memory requirement due to an increasingly finer grid and a possible memory leak. The memory estimate for the \verb|my_p4est_nodes_t| data structures would be very helpful, however not straightforward to implemented (and not implemented yet in \verb|p4est|).
 \item Two functions \verb|nodes_are_equal| and \verb|ghosts_are_equal| were added to \verb|p4est_utils| to check equality between two \verb|p4est_nodes_t| and between two \verb|p4est_ghost_t| structures, respectively. The two functions check that every single item is identical, ordered in the same way and that the local and global indices are identical. The tests are processed locally, an \verb|MPI_allreduce| call is required afterwards to test equality globally.
 \item A function \verb|index_of_node| was developped and added to \verb|p4est_utils|. This function takes any node and its tree index as defined within \verb|p4est|\footnote{In \texttt{p4est}, a node is defined as a specific \texttt{p4est\_quadrant\_t} object whose \texttt{level} field is assigned the value \texttt{P4EST\_MAX\_LEVEL} (which is defined as \texttt{P4EST\_QMAX\_LEVEL+1}, hence the unambiguous distinction with any actual quadrant). The regular \emph{logical} coordinates \texttt{x}, \texttt{y}, \texttt{z} and the tree index of the considered node make its localization unambiguous and robust (no floating point arithmetics error is possible).}\textsuperscript{,}\footnote{To ensure that the binary search finds any given point that exists, the node and its tree index \emph{must be canonicalized} before being passed to \texttt{index\_of\_node}. The \texttt{static} qualifier of the \texttt{p4est\_node\_canonicalize} function from \texttt{my\_p4est\_nodes.c} was removed so that it can be used from another context. This ensures consistency between the considered node and the nodes as they were defined when the \texttt{p4est\_nodes\_t} data structure was assembled. It is \emph{critical} if the node lies exactly on a tree boundary and its owning tree is not clear. Even if only one tree is used, the node must be \emph{clamped} inside the domain.} to perform a binary search based on the tree and Morton indices that are used to sort the nodes internally by \verb|p4est_nodes_t|. If the given node actually exists in the current forest, the function returns a positive answer along with its local node index. If the considered node does not exist and/or does not belong to the \verb|p4est_nodes_t| data structure, a negative answer is returned.\\ 
 Hence, this function finds any existing node indexed by the local partition in $\mathcal{O}\left( \log\left( N \right) \right)$ operations without the need of constructing a local hierarchy object, which was the typical former approach.
\end{itemize}

\subsection{Inner loop optimization}
\label{subsec:inner_loop_optimization}
The following modifications were implemented in order to optimize the possible inner loop for convergence of the Hodge variable $\Phi_{n+1}$ (see \fig \ref{fig:flow_chart}).
\begin{itemize}
 \item The Poisson solver for face-sampled (vector) fields \verb|my_p4est_poisson_faces_t| was modified in order to store in memory all \verb|P4EST_DIM| matrices that it builds. Previously, those matrices (and the corresponding preconditioned Krylov solvers) were built and destroyed immediately after the solve-step for the velocity component that was currently treated, one after the other. This impeded the possible reuse of the (exact same) Krylov solvers within such a possible inner loop, introducing an overhead cost for recalculating the very same matrices and solvers as previsouly destroyed. This modification applies for the interface-stress-balanced two-phase flows solver.
 \item The \verb|solve_viscosity| and \verb|solve_projection| functions within the Navier-Stokes solver now allow the user to possibly feed them with valid face-based and cell-based Poisson solvers and/or to retrieve them from the object for possible reuse. If the corresponding parameter is \verb|NULL| on input, the functions build the solvers internally and return them to the calling environment (instead of destroying it, right away). Therefore, the user can now possibly reuse the solvers within such an inner loop, avoiding the cost of re-assembling the linear system and re-building preconditioners.
 \item These functions also now allow the user to specify the kinds of Krylov solver and preconditioners they wish to use for their application. The default values for these fields are what used to be hard-coded internally, \textit{i.e.} a BiConjugate Gradient Stabilized (\verb|BiCGStab|) solver with Successive Over-Relaxation (\verb|SOR|) for the preconditioner. While \verb|BiCGStab| is the only safe choice for the solvers in the most general context (given that interface boundary conditions might break matrices' symmetries), there are better preconditioners to use than \verb|SOR| in general.
 \item All relevant backtraced semi-lagrangian points in \verb|solve_viscosity| are calculated \emph{simultaneously}, instead of considering faces dimension by dimension. This cuts down the number of calls to interpolation procedure by \verb|P4EST_DIM|. For that purpose, a capability to calculate trajectories of several lists of points at once has been added (to minimize the number of calls to \verb|trajectory_from_np1_to_n(m1)| which all used up to \verb|5*P4EST_DIM| interpolation procedures). In summary the number of calls to interpolation tools was cut down by a factor \verb|P4EST_DIM*PEST_DIM| which was terribly slow if the node neighbors were not initialied. All former relevant functions for semi-lagrangian backtracing of faces along one single Cartesian direction now call this generalized function (although they should be considered outdated from now on).
 \item All relevant backtraced semi-lagrangian points are stored in memory and calculated only once per inner loop for $\Phi_{n+1}$-convergence (\textit{i.e.,} not recalculated in case of subiterations).
\end{itemize}

\subsection{General time-stepping optimization in absence of grid update}
\label{subsec:general_time_stepping}
The following modifications were implemented in order to optimize time-stepping in case of absence of modifications in the computational grid (either enforced/desired by the user or as a result of the regular grid-update procedure if it reaches a conclusion that no grid update nor any modification to the levelset were made). 
\begin{itemize}
 \item The cell- and face-based Poisson solvers were modified to allow a simple update of the constant coefficient term. If the grid is not modified, the solvers from the previous time step are almost ready to be re-used: their memory allocation is still valid, their sparsity is unchanged and most off-diagonal terms are unchanged, only the terms affected by the constant coefficient need to be updated (most are on the diagonal of the matrices, but not necessarily all of them). The objects now allow to keep track of such states and update only the relevant terms within the matrices, which avoids costly reconstructions of the entire matrices.\\
 The \verb|update_from_tn_to_tnp1| function from the Navier-Stokes solver now returns a boolean value that is true if the solvers for viscosity and projection steps can be used again in the future. This flag can be true if and only if the grid has not been changed (nor its partitioning) and if the levelset has not been modified either. Otherwise, the solvers must be deleted and reconstructed!
 \item In case of no grid update, only the set of grid-defining objects corresponding to time step $n$ is kept in memory. The corresponding pointers associated with $\left(n-1\right)$-objects simply point to the $n$-objects. This was meant to ensure the re-usability of face- and cell-solvers for the following time step: since the grid is not modified, we keep it as such, we do not create a new one and we make the $\left(n-1\right)$-pointers point to the $n$-object. Data management was modified to avoid deletion of $\left( n-1 \right)$-objects if they happen to point to $n$-objects.
 \item No expensive cell-to-cell nor face-to-face interpolations are executed when no modification to the computational grid was made. 
\end{itemize}

\subsection{Non-restricted optimization of the solver}
\label{subsec:nonrestricted_optimization_of_the_solver}
The following modifications target general optimizaton of several distinct parts within the Navier-Stokes solver.
\begin{itemize}
 \item The node neighbors are \emph{initialized} right after the finalization of a new distributed grid and of the corresponding new distributed nodes. This speeds up significantly all calculations querying access to the node neighbors like the calculations of spatial first or second derivatives. Previously, the node neighbors were initialized only for data extension over the interface (in presence of an interface in the computational domain) at the very end of every time step.
 \item All \verb|P4EST_DIM| spatial second derivatives along Cartesian axes of the \verb|P4EST_DIM| velocity components (at nodes) at times $n$ and $\left( n-1 \right)$ are now stored in memory in order to accelerate all quadratic interpolation of velocity fields (required in the calculation of semi-lagrangian trajectories \emph{and} in the definition of the right-hand sides of the viscosity step).
 \item The solver now allows the user to choose between two different time steps satisfying either a \emph{global} or \emph{local} CFL condition. The global CFL requirement defines the time step as  
 $$\Delta t = \dfrac{\text{CFL}\, \min_{\Omega}\left( \Delta x, \Delta y, \Delta z \right)}{\max_{\Omega} \left\| \mathbf{v}_{n+1}\right\|_{2}}$$ 
 whereas, the local CFL requirement defines it as
 $$\Delta t = \text{CFL} \min_{\Omega}\left( \dfrac{\Delta x}{\left\| \mathbf{v}_{n+1}\right\|_{2}}, \dfrac{\Delta y}{\left\| \mathbf{v}_{n+1}\right\|_{2}}, \dfrac{\Delta z}{\left\| \mathbf{v}_{n+1}\right\|_{2}}\right)$$ 
 which ensures that the CFL conditions is satisfied locally everywhere. The former definition might be much more restrictive than the latter in simulations where the largest velocities are found in the coarsest regions and vice versa, \textit{e.g.} developing or laminar flows past a sphere, channel flows, etc. 
 \item For the very first iteration of the iterative grid-update procedure, the grid-related data structures are no longer built except for the a copy of the raw grid. Similarly, the node-sampled voriticy does not go through a dummy interpolation and the levelset function either if it is not modified.
 \item The iterative grid-update procedure no longer builds the ghost layer and the hierarchy at every iteration, except in case of smoke advection associated with smoke-based grid refinement since they are required for the resolution of the advection equation in this case. The final ghost layer is built only once, after completion of the grid update procedure.\\ 
 The construction of a ghost layer and the corresponding hierarchy were significant time-consuming parts of every grid update iteration. Regarding the ghost cells, they are relevant in this context for the sole purpose of identifying T-junction nodes that happen to lie exactly on the partition's border without being owned by itself. As for the hierarchy, it was used for easily finding any possible local node (owned or not by the current partition).\\
 Putting in balance the solver's behavior with the extra cost of creating a ghost layer and the extra \verb|MPI| communication it created, I decided to allow the grid update to miss the (supposedly few) such border T-junction nodes that are not owned. The rationale is that 
 \begin{itemize}
  \item a 2:1 grid balance is still enforced anyways so that such misses should not be too detrimental eventually;
  \item the iterative grid update procedure will very likely change the grid partition and such missed nodes could as well be indexed by the partition in the next iteration.
 \end{itemize}
 The usage of the costly-to-build hierarchy was replaced by the \verb|index_of_node| binary-search capability detailed in subsection \ref{subsec:general_interest}: if the node actually exists, its index is found in $\mathcal{O}\left( \log\left( N \right)\right)$ operations.
 \item When refining with smoke, the smoke advection is not recalculated after the grid update procedure if the 2:1 grid balancing operations turned out to leave the grid unchanged. Instead the current valid \verb|smoke_np1| vector is simply re-scattered to the new vector layout.
 \item The number of operations in the procedure to tag quadrant with their refining or coarsening status was minimized, and break statements are used to exit loops as soon as the (final) result is known.
\end{itemize}

\subsection{General optimization-related intel for the \texttt{casl\_p4est} developers}
I attempted to further optimize the execution time by changing all core linear algebra functions from the \verb|matrix_t| class to \verb|CBLAS| and \verb|LAPACKE| functions (mainly \verb|cblas_dgemv|, \verb|cblas_dgemm|, \verb|LAPACKE_dpotrf| and \verb|LAPACKE_dgels|). However, those core linear algebra functions are called as such when evaluating local moving-least-square interpolations and it turns out that the linear systems to be solved are not large enough to observe a significant improvement from the optimized cache management done under the hood by those optimized library. Actually, it even makes the execution slightly slower. I commented the modifications in the corresponding files in case someone else wants to ever try it again in the future...
\section{Incomplete modifications}

There was a mistake in the treatment of Neumann wall-boundary conditions on velocity components when an interface is close, this has been fixed. Starting from that observation, I decided to implement a feature to activate a possibility to account for second derivatives of Hodge variables in case of Neumann boundary conditions. If the corresponding flag is activated, those second derivatives are estimated, otherwise a value of $0$ is substituted.

However, it became overwhelmingly complicated and I had to give up by the end of the assembly of linear system and of its right-hand side. Therefore, the task is not completed yet but it is documented! As of now, only the unactivated flag value is acceptable!

\section{Showing unchanged behavior besides runtime}
First, the performance results are shown in 2D and 3D based on test cases that are designed to show an unchanged solver's behavior, yet a significant gain in runtime due to better computing strategies: namely the more efficient iterative grid-update procedure, calculations of semi-lagrangian backtraced points, and grouped interpolations procedure.

These two benchmark cases run the \emph{exact} same solver configurations. No cell-/face-based poisson solver is ever re-used but disposed of after every single use. The test cases involve advection of smoke (\textit{i.e.}, a passive scalar) and adaptive refinement is triggered by distance to the interface, local vorticity and local value of the passive scalar. Therefore, the grid update procedure must be unchanged as well\footnote{In absence of refinement based on local values of a passive scalar, the ghost layers are no longer created within the iterative grid-update procedure, which can lead to a modified behavior overall because of T-junctions lying exactly on partitions' boundaries but not owned by them}.

All runs have been processed on a machine running a dual Intel Xeon 6148 2.4 \SI{2666}{\mega\hertz} 20C CPU, with \SI{64}{\gibi\byte} DDR4.

\subsection{Flow past a cylinder}
\label{subsec:flow_past_cylinder_check}
For the two-dimensional example, we consider the flow past a static cylinder with advection of a passive scalar (smoke). The domain size is $\left[0, 32 \right]\times \left[ -8, 8\right]$. The cylinder is centered at $\left( 8, 0 \right)$ and has a radius of $r_{0} = 0.5$. The macro-mesh is made of $8$ trees along $x$ and $4$ trees along $y$ (aspect ratio of $1$ for all quadrants). The trees are refined with minimum refinement level $4$ and maximum refinement level $6$, with the following grid-refinement parameters:
\begin{itemize}
 \item the vorticity threshold value for triggering local refinement is $0.1\dfrac{\max_{\Omega}\left( \left\| \mathbf{v} \right\|_{2}\right)}{\max\left( \Delta x, \Delta y \right)}$;
 \item a uniform band of $2$ finest grid cells is enforced on either side of the surface of the cylinder;
 \item the threshold value for triggering smoke-based refinement is $0.5$.
\end{itemize}

The boundary conditions are 
\begin{itemize}
 \item homogeneous Dirichlet boundary condition at the inlet (left wall) and homogeneous Neumann on all other walls for the pressure;
 \item homogeneous Neumann for pressure on the surface of the cylinder;
 \item Dirichlet boundary condition $\mathbf{u} = u_{0}\mathbf{e}_{x} = \mathbf{e}_{x}$ for the velocity components on all walls but the outlet (right wall) on which homogeneous Neumann is enforced;
 \item homogeneous Dirichlet boundary condition for $\mathbf{u}$ on the surface of the cylinder (no-slip);
 \item the smoke is released into the computational domain by a unit Dirichlet boundary condition on part of the inlet wall: $x = 0$, $\left| y\right| < 0.5$.
\end{itemize}
The initial condition for the velocity field is $\left.\mathbf{u}\left( \mathbf{x}, t\right)\right|_{t = 0} = u_{0}\mathbf{e}_{x} = \mathbf{e}_{x}$. The initial boundary condition for the smoke is $0$ everywhere but in $\left[0, \, 2 \right]\times \left[ -1, \, 1\right]$ where it is $1$.

The mass density of the fluid is $\rho = 1$, its dynamic viscosity is set to $\mu = \dfrac{2 r_{0}\rho u_{0}}{\Reynolds} = \dfrac{1}{\Reynolds}$ where $\Reynolds = 350$ in the following tests. The simulations were run using $8$ cores with a second-order semi-lagrangian advection scheme, the CFL number is enforced to be smaller than or equal to $1$ (global criterion) until a simulation time of $400$ for a total number of $9950$ time steps. The number of grid cells grows from around $9,500$ to about $35,000$ as the K\'{a}rm\'{a}n street develops.

\subsubsection*{Results and performances}
The original solver completed the simulation in \SI{1}{\hour}\,\SI{41}{\minute}\,\SI{45}{\second} while the optimized solver completed the very same task with the exact same numerics in \SI{56}{\minute}\,\SI{19}{\second}, that is a \textbf{44.64\% gain}. The only difference observed in the exported results relates to slight differences in the very first residuals for the Hodge variable in $6$ out of $9,950$ time steps. These slight differences can be explained by a modification brought to the Moving-Least-Square cell-interpolation procedure that intends to make the procedure more robust and less sensitive to the grid size (comparison of dimensionally consistent double-precision real numbers). The results are illustrated in \fig \ref{fig:forces_flow_past_cylinder_check_unchanged} and \fig \ref{fig:grid_and_smoke_flow_past_cylinder_check_unchanged}.

\begin{figure}
  \centering
  \includegraphics[width=0.49\textwidth]{./results/flow_past_cylinder_smoke/reference/force_history.pdf}
  \includegraphics[width=0.49\textwidth]{./results/flow_past_cylinder_smoke/optimized_run/force_history.pdf}
 \caption{\label{fig:forces_flow_past_cylinder_check_unchanged} history of the non-dimensional force components for $300 < t < 400$ in the simulation described in subsection \ref{subsec:flow_past_cylinder_check}. Top: original solver; bottom: optimized solver.} 
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=0.95\textwidth]{./figures/flow_past_cylinder/check_unchanged_behavior/original_grid_and_smoke_t_is_300.pdf} \\
  \includegraphics[width=0.95\textwidth]{./figures/flow_past_cylinder/check_unchanged_behavior/optimized_grid_and_smoke_t_is_300.pdf}
 \caption{\label{fig:grid_and_smoke_flow_past_cylinder_check_unchanged} computational grid and color map of the advected smoke for $t = 300$ in the simulation described in subsection \ref{subsec:flow_past_cylinder_check}. Top: original solver; bottom: optimized solver.}
\end{figure}

\subsection{Flow past a sphere}
\label{subsec:flow_past_sphere_check}
For the three-dimensional example, we consider the flow past a static sphere, with advection of a passive scalar (smoke). The domain size is $\left[0, 32 \right]\times \left[ -8, 8\right] \times \left[ -8, 8\right]$. The sphere is centered at $\left( 8, 0, 0 \right)$ and has a radius of $r_{0} = 1$. The macro-mesh is made of $8$ trees along $x$, $4$ trees along $y$ and $4$ trees along $z$ (aspect ratio of $1$ for all quadrants). The trees are refined with minimum refinement level $4$ and maximum refinement level $6$, with the following grid-refinement parameters:
\begin{itemize}
 \item the vorticity threshold value for triggering local refinement is $0.1\dfrac{\max_{\Omega}\left( \left\| \mathbf{v} \right\|_{2}\right)}{\max\left( \Delta x, \Delta y \right)}$;
 \item a uniform band of $2$ finest grid cells is enforced on either side of the surface of the cylinder;
 \item the threshold value for triggering smoke-based refinement is $0.5$.
\end{itemize}

The boundary conditions are 
\begin{itemize}
 \item homogeneous Dirichlet boundary condition at the inlet ($x = 0$ wall) and homogeneous Neumann on all other walls for the pressure;
 \item homogeneous Neumann for pressure on the surface of the sphere;
 \item Dirichlet boundary condition $\mathbf{u} = u_{0}\mathbf{e}_{x} = \mathbf{e}_{x}$ for the velocity components on all walls but the outlet ($x = 32$ wall) on which homogeneous Neumann is enforced;
 \item homogeneous Dirichlet boundary condition for $\mathbf{u}$ on the surface of the sphere (no-slip);
 \item the smoke is released into the computational domain by a unit Dirichlet boundary condition on part of the inlet wall: $x = 0$, $\sqrt{y^2 + z^2} < 0.4$.
\end{itemize}
The initial condition for the velocity field is $\left.\mathbf{u}\left( \mathbf{x}, t\right)\right|_{t = 0} = u_{0}\mathbf{e}_{x} = \mathbf{e}_{x}$. The initial boundary condition for the smoke is $0$ everywhere but in $\left\{ 0 \leq x \leq 2\right\} \bigcap \left\{ \sqrt{ y^2 + z^2} < 0.4\right\}$ where it is $1$.

The mass density of the fluid is $\rho = 1$, its dynamic viscosity is set to $\mu = \dfrac{2 r_{0}\rho u_{0}}{\Reynolds} = \dfrac{2}{\Reynolds}$ where $\Reynolds = 350$ in the following tests. The simulations were run using $36$ cores with a second-order semi-lagrangian advection scheme, the CFL number is enforced to be smaller than or equal to $1$ (global criterion) until a simulation time of $200$ for a total of $3,852$ time steps. The number of grid cells grows from around $558,280$ to about $700,000$ as the wake develops behind the sphere.

\subsubsection*{Results and performances}

The original solver completed the simulation in \SI{1}{\day}\,\SI{11}{\hour}\,\SI{12}{\minute}\,\SI{56}{\second} while the optimized solver completed the very same task with the exact same numerics in \SI{23}{\hour}\,\SI{12}{\minute}\,\SI{41}{\second}, that is a \textbf{34.09\% gain}. The only difference observed in the exported results relates to slight differences in the very first residuals for the Hodge variable in $6$ out of $3,852$ time steps. These slight differences can be explained by a modification brought to the Moving-Least-Square cell-interpolation procedure that intends to make the procedure more robust and less sensitive to the grid size (comparison of dimensionally consistent double-precision real numbers). The results are illustrated in \fig \ref{fig:forces_flow_past_sphere_check_unchanged} and \fig \ref{fig:grid_and_smoke_flow_past_sphere_check_unchanged}.

\begin{figure}
  \centering
  \includegraphics[width=0.49\textwidth]{./results/flow_past_sphere_smoke/reference/force_history.pdf}
  \includegraphics[width=0.49\textwidth]{./results/flow_past_sphere_smoke/optimized_run/force_history.pdf}
 \caption{\label{fig:forces_flow_past_sphere_check_unchanged} history of the non-dimensional force components for $100 < t < 200$ in the simulation described in subsection \ref{subsec:flow_past_sphere_check}. Top: original solver; bottom: optimized solver. (The graphs for $C_{\mathrm{D}, y}$ and $C_{\mathrm{D}, z}$ overlap)} 
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=0.95\textwidth]{./figures/flow_past_sphere/check_unchanged_behavior/original_grid_and_smoke_t_is_200.pdf} \\ 
  \includegraphics[width=0.95\textwidth]{./figures/flow_past_sphere/check_unchanged_behavior/optimized_grid_and_smoke_t_is_200.pdf}
 \caption{\label{fig:grid_and_smoke_flow_past_sphere_check_unchanged} computational grid and color map of the advected smoke for $t = 200$ in the simulation described in subsection \ref{subsec:flow_past_sphere_check}. Top: original solver; bottom: optimized solver.}
\end{figure}

\section{Assessment of newly added features departing from the original solver}
In this section, we present and investigate other features that were added to the Navier-Stokes solver to possibly accelerate its execution. In particular, we consider the possibility to re-use the cell and face solvers with appropriate good initial guesses within the inner loop and for successive time steps (if the grid has not changed, the interface has not moved and the boundary conditions have not changed). Concurrently, we consider the possibility to restrict the number of grid updates: given that the cylinder/sphere is static, we can take advantage of the constant physical description of the problem.

\subsection{Flow past a cylinder with smoke and smoke-based refinement}
\label{subsec:assessment_cylinder_and_smoke}
For this analysis, we consider the same benchmark test case as described in subsection \ref{subsec:flow_past_cylinder_check}. However, instead of deleting the cell and face solvers right every single usage, the Navier-Stokes solver is now allowed to re-use cell and face solvers. We also increase the number of time steps between each grid update from $1$ to $2$, $4$ and $8$.

When using the same solvers and the same preconditioners as originally hard-coded within the Navier-Stokes solver, \textit{i.e.}, \verb|BiCGStab| solvers for both the velocity components and the Hodge variable with \verb|SOR| preconditioning), the execution time drops to \SI{53}{\minute}\,\SI{48}{\second}, \SI{44}{\minute}\,\SI{27}{\second}, \SI{37}{\minute}\,\SI{56}{\second} and \SI{32}{\minute}\,\SI{3}{\second} when updating the grid every $1$, $2$, $4$ and $8$ time step(s) respectively. The gain in execution time when the grid is updated at every time step comes mainly from reusing the cell and face solvers as well as their preconditioners with good initial guesses within the inner loop for convergence of the Hodge variable. In this case, the gain is very small, which can be explained by the very small cost of building a \verb|SOR| preconditoner and the large number of iterations required by the Krylov subspace solvers to converge when using such a preconditioner. The force components over the time period $300 < t < 400$ are illustrated in \fig \ref{fig:forces_flow_past_cylinder_reused_solvers_bcgs_and_pcsor}. A time shit in the force results appears when the grid updates are prevented from happening at every time step, however the frequency of the vortex shedding does not seem to be affected.

When using a Conjugate Gradient (\verb|CG|) solver for the Hoge variable (the linear system to be solved is symmetric positive definite in this case) and \verb|HYPRE| preconditioners instead, the execution time drops to \SI{48}{\minute}\,\SI{52}{\second}, \SI{37}{\minute}\,\SI{39}{\second}, \SI{31}{\minute}\,\SI{3}{\second} and \SI{26}{\minute}\,\SI{18}{\second} when updating the grid every $1$, $2$, $4$ and $8$ time step(s), respectively. When compared to the results obtained with the original (hardcoded) solvers, this shows that one can gain even more execution time by using better suited solvers along with optimal preconditioners\footnote{Although costly to build, the \texttt{HYPRE} preconditioner is \emph{optimal} in several ways: the number of iterations required for convergence is significantly smaller than for most other available (parallel) preconditioners and it is also almost independent under further grid refinement (very efficient clustering).}. The force components over the time period $300 < t < 400$ are illustrated in \fig \ref{fig:forces_flow_past_cylinder_reused_solvers_cg_and_hypre}: note the similarity between graphs from \fig \ref{fig:forces_flow_past_cylinder_reused_solvers_bcgs_and_pcsor} and and graphs from \fig \ref{fig:forces_flow_past_cylinder_reused_solvers_cg_and_hypre}.

\begin{figure}
  \centering
  \includegraphics[width=0.49\textwidth]{./results/flow_past_cylinder_smoke/optimized_run_reused_solvers_bcgs_and_pcsor/grid_update_1/force_history.pdf}
  \includegraphics[width=0.49\textwidth]{./results/flow_past_cylinder_smoke/optimized_run_reused_solvers_bcgs_and_pcsor/grid_update_2/force_history.pdf} \\
  \includegraphics[width=0.49\textwidth]{./results/flow_past_cylinder_smoke/optimized_run_reused_solvers_bcgs_and_pcsor/grid_update_4/force_history.pdf} 
  \includegraphics[width=0.49\textwidth]{./results/flow_past_cylinder_smoke/optimized_run_reused_solvers_bcgs_and_pcsor/grid_update_8/force_history.pdf} \\
 \caption{\label{fig:forces_flow_past_cylinder_reused_solvers_bcgs_and_pcsor} history of the non-dimensional force components for $300 < t < 400$ in the simulation described in subsection \ref{subsec:flow_past_cylinder_check}. These results have been obtained with \texttt{BiCGStab} solvers for the velocity components and the Hodge variable. All solvers are preconditioned using \texttt{SOR} preconditioning. The solvers and their preconditioners are re-used as long as they are valid. The grid is updated every $1$, $2$, $4$ and $8$ time step(s), from top left to bottom right figure.} 
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=0.49\textwidth]{./results/flow_past_cylinder_smoke/optimized_run_reused_solvers_cg_and_pchypre/grid_update_1/force_history.pdf} 
  \includegraphics[width=0.49\textwidth]{./results/flow_past_cylinder_smoke/optimized_run_reused_solvers_cg_and_pchypre/grid_update_2/force_history.pdf} \\
  \includegraphics[width=0.49\textwidth]{./results/flow_past_cylinder_smoke/optimized_run_reused_solvers_cg_and_pchypre/grid_update_4/force_history.pdf}
  \includegraphics[width=0.49\textwidth]{./results/flow_past_cylinder_smoke/optimized_run_reused_solvers_cg_and_pchypre/grid_update_8/force_history.pdf} \\
 \caption{\label{fig:forces_flow_past_cylinder_reused_solvers_cg_and_hypre} history of the non-dimensional force components for $300 < t < 400$ in the simulation described in subsection \ref{subsec:flow_past_cylinder_check}. These results have been obtained with a \texttt{CG} solver for the Hodge variable and \texttt{BiCGStab} solvers for the velocity components. All solvers are preconditioned using the \texttt{HYPRE} preconditioner. The solvers and their preconditioners are re-used as long as they are valid. The grid is updated every (from top left to bottom right figure): $1$ time step, $2$ time steps, $4$ time steps, $8$ time steps.} 
\end{figure}

\subsection{Flow past a sphere with smoke and smoke-based refinement}
\label{subsec:assessment_sphere_and_smoke}
For this analysis, we consider the same benchmark test case as described in subsection \ref{subsec:flow_past_sphere_check}. However, instead of deleting the cell and face solvers right every single usage, the Navier-Stokes solver is now allowed to re-use cell and face solvers. We also increase the number of time steps between each grid update from $1$ to $2$, $4$ and $8$.

When using the same solvers and the same preconditioners as originally hard-coded within the Navier-Stokes solver, \textit{i.e.}, \verb|BiCGStab| solvers for both the velocity components and the Hodge variable with \verb|SOR| preconditioning), the execution time drops to \SI{20}{\hour}\,\SI{32}{\minute}\,\SI{32}{\second}, \SI{13}{\hour}\,\SI{30}{\minute}\,\SI{52}{\second}, \SI{10}{\hour}\,\SI{18}{\minute}\,\SI{41}{\second} and \SI{8}{\hour}\,\SI{26}{\minute}\,\SI{31}{\second} when updating the grid every $1$, $2$, $4$ and $8$ time step(s) respectively. In this case, the gain is much more significant than in the two-dimensional case when solvers are reused, and can probably be explained by a significant reduction in the number of iterations required by the iterative solvers for convergence, when the Krylov solvers start from good initial guesses. The force components over the time period $100 < t < 200$ are illustrated in \fig \ref{fig:forces_flow_past_sphere_reused_solvers_bcgs_and_pcsor}. A time shit in the force results appears when the grid updates are prevented from happening at every time step, however the frequency of the vortex shedding does not seem to be affected.

When using a Conjugate Gradient (\verb|CG|) solver for the Hoge variable (the linear system to be solved is symmetric positive definite in this case) along with a \verb|HYPRE| preconditioner for the cell-solver\footnote{The linear systems for the face problems have a significant positive diagonal contribution and only a few iterations are required for convergence even with simple \texttt{SOR} preconditioners, it does not seem relevant to consider another kind of costly-to-build conditioner for threse problems} instead, the execution time drops to \SI{20}{\hour}\,\SI{44}{\minute}\,\SI{48}{\second}, \SI{14}{\hour}\,\SI{6}{\minute}\,\SI{54}{\second}, \SI{10}{\hour}\,\SI{4}{\minute}\,\SI{43}{\second} and \SI{8}{\hour}\,\SI{20}{\minute}\,\SI{40}{\second} when updating the grid every $1$, $2$, $4$ and $8$ time step(s), respectively. Contrary to the two-dimensional case, it seems like the complexity of building the \verb|HYPRE| preconditioner takes over when assessing the computational cost of solving the linear systems in this case: one gains execution time only when grid updates are prevented (allowing more reuse of the costly-to-build preconditioner). No general conclusion can be drawn from these results however since solver performances are case- and eigenvalues-dependent\footnote{Personally, I expect \texttt{HYPRE} to outperform \texttt{SOR} preconditioning when dealing with larger and finer grids because of the larger spectrum of eigenvalues to be clustered.}.

The force components over the time period $100 < t < 200$ are illustrated in \fig \ref{fig:forces_flow_past_sphere_reused_solvers_cg_and_hypre}: note the similarity between graphs from \fig \ref{fig:forces_flow_past_sphere_reused_solvers_bcgs_and_pcsor} and graphs from \fig \ref{fig:forces_flow_past_sphere_reused_solvers_cg_and_hypre}.

\begin{figure}
  \centering
  \includegraphics[width=0.49\textwidth]{./results/flow_past_sphere_smoke/optimized_run_reused_solvers_bcgs_and_pcsor/grid_update_1/force_history.pdf} 
  \includegraphics[width=0.49\textwidth]{./results/flow_past_sphere_smoke/optimized_run_reused_solvers_bcgs_and_pcsor/grid_update_2/force_history.pdf} \\
  \includegraphics[width=0.49\textwidth]{./results/flow_past_sphere_smoke/optimized_run_reused_solvers_bcgs_and_pcsor/grid_update_4/force_history.pdf}
  \includegraphics[width=0.49\textwidth]{./results/flow_past_sphere_smoke/optimized_run_reused_solvers_bcgs_and_pcsor/grid_update_8/force_history.pdf} \\
 \caption{\label{fig:forces_flow_past_sphere_reused_solvers_bcgs_and_pcsor} history of the non-dimensional force components for $100 < t < 200$ in the simulation described in subsection \ref{subsec:flow_past_sphere_check}. These results have been obtained with \texttt{BiCGStab} solvers for the velocity components and the Hodge variable. All solvers are preconditioned using \texttt{SOR} preconditioning. The solvers and their preconditioners are re-used as long as they are valid. The grid is updated every $1$, $2$, $4$ and $8$ time step(s), from top left to bottom right figure. The sign of the results for the $y$- and $z$- components was inverted in the top right figure, for comparison purposes.} 
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=0.49\textwidth]{./results/flow_past_sphere_smoke/optimized_run_reused_solvers_cg_and_pchypre/grid_update_1/force_history.pdf}
  \includegraphics[width=0.49\textwidth]{./results/flow_past_sphere_smoke/optimized_run_reused_solvers_cg_and_pchypre/grid_update_2/force_history.pdf} \\
  \includegraphics[width=0.49\textwidth]{./results/flow_past_sphere_smoke/optimized_run_reused_solvers_cg_and_pchypre/grid_update_4/force_history.pdf} 
  \includegraphics[width=0.49\textwidth]{./results/flow_past_sphere_smoke/optimized_run_reused_solvers_cg_and_pchypre/grid_update_8/force_history.pdf} \\
 \caption{\label{fig:forces_flow_past_sphere_reused_solvers_cg_and_hypre} history of the non-dimensional force components for $100 < t < 200$ in the simulation described in subsection \ref{subsec:flow_past_sphere_check}. These results have been obtained with a \texttt{CG} solver for the Hodge variable and \texttt{BiCGStab} solvers for the velocity components. The \texttt{CG} solver is preconditioned with the \texttt{HYPRE} preconditioner while the \texttt{BiCGStab} solvers are preconditioned using \texttt{SOR} preconditioners. The solvers and their preconditioners are re-used as long as they are valid. The grid is updated every (from top left to bottom right figure): $1$ time step, $2$ time steps, $4$ time steps, $8$ time steps, from top left to bottom right figure. Here again, the sign of the results for the $y$- and $z$- components was inverted in the top right figure, for comparison purposes.} 
\end{figure}

\section{A note on a possible change of behavior}

Finally, we want to assess and evaluate the possible change of behavior of the solver, in \emph{absence} of grid refinement based on a passively advected scalar. Indeed, the iterative grid update procedure was modified in order to avoid the (costly) creation of ghost layers for every iterative step. As detailed in subsection \ref{subsec:nonrestricted_optimization_of_the_solver}, this modification \emph{might} induce a very slight change of behavior in the solver: if the node value triggering the refinement criterion happens to be a T-junction at the border of the domain's partition but owned by another processor, this node would be missed locally and would not trigger local refinement. However, we expect that to be relatively rare and mitigated by the grid repartitioning (within the iterative process) and the 2:1 grid balancing.

In order to assess that possibility, we compared the original and optimized solver on the same test case as described in subsection \ref{subsec:flow_past_sphere_check}, but \emph{without} passive advection. For the entire run, the only difference observed in the exported results relates to slight differences in the very first residuals for the Hodge variable in $3$ out of $3,852$ time steps. These slight differences can be explained by a modification brought to the Moving-Least-Square cell-interpolation procedure that intends to make the procedure more robust and less sensitive to the grid size (comparison of dimensionally consistent double-precision real numbers). No difference in computational grids is reported, so the solver's behavior was not affected by the modified strategy.

In terms of performance, the original solver completed the simulation in \SI{1}{\day}\,\SI{9}{\hour}\,\SI{30}{\minute}\,\SI{4}{\second} while the optimized solver completed the very same task with the exact same numerics in \SI{22}{\hour}\,\SI{51}{\minute}\,\SI{30}{\second}, that is a \textbf{31.77\% gain}.

\begin{figure}
  \centering
  \includegraphics[width=0.49\textwidth]{./results/flow_past_sphere_smoke/reference/force_history.pdf}
  \includegraphics[width=0.49\textwidth]{./results/flow_past_sphere_smoke/optimized_run/force_history.pdf}
 \caption{\label{fig:forces_flow_past_sphere_check_unchanged} history of the non-dimensional force components for $100 < t < 200$ in the simulation described in subsection \ref{subsec:flow_past_sphere_check}. Top: original solver; bottom: optimized solver. (The graphs for $C_{\mathrm{D}, y}$ and $C_{\mathrm{D}, z}$ overlap)} 
\end{figure}

\newpage
\section{Summary}
The results of the optimization analyses are summarized in Tables \ref{tab:flow_past_cylinder_with_smoke} and \ref{tab:flow_past_sphere_with_smoke}.

\begin{landscape}

\begin{table}
  \newcommand{\mc}[3]{\multicolumn{#1}{#2}{#3}}
  \begin{center}
    \begin{tabular}[t]{|c|c|c|c|p{10mm}|p{25mm}|c|c|p{55mm}|}\hline
      N-S solver & cell-solver & face-solver & precond. & solver re-use & \mc{1}{c|}{grid update} & execution time & gain & \mc{1}{c|}{notes}\\\hhline{|=|=|=|=|=|=|=|=|=|}
      original & \verb|BiCGStab| & \verb|BiCGStab| & SOR & \mc{1}{c|}{no} & every time step & \SI{1}{\hour}\,\SI{41}{\minute}\,\SI{45}{\second} & 0\% & reference results (before significant code changes) \\\hline
      optimized & \verb|BiCGStab| & \verb|BiCGStab| & SOR & \mc{1}{c|}{no} & every time step & \SI{56}{\minute}\,\SI{19}{\second} & 44.64\% & exact same results except for a few initial residuals of the Hodge variable ($6$ out of $9,950$ time steps)\\\hhline{|=|=|=|=|=|=|=|=|=|}
      optimized & \verb|BiCGStab| & \verb|BiCGStab| & SOR & \mc{1}{c|}{yes} & every time step & \SI{53}{\minute}\,\SI{48}{\second} & 47.13\% & same numerics as originally, results are almost exactly the same \\\hline
      optimized & \verb|BiCGStab| & \verb|BiCGStab| & SOR & \mc{1}{c|}{yes} & once every $2$ time steps & \SI{44}{\minute}\,\SI{27}{\second} & 56.31\% & time shift but frequency of vortex shedding unchanged\\\hline
      optimized & \verb|BiCGStab| & \verb|BiCGStab| & SOR & \mc{1}{c|}{yes} & once every $4$ time steps & \SI{37}{\minute}\,\SI{56}{\second} & 62.72\% & same as above \\\hline
      optimized & \verb|BiCGStab| & \verb|BiCGStab| & SOR & \mc{1}{c|}{yes} & once every $8$ time steps & \SI{32}{\minute}\,\SI{3}{\second}  & 68.5\% & same as above \\\hhline{|=|=|=|=|=|=|=|=|=|}
      optimized & CG & \verb|BiCGStab| & \verb|HYPRE| & \mc{1}{c|}{yes} & every time step           & \SI{48}{\minute}\,\SI{52}{\second} & 51.97\% & different numerics than originally, but results are almost exactly the same \\\hline
      optimized & CG & \verb|BiCGStab| & \verb|HYPRE| & \mc{1}{c|}{yes} & once every $2$ time steps & \SI{37}{\minute}\,\SI{39}{\second} & 63.00\% & time shift but frequency of vortex shedding unchanged \\\hline
      optimized & CG & \verb|BiCGStab| & \verb|HYPRE| & \mc{1}{c|}{yes} & once every $4$ time steps & \SI{31}{\minute}\,\SI{3}{\second} & 69.48 \% & same as above \\\hline
      optimized & CG & \verb|BiCGStab| & \verb|HYPRE| & \mc{1}{c|}{yes} & once every $8$ time steps & \SI{26}{\minute}\,\SI{18}{\second} & 74.15\% & same as above \\\hline
    \end{tabular}
  \end{center}
  \caption{\label{tab:flow_past_cylinder_with_smoke} summary of the runtime-optimization results for two-dimensional flow past a cylinder with advection of smoke and smoke-based grid refinement.}
\end{table}

\begin{table}
  \newcommand{\mc}[3]{\multicolumn{#1}{#2}{#3}}
  \begin{center}
    \begin{tabular}[t]{|c|c|c|c|p{10mm}|p{25mm}|c|c|p{55mm}|}\hline
      N-S solver & cell-solver & face-solver & precond. & solver re-use & \mc{1}{c|}{grid update} & execution time & gain & \mc{1}{c|}{notes}\\\hhline{|=|=|=|=|=|=|=|=|=|}
      original & \verb|BiCGStab| & \verb|BiCGStab| & SOR & \mc{1}{c|}{no} & every time step &  \SI{1}{\day}\,\SI{11}{\hour}\,\SI{12}{\minute}\,\SI{56}{\second} & 0\% & reference results (before significant code changes) \\\hline
      optimized & \verb|BiCGStab| & \verb|BiCGStab| & SOR & \mc{1}{c|}{no} & every time step & \SI{23}{\hour}\,\SI{12}{\minute}\,\SI{41}{\second} & 34.09\% & exact same results except for a few initial residuals of the Hodge variable ($6$ out of $3,852$ time steps)\\\hhline{|=|=|=|=|=|=|=|=|=|}
      optimized & \verb|BiCGStab| & \verb|BiCGStab| & SOR & \mc{1}{c|}{yes} & every time step & \SI{20}{\hour}\,\SI{32}{\minute}\,\SI{32}{\second} & 41.67\% & similar numerics as above, results are almost exactly the same \\\hline
      optimized & \verb|BiCGStab| & \verb|BiCGStab| & SOR & \mc{1}{c|}{yes} & once every $2$ time steps & \SI{13}{\hour}\,\SI{30}{\minute}\,\SI{52}{\second} & 61.62\% & time shift but frequency of vortex shedding unchanged (sign inversion for $y$ and $z$)\\\hline
      optimized & \verb|BiCGStab| & \verb|BiCGStab| & SOR & \mc{1}{c|}{yes} & once every $4$ time steps & \SI{10}{\hour}\,\SI{18}{\minute}\,\SI{41}{\second} & 70.72\% & same as above (no sign inversion) \\\hline
      optimized & \verb|BiCGStab| & \verb|BiCGStab| & SOR & \mc{1}{c|}{yes} & once every $8$ time steps & \SI{8}{\hour}\,\SI{26}{\minute}\,\SI{31}{\second} & 76.03\% & same as above (no sign inversion) \\\hhline{|=|=|=|=|=|=|=|=|=|}
      optimized & CG & \verb|BiCGStab| & \verb|HYPRE| & \mc{1}{c|}{yes} & every time step           & \SI{20}{\hour}\,\SI{44}{\minute}\,\SI{48}{\second} & 41.09\% & different numerics than originally, but results are almost exactly the same \\\hline
      optimized & CG & \verb|BiCGStab| & \verb|HYPRE| & \mc{1}{c|}{yes} & once every $2$ time steps & \SI{14}{\hour}\,\SI{6}{\minute}\,\SI{54}{\second} & 59.92\% & time shift but frequency of vortex shedding unchanged (sign inversion for $y$ and $z$) \\\hline
      optimized & CG & \verb|BiCGStab| & \verb|HYPRE| & \mc{1}{c|}{yes} & once every $4$ time steps & \SI{10}{\hour}\,\SI{4}{\minute}\,\SI{43}{\second} & 71.38\% & same as above (no sign inversion) \\\hline
      optimized & CG & \verb|BiCGStab| & \verb|HYPRE| & \mc{1}{c|}{yes} & once every $8$ time steps & \SI{8}{\hour}\,\SI{20}{\minute}\,\SI{40}{\second} & 76.30\% & same as above (no sign inversion) \\\hline
    \end{tabular}
  \end{center}
  \caption{\label{tab:flow_past_sphere_with_smoke} summary of the runtime-optimization results for three-dimensional flow past a sphere with advection of smoke and smoke-based grid refinement. When the \texttt{HYPRE} preconditioner is used, it is only for the cell solver. }
\end{table}
\end{landscape}

\end{document}

### 2019-08-05 examples/navier_stokes/main_flow_past_dendrites_2d.cpp
Elyce's comments + fixes

@@ -44,9 +44,9 @@
 using namespace std;
 
 const double xmin = 0.0;
-const double xmax = 32.0;
-const double ymin = -8.0;
-const double ymax = 8.0;
+const double xmax = 32.0; //
+const double ymin = 0.0;//
+const double ymax = 16.0;//
 #ifdef P4_TO_P8
 const double zmin = -8.0;
 const double zmax = 8.0;
@@ -56,9 +56,9 @@ const double zmax = 8.0;
 #ifdef P4_TO_P8
 const double r0 = 1.0;
 #else
-const double r0 = 0.5;
+const double r0 =0.5;//
 #endif
-const double u0 = 1.0;
+const double u0 = 1.0;//
 
 #ifdef P4_TO_P8
 
@@ -275,7 +275,7 @@ class INIT_SMOKE : public CF_2
 public:
   double operator()(double x, double y) const
   {
-    return (x<2 && y>-1 && y<1) ? 1 : 0;
+    return (x<(xmax-xmin)/15.0 ) && (y>(ymax - ymin)*5.0/16.0) && (y<((3.0/4.0)*(ymax - ymin))) ? 1 : 0;
   }
 } init_smoke;
 
@@ -285,7 +285,7 @@ class BC_SMOKE : public CF_2
 public:
   double operator()(double x, double y) const
   {
-    return(fabs(x-xmin)<EPS && y>-.5 && y<.5) ? 1 : 0;
+    return(fabs(x-xmin)<EPS && (y>(ymax - ymin)*5.0/16.0) && (y<((3.0/4.0)*(ymax - ymin)))) ? 1 : 0;
   }
 } bc_smoke;
 
@@ -295,8 +295,92 @@ class LEVEL_SET: public CF_2
 public:
   LEVEL_SET() { lip = 1.2; }
   double operator()(double x, double y) const
+
   {
-    return r0 - sqrt(SQR(x-(xmax+xmin)/4) + SQR(y-(ymax+ymin)/2));
+      // Generates a "fake dendrite" via superposition of ellipses
+      double LS = 0.0;
+
+      // Initialize variables
+      double rx = 1.0, yc = 0.0, xc_0 = 1.65, xc_inc = 1.65, xc;
[…]

### 2019-08-05 src/my_p4est_poisson_cells.cpp

@@ -390,26 +390,33 @@ void my_p4est_poisson_cells_t::setup_negative_laplace_matrix()
   std::vector<p4est_quadrant_t> ngbd;
 
   for(p4est_topidx_t tree_idx = p4est->first_local_tree; tree_idx <= p4est->last_local_tree; ++tree_idx){
+      //E: Loop over each tree in the grid
     p4est_tree_t *tree = (p4est_tree_t*)sc_array_index(p4est->trees, tree_idx);
 
+    //E: Loop over each quadrant in the given tree:
     for (size_t q=0; q<tree->quadrants.elem_count; ++q){
       const p4est_quadrant_t *quad = (const p4est_quadrant_t*)sc_array_index(&tree->quadrants, q);
+
+      // E: Grab the local and global index of the current quadrant
       p4est_locidx_t quad_idx = q + tree->quadrants_offset;
       PetscInt quad_gloidx = quad_idx + p4est->global_first_quadrant[p4est->mpirank];
 
       p4est_locidx_t corners[P4EST_CHILDREN];
+      // E: Get the corner nodes of each quadrant
       for(int i=0; i<P4EST_CHILDREN; ++i)
         corners[i] = nodes->local_nodes[quad_idx*P4EST_CHILDREN + i];
 
+      // E: Get the average value of phi for the given quadrant (average over the four corners)
       double phi_q = phi_cell(quad_idx, phi_p);
 
+      //E: Get the dx and dy associated with the given cell
       double dtmp = (double)P4EST_QUADRANT_LEN(quad->level)/(double)P4EST_ROOT_LEN;
       double dx = (xyz_max[0]-xyz_min[0]) * dtmp;
       double dy = (xyz_max[1]-xyz_min[1]) * dtmp;
   #ifdef P4_TO_P8
       double dz = (xyz_max[2]-xyz_min[2]) * dtmp;
   #endif
-
+      // E: Get the coordiantes of the center of the given quadrant, and store these coordinates
       double x = quad_x_fr_q(quad_idx, tree_idx, p4est, ghost);
       double y = quad_y_fr_q(quad_idx, tree_idx, p4est, ghost);
 #ifdef P4_TO_P8
@@ -421,10 +428,12 @@ void my_p4est_poisson_cells_t::setup_negative_laplace_matrix()
                   #endif
                        };
 
+      // E: Initialize a boolean to check if the LSF is positive at each corner of the quadrant, then loop through and check if all corners are positive
       bool all_pos = true;
[…]

### 2019-08-05 src/my_p4est_poisson_faces.cpp

@@ -1050,7 +1050,7 @@ void my_p4est_poisson_faces_t::preallocate_matrix(int dir)
 
 
 void my_p4est_poisson_faces_t::setup_linear_system(int dir)
-{
+{ //E: Sets up the linear system for a given direction --
   PetscErrorCode ierr;
   ierr = PetscLogEventBegin(log_my_p4est_poisson_faces_setup_linear_system, A, rhs[dir], 0, 0); CHKERRXX(ierr);
 
@@ -1067,8 +1067,8 @@ void my_p4est_poisson_faces_t::setup_linear_system(int dir)
   p4est_topidx_t tree_idx;
   matrix_has_nullspace[dir] = true;
 
-  int dir_m = 2*dir;
-  int dir_p = 2*dir+1;
+  int dir_m = 2*dir; // E: Set an index for the minus direction face in the given direction (ie. x, y, z)
+  int dir_p = 2*dir+1; // E: Set an index for the plus direction face in the given direction (ie. x, y, z)
 
   const PetscScalar *face_is_well_defined_p;
   ierr = VecGetArrayRead(face_is_well_defined[dir], &face_is_well_defined_p); CHKERRXX(ierr);
@@ -1089,16 +1089,18 @@ void my_p4est_poisson_faces_t::setup_linear_system(int dir)
   my_p4est_interpolation_faces_t interp_dxyz_hodge(ngbd_n, faces);
   interp_dxyz_hodge.set_input(dxyz_hodge[dir], dir, 1, face_is_well_defined[dir]);
 
+  // E: Loop over all faces in the given direction (ie. all x faces on local processor, or all y faces on local processor)
   for(p4est_locidx_t f_idx=0; f_idx<faces->num_local[dir]; ++f_idx)
   {
     null_space_p[f_idx] = 1;
-    p4est_gloidx_t f_idx_g = f_idx + proc_offset[dir][p4est->mpirank];
+    p4est_gloidx_t f_idx_g = f_idx + proc_offset[dir][p4est->mpirank]; //E: Grab the global index for the given face
 
-    faces->f2q(f_idx, dir, quad_idx, tree_idx);
+    faces->f2q(f_idx, dir, quad_idx, tree_idx); //E: Grabs the local index of quadrant that owns the face (in the given direction) (since faces are accounted for by their index and direction)
 
     p4est_tree_t *tree = (p4est_tree_t*) sc_array_index(p4est->trees, tree_idx);
[…]

### 2019-08-05 src/my_p4est_solve_lsqr.cpp

@@ -226,7 +226,7 @@ void solve_lsqr_system(matrix_t &A, vector<double> p[], unsigned int n_vectors, 
     if( (unsigned int) A.num_rows() != p[k].size() )
       throw std::invalid_argument("[CASL_ERROR]: solve_lsqr_system(...): the matrix and (one of) the right hand side(s) don't have the same size");
 #endif
-  P4EST_ASSERT(nconstraints<P4EST_DIM);
+  P4EST_ASSERT(nconstraints<=P4EST_DIM);
   int m = A.num_rows();
   matrix_t *M = new matrix_t();
   vector<double> Atp[n_vectors];

### 2019-08-05 examples/multialloy/main_2d.cpp
killing bugs in 3D

@@ -81,7 +81,7 @@ DEFINE_PARAMETER(pl, int, lmin, 5, "Min level of the tree");
 DEFINE_PARAMETER(pl, int, lmax, 5, "Max level of the tree");
 #else
 DEFINE_PARAMETER(pl, int, lmin, 6, "Min level of the tree");
-DEFINE_PARAMETER(pl, int, lmax, 10, "Max level of the tree");
+DEFINE_PARAMETER(pl, int, lmax, 9, "Max level of the tree");
 #endif
 
 DEFINE_PARAMETER(pl, double, lip, 1.75, "");
@@ -92,12 +92,12 @@ DEFINE_PARAMETER(pl, double, lip, 1.75, "");
 DEFINE_PARAMETER(pl, bool, use_points_on_interface,   false, "");
 DEFINE_PARAMETER(pl, bool, use_superconvergent_robin, false, "");
 
-DEFINE_PARAMETER(pl, int,    update_c0_robin, 1, "Solve for c0 using Robin BC: 0 - never, 1 - once, 2 - always");
-DEFINE_PARAMETER(pl, int,    num_time_layers, 3, "");
+DEFINE_PARAMETER(pl, int,    update_c0_robin, 0, "Solve for c0 using Robin BC: 0 - never, 1 - once, 2 - always");
+DEFINE_PARAMETER(pl, int,    num_time_layers, 2, "");
 DEFINE_PARAMETER(pl, int,    pin_every_n_iterations, 20, "");
 DEFINE_PARAMETER(pl, int,    max_iterations,   10, "");
-DEFINE_PARAMETER(pl, int,    front_smoothing,   1, "");
-DEFINE_PARAMETER(pl, double, bc_tolerance,      1.e-5, "");
+DEFINE_PARAMETER(pl, int,    front_smoothing,   3, "");
+DEFINE_PARAMETER(pl, double, bc_tolerance,      1.e-12, "");
 DEFINE_PARAMETER(pl, double, cfl_number, 0.3, "");
 DEFINE_PARAMETER(pl, double, phi_thresh, 0.1, "");
 
@@ -248,9 +248,9 @@ void set_alloy_parameters()
       part_coeff_0     = 0.86;
       part_coeff_1     = 0.86;
 
-      eps_c = 0;
+      eps_c = 1.e-3/melting_temp;
       eps_v = 0;
-      eps_a = 0.05;
+      eps_a = 0.0;
       break;
     case 7:// Ni - 0.3at%Cu - 0.1at%Cu
[…]

### 2019-08-05 src/my_p4est_level_set.cpp

@@ -10136,7 +10136,7 @@ double my_p4est_level_set_t::advect_in_normal_direction_with_contact_angle(const
     VecCopyGhost(cos_angle, flux);
     VecPointwiseMultGhost(flux, flux, surf_tns);
 
-    solver.add_boundary(MLS_INTERSECTION, phi_wall, NULL, NULL, NEUMANN, interp, zero_cf);
+    solver.add_boundary(MLS_INTERSECTION, phi_wall, NULL, NEUMANN, interp, zero_cf);
   }
 
   solver.set_use_sc_scheme(0);

### 2019-08-05 src/my_p4est_multialloy.cpp

@@ -630,8 +630,6 @@ void my_p4est_multialloy_t::update_grid()
   bc_error_.destroy();
   bc_error_.create(front_phi_.vec);
 
-
-//  std::cout << "HERE\n";
   if (num_seeds_ == 1)
   {
     seed_map_.destroy();

### 2019-08-05 src/my_p4est_poisson_nodes_mls.cpp

@@ -867,13 +867,13 @@ void my_p4est_poisson_nodes_mls_t::setup_linear_system(bool setup_rhs)
           case FINITE_DIFFERENCE: break;
           case FINITE_VOLUME:
             interpolators_prepare(n);
-            construct_finite_volume(fv, n, p4est_, nodes_, bdry_phi_cf_, bdry_.opn, integration_order_, cube_refinement_, use_sc_scheme_, phi_perturbation_);
+            construct_finite_volume(fv, n, p4est_, nodes_, bdry_phi_cf_, bdry_.opn, integration_order_, cube_refinement_, 1, phi_perturbation_);
             bdry_fvs_->push_back(fv);
             bdry_node_to_fv_[n] = bdry_fvs_->size()-1;
           break;
           case IMMERSED_INTERFACE:
             interpolators_prepare(n);
-            construct_finite_volume(fv, n, p4est_, nodes_, infc_phi_cf_, infc_.opn, integration_order_, cube_refinement_, use_sc_scheme_, phi_perturbation_);
+            construct_finite_volume(fv, n, p4est_, nodes_, infc_phi_cf_, infc_.opn, integration_order_, cube_refinement_, 1, phi_perturbation_);
             infc_fvs_->push_back(fv);
             infc_node_to_fv_[n] = infc_fvs_->size()-1;
           break;
@@ -912,7 +912,7 @@ void my_p4est_poisson_nodes_mls_t::setup_linear_system(bool setup_rhs)
           case FINITE_VOLUME:
           {
             if (finite_volumes_initialized_) fv = bdry_fvs_->at(bdry_node_to_fv_[n]);
[…]

### 2019-08-05 src/my_p4est_poisson_nodes_multialloy.cpp

@@ -694,7 +694,7 @@ void my_p4est_poisson_nodes_multialloy_t::solve_psi_c0()
       pw_psi_c0_values_[idx] =
           -( conc_term + eps_v
              + latent_heat_*(1.+eps_c*kappa)*pt->interpolate(node_neighbors_, psi_t_.ptr)
-             ) /(1.-part_coeff_[0])/c_gamma_all[0];
+             ) /(1.-part_coeff_[0])/pw_c0_values_[idx];
     }
   }
 
@@ -1156,7 +1156,7 @@ void my_p4est_poisson_nodes_multialloy_t::compute_pw_bc_values(int start, int nu
       interp_local.set_input(front_curvature_.vec, linear);
       kappa_pr = interp_local.value(xyz_pr);
 
-      eps_c_pr = eps_c_[seed_map_.ptr[n]]->value(normal);
+      eps_c_pr = 0.*eps_c_[seed_map_.ptr[n]]->value(normal);
 
 //      vn_pr = vn_exact_->value(xyz_pr);
 
@@ -1185,7 +1185,7 @@ void my_p4est_poisson_nodes_multialloy_t::compute_pw_bc_values(int start, int nu
       interp_local.set_input(front_curvature_.vec, linear);
       kappa_cd = interp_local.value(xyz_cd);
 
-      eps_c_cd = eps_c_[seed_map_.ptr[n]]->value(normal);
+      eps_c_cd = 0.*eps_c_[seed_map_.ptr[n]]->value(normal);
 
       pw_t_flx_jump_integr_[idx] = front_temp_flux_jump_->value(xyz_cd) - latent_heat_*(1.0+eps_c_cd*kappa_cd)*vn_cd;
     }
@@ -1458,7 +1458,7 @@ void my_p4est_poisson_nodes_multialloy_t::adjust_c0_gamma(bool simple)
         normal[dim] = pt->interpolate(node_neighbors_, front_normal_.ptr[dim]);
         vn += pt->interpolate(node_neighbors_, c0d_.ptr[dim])*normal[dim];
       }
-      vn = ( (conc_diff_[0]*vn - front_conc_flux_[0]->value(xyz))/(1.-part_coeff_[0])/c_all[0] );
+      vn = ( (conc_diff_[0]*vn - front_conc_flux_[0]->value(xyz))/(1.-part_coeff_[0])/pw_c0_values_[idx] );
 
       // curvature
       kappa = pt->interpolate(node_neighbors_, front_curvature_.ptr);

### 2019-08-05 src/my_p8est_multialloy.h

@@ -2,6 +2,6 @@
 #define MY_P8EST_MULTIALLOY_H
 
 #include <src/my_p4est_to_p8est.h>
-#include "my_p4est_multiallo.h"
+#include "my_p4est_multialloy.h"
 
 #endif /* MY_P8EST_MULTIALLOY_H */

### 2019-08-05 src/my_p4est_interpolation.h
fixing merge issues

@@ -101,6 +101,7 @@ public:
   void set_input(Vec& F) {set_input(&F, 1);}
 
   void add_point(p4est_locidx_t locidx, const double *xyz);
+  void add_point_local(p4est_locidx_t locidx, const double *xyz);
 
   // interpolation methods
   void interpolate(Vec Fo)
@@ -115,6 +116,8 @@ public:
     interpolate(&Fo, 1);
   }
   void interpolate(double * const *Fo, unsigned int n_functions);
+
+  void interpolate_local(double *Fo_p);
 
 #ifdef P4_TO_P8
   virtual void operator()(double x, double y, double z, double* results) const = 0;

### 2019-08-05 src/my_p4est_interpolation_nodes.cpp

@@ -202,9 +202,9 @@ void my_p4est_interpolation_nodes_t::operator ()(double x, double y, double* res
     } else if (method == quadratic_non_oscillatory) {
       quadratic_non_oscillatory_interpolation(p4est, best_match.p.piggy3.which_tree, best_match, f, fdd, xyz_p, results, n_functions);
     } else if (method == quadratic_non_oscillatory_continuous_v1) {
-      value = quadratic_non_oscillatory_continuous_v1_interpolation(p4est, best_match.p.piggy3.which_tree, best_match, f, fdd, xyz_p);
+        quadratic_non_oscillatory_continuous_v1_interpolation(p4est, best_match.p.piggy3.which_tree, best_match, f, fdd, xyz_p, results, n_functions);
     } else if (method == quadratic_non_oscillatory_continuous_v2) {
-      value = quadratic_non_oscillatory_continuous_v2_interpolation(p4est, best_match.p.piggy3.which_tree, best_match, f, fdd, xyz_p);
+        quadratic_non_oscillatory_continuous_v2_interpolation(p4est, best_match.p.piggy3.which_tree, best_match, f, fdd, xyz_p, results, n_functions);
     }
     
   } else {
@@ -336,7 +336,7 @@ void my_p4est_interpolation_nodes_t::interpolate(const p4est_quadrant_t &quad, c
       ierr = VecRestoreArrayRead(Fi[k], &Fi_p[k]); CHKERRXX(ierr);
     }
 
-    if      (method==quadratic)
+    if(method==quadratic)
     {
       quadratic_interpolation(p4est, quad.p.piggy3.which_tree, quad, f, fdd, xyz_p, results, n_functions);
       return;
@@ -347,9 +347,15 @@ void my_p4est_interpolation_nodes_t::interpolate(const p4est_quadrant_t &quad, c
       return;
     }
     else if (method==quadratic_non_oscillatory_continuous_v1)
-      return quadratic_non_oscillatory_continuous_v1_interpolation(p4est, quad.p.piggy3.which_tree, quad, f, fdd, xyz_p);
+    {
+        quadratic_non_oscillatory_continuous_v1_interpolation(p4est, quad.p.piggy3.which_tree, quad, f, fdd, xyz_p, results, n_functions);
+      return;
+    }
     else if (method==quadratic_non_oscillatory_continuous_v2)
[…]

### 2019-08-05 src/my_p4est_utils.cpp

@@ -45,102 +45,6 @@
 
 std::vector<InterpolatingFunctionLogEntry> InterpolatingFunctionLogger::entries;
 
-WallBC2D::~WallBC2D() {};
-WallBC3D::~WallBC3D() {};
-
-bool index_of_node(const p4est_quadrant_t *n, p4est_nodes_t* nodes, p4est_locidx_t& idx)
-{
-#ifdef P4EST_DEBUG
-  int clamped = 1;
-#endif
-  P4EST_ASSERT(p4est_quadrant_is_node(n, clamped));
-  unsigned int idx_l, idx_u, idx_m;
-  const p4est_indep_t *node_l, *node_u, *node_m;
-  // check if the candidate can be in the locally owned nodes first, or in the ghost ones
-  idx_l   = 0;
-  idx_u   = nodes->num_owned_indeps-1;
-  node_l  = (const p4est_indep_t*) sc_array_index(&nodes->indep_nodes, idx_l);
-  node_u  = (const p4est_indep_t*) sc_array_index(&nodes->indep_nodes, idx_u);
-  if((p4est_quadrant_compare_piggy(node_l, n) > 0) || (p4est_quadrant_compare_piggy(node_u, n) < 0))
-    goto lookup_in_ghost_nodes;
-  while((p4est_quadrant_compare_piggy(node_l, n) <= 0) && (p4est_quadrant_compare_piggy(node_u, n) >= 0))
-  {
-    if(!p4est_quadrant_compare_piggy(node_l, n))
-    {
-      idx = idx_l;
-      return true;
-    }
-    if(!p4est_quadrant_compare_piggy(node_u, n))
-    {
-      idx = idx_u;
-      return true;
-    }
-    if(idx_u-idx_l == 1)
-      break;
-    idx_m   = (idx_l + idx_u)/2;
-    node_m  = (const p4est_indep_t*) sc_array_index(&nodes->indep_nodes, idx_m);
-    P4EST_ASSERT((p4est_quadrant_compare_piggy(node_l, node_m) <= 0) && (p4est_quadrant_compare_piggy(node_u, node_m) >= 0));
[…]

### 2019-08-05 src/my_p4est_utils.h

@@ -5,16 +5,12 @@
 #ifdef P4_TO_P8
 #include <p8est.h>
 #include <p8est_nodes.h>
-#include <src/my_p8est_nodes.h>
 #include <p8est_ghost.h>
-#include <p8est_bits.h>
 #include <src/my_p8est_refine_coarsen.h>
 #else
 #include <p4est.h>
 #include <p4est_nodes.h>
-#include <src/my_p4est_nodes.h>
 #include <p4est_ghost.h>
-#include <p4est_bits.h>
 #include <src/my_p4est_refine_coarsen.h>
 #endif
 #include <src/petsc_logging.h>
@@ -108,7 +104,7 @@ struct quad_neighbor_nodes_of_node_t;
 #define ZFOR(a)
 #endif
 
-enum cf_value_type_t { VAL, DDX, DDY, DDZ, LAP, CUR };
+enum cf_value_type_t { VAL, DDX, DDY, DDZ, LAP };
 
 enum mls_opn_t { MLS_INTERSECTION = 0, MLS_ADDITION = 1, MLS_COLORATION = 2, MLS_INT = MLS_INTERSECTION, MLS_ADD = MLS_ADDITION };
 
@@ -374,8 +370,7 @@ enum {
   WALL_0p0 = -4,
   WALL_00m = -5,
   WALL_00p = -6,
-  INTERFACE = -7,
-  WALL_parallel_to_face = -8 // to allow for Dirichlet wall boundary conditions on the face_solver even with rectangular grids
+  INTERFACE = -7
 };
 
 typedef enum {
@@ -402,7 +397,6 @@ class WallBC2D
 public:
   virtual BoundaryConditionType operator()( double x, double y ) const=0 ;
   double value(double *xyz) const {return this->operator ()(xyz[0], xyz[1]);}
-  virtual ~WallBC2D() = 0;
 };
 
 class WallBC3D
@@ -410,7 +404,6 @@ class WallBC3D
 public:
   virtual BoundaryConditionType operator()( double x, double y, double z ) const=0 ;
   double value(double *xyz) const {return this->operator ()(xyz[0], xyz[1], xyz[2]);}
-  virtual ~WallBC3D() = 0;
 };
 
 #ifdef P4_TO_P8
@@ -533,11 +526,6 @@ public:
 #endif
     return p_InterfaceValue->operator ()(x,y);
   }
-  inline double  interfaceValue(double xyz_[]) const
-  {
[…]

### 2019-08-05 src/my_p4est_level_set.cpp
fixing merge issues

@@ -6182,1361 +6182,1361 @@ void my_p4est_level_set_t::extend_Over_Interface_TVD( Vec phi, Vec mask, Vec q, 
 }
 
 
-void my_p4est_level_set_t::extend_Over_Interface_TVD(Vec phi, Vec q, my_p4est_poisson_nodes_t *solver, int iterations, int order, Vec normal[]) const
-{
-#ifdef CASL_THROWS
-  if(order!=0 && order!=1 && order!=2) throw std::invalid_argument("[CASL_ERROR]: my_p4est_level_set_t->extend_Over_Interface_TVD: order must be 0, 1 or 2.");
-#endif
-  PetscErrorCode ierr;
-  ierr = PetscLogEventBegin(log_my_p4est_level_set_extend_over_interface_TVD, phi, q, 0, 0); CHKERRXX(ierr);
-
-  double *phi_p;
-  ierr = VecGetArray(phi, &phi_p); CHKERRXX(ierr);
-
-  Vec qn, qnn;
-  double *q_p, *qn_p, *qnn_p;
-  Vec b_qn_well_defined;
-  Vec b_qnn_well_defined;
-  double *b_qn_well_defined_p;
-  double *b_qnn_well_defined_p;
-
-  Vec tmp, tmp_loc;
-  Vec qnn_loc, qn_loc, q_loc;
-  ierr = VecDuplicate(phi, &tmp); CHKERRXX(ierr);
-  double *tmp_p;
-
-  double dxyz[P4EST_DIM];
-
-  dxyz_min(p4est, dxyz);
-
-#ifdef P4_TO_P8
-  double diag = sqrt(SQR(dxyz[0]) + SQR(dxyz[1]) + SQR(dxyz[2]));
-#else
-  double diag = sqrt(SQR(dxyz[0]) + SQR(dxyz[1]));
-#endif
-
-  double rel_thresh = 1.e-2;
-
-  /* init the neighborhood information if needed */
-  /* NOTE: from now on the neighbors will be initialized ... do we want to clear them
-   * at the end of this function if they were not initialized beforehand ?
-   */
-  ngbd->init_neighbors();
-
-  /* compute the normals */
-  double *normal_p[P4EST_DIM];
-
-  if (normal != NULL)
[…]

### 2019-08-05 src/my_p4est_level_set.h

@@ -181,7 +181,7 @@ public:
   /* extend a quantity over the interface with the TVD algorithm using mask */
   void extend_Over_Interface_TVD(Vec phi, Vec mask, Vec q, int iterations=20, int order=2) const;
 
-  void extend_Over_Interface_TVD(Vec phi, Vec q, my_p4est_poisson_nodes_t *solver, int iterations=20, int order=2, Vec normal[P4EST_DIM] = NULL) const;
+//  void extend_Over_Interface_TVD(Vec phi, Vec q, my_p4est_poisson_nodes_t *solver, int iterations=20, int order=2, Vec normal[P4EST_DIM] = NULL) const;
 
   void extend_Over_Interface_TVD_not_parallel(Vec phi, Vec q, int iterations=20, int order=2) const;
 

### 2019-08-05 src/my_p4est_poisson_nodes.h

@@ -1,5 +1,5 @@
-#ifndef MY_P4EST_POISSON_NODES_H
-#define MY_P4EST_POISSON_NODES_H
+#ifndef MY_P4EST_POISSON_NODEs_H
+#define MY_P4EST_POISSON_NODEs_H
 
 #include <petsc.h>
 
@@ -17,7 +17,6 @@
 
 class my_p4est_poisson_nodes_t
 {
-  static const int cube_refinement = 1;
   const my_p4est_node_neighbors_t *node_neighbors_;
 
   // p4est objects
@@ -26,13 +25,11 @@ class my_p4est_poisson_nodes_t
   p4est_ghost_t *ghost;
   my_p4est_brick_t *myb_;
   my_p4est_interpolation_nodes_t phi_interp;
-//  my_p4est_interpolation_nodes_t robin_coef_interp;
 
   bool neumann_wall_first_order;
   double mu_, diag_add_;
   bool is_matrix_computed;
   int matrix_has_nullspace;
-  double dxyz_m[P4EST_DIM];
   double dx_min, dy_min, d_min, diag_min;
 #ifdef P4_TO_P8
   double dz_min;
@@ -51,26 +48,20 @@ class my_p4est_poisson_nodes_t
   p4est_gloidx_t fixed_value_idx_l;
   bool is_phi_dd_owned, is_mue_dd_owned;
   Vec rhs_, phi_, add_, mue_, phi_xx_, phi_yy_, mue_xx_, mue_yy_;
-//  Vec robin_coef_;
+  Vec robin_coef_;
 #ifdef P4_TO_P8
   Vec phi_zz_, mue_zz_;
 #endif
   KSP ksp;
   PetscErrorCode ierr;
 
-  Vec mask;
-  std::vector<double> scalling;
-  bool keep_scalling;
-
-  bool variable_mu;
-
-  bool use_refined_cube;
-  bool use_pointwise_dirichlet;
-  bool use_continuous_stencil;
-
-  bool new_pc;
-
   void preallocate_matrix();
+
+  void setup_negative_laplace_matrix_neumann_wall_1st_order();
+  void setup_negative_laplace_rhsvec_neumann_wall_1st_order();
+
+  void setup_negative_laplace_matrix();
+  void setup_negative_laplace_rhsvec();
 
   void setup_negative_variable_coeff_laplace_matrix();
   void setup_negative_variable_coeff_laplace_rhsvec();
@@ -104,9 +95,8 @@ public:
 #else
[…]

### 2019-08-05 src/my_p4est_refine_coarsen.cpp

@@ -497,12 +497,12 @@ coarsen_marked_quadrants(p4est_t *p4est, p4est_topidx_t which_tree, p4est_quadra
   return P4EST_FALSE;
 }
 
-void splitting_criteria_tag_t::tag_quadrant(p4est_t *p4est, p4est_quadrant_t *quad, p4est_topidx_t which_tree, const double *f, bool finest_in_negative_flag){
+void splitting_criteria_tag_t::tag_quadrant(p4est_t *p4est, p4est_quadrant_t *quad, p4est_topidx_t which_tree, const double* f) {
   if (quad->level < min_lvl) {
     quad->p.user_int = REFINE_QUADRANT;
 
   } else if (quad->level > max_lvl) {
-    quad->p.user_int = COARSEN_QUADRANT;    
+    quad->p.user_int = COARSEN_QUADRANT;
 
   } else {
     p4est_topidx_t v_m = p4est->connectivity->tree_to_vertex[P4EST_CHILDREN*which_tree + 0];
@@ -534,25 +534,23 @@ void splitting_criteria_tag_t::tag_quadrant(p4est_t *p4est, p4est_quadrant_t *qu
 #endif
 
     // refinement based on distance
-    bool refine = false, coarsen = true;
+                bool refine = false, coarsen = true;
 
-    if(finest_in_negative_flag)
-      for (short i = 0; i < P4EST_CHILDREN; i++) {
-        refine  = refine  || (f[i] <= 0.5*lip*d && quad->level < max_lvl);
-        coarsen = coarsen && (f[i] >= 1.0*lip*d && quad->level > min_lvl);
-      }
-    else
-      for (short i = 0; i < P4EST_CHILDREN; i++) {
-        refine  = refine  || (fabs(f[i]) <= 0.5*lip*d && quad->level < max_lvl);
-        coarsen = coarsen && (fabs(f[i]) >= 1.0*lip*d && quad->level > min_lvl);
-      }
+    for (short i = 0; i < P4EST_CHILDREN; i++) {
+                        refine  = refine  || (fabs(f[i]) <= 0.5*lip*d && quad->level < max_lvl);
+                        coarsen = coarsen && (fabs(f[i]) >= 1.0*lip*d && quad->level > min_lvl);
+                }
 
-    if (refine)
-      quad->p.user_int = REFINE_QUADRANT;
-    else if (coarsen)
[…]

### 2019-08-05 src/my_p4est_refine_coarsen.h

@@ -129,7 +129,7 @@ protected:
 	static void init_fn   (p4est_t* p4est, p4est_topidx_t which_tree, p4est_quadrant_t*  quad);
 	static int  refine_fn (p4est_t* p4est, p4est_topidx_t which_tree, p4est_quadrant_t*  quad);
 	static int  coarsen_fn(p4est_t* p4est, p4est_topidx_t which_tree, p4est_quadrant_t** quad);
-	
+
   void tag_quadrant(p4est_t* p4est, p4est_quadrant_t* quad, p4est_topidx_t which_tree, const double* f);
   void tag_quadrant_inside(p4est_t* p4est, p4est_quadrant_t* quad, p4est_topidx_t which_tree, const double* f);
   bool refine_only_inside;

### 2019-08-05 src/my_p4est_utils.cpp

@@ -45,6 +45,99 @@
 
 std::vector<InterpolatingFunctionLogEntry> InterpolatingFunctionLogger::entries;
 
+bool index_of_node(const p4est_quadrant_t *n, p4est_nodes_t* nodes, p4est_locidx_t& idx)
+{
+#ifdef P4EST_DEBUG
+  int clamped = 1;
+#endif
+  P4EST_ASSERT(p4est_quadrant_is_node(n, clamped));
+  unsigned int idx_l, idx_u, idx_m;
+  const p4est_indep_t *node_l, *node_u, *node_m;
+  // check if the candidate can be in the locally owned nodes first, or in the ghost ones
+  idx_l   = 0;
+  idx_u   = nodes->num_owned_indeps-1;
+  node_l  = (const p4est_indep_t*) sc_array_index(&nodes->indep_nodes, idx_l);
+  node_u  = (const p4est_indep_t*) sc_array_index(&nodes->indep_nodes, idx_u);
+  if((p4est_quadrant_compare_piggy(node_l, n) > 0) || (p4est_quadrant_compare_piggy(node_u, n) < 0))
+    goto lookup_in_ghost_nodes;
+  while((p4est_quadrant_compare_piggy(node_l, n) <= 0) && (p4est_quadrant_compare_piggy(node_u, n) >= 0))
+  {
+    if(!p4est_quadrant_compare_piggy(node_l, n))
+    {
+      idx = idx_l;
+      return true;
+    }
+    if(!p4est_quadrant_compare_piggy(node_u, n))
+    {
+      idx = idx_u;
+      return true;
+    }
+    if(idx_u-idx_l == 1)
+      break;
+    idx_m   = (idx_l + idx_u)/2;
+    node_m  = (const p4est_indep_t*) sc_array_index(&nodes->indep_nodes, idx_m);
+    P4EST_ASSERT((p4est_quadrant_compare_piggy(node_l, node_m) <= 0) && (p4est_quadrant_compare_piggy(node_u, node_m) >= 0));
+    if(p4est_quadrant_compare_piggy(node_m, n) <0)
+    {
+      idx_l   = idx_m;
+      node_l  = node_m;
+    }
+    else if (p4est_quadrant_compare_piggy(node_m, n) >0)
+    {
+      idx_u   = idx_m;
+      node_u  = node_m;
+    }
+    else
+    {
[…]

### 2019-08-05 src/my_p4est_utils.h

@@ -104,7 +104,7 @@ struct quad_neighbor_nodes_of_node_t;
 #define ZFOR(a)
 #endif
 
-enum cf_value_type_t { VAL, DDX, DDY, DDZ, LAP };
+enum cf_value_type_t { VAL, DDX, DDY, DDZ, LAP, CUR };
 
 enum mls_opn_t { MLS_INTERSECTION = 0, MLS_ADDITION = 1, MLS_COLORATION = 2, MLS_INT = MLS_INTERSECTION, MLS_ADD = MLS_ADDITION };
 
@@ -414,7 +414,6 @@ public:
 #define BoundaryConditionsDIM BoundaryConditions2D
 #endif
 
-
 class BoundaryConditions2D
 {
 private:
@@ -721,12 +720,24 @@ public:
 };
 
 /*!
+ * \brief index_of_node finds the (local) index of a node as defined within p4est, i.e. as a pest_quadrant_t structure whose level is P4EST_MAXLEVEL!
+ *        The method uses a binary search through the provided nodes: its complexity is O(log(N_nodes)).
+ *        The given node MUST MANDATORILY be canonicalized before being passed to this function to ensure consistency with the provided nodes: use
+ *        p4est_node_canonicalize beforehand!
+ * \param [in]    n node whose local index is queried!
+ * \param [in]    nodes the nodes data structure
+ * \param [inout] idx the local index of the node on output if found, undefined if not found (i.e. if the returned value is false)
+ * \return true if the queried node exists and was found in the nodes (i.e. if the idx is valid), false otherwise.
+ */
+bool index_of_node(const p4est_quadrant_t *n, p4est_nodes_t* nodes, p4est_locidx_t& idx);
+
+/*!
  * \brief linear_interpolation performs linear interpolation for a point
  * \param [in]    p4est the forest
  * \param [in]    tree_id the current tree that owns the quadrant
  * \param [in]    quad the current quarant
  * \param [in]    F a simple C-style array of size n_results*P4EST_CHILDREN, containing the values of the n_vecs function(s) at the vertices of the quadrant. __MUST__ be z-ordered
[…]

### 2019-08-05 examples/poisson_jump_voronoi/main_jump_2d.cpp
Merge remote-tracking branch 'origin/multialloy' into nonlinearPB

@@ -48,9 +48,9 @@
 
 using namespace std;
 
-int lmin = 3;
-int lmax = 10;
-int nb_splits = 3;
+int lmin = 2;
+int lmax = 5;
+int nb_splits = 5;
 
 int k1 = 1;
 int k2 = 1;
@@ -67,7 +67,7 @@ bool save_stats = true;
 bool check_partition = false;
 
 const int wave_number = 1;
-const double mu_value = 4.3;
+const double mu_value = 1;
 const double mu_ratio = 10.0;
 
 struct domain{

### 2019-08-05 examples/poisson_nodes_mls/main_2d.cpp

@@ -273,7 +273,7 @@ DEFINE_PARAMETER(pl, bool, save_matrix_ascii,  0, "Save the matrix in ASCII MATL
 DEFINE_PARAMETER(pl, bool, save_matrix_binary, 0, "Save the matrix in BINARY MATLAB format");
 DEFINE_PARAMETER(pl, bool, save_convergence,   0, "Save convergence results");
 
-DEFINE_PARAMETER(pl, int, n_example, 1, "Predefined example");
+DEFINE_PARAMETER(pl, int, n_example, 3, "Predefined example");
 
 void set_example(int n_example)
 {

### 2019-08-05 src/my_p4est_level_set.cpp

@@ -10889,7 +10889,7 @@ double my_p4est_level_set_t::advect_in_normal_direction_with_contact_angle(const
     VecCopyGhost(cos_angle, flux);
     VecPointwiseMultGhost(flux, flux, surf_tns);
 
-    solver.add_boundary(MLS_INTERSECTION, phi_wall, NULL, NULL, NEUMANN, interp, zero_cf);
+    solver.add_boundary(MLS_INTERSECTION, phi_wall, NULL, NEUMANN, interp, zero_cf);
   }
 
   solver.set_use_sc_scheme(0);

### 2019-08-05 src/my_p4est_poisson_nodes_mls.cpp

@@ -867,13 +867,13 @@ void my_p4est_poisson_nodes_mls_t::setup_linear_system(bool setup_rhs)
           case FINITE_DIFFERENCE: break;
           case FINITE_VOLUME:
             interpolators_prepare(n);
-            construct_finite_volume(fv, n, p4est_, nodes_, bdry_phi_cf_, bdry_.opn, integration_order_, cube_refinement_, use_sc_scheme_, phi_perturbation_);
+            construct_finite_volume(fv, n, p4est_, nodes_, bdry_phi_cf_, bdry_.opn, integration_order_, cube_refinement_, 1, phi_perturbation_);
             bdry_fvs_->push_back(fv);
             bdry_node_to_fv_[n] = bdry_fvs_->size()-1;
           break;
           case IMMERSED_INTERFACE:
             interpolators_prepare(n);
-            construct_finite_volume(fv, n, p4est_, nodes_, infc_phi_cf_, infc_.opn, integration_order_, cube_refinement_, use_sc_scheme_, phi_perturbation_);
+            construct_finite_volume(fv, n, p4est_, nodes_, infc_phi_cf_, infc_.opn, integration_order_, cube_refinement_, 1, phi_perturbation_);
             infc_fvs_->push_back(fv);
             infc_node_to_fv_[n] = infc_fvs_->size()-1;
           break;
@@ -2645,60 +2645,60 @@ void my_p4est_poisson_nodes_mls_t::discretize_dirichlet(bool setup_rhs, p4est_lo
       else                          w_00p += -2.0*wk/d_00p/(d_00m+d_00p);
 
       if(!is_interface[dir::f_m00] && !is_wall[dir::f_m00]) {
[…]

### 2019-08-05 src/my_p4est_poisson_nodes_multialloy.cpp

@@ -673,7 +673,7 @@ void my_p4est_poisson_nodes_multialloy_t::solve_psi_c0()
       pw_psi_c0_values_[idx] =
           -( conc_term + eps_v
              + latent_heat_*(1.+eps_c*kappa)*pt->interpolate(node_neighbors_, psi_t_.ptr)
-             ) /(1.-part_coeff_[0])/c_gamma_all[0];
+             ) /(1.-part_coeff_[0])/pw_c0_values_[idx];
     }
   }
 
@@ -1251,7 +1251,7 @@ void my_p4est_poisson_nodes_multialloy_t::compute_pw_bc_values(int start, int nu
       interp_local.set_input(front_curvature_.vec, linear);
       kappa_pr = interp_local.value(xyz_pr);
 
-      eps_c_pr = eps_c_[seed_map_.ptr[n]]->value(normal);
+      eps_c_pr = 0*eps_c_[seed_map_.ptr[n]]->value(normal);
 
 //      vn_pr = vn_exact_->value(xyz_pr);
 
@@ -1280,7 +1280,7 @@ void my_p4est_poisson_nodes_multialloy_t::compute_pw_bc_values(int start, int nu
       interp_local.set_input(front_curvature_.vec, linear);
       kappa_cd = interp_local.value(xyz_cd);
 
-      eps_c_cd = eps_c_[seed_map_.ptr[n]]->value(normal);
+      eps_c_cd = 0*eps_c_[seed_map_.ptr[n]]->value(normal);
 
       pw_t_flx_jump_integr_[idx] = front_temp_flux_jump_->value(xyz_cd) - latent_heat_*(1.0+eps_c_cd*kappa_cd)*vn_cd;
     }
@@ -1553,7 +1553,7 @@ void my_p4est_poisson_nodes_multialloy_t::adjust_c0_gamma(bool simple)
         normal[dim] = pt->interpolate(node_neighbors_, front_normal_.ptr[dim]);
         vn += pt->interpolate(node_neighbors_, c0d_.ptr[dim])*normal[dim];
       }
-      vn = ( (conc_diff_[0]*vn - front_conc_flux_[0]->value(xyz))/(1.-part_coeff_[0])/c_all[0] );
+      vn = ( (conc_diff_[0]*vn - front_conc_flux_[0]->value(xyz))/(1.-part_coeff_[0])/pw_c0_values_[idx] );
 
       // curvature
       kappa = pt->interpolate(node_neighbors_, front_curvature_.ptr);

### 2019-08-05 src/my_p4est_shapes.h

@@ -788,7 +788,7 @@ public:
 
         for (int i = 0; i < N; ++i)
         {
-          perturb_0 +=  beta[i]*сos(n[i]*(t - theta[i]));
+           perturb_0 +=  beta[i]*cos(n[i]*(t - theta[i]));
         }
 
         return inside*( r -r0*(1.+ perturb_0*pow(r2/r,p)) );
@@ -803,7 +803,7 @@ public:
 
         for (int i = 0; i < N; ++i)
         {
-          perturb_0 +=  beta[i]*сos(n[i]*(t - theta[i]));
+          perturb_0 +=  beta[i]*cos(n[i]*(t - theta[i]));
           perturb_1 += -beta[i]*sin(n[i]*(t - theta[i]))*n[i];
         }
 
@@ -826,9 +826,9 @@ public:
 
         for (int i = 0; i < N; ++i)
         {
-          perturb_0 +=  beta[i]*сos(n[i]*(t - theta[i]));
-          perturb_1 += -beta[i]*sin(n[i]*(t - theta[i]))*n[i];
-          perturb_2 += -beta[i]*сos(n[i]*(t - theta[i]))*n[i]*n[i];
+          perturb_0 +=  beta[i]* cos(n[i]*(t - theta[i]));
+          perturb_1 += -beta[i]* sin(n[i]*(t - theta[i]))*n[i];
+          perturb_2 += -beta[i]* cos(n[i]*(t - theta[i]))*n[i]*n[i];
         }
 
         double t_x = - Y/r/r;
@@ -862,6 +862,8 @@ public:
         double r2_yz = 0;
         double r2_zx = 0;
 
+        double e= perturb_0;
+
         double e_x = perturb_1*t_x;
         double e_y = perturb_1*t_y;
         double e_z = 0;
@@ -880,28 +882,28 @@ public:
[…]
